### 2019-03-14 ACII_2019_paper/ACII_2019_paper.Rmd
janitor

---
title: The relation betweem feeling and expressing emotions. A comparison between sujective feeling and facial expression recognition of spontaneous emotions. 
affiliation:
  ## use one only of the following
  # author-columnar: true         ## one column per author
  institution-columnar: true  ## one column per institution (multiple autors eventually)
  # wide: true                  ## one column wide author/affiliation fields

  institution:
    - name: Dublin City University
      location: Dublin, Ireland
      email: damien.dupre@dcu.ie
      mark: 1
      author:
        - name: Damien Dupré
    - name: Grenoble Alpes University
      location: Grenoble, France
      mark: 2
      author:
        - name: Anna Tcherkassof
          email: anna.tcherkassof@univ-grenoble-alpes.fr
abstract: |
  The abstract goes here.
  On multiple lines eventually.

bibliography: mybibfile.bib
output: rticles::ieee_article
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  cache.extra = knitr::rand_seed, 
  message = FALSE, 
  warning = FALSE, 
  error = FALSE, 
  echo = FALSE, 
  fig.align="center",
  fig.pos = 'H')
#
set.seed(123)# Seed for random number generation
options(scipen = 999)# Disable scientific number format
#
library(tidyverse) # Data wrangling & grammar of graphics
library(here) # Locate files relative to project root
```

```{r data}
# self_report_data
self_report_data <- readr::read_rds(here::here("data/human_self_report_data/Base_Finale_20090114.rds"))
# metadata_video
metadata_video <- readr::read_rds(here::here("metadata_video.rds"))
# human_recognition_data
source(here::here("script/data_wrangling_human_recognition.R"))
```

Introduction
=============

With the developement of commercial automatic facial expression recognition tools, industries and governments are gradually implementing this technology in order to track humans' emotions in various scenarios (e.g., marketing, heathcare, automotive to name a few). Beside the ethical question of measuring human emotions, these tools are challenging psychological theories about infering emotions from facial expressions. When automatic facial expression recognition tools are used to measure human emotions, they rest on the premise that facial expressions provide a direct access to expressers' emotions. However the link between emotion felt and facial expressions is far from being established and still remain a hot topic in psychology research.

## The link between emotion and facial expression thought the behaviorist approach
Based on the behaviorist approach initiated by Darwin in *The Expression of the Emotions in Man and Animals* [@darwin1872expression], facial expression are conceived as a genuine display of individuals inner emotional state. This hypothesis is used as a basis for the Basic Emotion Theory (BET) which states that a set of six or seven emotions are universally displayed and are genetically hardwired [@ekman1992argument]. According to this view, "when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including [...] facial expressions." [@ekman2007directed, p49]. Even if this theory obtained a popular support, it fails to explain how individuals can feel emotions without expressing them and how individuals can express emotions without feeling them.

## The link between emotion and facial expression thought the social constructivist approach
Detractors of the Basic Emotion Theory are perceiving emotion not as genetically hardwired but as a learnt association between a given situation and an appropriate response [@averill1980constructivist]. For the tenants of the constructivist approach, emotions are "concepts" based on past experiences and which are "a collection of embodied, whole brain representations that predict what is about to happen in the sensory environment, what the best action is to deal with impending events, and their consequences for allostasis" [@barrett2017theory, p12]. Following this assumption, faces are used as tools to display signals in social interactions [@crivelli2018facial]. These signals can convey individuals' motivations and readiness [@frijda1997facial] or social messages [@fridlund1995human].

This current paper investigates the link between the subjective feeling of emotions and their recognition from facial expressions. If emotions are hardwired, individuals emotional subjective feeling should be correlated to the recognition of facial expressions from both human observers and automatic classifiers whereas if emotions are social constructs, no correlation between subjective feeling and facial expression recognition should be observed.

Method
=============

To evaluate the link between subjective feeling of emotions and their recognition from facial expressions, participants were recruited to perform an emotion elicitation task while their facial expression was video recorded. Then, the videos was shown to human observer and analysed by an automatic classifier in order to identify which emotion was displayed.

## Emotion Elicitation

```{r self_report_method}
self_report_gender <- self_report_data %>% 
  dplyr::group_by(Genre_C) %>% 
  dplyr::summarise(n= n())

self_report_age <- self_report_data %>%
  dplyr::summarise(Age_m = mean(Age), Age_sd = sd(Age)) %>% 
  dplyr::mutate_all(round,1)
```

For the emotion elicitation experiment, `r nrow(self_report_data)` French participants (`r self_report_gender[self_report_gender$Genre_C == "F","n"]` females, `r self_report_gender[self_report_gender$Genre_C == "H","n"]` males, *M*age = `r self_report_age[,"Age_m"]`, *SD*age = `r self_report_age[,"Age_sd"]`) were recruited to perform one out of 11 emotions eliciation tasks designed to trigger a positive, negative or neutral emotional state [see @tcherkassof2013dynemo for a description of tasks and procedure]. Participants' face were recorded using an hidden camera resulting `r nrow(self_report_data)` front facing 768x576 videos varying from `r min(metadata_video$ffprobe_duration)`s to `r max(metadata_video$ffprobe_duration)`s.

After their emotion eliciation task the participants had to rate their emotional state during the task on a likert scale from 0 ("not at all") to 5 ("strongly") the six "basic" emotions (i.e., *Anger*, *Disgust*, *Fear*, *Happiness*, *Surprise* and *Sadness*) as well as six "non-basic" emotions (i.e., *Pride*, *Curiosity*, *Boredom*, *Shame*, *Humiliation*, and *Disappointement*). The emotion scales rated as the highest is considered as describing participants subjective feeling.

Finally, a debriefing session was perform to emsure that participants were not durably affected by the emotion eliciation task. All the participants gave their agreement on their data and video to be processed for research puropose only.

## Human Facial Expression Recognition

```{r human_recognition_method}
unique_ppt_count <- human_recognition_data %>% 
  dplyr::group_by(source,SEXE_Juge, C_Juge) %>% 
  dplyr::summarise(n = n())

unique_video_count <- human_recognition_data %>% 
  dplyr::select(source,SEXE_Juge, C_Juge, C_Video) %>% 
  unique() %>% 
  dplyr::group_by(C_Video) %>% 
  dplyr::summarise(n = n())
```

For the human facial expression recognition method, `r nrow(unique_ppt_count)` student participants were recruited to annotate `r nrow(unique_video_count)` out of the `r nrow(self_report_data)` video, therefore only the `r nrow(unique_video_count)` annotated videos will be analysed in this paper. Each participants had to annotate between `r min(unique_ppt_count$n)` and `r max(unique_ppt_count$n)` videos resulting that ach video was annotated `r mean(unique_video_count$n) %>% round(0)` times on average (*SD* = `r sd(unique_video_count$n) %>% round(0)`).

The annotation of facial expressions was performed on-site using *Oudjat*, a software for designing video annotation experiements [@dupre2015oudjat]. For each video, the annotation procedure hat two steps. First, the participants had to identify the emotional sequences by pressing the space bar of their keyboard to indicate the begining and the end of the emotional sequences while watching the video. Second, the participants watched each emotional sequence previously identified and had to label the sequence using one of the 12 emotions proposed including six "basic" emotions (i.e., *Anger*, *Disgust*, *Fear*, *Happiness*, *Surprise* and *Sadness*) and six "non-basic" emotions (i.e., *Pride*, *Curiosity*, *Boredom*, *Shame*, *Humiliation*, and *Disappointement*). They also had the possibility to indicate that the sequence was not expressing one of the proposed emotion.

This annotation procedure results in a uni-dimesional time-serie for each video per human observer identifying for each second of the video which emotion was recognized. Then time-series corresponding to the same video were aggreated to calculate the proprotion of human observers for each second of the video per emotional label. The sum of each label proportion per second was used as a score to determine which labels corresponds to the overall video (i.e., the highest score).

## Automatic Facial Expression Recognition

The `r nrow(unique_video_count)` annotated video were processed with Affdex (SDK v3.4.1). Affdex is an automatic facial expression recognition classifier developed and distributed by Affectiva is a spinoff company resulting from the research activities of MIT media lab created in 2009 [@mcduff2016affdex]. Affdex’s algorithm uses Histogram of Oriented Gradient (HOG) features and Support Vector Machine (SVM) classifiers in order to recognize facial expressions. For each video frame, Affdex identify the probability of the face as expressing one of the six "basic" emotions (i.e., *Anger*, *Disgust*, *Fear*, *Happiness*, *Surprise* and *Sadness*) as well as additional psychological states such as *Valence*, *Engagement* or *Contempt*, and facial features such as *Cheek Raise*, *Eye Widen* or *Jaw Drop*.

To determine which of the six "basic" emotion can be used to identify each video, the recognition probability for each label by frame was converted into odd ratio by frame [@dente2017measures]. The highest sum of each odd ratio time-serie defines the label recognized by the automatic classifier.

Results
============

Whereas the self-reports, the human annotations and the automatic recognition include data on "non-basic" emotions and features, the analysis is performed using only the six "basic" emotions in order to compare them.

## Correlation between self-report and human facial expression recognition

## Correlation between self-report and automatic facial expression recognition

Conclusion
============
The conclusion goes here.

Acknowledgment {#acknowledgment}
==============

The authors would like to thank Brigitte Meillon and Jean Michel Adam who developed the software used to collect and preprocess human observer annotations.

References {#references .numbered}
==========

### 2019-03-14 ACII_2019_paper/ACII_2019_paper.tex
janitor

%% template.tex
%% from
%% bare_conf.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE!
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[conference,final,]{IEEEtran}
% Some Computer Society conferences also require the compsoc mode option,
% but others use the standard conference format.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at:
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%

%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

%% BEGIN MY ADDITIONS %%

\usepackage[unicode=true]{hyperref}

\hypersetup{
            pdftitle={The relation betweem feeling and expressing emotions. A comparison between sujective feeling and facial expression recognition of spontaneous emotions.},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls

% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}

% Pandoc syntax highlighting

% Pandoc header

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%% END MY ADDITIONS %%

\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{The relation betweem feeling and expressing emotions. A comparison
between sujective feeling and facial expression recognition of
spontaneous emotions.}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations

\author{

%% ---- classic IEEETrans wide authors' list ----------------
 % -- end affiliation.wide
%% ----------------------------------------------------------

%% ---- classic IEEETrans one column per institution --------
 %% -- beg if/affiliation.institution-columnar
\IEEEauthorblockN{
  %% -- beg for/affiliation.institution.author
Damien Dupré %% -- end for/affiliation.institution.author
}
\IEEEauthorblockA{Dublin City University\\
Dublin, Ireland
\\damien.dupre@dcu.ie
}
\and
\IEEEauthorblockN{
  %% -- beg for/affiliation.institution.author
Anna Tcherkassof %% -- end for/affiliation.institution.author
}
\IEEEauthorblockA{Grenoble Alpes University\\
Grenoble, France
  %% -- beg for/affiliation.institution.author
\\anna.tcherkassof@univ-grenoble-alpes.fr
 %% -- end for/affiliation.institution.author
}
 %% -- end for/affiliation.institution
 %% -- end if/affiliation.institution-columnar
%% ----------------------------------------------------------

%% ---- one column per author, classic/default IEEETrans ----
 %% -- end if/affiliation.institution-columnar
%% ----------------------------------------------------------

}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
%
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3},
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The abstract goes here. On multiple lines eventually.
\end{abstract}

% no keywords

% use for special paper notices

% make the title area
\maketitle

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

With the developement of commercial automatic facial expression
recognition tools, industries and governments are gradually implementing
this technology in order to track humans' emotions in various scenarios
(e.g., marketing, heathcare, automotive to name a few). Beside the
ethical question of measuring human emotions, these tools are
challenging psychological theories about infering emotions from facial
expressions. When automatic facial expression recognition tools are used
to measure human emotions, they rest on the premise that facial
expressions provide a direct access to expressers' emotions. However the
link between emotion felt and facial expressions is far from being
established and still remain a hot topic in psychology research.

\hypertarget{the-link-between-emotion-and-facial-expression-thought-the-behaviorist-approach}{%
\subsection{The link between emotion and facial expression thought the
behaviorist
approach}\label{the-link-between-emotion-and-facial-expression-thought-the-behaviorist-approach}}

Based on the behaviorist approach initiated by Darwin in \emph{The
Expression of the Emotions in Man and Animals} (Darwin 1872), facial
expression are conceived as a genuine display of individuals inner
emotional state. This hypothesis is used as a basis for the Basic
Emotion Theory (BET) which states that a set of six or seven emotions
are universally displayed and are genetically hardwired (Ekman 1992).
According to this view, ``when emotions are aroused by perception of a
social event, a set of central commands produce patterned
emotion-specific changes in multiple systems, including {[}\ldots{}{]}
facial expressions.'' (Ekman 2007, p49). Even if this theory obtained a
popular support, it fails to explain how individuals can feel emotions
without expressing them and how individuals can express emotions without
feeling them.

\hypertarget{the-link-between-emotion-and-facial-expression-thought-the-social-constructivist-approach}{%
\subsection{The link between emotion and facial expression thought the
social constructivist
approach}\label{the-link-between-emotion-and-facial-expression-thought-the-social-constructivist-approach}}

Detractors of the Basic Emotion Theory are perceiving emotion not as
genetically hardwired but as a learnt association between a given
situation and an appropriate response (Averill 1980). For the tenants of
the constructivist approach, emotions are ``concepts'' based on past
experiences and which are ``a collection of embodied, whole brain
representations that predict what is about to happen in the sensory
environment, what the best action is to deal with impending events, and
their consequences for allostasis'' (Barrett 2017, p12). Following this
assumption, faces are used as tools to display signals in social
interactions (Crivelli and Fridlund 2018). These signals can convey
individuals' motivations and readiness (Frijda and Tcherkassof 1997) or
social messages (Fridlund and Rosenberg 1995).

This current paper investigates the link between the subjective feeling
of emotions and their recognition from facial expressions. If emotions
are hardwired, individuals emotional subjective feeling should be
correlated to the recognition of facial expressions from both human
observers and automatic classifiers whereas if emotions are social
constructs, no correlation between subjective feeling and facial
expression recognition should be observed.

\hypertarget{method}{%
\section{Method}\label{method}}

To evaluate the link between subjective feeling of emotions and their
recognition from facial expressions, participants were recruited to
perform an emotion elicitation task while their facial expression was
video recorded. Then, the videos was shown to human observer and
analysed by an automatic classifier in order to identify which emotion
was displayed.

\hypertarget{emotion-elicitation}{%
\subsection{Emotion Elicitation}\label{emotion-elicitation}}

For the emotion elicitation experiment, 358 French participants (182
females, 176 males, \emph{M}age = 47.9, \emph{SD}age = 9.2) were
recruited to perform one out of 11 emotions eliciation tasks designed to
trigger a positive, negative or neutral emotional state (see Tcherkassof
et al. 2013 for a description of tasks and procedure). Participants'
face were recorded using an hidden camera resulting 358 front facing
768x576 videos varying from 1s to 1479s.

After their emotion eliciation task the participants had to rate their
emotional state during the task on a likert scale from 0 (``not at
all'') to 5 (``strongly'') the six ``basic'' emotions (i.e.,
\emph{Anger}, \emph{Disgust}, \emph{Fear}, \emph{Happiness},
\emph{Surprise} and \emph{Sadness}) as well as six ``non-basic''
emotions (i.e., \emph{Pride}, \emph{Curiosity}, \emph{Boredom},
\emph{Shame}, \emph{Humiliation}, and \emph{Disappointement}). The
emotion scales rated as the highest is considered as describing
participants subjective feeling.

Finally, a debriefing session was perform to emsure that participants
were not durably affected by the emotion eliciation task. All the
participants gave their agreement on their data and video to be
processed for research puropose only.

\hypertarget{human-facial-expression-recognition}{%
\subsection{Human Facial Expression
Recognition}\label{human-facial-expression-recognition}}

For the human facial expression recognition method, 1383 student
participants were recruited to annotate 232 out of the 358 video,
therefore only the 232 annotated videos will be analysed in this paper.
Each participants had to annotate between 1 and 177 videos resulting
that ach video was annotated 29 times on average (\emph{SD} = 12).

The annotation of facial expressions was performed on-site using
\emph{Oudjat}, a software for designing video annotation experiements
(Dupré et al. 2015). For each video, the annotation procedure hat two
steps. First, the participants had to identify the emotional sequences
by pressing the space bar of their keyboard to indicate the begining and
the end of the emotional sequences while watching the video. Second, the
participants watched each emotional sequence previously identified and
had to label the sequence using one of the 12 emotions proposed
including six ``basic'' emotions (i.e., \emph{Anger}, \emph{Disgust},
\emph{Fear}, \emph{Happiness}, \emph{Surprise} and \emph{Sadness}) and
six ``non-basic'' emotions (i.e., \emph{Pride}, \emph{Curiosity},
\emph{Boredom}, \emph{Shame}, \emph{Humiliation}, and
\emph{Disappointement}). They also had the possibility to indicate that
the sequence was not expressing one of the proposed emotion.

This annotation procedure results in a uni-dimesional time-serie for
each video per human observer identifying for each second of the video
which emotion was recognized. Then time-series corresponding to the same
video were aggreated to calculate the proprotion of human observers for
each second of the video per emotional label. The sum of each label
proportion per second was used as a score to determine which labels
corresponds to the overall video (i.e., the highest score).

\hypertarget{automatic-facial-expression-recognition}{%
\subsection{Automatic Facial Expression
Recognition}\label{automatic-facial-expression-recognition}}

The 232 annotated video were processed with Affdex (SDK v3.4.1). Affdex
is an automatic facial expression recognition classifier developed and
distributed by Affectiva is a spinoff company resulting from the
research activities of MIT media lab created in 2009 (McDuff et al.
2016). Affdex's algorithm uses Histogram of Oriented Gradient (HOG)
features and Support Vector Machine (SVM) classifiers in order to
recognize facial expressions. For each video frame, Affdex identify the
probability of the face as expressing one of the six ``basic'' emotions
(i.e., \emph{Anger}, \emph{Disgust}, \emph{Fear}, \emph{Happiness},
\emph{Surprise} and \emph{Sadness}) as well as additional psychological
states such as \emph{Valence}, \emph{Engagement} or \emph{Contempt}, and
facial features such as \emph{Cheek Raise}, \emph{Eye Widen} or
\emph{Jaw Drop}.

To determine which of the six ``basic'' emotion can be used to identify
each video, the recognition probability for each label by frame was
converted into odd ratio by frame (Dente et al. 2017). The highest sum
of each odd ratio time-serie defines the label recognized by the
automatic classifier.

\hypertarget{results}{%
\section{Results}\label{results}}

Whereas the self-reports, the human annotations and the automatic
recognition include data on ``non-basic'' emotions and features, the
analysis is performed using only the six ``basic'' emotions in order to
compare them.

\hypertarget{correlation-between-self-report-and-human-facial-expression-recognition}{%
\subsection{Correlation between self-report and human facial expression
recognition}\label{correlation-between-self-report-and-human-facial-expression-recognition}}

\hypertarget{correlation-between-self-report-and-automatic-facial-expression-recognition}{%
\subsection{Correlation between self-report and automatic facial
expression
recognition}\label{correlation-between-self-report-and-automatic-facial-expression-recognition}}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

The conclusion goes here.

\hypertarget{acknowledgment}{%
\section{Acknowledgment}\label{acknowledgment}}

The authors would like to thank Brigitte Meillon and Jean Michel Adam
who developed the software used to collect and preprocess human observer
annotations.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-averill1980constructivist}{}%
Averill, James R. 1980. ``A Constructivist View of Emotion.'' In
\emph{Theories of Emotion}, 305--39. Elsevier.

\leavevmode\hypertarget{ref-barrett2017theory}{}%
Barrett, Lisa Feldman. 2017. ``The Theory of Constructed Emotion: An
Active Inference Account of Interoception and Categorization.''
\emph{Social Cognitive and Affective Neuroscience} 12 (1): 1--23.

\leavevmode\hypertarget{ref-crivelli2018facial}{}%
Crivelli, Carlos, and Alan J Fridlund. 2018. ``Facial Displays Are Tools
for Social Influence.'' \emph{Trends in Cognitive Sciences} 22 (5):
388--99.

\leavevmode\hypertarget{ref-darwin1872expression}{}%
Darwin, Charles. 1872. \emph{The Expression of the Emotions in Man and
Animals}. London, UK: John Murray, London UK.

\leavevmode\hypertarget{ref-dente2017measures}{}%
Dente, Pasquale, Dennis Küster, Lina Skora, and E Krumhuber. 2017.
``Measures and Metrics for Automatic Emotion Classification via Facet.''
In \emph{Proceedings of the Conference on the Study of Artificial
Intelligence and Simulation of Behaviour (Aisb)}, 160--63.

\leavevmode\hypertarget{ref-dupre2015oudjat}{}%
Dupré, Damien, Daniel Akpan, Elena Elias, Jean-Michel Adam, Brigitte
Meillon, Nicolas Bonnefond, Michel Dubois, and Anna Tcherkassof. 2015.
``Oudjat: A Configurable and Usable Annotation Tool for the Study of
Facial Expressions of Emotion.'' \emph{International Journal of
Human-Computer Studies} 83: 51--61.

\leavevmode\hypertarget{ref-ekman1992argument}{}%
Ekman, Paul. 1992. ``An Argument for Basic Emotions.'' \emph{Cognition
\& Emotion} 6 (3-4): 169--200.

\leavevmode\hypertarget{ref-ekman2007directed}{}%
---------. 2007. ``The Directed Facial Action Task.'' \emph{Handbook of
Emotion Elicitation and Assessment} 47: 53.

\leavevmode\hypertarget{ref-fridlund1995human}{}%
Fridlund, Alan J, and Erika L Rosenberg. 1995. ``Human Facial
Expression: An Evolutionary View.'' \emph{Nature} 373 (6515): 569--69.

\leavevmode\hypertarget{ref-frijda1997facial}{}%
Frijda, Nico H, and Anna Tcherkassof. 1997. ``Facial Expressions as
Modes of Action Readiness.'' \emph{The Psychology of Facial Expression},
78--102.

\leavevmode\hypertarget{ref-mcduff2016affdex}{}%
McDuff, Daniel, Abdelrahman Mahmoud, Mohammad Mavadati, May Amr, Jay
Turcot, and Rana el Kaliouby. 2016. ``AFFDEX Sdk: A Cross-Platform
Real-Time Multi-Face Expression Recognition Toolkit.'' In
\emph{Proceedings of the 2016 Chi Conference Extended Abstracts on Human
Factors in Computing Systems}, 3723--6. ACM.

\leavevmode\hypertarget{ref-tcherkassof2013dynemo}{}%
Tcherkassof, Anna, Damien Dupré, Brigitte Meillon, Nadine Mandran,
Michel Dubois, and Jean-Michel Adam. 2013. ``DynEmo: A Video Database of
Natural Facial Expressions of Emotions.'' \emph{The International
Journal of Multimedia \& Its Applications} 5 (5): 61--80.

\end{document}

### 2019-03-14 script/data_wangling_self_report.R

@@ -2,10 +2,11 @@
 library(here)
 library(tidyverse)
 library(data.table)
+library(janitor)
 # upload data ------------------------------------------------------------------
-data_sr <- read_rds(here::here("data/human_self_report_data/Base_Finale_20090114.rds"))
+self_report_data <- readr::read_rds(here::here("data/human_self_report_data/Base_Finale_20090114.rds"))
 
-test <- data_sr %>%
+test <- self_report_data %>%
   dplyr::select(
     "N° sujet__C",
     "Induction_C",
@@ -23,6 +24,10 @@ test <- data_sr %>%
     "DECU",
     "AUTRE"
   )
+
+test <- test %>% 
+  janitor::clean_names()
+
 
 test2 <- test %>%
   tidyr::gather(
@@ -92,3 +97,7 @@ test5 %>%
   theme_minimal() +
   theme(text = element_text(family = "serif", size=8))
   
+data_sr <- data_sr %>% 
+  dplyr::filter(!is.na(`N° sujet__C`))
+
+write_rds(data_sr, here::here("data/human_self_report_data/Base_Finale_20090114.rds"))

### 2019-03-14 script/data_wrangling_human_recognition.R

@@ -1,86 +1,5 @@
 library(tidyverse)
 library(splitstackshape)
-# ##############################################################################
-# test <- data.frame(ID = c("a","b","c"),
-#                    start_time = c(0,20,50),
-#                    stop_time = c(10,35,60))
-# timeline <- data.frame(timeline = 0:100)
-# 
-# z1 <- data.frame(
-#   time = seq(from = test$start_time[1], to = test$stop_time[1], by = 1),
-#   label = 1)
-# 
-# z2 <- seq(from = test$start_time[2], to = test$stop_time[2], by = 1)
-# 
-# ##############################################################################
-# test2 <- test %>% 
-#   mutate(timeline = map2(start_time, stop_time, `:`)) %>% 
-#   select(-start_time, -stop_time) %>% 
-#   unnest
-# 
-# test3 <- dplyr::full_join(timeline,test2,by = "timeline")
-# ##############################################################################
-# data <- read.csv2("C:/Users/dupred/Desktop/DynEmo_human_recognition/Data_Pretraitees_03_11/Data_Pretraitees_03_11/final_2.csv")
-# 
-# head_data <- head(data) %>% 
-#   dplyr::mutate(C_Video = stringr::str_replace(C_Video,".mpg",""))
-# 
-# str(head_data)
-# 
-# head_data <- dplyr::left_join(head_data,metadata_video, by = "C_Video")
-# 
-# str(metadata_video)
-# 
-# 
-# test4 <- head_data %>% 
-#   mutate(timeline = map2(TC_Debut, TC_Fin, `:`)) %>% 
-#   select(-TC_Debut, -TC_Fin) %>% 
-#   unnest
-# # DVD10_1 = 259.16032s -> 259160.32ms 
-# 
-# timeline <- data.frame(timeline = 0:259160)
-# 
-# test6 <- dplyr::full_join(timeline,test4,by = "timeline")
-# 
-# 
-# ##########################
-# 
-# unique_ppt <- head_data %>% 
-#   dplyr::select(-E_Detectee, -TC_Debut, -TC_Fin) %>% 
-#   unique()
-# 
-# test <- unique_ppt[rep(1:nrow(unique_ppt), unique_ppt$ffprobe_duration),] %>% 
-#   dplyr::mutate(timeline = 1:unique_ppt$ffprobe_duration)
-# 
[…]

### 2019-03-14 script/get_video_duration.R

@@ -28,5 +28,6 @@ metadata_video <- data.frame(list_video, ffprobe_duration) %>%
   dplyr::mutate(ffprobe_duration = round(ffprobe_duration,0)) %>% 
   dplyr::select(-list_video)
 
+write_rds(metadata_video, "metadata_video.rds")
 
 

### 2019-03-15 ACII_2019_paper/ACII_2019_paper.Rmd
acii paper

@@
   On multiple lines eventually.
-
+header-includes:
+  - \usepackage{booktabs}
+  - \usepackage{float}
+  - \usepackage{tabu}
 bibliography: mybibfile.bib
@@
   echo = FALSE,
-  fig.align="center",
-  fig.pos = 'H')
+  fig.pos='H')
 #
@@
 #
+library(kableExtra)
 library(tidyverse) # Data wrangling & grammar of graphics
@@
 # self_report_data
-self_report_data <- readr::read_rds(here::here("data/human_self_report_data/Base_Finale_20090114.rds"))
-# metadata_video
-metadata_video <- readr::read_rds(here::here("metadata_video.rds"))
+source(here::here("script/data_wrangling_self_report.R"))
+self_report_score <- self_report_score %>% 
+  dplyr::select(C_Video, sr_emotion = emotion, sr_score = value)
 # human_recognition_data
 source(here::here("script/data_wrangling_human_recognition.R"))
+human_recognition_score <- human_recognition_score %>% 
+  dplyr::select(C_Video, hr_emotion = emotion, hr_score = confidence_score)
+# automatic_recognition_data
+source(here::here("script/data_wrangling_automatic_recognition.R"))
+automatic_recognition_score <- automatic_recognition_score %>% 
+  dplyr::select(C_Video, ar_emotion = emotion, ar_score = confidence_score) %>% 
+  dplyr::filter(C_Video %in% ppt_per_video$C_Video)
 ```
@@
 =============
+```{r}
+# note to myself: instead of calculating the proportion on all the video, calculate the marginal proportion
+```
 
-With the developement of commercial automatic facial expression recognition tools, industries and governments are gradually implementing this technology in order to track humans' emotions in various scenarios (e.g., marketing, heathcare, automotive to name a few).
+With the developement of commercial automatic facial expression recognition tools [see @dupre2018accuracy for a non-exhaustive list of available tools], industries and governments are gradually implementing this technology in order to track humans' emotions in various scenarios (e.g., marketing, heathcare, automotive to name a few).
 Beside the ethical question of measuring human emotions, these tools are challenging psychological theories about infering emotions from facial expressions.
@@
 self_report_gender <- self_report_data %>% 
-  dplyr::group_by(Genre_C) %>% 
+  dplyr::group_by(genre_c) %>% 
   dplyr::summarise(n= n())
@@
 self_report_age <- self_report_data %>%
-  dplyr::summarise(Age_m = mean(Age), Age_sd = sd(Age)) %>% 
+  dplyr::summarise(Age_m = mean(age), Age_sd = sd(age)) %>% 
   dplyr::mutate_all(round,1)
@@
 
-For the emotion elicitation experiment, `r nrow(self_report_data)` French participants (`r self_report_gender[self_report_gender$Genre_C == "F","n"]` females, `r self_report_gender[self_report_gender$Genre_C == "H","n"]` males, *M*age = `r self_report_age[,"Age_m"]`, *SD*age = `r self_report_age[,"Age_sd"]`) were recruited to perform one out of 11 emotions eliciation tasks designed to trigger a positive, negative or neutral emotional state [see @tcherkassof2013dynemo for a description of tasks and procedure].
+For the emotion elicitation experiment, `r nrow(self_report_data)` French participants (`r self_report_gender[self_report_gender$genre_c == "F","n"]` females, `r self_report_gender[self_report_gender$genre_c == "H","n"]` males, *M*age = `r self_report_age[,"Age_m"]`, *SD*age = `r self_report_age[,"Age_sd"]`) were recruited to perform one out of 11 emotions eliciation tasks designed to trigger a positive, negative or neutral emotional state [see @tcherkassof2013dynemo for a description of tasks and procedure].
 Participants' face were recorded using an hidden camera resulting `r nrow(self_report_data)` front facing 768x576 videos varying from `r min(metadata_video$ffprobe_duration)`s to `r max(metadata_video$ffprobe_duration)`s.
 
-After their emotion eliciation task the participants had to rate their emotional state during the task on a likert scale from 0 ("not at all") to 5 ("strongly") the six "basic" emotions (i.e., *Anger*, *Disgust*, *Fear*, *Happiness*, *Surprise* and *Sadness*) as well as six "non-basic" emotions (i.e., *Pride*, *Curiosity*, *Boredom*, *Shame*, *Humiliation*, and *Disappointement*).
-The emotion scales rated as the highest is considered as describing participants subjective feeling.
+After their emotion eliciation task the participants had to rate their emotional state during the task on a likert scale from 0 ("not at all") to 5 ("strongly") the six "basic" emotions (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as six "non-basic" emotions (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointement*).
 
@@
 First, the participants had to identify the emotional sequences by pressing the space bar of their keyboard to indicate the begining and the end of the emotional sequences while watching the video.
-Second, the participants watched each emotional sequence previously identified and had to label the sequence using one of the 12 emotions proposed including six "basic" emotions (i.e., *Anger*, *Disgust*, *Fear*, *Happiness*, *Surprise* and *Sadness*) and six "non-basic" emotions (i.e., *Pride*, *Curiosity*, *Boredom*, *Shame*, *Humiliation*, and *Disappointement*).
+Second, the participants watched each emotional sequence previously identified and had to label the sequence using one of the 12 emotions proposed including six "basic" emotions (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) and six "non-basic" emotions (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointement*).
 They also had the possibility to indicate that the sequence was not expressing one of the proposed emotion.
@@
 The sum of each label proportion per second was used as a score to determine which labels corresponds to the overall video (i.e., the highest score).
+In case of more than one label having the maximum value, the emotion is discribed as undetermined.
 
@@
 Affdex’s algorithm uses Histogram of Oriented Gradient (HOG) features and Support Vector Machine (SVM) classifiers in order to recognize facial expressions.
-For each video frame, Affdex identify the probability of the face as expressing one of the six "basic" emotions (i.e., *Anger*, *Disgust*, *Fear*, *Happiness*, *Surprise* and *Sadness*) as well as additional psychological states such as *Valence*, *Engagement* or *Contempt*, and facial features such as *Cheek Raise*, *Eye Widen* or *Jaw Drop*.
+For each video frame, Affdex identify the probability of the face as expressing one of the six "basic" emotions (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as additional psychological states such as *valence*, *engagement* or *contempt*, and facial features such as *cheek raise*, *eye widen* or *jaw drop*.
 
@@
 Whereas the self-reports, the human annotations and the automatic recognition include data on "non-basic" emotions and features, the analysis is performed using only the six "basic" emotions in order to compare them.
+The maximum score for self-reports, human annotations and automatic recognition is used to label the video.
+In case of more than one label obtaining the maximum value, the video is labeled as undetermined.
 
@@
 
+```{r sr_hr}
+comparison_sr_hr <- dplyr::inner_join(self_report_score,human_recognition_score, by = "C_Video") %>% 
+  dplyr::ungroup() %>% 
+  dplyr::mutate_if(is.character,as.factor)
+
+confusionMatrix_sr_hr <- caret::confusionMatrix(
+  data = comparison_sr_hr$hr_emotion,
+  reference = comparison_sr_hr$sr_emotion
+  )
+```
+
+Emotions self-reported as being characteristic of the elicitation are compared with the emotion recognized by the human observers in a confusion matrix (Figure \ref{fig:confusionMatrix_sr_hr}).
+
+```{r confusionMatrix_sr_hr, fig.height=4, fig.cap="Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the human observers."}
+confusionMatrix_sr_hr_freq <- confusionMatrix_sr_hr$table %>% 
+  as.data.frame() %>% 
+  dplyr::rename(hr_emotion = Prediction, sr_emotion = Reference) %>% 
+  dplyr::mutate(prop_video = Freq/sum(Freq))
+
+confusionMatrix_sr_hr_freq %>% 
+  ggplot(mapping = aes(x = sr_emotion, y = hr_emotion)) +
+  geom_tile(aes(fill = Freq), colour = "white") +
+  geom_text(
+    aes(label = Freq),
+    color = "black",
+    family="serif",
+    size = 4,
+    parse = FALSE, 
+    lineheight = 0.7
+  ) +
+  scale_fill_gradient(
+    name = "Frequency",
+    low = "white", 
+    high = "red", 
+    ) +
+  scale_x_discrete(name = "Self-Reports") +
+  scale_y_discrete(name = "Human Observers") +
+  theme_minimal() +
+  theme(text = element_text(size=16,family="serif"),
+        axis.text.x = element_text(size=16, angle = 45, hjust = 0.75,vjust=0.9),
+        axis.text.y = element_text(size=16, hjust = 0.5),
+        axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
+        legend.text = element_text(size=8),
+        legend.position = "none")
+```
+
+The result of the confusion matrix show a low agreement between emotion felt during the elecitation and emotion recognized by the human annotators (Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "happiness", "prop_video"])`), *surprise* (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "surprise", "prop_video"])`) and *disgust* (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_video"])`).
+Sensitivity, specificity, precision and F1 score for each emotion can be found Table \ref{tab:confusionTable_sr_hr}.
+Interestingly human annotators seem to recognize as *surprise* videos in which *happiness* was the highest self-reported emotion (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "happiness", "prop_video"])`), and in a lower instance *happiness* videos in which *surprise* was the highest self-reported emotion (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "surprise", "prop_video"])`).
+
+```{r confusionTable_sr_hr, results="asis"}
+confusionMatrix_sr_hr_table <- confusionMatrix_sr_hr %>%
+  magrittr::use_series(byClass) %>%
+  as.data.frame() %>%
+  tibble::rownames_to_column("Emotion") %>%
+  dplyr::mutate(Emotion = gsub(pattern = "Class: ",replacement = "",x = Emotion)) %>% 
+  dplyr::select(Emotion, Sensitivity, Specificity, Precision, F1) %>% 
+  dplyr::mutate_if(is.numeric,round,2) %>% 
+  replace(is.na(.), "\\textit{na.}")
+
+confusionMatrix_sr_hr_table %>%
+  knitr::kable(
+    "latex",
+    caption = "Agreement accuracy metrics for each emotion. ",
+    booktabs = TRUE,
+    digits = 2,
+    linesep = "",
+    escape = FALSE,
+    align = c('l',rep('c', 4))
+  ) %>%
+  kableExtra::kable_styling(
+    font_size = 8,
+    full_width = TRUE,
+    latex_options = c("hold_position")) %>%
+   add_footnote("Note. \\textit{na.} values are produced when not enough data are available to compute accuracy indicators.", notation="none", escape = FALSE)
+```
+
+However, the self-report show a very high proportion of undetermined emotional states which reveals not only the possiblility of the emotion elicitation tasks to trigger more than one emotion but also the potential limit of using 6-points likert scales for which the participants can easily score to the maximum for more than one emotion.
+
 ## Correlation between self-report and automatic facial expression recognition
+
+```{r sr_ar}
+comparison_sr_ar <- dplyr::inner_join(
+    self_report_score, 
+    automatic_recognition_score, 
+    by = "C_Video") %>% 
+  dplyr::ungroup() %>% 
+  dplyr::mutate_if(is.character,as.factor)
+
+dat_table <- table(
+  comparison_sr_ar$ar_emotion,
+  comparison_sr_ar$sr_emotion
+)
+emotions <- c("anger","disgust","fear","happiness","sadness","surprise", "undetermined")
+emo_table <- matrix(
+  0,
+  length(emotions),
+  length(emotions),
+  dimnames = list(emotions, emotions))
+emo_table[row.names(dat_table), colnames(dat_table)] <- dat_table
+
+confusionMatrix_sr_ar <- caret::confusionMatrix(as.table(emo_table))
+```
+
+As in the previous analysis, emotions self-reported as being characteristic of the elicitation are compared with the emotion recognized by the automatic classifier in a confusion matrix (Figure \ref{fig:confusionMatrix_sr_ar}).
+
+```{r confusionMatrix_sr_ar, fig.height=4, fig.cap="Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the automatic classifier."}
+confusionMatrix_sr_ar_freq <- confusionMatrix_sr_ar$table %>% 
+  as.data.frame() %>% 
+  dplyr::rename(ar_emotion = Var1, sr_emotion = Var2) %>% 
+  dplyr::mutate(prop_video = Freq/sum(Freq))
+
+confusionMatrix_sr_ar_freq %>% 
+  ggplot(mapping = aes(x = sr_emotion, y = ar_emotion)) +
+  geom_tile(aes(fill = Freq), colour = "white") +
+  geom_text(
+    aes(label = Freq),
+    color = "black",
+    family="serif",
+    size = 4,
+    parse = FALSE, 
+    lineheight = 0.7
+  ) +
+  scale_fill_gradient(
+    name = "Frequency",
+    low = "white", 
+    high = "red", 
+    ) +
+  scale_x_discrete(name = "Self-Reports") +
+  scale_y_discrete(name = "Automatic Classifier") +
+  theme_minimal() +
+  theme(text = element_text(size=16,family="serif"),
+        axis.text.x = element_text(size=16, angle = 45, hjust = 0.75,vjust=0.9),
+        axis.text.y = element_text(size=16, hjust = 0.5),
+        axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
+        legend.text = element_text(size=8),
+        legend.position = "none")
+```
+
+Results obtained for the comparison between emotions self-reported and recognised by the automatic classifier are similar to the ones with human observers (Table \ref{tab:confusionTable_sr_ar}).
+Overall there is a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_ar$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_ar$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_ar$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_video"])`) and *surprise* (`r scales::percent(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "surprise" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_video"])`).
+Surprisingly the automatic classifier incorrectly recognized as *disgust* an important proportion of videos in which *happiness* was the highest self-reported emotion (`r scales::percent(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_video"])`).
+In parallel, the automatic classifier recognised as *happiness* and *disgust* videos in which *surprise* was the highest self-reported emotion (respectively `r scales::percent(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_video"])` and `r scales::percent(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_video"])`).
+
+```{r confusionTable_sr_ar, results="asis"}
+confusionMatrix_sr_ar_table <- confusionMatrix_sr_ar %>%
+  magrittr::use_series(byClass) %>%
+  as.data.frame() %>%
+  tibble::rownames_to_column("Emotion") %>%
+  dplyr::mutate(Emotion = gsub(pattern = "Class: ",replacement = "",x = Emotion)) %>% 
+  dplyr::select(Emotion, Sensitivity, Specificity, Precision, F1) %>% 
+  dplyr::mutate_if(is.numeric,round,2) %>% 
+  replace(is.na(.), "\\textit{na.}")
+
+confusionMatrix_sr_ar_table %>%
+  knitr::kable(
+    "latex",
+    caption = "Agreement accuracy metrics for each emotion.",
+    booktabs = T,
+    digits = 2,
+    linesep = "",
+    escape = FALSE,
+    align = c('l',rep('c', 4))
+  ) %>%
+  kableExtra::kable_styling(
+    font_size = 8,
+    full_width = TRUE,
+    latex_options = c("hold_position")) %>%
+   add_footnote("Note. \\textit{na.} values are produced when not enough data are available to compute accuracy indicators.", notation="none", escape = FALSE)
+```
+
+A comparable explaination can be provided as the level of undetermined emotion are very high for the self reports.
 
@@
 ============
-The conclusion goes here.
+Despit being one on the most investigated question in affective science, the link between emotion felt and facial expression is a hot topic and no clear evidence have been found to definively answer it.
+However, with the growing interest of industries and govenment to monitor individual's psychological states, evidences are showing that facial expressions are in reality not expressing emotions [@mckeown2013analogical].
+This research aimed to provide some empirical data to the question.
+The subjective feeling of participants was compared with human recognition on one side and automatic recognition on the other side.
+The results reveals a low accuracy for both humans and automatic classifier to accurately identify the inner emotional states of these individuals based on their facial expressions.
+
+Some limitations to this process should be stated over the use of self-reports to evalute individual's subjective feelings.
+Accessing to the inner subjective feeling can be biaised if not impossible.
+Moreover the laboratory setting can trigger ambiguous and "non-basic" emotions which were not analysed in this research.
+The procedure use for human annotation can also be incriminated.
+Instead of asking the human annotators to provide an unique label, a more sutble approach was choosen to mimic results provided by the automatic classifier.
+In this regard, the results of the human annotation could have been more ambiguous because it is not the natural way that people are infering human emotions.
+Finaly, the automatic classifer algorithm can also be problematic.
+Based on training datasets which are most of the time using prototypical facial expression of the "basic" emotions, the algorithm to classify facial expressions can be held in check by the spontanous facial expressions analysed.
+
+Considering the above, the results provides an additional evidence that individuals' subjective feeling can not be infered from facial expressions and in our case invalidate the hypothesis of hardwired emotions.
+This result suggests that automatic facial expression recognition tools should be focused on evaluating facial morphology features such as action units rather than infering potentional emotional or affective states.
 

### 2019-03-15 ACII_2019_paper/ACII_2019_paper.tex
acii paper

@@
 
+\usepackage{graphicx}
+% We will generate all images so they have a width \maxwidth. This means
+% that they will get their normal width if they fit onto the page, but
+% are scaled down if they would overflow the margins.
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
+\else\Gin@nat@width\fi}
+\makeatother
+\let\Oldincludegraphics\includegraphics
+\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
+
 \usepackage[unicode=true]{hyperref}
@@
 % Pandoc header
+\usepackage{booktabs}
+\usepackage{float}
+\usepackage{tabu}
 
@@
 With the developement of commercial automatic facial expression
-recognition tools, industries and governments are gradually implementing
+recognition tools (see Dupré et al. 2018 for a non-exhaustive list of
+available tools), industries and governments are gradually implementing
 this technology in order to track humans' emotions in various scenarios
@@
 all'') to 5 (``strongly'') the six ``basic'' emotions (i.e.,
-\emph{Anger}, \emph{Disgust}, \emph{Fear}, \emph{Happiness},
-\emph{Surprise} and \emph{Sadness}) as well as six ``non-basic''
-emotions (i.e., \emph{Pride}, \emph{Curiosity}, \emph{Boredom},
-\emph{Shame}, \emph{Humiliation}, and \emph{Disappointement}). The
-emotion scales rated as the highest is considered as describing
-participants subjective feeling.
+\emph{anger}, \emph{disgust}, \emph{fear}, \emph{happiness},
+\emph{surprise} and \emph{sadness}) as well as six ``non-basic''
+emotions (i.e., \emph{pride}, \emph{curiosity}, \emph{boredom},
+\emph{shame}, \emph{humiliation}, and \emph{disappointement}).
 
@@
 had to label the sequence using one of the 12 emotions proposed
-including six ``basic'' emotions (i.e., \emph{Anger}, \emph{Disgust},
-\emph{Fear}, \emph{Happiness}, \emph{Surprise} and \emph{Sadness}) and
-six ``non-basic'' emotions (i.e., \emph{Pride}, \emph{Curiosity},
-\emph{Boredom}, \emph{Shame}, \emph{Humiliation}, and
-\emph{Disappointement}). They also had the possibility to indicate that
+including six ``basic'' emotions (i.e., \emph{anger}, \emph{disgust},
+\emph{fear}, \emph{happiness}, \emph{surprise} and \emph{sadness}) and
+six ``non-basic'' emotions (i.e., \emph{pride}, \emph{curiosity},
+\emph{boredom}, \emph{shame}, \emph{humiliation}, and
+\emph{disappointement}). They also had the possibility to indicate that
 the sequence was not expressing one of the proposed emotion.
@@
 proportion per second was used as a score to determine which labels
-corresponds to the overall video (i.e., the highest score).
+corresponds to the overall video (i.e., the highest score). In case of
+more than one label having the maximum value, the emotion is discribed
+as undetermined.
 
@@
 probability of the face as expressing one of the six ``basic'' emotions
-(i.e., \emph{Anger}, \emph{Disgust}, \emph{Fear}, \emph{Happiness},
-\emph{Surprise} and \emph{Sadness}) as well as additional psychological
-states such as \emph{Valence}, \emph{Engagement} or \emph{Contempt}, and
-facial features such as \emph{Cheek Raise}, \emph{Eye Widen} or
-\emph{Jaw Drop}.
+(i.e., \emph{anger}, \emph{disgust}, \emph{fear}, \emph{happiness},
+\emph{surprise} and \emph{sadness}) as well as additional psychological
+states such as \emph{valence}, \emph{engagement} or \emph{contempt}, and
+facial features such as \emph{cheek raise}, \emph{eye widen} or
+\emph{jaw drop}.
 
@@
 analysis is performed using only the six ``basic'' emotions in order to
-compare them.
+compare them. The maximum score for self-reports, human annotations and
+automatic recognition is used to label the video. In case of more than
+one label obtaining the maximum value, the video is labeled as
+undetermined.
 
@@
 recognition}\label{correlation-between-self-report-and-human-facial-expression-recognition}}
+
+Emotions self-reported as being characteristic of the elicitation are
+compared with the emotion recognized by the human observers in a
+confusion matrix (Figure \ref{fig:confusionMatrix_sr_hr}).
+
+\begin{figure}
+\centering
+\includegraphics{ACII_2019_paper_files/figure-latex/confusionMatrix_sr_hr-1.pdf}
+\caption{Confusion matrix of between the emotion self-reported as being
+characteristic of the elicitation with the emotion recognized by the
+human observers.}
+\end{figure}
+
+The result of the confusion matrix show a low agreement between emotion
+felt during the elecitation and emotion recognized by the human
+annotators (Accuracy = 0.27, 95\%CI{[}0.21,0.33{]}; Kappa = 0.11) except
+for \emph{happiness} (15.2\%), \emph{surprise} (8.26\%) and
+\emph{disgust} (1.74\%). Sensitivity, specificity, precision and F1
+score for each emotion can be found Table
+\ref{tab:confusionTable_sr_hr}. Interestingly human annotators seem to
+recognize as \emph{surprise} videos in which \emph{happiness} was the
+highest self-reported emotion (10.9\%), and in a lower instance
+\emph{happiness} videos in which \emph{surprise} was the highest
+self-reported emotion (4.78\%).
+
+\begin{table}[!h]
+
+\caption{\label{tab:confusionTable_sr_hr}Agreement accuracy metrics for each emotion. }
+\centering
+\fontsize{8}{10}\selectfont
+\begin{tabu} to \linewidth {>{\raggedright}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
+\toprule
+Emotion & Sensitivity & Specificity & Precision & F1\\
+\midrule
+anger & 0.43 & 0.92 & 0.14 & 0.21\\
+disgust & 1.00 & 0.93 & 0.20 & 0.33\\
+fear & 0.00 & 0.96 & 0.00 & \textit{na.}\\
+happiness & 0.49 & 0.73 & 0.45 & 0.47\\
+sadness & 0.08 & 0.91 & 0.05 & 0.06\\
+surprise & 0.44 & 0.67 & 0.23 & 0.31\\
+undetermined & 0.00 & 0.99 & 0.00 & \textit{na.}\\
+\bottomrule
+\multicolumn{5}{l}{\textsuperscript{} Note. \textit{na.} values are produced when not enough}\\
+\multicolumn{5}{l}{data are available to compute accuracy indicators.}\\
+\end{tabu}
+\end{table}
+
+However, the self-report show a very high proportion of undetermined
+emotional states which reveals not only the possiblility of the emotion
+elicitation tasks to trigger more than one emotion but also the
+potential limit of using 6-points likert scales for which the
+participants can easily score to the maximum for more than one emotion.
 
@@
 
+As in the previous analysis, emotions self-reported as being
+characteristic of the elicitation are compared with the emotion
+recognized by the automatic classifier in a confusion matrix (Figure
+\ref{fig:confusionMatrix_sr_ar}).
+
+\begin{figure}
+\centering
+\includegraphics{ACII_2019_paper_files/figure-latex/confusionMatrix_sr_ar-1.pdf}
+\caption{Confusion matrix of between the emotion self-reported as being
+characteristic of the elicitation with the emotion recognized by the
+automatic classifier.}
+\end{figure}
+
+Results obtained for the comparison between emotions self-reported and
+recognised by the automatic classifier are similar to the ones with
+human observers (Table \ref{tab:confusionTable_sr_ar}). Overall there is
+a low agreement between emotion self-reported and emotion recognized by
+the automatic classifier (Accuracy = 0.19, 95\%CI{[}0.14,0.24{]}; Kappa
+= 0.05) except for \emph{happiness} (13.9\%) and \emph{surprise}
+(3.48\%). Surprisingly the automatic classifier incorrectly recognized
+as \emph{disgust} an important proportion of videos in which
+\emph{happiness} was the highest self-reported emotion (7.83\%). In
+parallel, the automatic classifier recognised as \emph{happiness} and
+\emph{disgust} videos in which \emph{surprise} was the highest
+self-reported emotion (respectively 5.22\% and 6.09\%).
+
+\begin{table}[!h]
+
+\caption{\label{tab:confusionTable_sr_ar}Agreement accuracy metrics for each emotion.}
+\centering
+\fontsize{8}{10}\selectfont
+\begin{tabu} to \linewidth {>{\raggedright}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
+\toprule
+Emotion & Sensitivity & Specificity & Precision & F1\\
+\midrule
+anger & 0.00 & 0.89 & 0 & \textit{na.}\\
+disgust & 0.25 & 0.70 & 0.01 & 0.03\\
+fear & 0.00 & 0.93 & 0 & \textit{na.}\\
+happiness & 0.44 & 0.71 & 0.41 & 0.43\\
+sadness & 0.15 & 0.94 & 0.14 & 0.15\\
+surprise & 0.19 & 0.89 & 0.28 & 0.22\\
+undetermined & 0.00 & 1.00 & \textit{na.} & \textit{na.}\\
+\bottomrule
+\multicolumn{5}{l}{\textsuperscript{} Note. \textit{na.} values are produced when not enough data are}\\
+\multicolumn{5}{l}{available to compute accuracy indicators.}\\
+\end{tabu}
+\end{table}
+
+A comparable explaination can be provided as the level of undetermined
+emotion are very high for the self reports.
+
 \hypertarget{conclusion}{%
@@
 
-The conclusion goes here.
+Despit being one on the most investigated question in affective science,
+the link between emotion felt and facial expression is a hot topic and
+no clear evidence have been found to definively answer it. However, with
+the growing interest of industries and govenment to monitor individual's
+psychological states, evidences are showing that facial expressions are
+in reality not expressing emotions (McKeown 2013). This research aimed
+to provide some empirical data to the question. The subjective feeling
+of participants was compared with human recognition on one side and
+automatic recognition on the other side. The results reveals a low
+accuracy for both humans and automatic classifier to accurately identify
+the inner emotional states of these individuals based on their facial
+expressions.
+
+Some limitations to this process should be stated over the use of
+self-reports to evalute individual's subjective feelings. Accessing to
+the inner subjective feeling can be biaised if not impossible. Moreover
+the laboratory setting can trigger ambiguous and ``non-basic'' emotions
+which were not analysed in this research. The procedure use for human
+annotation can also be incriminated. Instead of asking the human
+annotators to provide an unique label, a more sutble approach was
+choosen to mimic results provided by the automatic classifier. In this
+regard, the results of the human annotation could have been more
+ambiguous because it is not the natural way that people are infering
+human emotions. Finaly, the automatic classifer algorithm can also be
+problematic. Based on training datasets which are most of the time using
+prototypical facial expression of the ``basic'' emotions, the algorithm
+to classify facial expressions can be held in check by the spontanous
+facial expressions analysed.
+
+Considering the above, the results provides an additional evidence that
+individuals' subjective feeling can not be infered from facial
+expressions and in our case invalidate the hypothesis of hardwired
+emotions. This result suggests that automatic facial expression
+recognition tools should be focused on evaluating facial morphology
+features such as action units rather than infering potentional emotional
+or affective states.
 
@@
 
+\leavevmode\hypertarget{ref-dupre2018accuracy}{}%
+Dupré, Damien, Nicole Andelic, Gawain Morrison, and Gary McKeown. 2018.
+``Accuracy of Three Commercial Automatic Emotion Recognition Systems
+Across Different Individuals and Their Facial Expressions.'' In
+\emph{2018 Ieee International Conference on Pervasive Computing and
+Communications Workshops (Percom Workshops)}, 627--32. IEEE.
+
 \leavevmode\hypertarget{ref-ekman1992argument}{}%
@@
 
+\leavevmode\hypertarget{ref-mckeown2013analogical}{}%
+McKeown, Gary J. 2013. ``The Analogical Peacock Hypothesis: The Sexual
+Selection of Mind-Reading and Relational Cognition in Human
+Communication.'' \emph{Review of General Psychology} 17 (3): 267--87.
+
 \leavevmode\hypertarget{ref-tcherkassof2013dynemo}{}%

### 2019-03-15 script/data_wangling_self_report.R

@@ -1,103 +1,0 @@
-# setup ------------------------------------------------------------------------
-library(here)
-library(tidyverse)
-library(data.table)
-library(janitor)
-# upload data ------------------------------------------------------------------
-self_report_data <- readr::read_rds(here::here("data/human_self_report_data/Base_Finale_20090114.rds"))
-
-test <- self_report_data %>%
-  dplyr::select(
-    "N° sujet__C",
-    "Induction_C",
-    "CURIOSITE/INTERET",
-    "ENNUI",
-    "REPULSION/DEGOUT",
-    "FRAYEUR",
-    "GAIETE/AMUSEMENT",
-    "HONTE",
-    "FIERTE/SATISFACTION",
-    "STUPEFACTION/SURPRISE",
-    "EMU",
-    "ENERVEE/IRRITATION",
-    "VEXE/VEXATION",
-    "DECU",
-    "AUTRE"
-  )
-
-test <- test %>% 
-  janitor::clean_names()
-
-
-test2 <- test %>%
-  tidyr::gather(
-    emotion,
-    value,
-    c(
-      "CURIOSITE/INTERET",
-      "ENNUI",
-      "REPULSION/DEGOUT",
-      "FRAYEUR",
-      "GAIETE/AMUSEMENT",
-      "HONTE",
-      "FIERTE/SATISFACTION",
-      "STUPEFACTION/SURPRISE",
-      "EMU",
-      "ENERVEE/IRRITATION",
-      "VEXE/VEXATION",
-      "DECU",
-      "AUTRE"
-    )
-  ) %>%
-  dplyr::group_by(`N° sujet__C`) %>%
-  dplyr::filter(value == max(value))
-
-ggplot(data = test2, mapping = aes(x = Induction_C, y = emotion)) +
-  geom_tile(aes(fill = value), colour = "white") +
-  scale_fill_gradient2(low = "blue",mid = "green",high = "red", midpoint = 4) +
-  theme_minimal() +
-  theme(text = element_text(family = "serif", size=8))
-
-test3 <- test2 %>%
-  dplyr::group_by(Induction_C) %>%
-  dplyr::count(emotion)
-
-unique(test3$Induction_C)
-
-list_emotion_sr <- c(
-  "CURIOSITE/INTERET",
-  "ENNUI",
-  "REPULSION/DEGOUT",
-  "FRAYEUR",
[…]

### 2019-03-15 script/data_wrangling_automatic_recognition.R

@@ -1,10 +1,40 @@
 library(tidyverse)
 
-data <- here::here("./data/automatic_recognition_data") %>% 
-  fs::dir_ls(regexp = "\\.csv$") %>% 
-  purrr::map_dfr(readr::read_csv, .id = "source") %>% 
-  dplyr::mutate(filename = base::basename(source))
+# automatic_recognition_data <- here::here("./data/automatic_recognition_data/automatic_recognition_data") %>% 
+#   fs::dir_ls(regexp = "\\.csv$") %>% 
+#   purrr::map_dfr(readr::read_csv, .id = "source") %>% 
+#   dplyr::mutate(filename = base::basename(source))
+# 
+# readr::write_rds(automatic_recognition_data, here::here("./data/automatic_recognition_data/automatic_recognition_data.rds"))
 
-readr::write_rds(data, here::here("./data/automatic_recognition_data.rds"))
+automatic_recognition_data <- readr::read_rds(here::here("./data/automatic_recognition_data/automatic_recognition_data.rds"))
 
-data <- readr::read_rds(here::here("./data/automatic_recognition_data.rds"))
+automatic_recognition_score <- automatic_recognition_data %>% 
+  dplyr::select(C_Video = filename, TimeStamp, happiness = joy, fear, disgust, sadness, anger, surprise) %>% 
+  dplyr::mutate(C_Video = stringr::str_replace(C_Video,".csv","")) %>% 
+  na.omit() %>% 
+  tidyr::gather(emotion, value, c("happiness", "fear", "disgust", "sadness", "anger", "surprise")) %>% 
+  dplyr::mutate(probabilities = value/100) %>% 
+  dplyr::mutate(odds = probabilities/(1-probabilities)) %>% 
+  dplyr::select(C_Video, TimeStamp, emotion, odds) %>% 
+  dplyr::group_by(C_Video, emotion) %>%
+  dplyr::summarise(sum_emotion = sum(odds)) %>% 
+  dplyr::group_by(C_Video) %>% 
+  dplyr::mutate(sum_video = sum(sum_emotion)) %>% 
+  dplyr::mutate(confidence_score = sum_emotion/sum_video) %>% 
+  dplyr::filter(confidence_score == max(confidence_score))
+
+list_ties <- automatic_recognition_score %>%
+  dplyr::group_by(C_Video) %>%
+  dplyr::summarise(n = n()) %>% 
[…]

### 2019-03-15 script/data_wrangling_human_recognition.R

@@ -1,5 +1,8 @@
+# setup ------------------------------------------------------------------------
 library(tidyverse)
 library(splitstackshape)
+# upload data ------------------------------------------------------------------
+metadata_video <- readr::read_rds(here::here("data/human_recognition_data/metadata_video.rds"))
 
 human_recognition_data <- here::here("./data/human_recognition_data") %>% 
   fs::dir_ls(regexp = "\\.csv$") %>% 
@@ -15,8 +18,10 @@ human_recognition_data <- here::here("./data/human_recognition_data") %>%
   )) %>% 
   dplyr::mutate(source = base::basename(source)) %>% 
   dplyr::mutate_at(c("TC_Debut", "TC_Fin"), function(x){round(x/1000,0)}) %>% # time form millisec to sec
-  dplyr::mutate(C_Video = stringr::str_replace(C_Video,".mpg",""))
+  dplyr::mutate(C_Video = stringr::str_replace(C_Video,".mpg","")) %>% 
+  dplyr::mutate(SEXE_Juge = ifelse(is.na(SEXE_Juge),"N",SEXE_Juge))
 
+# data wrangling ---------------------------------------------------------------
 unique_ppt_count <- human_recognition_data %>% 
   dplyr::select(source,SEXE_Juge, C_Juge) %>% 
   unique()
@@ -53,7 +58,7 @@ final_timeline <- dplyr::left_join(full_timeline,data_timeline, by = c("source",
   dplyr::mutate(Ennui = ifelse(E_Detectee == "Ennui",1,0)) %>% 
   dplyr::mutate(Fierte = ifelse(E_Detectee == "Fierte",1,0)) %>% 
   dplyr::mutate(Frayeur = ifelse(E_Detectee == "Frayeur",1,0)) %>% 
-  dplyr::mutate(Gaiete  = ifelse(E_Detectee == "Gaiete ",1,0)) %>% 
+  dplyr::mutate(Gaiete  = ifelse(E_Detectee == "Gaiete",1,0)) %>% 
   dplyr::mutate(Honte = ifelse(E_Detectee == "Honte",1,0)) %>% 
   dplyr::mutate(RAS = ifelse(E_Detectee == "RAS",1,0)) %>% 
   dplyr::mutate(Stupefaction = ifelse(E_Detectee == "Stupefaction",1,0)) %>% 
@@ -74,5 +79,46 @@ average_timeline <- final_timeline %>%
[…]

### 2019-03-15 script/data_wrangling_self_report.R

@@ -1,0 +1,175 @@
+# setup ------------------------------------------------------------------------
+library(here)
+library(tidyverse)
+library(janitor)
+# upload data ------------------------------------------------------------------
+self_report_data <- readr::read_rds(here::here("data/human_self_report_data/Base_Finale_20090114.rds")) %>% 
+  janitor::clean_names() %>% 
+  dplyr::mutate(C_Video = stringr::str_replace(fichier_video_v,".mpg","")) %>% 
+  dplyr::select(
+    n_sujet_c,
+    age,
+    induction_c,
+    genre_c,
+    duree_v,
+    C_Video,          
+    curiosite_interet,
+    ennui,
+    repulsion_degout,
+    frayeur,
+    gaiete_amusement,
+    honte,
+    fierte_satisfaction,
+    stupefaction_surprise,
+    emu,
+    enervee_irritation,   
+    vexe_vexation,
+    decu
+  ) %>%
+  dplyr::mutate(induction_c = case_when(
+    induction_c == "AM" ~ "PT_01",
+    induction_c == "DE" ~ "NT_05",
+    induction_c == "EM" ~ "NT_06",
+    induction_c == "EN" ~ "NT_07",
+    induction_c == "FR" ~ "NT_08",
+    induction_c == "HO" ~ "NT_10",
+    induction_c == "IN" ~ "PT_02",
+    induction_c == "IR" ~ "NT_09",
+    induction_c == "NE" ~ "Neutral",
+    induction_c == "SA" ~ "PT_03",
+    induction_c == "SU" ~ "PT_04"
+    )
+  )
+
+# data wrangling ---------------------------------------------------------------
+
+self_report_tidy <- self_report_data %>%
+  dplyr::select(
+    n_sujet_c,
+    C_Video,
+    induction_c,
+    repulsion_degout,
+    frayeur,
+    gaiete_amusement,
+    stupefaction_surprise,
+    emu,
+    enervee_irritation
+  ) %>% 
+  tidyr::gather(
+    emotion,
+    value,
+    c(
+      "repulsion_degout",
+      "frayeur",
+      "gaiete_amusement",
+      "stupefaction_surprise",
+      "emu",
+      "enervee_irritation"  
+    )
+  ) %>%
+  dplyr::mutate(emotion = case_when(
[…]

### 2019-03-15 script/get_video_duration.R

@@ -28,6 +28,6 @@ metadata_video <- data.frame(list_video, ffprobe_duration) %>%
   dplyr::mutate(ffprobe_duration = round(ffprobe_duration,0)) %>% 
   dplyr::select(-list_video)
 
-write_rds(metadata_video, "metadata_video.rds")
+# write_rds(metadata_video, "metadata_video.rds")
 
 

### 2019-03-15 ACII_2019_paper/ACII_2019_paper.Rmd
typos

@@
 
-With the developement of commercial automatic facial expression recognition tools [see @dupre2018accuracy for a non-exhaustive list of available tools], industries and governments are gradually implementing this technology in order to track humans' emotions in various scenarios (e.g., marketing, heathcare, automotive to name a few).
-Beside the ethical question of measuring human emotions, these tools are challenging psychological theories about infering emotions from facial expressions.
+With the development of commercial automatic facial expression recognition tools [see @dupre2018accuracy for a non-exhaustive list of available tools], industries and governments are gradually implementing this technology in order to track humans' emotions in various scenarios (e.g., marketing, healthcare, automotive to name a few).
+Beside the ethical question of measuring human emotions, these tools are challenging psychological theories about inferring emotions from facial expressions.
 When automatic facial expression recognition tools are used to measure human emotions, they rest on the premise that facial expressions provide a direct access to expressers' emotions.
@@
 
-For the emotion elicitation experiment, `r nrow(self_report_data)` French participants (`r self_report_gender[self_report_gender$genre_c == "F","n"]` females, `r self_report_gender[self_report_gender$genre_c == "H","n"]` males, *M*age = `r self_report_age[,"Age_m"]`, *SD*age = `r self_report_age[,"Age_sd"]`) were recruited to perform one out of 11 emotions eliciation tasks designed to trigger a positive, negative or neutral emotional state [see @tcherkassof2013dynemo for a description of tasks and procedure].
+For the emotion elicitation experiment, `r nrow(self_report_data)` French participants (`r self_report_gender[self_report_gender$genre_c == "F","n"]` females, `r self_report_gender[self_report_gender$genre_c == "H","n"]` males, *M*age = `r self_report_age[,"Age_m"]`, *SD*age = `r self_report_age[,"Age_sd"]`) were recruited to perform one out of 11 emotions elicitation tasks designed to trigger a positive, negative or neutral emotional state [see @tcherkassof2013dynemo for a description of tasks and procedure].
 Participants' face were recorded using an hidden camera resulting `r nrow(self_report_data)` front facing 768x576 videos varying from `r min(metadata_video$ffprobe_duration)`s to `r max(metadata_video$ffprobe_duration)`s.
 
-After their emotion eliciation task the participants had to rate their emotional state during the task on a likert scale from 0 ("not at all") to 5 ("strongly") the six "basic" emotions (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as six "non-basic" emotions (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointement*).
+After their emotion elicitation task the participants had to rate their emotional state during the task on a likert scale from 0 ("not at all") to 5 ("strongly") the six "basic" emotions (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as six "non-basic" emotions (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
 
-Finally, a debriefing session was perform to emsure that participants were not durably affected by the emotion eliciation task.
-All the participants gave their agreement on their data and video to be processed for research puropose only.
+Finally, a debriefing session was perform to ensure that participants were not durably affected by the emotion elicitation task.
+All the participants gave their agreement on their data and video to be processed for research purpose only.
 
@@
 For the human facial expression recognition method, `r nrow(unique_ppt_count)` student participants were recruited to annotate `r nrow(unique_video_count)` out of the `r nrow(self_report_data)` video, therefore only the `r nrow(unique_video_count)` annotated videos will be analysed in this paper.
-Each participants had to annotate between `r min(unique_ppt_count$n)` and `r max(unique_ppt_count$n)` videos resulting that ach video was annotated `r mean(unique_video_count$n) %>% round(0)` times on average (*SD* = `r sd(unique_video_count$n) %>% round(0)`).
+Each participants had to annotate between `r min(unique_ppt_count$n)` and `r max(unique_ppt_count$n)` videos resulting that each video was annotated `r mean(unique_video_count$n) %>% round(0)` times on average (*SD* = `r sd(unique_video_count$n) %>% round(0)`).
 
-The annotation of facial expressions was performed on-site using *Oudjat*, a software for designing video annotation experiements [@dupre2015oudjat].
+The annotation of facial expressions was performed on-site using *Oudjat*, a software for designing video annotation experiments [@dupre2015oudjat].
 For each video, the annotation procedure hat two steps.
-First, the participants had to identify the emotional sequences by pressing the space bar of their keyboard to indicate the begining and the end of the emotional sequences while watching the video.
-Second, the participants watched each emotional sequence previously identified and had to label the sequence using one of the 12 emotions proposed including six "basic" emotions (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) and six "non-basic" emotions (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointement*).
+First, the participants had to identify the emotional sequences by pressing the space bar of their keyboard to indicate the beginning and the end of the emotional sequences while watching the video.
+Second, the participants watched each emotional sequence previously identified and had to label the sequence using one of the 12 emotions proposed including six "basic" emotions (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) and six "non-basic" emotions (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
 They also had the possibility to indicate that the sequence was not expressing one of the proposed emotion.
 
-This annotation procedure results in a uni-dimesional time-serie for each video per human observer identifying for each second of the video which emotion was recognized.
-Then time-series corresponding to the same video were aggreated to calculate the proprotion of human observers for each second of the video per emotional label.
+This annotation procedure results in a uni-dimensional time-series for each video per human observer identifying for each second of the video which emotion was recognized.
+Then time-series corresponding to the same video were aggregated to calculate the proportion of human observers for each second of the video per emotional label.
 The sum of each label proportion per second was used as a score to determine which labels corresponds to the overall video (i.e., the highest score).
-In case of more than one label having the maximum value, the emotion is discribed as undetermined.
+In case of more than one label having the maximum value, the emotion is described as undetermined.
 
@@
 The `r nrow(unique_video_count)` annotated video were processed with Affdex (SDK v3.4.1).
-Affdex is an automatic facial expression recognition classifier developed and distributed by Affectiva is a spinoff company resulting from the research activities of MIT media lab created in 2009 [@mcduff2016affdex].
+Affdex is an automatic facial expression recognition classifier developed and distributed by Affectiva is a spin-off company resulting from the research activities of MIT media lab created in 2009 [@mcduff2016affdex].
 Affdex’s algorithm uses Histogram of Oriented Gradient (HOG) features and Support Vector Machine (SVM) classifiers in order to recognize facial expressions.
@@
 To determine which of the six "basic" emotion can be used to identify each video, the recognition probability for each label by frame was converted into odd ratio by frame [@dente2017measures].
-The highest sum of each odd ratio time-serie defines the label recognized by the automatic classifier.
+The highest sum of each odd ratio time-series defines the label recognized by the automatic classifier.
 
@@
 
-The result of the confusion matrix show a low agreement between emotion felt during the elecitation and emotion recognized by the human annotators (Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "happiness", "prop_video"])`), *surprise* (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "surprise", "prop_video"])`) and *disgust* (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_video"])`).
+The result of the confusion matrix show a low agreement between emotion felt during the elicitation and emotion recognized by the human annotators (Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "happiness", "prop_video"])`), *surprise* (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "surprise", "prop_video"])`) and *disgust* (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_video"])`).
 Sensitivity, specificity, precision and F1 score for each emotion can be found Table \ref{tab:confusionTable_sr_hr}.
@@
 
-However, the self-report show a very high proportion of undetermined emotional states which reveals not only the possiblility of the emotion elicitation tasks to trigger more than one emotion but also the potential limit of using 6-points likert scales for which the participants can easily score to the maximum for more than one emotion.
+However, the self-report show a very high proportion of undetermined emotional states which reveals not only the possibility of the emotion elicitation tasks to trigger more than one emotion but also the potential limit of using 6-points likert scales for which the participants can easily score to the maximum for more than one emotion.
 
@@
 
-Results obtained for the comparison between emotions self-reported and recognised by the automatic classifier are similar to the ones with human observers (Table \ref{tab:confusionTable_sr_ar}).
+Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are similar to the ones with human observers (Table \ref{tab:confusionTable_sr_ar}).
 Overall there is a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_ar$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_ar$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_ar$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_video"])`) and *surprise* (`r scales::percent(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "surprise" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_video"])`).
 Surprisingly the automatic classifier incorrectly recognized as *disgust* an important proportion of videos in which *happiness* was the highest self-reported emotion (`r scales::percent(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_video"])`).
-In parallel, the automatic classifier recognised as *happiness* and *disgust* videos in which *surprise* was the highest self-reported emotion (respectively `r scales::percent(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_video"])` and `r scales::percent(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_video"])`).
+In parallel, the automatic classifier recognized as *happiness* and *disgust* videos in which *surprise* was the highest self-reported emotion (respectively `r scales::percent(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_video"])` and `r scales::percent(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_video"])`).
 
@@
 
-A comparable explaination can be provided as the level of undetermined emotion are very high for the self reports.
+A comparable explanation can be provided as the level of undetermined emotion are very high for the self reports.
 
@@
 ============
-Despit being one on the most investigated question in affective science, the link between emotion felt and facial expression is a hot topic and no clear evidence have been found to definively answer it.
-However, with the growing interest of industries and govenment to monitor individual's psychological states, evidences are showing that facial expressions are in reality not expressing emotions [@mckeown2013analogical].
+Despite being one on the most investigated question in affective science, the link between emotion felt and facial expression is a hot topic and no clear evidence have been found to definitely answer it.
+However, with the growing interest of industries and government to monitor individual's psychological states, evidences are showing that facial expressions are in reality not expressing emotions [@mckeown2013analogical].
 This research aimed to provide some empirical data to the question.
@@
 
-Some limitations to this process should be stated over the use of self-reports to evalute individual's subjective feelings.
-Accessing to the inner subjective feeling can be biaised if not impossible.
+Some limitations to this process should be stated over the use of self-reports to evaluate individual's subjective feelings.
+Accessing to the inner subjective feeling can be biased if not impossible.
 Moreover the laboratory setting can trigger ambiguous and "non-basic" emotions which were not analysed in this research.
 The procedure use for human annotation can also be incriminated.
-Instead of asking the human annotators to provide an unique label, a more sutble approach was choosen to mimic results provided by the automatic classifier.
-In this regard, the results of the human annotation could have been more ambiguous because it is not the natural way that people are infering human emotions.
-Finaly, the automatic classifer algorithm can also be problematic.
-Based on training datasets which are most of the time using prototypical facial expression of the "basic" emotions, the algorithm to classify facial expressions can be held in check by the spontanous facial expressions analysed.
+Instead of asking the human annotators to provide an unique label, a more subtle approach was chosen to mimic results provided by the automatic classifier.
+In this regard, the results of the human annotation could have been more ambiguous because it is not the natural way that people are inferring human emotions.
+Finally, the automatic classifier algorithm can also be problematic.
+Based on training datasets which are most of the time using prototypical facial expression of the "basic" emotions, the algorithm to classify facial expressions can be held in check by the spontaneous facial expressions analysed.
 
-Considering the above, the results provides an additional evidence that individuals' subjective feeling can not be infered from facial expressions and in our case invalidate the hypothesis of hardwired emotions.
-This result suggests that automatic facial expression recognition tools should be focused on evaluating facial morphology features such as action units rather than infering potentional emotional or affective states.
+Considering the above, the results provides an additional evidence that individuals' subjective feeling can not be inferred from facial expressions and in our case invalidate the hypothesis of hardwired emotions.
+This result suggests that automatic facial expression recognition tools should be focused on evaluating facial morphology features such as action units rather than inferring potential emotional or affective states.
 

### 2019-03-15 ACII_2019_paper/ACII_2019_paper.tex
typos

@@
 
-With the developement of commercial automatic facial expression
+With the development of commercial automatic facial expression
 recognition tools (see Dupré et al. 2018 for a non-exhaustive list of
@@
 this technology in order to track humans' emotions in various scenarios
-(e.g., marketing, heathcare, automotive to name a few). Beside the
+(e.g., marketing, healthcare, automotive to name a few). Beside the
 ethical question of measuring human emotions, these tools are
-challenging psychological theories about infering emotions from facial
+challenging psychological theories about inferring emotions from facial
 expressions. When automatic facial expression recognition tools are used
@@
 females, 176 males, \emph{M}age = 47.9, \emph{SD}age = 9.2) were
-recruited to perform one out of 11 emotions eliciation tasks designed to
-trigger a positive, negative or neutral emotional state (see Tcherkassof
-et al. 2013 for a description of tasks and procedure). Participants'
-face were recorded using an hidden camera resulting 358 front facing
-768x576 videos varying from 1s to 1479s.
+recruited to perform one out of 11 emotions elicitation tasks designed
+to trigger a positive, negative or neutral emotional state (see
+Tcherkassof et al. 2013 for a description of tasks and procedure).
+Participants' face were recorded using an hidden camera resulting 358
+front facing 768x576 videos varying from 1s to 1479s.
 
-After their emotion eliciation task the participants had to rate their
+After their emotion elicitation task the participants had to rate their
 emotional state during the task on a likert scale from 0 (``not at
@@
 emotions (i.e., \emph{pride}, \emph{curiosity}, \emph{boredom},
-\emph{shame}, \emph{humiliation}, and \emph{disappointement}).
+\emph{shame}, \emph{humiliation}, and \emph{disappointment}).
 
-Finally, a debriefing session was perform to emsure that participants
-were not durably affected by the emotion eliciation task. All the
+Finally, a debriefing session was perform to ensure that participants
+were not durably affected by the emotion elicitation task. All the
 participants gave their agreement on their data and video to be
-processed for research puropose only.
+processed for research purpose only.
 
@@
 Each participants had to annotate between 1 and 177 videos resulting
-that ach video was annotated 29 times on average (\emph{SD} = 12).
+that each video was annotated 29 times on average (\emph{SD} = 12).
 
 The annotation of facial expressions was performed on-site using
-\emph{Oudjat}, a software for designing video annotation experiements
+\emph{Oudjat}, a software for designing video annotation experiments
 (Dupré et al. 2015). For each video, the annotation procedure hat two
 steps. First, the participants had to identify the emotional sequences
-by pressing the space bar of their keyboard to indicate the begining and
-the end of the emotional sequences while watching the video. Second, the
-participants watched each emotional sequence previously identified and
-had to label the sequence using one of the 12 emotions proposed
+by pressing the space bar of their keyboard to indicate the beginning
+and the end of the emotional sequences while watching the video. Second,
+the participants watched each emotional sequence previously identified
+and had to label the sequence using one of the 12 emotions proposed
 including six ``basic'' emotions (i.e., \emph{anger}, \emph{disgust},
@@
 \emph{boredom}, \emph{shame}, \emph{humiliation}, and
-\emph{disappointement}). They also had the possibility to indicate that
+\emph{disappointment}). They also had the possibility to indicate that
 the sequence was not expressing one of the proposed emotion.
 
-This annotation procedure results in a uni-dimesional time-serie for
+This annotation procedure results in a uni-dimensional time-series for
 each video per human observer identifying for each second of the video
 which emotion was recognized. Then time-series corresponding to the same
-video were aggreated to calculate the proprotion of human observers for
+video were aggregated to calculate the proportion of human observers for
 each second of the video per emotional label. The sum of each label
@@
 corresponds to the overall video (i.e., the highest score). In case of
-more than one label having the maximum value, the emotion is discribed
+more than one label having the maximum value, the emotion is described
 as undetermined.
@@
 is an automatic facial expression recognition classifier developed and
-distributed by Affectiva is a spinoff company resulting from the
+distributed by Affectiva is a spin-off company resulting from the
 research activities of MIT media lab created in 2009 (McDuff et al.
@@
 converted into odd ratio by frame (Dente et al. 2017). The highest sum
-of each odd ratio time-serie defines the label recognized by the
+of each odd ratio time-series defines the label recognized by the
 automatic classifier.
@@
 The result of the confusion matrix show a low agreement between emotion
-felt during the elecitation and emotion recognized by the human
+felt during the elicitation and emotion recognized by the human
 annotators (Accuracy = 0.27, 95\%CI{[}0.21,0.33{]}; Kappa = 0.11) except
@@
 However, the self-report show a very high proportion of undetermined
-emotional states which reveals not only the possiblility of the emotion
+emotional states which reveals not only the possibility of the emotion
 elicitation tasks to trigger more than one emotion but also the
@@
 Results obtained for the comparison between emotions self-reported and
-recognised by the automatic classifier are similar to the ones with
+recognized by the automatic classifier are similar to the ones with
 human observers (Table \ref{tab:confusionTable_sr_ar}). Overall there is
@@
 \emph{happiness} was the highest self-reported emotion (7.83\%). In
-parallel, the automatic classifier recognised as \emph{happiness} and
+parallel, the automatic classifier recognized as \emph{happiness} and
 \emph{disgust} videos in which \emph{surprise} was the highest
@@
 
-A comparable explaination can be provided as the level of undetermined
+A comparable explanation can be provided as the level of undetermined
 emotion are very high for the self reports.
@@
 
-Despit being one on the most investigated question in affective science,
-the link between emotion felt and facial expression is a hot topic and
-no clear evidence have been found to definively answer it. However, with
-the growing interest of industries and govenment to monitor individual's
-psychological states, evidences are showing that facial expressions are
-in reality not expressing emotions (McKeown 2013). This research aimed
-to provide some empirical data to the question. The subjective feeling
-of participants was compared with human recognition on one side and
-automatic recognition on the other side. The results reveals a low
-accuracy for both humans and automatic classifier to accurately identify
-the inner emotional states of these individuals based on their facial
-expressions.
+Despite being one on the most investigated question in affective
+science, the link between emotion felt and facial expression is a hot
+topic and no clear evidence have been found to definitely answer it.
+However, with the growing interest of industries and government to
+monitor individual's psychological states, evidences are showing that
+facial expressions are in reality not expressing emotions (McKeown
+2013). This research aimed to provide some empirical data to the
+question. The subjective feeling of participants was compared with human
+recognition on one side and automatic recognition on the other side. The
+results reveals a low accuracy for both humans and automatic classifier
+to accurately identify the inner emotional states of these individuals
+based on their facial expressions.
 
 Some limitations to this process should be stated over the use of
-self-reports to evalute individual's subjective feelings. Accessing to
-the inner subjective feeling can be biaised if not impossible. Moreover
+self-reports to evaluate individual's subjective feelings. Accessing to
+the inner subjective feeling can be biased if not impossible. Moreover
 the laboratory setting can trigger ambiguous and ``non-basic'' emotions
@@
 annotation can also be incriminated. Instead of asking the human
-annotators to provide an unique label, a more sutble approach was
-choosen to mimic results provided by the automatic classifier. In this
-regard, the results of the human annotation could have been more
-ambiguous because it is not the natural way that people are infering
-human emotions. Finaly, the automatic classifer algorithm can also be
+annotators to provide an unique label, a more subtle approach was chosen
+to mimic results provided by the automatic classifier. In this regard,
+the results of the human annotation could have been more ambiguous
+because it is not the natural way that people are inferring human
+emotions. Finally, the automatic classifier algorithm can also be
 problematic. Based on training datasets which are most of the time using
 prototypical facial expression of the ``basic'' emotions, the algorithm
-to classify facial expressions can be held in check by the spontanous
+to classify facial expressions can be held in check by the spontaneous
 facial expressions analysed.
@@
 Considering the above, the results provides an additional evidence that
-individuals' subjective feeling can not be infered from facial
+individuals' subjective feeling can not be inferred from facial
 expressions and in our case invalidate the hypothesis of hardwired
@@
 recognition tools should be focused on evaluating facial morphology
-features such as action units rather than infering potentional emotional
+features such as action units rather than inferring potential emotional
 or affective states.

### 2019-03-16 ACII_2019_paper/ACII_2019_paper.Rmd
summary

@@
 ---
-title: The relation betweem feeling and expressing emotions. A comparison between sujective feeling and facial expression recognition of spontaneous emotions. 
+title: Are facial expressions the genuine display of individuals' subjective feeling? A comparison with human and automatic recognition. 
 affiliation:
@@
 abstract: |
-  The abstract goes here.
-  On multiple lines eventually.
+  While it has been taken for granted in the development of several automatic facial expression recognition tools, the question of the link between subjective feelings and facial expressions is still a subject of debate.
+On one hand the behaviorist approach conceives emotions as genetically hardwired and therefore being genuinely displayed through facial expressions.
+On the other hand the constructivist approach conceives emotions are socially constructed and facial expression as social messages that are not related to emotions.
+In order to evaluate the link between the subjective feeling of emotions and their recognition based on facial expression, 232 videos of participants recruited to perform an emotion elicitation task were annotated by 1383 human observers as well as by an automatic facial expression classifier.
+The results show a low accuracy of human observers and of the automatic classifier to infer from the facial expression the subjective feeling of the participants recorded.
+Based on these results, the hypothesis of genetically hardwired emotion genuinely display is difficult to support whereas the idea of emotion socially constructed and facial expression as display of social messages appears to be more likely.
+Then, the way to infer emotional and mental state based on facial expressions should be questioned.
 header-includes:
@@
 Based on the behaviorist approach initiated by Darwin in *The Expression of the Emotions in Man and Animals* [@darwin1872expression], facial expression are conceived as a genuine display of individuals inner emotional state.
-This hypothesis is used as a basis for the Basic Emotion Theory (BET) which states that a set of six or seven emotions are universally displayed and are genetically hardwired [@ekman1992argument].
+This hypothesis is used as a basis for the Basic Emotion Theory (BET) which states that a set of six or seven emotions are universally displayed and are genetically hardwired not only in humans [@ekman1992argument] but also in different animal species [@de2019mama].
 According to this view, "when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including [...] facial expressions." [@ekman2007directed, p49].
@@
 ## The link between emotion and facial expression thought the social constructivist approach
-Detractors of the Basic Emotion Theory are perceiving emotion not as genetically hardwired but as a learnt association between a given situation and an appropriate response [@averill1980constructivist].
+Detractors of the Basic Emotion Theory are perceiving emotion not as genetically hardwired but as a learnt association between a given situation and an appropriate response [@averill1980constructivist, @barrett2017emotions].
 For the tenants of the constructivist approach, emotions are "concepts" based on past experiences and which are "a collection of embodied, whole brain representations that predict what is about to happen in the sensory environment, what the best action is to deal with impending events, and their consequences for allostasis" [@barrett2017theory, p12].

### 2019-03-16 ACII_2019_paper/ACII_2019_paper.tex
summary

@@
 \hypersetup{
-            pdftitle={The relation betweem feeling and expressing emotions. A comparison between sujective feeling and facial expression recognition of spontaneous emotions.},
+            pdftitle={Are facial expressions the genuine display of individuals' subjective feeling? A comparison with human and automatic recognition.},
             pdfborder={0 0 0},
@@
 % Do not put math or special symbols in the title.
-\title{The relation betweem feeling and expressing emotions. A comparison
-between sujective feeling and facial expression recognition of
-spontaneous emotions.}
+\title{Are facial expressions the genuine display of individuals' subjective
+feeling? A comparison with human and automatic recognition.}
 
@@
 \begin{abstract}
-The abstract goes here. On multiple lines eventually.
+While it has been taken for granted in the development of several
+automatic facial expression recognition tools, the question of the link
+between subjective feelings and facial expressions is still a subject of
+debate. On one hand the behaviorist approach conceives emotions as
+genetically hardwired and therefore being genuinely displayed through
+facial expressions. On the other hand the constructivist approach
+conceives emotions are socially constructed and facial expression as
+social messages that are not related to emotions. In order to evaluate
+the link between the subjective feeling of emotions and their
+recognition based on facial expression, 232 videos of participants
+recruited to perform an emotion elicitation task were annotated by 1383
+human observers as well as by an automatic facial expression classifier.
+The results show a low accuracy of human observers and of the automatic
+classifier to infer from the facial expression the subjective feeling of
+the participants recorded. Based on these results, the hypothesis of
+genetically hardwired emotion genuinely display is difficult to support
+whereas the idea of emotion socially constructed and facial expression
+as display of social messages appears to be more likely. Then, the way
+to infer emotional and mental state based on facial expressions should
+be questioned.
 \end{abstract}
@@
 Emotion Theory (BET) which states that a set of six or seven emotions
-are universally displayed and are genetically hardwired (Ekman 1992).
+are universally displayed and are genetically hardwired not only in
+humans (Ekman 1992) but also in different animal species (Waal 2019).
 According to this view, ``when emotions are aroused by perception of a
@@
 genetically hardwired but as a learnt association between a given
-situation and an appropriate response (Averill 1980). For the tenants of
-the constructivist approach, emotions are ``concepts'' based on past
-experiences and which are ``a collection of embodied, whole brain
-representations that predict what is about to happen in the sensory
-environment, what the best action is to deal with impending events, and
-their consequences for allostasis'' (Barrett 2017, p12). Following this
-assumption, faces are used as tools to display signals in social
-interactions (Crivelli and Fridlund 2018). These signals can convey
-individuals' motivations and readiness (Frijda and Tcherkassof 1997) or
-social messages (Fridlund and Rosenberg 1995).
+situation and an appropriate response (Averill 1980,
+@barrett2017emotions). For the tenants of the constructivist approach,
+emotions are ``concepts'' based on past experiences and which are ``a
+collection of embodied, whole brain representations that predict what is
+about to happen in the sensory environment, what the best action is to
+deal with impending events, and their consequences for allostasis''
+(Barrett 2017b, p12). Following this assumption, faces are used as tools
+to display signals in social interactions (Crivelli and Fridlund 2018).
+These signals can convey individuals' motivations and readiness (Frijda
+and Tcherkassof 1997) or social messages (Fridlund and Rosenberg 1995).
 
@@
 
+\leavevmode\hypertarget{ref-barrett2017emotions}{}%
+Barrett, Lisa Feldman. 2017a. \emph{How Emotions Are Made: The Secret
+Life of the Brain}. Houghton Mifflin Harcourt.
+
 \leavevmode\hypertarget{ref-barrett2017theory}{}%
-Barrett, Lisa Feldman. 2017. ``The Theory of Constructed Emotion: An
-Active Inference Account of Interoception and Categorization.''
-\emph{Social Cognitive and Affective Neuroscience} 12 (1): 1--23.
+---------. 2017b. ``The Theory of Constructed Emotion: An Active
+Inference Account of Interoception and Categorization.'' \emph{Social
+Cognitive and Affective Neuroscience} 12 (1): 1--23.
 
@@
 
+\leavevmode\hypertarget{ref-de2019mama}{}%
+Waal, Frans de. 2019. \emph{Mama's Last Hug: Animal Emotions and What
+They Teach Us About Ourselves}. Granta Books.
+
 \end{document}

### 2019-03-16 ACII_2019_paper/ACII_2019_paper.tex
word

@@
 
+\hypertarget{comparison-between-human-and-automatic-recognition}{%
+\subsection{Comparison between human and automatic
+recognition}\label{comparison-between-human-and-automatic-recognition}}
+
+As previously mentioned, the accuracy of humans observers and the
+automatic classifier have some similarities. In order to compare which
+recognition has the highest accuracy a Receiver Operating Characteristic
+(ROC) curve was calculated (Figure \ref{fig:roc}).
+
+\begin{figure}
+\centering
+\includegraphics{ACII_2019_paper_files/figure-latex/roc-1.pdf}
+\caption{ROC curve comparing the accuracy in inferring subjective
+feelings from facial expressions by human observers and automatic
+recognition.}
+\end{figure}
+
+The ROC curve and its Area Under the Curve (AUC) values shows that the
+automatic classifier are more accurate than human observers in inferring
+subjective feeling from facial expressions (human AUC = 0.57 \emph{vs.}
+automatic AUC = 0.66).
+
 \hypertarget{conclusion}{%

### 2019-03-16 ACII_2019_paper/ACII_2019_word.Rmd
word

--- 2019-03-16 ACII_2019_paper/ACII_2019_paper.Rmd
@@
 ---
-title: Are facial expressions the genuine display of individuals' subjective feeling? A comparison with human and automatic recognition. 
-affiliation:
-  ## use one only of the following
-  # author-columnar: true         ## one column per author
-  institution-columnar: true  ## one column per institution (multiple autors eventually)
-  # wide: true                  ## one column wide author/affiliation fields
-
-  institution:
-    - name: Dublin City University
-      location: Dublin, Ireland
-      email: damien.dupre@dcu.ie
-      mark: 1
-      author:
-        - name: Damien Dupré
-    - name: Grenoble Alpes University
-      location: Grenoble, France
-      mark: 2
-      author:
-        - name: Anna Tcherkassof
-          email: anna.tcherkassof@univ-grenoble-alpes.fr
+title: "Are facial expressions the genuine display of individuals' subjective feeling? A comparison of human and automatic recognition. "
+output: word_document
+bibliography: mybibfile.bib
 abstract: |
@@
 Then, the way to infer emotional and mental state based on facial expressions should be questioned.
-header-includes:
-  - \usepackage{booktabs}
-  - \usepackage{float}
-  - \usepackage{tabu}
-bibliography: mybibfile.bib
-output: rticles::ieee_article
 ---
@@
 library(here) # Locate files relative to project root
+library(caret)
+library(plotROC)
+library(pROC)
+library(pander)
+library(papaja)
+library(captioner)
+table_nums <- captioner::captioner(prefix = "Table")
+table_nums(
+  name = "confusionTable_sr_hr", 
+  caption = "Agreement accuracy metrics for each emotion.")
+table_nums(
+  name = "confusionTable_sr_ar", 
+  caption = "Agreement accuracy metrics for each emotion.")
+#
+fig_nums <- captioner::captioner()
+
+fig_nums(
+  name = "confusionMatrix_sr_hr", 
+  caption = "Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the human observers.")
+fig_nums(
+  name = "confusionMatrix_sr_ar", 
+  caption = "Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the automatic classifier.")
+fig_nums(
+  name = "roc", 
+  caption = "ROC curve comparing the accuracy in inferring subjective feelings from facial expressions by human observers and automatic recognition.")
 ```
@@
 ```
-
 With the development of commercial automatic facial expression recognition tools [see @dupre2018accuracy for a non-exhaustive list of available tools], industries and governments are gradually implementing this technology in order to track humans' emotions in various scenarios (e.g., marketing, healthcare, automotive to name a few).
@@
 
-Emotions self-reported as being characteristic of the elicitation are compared with the emotion recognized by the human observers in a confusion matrix (Figure \ref{fig:confusionMatrix_sr_hr}).
+Emotions self-reported as being characteristic of the elicitation are compared with the emotion recognized by the human observers in a confusion matrix (`r fig_nums("confusionMatrix_sr_hr", display = "cite")`).
 
-```{r confusionMatrix_sr_hr, fig.height=4, fig.cap="Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the human observers."}
+```{r confusionMatrix_sr_hr, fig.height=4, fig.cap=fig_nums("confusionMatrix_sr_hr")}
 confusionMatrix_sr_hr_freq <- confusionMatrix_sr_hr$table %>% 
@@
 The result of the confusion matrix show a low agreement between emotion felt during the elicitation and emotion recognized by the human annotators (Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "happiness", "prop_video"])`), *surprise* (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "surprise", "prop_video"])`) and *disgust* (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_video"])`).
-Sensitivity, specificity, precision and F1 score for each emotion can be found Table \ref{tab:confusionTable_sr_hr}.
+Sensitivity, specificity, precision and F1 score for each emotion can be found `r table_nums("confusionTable_sr_hr", display = "cite")`.
 Interestingly human annotators seem to recognize as *surprise* videos in which *happiness* was the highest self-reported emotion (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "happiness", "prop_video"])`), and in a lower instance *happiness* videos in which *surprise* was the highest self-reported emotion (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "surprise", "prop_video"])`).
@@
 
+# confusionMatrix_sr_hr_table %>%
+#   knitr::kable(
+#     "latex",
+#     caption = "Agreement accuracy metrics for each emotion. ",
+#     booktabs = TRUE,
+#     digits = 2,
+#     linesep = "",
+#     escape = FALSE,
+#     align = c('l',rep('c', 4))
+#   ) %>%
+#   kableExtra::kable_styling(
+#     font_size = 8,
+#     full_width = TRUE,
+#     latex_options = c("hold_position")) %>%
+#    add_footnote("Note. \\textit{na.} values are produced when not enough data are available to compute accuracy indicators.", notation="none", escape = FALSE)
+
 confusionMatrix_sr_hr_table %>%
-  knitr::kable(
-    "latex",
-    caption = "Agreement accuracy metrics for each emotion. ",
-    booktabs = TRUE,
-    digits = 2,
-    linesep = "",
-    escape = FALSE,
-    align = c('l',rep('c', 4))
-  ) %>%
-  kableExtra::kable_styling(
-    font_size = 8,
-    full_width = TRUE,
-    latex_options = c("hold_position")) %>%
-   add_footnote("Note. \\textit{na.} values are produced when not enough data are available to compute accuracy indicators.", notation="none", escape = FALSE)
+  papaja::apa_table(
+    caption = table_nums("confusionTable_sr_hr"),
+    row.names = FALSE,
+    format = "word",
+    escape = FALSE
+    )
 ```
@@
 
-As in the previous analysis, emotions self-reported as being characteristic of the elicitation are compared with the emotion recognized by the automatic classifier in a confusion matrix (Figure \ref{fig:confusionMatrix_sr_ar}).
+As in the previous analysis, emotions self-reported as being characteristic of the elicitation are compared with the emotion recognized by the automatic classifier in a confusion matrix (`r fig_nums("confusionMatrix_sr_ar", display = "cite")`).
 
-```{r confusionMatrix_sr_ar, fig.height=4, fig.cap="Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the automatic classifier."}
+```{r confusionMatrix_sr_ar, fig.height=4, fig.cap=fig_nums("confusionMatrix_sr_ar")}
 confusionMatrix_sr_ar_freq <- confusionMatrix_sr_ar$table %>% 
@@
 
-Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are similar to the ones with human observers (Table \ref{tab:confusionTable_sr_ar}).
+Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are similar to the ones with human observers (`r table_nums("confusionTable_sr_ar", display = "cite")`).
 Overall there is a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_ar$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_ar$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_ar$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_video"])`) and *surprise* (`r scales::percent(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "surprise" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_video"])`).
@@
 
+# confusionMatrix_sr_ar_table %>%
+#   knitr::kable(
+#     "latex",
+#     caption = "Agreement accuracy metrics for each emotion.",
+#     booktabs = T,
+#     digits = 2,
+#     linesep = "",
+#     escape = FALSE,
+#     align = c('l',rep('c', 4))
+#   ) %>%
+#   kableExtra::kable_styling(
+#     font_size = 8,
+#     full_width = TRUE,
+#     latex_options = c("hold_position")) %>%
+#    add_footnote("Note. \\textit{na.} values are produced when not enough data are available to compute accuracy indicators.", notation="none", escape = FALSE)
+
 confusionMatrix_sr_ar_table %>%
-  knitr::kable(
-    "latex",
-    caption = "Agreement accuracy metrics for each emotion.",
-    booktabs = T,
-    digits = 2,
-    linesep = "",
-    escape = FALSE,
-    align = c('l',rep('c', 4))
-  ) %>%
-  kableExtra::kable_styling(
-    font_size = 8,
-    full_width = TRUE,
-    latex_options = c("hold_position")) %>%
-   add_footnote("Note. \\textit{na.} values are produced when not enough data are available to compute accuracy indicators.", notation="none", escape = FALSE)
+  papaja::apa_table(
+    caption = table_nums("confusionTable_sr_ar"),
+    row.names = FALSE,
+    format = "word",
+    escape = FALSE
+    )
 ```
@@
 A comparable explanation can be provided as the level of undetermined emotion are very high for the self reports.
+
+## Comparison between human and automatic recognition
+
+As previously mentioned, the accuracy of humans observers and the automatic classifier have some similarities.
+In order to compare which recognition has the highest accuracy a Receiver Operating Characteristic (ROC) curve was calculated (`r fig_nums("roc", display = "cite")`).
+
+```{r roc, fig.height=6, fig.cap=fig_nums("roc")}
+roc_sr_hr <- comparison_sr_hr %>% 
+  dplyr::mutate_if(is.factor, as.character) %>% 
+  dplyr::mutate(correct_classification = case_when(
+    sr_emotion == hr_emotion ~ 1,
+    sr_emotion != hr_emotion ~ 0
+  )) %>% 
+  dplyr::rename(rec_score = hr_score) %>% 
+  dplyr::mutate(rec_type = "human")
+
+roc_sr_ar <- comparison_sr_ar %>% 
+  dplyr::mutate_if(is.factor, as.character) %>% 
+  dplyr::mutate(correct_classification = case_when(
+    sr_emotion == ar_emotion ~ 1,
+    sr_emotion != ar_emotion ~ 0
+  )) %>% 
+  dplyr::rename(rec_score = ar_score) %>% 
+  dplyr::mutate(rec_type = "automatic")
+
+auc <- dplyr::bind_rows(roc_sr_hr,roc_sr_ar) %>% 
+  tidyr::nest(-rec_type) %>%
+  dplyr::mutate(rec_type = as.factor(rec_type),
+                roc_obj = purrr::map(data, ~ roc(.$correct_classification,.$rec_score)),
+                roc_value = purrr::map(roc_obj,auc)) %>%
+  tidyr::unnest(roc_value) %>% 
+  dplyr::select(rec_type,roc_value) %>% 
+  dplyr::mutate(roc_value = round(roc_value,2))
+
+dplyr::bind_rows(roc_sr_hr,roc_sr_ar) %>% 
+  ggplot(aes(d = correct_classification, m = rec_score, color = rec_type)) + 
+  geom_roc(labels = FALSE,pointsize = 0) +
+  style_roc(xlab = "False Positive Rate",
+            ylab = "True Positive Rate", 
+            theme = theme_minimal) + 
+  scale_color_manual(
+    "Classifiers",
+    values = c("#e41a1c","#377eb8")
+  ) +
+  theme(text = element_text(size=18,family="serif"),
+        strip.text.x = element_text(face="bold",size=18),
+        axis.title.x = element_text(margin = margin(t = 20, r = 0, b = 0, l = 0)),
+        axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
+        axis.text.x = element_text(size = 10),
+        axis.text.y = element_text(size = 10),
+        legend.text = element_text(size=10),
+        legend.position = "bottom") + 
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed")
+```
+
+The ROC curve and its Area Under the Curve (AUC) values shows that the automatic classifier are more accurate than human observers in inferring subjective feeling from facial expressions (human AUC = `r auc[auc$rec_type == "human","roc_value"]` *vs.* automatic AUC = `r auc[auc$rec_type == "automatic","roc_value"]`).
 

### 2019-03-17 ACII_2019_paper/ACII_2019_word.Rmd
picture dynemo

@@
 fig_nums <- captioner::captioner()
-
+fig_nums(
+  name = "dynemo_img", 
+  caption = "Example of a front facing recording synced with the full view of the participant and the elicitation task.
+All participants gave their consent for their data to be processed in non-commercial research.
+This picture is taken from a pilot with projects collaborators and all gave a consent for the publication of their photos and videos.")
 fig_nums(
@@
 With the development of commercial automatic facial expression recognition tools [see @dupre2018accuracy for a non-exhaustive list of available tools], industries and governments are gradually implementing this technology in order to track humans' emotions in various scenarios (e.g., marketing, healthcare, automotive to name a few).
-Beside the ethical question of measuring human emotions, these tools are challenging psychological theories about inferring emotions from facial expressions.
-When automatic facial expression recognition tools are used to measure human emotions, they rest on the premise that facial expressions provide a direct access to expressers' emotions.
-However the link between emotion felt and facial expressions is far from being established and still remain a hot topic in psychology research.
+This technoloogy rests on the premise that facial expressions provide a direct access to individuals' subjective feeling.
+Even if this premise is central to the modern mainstream approach of human emotion, recent research in affective science are challenging it.
+After presenting the arguments supporting both sides, an experiment testing these hypotheses will be presented and its results analyzed in order to provide empirical evidence to contribute to answer the question.
 
-## The link between emotion and facial expression thought the behaviorist approach
+## The link between emotion and facial expression through the behaviorist approach
 Based on the behaviorist approach initiated by Darwin in *The Expression of the Emotions in Man and Animals* [@darwin1872expression], facial expression are conceived as a genuine display of individuals inner emotional state.
-This hypothesis is used as a basis for the Basic Emotion Theory (BET) which states that a set of six or seven emotions are universally displayed and are genetically hardwired not only in humans [@ekman1992argument] but also in different animal species [@de2019mama].
+This hypothesis is used as a basis for the Basic Emotion Theory (BET) which states that a set of six emotions are universally displayed and are genetically hardwired not only in humans [@ekman1992argument] but also in different animal species [@de2019mama].
 According to this view, "when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including [...] facial expressions." [@ekman2007directed, p49].
-Even if this theory obtained a popular support, it fails to explain how individuals can feel emotions without expressing them and how individuals can express emotions without feeling them.
+To cope with critics, several amendments have been made to the BET, increasing the number of basic emotions from six to seven  [@ekman1988universality] as well as adding the concept of "display rules" to explain cultural differences in the management of facial expressions [@ekman1987universals].
+Even if this theory obtained a popular support, it fails to explain how individuals can feel emotions without expressing them and how individuals can express emotions without feeling them, cases for which display rules cannot be applied to [@kraut1979social, @duran2017coherence].
 
-## The link between emotion and facial expression thought the social constructivist approach
+## The link between emotion and facial expression through the social constructivist approach
 Detractors of the Basic Emotion Theory are perceiving emotion not as genetically hardwired but as a learnt association between a given situation and an appropriate response [@averill1980constructivist, @barrett2017emotions].
@@
 These signals can convey individuals' motivations and readiness [@frijda1997facial] or social messages [@fridlund1995human].
+Therefore, facial expressions are thought as behaviors which meaning is inferred by the observer.
 
@@
 
-For the emotion elicitation experiment, `r nrow(self_report_data)` French participants (`r self_report_gender[self_report_gender$genre_c == "F","n"]` females, `r self_report_gender[self_report_gender$genre_c == "H","n"]` males, *M*age = `r self_report_age[,"Age_m"]`, *SD*age = `r self_report_age[,"Age_sd"]`) were recruited to perform one out of 11 emotions elicitation tasks designed to trigger a positive, negative or neutral emotional state [see @tcherkassof2013dynemo for a description of tasks and procedure].
-Participants' face were recorded using an hidden camera resulting `r nrow(self_report_data)` front facing 768x576 videos varying from `r min(metadata_video$ffprobe_duration)`s to `r max(metadata_video$ffprobe_duration)`s.
+For the emotion elicitation experiment, `r nrow(self_report_data)` French participants (`r self_report_gender[self_report_gender$genre_c == "F","n"]` females, `r self_report_gender[self_report_gender$genre_c == "H","n"]` males, *M*age = `r self_report_age[,"Age_m"]`, *SD*age = `r self_report_age[,"Age_sd"]`) were recruited to perform one out of 11 emotion elicitation tasks designed to trigger a positive, negative or neutral emotional state.
+Participants' face were recorded using an hidden camera resulting `r nrow(self_report_data)` front facing 768x576 videos varying from `r min(metadata_video$ffprobe_duration)`s to `r max(metadata_video$ffprobe_duration)`s (`r fig_nums("dynemo_img", display = "cite")`).
+These recordings are constituting the DynEmo database [see @tcherkassof2013dynemo for a full description of tasks and procedure].
+
+```{r dynemo_img, fig.cap=fig_nums("dynemo_img")}
+knitr::include_graphics("./img/dynemo_img.jpg")
+```
 
@@
 Finally, a debriefing session was perform to ensure that participants were not durably affected by the emotion elicitation task.
+The debriefing was also used to check that participants did not guess the real purpose of the experiment (e.g.
+being filmed while they were performing an emotional elicitation task) to guarantee facial expressions genuineness.
 All the participants gave their agreement on their data and video to be processed for research purpose only.
+
+and did not guess the rela purpose of the task (e.g. being filmed while they were performing an emotional indiuction task). garantuee of genuine facial emotional expressions.
 

### 2019-03-18 ACII_2019_paper/ACII_2019_paper.Rmd
conclusion

@@
 ---
-title: Are facial expressions the genuine display of individuals' subjective feeling? A comparison with human and automatic recognition. 
+title: Are facial expressions the genuine display of individuals' subjective feeling? A comparison of human and automatic recognition. 
 affiliation:
@@
 library(here) # Locate files relative to project root
+library(caret)
+library(plotROC)
+library(pROC)
 ```
@@
 
-After their emotion elicitation task the participants had to rate their emotional state during the task on a likert scale from 0 ("not at all") to 5 ("strongly") the six "basic" emotions (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as six "non-basic" emotions (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
+After their emotion elicitation task the participants had to rate their subjective feeling during the task on likert scales from 0 ("not at all") to 5 ("strongly") according six "basic" emotion labels (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as six "non-basic" emotion labels (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
 
@@
 First, the participants had to identify the emotional sequences by pressing the space bar of their keyboard to indicate the beginning and the end of the emotional sequences while watching the video.
-Second, the participants watched each emotional sequence previously identified and had to label the sequence using one of the 12 emotions proposed including six "basic" emotions (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) and six "non-basic" emotions (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
+Second, the participants watched each emotional sequence previously identified and had to label the sequence using one of the 12 emotion labels proposed including six "basic" emotion labels (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) and six "non-basic" emotion labels (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
 They also had the possibility to indicate that the sequence was not expressing one of the proposed emotion.
@@
 Affdex’s algorithm uses Histogram of Oriented Gradient (HOG) features and Support Vector Machine (SVM) classifiers in order to recognize facial expressions.
-For each video frame, Affdex identify the probability of the face as expressing one of the six "basic" emotions (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as additional psychological states such as *valence*, *engagement* or *contempt*, and facial features such as *cheek raise*, *eye widen* or *jaw drop*.
+For each video frame, Affdex identify the probability of the face to express each of the six "basic" emotion labels (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as additional psychological states such as *valence*, *engagement* or *contempt*, and facial features such as *cheek raise*, *eye widen* or *jaw drop*.
 
-To determine which of the six "basic" emotion can be used to identify each video, the recognition probability for each label by frame was converted into odd ratio by frame [@dente2017measures].
+To determine which of the six emotion label can be used to identify each video, the recognition probability for each label by frame was converted into odd ratio by frame [@dente2017measures].
 The highest sum of each odd ratio time-series defines the label recognized by the automatic classifier.
@@
 
-Whereas the self-reports, the human annotations and the automatic recognition include data on "non-basic" emotions and features, the analysis is performed using only the six "basic" emotions in order to compare them.
+Whereas the self-reports, the human annotations and the automatic recognition include data on "non-basic" emotion labels and features, the analysis is performed using only the six "basic" emotion labels in order to compare them.
 The maximum score for self-reports, human annotations and automatic recognition is used to label the video.
@@
     "latex",
-    caption = "Agreement accuracy metrics for each emotion. ",
+    caption = "Agreement accuracy metrics for each emotion.",
     booktabs = TRUE,
@@
 
+## Comparison between human and automatic recognition
+
+As previously mentioned, the accuracy of humans observers and the automatic classifier have some similarities.
+In order to compare which recognition has the highest accuracy a Receiver Operating Characteristic (ROC) curve was calculated (Figure \ref{fig:roc}).
+
+```{r roc, fig.height=6, fig.cap="ROC curve comparing the accuracy in inferring subjective feelings from facial expressions by human observers and automatic recognition."}
+roc_sr_hr <- comparison_sr_hr %>% 
+  dplyr::mutate_if(is.factor, as.character) %>% 
+  dplyr::mutate(correct_classification = case_when(
+    sr_emotion == hr_emotion ~ 1,
+    sr_emotion != hr_emotion ~ 0
+  )) %>% 
+  dplyr::rename(rec_score = hr_score) %>% 
+  dplyr::mutate(rec_type = "human")
+
+roc_sr_ar <- comparison_sr_ar %>% 
+  dplyr::mutate_if(is.factor, as.character) %>% 
+  dplyr::mutate(correct_classification = case_when(
+    sr_emotion == ar_emotion ~ 1,
+    sr_emotion != ar_emotion ~ 0
+  )) %>% 
+  dplyr::rename(rec_score = ar_score) %>% 
+  dplyr::mutate(rec_type = "automatic")
+
+auc <- dplyr::bind_rows(roc_sr_hr,roc_sr_ar) %>% 
+  tidyr::nest(-rec_type) %>%
+  dplyr::mutate(rec_type = as.factor(rec_type),
+                roc_obj = purrr::map(data, ~ roc(.$correct_classification,.$rec_score)),
+                roc_value = purrr::map(roc_obj,auc)) %>%
+  tidyr::unnest(roc_value) %>% 
+  dplyr::select(rec_type,roc_value) %>% 
+  dplyr::mutate(roc_value = round(roc_value,2))
+
+dplyr::bind_rows(roc_sr_hr,roc_sr_ar) %>% 
+  ggplot(aes(d = correct_classification, m = rec_score, color = rec_type)) + 
+  geom_roc(labels = FALSE,pointsize = 0) +
+  style_roc(xlab = "False Positive Rate",
+            ylab = "True Positive Rate", 
+            theme = theme_minimal) + 
+  scale_color_manual(
+    "Classifiers",
+    values = c("#e41a1c","#377eb8")
+  ) +
+  theme(text = element_text(size=18,family="serif"),
+        strip.text.x = element_text(face="bold",size=18),
+        axis.title.x = element_text(margin = margin(t = 20, r = 0, b = 0, l = 0)),
+        axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
+        axis.text.x = element_text(size = 10),
+        axis.text.y = element_text(size = 10),
+        legend.text = element_text(size=10),
+        legend.position = "bottom") + 
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed")
+```
+
+The ROC curve and its Area Under the Curve (AUC) values shows that the automatic classifier are more accurate than human observers in inferring subjective feeling from facial expressions (human AUC = `r auc[auc$rec_type == "human","roc_value"]` *vs.* automatic AUC = `r auc[auc$rec_type == "automatic","roc_value"]`).
+
 Conclusion
@@
 Some limitations to this process should be stated over the use of self-reports to evaluate individual's subjective feelings.
-Accessing to the inner subjective feeling can be biased if not impossible.
-Moreover the laboratory setting can trigger ambiguous and "non-basic" emotions which were not analysed in this research.
-The procedure use for human annotation can also be incriminated.
+Accessing the inner subjective feeling can be biased if not impossible.
+Moreover the laboratory setting can trigger ambiguous and "non-basic" emotion labels which were not analysed in this research.
+The procedure used for human annotation can also be open to dispute.
 Instead of asking the human annotators to provide an unique label, a more subtle approach was chosen to mimic results provided by the automatic classifier.
@@
 Finally, the automatic classifier algorithm can also be problematic.
-Based on training datasets which are most of the time using prototypical facial expression of the "basic" emotions, the algorithm to classify facial expressions can be held in check by the spontaneous facial expressions analysed.
+Based on training datasets which are most of the time using prototypical facial expression of the "basic" emotion labels, the algorithm to classify facial expressions can be held in check by the spontaneity of the expressions analysed which are less likely to be prototypical.
 
 Considering the above, the results provides an additional evidence that individuals' subjective feeling can not be inferred from facial expressions and in our case invalidate the hypothesis of hardwired emotions.
-This result suggests that automatic facial expression recognition tools should be focused on evaluating facial morphology features such as action units rather than inferring potential emotional or affective states.
+Advancements in identifying "non-basic" emotion labels [@mcduff2016discovering] as well as non-prototypical facial expression have been made in the developpement of automatic facial expression recognition tools, however this result suggests that automatic facial expression recognition tools should be focused on evaluating facial morphology features such as action units [e.g., OpenFace @baltruvsaitis2016openface] rather than inferring potential emotional or affective states.
 

### 2019-03-18 ACII_2019_paper/ACII_2019_word.Rmd
conclusion

@@
 ```{r dynemo_img, fig.cap=fig_nums("dynemo_img")}
-knitr::include_graphics("./img/dynemo_img.jpg")
+knitr::include_graphics(here::here("./img/dynemo_img.jpg"))
 ```
 
-After their emotion elicitation task the participants had to rate their emotional state during the task on a likert scale from 0 ("not at all") to 5 ("strongly") the six "basic" emotions (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as six "non-basic" emotions (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
+After their emotion elicitation task the participants had to rate their subjective feeling during the task on likert scales from 0 ("not at all") to 5 ("strongly") according six "basic" emotion labels (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as six "non-basic" emotion labels (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
 
@@
 First, the participants had to identify the emotional sequences by pressing the space bar of their keyboard to indicate the beginning and the end of the emotional sequences while watching the video.
-Second, the participants watched each emotional sequence previously identified and had to label the sequence using one of the 12 emotions proposed including six "basic" emotions (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) and six "non-basic" emotions (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
+Second, the participants watched each emotional sequence previously identified and had to label the sequence using one of the 12 emotions proposed including six "basic" emotion labels (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) and six "non-basic" emotion labels (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
 They also had the possibility to indicate that the sequence was not expressing one of the proposed emotion.
@@
 Affdex’s algorithm uses Histogram of Oriented Gradient (HOG) features and Support Vector Machine (SVM) classifiers in order to recognize facial expressions.
-For each video frame, Affdex identify the probability of the face as expressing one of the six "basic" emotions (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as additional psychological states such as *valence*, *engagement* or *contempt*, and facial features such as *cheek raise*, *eye widen* or *jaw drop*.
+For each video frame, Affdex identify the probability of the face as expressing one of the six "basic" emotion labels (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as additional psychological states such as *valence*, *engagement* or *contempt*, and facial features such as *cheek raise*, *eye widen* or *jaw drop*.
 
@@
 
-Emotions self-reported as being characteristic of the elicitation are compared with the emotion recognized by the human observers in a confusion matrix (`r fig_nums("confusionMatrix_sr_hr", display = "cite")`).
+Participants' subjective feeling is compared with human observers recognition in a confusion matrix (`r fig_nums("confusionMatrix_sr_hr", display = "cite")`).
+Each emotion label used to describe participants subjective feeling (i.e., the label rated with the highest value) are compared with the emotion labels which was rated with the highest score by human observers
 
@@
 
-However, the self-report show a very high proportion of undetermined emotional states which reveals not only the possibility of the emotion elicitation tasks to trigger more than one emotion but also the potential limit of using 6-points likert scales for which the participants can easily score to the maximum for more than one emotion.
+However, the self-report show a very high proportion of undetermined emotional states which reveals the potential limit of using 6-points likert scales for which the participants can easily score to the maximum for more than one emotion.
 
@@
 
-As in the previous analysis, emotions self-reported as being characteristic of the elicitation are compared with the emotion recognized by the automatic classifier in a confusion matrix (`r fig_nums("confusionMatrix_sr_ar", display = "cite")`).
+Similarly to the previous analysis, a confusion matrix was used to compare participants subjective feeling with the emotion label recognized by the automatic classifier (`r fig_nums("confusionMatrix_sr_ar", display = "cite")`).
 
@@
 The results reveals a low accuracy for both humans and automatic classifier to accurately identify the inner emotional states of these individuals based on their facial expressions.
+Moreover it appeared that automatic classifier obtain a higher accuracy to recognize the corresponding emotion label than humans observers are.
+A possible explaination is that human oberservers are usually not using emotion label to recognize emotion and this type of paradigm provides non accurate results.
+A second explaination would rely on the type of annotation procedure which is different of the classic paradigm in which observers are asked to attribuate a label to the video.
+Instead, the human observers had to annotate sequences of the video.
+Whereas this paradigm is longer and more complicated, it can lead to more robust results in reducing the forced-choice biais [@russell1993forced].
+However this procedure can also reduce the human obersvers accuracy.
 
 Some limitations to this process should be stated over the use of self-reports to evaluate individual's subjective feelings.
-Accessing to the inner subjective feeling can be biased if not impossible.
-Moreover the laboratory setting can trigger ambiguous and "non-basic" emotions which were not analysed in this research.
-The procedure use for human annotation can also be incriminated.
+Accessing the inner subjective feeling can be biased if not impossible.
+Moreover the laboratory setting can trigger ambiguous and "non-basic" emotion labels which were not analysed in this research.
+The procedure used for human annotation can also be open to dispute.
 Instead of asking the human annotators to provide an unique label, a more subtle approach was chosen to mimic results provided by the automatic classifier.

### 2019-03-18 script/data_wrangling_human_recognition.R

@@ -23,8 +23,10 @@ human_recognition_data <- here::here("./data/human_recognition_data") %>%
 
 # data wrangling ---------------------------------------------------------------
 unique_ppt_count <- human_recognition_data %>% 
-  dplyr::select(source,SEXE_Juge, C_Juge) %>% 
-  unique()
+  dplyr::group_by(source,SEXE_Juge, C_Juge, C_Video) %>% 
+  dplyr::summarise(n = 1) %>% 
+  dplyr::group_by(source,SEXE_Juge, C_Juge) %>% 
+  dplyr::summarise(n = n())
 
 unique_ppt <- human_recognition_data %>% 
   dplyr::select(-E_Detectee, -TC_Debut, -TC_Fin) %>% 

### 2019-03-18 ACII_2019_paper/ACII_2019_word.Rmd
word version

@@
   caption = "Example of a front facing recording synced with the full view of the participant and the elicitation task.
-All participants gave their consent for their data to be processed in non-commercial research.
 This picture is taken from a pilot with projects collaborators and all gave a consent for the publication of their photos and videos.")
@@
 All the participants gave their agreement on their data and video to be processed for research purpose only.
-
-and did not guess the rela purpose of the task (e.g. being filmed while they were performing an emotional indiuction task). garantuee of genuine facial emotional expressions.
 
@@
 A possible explaination is that human oberservers are usually not using emotion label to recognize emotion and this type of paradigm provides non accurate results.
-A second explaination would rely on the type of annotation procedure which is different of the classic paradigm in which observers are asked to attribuate a label to the video.
-Instead, the human observers had to annotate sequences of the video.
-Whereas this paradigm is longer and more complicated, it can lead to more robust results in reducing the forced-choice biais [@russell1993forced].
-However this procedure can also reduce the human obersvers accuracy.
 
@@
 Accessing the inner subjective feeling can be biased if not impossible.
-Moreover the laboratory setting can trigger ambiguous and "non-basic" emotion labels which were not analysed in this research.
-The procedure used for human annotation can also be open to dispute.
+Moreover the procedure used for human annotation can also be open to dispute.
 Instead of asking the human annotators to provide an unique label, a more subtle approach was chosen to mimic results provided by the automatic classifier.
-In this regard, the results of the human annotation could have been more ambiguous because it is not the natural way that people are inferring human emotions.
-Finally, the automatic classifier algorithm can also be problematic.
-Based on training datasets which are most of the time using prototypical facial expression of the "basic" emotion labels, the algorithm to classify facial expressions can be held in check by the spontaneity of the expressions analysed which are less likely to be prototypical.
+Whereas this paradigm is longer and more complicated, it can lead to more robust results in reducing the forced-choice biais [@russell1993forced].
+However this procedure can also reduce the human obersvers accuracy.
+In this regard, the results of the human annotation could have been more ambiguous because it is not the natural way that people are inferring meaning from facial expressions.
+An alternative explaination relies in reducing the recognition biais involved in the classic recognition paradigm.
+Classic forced-choice paradigms obtain artificial high results, thus by using a more evolved approach observers' accuracy may have been reduced.
 
 Considering the above, the results provides an additional evidence that individuals' subjective feeling can not be inferred from facial expressions and in our case invalidate the hypothesis of hardwired emotions.
-Advancements in identifying "non-basic" emotion labels [@mcduff2016discovering] as well as non-prototypical facial expression have been made in the developpement of automatic facial expression recognition tools, however this result suggests that automatic facial expression recognition tools should be focused on evaluating facial morphology features such as action units [e.g., OpenFace @baltruvsaitis2016openface] rather than inferring potential emotional or affective states.
+Even if emotions were hardwired, in everyday life one does not observe prototypical facial expressions and therefore research should be focused on analysing non-prototypical facial expressions.
+Advancements in identifying "non-basic" emotion labels [@mcduff2016discovering] as well as non-prototypical facial expression have been made in the developpement of automatic facial expression recognition tools.
+However this result suggests that automatic facial expression recognition tools should evaluate facial morphology features such as action units [e.g., OpenFace @baltruvsaitis2016openface] rather than inferring potential emotional or affective states.
 

### 2019-03-19 ACII_2019_paper/ACII_2019_word.Rmd
prop_sr

@@
 =============
-```{r}
-# note to myself: instead of calculating the proportion on all the video, calculate the marginal proportion
-```
+
 With the development of commercial automatic facial expression recognition tools [see @dupre2018accuracy for a non-exhaustive list of available tools], industries and governments are gradually implementing this technology in order to track humans' emotions in various scenarios (e.g., marketing, healthcare, automotive to name a few).
@@
   dplyr::rename(hr_emotion = Prediction, sr_emotion = Reference) %>% 
-  dplyr::mutate(prop_video = Freq/sum(Freq))
+  dplyr::mutate(prop_video = Freq/sum(Freq)) %>% 
+  dplyr::group_by(sr_emotion) %>% 
+  dplyr::mutate(prop_sr = Freq/sum(Freq)) %>% 
+  dplyr::ungroup()
 
@@
 
-The result of the confusion matrix show a low agreement between emotion felt during the elicitation and emotion recognized by the human annotators (Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "happiness", "prop_video"])`), *surprise* (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "surprise", "prop_video"])`) and *disgust* (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_video"])`).
+The result of the confusion matrix show a low agreement between emotion felt during the elicitation and emotion recognized by the human annotators (Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "happiness", "prop_sr"]))` of the video self-reported), *surprise* (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported) and *disgust* (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_sr"]))` of the videos self-reported).
 Sensitivity, specificity, precision and F1 score for each emotion can be found `r table_nums("confusionTable_sr_hr", display = "cite")`.
-Interestingly human annotators seem to recognize as *surprise* videos in which *happiness* was the highest self-reported emotion (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "happiness", "prop_video"])`), and in a lower instance *happiness* videos in which *surprise* was the highest self-reported emotion (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "surprise", "prop_video"])`).
+Interestingly human annotators seem to recognize as *surprise* videos in which *happiness* was the highest self-reported emotion (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "happiness", "prop_sr"]))` of the videos self-reported), and in a lower instance *happiness* videos in which *surprise* was the highest self-reported emotion (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
 
@@
   dplyr::rename(ar_emotion = Var1, sr_emotion = Var2) %>% 
-  dplyr::mutate(prop_video = Freq/sum(Freq))
+  dplyr::mutate(prop_video = Freq/sum(Freq)) %>% 
+  dplyr::group_by(sr_emotion) %>% 
+  dplyr::mutate(prop_sr = Freq/sum(Freq)) %>% 
+  dplyr::ungroup()
 
@@
 Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are similar to the ones with human observers (`r table_nums("confusionTable_sr_ar", display = "cite")`).
-Overall there is a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_ar$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_ar$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_ar$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_video"])`) and *surprise* (`r scales::percent(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "surprise" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_video"])`).
-Surprisingly the automatic classifier incorrectly recognized as *disgust* an important proportion of videos in which *happiness* was the highest self-reported emotion (`r scales::percent(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_video"])`).
-In parallel, the automatic classifier recognized as *happiness* and *disgust* videos in which *surprise* was the highest self-reported emotion (respectively `r scales::percent(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_video"])` and `r scales::percent(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_video"])`).
+Overall there is a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_ar$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_ar$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_ar$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` of the video self-reported) and *surprise* (`r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "surprise" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
+Surprisingly the automatic classifier incorrectly recognized as *disgust* an important proportion of videos in which *happiness* was the highest self-reported emotion (`r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` of the videos self-reported).
+In parallel, the automatic classifier recognized as *happiness* and *disgust* videos in which *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
 

### 2019-03-21 ACII_2019_paper/ACII_2019_word.Rmd
EQ

@@
 ---
-title: "Are facial expressions the genuine display of individuals' subjective feeling? A comparison of human and automatic recognition. "
+title: "Coherence between subjective feeling and facial expression: a time-series analysis of human and automatic recognition."
 output: word_document
@@
 fig_nums(
-  name = "roc", 
-  caption = "ROC curve comparing the accuracy in inferring subjective feelings from facial expressions by human observers and automatic recognition.")
+  name = "confusionMatrix_hr_ar", 
+  caption = "Confusion matrix of between the human obersers and automatic classifier recognition.")
 ```
@@
 This annotation procedure results in a uni-dimensional time-series for each video per human observer identifying for each second of the video which emotion was recognized.
-Then time-series corresponding to the same video were aggregated to calculate the proportion of human observers for each second of the video per emotional label.
+Then time-series corresponding to the same video were aggregated to calculate the proportion of human observers for each second of the video per emotional label (EQ1).
+
+$$x_{video_{i}.label_{i}.t_{i}} = \frac{n_{video_{i}.label_{i}.t_{i}}}{n_{video_{i}}}$$
+
 The sum of each label proportion per second was used as a score to determine which labels corresponds to the overall video (i.e., the highest score).
+
+$$x_{video.label.t_{i}} = \frac{n_{video.label.t_{i}}}{n_{video}}$$
+
 In case of more than one label having the maximum value, the emotion is described as undetermined.
@@
 
-Whereas the self-reports, the human annotations and the automatic recognition include data on "non-basic" emotion labels and features, the analysis is performed using only the six "basic" emotion labels in order to compare them.
+Whereas the self-reports, the human observation and the automatic recognition include data on "non-basic" emotion labels and features, the analysis is performed using only the six "basic" emotion labels in order to compare them.
 The maximum score for self-reports, human annotations and automatic recognition is used to label the video.
@@
   dplyr::ungroup() %>% 
+  dplyr::filter(hr_emotion != "undetermined") %>% 
+  dplyr::filter(sr_emotion != "undetermined") %>% 
   dplyr::mutate_if(is.character,as.factor)
@@
   dplyr::rename(hr_emotion = Prediction, sr_emotion = Reference) %>% 
+  dplyr::mutate(n_video = sum(Freq)) %>% 
   dplyr::mutate(prop_video = Freq/sum(Freq)) %>% 
   dplyr::group_by(sr_emotion) %>% 
+  dplyr::mutate(n_sr = sum(Freq)) %>%
   dplyr::mutate(prop_sr = Freq/sum(Freq)) %>% 
@@
   ggplot(mapping = aes(x = sr_emotion, y = hr_emotion)) +
-  geom_tile(aes(fill = Freq), colour = "white") +
+  geom_tile(aes(fill = prop_sr), colour = "white") +
   geom_text(
-    aes(label = Freq),
+    aes(label = paste0(scales::percent(prop_sr), "\n(", Freq,"/",n_sr, ")")),
     color = "black",
+    size = 4,
     family="serif",
-    size = 4,
     parse = FALSE,
-    lineheight = 0.7
-  ) +
+    lineheight = 0.7)+
   scale_fill_gradient(
-    name = "Frequency",
+    name = "Proportion of Recognized\nSelf-Reported Emotion",
     low = "white", 
     high = "red", 
+    labels = scales::percent,
+    limits = c(0,1)
   ) +
@@
         legend.text = element_text(size=8),
-        legend.position = "none")
+        legend.position = "bottom")
 ```
 
-The result of the confusion matrix show a low agreement between emotion felt during the elicitation and emotion recognized by the human annotators (Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "happiness", "prop_sr"]))` of the video self-reported), *surprise* (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported) and *disgust* (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_sr"]))` of the videos self-reported).
-Sensitivity, specificity, precision and F1 score for each emotion can be found `r table_nums("confusionTable_sr_hr", display = "cite")`.
-Interestingly human annotators seem to recognize as *surprise* videos in which *happiness* was the highest self-reported emotion (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "happiness", "prop_sr"]))` of the videos self-reported), and in a lower instance *happiness* videos in which *surprise* was the highest self-reported emotion (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
+The result of the confusion matrix show a moderate agreement between emotion felt during the elicitation and emotion recognized by the human observers (Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *disgust* (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_sr"]))` of the videos self-reported).
+Sensitivity, specificity, precision and F1 score for each emotion reveals that *happiness* has the highest coherence ratio whereas *sadness* has the lowest coherence ratio between true positives and false positives `r table_nums("confusionTable_sr_hr", display = "cite")`.
+Interestingly human observers seem to recognize *surprise* expressed in videos where *anger*, *fear* *happiness* and *sadness* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "anger", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "fear", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "happiness", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "sadness", "prop_sr"]))` of the videos self-reported), and in a lower instance *happiness* was recognized in videos where *fear* and *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "fear", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
 
@@
 ```{r sr_ar}
-comparison_sr_ar <- dplyr::inner_join(
-    self_report_score, 
-    automatic_recognition_score, 
-    by = "C_Video") %>% 
+comparison_sr_ar <- dplyr::inner_join(self_report_score,automatic_recognition_score, by = "C_Video") %>% 
   dplyr::ungroup() %>% 
+  dplyr::filter(ar_emotion != "undetermined") %>% 
+  dplyr::filter(sr_emotion != "undetermined") %>% 
   dplyr::mutate_if(is.character,as.factor)
 
-dat_table <- table(
-  comparison_sr_ar$ar_emotion,
-  comparison_sr_ar$sr_emotion
+confusionMatrix_sr_ar <- caret::confusionMatrix(
+  data = comparison_sr_ar$ar_emotion,
+  reference = comparison_sr_ar$sr_emotion
   )
-emotions <- c("anger","disgust","fear","happiness","sadness","surprise", "undetermined")
-emo_table <- matrix(
-  0,
-  length(emotions),
-  length(emotions),
-  dimnames = list(emotions, emotions))
-emo_table[row.names(dat_table), colnames(dat_table)] <- dat_table
-
-confusionMatrix_sr_ar <- caret::confusionMatrix(as.table(emo_table))
 ```
@@
   as.data.frame() %>% 
-  dplyr::rename(ar_emotion = Var1, sr_emotion = Var2) %>% 
+  dplyr::rename(ar_emotion = Prediction, sr_emotion = Reference) %>% 
+  dplyr::mutate(n_video = sum(Freq)) %>% 
   dplyr::mutate(prop_video = Freq/sum(Freq)) %>% 
   dplyr::group_by(sr_emotion) %>% 
+  dplyr::mutate(n_sr = sum(Freq)) %>%
   dplyr::mutate(prop_sr = Freq/sum(Freq)) %>% 
@@
   ggplot(mapping = aes(x = sr_emotion, y = ar_emotion)) +
-  geom_tile(aes(fill = Freq), colour = "white") +
+  geom_tile(aes(fill = prop_sr), colour = "white") +
   geom_text(
-    aes(label = Freq),
+    aes(label = paste0(scales::percent(prop_sr), "\n(", Freq,"/",n_sr, ")")),
     color = "black",
+    size = 4,
     family="serif",
-    size = 4,
     parse = FALSE,
-    lineheight = 0.7
-  ) +
+    lineheight = 0.7)+
   scale_fill_gradient(
-    name = "Frequency",
+    name = "Proportion of Recognized\nSelf-Reported Emotion",
     low = "white", 
     high = "red", 
+    labels = scales::percent,
+    limits = c(0,1)
   ) +
@@
         legend.text = element_text(size=8),
-        legend.position = "none")
+        legend.position = "bottom")
 ```
@@
 Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are similar to the ones with human observers (`r table_nums("confusionTable_sr_ar", display = "cite")`).
-Overall there is a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_ar$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_ar$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_ar$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` of the video self-reported) and *surprise* (`r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "surprise" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
-Surprisingly the automatic classifier incorrectly recognized as *disgust* an important proportion of videos in which *happiness* was the highest self-reported emotion (`r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` of the videos self-reported).
-In parallel, the automatic classifier recognized as *happiness* and *disgust* videos in which *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
+Overall there is a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_ar$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_ar$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_ar$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` of the video self-reported).
+Surprisingly the automatic classifier incorrectly recognized as *disgust* an important proportion of videos in which *anger*, *happiness* and *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "anger", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
+In parallel, the automatic classifier recognized as *happiness* videos in which *fear* and *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "fear", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
 
@@
 
-As previously mentioned, the accuracy of humans observers and the automatic classifier have some similarities.
-In order to compare which recognition has the highest accuracy a Receiver Operating Characteristic (ROC) curve was calculated (`r fig_nums("roc", display = "cite")`).
+As previously mentioned, the accuracy of humans observers and the automatic classifier have some similarities in there mistakes.
+However human observers appears to be more accurate than the automatic classifier to recognize individuals' subjective feeling (human observer Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`; automatic classifier Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`.
+A third confusion matrix is used to compare similarities between human observers and automatic classifier (`r fig_nums("confusionMatrix_hr_ar", display = "cite")`).
 
-```{r roc, fig.height=6, fig.cap=fig_nums("roc")}
-roc_sr_hr <- comparison_sr_hr %>% 
-  dplyr::mutate_if(is.factor, as.character) %>% 
-  dplyr::mutate(correct_classification = case_when(
-    sr_emotion == hr_emotion ~ 1,
-    sr_emotion != hr_emotion ~ 0
-  )) %>% 
-  dplyr::rename(rec_score = hr_score) %>% 
-  dplyr::mutate(rec_type = "human")
+```{r confusionMatrix_hr_ar, fig.height=4, fig.cap=fig_nums("confusionMatrix_hr_ar")}
+comparison_hr_ar <- dplyr::inner_join(human_recognition_score,automatic_recognition_score, by = "C_Video") %>% 
+  dplyr::ungroup() %>% 
+  dplyr::filter(ar_emotion != "undetermined") %>% 
+  dplyr::filter(hr_emotion != "undetermined") %>% 
+  dplyr::mutate_if(is.character,as.factor)
 
-roc_sr_ar <- comparison_sr_ar %>% 
-  dplyr::mutate_if(is.factor, as.character) %>% 
-  dplyr::mutate(correct_classification = case_when(
-    sr_emotion == ar_emotion ~ 1,
-    sr_emotion != ar_emotion ~ 0
-  )) %>% 
-  dplyr::rename(rec_score = ar_score) %>% 
-  dplyr::mutate(rec_type = "automatic")
+confusionMatrix_hr_ar <- caret::confusionMatrix(
+  data = comparison_hr_ar$ar_emotion,
+  reference = comparison_hr_ar$hr_emotion
+)
 
-auc <- dplyr::bind_rows(roc_sr_hr,roc_sr_ar) %>% 
-  tidyr::nest(-rec_type) %>%
-  dplyr::mutate(rec_type = as.factor(rec_type),
-                roc_obj = purrr::map(data, ~ roc(.$correct_classification,.$rec_score)),
-                roc_value = purrr::map(roc_obj,auc)) %>%
-  tidyr::unnest(roc_value) %>% 
-  dplyr::select(rec_type,roc_value) %>% 
-  dplyr::mutate(roc_value = round(roc_value,2))
+confusionMatrix_hr_ar_freq <- confusionMatrix_hr_ar$table %>% 
+  as.data.frame() %>% 
+  dplyr::rename(ar_emotion = Prediction, hr_emotion = Reference) %>% 
+  dplyr::mutate(n_video = sum(Freq)) %>% 
+  dplyr::mutate(prop_video = Freq/sum(Freq)) %>% 
+  dplyr::group_by(hr_emotion) %>% 
+  dplyr::mutate(n_hr = sum(Freq)) %>%
+  dplyr::mutate(prop_hr = Freq/sum(Freq)) %>% 
+  dplyr::ungroup()
 
-dplyr::bind_rows(roc_sr_hr,roc_sr_ar) %>% 
-  ggplot(aes(d = correct_classification, m = rec_score, color = rec_type)) + 
-  geom_roc(labels = FALSE,pointsize = 0) +
-  style_roc(xlab = "False Positive Rate",
-            ylab = "True Positive Rate", 
-            theme = theme_minimal) + 
-  scale_color_manual(
-    "Classifiers",
-    values = c("#e41a1c","#377eb8")
+confusionMatrix_hr_ar_freq %>% 
+  ggplot(mapping = aes(x = hr_emotion, y = ar_emotion)) +
+  geom_tile(aes(fill = prop_hr), colour = "white") +
+  geom_text(
+    aes(label = paste0(scales::percent(prop_hr), "\n(", Freq,"/",n_hr, ")")),
+    color = "black",
+    size = 4,
+    family="serif",
+    parse = FALSE,
+    lineheight = 0.7)+
+  scale_fill_gradient(
+    name = "Proportion of Recognized\nEmotion",
+    low = "white", 
+    high = "red", 
+    labels = scales::percent,
+    limits = c(0,1)
   ) +
-  theme(text = element_text(size=18,family="serif"),
-        strip.text.x = element_text(face="bold",size=18),
-        axis.title.x = element_text(margin = margin(t = 20, r = 0, b = 0, l = 0)),
+  scale_x_discrete(name = "Human Observers") +
+  scale_y_discrete(name = "Automatic Classifier") +
+  theme_minimal() +
+  theme(text = element_text(size=16,family="serif"),
+        axis.text.x = element_text(size=16, angle = 45, hjust = 0.75,vjust=0.9),
+        axis.text.y = element_text(size=16, hjust = 0.5),
         axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
-        axis.text.x = element_text(size = 10),
-        axis.text.y = element_text(size = 10),
-        legend.text = element_text(size=10),
-        legend.position = "bottom") + 
-  geom_abline(intercept = 0, slope = 1, linetype = "dashed")
+        legend.text = element_text(size=8),
+        legend.position = "bottom")
 ```
 
-The ROC curve and its Area Under the Curve (AUC) values shows that the automatic classifier are more accurate than human observers in inferring subjective feeling from facial expressions (human AUC = `r auc[auc$rec_type == "human","roc_value"]` *vs.* automatic AUC = `r auc[auc$rec_type == "automatic","roc_value"]`).
+The overall agreement between human observers and the automatic classifier is in fact very low (Kappa = `r round(confusionMatrix_hr_ar$overall["Kappa"],2)`).
+Except for *happiness* and *disgust* (respectively `r scales::percent(as.numeric(confusionMatrix_hr_ar_freq[confusionMatrix_hr_ar_freq$ar_emotion == "happiness" & confusionMatrix_hr_ar_freq$hr_emotion == "happiness", "prop_hr"]))` and `r scales::percent(as.numeric(confusionMatrix_hr_ar_freq[confusionMatrix_hr_ar_freq$ar_emotion == "disgust" & confusionMatrix_hr_ar_freq$hr_emotion == "disgust", "prop_hr"]))` of common labelling), there is no clear common pattern.
+Moreover, the automatic classifier has a tendency to label as *disgust* videos labeled as *sadness* by human observers and as *happiness* videos labaled as *fear* by human observers
 
@@
 ============
-Despite being one on the most investigated question in affective science, the link between emotion felt and facial expression is a hot topic and no clear evidence have been found to definitely answer it.
+Despite being one on the most investigated question in affective science, the coherence between emotion felt and facial expression is a hot topic and no clear evidence have been found to definitely answer it.
 However, with the growing interest of industries and government to monitor individual's psychological states, evidences are showing that facial expressions are in reality not expressing emotions [@mckeown2013analogical].
@@
 The results reveals a low accuracy for both humans and automatic classifier to accurately identify the inner emotional states of these individuals based on their facial expressions.
-Moreover it appeared that automatic classifier obtain a higher accuracy to recognize the corresponding emotion label than humans observers are.
-A possible explaination is that human oberservers are usually not using emotion label to recognize emotion and this type of paradigm provides non accurate results.
+Moreover it appeared that human observers obtain a higher accuracy to recognize the corresponding emotion label than the automatic classifier is.
+A possible explaination is that human oberservers are assessing expressing sequences while having access to the full video which provides a context whereas the automatic classifier only assess the videos frame by frame.
 
@@
 Accessing the inner subjective feeling can be biased if not impossible.
-Moreover the procedure used for human annotation can also be open to dispute.
+Moreover the procedure used for human observation can also be open to dispute.
 Instead of asking the human annotators to provide an unique label, a more subtle approach was chosen to mimic results provided by the automatic classifier.
@@
 However this procedure can also reduce the human obersvers accuracy.
-In this regard, the results of the human annotation could have been more ambiguous because it is not the natural way that people are inferring meaning from facial expressions.
+In this regard, the results of the human observation could have been more ambiguous because it is not the natural way that people are inferring meaning from facial expressions.
 An alternative explaination relies in reducing the recognition biais involved in the classic recognition paradigm.
@@
 
-The authors would like to thank Brigitte Meillon and Jean Michel Adam who developed the software used to collect and preprocess human observer annotations.
+The authors would like to thank Brigitte Meillon and Jean Michel Adam who developed the software used to collect and preprocess human observer results.
 

### 2019-04-02 ACII_2019_paper/ACII_2019_word.Rmd
eq
typo

@@
 To cope with critics, several amendments have been made to the BET, increasing the number of basic emotions from six to seven  [@ekman1988universality] as well as adding the concept of "display rules" to explain cultural differences in the management of facial expressions [@ekman1987universals].
-Even if this theory obtained a popular support, it fails to explain how individuals can feel emotions without expressing them and how individuals can express emotions without feeling them, cases for which display rules cannot be applied to [@kraut1979social, @duran2017coherence].
+Even if this theory obtained a popular support, it fails to explain how individuals can feel emotions without expressing them and how individuals can express emotions without feeling them, cases for which display rules cannot be applied to [@kraut1979social; @duran2017coherence].
 
 ## The link between emotion and facial expression through the social constructivist approach
-Detractors of the Basic Emotion Theory are perceiving emotion not as genetically hardwired but as a learnt association between a given situation and an appropriate response [@averill1980constructivist, @barrett2017emotions].
+Detractors of the Basic Emotion Theory are perceiving emotion not as genetically hardwired but as a learnt association between a given situation and an appropriate response [@averill1980constructivist; @barrett2017emotions].
 For the tenants of the constructivist approach, emotions are "concepts" based on past experiences and which are "a collection of embodied, whole brain representations that predict what is about to happen in the sensory environment, what the best action is to deal with impending events, and their consequences for allostasis" [@barrett2017theory, p12].
@@
 
-$$x_{video_{i}.label_{i}.t_{i}} = \frac{n_{video_{i}.label_{i}.t_{i}}}{n_{video_{i}}}$$
+$$x_{video_{i}.label_{j}.t_{k}} = \frac{n_{video_{i}.label_{j}.t_{i}}}{n_{video_{k}}}$$
 
@@
 
-$$x_{video.label.t_{i}} = \frac{n_{video.label.t_{i}}}{n_{video}}$$
+$$label_{video_{i}} = max{\frac{n_{video.label.t_{i}}}{n_{video}}}$$
 

### 2019-04-05 ACII_2019_paper/ACII_2019_paper.Rmd
latex

--- 2019-03-15 ACII_2019_paper/ACII_2019_paper.Rmd
@@
 ---
-title: The relation betweem feeling and expressing emotions. A comparison between sujective feeling and facial expression recognition of spontaneous emotions. 
+title: "Coherence between subjective feeling and facial expression: A comparison of human and automatic recognition" 
 affiliation:
-  ## use one only of the following
-  # author-columnar: true         ## one column per author
-  institution-columnar: true  ## one column per institution (multiple autors eventually)
-  # wide: true                  ## one column wide author/affiliation fields
+  institution-columnar: true
 
@@
 abstract: |
-  The abstract goes here.
-  On multiple lines eventually.
+  While it has been taken for granted in the development of several automatic facial expression recognition tools, the question of the link between subjective feelings and facial expressions is still a subject of debate.
+On one hand the behaviorist approach conceives emotions as genetically hardwired and therefore being genuinely displayed through facial expressions.
+On the other hand the constructivist approach conceives emotions as socially constructed, the emotional meaning of a facial expression being inferred by the observer.
+In order to evaluate the link between the subjective feeling of emotions and their recognition based on facial expression, 232 videos of encoders recruited to carry out an emotion elicitation task were annotated by 1383 human observers as well as by an automatic facial expression classifier.
+Results show a low accuracy of human observers and of the automatic classifier to infer the subjective feeling from the facial expressions displayed by encoders .
+They also show a weak link between self-reported emotional states and facial emotional displays.
+Based on these results, the hypothesis of genetically hardwired emotion genuinely display is difficult to support whereas the idea of emotion and facial expression socially constructed appears to be more likely.
+Then automatic emotion recognition tools based on facial expressions should be questioned.
 header-includes:
@@
   - \usepackage{tabu}
+
 bibliography: mybibfile.bib
@@
   echo = FALSE,
-  fig.pos='H')
+  fig.pos = 'H')
 #
@@
 =============
-```{r}
-# note to myself: instead of calculating the proportion on all the video, calculate the marginal proportion
-```
 
-With the development of commercial automatic facial expression recognition tools [see @dupre2018accuracy for a non-exhaustive list of available tools], industries and governments are gradually implementing this technology in order to track humans' emotions in various scenarios (e.g., marketing, healthcare, automotive to name a few).
-Beside the ethical question of measuring human emotions, these tools are challenging psychological theories about inferring emotions from facial expressions.
-When automatic facial expression recognition tools are used to measure human emotions, they rest on the premise that facial expressions provide a direct access to expressers' emotions.
-However the link between emotion felt and facial expressions is far from being established and still remain a hot topic in psychology research.
+With the development of commercial automatic facial expression recognition tools [see @dupre2018accuracy for a non-exhaustive list of available tools], industries and governments are gradually implementing this technology in order to track humans’ emotions in various scenarios (e.g., marketing, healthcare, automotive to name a few).
+This technoloogy rests on the premise that facial expressions provide a direct access to individuals’ subjective feeling.
+Even if this premise is central to the modern mainstream approach of human emotion, recent research in affective science are challenging it.
+Once the two competing approaches briefly described as well as the predictions they respectively entail, an experiment testing these hypotheses will be presented and its results analyzed in order to provide empirical evidence to contribute to answer the question.
 
-## The link between emotion and facial expression thought the behaviorist approach
-Based on the behaviorist approach initiated by Darwin in *The Expression of the Emotions in Man and Animals* [@darwin1872expression], facial expression are conceived as a genuine display of individuals inner emotional state.
-This hypothesis is used as a basis for the Basic Emotion Theory (BET) which states that a set of six or seven emotions are universally displayed and are genetically hardwired [@ekman1992argument].
-According to this view, "when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including [...] facial expressions." [@ekman2007directed, p49].
-Even if this theory obtained a popular support, it fails to explain how individuals can feel emotions without expressing them and how individuals can express emotions without feeling them.
+## The link between emotion and facial expression through the behaviorist approach
+Based on the behaviorist approach initiated by Darwin in *The Expression of the Emotions in Man and Animals* [@darwin1872expression], facial expressions are conceived as a genuine displays of individuals’ inner emotional state.
+This hypothesis is used as a basis for the Basic Emotion Theory (BET) which states that a set of six emotions are universally displayed and are genetically hardwired not only in humans [@ekman1992argument] but also in different animal species [@de2019mama].
+According to this view, “when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including […] facial expressions.” [@ekman2007directed, p49].
+To cope with critics, several amendments have been made to the BET, increasing the number of basic emotions from six to seven [@ekman1988universality] as well as adding the concept of “display rules” to explain cultural differences in the management of facial expressions [@ekman1987universals].
 
-## The link between emotion and facial expression thought the social constructivist approach
-Detractors of the Basic Emotion Theory are perceiving emotion not as genetically hardwired but as a learnt association between a given situation and an appropriate response [@averill1980constructivist].
-For the tenants of the constructivist approach, emotions are "concepts" based on past experiences and which are "a collection of embodied, whole brain representations that predict what is about to happen in the sensory environment, what the best action is to deal with impending events, and their consequences for allostasis" [@barrett2017theory, p12].
-Following this assumption, faces are used as tools to display signals in social interactions [@crivelli2018facial].
-These signals can convey individuals' motivations and readiness [@frijda1997facial] or social messages [@fridlund1995human].
+Even if this theory obtained a popular support, it fails to explain how individuals can feel emotions without expressing them and how individuals can express emotions without feeling them, in instances in which display rules cannot be called upon [@kraut1979social; @duran2017coherence].
+These evidences have led to an alternative conception, notably the social constructivist approach.
 
-This current paper investigates the link between the subjective feeling of emotions and their recognition from facial expressions.
-If emotions are hardwired, individuals emotional subjective feeling should be correlated to the recognition of facial expressions from both human observers and automatic classifiers whereas if emotions are social constructs, no correlation between subjective feeling and facial expression recognition should be observed.
+## The link between emotion and facial expression through the social constructivist approach
+Detractors of the Basic Emotion Theory consider emotion not as genetically hardwired but as a learnt association between a given situation and an appropriate response [@averill1980constructivist; @barrett2017emotions].
+For the tenants of the constructivist approach, emotions are “concepts” based on past experiences and which are “a collection of embodied, whole brain representations that predict what is about to happen in the sensory environment, what the best action is to deal with impending events, and their consequences for allostasis” [@barrett2017theory, p12].
+Following this assumption, faces are best conceived as tools displaying signals in social interactions [@crivelli2018facial].
+These signals can convey individuals’ motivations and readiness [@frijda1997facial] or social messages [@fridlund1995human].
+Therefore, facial expressions are thought as behaviors which meaning is inferred by the observer.
+Findings support this observer dependence [@lindquist2013s; @niedenthal2017embodied].
+They show that to make meaning of another person’s facial behavior, the perceiver relies in particular on her/his knowledge about emotion categories.
+
+## The recognition of emotional facial expressions
+Regarding emotional facial expression (EFE) recognition, the behaviorist approach assertions and the constructivist approach claims lead to two divergent predictions.
+The foremost postulates that, when triggered, each basic emotion is expressed by a prototypical face (non basic emotions being blends of the basic ones).
+Basic emotions are easily recognized by all human observers and emotional states are accessible by facial measurement.
+As the recognition of facial expressions is based on the identification of specific patterns of facial movements, it implies that EFE recognition of both human observers and automatic classifiers should be as accurate.
+The constructivist approach, in contrast, affirms that facial expressions do not provide a direct access to individuals’ subjective feeling.
+Even if the face possibly moves during an emotional episode, facial muscle movements are not linked in a one-to-one manner to a specific discrete emotional experience [@doyle2017language, p418].
+Instead, emotions are mentally constructed by the perceiver and mental categories of emotions are needed to accurately categorize facial movements.
+Therefore, since there is no emotional prototypical face, one should expect human observers’ superior ability to accurately recognize EFE as compared to automatic EFE recognition tools.
+
+The aim of the current paper is to investigate the link between the subjective feeling of emotions and its recognition from facial expressions in order to give credit either to the behaviorist plea or to the constructivist’s one.
+Contrary to most studies using posed and static EFE, the present study focuses on natural EFE.
+Spontaneous and dynamic facial reactions to emotional elicitations are under consideration to ensure the generalizability of the results to emotional behaviors in ordinary life.
 
@@
 
-To evaluate the link between subjective feeling of emotions and their recognition from facial expressions, participants were recruited to perform an emotion elicitation task while their facial expression was video recorded.
-Then, the videos was shown to human observer and analysed by an automatic classifier in order to identify which emotion was displayed.
+To evaluate the link between subjective feeling of emotions and their recognition from facial expressions, encoders were first recruited to perform an emotion elicitation task while their facial expression was video recorded.
+Then, the videos of the encoders’ faces were shown to human observers and were also analysed by an automatic classifier in order to identify which emotion was displayed.
 
@@
 
-For the emotion elicitation experiment, `r nrow(self_report_data)` French participants (`r self_report_gender[self_report_gender$genre_c == "F","n"]` females, `r self_report_gender[self_report_gender$genre_c == "H","n"]` males, *M*age = `r self_report_age[,"Age_m"]`, *SD*age = `r self_report_age[,"Age_sd"]`) were recruited to perform one out of 11 emotions elicitation tasks designed to trigger a positive, negative or neutral emotional state [see @tcherkassof2013dynemo for a description of tasks and procedure].
-Participants' face were recorded using an hidden camera resulting `r nrow(self_report_data)` front facing 768x576 videos varying from `r min(metadata_video$ffprobe_duration)`s to `r max(metadata_video$ffprobe_duration)`s.
+For the emotion elicitation experiment, `r nrow(self_report_data)` French participants (`r self_report_gender[self_report_gender$genre_c == "F","n"]` females, `r self_report_gender[self_report_gender$genre_c == "H","n"]` males, *M*age = `r self_report_age[,"Age_m"]`, *SD*age = `r self_report_age[,"Age_sd"]`) were recruited to perform one out of 11 emotion elicitation tasks designed to trigger a positive, a specific negative or a neutral emotional state.
+Encoders’ face were recorded using an hidden camera resulting `r nrow(self_report_data)` front facing 768x576 videos varying from `r min(metadata_video$ffprobe_duration)`s to `r max(metadata_video$ffprobe_duration)`s (Figure \ref{fig:dynemo_img}).
+These recordings form the DynEmo database [see @tcherkassof2013dynemo for a full description of tasks and procedure].
 
-After their emotion elicitation task the participants had to rate their emotional state during the task on a likert scale from 0 ("not at all") to 5 ("strongly") the six "basic" emotions (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as six "non-basic" emotions (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
+```{r dynemo_img, fig.height=3, fig.cap="\\label{fig:dynemo_img}Example of a front facing recording synced with the full view of the participant and the elicitation task.
+This picture is taken from a pilot with projects collaborators and all gave a consent for the publication of their photos and videos."}
+# knitr::include_graphics("C:/Users/dupred/Desktop/Projects/dynemo_feel_express/img/dynemo_img.png", auto_pdf = FALSE)
+raster::stack(here::here("./img/dynemo_img.jpg")) %>%
+  raster::plotRGB()
+```
 
-Finally, a debriefing session was perform to ensure that participants were not durably affected by the emotion elicitation task.
-All the participants gave their agreement on their data and video to be processed for research purpose only.
+After the emotion elicitation task the encoders rated their subjective feeling on Likert scales from 0 ("not at all") to \nolinebreak 5 \nolinebreak ("strongly") related to six “basic” emotion labels (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as six “non-basic” emotion labels (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
+
+Finally, a debriefing session was performed to ensure that encoders were not durably affected by the emotion elicitation task.
+The debriefing was also used to check that encoders did not guess the real purpose of the experiment (e.g.
+being filmed while they were performing an emotional elicitation task) to guarantee facial expressions’ genuineness.
+All encoders gave their agreement on their data and video to be processed for research purpose only.
 
@@
 
-For the human facial expression recognition method, `r nrow(unique_ppt_count)` student participants were recruited to annotate `r nrow(unique_video_count)` out of the `r nrow(self_report_data)` video, therefore only the `r nrow(unique_video_count)` annotated videos will be analysed in this paper.
-Each participants had to annotate between `r min(unique_ppt_count$n)` and `r max(unique_ppt_count$n)` videos resulting that each video was annotated `r mean(unique_video_count$n) %>% round(0)` times on average (*SD* = `r sd(unique_video_count$n) %>% round(0)`).
+For the human facial expression recognition method, `r nrow(unique_ppt_count)` \nolinebreak student participants were recruited to annotate `r nrow(unique_video_count)` out of the `r nrow(self_report_data)` videos, therefore only the `r nrow(unique_video_count)` annotated videos will be analysed in this paper.
+Because videos have different durations, participants had to annotate a series of video corresponding to 30min long in total.
+Each video was annotated `r mean(unique_video_count$n) %>% round(0)` times on average (*SD* = `r sd(unique_video_count$n) %>% round(0)`).
 
 The annotation of facial expressions was performed on-site using *Oudjat*, a software for designing video annotation experiments [@dupre2015oudjat].
-For each video, the annotation procedure hat two steps.
+For each video, the annotation procedure followed two steps.
 First, the participants had to identify the emotional sequences by pressing the space bar of their keyboard to indicate the beginning and the end of the emotional sequences while watching the video.
-Second, the participants watched each emotional sequence previously identified and had to label the sequence using one of the 12 emotions proposed including six "basic" emotions (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) and six "non-basic" emotions (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
-They also had the possibility to indicate that the sequence was not expressing one of the proposed emotion.
+Second, the participants watched each emotional sequence previously identified and labeled the sequence using one of the 12 \nolinebreak emotions proposed including six “basic” emotion labels (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) and six "non-basic" emotion labels (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
+They also had the possibility to indicate that the sequence was expressing none of the proposed emotion labels.
 
 This annotation procedure results in a uni-dimensional time-series for each video per human observer identifying for each second of the video which emotion was recognized.
-Then time-series corresponding to the same video were aggregated to calculate the proportion of human observers for each second of the video per emotional label.
-The sum of each label proportion per second was used as a score to determine which labels corresponds to the overall video (i.e., the highest score).
-In case of more than one label having the maximum value, the emotion is described as undetermined.
+Then, time-series corresponding to the same video were aggregated to calculate the proportion of human observers $x_{video_{i}.label_{j}.t_{k}}$ for each second of the video per emotional label (EQ\ref{eq:1}).
+
+\begin{equation}
+\label{eq:1}
+x_{video_{i}.label_{j}.t_{k}} = \frac{n_{video_{i}.label_{j}.t_{k}}}{n_{video_{k}}}
+\end{equation}
+
+where *i* is one of the `r nrow(unique_video_count)` videos, *j* is one of the six “basic” emotion labels, *k* for each second of the video.
 
@@
 Affdex’s algorithm uses Histogram of Oriented Gradient (HOG) features and Support Vector Machine (SVM) classifiers in order to recognize facial expressions.
-For each video frame, Affdex identify the probability of the face as expressing one of the six "basic" emotions (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as additional psychological states such as *valence*, *engagement* or *contempt*, and facial features such as *cheek raise*, *eye widen* or *jaw drop*.
+For each video frame, Affdex identifies the probability $p_{video_{i}.label_{j}.t_{k}}$ from 0 to 100 (rescaled to 0 to 1 for the analysis) of the face as expressing one of the six “basic” emotion labels (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as additional psychological states such as *valence*, *engagement* or *contempt*, and facial features such as *cheek raise*, *eye widen* or *jaw drop*.
 
-To determine which of the six "basic" emotion can be used to identify each video, the recognition probability for each label by frame was converted into odd ratio by frame [@dente2017measures].
-The highest sum of each odd ratio time-series defines the label recognized by the automatic classifier.
+For both human and automatic recognition, to determine which of the six “basic” emotions can be used to identify each video, the recognition probability for each label by frame was converted into odd ratio by label [@dente2017measures].
+The highest sum of each odd ratio time-series defines the label recognized by the automatic classifier (EQ\ref{eq:2} for human recognition and EQ\ref{eq:3} for automatic recognition).
+\begin{equation}
+\label{eq:2}
+video_{i}.label = \max\left(\frac{\sum_{k=1}^{n}x_{video_{i}.label_{j}.t_{k}}}{\sum_{k=1}^{n}x_{video_{i}.t_{k}}}\right)
+\end{equation}
+
+\begin{equation}
+\label{eq:3}
+video_{i}.label = \max\left(\frac{\sum_{k=1}^{f}p_{video_{i}.label_{j}.t_{k}}}{\sum_{k=1}^{f}p_{video_{i}.t_{k}}}\right)
+\end{equation}
+
+where *i* is one of the `r nrow(unique_video_count)` videos, *j* is one of the six “basic” emotion labels, *k* for each second of the *n* second video or for each frame of the *f* frame video.
 
@@
 
-Whereas the self-reports, the human annotations and the automatic recognition include data on "non-basic" emotions and features, the analysis is performed using only the six "basic" emotions in order to compare them.
+Since encoders’ self-reports, human annotations and the automatic recognition include data on “non-basic” emotion labels and features, the analysis is performed using only the six “basic” emotion labels in order to compare them.
 The maximum score for self-reports, human annotations and automatic recognition is used to label the video.
@@
 
-Emotions self-reported as being characteristic of the elicitation are compared with the emotion recognized by the human observers in a confusion matrix (Figure \ref{fig:confusionMatrix_sr_hr}).
+Encoders’ subjective feeling is compared with human observers recognition in a confusion matrix (Figure \ref{fig:confusionMatrix_sr_hr}).
+Each emotion label used to describe encoders’ self-reported subjective feeling (i.e., the label rated with the highest value) is compared with the emotion labels which were rated with the highest score by human observers.
 
-```{r confusionMatrix_sr_hr, fig.height=4, fig.cap="Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the human observers."}
+```{r confusionMatrix_sr_hr, fig.height=5, fig.cap="\\label{fig:confusionMatrix_sr_hr}Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the human observers."}
 confusionMatrix_sr_hr_freq <- confusionMatrix_sr_hr$table %>% 
@@
   dplyr::rename(hr_emotion = Prediction, sr_emotion = Reference) %>% 
-  dplyr::mutate(prop_video = Freq/sum(Freq))
+  dplyr::mutate(n_video = sum(Freq)) %>% 
+  dplyr::mutate(prop_video = Freq/sum(Freq)) %>% 
+  dplyr::group_by(sr_emotion) %>% 
+  dplyr::mutate(n_sr = sum(Freq)) %>%
+  dplyr::mutate(prop_sr = Freq/sum(Freq)) %>% 
+  dplyr::ungroup()
 
@@
         legend.text = element_text(size=8),
-        legend.position = "none")
+        legend.position = "bottom")
 ```
 
-The result of the confusion matrix show a low agreement between emotion felt during the elicitation and emotion recognized by the human annotators (Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "happiness", "prop_video"])`), *surprise* (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "surprise", "prop_video"])`) and *disgust* (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_video"])`).
-Sensitivity, specificity, precision and F1 score for each emotion can be found Table \ref{tab:confusionTable_sr_hr}.
-Interestingly human annotators seem to recognize as *surprise* videos in which *happiness* was the highest self-reported emotion (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "happiness", "prop_video"])`), and in a lower instance *happiness* videos in which *surprise* was the highest self-reported emotion (`r scales::percent(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "surprise", "prop_video"])`).
+Results of the confusion matrix show a moderate agreement between the emotion felt by the encoder during the elicitation and the emotion recognized by the human observers (Accuracy \nolinebreak = \nolinebreak `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *disgust* (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_sr"]))` of the videos self-reported).
+Sensitivity, specificity, precision and F1 score for each emotion reveals that *happiness* has the highest coherence ratio whereas *sadness* has the lowest coherence ratio between true positives and false positives (Table \ref{table:confusionTable_sr_hr}).
+Interestingly human observers seem to recognize *surprise* expressed in videos where *anger*, *fear* *happiness* and *sadness* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "anger", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "fear", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "happiness", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "sadness", "prop_sr"]))` of the videos self-reported), and in a lower instance *happiness* was recognized in videos where *fear* and *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "fear", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
 
@@
     "latex",
-    caption = "Agreement accuracy metrics for each emotion. ",
+    caption = "\\label{table:confusionTable_sr_hr}Agreement accuracy metrics for each emotion.",
     booktabs = TRUE,
@@
    add_footnote("Note. \\textit{na.} values are produced when not enough data are available to compute accuracy indicators.", notation="none", escape = FALSE)
+
+# confusionMatrix_sr_hr_table %>%
+#   papaja::apa_table(
+#     caption = table_nums("confusionTable_sr_hr"),
+#     row.names = FALSE,
+#     format = "word",
+#     escape = FALSE
+#     )
 ```
 
-However, the self-report show a very high proportion of undetermined emotional states which reveals not only the possibility of the emotion elicitation tasks to trigger more than one emotion but also the potential limit of using 6-points likert scales for which the participants can easily score to the maximum for more than one emotion.
+However, self-reports show a significant proportion of undetermined emotional states (`r scales::percent(nrow(list_automatic_ties)/358)` of the 358 videos) which reveals the potential limit of using 6-points likert scales to measure emotional self-reports.
+Indeed, encoders can easily score to the maximum for more than one emotion.
 
@@
 ```{r sr_ar}
-comparison_sr_ar <- dplyr::inner_join(
-    self_report_score, 
-    automatic_recognition_score, 
-    by = "C_Video") %>% 
+comparison_sr_ar <- dplyr::inner_join(self_report_score,automatic_recognition_score, by = "C_Video") %>% 
   dplyr::ungroup() %>% 
+  dplyr::filter(ar_emotion != "undetermined") %>% 
+  dplyr::filter(sr_emotion != "undetermined") %>% 
   dplyr::mutate_if(is.character,as.factor)
 
-dat_table <- table(
-  comparison_sr_ar$ar_emotion,
-  comparison_sr_ar$sr_emotion
+confusionMatrix_sr_ar <- caret::confusionMatrix(
+  data = comparison_sr_ar$ar_emotion,
+  reference = comparison_sr_ar$sr_emotion
   )
-emotions <- c("anger","disgust","fear","happiness","sadness","surprise", "undetermined")
-emo_table <- matrix(
-  0,
-  length(emotions),
-  length(emotions),
-  dimnames = list(emotions, emotions))
-emo_table[row.names(dat_table), colnames(dat_table)] <- dat_table
-
-confusionMatrix_sr_ar <- caret::confusionMatrix(as.table(emo_table))
 ```
 
-As in the previous analysis, emotions self-reported as being characteristic of the elicitation are compared with the emotion recognized by the automatic classifier in a confusion matrix (Figure \ref{fig:confusionMatrix_sr_ar}).
+Similarly to the previous analysis, a confusion matrix was used to compare encoders’ subjective feeling with the emotion label recognized by the automatic classifier (Figure \ref{fig:confusionMatrix_sr_ar}).
 
-```{r confusionMatrix_sr_ar, fig.height=4, fig.cap="Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the automatic classifier."}
+```{r confusionMatrix_sr_ar, fig.height=5, fig.cap="\\label{fig:confusionMatrix_sr_ar}Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the automatic classifier."}
 confusionMatrix_sr_ar_freq <- confusionMatrix_sr_ar$table %>% 
   as.data.frame() %>% 
-  dplyr::rename(ar_emotion = Var1, sr_emotion = Var2) %>% 
-  dplyr::mutate(prop_video = Freq/sum(Freq))
+  dplyr::rename(ar_emotion = Prediction, sr_emotion = Reference) %>% 
+  dplyr::mutate(n_video = sum(Freq)) %>% 
+  dplyr::mutate(prop_video = Freq/sum(Freq)) %>% 
+  dplyr::group_by(sr_emotion) %>% 
+  dplyr::mutate(n_sr = sum(Freq)) %>%
+  dplyr::mutate(prop_sr = Freq/sum(Freq)) %>% 
+  dplyr::ungroup()
 
@@
         legend.text = element_text(size=8),
-        legend.position = "none")
+        legend.position = "bottom")
 ```
 
-Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are similar to the ones with human observers (Table \ref{tab:confusionTable_sr_ar}).
-Overall there is a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_ar$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_ar$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_ar$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_video"])`) and *surprise* (`r scales::percent(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "surprise" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_video"])`).
-Surprisingly the automatic classifier incorrectly recognized as *disgust* an important proportion of videos in which *happiness* was the highest self-reported emotion (`r scales::percent(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_video"])`).
-In parallel, the automatic classifier recognized as *happiness* and *disgust* videos in which *surprise* was the highest self-reported emotion (respectively `r scales::percent(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_video"])` and `r scales::percent(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_video"])`).
+Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are somewhat similar to the ones with human observers (Table \ref{table:confusionTable_sr_ar}).
+Overall, a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy \nolinebreak = \nolinebreak `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_ar$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_ar$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_ar$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` of the video self-reported) is watched.
+Surprisingly the automatic classifier incorrectly recognized as *disgust* an important proportion of videos in which *anger*, *happiness* and *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "anger", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
+In parallel, the automatic classifier recognized as *happiness* videos in which *fear* and *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "fear", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
 
@@
     "latex",
-    caption = "Agreement accuracy metrics for each emotion.",
+    caption = "\\label{table:confusionTable_sr_ar}Agreement accuracy metrics for each emotion.",
     booktabs = T,
@@
    add_footnote("Note. \\textit{na.} values are produced when not enough data are available to compute accuracy indicators.", notation="none", escape = FALSE)
+
+# confusionMatrix_sr_ar_table %>%
+#   papaja::apa_table(
+#     caption = table_nums("confusionTable_sr_ar"),
+#     row.names = FALSE,
+#     format = "word",
+#     escape = FALSE
+#     )
 ```
 
-A comparable explanation can be provided as the level of undetermined emotion are very high for the self reports.
+A comparable explanation involing the amount of undertermined video based on self-reports can be provided as the level of undetermined emotions are very high for the self reports.
 
-Conclusion
+## Comparison between human and automatic recognition
+
+As previously mentioned, human observers appear to be more accurate than the automatic classifier to recognize individuals’ subjective feeling (human observer Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`; automatic classifier Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`).
+However, both make mistakes.
+A third confusion matrix is used to compare similarities (diagonal) and differences between human observers and automatic classifier in classifing the six emotion labels (Figure \nolinebreak \ref{fig:confusionMatrix_hr_ar}).
+
+```{r confusionMatrix_hr_ar, fig.height=5, fig.cap="\\label{fig:confusionMatrix_hr_ar}Comparison between the human observers and automatic classifier recognition for the classfication of emotion labels."}
+comparison_hr_ar <- dplyr::inner_join(human_recognition_score,automatic_recognition_score, by = "C_Video") %>% 
+  dplyr::ungroup() %>% 
+  dplyr::filter(ar_emotion != "undetermined") %>% 
+  dplyr::filter(hr_emotion != "undetermined") %>% 
+  dplyr::mutate_if(is.character,as.factor)
+
+confusionMatrix_hr_ar <- caret::confusionMatrix(
+  data = comparison_hr_ar$ar_emotion,
+  reference = comparison_hr_ar$hr_emotion
+)
+
+confusionMatrix_hr_ar_freq <- confusionMatrix_hr_ar$table %>% 
+  as.data.frame() %>% 
+  dplyr::rename(ar_emotion = Prediction, hr_emotion = Reference) %>% 
+  dplyr::mutate(n_video = sum(Freq)) %>% 
+  dplyr::mutate(prop_video = Freq/sum(Freq)) %>% 
+  dplyr::group_by(hr_emotion) %>% 
+  dplyr::mutate(n_hr = sum(Freq)) %>%
+  dplyr::mutate(prop_hr = Freq/sum(Freq)) %>% 
+  dplyr::ungroup()
+
+confusionMatrix_hr_ar_freq %>% 
+  ggplot(mapping = aes(x = hr_emotion, y = ar_emotion)) +
+  geom_tile(aes(fill = prop_hr), colour = "white") +
+  geom_text(
+    aes(label = paste0(scales::percent(prop_hr), "\n(", Freq,"/",n_hr, ")")),
+    color = "black",
+    size = 4,
+    family="serif",
+    parse = FALSE,
+    lineheight = 0.7)+
+  scale_fill_gradient(
+    name = "Proportion of Recognized\nEmotion",
+    low = "white", 
+    high = "red", 
+    labels = scales::percent,
+    limits = c(0,1)
+  ) +
+  scale_x_discrete(name = "Human Observers") +
+  scale_y_discrete(name = "Automatic Classifier") +
+  theme_minimal() +
+  theme(text = element_text(size=16,family="serif"),
+        axis.text.x = element_text(size=16, angle = 45, hjust = 0.75,vjust=0.9),
+        axis.text.y = element_text(size=16, hjust = 0.5),
+        axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
+        legend.text = element_text(size=8),
+        legend.position = "bottom")
+```
+
+The overall agreement between human observers and the automatic classifier is in fact very low (Kappa = `r round(confusionMatrix_hr_ar$overall["Kappa"],2)`).
+Except for *happiness* and *disgust* (respectively `r scales::percent(as.numeric(confusionMatrix_hr_ar_freq[confusionMatrix_hr_ar_freq$ar_emotion == "happiness" & confusionMatrix_hr_ar_freq$hr_emotion == "happiness", "prop_hr"]))` and `r scales::percent(as.numeric(confusionMatrix_hr_ar_freq[confusionMatrix_hr_ar_freq$ar_emotion == "disgust" & confusionMatrix_hr_ar_freq$hr_emotion == "disgust", "prop_hr"]))` of common labelling), there is no clear common pattern.
+Moreover, the automatic classifier has a tendency to label as *disgust* videos labeled as *sadness* by human observers and as *happiness* videos labaled as *fear* by human observers
+
+Discussion and Conclusion
 ============
-Despite being one on the most investigated question in affective science, the link between emotion felt and facial expression is a hot topic and no clear evidence have been found to definitely answer it.
-However, with the growing interest of industries and government to monitor individual's psychological states, evidences are showing that facial expressions are in reality not expressing emotions [@mckeown2013analogical].
-This research aimed to provide some empirical data to the question.
-The subjective feeling of participants was compared with human recognition on one side and automatic recognition on the other side.
-The results reveals a low accuracy for both humans and automatic classifier to accurately identify the inner emotional states of these individuals based on their facial expressions.
 
-Some limitations to this process should be stated over the use of self-reports to evaluate individual's subjective feelings.
-Accessing to the inner subjective feeling can be biased if not impossible.
-Moreover the laboratory setting can trigger ambiguous and "non-basic" emotions which were not analysed in this research.
-The procedure use for human annotation can also be incriminated.
-Instead of asking the human annotators to provide an unique label, a more subtle approach was chosen to mimic results provided by the automatic classifier.
-In this regard, the results of the human annotation could have been more ambiguous because it is not the natural way that people are inferring human emotions.
-Finally, the automatic classifier algorithm can also be problematic.
-Based on training datasets which are most of the time using prototypical facial expression of the "basic" emotions, the algorithm to classify facial expressions can be held in check by the spontaneous facial expressions analysed.
+Despite being one on the most investigated question in affective science, the coherence between emotion felt and facially displayed and facial expression recognized is a hot topic.
+To date no clear evidence has been found to definitely answer it.
+Yet, with the growing interest of industries and government to monitor individual’s psychological states, this issue is under high pressure.
+The present research aims to provide some empirical data to the question.
+The faces encoders displayed when confronted with an emotional eliciting task were submitted both to human and to automatic recognition.
+The criterion for recognition accuracy was the subjective feeling self-reported by the encoder once the elicitation task carried out.
+Results first reveal a low coherence between emotion felt and facial expression displayed.
+Secondly, results show a low accuracy for both humans and automatic classifier in identifying the inner emotional states of these encoders based on their facial expressions.
+Thirdly, human observers prove to be better at recognizing the emotion facially expressed than the automatic recognition tool is.
 
-Considering the above, the results provides an additional evidence that individuals' subjective feeling can not be inferred from facial expressions and in our case invalidate the hypothesis of hardwired emotions.
-This result suggests that automatic facial expression recognition tools should be focused on evaluating facial morphology features such as action units rather than inferring potential emotional or affective states.
+Such results support the hypothesis advanced by some authors of a low emotion–expression coherence [e.g., @kappas2003facial].
+In many instances, facial displays are not associated with a concordant emotional state, even any emotional state at all [@bonanno2004brief; @fernandez2013emotion].
+More and more evidences are showing that facial expressions are in reality not expressing emotions [@mckeown2013analogical].
+As well as other nonverbal behaviors, facial movements are not only assumed to be determined by emotion but also by various other causes, such as psychological states (e.g., \nolinebreak motivations or pain), to say nothing of social context and sociocultural norms [such as “display rules”, @ekman1987universals].
+This multiple determination excludes any possibility of drawing an inference from facial activity on the underlying psychological state (emotional or other).
+Beyond the present observations showing a weak coherence between subjective feelings and spontaneous facial expressions, this study provides some light on the controversy between the behaviorist and the constructivist approaches.
+The first of those two lines of thinking [@ekman1992argument] assumes that expressions of emotion are brief and coherent patterns of facial muscle movements that co-vary with discrete subjective experiences.
+Emotions and their related prototypical facial behavior are universal because they are considered as innate mechanisms allowing individuals to respond adaptively to evolutionary significant events (threats, opportunities…) encountered in the environment.
+Instead of viewing emotions as natural kinds [@barrett2017theory], the constructivist approach supposes that emotions are social constructions.
+The specific emotion categories identified on the face of another person are the result of categorization.
+In other words, the emotions that are recognized by the observer are constructed in her/his mind, primarily based on her/his conceptual knowledge of emotion.
+Therefore, facial movements do not express specific emotions.
+It is the observer that infers the emotional meaning of the facial expression.
+As a consequence, one can predict from the first line of thinking that individuals’ emotional subjective feeling should be correlated to the recognition of facial expressions from both human observers and automatic classifiers whereas if emotions are social constructs, as stated by the second line of thinking, human observer’s should be better at perceiving emotions in face than automatic classifiers.
+Present results plead in favor of the latter stance.
+They show that human observers are more accurate than the automatic recognition tool to identify individuals’ subjective feeling on the basis of their face.
+Moreover, mistakes made by human observers look less arbitrary to the ones made by the classifier.
+For instance, even if mix-up between disgust and anger is sometimes reported in recognition studies, confusions such as the ones produced by the classifier have never been noted for human observers.
+A possible explanation is that human observers are assessing dynamic expressive sequences.
+They have access to the full video which provides a kinetic context whereas the automatic classifier only assesses the videos frame by frame.
+Further research is needed to shed light on this issue.
+
+Some limitations should be stated, notably regarding the use of self-reports to evaluate encoders’ subjective feelings.
+Accessing the inner subjective feeling can be biased if not impossible.
+Moreover the procedure used for human observation can also be open to dispute.
+Instead of asking the human annotators to provide a unique label, a more subtle approach was chosen to mimic results provided by the automatic classifier .
+Whereas this paradigm is longer and more complicated, it can lead to more robust results in reducing the forced-choice biais [@russell1993forced].
+However this procedure can also reduce the human observers accuracy.
+In this regard, the results of the human observation could have been more ambiguous because it is not the natural way that people are inferring meaning from facial expressions.
+An alternative explanation relies in reducing the recognition biais involved in the classic recognition paradigm.
+Classic forced-choice paradigms obtain artificial high results, thus by using a more evolved approach observers’ accuracy may have been lowered.
+
+Considering the above, the results provide additional evidence that individuals’ subjective feeling can not be inferred from facial expressions and in our case invalidate the hypothesis of hardwired emotions unambiguously displayed on the face.
+Even if emotions were hardwired, in everyday life one does not observe prototypical facial expressions and therefore research should be focused on analysing non-prototypical facial expressions.
+Advancements in identifying “non-basic” emotion labels [@mcduff2016discovering] as well as non-prototypical facial expression have been made in the developpement of automatic facial expression recognition tools.
+However this result suggests that automatic facial expression recognition tools should merely evaluate facial morphology features such as action units [e.g., OpenFace presented in @baltruvsaitis2016openface] rather than inferring supposedly emotional or affective states.
 

### 2019-04-05 ACII_2019_paper/ACII_2019_paper.tex
latex

--- 2019-03-16 ACII_2019_paper/ACII_2019_paper.tex
@@
 \hypersetup{
-            pdftitle={Are facial expressions the genuine display of individuals' subjective feeling? A comparison with human and automatic recognition.},
+            pdftitle={Coherence between subjective feeling and facial expression: A comparison of human and automatic recognition},
             pdfborder={0 0 0},
@@
 % Do not put math or special symbols in the title.
-\title{Are facial expressions the genuine display of individuals' subjective
-feeling? A comparison with human and automatic recognition.}
+\title{Coherence between subjective feeling and facial expression: A comparison
+of human and automatic recognition}
 
@@
 facial expressions. On the other hand the constructivist approach
-conceives emotions are socially constructed and facial expression as
-social messages that are not related to emotions. In order to evaluate
+conceives emotions as socially constructed, the emotional meaning of a
+facial expression being inferred by the observer. In order to evaluate
 the link between the subjective feeling of emotions and their
-recognition based on facial expression, 232 videos of participants
-recruited to perform an emotion elicitation task were annotated by 1383
-human observers as well as by an automatic facial expression classifier.
-The results show a low accuracy of human observers and of the automatic
-classifier to infer from the facial expression the subjective feeling of
-the participants recorded. Based on these results, the hypothesis of
-genetically hardwired emotion genuinely display is difficult to support
-whereas the idea of emotion socially constructed and facial expression
-as display of social messages appears to be more likely. Then, the way
-to infer emotional and mental state based on facial expressions should
-be questioned.
+recognition based on facial expression, 232 videos of encoders recruited
+to carry out an emotion elicitation task were annotated by 1383 human
+observers as well as by an automatic facial expression classifier.
+Results show a low accuracy of human observers and of the automatic
+classifier to infer the subjective feeling from the facial expressions
+displayed by encoders . They also show a weak link between self-reported
+emotional states and facial emotional displays. Based on these results,
+the hypothesis of genetically hardwired emotion genuinely display is
+difficult to support whereas the idea of emotion and facial expression
+socially constructed appears to be more likely. Then automatic emotion
+recognition tools based on facial expressions should be questioned.
 \end{abstract}
@@
 this technology in order to track humans' emotions in various scenarios
-(e.g., marketing, healthcare, automotive to name a few). Beside the
-ethical question of measuring human emotions, these tools are
-challenging psychological theories about inferring emotions from facial
-expressions. When automatic facial expression recognition tools are used
-to measure human emotions, they rest on the premise that facial
-expressions provide a direct access to expressers' emotions. However the
-link between emotion felt and facial expressions is far from being
-established and still remain a hot topic in psychology research.
+(e.g., marketing, healthcare, automotive to name a few). This
+technoloogy rests on the premise that facial expressions provide a
+direct access to individuals' subjective feeling. Even if this premise
+is central to the modern mainstream approach of human emotion, recent
+research in affective science are challenging it. Once the two competing
+approaches briefly described as well as the predictions they
+respectively entail, an experiment testing these hypotheses will be
+presented and its results analyzed in order to provide empirical
+evidence to contribute to answer the question.
 
-\hypertarget{the-link-between-emotion-and-facial-expression-thought-the-behaviorist-approach}{%
-\subsection{The link between emotion and facial expression thought the
+\hypertarget{the-link-between-emotion-and-facial-expression-through-the-behaviorist-approach}{%
+\subsection{The link between emotion and facial expression through the
 behaviorist
-approach}\label{the-link-between-emotion-and-facial-expression-thought-the-behaviorist-approach}}
+approach}\label{the-link-between-emotion-and-facial-expression-through-the-behaviorist-approach}}
 
@@
 Expression of the Emotions in Man and Animals} (Darwin 1872), facial
-expression are conceived as a genuine display of individuals inner
+expressions are conceived as a genuine displays of individuals' inner
 emotional state. This hypothesis is used as a basis for the Basic
-Emotion Theory (BET) which states that a set of six or seven emotions
-are universally displayed and are genetically hardwired not only in
-humans (Ekman 1992) but also in different animal species (Waal 2019).
-According to this view, ``when emotions are aroused by perception of a
-social event, a set of central commands produce patterned
-emotion-specific changes in multiple systems, including {[}\ldots{}{]}
-facial expressions.'' (Ekman 2007, p49). Even if this theory obtained a
-popular support, it fails to explain how individuals can feel emotions
-without expressing them and how individuals can express emotions without
-feeling them.
+Emotion Theory (BET) which states that a set of six emotions are
+universally displayed and are genetically hardwired not only in humans
+(Ekman 1992) but also in different animal species (Waal 2019). According
+to this view, ``when emotions are aroused by perception of a social
+event, a set of central commands produce patterned emotion-specific
+changes in multiple systems, including {[}\ldots{}{]} facial
+expressions.'' (Ekman 2007, p49). To cope with critics, several
+amendments have been made to the BET, increasing the number of basic
+emotions from six to seven (Ekman and Heider 1988) as well as adding the
+concept of ``display rules'' to explain cultural differences in the
+management of facial expressions (Ekman et al. 1987).
 
-\hypertarget{the-link-between-emotion-and-facial-expression-thought-the-social-constructivist-approach}{%
-\subsection{The link between emotion and facial expression thought the
+Even if this theory obtained a popular support, it fails to explain how
+individuals can feel emotions without expressing them and how
+individuals can express emotions without feeling them, in instances in
+which display rules cannot be called upon (Kraut and Johnston 1979;
+Durán, Reisenzein, and Fernández-Dols 2017). These evidences have led to
+an alternative conception, notably the social constructivist approach.
+
+\hypertarget{the-link-between-emotion-and-facial-expression-through-the-social-constructivist-approach}{%
+\subsection{The link between emotion and facial expression through the
 social constructivist
-approach}\label{the-link-between-emotion-and-facial-expression-thought-the-social-constructivist-approach}}
+approach}\label{the-link-between-emotion-and-facial-expression-through-the-social-constructivist-approach}}
 
-Detractors of the Basic Emotion Theory are perceiving emotion not as
+Detractors of the Basic Emotion Theory consider emotion not as
 genetically hardwired but as a learnt association between a given
-situation and an appropriate response (Averill 1980,
-@barrett2017emotions). For the tenants of the constructivist approach,
-emotions are ``concepts'' based on past experiences and which are ``a
-collection of embodied, whole brain representations that predict what is
-about to happen in the sensory environment, what the best action is to
-deal with impending events, and their consequences for allostasis''
-(Barrett 2017b, p12). Following this assumption, faces are used as tools
-to display signals in social interactions (Crivelli and Fridlund 2018).
-These signals can convey individuals' motivations and readiness (Frijda
-and Tcherkassof 1997) or social messages (Fridlund and Rosenberg 1995).
+situation and an appropriate response (Averill 1980; Barrett 2017a). For
+the tenants of the constructivist approach, emotions are ``concepts''
+based on past experiences and which are ``a collection of embodied,
+whole brain representations that predict what is about to happen in the
+sensory environment, what the best action is to deal with impending
+events, and their consequences for allostasis'' (Barrett 2017b, p12).
+Following this assumption, faces are best conceived as tools displaying
+signals in social interactions (Crivelli and Fridlund 2018). These
+signals can convey individuals' motivations and readiness (Frijda and
+Tcherkassof 1997) or social messages (Fridlund and Rosenberg 1995).
+Therefore, facial expressions are thought as behaviors which meaning is
+inferred by the observer. Findings support this observer dependence
+(Lindquist and Gendron 2013; Niedenthal et al. 2017). They show that to
+make meaning of another person's facial behavior, the perceiver relies
+in particular on her/his knowledge about emotion categories.
 
-This current paper investigates the link between the subjective feeling
-of emotions and their recognition from facial expressions. If emotions
-are hardwired, individuals emotional subjective feeling should be
-correlated to the recognition of facial expressions from both human
-observers and automatic classifiers whereas if emotions are social
-constructs, no correlation between subjective feeling and facial
-expression recognition should be observed.
+\hypertarget{the-recognition-of-emotional-facial-expressions}{%
+\subsection{The recognition of emotional facial
+expressions}\label{the-recognition-of-emotional-facial-expressions}}
+
+Regarding emotional facial expression (EFE) recognition, the behaviorist
+approach assertions and the constructivist approach claims lead to two
+divergent predictions. The foremost postulates that, when triggered,
+each basic emotion is expressed by a prototypical face (non basic
+emotions being blends of the basic ones). Basic emotions are easily
+recognized by all human observers and emotional states are accessible by
+facial measurement. As the recognition of facial expressions is based on
+the identification of specific patterns of facial movements, it implies
+that EFE recognition of both human observers and automatic classifiers
+should be as accurate. The constructivist approach, in contrast, affirms
+that facial expressions do not provide a direct access to individuals'
+subjective feeling. Even if the face possibly moves during an emotional
+episode, facial muscle movements are not linked in a one-to-one manner
+to a specific discrete emotional experience (Doyle and Lindquist 2017,
+p418). Instead, emotions are mentally constructed by the perceiver and
+mental categories of emotions are needed to accurately categorize facial
+movements. Therefore, since there is no emotional prototypical face, one
+should expect human observers' superior ability to accurately recognize
+EFE as compared to automatic EFE recognition tools.
+
+The aim of the current paper is to investigate the link between the
+subjective feeling of emotions and its recognition from facial
+expressions in order to give credit either to the behaviorist plea or to
+the constructivist's one. Contrary to most studies using posed and
+static EFE, the present study focuses on natural EFE. Spontaneous and
+dynamic facial reactions to emotional elicitations are under
+consideration to ensure the generalizability of the results to emotional
+behaviors in ordinary life.
 
@@
 To evaluate the link between subjective feeling of emotions and their
-recognition from facial expressions, participants were recruited to
+recognition from facial expressions, encoders were first recruited to
 perform an emotion elicitation task while their facial expression was
-video recorded. Then, the videos was shown to human observer and
-analysed by an automatic classifier in order to identify which emotion
-was displayed.
+video recorded. Then, the videos of the encoders' faces were shown to
+human observers and were also analysed by an automatic classifier in
+order to identify which emotion was displayed.
 
@@
 females, 176 males, \emph{M}age = 47.9, \emph{SD}age = 9.2) were
-recruited to perform one out of 11 emotions elicitation tasks designed
-to trigger a positive, negative or neutral emotional state (see
-Tcherkassof et al. 2013 for a description of tasks and procedure).
-Participants' face were recorded using an hidden camera resulting 358
-front facing 768x576 videos varying from 1s to 1479s.
+recruited to perform one out of 11 emotion elicitation tasks designed to
+trigger a positive, a specific negative or a neutral emotional state.
+Encoders' face were recorded using an hidden camera resulting 358 front
+facing 768x576 videos varying from 1s to 1479s (Figure
+\ref{fig:dynemo_img}). These recordings form the DynEmo database (see
+Tcherkassof et al. 2013 for a full description of tasks and procedure).
 
-After their emotion elicitation task the participants had to rate their
-emotional state during the task on a likert scale from 0 (``not at
-all'') to 5 (``strongly'') the six ``basic'' emotions (i.e.,
-\emph{anger}, \emph{disgust}, \emph{fear}, \emph{happiness},
-\emph{surprise} and \emph{sadness}) as well as six ``non-basic''
-emotions (i.e., \emph{pride}, \emph{curiosity}, \emph{boredom},
+\begin{figure}
+\centering
+\includegraphics{ACII_2019_paper_files/figure-latex/dynemo_img-1.pdf}
+\caption{\label{fig:dynemo_img}Example of a front facing recording
+synced with the full view of the participant and the elicitation task.
+This picture is taken from a pilot with projects collaborators and all
+gave a consent for the publication of their photos and videos.}
+\end{figure}
+
+After the emotion elicitation task the encoders rated their subjective
+feeling on Likert scales from 0 (``not at all'') to \nolinebreak 5
+\nolinebreak (``strongly'') related to six ``basic'' emotion labels
+(i.e., \emph{anger}, \emph{disgust}, \emph{fear}, \emph{happiness},
+\emph{surprise} and \emph{sadness}) as well as six ``non-basic'' emotion
+labels (i.e., \emph{pride}, \emph{curiosity}, \emph{boredom},
 \emph{shame}, \emph{humiliation}, and \emph{disappointment}).
 
-Finally, a debriefing session was perform to ensure that participants
-were not durably affected by the emotion elicitation task. All the
-participants gave their agreement on their data and video to be
-processed for research purpose only.
+Finally, a debriefing session was performed to ensure that encoders were
+not durably affected by the emotion elicitation task. The debriefing was
+also used to check that encoders did not guess the real purpose of the
+experiment (e.g.~being filmed while they were performing an emotional
+elicitation task) to guarantee facial expressions' genuineness. All
+encoders gave their agreement on their data and video to be processed
+for research purpose only.
 
@@
 
-For the human facial expression recognition method, 1383 student
-participants were recruited to annotate 232 out of the 358 video,
-therefore only the 232 annotated videos will be analysed in this paper.
-Each participants had to annotate between 1 and 177 videos resulting
-that each video was annotated 29 times on average (\emph{SD} = 12).
+For the human facial expression recognition method, 1383
+\nolinebreak student participants were recruited to annotate 232 out of
+the 358 videos, therefore only the 232 annotated videos will be analysed
+in this paper. Because videos have different durations, participants had
+to annotate a series of video corresponding to 30min long in total. Each
+video was annotated 29 times on average (\emph{SD} = 12).
 
@@
 \emph{Oudjat}, a software for designing video annotation experiments
-(Dupré et al. 2015). For each video, the annotation procedure hat two
-steps. First, the participants had to identify the emotional sequences
-by pressing the space bar of their keyboard to indicate the beginning
-and the end of the emotional sequences while watching the video. Second,
-the participants watched each emotional sequence previously identified
-and had to label the sequence using one of the 12 emotions proposed
-including six ``basic'' emotions (i.e., \emph{anger}, \emph{disgust},
-\emph{fear}, \emph{happiness}, \emph{surprise} and \emph{sadness}) and
-six ``non-basic'' emotions (i.e., \emph{pride}, \emph{curiosity},
-\emph{boredom}, \emph{shame}, \emph{humiliation}, and
-\emph{disappointment}). They also had the possibility to indicate that
-the sequence was not expressing one of the proposed emotion.
+(Dupré et al. 2015). For each video, the annotation procedure followed
+two steps. First, the participants had to identify the emotional
+sequences by pressing the space bar of their keyboard to indicate the
+beginning and the end of the emotional sequences while watching the
+video. Second, the participants watched each emotional sequence
+previously identified and labeled the sequence using one of the 12
+\nolinebreak emotions proposed including six ``basic'' emotion labels
+(i.e., \emph{anger}, \emph{disgust}, \emph{fear}, \emph{happiness},
+\emph{surprise} and \emph{sadness}) and six ``non-basic'' emotion labels
+(i.e., \emph{pride}, \emph{curiosity}, \emph{boredom}, \emph{shame},
+\emph{humiliation}, and \emph{disappointment}). They also had the
+possibility to indicate that the sequence was expressing none of the
+proposed emotion labels.
 
@@
 each video per human observer identifying for each second of the video
-which emotion was recognized. Then time-series corresponding to the same
-video were aggregated to calculate the proportion of human observers for
-each second of the video per emotional label. The sum of each label
-proportion per second was used as a score to determine which labels
-corresponds to the overall video (i.e., the highest score). In case of
-more than one label having the maximum value, the emotion is described
-as undetermined.
+which emotion was recognized. Then, time-series corresponding to the
+same video were aggregated to calculate the proportion of human
+observers \(x_{video_{i}.label_{j}.t_{k}}\) for each second of the video
+per emotional label (EQ\ref{eq:1}).
+
+\begin{equation}
+\label{eq:1}
+x_{video_{i}.label_{j}.t_{k}} = \frac{n_{video_{i}.label_{j}.t_{k}}}{n_{video_{k}}}
+\end{equation}
+
+where \emph{i} is one of the 232 videos, \emph{j} is one of the six
+``basic'' emotion labels, \emph{k} for each second of the video.
 
@@
 features and Support Vector Machine (SVM) classifiers in order to
-recognize facial expressions. For each video frame, Affdex identify the
-probability of the face as expressing one of the six ``basic'' emotions
-(i.e., \emph{anger}, \emph{disgust}, \emph{fear}, \emph{happiness},
-\emph{surprise} and \emph{sadness}) as well as additional psychological
-states such as \emph{valence}, \emph{engagement} or \emph{contempt}, and
-facial features such as \emph{cheek raise}, \emph{eye widen} or
-\emph{jaw drop}.
+recognize facial expressions. For each video frame, Affdex identifies
+the probability \(p_{video_{i}.label_{j}.t_{k}}\) from 0 to 100
+(rescaled to 0 to 1 for the analysis) of the face as expressing one of
+the six ``basic'' emotion labels (i.e., \emph{anger}, \emph{disgust},
+\emph{fear}, \emph{happiness}, \emph{surprise} and \emph{sadness}) as
+well as additional psychological states such as \emph{valence},
+\emph{engagement} or \emph{contempt}, and facial features such as
+\emph{cheek raise}, \emph{eye widen} or \emph{jaw drop}.
 
-To determine which of the six ``basic'' emotion can be used to identify
-each video, the recognition probability for each label by frame was
-converted into odd ratio by frame (Dente et al. 2017). The highest sum
-of each odd ratio time-series defines the label recognized by the
-automatic classifier.
+For both human and automatic recognition, to determine which of the six
+``basic'' emotions can be used to identify each video, the recognition
+probability for each label by frame was converted into odd ratio by
+label (Dente et al. 2017). The highest sum of each odd ratio time-series
+defines the label recognized by the automatic classifier (EQ\ref{eq:2}
+for human recognition and EQ\ref{eq:3} for automatic recognition).
+\begin{equation}
+\label{eq:2}
+video_{i}.label = \max\left(\frac{\sum_{k=1}^{n}x_{video_{i}.label_{j}.t_{k}}}{\sum_{k=1}^{n}x_{video_{i}.t_{k}}}\right)
+\end{equation}
+
+\begin{equation}
+\label{eq:3}
+video_{i}.label = \max\left(\frac{\sum_{k=1}^{f}p_{video_{i}.label_{j}.t_{k}}}{\sum_{k=1}^{f}p_{video_{i}.t_{k}}}\right)
+\end{equation}
+
+where \emph{i} is one of the 232 videos, \emph{j} is one of the six
+``basic'' emotion labels, \emph{k} for each second of the \emph{n}
+second video or for each frame of the \emph{f} frame video.
 
@@
 
-Whereas the self-reports, the human annotations and the automatic
-recognition include data on ``non-basic'' emotions and features, the
-analysis is performed using only the six ``basic'' emotions in order to
-compare them. The maximum score for self-reports, human annotations and
-automatic recognition is used to label the video. In case of more than
-one label obtaining the maximum value, the video is labeled as
-undetermined.
+Since encoders' self-reports, human annotations and the automatic
+recognition include data on ``non-basic'' emotion labels and features,
+the analysis is performed using only the six ``basic'' emotion labels in
+order to compare them. The maximum score for self-reports, human
+annotations and automatic recognition is used to label the video. In
+case of more than one label obtaining the maximum value, the video is
+labeled as undetermined.
 
@@
 
-Emotions self-reported as being characteristic of the elicitation are
-compared with the emotion recognized by the human observers in a
-confusion matrix (Figure \ref{fig:confusionMatrix_sr_hr}).
+Encoders' subjective feeling is compared with human observers
+recognition in a confusion matrix (Figure
+\ref{fig:confusionMatrix_sr_hr}). Each emotion label used to describe
+encoders' self-reported subjective feeling (i.e., the label rated with
+the highest value) is compared with the emotion labels which were rated
+with the highest score by human observers.
 
@@
 \includegraphics{ACII_2019_paper_files/figure-latex/confusionMatrix_sr_hr-1.pdf}
-\caption{Confusion matrix of between the emotion self-reported as being
-characteristic of the elicitation with the emotion recognized by the
-human observers.}
+\caption{\label{fig:confusionMatrix_sr_hr}Confusion matrix of between
+the emotion self-reported as being characteristic of the elicitation
+with the emotion recognized by the human observers.}
 \end{figure}
 
-The result of the confusion matrix show a low agreement between emotion
-felt during the elicitation and emotion recognized by the human
-annotators (Accuracy = 0.27, 95\%CI{[}0.21,0.33{]}; Kappa = 0.11) except
-for \emph{happiness} (15.2\%), \emph{surprise} (8.26\%) and
-\emph{disgust} (1.74\%). Sensitivity, specificity, precision and F1
-score for each emotion can be found Table
-\ref{tab:confusionTable_sr_hr}. Interestingly human annotators seem to
-recognize as \emph{surprise} videos in which \emph{happiness} was the
-highest self-reported emotion (10.9\%), and in a lower instance
-\emph{happiness} videos in which \emph{surprise} was the highest
-self-reported emotion (4.78\%).
+Results of the confusion matrix show a moderate agreement between the
+emotion felt by the encoder during the elicitation and the emotion
+recognized by the human observers (Accuracy \nolinebreak =
+\nolinebreak 0.43, 95\%CI{[}0.35,0.52{]}; Kappa = 0.19) except for
+\emph{disgust} (100\% of the videos self-reported). Sensitivity,
+specificity, precision and F1 score for each emotion reveals that
+\emph{happiness} has the highest coherence ratio whereas \emph{sadness}
+has the lowest coherence ratio between true positives and false
+positives (Table \ref{table:confusionTable_sr_hr}). Interestingly human
+observers seem to recognize \emph{surprise} expressed in videos where
+\emph{anger}, \emph{fear} \emph{happiness} and \emph{sadness} was the
+highest self-reported emotion (respectively 28.6\%, 33.3\%, 34.7\% and
+33.3\% of the videos self-reported), and in a lower instance
+\emph{happiness} was recognized in videos where \emph{fear} and
+\emph{surprise} was the highest self-reported emotion (respectively
+50.0\% and 25.6\% of the videos self-reported).
 
@@
 
-\caption{\label{tab:confusionTable_sr_hr}Agreement accuracy metrics for each emotion. }
+\caption{\label{tab:confusionTable_sr_hr}\label{table:confusionTable_sr_hr}Agreement accuracy metrics for each emotion.}
 \centering
@@
 \midrule
-anger & 0.43 & 0.92 & 0.14 & 0.21\\
-disgust & 1.00 & 0.93 & 0.20 & 0.33\\
-fear & 0.00 & 0.96 & 0.00 & \textit{na.}\\
-happiness & 0.49 & 0.73 & 0.45 & 0.47\\
-sadness & 0.08 & 0.91 & 0.05 & 0.06\\
-surprise & 0.44 & 0.67 & 0.23 & 0.31\\
-undetermined & 0.00 & 0.99 & 0.00 & \textit{na.}\\
+anger & 0.43 & 0.93 & 0.23 & 0.3\\
+disgust & 1.00 & 0.94 & 0.33 & 0.5\\
+fear & 0.00 & 0.97 & 0.00 & \textit{na.}\\
+happiness & 0.49 & 0.78 & 0.69 & 0.57\\
+sadness & 0.08 & 0.92 & 0.08 & 0.08\\
+surprise & 0.44 & 0.67 & 0.37 & 0.4\\
 \bottomrule
@@
 
-However, the self-report show a very high proportion of undetermined
-emotional states which reveals not only the possibility of the emotion
-elicitation tasks to trigger more than one emotion but also the
-potential limit of using 6-points likert scales for which the
-participants can easily score to the maximum for more than one emotion.
+However, self-reports show a significant proportion of undetermined
+emotional states (35.2\% of the 358 videos) which reveals the potential
+limit of using 6-points likert scales to measure emotional self-reports.
+Indeed, encoders can easily score to the maximum for more than one
+emotion.
 
@@
 
-As in the previous analysis, emotions self-reported as being
-characteristic of the elicitation are compared with the emotion
-recognized by the automatic classifier in a confusion matrix (Figure
-\ref{fig:confusionMatrix_sr_ar}).
+Similarly to the previous analysis, a confusion matrix was used to
+compare encoders' subjective feeling with the emotion label recognized
+by the automatic classifier (Figure \ref{fig:confusionMatrix_sr_ar}).
 
@@
 \includegraphics{ACII_2019_paper_files/figure-latex/confusionMatrix_sr_ar-1.pdf}
-\caption{Confusion matrix of between the emotion self-reported as being
-characteristic of the elicitation with the emotion recognized by the
-automatic classifier.}
+\caption{\label{fig:confusionMatrix_sr_ar}Confusion matrix of between
+the emotion self-reported as being characteristic of the elicitation
+with the emotion recognized by the automatic classifier.}
 \end{figure}
@@
 Results obtained for the comparison between emotions self-reported and
-recognized by the automatic classifier are similar to the ones with
-human observers (Table \ref{tab:confusionTable_sr_ar}). Overall there is
+recognized by the automatic classifier are somewhat similar to the ones
+with human observers (Table \ref{table:confusionTable_sr_ar}). Overall,
 a low agreement between emotion self-reported and emotion recognized by
-the automatic classifier (Accuracy = 0.19, 95\%CI{[}0.14,0.24{]}; Kappa
-= 0.05) except for \emph{happiness} (13.9\%) and \emph{surprise}
-(3.48\%). Surprisingly the automatic classifier incorrectly recognized
-as \emph{disgust} an important proportion of videos in which
-\emph{happiness} was the highest self-reported emotion (7.83\%). In
-parallel, the automatic classifier recognized as \emph{happiness} and
-\emph{disgust} videos in which \emph{surprise} was the highest
-self-reported emotion (respectively 5.22\% and 6.09\%).
+the automatic classifier (Accuracy \nolinebreak = \nolinebreak 0.3,
+95\%CI{[}0.22,0.38{]}; Kappa = 0.07) except for \emph{happiness} (44.4\%
+of the video self-reported) is watched. Surprisingly the automatic
+classifier incorrectly recognized as \emph{disgust} an important
+proportion of videos in which \emph{anger}, \emph{happiness} and
+\emph{surprise} was the highest self-reported emotion (respectively
+42.9\%, 25.0\% and 32.6\% of the videos self-reported). In parallel, the
+automatic classifier recognized as \emph{happiness} videos in which
+\emph{fear} and \emph{surprise} was the highest self-reported emotion
+(respectively 66.7\% and 32.6\% of the videos self-reported).
 
@@
 
-\caption{\label{tab:confusionTable_sr_ar}Agreement accuracy metrics for each emotion.}
+\caption{\label{tab:confusionTable_sr_ar}\label{table:confusionTable_sr_ar}Agreement accuracy metrics for each emotion.}
 \centering
@@
 \midrule
-anger & 0.00 & 0.89 & 0 & \textit{na.}\\
-disgust & 0.25 & 0.70 & 0.01 & 0.03\\
-fear & 0.00 & 0.93 & 0 & \textit{na.}\\
-happiness & 0.44 & 0.71 & 0.41 & 0.43\\
-sadness & 0.15 & 0.94 & 0.14 & 0.15\\
-surprise & 0.19 & 0.89 & 0.28 & 0.22\\
-undetermined & 0.00 & 1.00 & \textit{na.} & \textit{na.}\\
+anger & 0.00 & 0.90 & 0.00 & \textit{na.}\\
+disgust & 0.25 & 0.72 & 0.02 & 0.04\\
+fear & 0.00 & 0.94 & 0.00 & \textit{na.}\\
+happiness & 0.44 & 0.73 & 0.62 & 0.52\\
+sadness & 0.15 & 0.95 & 0.25 & 0.19\\
+surprise & 0.19 & 0.86 & 0.36 & 0.25\\
 \bottomrule
-\multicolumn{5}{l}{\textsuperscript{} Note. \textit{na.} values are produced when not enough data are}\\
-\multicolumn{5}{l}{available to compute accuracy indicators.}\\
+\multicolumn{5}{l}{\textsuperscript{} Note. \textit{na.} values are produced when not enough}\\
+\multicolumn{5}{l}{data are available to compute accuracy indicators.}\\
 \end{tabu}
@@
 
-A comparable explanation can be provided as the level of undetermined
-emotion are very high for the self reports.
+A comparable explanation involing the amount of undertermined video
+based on self-reports can be provided as the level of undetermined
+emotions are very high for the self reports.
 
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
+\hypertarget{comparison-between-human-and-automatic-recognition}{%
+\subsection{Comparison between human and automatic
+recognition}\label{comparison-between-human-and-automatic-recognition}}
+
+As previously mentioned, human observers appear to be more accurate than
+the automatic classifier to recognize individuals' subjective feeling
+(human observer Accuracy = 0.43; automatic classifier Accuracy = 0.3).
+However, both make mistakes. A third confusion matrix is used to compare
+similarities (diagonal) and differences between human observers and
+automatic classifier in classifing the six emotion labels (Figure
+\nolinebreak \ref{fig:confusionMatrix_hr_ar}).
+
+\begin{figure}
+\centering
+\includegraphics{ACII_2019_paper_files/figure-latex/confusionMatrix_hr_ar-1.pdf}
+\caption{\label{fig:confusionMatrix_hr_ar}Comparison between the human
+observers and automatic classifier recognition for the classfication of
+emotion labels.}
+\end{figure}
+
+The overall agreement between human observers and the automatic
+classifier is in fact very low (Kappa = 0.18). Except for
+\emph{happiness} and \emph{disgust} (respectively 64.9\% and 50.0\% of
+common labelling), there is no clear common pattern. Moreover, the
+automatic classifier has a tendency to label as \emph{disgust} videos
+labeled as \emph{sadness} by human observers and as \emph{happiness}
+videos labaled as \emph{fear} by human observers
+
+\hypertarget{discussion-and-conclusion}{%
+\section{Discussion and Conclusion}\label{discussion-and-conclusion}}
 
 Despite being one on the most investigated question in affective
-science, the link between emotion felt and facial expression is a hot
-topic and no clear evidence have been found to definitely answer it.
-However, with the growing interest of industries and government to
-monitor individual's psychological states, evidences are showing that
-facial expressions are in reality not expressing emotions (McKeown
-2013). This research aimed to provide some empirical data to the
-question. The subjective feeling of participants was compared with human
-recognition on one side and automatic recognition on the other side. The
-results reveals a low accuracy for both humans and automatic classifier
-to accurately identify the inner emotional states of these individuals
-based on their facial expressions.
+science, the coherence between emotion felt and facially displayed and
+facial expression recognized is a hot topic. To date no clear evidence
+has been found to definitely answer it. Yet, with the growing interest
+of industries and government to monitor individual's psychological
+states, this issue is under high pressure. The present research aims to
+provide some empirical data to the question. The faces encoders
+displayed when confronted with an emotional eliciting task were
+submitted both to human and to automatic recognition. The criterion for
+recognition accuracy was the subjective feeling self-reported by the
+encoder once the elicitation task carried out. Results first reveal a
+low coherence between emotion felt and facial expression displayed.
+Secondly, results show a low accuracy for both humans and automatic
+classifier in identifying the inner emotional states of these encoders
+based on their facial expressions. Thirdly, human observers prove to be
+better at recognizing the emotion facially expressed than the automatic
+recognition tool is.
 
-Some limitations to this process should be stated over the use of
-self-reports to evaluate individual's subjective feelings. Accessing to
-the inner subjective feeling can be biased if not impossible. Moreover
-the laboratory setting can trigger ambiguous and ``non-basic'' emotions
-which were not analysed in this research. The procedure use for human
-annotation can also be incriminated. Instead of asking the human
-annotators to provide an unique label, a more subtle approach was chosen
-to mimic results provided by the automatic classifier. In this regard,
-the results of the human annotation could have been more ambiguous
-because it is not the natural way that people are inferring human
-emotions. Finally, the automatic classifier algorithm can also be
-problematic. Based on training datasets which are most of the time using
-prototypical facial expression of the ``basic'' emotions, the algorithm
-to classify facial expressions can be held in check by the spontaneous
-facial expressions analysed.
+Such results support the hypothesis advanced by some authors of a low
+emotion--expression coherence (e.g., Kappas 2003). In many instances,
+facial displays are not associated with a concordant emotional state,
+even any emotional state at all (Bonanno and Keltner 2004;
+Fernández-Dols and Crivelli 2013). More and more evidences are showing
+that facial expressions are in reality not expressing emotions (McKeown
+2013). As well as other nonverbal behaviors, facial movements are not
+only assumed to be determined by emotion but also by various other
+causes, such as psychological states (e.g., \nolinebreak motivations or
+pain), to say nothing of social context and sociocultural norms (such as
+``display rules'', Ekman et al. 1987). This multiple determination
+excludes any possibility of drawing an inference from facial activity on
+the underlying psychological state (emotional or other). Beyond the
+present observations showing a weak coherence between subjective
+feelings and spontaneous facial expressions, this study provides some
+light on the controversy between the behaviorist and the constructivist
+approaches. The first of those two lines of thinking (Ekman 1992)
+assumes that expressions of emotion are brief and coherent patterns of
+facial muscle movements that co-vary with discrete subjective
+experiences. Emotions and their related prototypical facial behavior are
+universal because they are considered as innate mechanisms allowing
+individuals to respond adaptively to evolutionary significant events
+(threats, opportunities\ldots{}) encountered in the environment. Instead
+of viewing emotions as natural kinds (Barrett 2017b), the constructivist
+approach supposes that emotions are social constructions. The specific
+emotion categories identified on the face of another person are the
+result of categorization. In other words, the emotions that are
+recognized by the observer are constructed in her/his mind, primarily
+based on her/his conceptual knowledge of emotion. Therefore, facial
+movements do not express specific emotions. It is the observer that
+infers the emotional meaning of the facial expression. As a consequence,
+one can predict from the first line of thinking that individuals'
+emotional subjective feeling should be correlated to the recognition of
+facial expressions from both human observers and automatic classifiers
+whereas if emotions are social constructs, as stated by the second line
+of thinking, human observer's should be better at perceiving emotions in
+face than automatic classifiers. Present results plead in favor of the
+latter stance. They show that human observers are more accurate than the
+automatic recognition tool to identify individuals' subjective feeling
+on the basis of their face. Moreover, mistakes made by human observers
+look less arbitrary to the ones made by the classifier. For instance,
+even if mix-up between disgust and anger is sometimes reported in
+recognition studies, confusions such as the ones produced by the
+classifier have never been noted for human observers. A possible
+explanation is that human observers are assessing dynamic expressive
+sequences. They have access to the full video which provides a kinetic
+context whereas the automatic classifier only assesses the videos frame
+by frame. Further research is needed to shed light on this issue.
 
-Considering the above, the results provides an additional evidence that
+Some limitations should be stated, notably regarding the use of
+self-reports to evaluate encoders' subjective feelings. Accessing the
+inner subjective feeling can be biased if not impossible. Moreover the
+procedure used for human observation can also be open to dispute.
+Instead of asking the human annotators to provide a unique label, a more
+subtle approach was chosen to mimic results provided by the automatic
+classifier . Whereas this paradigm is longer and more complicated, it
+can lead to more robust results in reducing the forced-choice biais
+(Russell 1993). However this procedure can also reduce the human
+observers accuracy. In this regard, the results of the human observation
+could have been more ambiguous because it is not the natural way that
+people are inferring meaning from facial expressions. An alternative
+explanation relies in reducing the recognition biais involved in the
+classic recognition paradigm. Classic forced-choice paradigms obtain
+artificial high results, thus by using a more evolved approach
+observers' accuracy may have been lowered.
+
+Considering the above, the results provide additional evidence that
 individuals' subjective feeling can not be inferred from facial
 expressions and in our case invalidate the hypothesis of hardwired
-emotions. This result suggests that automatic facial expression
-recognition tools should be focused on evaluating facial morphology
-features such as action units rather than inferring potential emotional
-or affective states.
+emotions unambiguously displayed on the face. Even if emotions were
+hardwired, in everyday life one does not observe prototypical facial
+expressions and therefore research should be focused on analysing
+non-prototypical facial expressions. Advancements in identifying
+``non-basic'' emotion labels (McDuff 2016) as well as non-prototypical
+facial expression have been made in the developpement of automatic
+facial expression recognition tools. However this result suggests that
+automatic facial expression recognition tools should merely evaluate
+facial morphology features such as action units (e.g., OpenFace
+presented in Baltrušaitis, Robinson, and Morency 2016) rather than
+inferring supposedly emotional or affective states.
 
@@
 who developed the software used to collect and preprocess human observer
-annotations.
+results.
 
@@
 
+\leavevmode\hypertarget{ref-baltruvsaitis2016openface}{}%
+Baltrušaitis, Tadas, Peter Robinson, and Louis-Philippe Morency. 2016.
+``Openface: An Open Source Facial Behavior Analysis Toolkit.'' In
+\emph{2016 Ieee Winter Conference on Applications of Computer Vision
+(Wacv)}, 1--10. IEEE.
+
 \leavevmode\hypertarget{ref-barrett2017emotions}{}%
@@
 Cognitive and Affective Neuroscience} 12 (1): 1--23.
+
+\leavevmode\hypertarget{ref-bonanno2004brief}{}%
+Bonanno, George, and Dacher Keltner. 2004. ``Brief Report the Coherence
+of Emotion Systems: Comparing `on-Line' Measures of Appraisal and Facial
+Expressions, and Self-Report.'' \emph{Cognition and Emotion} 18 (3):
+431--44.
 
@@
 
+\leavevmode\hypertarget{ref-doyle2017language}{}%
+Doyle, Cameron M, and Kristen A Lindquist. 2017. ``Language and Emotion:
+Hypotheses on the Constructed Nature of Emotion Perception.''
+
 \leavevmode\hypertarget{ref-dupre2015oudjat}{}%
@@
 
+\leavevmode\hypertarget{ref-duran2017coherence}{}%
+Durán, Juan I, Rainer Reisenzein, and José-Miguel Fernández-Dols. 2017.
+``Coherence Between Emotions and Facial Expressions.'' \emph{The Science
+of Facial Expression}, 107--29.
+
 \leavevmode\hypertarget{ref-ekman1992argument}{}%
@@
 Emotion Elicitation and Assessment} 47: 53.
+
+\leavevmode\hypertarget{ref-ekman1987universals}{}%
+Ekman, Paul, Wallace V Friesen, Maureen O'sullivan, Anthony Chan, Irene
+Diacoyanni-Tarlatzis, Karl Heider, Rainer Krause, et al. 1987.
+``Universals and Cultural Differences in the Judgments of Facial
+Expressions of Emotion.'' \emph{Journal of Personality and Social
+Psychology} 53 (4): 712.
+
+\leavevmode\hypertarget{ref-ekman1988universality}{}%
+Ekman, Paul, and Karl G Heider. 1988. ``The Universality of a Contempt
+Expression: A Replication.'' \emph{Motivation and Emotion} 12 (3):
+303--8.
+
+\leavevmode\hypertarget{ref-fernandez2013emotion}{}%
+Fernández-Dols, José-Miguel, and Carlos Crivelli. 2013. ``Emotion and
+Expression: Naturalistic Studies.'' \emph{Emotion Review} 5 (1): 24--29.
 
@@
 78--102.
+
+\leavevmode\hypertarget{ref-kappas2003facial}{}%
+Kappas, Arvid. 2003. ``What Facial Activity Can and Cannot Tell Us About
+Emotions.'' In \emph{The Human Face}, 215--34. Springer.
+
+\leavevmode\hypertarget{ref-kraut1979social}{}%
+Kraut, Robert E, and Robert E Johnston. 1979. ``Social and Emotional
+Messages of Smiling: An Ethological Approach.'' \emph{Journal of
+Personality and Social Psychology} 37 (9): 1539.
+
+\leavevmode\hypertarget{ref-lindquist2013s}{}%
+Lindquist, Kristen A, and Maria Gendron. 2013. ``What's in a Word?
+Language Constructs Emotion Perception.'' \emph{Emotion Review} 5 (1):
+66--71.
+
+\leavevmode\hypertarget{ref-mcduff2016discovering}{}%
+McDuff, Daniel. 2016. ``Discovering Facial Expressions for States of
+Amused, Persuaded, Informed, Sentimental and Inspired.'' In
+\emph{Proceedings of the 18th Acm International Conference on Multimodal
+Interaction}, 71--75. ACM.
 
@@
 
+\leavevmode\hypertarget{ref-niedenthal2017embodied}{}%
+Niedenthal, Paula M, Adrienne Wood, Magdalena Rychlowska, and Sebastian
+Korb. 2017. ``Embodied Simulation in Decoding Facial Expression.''
+\emph{The Science of Facial Expression}, 397.
+
+\leavevmode\hypertarget{ref-russell1993forced}{}%
+Russell, James A. 1993. ``Forced-Choice Response Format in the Study of
+Facial Expression.'' \emph{Motivation and Emotion} 17 (1): 41--51.
+
 \leavevmode\hypertarget{ref-tcherkassof2013dynemo}{}%

### 2019-04-05 ACII_2019_paper/ACII_2019_word.Rmd
latex

--- 2019-03-21 ACII_2019_paper/ACII_2019_word.Rmd
@@
 ---
-title: "Coherence between subjective feeling and facial expression: a time-series analysis of human and automatic recognition."
+title: "Coherence between subjective feeling and facial expression: a comparison of human and automatic recognition."
 output: word_document
@@
 On one hand the behaviorist approach conceives emotions as genetically hardwired and therefore being genuinely displayed through facial expressions.
-On the other hand the constructivist approach conceives emotions are socially constructed and facial expression as social messages that are not related to emotions.
-In order to evaluate the link between the subjective feeling of emotions and their recognition based on facial expression, 232 videos of participants recruited to perform an emotion elicitation task were annotated by 1383 human observers as well as by an automatic facial expression classifier.
-The results show a low accuracy of human observers and of the automatic classifier to infer from the facial expression the subjective feeling of the participants recorded.
-Based on these results, the hypothesis of genetically hardwired emotion genuinely display is difficult to support whereas the idea of emotion socially constructed and facial expression as display of social messages appears to be more likely.
-Then, the way to infer emotional and mental state based on facial expressions should be questioned.
+On the other hand the constructivist approach conceives emotions as socially constructed, the emotional meaning of a facial expression being inferred by the observer.
+In order to evaluate the link between the subjective feeling of emotions and their recognition based on facial expression, 232 videos of encoders recruited to carry out an emotion elicitation task were annotated by 1383 human observers as well as by an automatic facial expression classifier.
+Results show a low accuracy of human observers and of the automatic classifier to infer the subjective feeling from the facial expressions displayed by encoders .
+They also show a weak link between self-reported emotional states and facial emotional displays.
+Based on these results, the hypothesis of genetically hardwired emotion genuinely display is difficult to support whereas the idea of emotion and facial expression socially constructed appears to be more likely.
+Then automatic emotion recognition tools  based on facial expressions should be questioned.
+
 ---
@@
 
-With the development of commercial automatic facial expression recognition tools [see @dupre2018accuracy for a non-exhaustive list of available tools], industries and governments are gradually implementing this technology in order to track humans' emotions in various scenarios (e.g., marketing, healthcare, automotive to name a few).
-This technoloogy rests on the premise that facial expressions provide a direct access to individuals' subjective feeling.
+With the development of commercial automatic facial expression recognition tools [see @dupre2018accuracy for a non-exhaustive list of available tools]), industries and governments are gradually implementing this technology in order to track humans’ emotions in various scenarios (e.g., marketing, healthcare, automotive to name a few).
+This technoloogy rests on the premise that facial expressions provide a direct access to individuals’ subjective feeling.
 Even if this premise is central to the modern mainstream approach of human emotion, recent research in affective science are challenging it.
-After presenting the arguments supporting both sides, an experiment testing these hypotheses will be presented and its results analyzed in order to provide empirical evidence to contribute to answer the question.
+Once the two competing approaches briefly described as well as the predictions they respectively entail, an experiment testing these hypotheses will be presented and its results analyzed in order to provide empirical evidence to contribute to answer the question.
 
 ## The link between emotion and facial expression through the behaviorist approach
-Based on the behaviorist approach initiated by Darwin in *The Expression of the Emotions in Man and Animals* [@darwin1872expression], facial expression are conceived as a genuine display of individuals inner emotional state.
+Based on the behaviorist approach initiated by Darwin in *The Expression of the Emotions in Man and Animals* [@darwin1872expression], facial expressions are conceived as a genuine displays of individuals’ inner emotional state.
 This hypothesis is used as a basis for the Basic Emotion Theory (BET) which states that a set of six emotions are universally displayed and are genetically hardwired not only in humans [@ekman1992argument] but also in different animal species [@de2019mama].
-According to this view, "when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including [...] facial expressions." [@ekman2007directed, p49].
-To cope with critics, several amendments have been made to the BET, increasing the number of basic emotions from six to seven  [@ekman1988universality] as well as adding the concept of "display rules" to explain cultural differences in the management of facial expressions [@ekman1987universals].
-Even if this theory obtained a popular support, it fails to explain how individuals can feel emotions without expressing them and how individuals can express emotions without feeling them, cases for which display rules cannot be applied to [@kraut1979social, @duran2017coherence].
+According to this view, “when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including […] facial expressions.” [@ekman2007directed, p49]).
+To cope with critics, several amendments have been made to the BET, increasing the number of basic emotions from six to seven [@ekman1988universality] ) as well as adding the concept of “display rules” to explain cultural differences in the management of facial expressions [@ekman1987universals].
+
+Even if this theory obtained a popular support, it fails to explain how individuals can feel emotions without expressing them and how individuals can express emotions without feeling them, in instances in which display rules cannot be called upon [@kraut1979social; @duran2017coherence].
+These evidences have led to an alternative conception, notably the social constructivist approach.
 
 ## The link between emotion and facial expression through the social constructivist approach
-Detractors of the Basic Emotion Theory are perceiving emotion not as genetically hardwired but as a learnt association between a given situation and an appropriate response [@averill1980constructivist, @barrett2017emotions].
-For the tenants of the constructivist approach, emotions are "concepts" based on past experiences and which are "a collection of embodied, whole brain representations that predict what is about to happen in the sensory environment, what the best action is to deal with impending events, and their consequences for allostasis" [@barrett2017theory, p12].
-Following this assumption, faces are used as tools to display signals in social interactions [@crivelli2018facial].
-These signals can convey individuals' motivations and readiness [@frijda1997facial] or social messages [@fridlund1995human].
+Detractors of the Basic Emotion Theory consider emotion not as genetically hardwired but as a learnt association between a given situation and an appropriate response [@averill1980constructivist; @barrett2017emotions]).
+For the tenants of the constructivist approach, emotions are “concepts” based on past experiences and which are “a collection of embodied, whole brain representations that predict what is about to happen in the sensory environment, what the best action is to deal with impending events, and their consequences for allostasis” [@barrett2017theory, p12].
+Following this assumption, faces are best conceived as tools displaying signals in social interactions [@crivelli2018facial].
+These signals can convey individuals’ motivations and readiness [@frijda1997facial] or social messages [@fridlund1995human].
 Therefore, facial expressions are thought as behaviors which meaning is inferred by the observer.
+Findings support this observer dependence [@lindquist2013s; @niedenthal2017embodied].
+They show that to make meaning of another person’s facial behavior, the perceiver relies in particular on her/his knowledge about emotion categories.
 
-This current paper investigates the link between the subjective feeling of emotions and their recognition from facial expressions.
-If emotions are hardwired, individuals emotional subjective feeling should be correlated to the recognition of facial expressions from both human observers and automatic classifiers whereas if emotions are social constructs, no correlation between subjective feeling and facial expression recognition should be observed.
+## The recognition of emotional facial expressions
+Regarding emotional facial expression (EFE) recognition, the behaviorist approach assertions and the constructivist approach claims lead to two divergent predictions.
+The foremost postulates that, when triggered, each basic emotion is expressed by a prototypical face (non basic emotions being blends of the basic ones).
+Basic emotions are easily recognized by all human observers and emotional states are accessible by facial measurement.
+As the recognition of facial expressions is based on the identification of specific patterns of facial movements, it implies that EFE recognition of both human observers and automatic classifiers should be as accurate.
+The constructivist approach, in contrast, affirms that facial expressions do not provide a direct access to individuals’ subjective feeling.
+Even if the face possibly moves during an emotional episode, facial muscle movements are not linked in a one-to-one manner to a specific discrete emotional experience [@doyle2017language, p418].
+Instead, emotions are mentally constructed by the perceiver and mental categories of emotions are needed to accurately categorize facial movements.
+Therefore, since there is no emotional prototypical face, one should expect human observers’ superior ability to accurately recognize EFE as compared to automatic EFE recognition tools.
+
+The aim of the current paper is to investigate the link between the subjective feeling of emotions and its recognition from facial expressions in order to give credit either to the behaviorist plea or to the constructivist’s one.
+Contrary to most studies using posed and static EFE, the present study focuses on natural EFE.
+Spontaneous and dynamic facial reactions to emotional elicitations are under consideration to ensure the generalizability of the results to emotional behaviors in ordinary life.
 
@@
 
-For the emotion elicitation experiment, `r nrow(self_report_data)` French participants (`r self_report_gender[self_report_gender$genre_c == "F","n"]` females, `r self_report_gender[self_report_gender$genre_c == "H","n"]` males, *M*age = `r self_report_age[,"Age_m"]`, *SD*age = `r self_report_age[,"Age_sd"]`) were recruited to perform one out of 11 emotion elicitation tasks designed to trigger a positive, negative or neutral emotional state.
-Participants' face were recorded using an hidden camera resulting `r nrow(self_report_data)` front facing 768x576 videos varying from `r min(metadata_video$ffprobe_duration)`s to `r max(metadata_video$ffprobe_duration)`s (`r fig_nums("dynemo_img", display = "cite")`).
-These recordings are constituting the DynEmo database [see @tcherkassof2013dynemo for a full description of tasks and procedure].
+For the emotion elicitation experiment, `r nrow(self_report_data)` French participants (`r self_report_gender[self_report_gender$genre_c == "F","n"]` females, `r self_report_gender[self_report_gender$genre_c == "H","n"]` males, *M*age = `r self_report_age[,"Age_m"]`, *SD*age = `r self_report_age[,"Age_sd"]`) were recruited to perform one out of 11 emotion elicitation tasks designed to trigger a positive, a specific negative or a neutral emotional state.
+Encoders’ face were recorded using an hidden camera resulting `r nrow(self_report_data)` front facing 768x576 videos varying from `r min(metadata_video$ffprobe_duration)`s to `r max(metadata_video$ffprobe_duration)`s (`r fig_nums("dynemo_img", display = "cite")`).
+These recordings form the DynEmo database [see @tcherkassof2013dynemo for a full description of tasks and procedure].
 
@@
 
-After their emotion elicitation task the participants had to rate their subjective feeling during the task on likert scales from 0 ("not at all") to 5 ("strongly") according six "basic" emotion labels (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as six "non-basic" emotion labels (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
+After the emotion elicitation task the encoders rated their subjective feeling on Likert scales from 0 ("not at all") to 5 ("strongly") related to six “basic” emotion labels (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as six “non-basic” emotion labels (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
 
-Finally, a debriefing session was perform to ensure that participants were not durably affected by the emotion elicitation task.
-The debriefing was also used to check that participants did not guess the real purpose of the experiment (e.g.
-being filmed while they were performing an emotional elicitation task) to guarantee facial expressions genuineness.
-All the participants gave their agreement on their data and video to be processed for research purpose only.
+Finally, a debriefing session was performed to ensure that encoders were not durably affected by the emotion elicitation task.
+The debriefing was also used to check that encoders did not guess the real purpose of the experiment (e.g.
+being filmed while they were performing an emotional elicitation task) to guarantee facial expressions’ genuineness.
+All encoders gave their agreement on their data and video to be processed for research purpose only.
 
@@
 For the human facial expression recognition method, `r nrow(unique_ppt_count)` student participants were recruited to annotate `r nrow(unique_video_count)` out of the `r nrow(self_report_data)` video, therefore only the `r nrow(unique_video_count)` annotated videos will be analysed in this paper.
-Each participants had to annotate between `r min(unique_ppt_count$n)` and `r max(unique_ppt_count$n)` videos resulting that each video was annotated `r mean(unique_video_count$n) %>% round(0)` times on average (*SD* = `r sd(unique_video_count$n) %>% round(0)`).
+Because videos have different durations, participants had to annotate a series of video corresponding to 30min long in total.
+Each video was annotated `r mean(unique_video_count$n) %>% round(0)` times on average (*SD* = `r sd(unique_video_count$n) %>% round(0)`).
 
 The annotation of facial expressions was performed on-site using *Oudjat*, a software for designing video annotation experiments [@dupre2015oudjat].
-For each video, the annotation procedure hat two steps.
+For each video, the annotation procedure followed two steps.
 First, the participants had to identify the emotional sequences by pressing the space bar of their keyboard to indicate the beginning and the end of the emotional sequences while watching the video.
-Second, the participants watched each emotional sequence previously identified and had to label the sequence using one of the 12 emotions proposed including six "basic" emotion labels (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) and six "non-basic" emotion labels (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
-They also had the possibility to indicate that the sequence was not expressing one of the proposed emotion.
+Second, the participants watched each emotional sequence previously identified and labeled the sequence using one of the 12 emotions proposed including six “basic” emotion labels (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) and six "non-basic" emotion labels (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
+They also had the possibility to indicate that the sequence was expressing none of the proposed emotion labels.
 
 This annotation procedure results in a uni-dimensional time-series for each video per human observer identifying for each second of the video which emotion was recognized.
-Then time-series corresponding to the same video were aggregated to calculate the proportion of human observers for each second of the video per emotional label (EQ1).
+Then, time-series corresponding to the same video were aggregated to calculate the proportion of human observers $x_{video_{i}.label_{j}.t_{k}}$ for each second of the video per emotional label \ref{eq:1}.
 
-$$x_{video_{i}.label_{i}.t_{i}} = \frac{n_{video_{i}.label_{i}.t_{i}}}{n_{video_{i}}}$$
+\begin{equation}
+ \label{eq:1}
+x_{video_{i}.label_{j}.t_{k}} = \frac{n_{video_{i}.label_{j}.t_{k}}}{n_{video_{k}}}
+\end{equation}
 
-The sum of each label proportion per second was used as a score to determine which labels corresponds to the overall video (i.e., the highest score).
-
-$$x_{video.label.t_{i}} = \frac{n_{video.label.t_{i}}}{n_{video}}$$
-
-In case of more than one label having the maximum value, the emotion is described as undetermined.
+where *i* is one of the `r nrow(unique_video_count)` videos, *j* is one of the six “basic” emotion labels, *k* for each second of the video.
 
@@
 Affdex’s algorithm uses Histogram of Oriented Gradient (HOG) features and Support Vector Machine (SVM) classifiers in order to recognize facial expressions.
-For each video frame, Affdex identify the probability of the face as expressing one of the six "basic" emotion labels (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as additional psychological states such as *valence*, *engagement* or *contempt*, and facial features such as *cheek raise*, *eye widen* or *jaw drop*.
+For each video frame, Affdex identifies the probability $p_{video_{i}.label_{j}.t_{k}}$ from 0 to 100 (rescaled to 0 to 1 for the analysis) of the face as expressing one of the six “basic” emotion labels (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as additional psychological states such as *valence*, *engagement* or *contempt*, and facial features such as *cheek raise*, *eye widen* or *jaw drop*.
 
-To determine which of the six "basic" emotion can be used to identify each video, the recognition probability for each label by frame was converted into odd ratio by frame [@dente2017measures].
-The highest sum of each odd ratio time-series defines the label recognized by the automatic classifier.
+For both human and automatic recognition, to determine which of the six “basic” emotions can be used to identify each video, the recognition probability for each label by frame was converted into odd ratio by label [@dente2017measures].
+The highest sum of each odd ratio time-series defines the label recognized by the automatic classifier (\ref{eq:1} for human recognition and \ref{eq:1} for automatic recognition).
+\begin{equation}
+ \label{eq:2}
+video_{i}.label = \max\left(\frac{\sum_{k=1}^{n}x_{video_{i}.label_{j}.t_{k}}}{\sum_{k=1}^{n}x_{video_{i}.t_{k}}}\right)
+\end{equation}
+
+\begin{equation}
+ \label{eq:3}
+video_{i}.label = \max\left(\frac{\sum_{k=1}^{f}p_{video_{i}.label_{j}.t_{k}}}{\sum_{k=1}^{f}p_{video_{i}.t_{k}}}\right)
+\end{equation}
+
+where *i* is one of the `r nrow(unique_video_count)` videos, *j* is one of the six “basic” emotion labels, *k* for each second of the *n* second video or for each frame of the *f* frame video.
 
@@
 
-Whereas the self-reports, the human observation and the automatic recognition include data on "non-basic" emotion labels and features, the analysis is performed using only the six "basic" emotion labels in order to compare them.
+Since encoders’ self-reports, human annotations and the automatic recognition include data on “non-basic” emotion labels and features, the analysis is performed using only the six “basic” emotion labels in order to compare them.
 The maximum score for self-reports, human annotations and automatic recognition is used to label the video.
@@
 
-Participants' subjective feeling is compared with human observers recognition in a confusion matrix (`r fig_nums("confusionMatrix_sr_hr", display = "cite")`).
-Each emotion label used to describe participants subjective feeling (i.e., the label rated with the highest value) are compared with the emotion labels which was rated with the highest score by human observers
+Encoders’ subjective feeling is compared with human observers recognition in a confusion matrix (`r fig_nums("confusionMatrix_sr_hr", display = "cite")`).
+Each emotion label used to describe encoders’ self-reported subjective feeling (i.e., the label rated with the highest value) is compared with the emotion labels which were rated with the highest score by human observers.
 
@@
 
-The result of the confusion matrix show a moderate agreement between emotion felt during the elicitation and emotion recognized by the human observers (Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *disgust* (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_sr"]))` of the videos self-reported).
+Results of the confusion matrix show a moderate agreement between the emotion felt by the encoder during the elicitation and the emotion recognized by the human observers (Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *disgust* (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_sr"]))` of the videos self-reported).
 Sensitivity, specificity, precision and F1 score for each emotion reveals that *happiness* has the highest coherence ratio whereas *sadness* has the lowest coherence ratio between true positives and false positives `r table_nums("confusionTable_sr_hr", display = "cite")`.
@@
 
-However, the self-report show a very high proportion of undetermined emotional states which reveals the potential limit of using 6-points likert scales for which the participants can easily score to the maximum for more than one emotion.
+However, self-reports show a very high proportion  of undetermined emotional states which reveals the potential limit of using 6-points likert scales to measure emotional self-reports.
+Indeed, encoders can easily score to the maximum for more than one emotion.
 
@@
 
-Similarly to the previous analysis, a confusion matrix was used to compare participants subjective feeling with the emotion label recognized by the automatic classifier (`r fig_nums("confusionMatrix_sr_ar", display = "cite")`).
+Similarly to the previous analysis, a confusion matrix was used to compare encoders’ subjective feeling with the emotion label recognized by the automatic classifier (`r fig_nums("confusionMatrix_sr_ar", display = "cite")`).
 
@@
 
-Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are similar to the ones with human observers (`r table_nums("confusionTable_sr_ar", display = "cite")`).
-Overall there is a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_ar$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_ar$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_ar$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` of the video self-reported).
+Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are somewhat similar to the ones with human observers (`r table_nums("confusionTable_sr_ar", display = "cite")`).
+Overall, a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_ar$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_ar$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_ar$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` of the video self-reported) is watched.
 Surprisingly the automatic classifier incorrectly recognized as *disgust* an important proportion of videos in which *anger*, *happiness* and *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "anger", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
@@
 
-A comparable explanation can be provided as the level of undetermined emotion are very high for the self reports.
+A comparable explanation  can be provided as the level of undetermined emotions are very high for the self reports.
 
@@
 
-As previously mentioned, the accuracy of humans observers and the automatic classifier have some similarities in there mistakes.
-However human observers appears to be more accurate than the automatic classifier to recognize individuals' subjective feeling (human observer Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`; automatic classifier Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`.
+As previously mentioned, human observers appear to be more accurate than the automatic classifier to recognize individuals’ subjective feeling (human observer Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`; automatic classifier Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`.
+).
+However, both make mistakes.
 A third confusion matrix is used to compare similarities between human observers and automatic classifier (`r fig_nums("confusionMatrix_hr_ar", display = "cite")`).
@@
 ============
-Despite being one on the most investigated question in affective science, the coherence between emotion felt and facial expression is a hot topic and no clear evidence have been found to definitely answer it.
-However, with the growing interest of industries and government to monitor individual's psychological states, evidences are showing that facial expressions are in reality not expressing emotions [@mckeown2013analogical].
-This research aimed to provide some empirical data to the question.
-The subjective feeling of participants was compared with human recognition on one side and automatic recognition on the other side.
-The results reveals a low accuracy for both humans and automatic classifier to accurately identify the inner emotional states of these individuals based on their facial expressions.
-Moreover it appeared that human observers obtain a higher accuracy to recognize the corresponding emotion label than the automatic classifier is.
-A possible explaination is that human oberservers are assessing expressing sequences while having access to the full video which provides a context whereas the automatic classifier only assess the videos frame by frame.
 
-Some limitations to this process should be stated over the use of self-reports to evaluate individual's subjective feelings.
+Despite being one on the most investigated question in affective science, the coherence between emotion felt and facially displayed and facial expression recognized is a hot topic.
+To date no clear evidence has been found to definitely answer it.
+Yet, with the growing interest of industries and government to monitor individual’s psychological states, this issue is under high pressure.
+The present research aims to provide some empirical data to the question.
+The faces encoders displayed when confronted with an emotional eliciting task were submitted both to human and to automatic recognition.
+The criterion for recognition accuracy was the subjective feeling self-reported by the encoder once the elicitation task carried out.
+Results first reveal a low coherence between emotion felt and facial expression displayed.
+Secondly, results show a low accuracy for both humans and automatic classifier in identifying the inner emotional states of these encoders based on their facial expressions.
+Thirdly, human observers prove to be better at recognizing the emotion facially expressed than the automatic recognition tool is.
+
+Such results support the hypothesis advanced by some authors of a low emotion–expression coherence [e.g., @kappas2003facial].
+In many instances, facial displays are not associated with a concordant emotional state, even any emotional state at all [@bonanno2004brief; @fernandez2013emotion].
+More and more evidences are showing that facial expressions are in reality not expressing emotions [@mckeown2013analogical].
+As well as other nonverbal behaviors, facial movements are not only assumed to be determined by emotion but also by various other causes, such as psychological states (e.g., motivations or pain), to say nothing of social context and sociocultural norms [such as “display rules”, @ekman1987universals].
+This multiple determination excludes any possibility of drawing an inference from facial activity on the underlying psychological state (emotional or other).
+Beyond the present observations showing a weak coherence between subjective feelings and spontaneous facial expressions, this study provides some light on the controversy between the behaviorist and the constructivist approaches.
+The first of those two lines of thinking [@ekman1992argument] assumes that expressions of emotion are brief and coherent patterns of facial muscle movements that co-vary with discrete subjective experiences.
+Emotions and their related prototypical facial behavior are universal because they are considered as innate mechanisms allowing individuals to respond adaptively to evolutionary significant events (threats, opportunities…) encountered in the environment.
+Instead of viewing emotions as natural kinds [@barrett2017theory], the constructivist approach supposes that emotions are social constructions.
+The specific emotion categories identified on the face of another person are the result of categorization.
+In other words, the emotions that are recognized by the observer are constructed in her/his mind, primarily based on her/his conceptual knowledge of emotion.
+Therefore, facial movements do not express specific emotions.
+It is the observer that infers the emotional meaning of the facial expression.
+As a consequence, one can predict from the first line of thinking that individuals’ emotional subjective feeling should be correlated to the recognition of facial expressions from both human observers and automatic classifiers whereas if emotions are social constructs, as stated by the second line of thinking, human observer’s should be better at perceiving emotions in face than automatic classifiers.
+Present results plead in favor of the latter stance.
+They show that human observers are more accurate than the automatic recognition tool  to identify individuals’ subjective feeling on the basis of their face.
+Moreover, mistakes made by human observers look less arbitrary to the ones made by the classifier.
+For instance, even if mix-up between disgust and anger is sometimes reported in recognition studies, confusions such as the ones produced by the classifier have never been noted for human observers.
+A possible explanation is that human observers are assessing dynamic expressive sequences.
+They have access to the full video which provides a kinetic context whereas the automatic classifier only assesses the videos frame by frame.
+Further research is needed to shed light on this issue.
+
+Some limitations should be stated, notably regarding the use of self-reports to evaluate encoders’ subjective feelings.
 Accessing the inner subjective feeling can be biased if not impossible.
 Moreover the procedure used for human observation can also be open to dispute.
-Instead of asking the human annotators to provide an unique label, a more subtle approach was chosen to mimic results provided by the automatic classifier.
+Instead of asking the human annotators to provide a unique label, a more subtle approach was chosen to mimic results provided by the automatic classifier .
 Whereas this paradigm is longer and more complicated, it can lead to more robust results in reducing the forced-choice biais [@russell1993forced].
-However this procedure can also reduce the human obersvers accuracy.
+However this procedure can also reduce the human observers accuracy.
 In this regard, the results of the human observation could have been more ambiguous because it is not the natural way that people are inferring meaning from facial expressions.
-An alternative explaination relies in reducing the recognition biais involved in the classic recognition paradigm.
-Classic forced-choice paradigms obtain artificial high results, thus by using a more evolved approach observers' accuracy may have been reduced.
+An alternative explanation relies in reducing the recognition biais involved in the classic recognition paradigm.
+Classic forced-choice paradigms obtain artificial high results, thus by using a more evolved approach observers’ accuracy may have been lowered.
 
-Considering the above, the results provides an additional evidence that individuals' subjective feeling can not be inferred from facial expressions and in our case invalidate the hypothesis of hardwired emotions.
+Considering the above, the results provide additional evidence that individuals’ subjective feeling can not be inferred from facial expressions and in our case invalidate the hypothesis of hardwired emotions unambiguously displayed on the face.
 Even if emotions were hardwired, in everyday life one does not observe prototypical facial expressions and therefore research should be focused on analysing non-prototypical facial expressions.
-Advancements in identifying "non-basic" emotion labels [@mcduff2016discovering] as well as non-prototypical facial expression have been made in the developpement of automatic facial expression recognition tools.
-However this result suggests that automatic facial expression recognition tools should evaluate facial morphology features such as action units [e.g., OpenFace @baltruvsaitis2016openface] rather than inferring potential emotional or affective states.
+Advancements in identifying “non-basic” emotion labels [@mcduff2016discovering] as well as non-prototypical facial expression have been made in the developpement of automatic facial expression recognition tools.
+However this result suggests that automatic facial expression recognition tools should merely evaluate facial morphology features such as action units [e.g., OpenFace presented in @baltruvsaitis2016openface] rather than inferring supposedly emotional or affective states.
 

### 2019-04-05 ACII_2019_paper/ACII_2019_paper.Rmd
corr_test

@@
 
-```{r confusionMatrix_hr_ar, fig.height=5, fig.cap="\\label{fig:confusionMatrix_hr_ar}Comparison between the human observers and automatic classifier recognition for the classfication of emotion labels."}
+```{r confusionMatrix_hr_ar, fig.height=5, fig.cap="\\label{fig:confusionMatrix_hr_ar}Proportion of emotion labels classified by human observers which are recognized by the automatif classifier."}
 comparison_hr_ar <- dplyr::inner_join(human_recognition_score,automatic_recognition_score, by = "C_Video") %>% 
@@
   scale_fill_gradient(
-    name = "Proportion of Recognized\nEmotion",
+    name = "Proportion Recognized Emotion",
     low = "white", 

### 2019-04-05 ACII_2019_paper/ACII_2019_paper.tex
corr_test

@@
 \includegraphics{ACII_2019_paper_files/figure-latex/confusionMatrix_hr_ar-1.pdf}
-\caption{\label{fig:confusionMatrix_hr_ar}Comparison between the human
-observers and automatic classifier recognition for the classfication of
-emotion labels.}
+\caption{\label{fig:confusionMatrix_hr_ar}Proportion of emotion labels
+classified by human observers which are recognized by the automatif
+classifier.}
 \end{figure}

### 2019-04-05 corr_test.R

@@ -1,0 +1,83 @@
+list_video <- self_report_score$C_Video
+list_emotion <- c("disgust","fear","happiness","surprise","sadness","anger")
+
+grid_data <- expand.grid(list_video = list_video, list_emotion = list_emotion)
+
+self_report_result <- self_report_score %>% 
+  dplyr::filter(sr_emotion != "undetermined") %>% 
+  dplyr::mutate(sr_score = 1)
+
+human_recognition_result <- human_recognition_score %>% 
+  dplyr::filter(hr_emotion != "undetermined") %>% 
+  dplyr::mutate(hr_score = 1)
+
+automatic_recognition_result <- automatic_recognition_score %>% 
+  dplyr::filter(ar_emotion != "undetermined") %>% 
+  dplyr::mutate(ar_score = 1)
+
+recognition_result <- grid_data %>% 
+  dplyr::left_join(self_report_result, by = c(
+    "list_video" = "C_Video",
+    "list_emotion" = "sr_emotion"
+    )
+  ) %>% 
+  dplyr::left_join(human_recognition_result, by = c(
+    "list_video" = "C_Video",
+    "list_emotion" = "hr_emotion"
+    )
+  ) %>% 
+  dplyr::left_join(automatic_recognition_result, by = c(
+    "list_video" = "C_Video",
+    "list_emotion" = "ar_emotion"
+    )
+  ) %>% 
+  replace(is.na(.), 0)
+
+recognition_result %>% 
+  dplyr::select(-list_video, -list_emotion) %>% 
+  cor(method = c("pearson")) %>%
+  ggcorrplot::ggcorrplot(
+    type = "lower",
+    outline.col = "white", 
+    lab = TRUE, 
+    legend.title = "Correlation\n(Pearson)", 
+    lab_size = 4)
+
+recognition_result %>% 
+  ggplot(aes(d = sr_score, m = hr_score)) + 
+  plotROC::geom_roc(labels = FALSE,pointsize = 0) 
+
+recognition_result %>% 
+  ggplot(aes(d = sr_score, m = ar_score)) + 
+  plotROC::geom_roc(labels = FALSE,pointsize = 0) 
+
+
+human_roc <- pROC::roc(recognition_result$sr_score, recognition_result$hr_score)
+human_auc <- pROC::auc(human_roc)
+human_auc
+
+automatic_roc <- pROC::roc(recognition_result$sr_score, recognition_result$ar_score)
[…]

### 2019-04-08 ACII_2019_paper/ACII_2019_paper.Rmd
final version

@@
   - \usepackage{tabu}
+  - \usepackage{wrapfig}
+  - \floatplacement{figure}{H}
+  - \floatplacement{table}{H}
 
@@
 output: rticles::ieee_article
+fig_caption: yes
+keep_tex: yes
+includes:
+  in_header: preamble-latex.tex
 ---
@@
   echo = FALSE,
-  fig.pos = 'H')
+  fig.pos = "H")
 #
@@
 library(pROC)
+library(xtable)
+library(Hmisc)
 ```
@@
 Following this assumption, faces are best conceived as tools displaying signals in social interactions [@crivelli2018facial].
-These signals can convey individuals’ motivations and readiness [@frijda1997facial] or social messages [@fridlund1995human].
+These signals can convey individuals’ motivations and readiness [@frijda1997facial] or social messages [@fridlund1994human].
 Therefore, facial expressions are thought as behaviors which meaning is inferred by the observer.
@@
 
-## Correlation between self-report and human facial expression recognition
+```{r corr-matrix}
+# list_video <- self_report_score$C_Video
+# list_emotion <- c("disgust","fear","happiness","surprise","sadness","anger")
+# 
+# grid_data <- expand.grid(list_video = list_video, list_emotion = list_emotion)
+# 
+# self_report_result <- self_report_score %>% 
+#   dplyr::filter(sr_emotion != "undetermined") %>% 
+#   dplyr::mutate(sr_score = 1)
+# 
+# human_recognition_result <- human_recognition_score %>% 
+#   dplyr::filter(hr_emotion != "undetermined") %>% 
+#   dplyr::mutate(hr_score = 1)
+# 
+# automatic_recognition_result <- automatic_recognition_score %>% 
+#   dplyr::filter(ar_emotion != "undetermined") %>% 
+#   dplyr::mutate(ar_score = 1)
+# 
+# recognition_result <- grid_data %>% 
+#   dplyr::left_join(self_report_result, by = c(
+#     "list_video" = "C_Video",
+#     "list_emotion" = "sr_emotion"
+#     )
+#   ) %>% 
+#   dplyr::left_join(human_recognition_result, by = c(
+#     "list_video" = "C_Video",
+#     "list_emotion" = "hr_emotion"
+#     )
+#   ) %>% 
+#   dplyr::left_join(automatic_recognition_result, by = c(
+#     "list_video" = "C_Video",
+#     "list_emotion" = "ar_emotion"
+#     )
+#   ) %>% 
+#   replace(is.na(.), 0) %>% 
+#   dplyr::select(-list_video, -list_emotion)
+# 
+# 
+#   x <- as.matrix(recognition_result)
+#   R <- Hmisc::rcorr(x)$r
+#   p <- Hmisc::rcorr(x)$P
+#   
+#   ## define notions for significance levels; spacing is important.
+#   mystars <- ifelse(p < .001, "***", ifelse(p < .01, "** ", ifelse(p < .05, "* ", " ")))
+#   
+#   ## trunctuate the matrix that holds the correlations to two decimal
+#   R <- format(round(cbind(rep(-1.11, ncol(x)), R), 2))[,-1]
+#   
+#   ## build a new matrix that includes the correlations with their apropriate stars
+#   Rnew <- matrix(paste(R, mystars, sep=""), ncol=ncol(x))
+#   diag(Rnew) <- paste(diag(R), " ", sep="")
+#   rownames(Rnew) <- colnames(x)
+#   colnames(Rnew) <- paste(colnames(x), "", sep="")
+#   colnames(x)
+#   ## remove upper triangle
+#   Rnew <- as.matrix(Rnew)
+#   Rnew[upper.tri(Rnew, diag = TRUE)] <- ""
+#   Rnew <- as.data.frame(Rnew)
+#   ## remove last column and return the matrix (which is now a data frame)
+#   Rnew <- cbind(Rnew[1:length(Rnew)-1])
+# 
+# 
+# xtable(Rnew)
+```
+
+## Self-reported emotions recognized by human observers
 
@@
     "latex",
-    caption = "\\label{table:confusionTable_sr_hr}Agreement accuracy metrics for each emotion.",
+    caption = "\\label{table:confusionTable_sr_hr}Human recognition accuracy metrics for each emotion.",
     booktabs = TRUE,
@@
     full_width = TRUE, 
-    latex_options = c("hold_position")) %>%
+    latex_options = "HOLD_position") %>%
    add_footnote("Note. \\textit{na.} values are produced when not enough data are available to compute accuracy indicators.", notation="none", escape = FALSE)
@@
 
-## Correlation between self-report and automatic facial expression recognition
+## Self-reported emotions recognized by the automatic classifier
 
@@
     "latex",
-    caption = "\\label{table:confusionTable_sr_ar}Agreement accuracy metrics for each emotion.",
+    caption = "\\label{table:confusionTable_sr_ar}Autonatic recognition accuracy metrics for each emotion.",
     booktabs = T,
@@
 Except for *happiness* and *disgust* (respectively `r scales::percent(as.numeric(confusionMatrix_hr_ar_freq[confusionMatrix_hr_ar_freq$ar_emotion == "happiness" & confusionMatrix_hr_ar_freq$hr_emotion == "happiness", "prop_hr"]))` and `r scales::percent(as.numeric(confusionMatrix_hr_ar_freq[confusionMatrix_hr_ar_freq$ar_emotion == "disgust" & confusionMatrix_hr_ar_freq$hr_emotion == "disgust", "prop_hr"]))` of common labelling), there is no clear common pattern.
-Moreover, the automatic classifier has a tendency to label as *disgust* videos labeled as *sadness* by human observers and as *happiness* videos labaled as *fear* by human observers
+Moreover, the automatic classifier has a tendency to label as *disgust* videos labeled as *sadness* by human observers and as *happiness* videos labeled as *fear* by human observers
 
@@
 Instead of asking the human annotators to provide a unique label, a more subtle approach was chosen to mimic results provided by the automatic classifier .
-Whereas this paradigm is longer and more complicated, it can lead to more robust results in reducing the forced-choice biais [@russell1993forced].
+Whereas this paradigm is longer and more complicated, it can lead to more robust results in reducing the forced-choice bais [@russell1993forced].
 However this procedure can also reduce the human observers accuracy.
 In this regard, the results of the human observation could have been more ambiguous because it is not the natural way that people are inferring meaning from facial expressions.
-An alternative explanation relies in reducing the recognition biais involved in the classic recognition paradigm.
+An alternative explanation relies in reducing the recognition bais involved in the classic recognition paradigm.
 Classic forced-choice paradigms obtain artificial high results, thus by using a more evolved approach observers’ accuracy may have been lowered.

### 2019-04-08 ACII_2019_paper/ACII_2019_paper.tex
final version

@@
 \usepackage{tabu}
+\usepackage{wrapfig}
+\floatplacement{figure}{H}
+\floatplacement{table}{H}
 
@@
 signals can convey individuals' motivations and readiness (Frijda and
-Tcherkassof 1997) or social messages (Fridlund and Rosenberg 1995).
-Therefore, facial expressions are thought as behaviors which meaning is
-inferred by the observer. Findings support this observer dependence
-(Lindquist and Gendron 2013; Niedenthal et al. 2017). They show that to
-make meaning of another person's facial behavior, the perceiver relies
-in particular on her/his knowledge about emotion categories.
+Tcherkassof 1997) or social messages (Fridlund 1994). Therefore, facial
+expressions are thought as behaviors which meaning is inferred by the
+observer. Findings support this observer dependence (Lindquist and
+Gendron 2013; Niedenthal et al. 2017). They show that to make meaning of
+another person's facial behavior, the perceiver relies in particular on
+her/his knowledge about emotion categories.
 
@@
 
-\hypertarget{correlation-between-self-report-and-human-facial-expression-recognition}{%
-\subsection{Correlation between self-report and human facial expression
-recognition}\label{correlation-between-self-report-and-human-facial-expression-recognition}}
+\hypertarget{self-reported-emotions-recognized-by-human-observers}{%
+\subsection{Self-reported emotions recognized by human
+observers}\label{self-reported-emotions-recognized-by-human-observers}}
 
@@
 
-\begin{table}[!h]
-
-\caption{\label{tab:confusionTable_sr_hr}\label{table:confusionTable_sr_hr}Agreement accuracy metrics for each emotion.}
+\begin{table}[H]
+
+\caption{\label{tab:confusionTable_sr_hr}\label{table:confusionTable_sr_hr}Human recognition accuracy metrics for each emotion.}
 \centering
@@
 
-\hypertarget{correlation-between-self-report-and-automatic-facial-expression-recognition}{%
-\subsection{Correlation between self-report and automatic facial
-expression
-recognition}\label{correlation-between-self-report-and-automatic-facial-expression-recognition}}
+\hypertarget{self-reported-emotions-recognized-by-the-automatic-classifier}{%
+\subsection{Self-reported emotions recognized by the automatic
+classifier}\label{self-reported-emotions-recognized-by-the-automatic-classifier}}
 
@@
 
-\caption{\label{tab:confusionTable_sr_ar}\label{table:confusionTable_sr_ar}Agreement accuracy metrics for each emotion.}
+\caption{\label{tab:confusionTable_sr_ar}\label{table:confusionTable_sr_ar}Autonatic recognition accuracy metrics for each emotion.}
 \centering
@@
 labeled as \emph{sadness} by human observers and as \emph{happiness}
-videos labaled as \emph{fear} by human observers
+videos labeled as \emph{fear} by human observers
 
@@
 classifier . Whereas this paradigm is longer and more complicated, it
-can lead to more robust results in reducing the forced-choice biais
+can lead to more robust results in reducing the forced-choice bais
 (Russell 1993). However this procedure can also reduce the human
@@
 people are inferring meaning from facial expressions. An alternative
-explanation relies in reducing the recognition biais involved in the
+explanation relies in reducing the recognition bais involved in the
 classic recognition paradigm. Classic forced-choice paradigms obtain
@@
 Averill, James R. 1980. ``A Constructivist View of Emotion.'' In
-\emph{Theories of Emotion}, 305--39. Elsevier.
+\emph{Theories of Emotion}, edited by Robert Plutchik and Henry
+Kellerman, 305--39. New York, NY: Academic Press.
 
@@
 ``Openface: An Open Source Facial Behavior Analysis Toolkit.'' In
-\emph{2016 Ieee Winter Conference on Applications of Computer Vision
-(Wacv)}, 1--10. IEEE.
+\emph{Proceedings of the Winter Conference on Applications of Computer
+Vision}, 1--10. \url{https://doi.org/10.1109/WACV.2016.7477553}.
 
@@
 Barrett, Lisa Feldman. 2017a. \emph{How Emotions Are Made: The Secret
-Life of the Brain}. Houghton Mifflin Harcourt.
+Life of the Brain}. Boston, MA: Houghton Mifflin Harcourt.
 
@@
 Cognitive and Affective Neuroscience} 12 (1): 1--23.
+\url{https://doi.org/10.1093/scan/nsw154}.
 
 \leavevmode\hypertarget{ref-bonanno2004brief}{}%
-Bonanno, George, and Dacher Keltner. 2004. ``Brief Report the Coherence
-of Emotion Systems: Comparing `on-Line' Measures of Appraisal and Facial
+Bonanno, George, and Dacher Keltner. 2004. ``The Coherence of Emotion
+Systems: Comparing `on-Line' Measures of Appraisal and Facial
 Expressions, and Self-Report.'' \emph{Cognition and Emotion} 18 (3):
-431--44.
+431--44. \url{https://doi.org/10.1080/02699930341000149}.
 
@@
 for Social Influence.'' \emph{Trends in Cognitive Sciences} 22 (5):
-388--99.
+388--99. \url{https://doi.org/10.1016/j.tics.2018.02.006}.
 
@@
 Darwin, Charles. 1872. \emph{The Expression of the Emotions in Man and
-Animals}. London, UK: John Murray, London UK.
+Animals}. London, UK: John Murray.
 
@@
 In \emph{Proceedings of the Conference on the Study of Artificial
-Intelligence and Simulation of Behaviour (Aisb)}, 160--63.
+Intelligence and Simulation of Behaviour}, 160--63.
 
@@
 Doyle, Cameron M, and Kristen A Lindquist. 2017. ``Language and Emotion:
-Hypotheses on the Constructed Nature of Emotion Perception.''
+Hypotheses on the Constructed Nature of Emotion Perception.'' In
+\emph{The Science of Facial Expression}, edited by José-Miguel
+Fernández-Dols and James A Russell, 415--32. New York, NY: Oxford
+University Press.
 
@@
 Human-Computer Studies} 83: 51--61.
+\url{https://doi.org/10.1016/j.ijhcs.2015.05.010}.
 
@@
 Across Different Individuals and Their Facial Expressions.'' In
-\emph{2018 Ieee International Conference on Pervasive Computing and
-Communications Workshops (Percom Workshops)}, 627--32. IEEE.
+\emph{Proceedings of the International Conference on Pervasive Computing
+and Communications}, 627--32.
+\url{https://doi.org/10.1109/PERCOMW.2018.8480127}.
 
@@
 Durán, Juan I, Rainer Reisenzein, and José-Miguel Fernández-Dols. 2017.
-``Coherence Between Emotions and Facial Expressions.'' \emph{The Science
-of Facial Expression}, 107--29.
+``Coherence Between Emotions and Facial Expressions.'' Edited by
+José-Miguel Fernández-Dols and James A Russell, 107--29.
 
@@
 \& Emotion} 6 (3-4): 169--200.
+\url{https://doi.org/10.1080/02699939208411068}.
 
 \leavevmode\hypertarget{ref-ekman2007directed}{}%
----------. 2007. ``The Directed Facial Action Task.'' \emph{Handbook of
-Emotion Elicitation and Assessment} 47: 53.
+---------. 2007. ``The Directed Facial Action Task.'' In \emph{Handbook
+of Emotion Elicitation and Assessment}, edited by James A. Coan and John
+J. B. Allen, 47--53. New York, NY: Oxford University Press.
 
@@
 Expressions of Emotion.'' \emph{Journal of Personality and Social
-Psychology} 53 (4): 712.
+Psychology} 53 (4): 712--17.
+\url{https://doi.org/10.1037/0022-3514.53.4.712}.
 
@@
 Expression: A Replication.'' \emph{Motivation and Emotion} 12 (3):
-303--8.
+303--8. \url{https://doi.org/10.1007/BF00993116}.
 
@@
 Expression: Naturalistic Studies.'' \emph{Emotion Review} 5 (1): 24--29.
-
-\leavevmode\hypertarget{ref-fridlund1995human}{}%
-Fridlund, Alan J, and Erika L Rosenberg. 1995. ``Human Facial
-Expression: An Evolutionary View.'' \emph{Nature} 373 (6515): 569--69.
+\url{https://doi.org/10.1177/1754073912457229}.
+
+\leavevmode\hypertarget{ref-fridlund1994human}{}%
+Fridlund, Alan J. 1994. \emph{Human Facial Expression: An Evolutionary
+View}. San Diego, CA: Academic Press.
 
@@
 Frijda, Nico H, and Anna Tcherkassof. 1997. ``Facial Expressions as
-Modes of Action Readiness.'' \emph{The Psychology of Facial Expression},
-78--102.
+Modes of Action Readiness.'' In \emph{The Psychology of Facial
+Expression}, edited by James A Russell and José-Miguel Fernández-Dols,
+78--102. Cambridge, UK: Cambridge University Press.
 
@@
 Kappas, Arvid. 2003. ``What Facial Activity Can and Cannot Tell Us About
-Emotions.'' In \emph{The Human Face}, 215--34. Springer.
+Emotions.'' In \emph{The Human Face}, edited by Mary Katsikitis,
+215--34. Boston, MA: Springer.
 
@@
 Messages of Smiling: An Ethological Approach.'' \emph{Journal of
-Personality and Social Psychology} 37 (9): 1539.
+Personality and Social Psychology} 37 (9): 1539--53.
 
@@
 Language Constructs Emotion Perception.'' \emph{Emotion Review} 5 (1):
-66--71.
+66--71. \url{https://doi.org/10.1177/1754073912451351}.
 
@@
 Amused, Persuaded, Informed, Sentimental and Inspired.'' In
-\emph{Proceedings of the 18th Acm International Conference on Multimodal
-Interaction}, 71--75. ACM.
+\emph{Proceedings of the International Conference on Multimodal
+Interaction}, 71--75. \url{https://doi.org/10.1145/2993148.2993192}.
 
@@
 Real-Time Multi-Face Expression Recognition Toolkit.'' In
-\emph{Proceedings of the 2016 Chi Conference Extended Abstracts on Human
-Factors in Computing Systems}, 3723--6. ACM.
+\emph{Proceedings of the Chi Conference on Human Factors in Computing
+Systems}, 3723--6. \url{https://doi.org/10.1145/2851581.2890247}.
 
@@
 Communication.'' \emph{Review of General Psychology} 17 (3): 267--87.
+\url{https://doi.org/10.1037/a0032631}.
 
@@
 Niedenthal, Paula M, Adrienne Wood, Magdalena Rychlowska, and Sebastian
-Korb. 2017. ``Embodied Simulation in Decoding Facial Expression.''
-\emph{The Science of Facial Expression}, 397.
+Korb. 2017. ``Embodied Simulation in Decoding Facial Expression.'' In
+\emph{The Science of Facial Expression}, edited by José-Miguel
+Fernández-Dols and James A Russell, 397--414. New York, NY: Oxford
+University Press.
 
@@
 Facial Expression.'' \emph{Motivation and Emotion} 17 (1): 41--51.
+\url{https://doi.org/10.1007/BF00995206}.
 
@@
 Journal of Multimedia \& Its Applications} 5 (5): 61--80.
+\url{https://doi.org/10.5121/ijma.2013.5505}.
 
@@
 Waal, Frans de. 2019. \emph{Mama's Last Hug: Animal Emotions and What
-They Teach Us About Ourselves}. Granta Books.
+They Teach Us About Ourselves}. London, UK: Granta Books.
 

### 2019-04-08 corr_test.R

@@ -61,6 +61,12 @@ automatic_auc <- pROC::auc(automatic_roc)
 automatic_auc
 
 
+# ggplot(recognition_result, aes(x=sr_score, y=ar_score)) +
+#   geom_point() +
+#   geom_smooth(method = "glm", 
+#               method.args = list(family = "binomial"), 
+#               se = FALSE) 
+
 human_logistic <- lme4::glmer(
   sr_score ~ hr_score + (1 | list_video) + (1 | list_emotion), 
   data = recognition_result, 
@@ -80,4 +86,5 @@ automatic_logistic <- lme4::glmer(
 summary(automatic_logistic)
 
 # see prediction accuracy
+
 

### 2019-04-08 corr_test_all.R

@@ -1,0 +1,51 @@
+self_report_score_all <- self_report_tidy %>% 
+  dplyr::select(C_Video, emotion, sr_value = value)
+
+human_recognition_score_all <- average_timeline %>%
+  tidyr::gather(emotion,value, c("disgust","sadness","anger", "fear", "happiness", "surprise")) %>% 
+  dplyr::group_by(C_Video, emotion) %>%
+  dplyr::summarise(sum_emotion = sum(value)) %>% 
+  dplyr::group_by(C_Video) %>% 
+  dplyr::mutate(sum_video = sum(sum_emotion)) %>% 
+  dplyr::mutate(confidence_score = sum_emotion/sum_video) %>% 
+  dplyr::select(C_Video, emotion, hr_value = confidence_score)
+
+automatic_recognition_score_all <- automatic_recognition_data %>% 
+  dplyr::select(C_Video = filename, TimeStamp, happiness = joy, fear, disgust, sadness, anger, surprise) %>% 
+  dplyr::mutate(C_Video = stringr::str_replace(C_Video,".csv","")) %>% 
+  na.omit() %>% 
+  tidyr::gather(emotion, value, c("happiness", "fear", "disgust", "sadness", "anger", "surprise")) %>% 
+  dplyr::mutate(probabilities = value/100) %>% 
+  dplyr::mutate(odds = probabilities/(1-probabilities)) %>% 
+  dplyr::select(C_Video, TimeStamp, emotion, odds) %>% 
+  dplyr::group_by(C_Video, emotion) %>%
+  dplyr::summarise(sum_emotion = sum(odds)) %>% 
+  dplyr::group_by(C_Video) %>% 
+  dplyr::mutate(sum_video = sum(sum_emotion)) %>% 
+  dplyr::mutate(confidence_score = sum_emotion/sum_video) %>% 
+  dplyr::select(C_Video, emotion, ar_value = confidence_score)
+
+recognition_result_all <- self_report_score_all %>% 
+  dplyr::left_join(human_recognition_score_all, by = c(
+    "C_Video",
+    "emotion"
+  )
+  ) %>% 
+  dplyr::left_join(automatic_recognition_score_all, by = c(
+    "C_Video",
+    "emotion"
+  )
+  ) %>% 
+  replace(is.na(.), 0)
+
+recognition_result_all %>% 
+  dplyr::select(-C_Video, -emotion) %>% 
+  cor(method = c("pearson")) %>%
[…]

### 2019-04-08 corr_test_probabilities.R

@@ -1,0 +1,112 @@
+list_video <- self_report_score$C_Video
+list_emotion <- c("disgust","fear","happiness","surprise","sadness","anger")
+
+grid_data <- expand.grid(list_video = list_video, list_emotion = list_emotion)
+
+self_report_result <- self_report_score %>% 
+  dplyr::filter(sr_emotion != "undetermined") %>% 
+  dplyr::mutate(sr_score = 1)
+
+human_recognition_score_all <- average_timeline %>%
+  tidyr::gather(emotion,value, c("disgust","sadness","anger", "fear", "happiness", "surprise")) %>% 
+  dplyr::group_by(C_Video, emotion) %>%
+  dplyr::summarise(sum_emotion = sum(value)) %>% 
+  dplyr::group_by(C_Video) %>% 
+  dplyr::mutate(sum_video = sum(sum_emotion)) %>% 
+  dplyr::mutate(confidence_score = sum_emotion/sum_video) %>% 
+  dplyr::select(C_Video, emotion, hr_value = confidence_score)
+
+automatic_recognition_score_all <- automatic_recognition_data %>% 
+  dplyr::select(C_Video = filename, TimeStamp, happiness = joy, fear, disgust, sadness, anger, surprise) %>% 
+  dplyr::mutate(C_Video = stringr::str_replace(C_Video,".csv","")) %>% 
+  na.omit() %>% 
+  tidyr::gather(emotion, value, c("happiness", "fear", "disgust", "sadness", "anger", "surprise")) %>% 
+  dplyr::mutate(probabilities = value/100) %>% 
+  dplyr::mutate(odds = probabilities/(1-probabilities)) %>% 
+  dplyr::select(C_Video, TimeStamp, emotion, odds) %>% 
+  dplyr::group_by(C_Video, emotion) %>%
+  dplyr::summarise(sum_emotion = sum(odds)) %>% 
+  dplyr::group_by(C_Video) %>% 
+  dplyr::mutate(sum_video = sum(sum_emotion)) %>% 
+  dplyr::mutate(confidence_score = sum_emotion/sum_video) %>% 
+  dplyr::select(C_Video, emotion, ar_value = confidence_score)
+
+recognition_result_probabilities <- grid_data %>% 
+  dplyr::left_join(self_report_result, by = c(
+    "list_video" = "C_Video",
+    "list_emotion" = "sr_emotion"
+  )
+  ) %>% 
+  dplyr::left_join(human_recognition_score_all, by = c(
[…]

### 2019-04-15 ACII_2019_paper/ACII_2019_paper.Rmd
new version

--- 2019-04-05 ACII_2019_paper/ACII_2019_paper.Rmd
@@
 Then automatic emotion recognition tools based on facial expressions should be questioned.
+
 header-includes:
@@
   - \usepackage{tabu}
+  - \usepackage{wrapfig}
+  - \usepackage[none]{hyphenat}
+  - \floatplacement{figure}{H}
 
@@
 output: rticles::ieee_article
+csl: http://www.zotero.org/styles/ieee
+fig_caption: yes
+keep_tex: yes
+number_sections: yes
+includes:
+  in_header: preamble-latex.tex
 ---
 
+\begin{IEEEkeywords}
+Facial expression, self-report, human observer, automatic recognition.
+\end{IEEEkeywords}
+
 ```{r setup, include = FALSE}
+# options ----------------------------------------------------------------------
 knitr::opts_chunk$set(
@@
   error = FALSE, 
-  echo = FALSE,
-  fig.pos = 'H')
-#
+  echo = FALSE)
+
 set.seed(123)# Seed for random number generation
 options(scipen = 999)# Disable scientific number format
-#
-library(kableExtra)
+
+# packages ----------------------------------------------------------------------
+library(kableExtra) # Tables
 library(tidyverse) # Data wrangling & grammar of graphics
 library(here) # Locate files relative to project root
-library(caret)
-library(plotROC)
-library(pROC)
+library(caret) # Confusion matrices
+library(papaja) # Correlation results
 ```
@@
 
-Introduction
+I. Introduction
 =============
 
-With the development of commercial automatic facial expression recognition tools [see @dupre2018accuracy for a non-exhaustive list of available tools], industries and governments are gradually implementing this technology in order to track humans’ emotions in various scenarios (e.g., marketing, healthcare, automotive to name a few).
+With the development of commercial automatic facial expression recognition tools [see @dupre2018accuracy for a non-exhaustive list of available tools], industries and governments are gradually implementing this technology in order to track humans’ emotions in various scenarios (*e.g.*, marketing, healthcare, automotive to name a few).
 This technoloogy rests on the premise that facial expressions provide a direct access to individuals’ subjective feeling.
@@
 
-## The link between emotion and facial expression through the behaviorist approach
+## A. The Behaviorist Approach
 Based on the behaviorist approach initiated by Darwin in *The Expression of the Emotions in Man and Animals* [@darwin1872expression], facial expressions are conceived as a genuine displays of individuals’ inner emotional state.
 This hypothesis is used as a basis for the Basic Emotion Theory (BET) which states that a set of six emotions are universally displayed and are genetically hardwired not only in humans [@ekman1992argument] but also in different animal species [@de2019mama].
-According to this view, “when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including […] facial expressions.” [@ekman2007directed, p49].
+According to this view, *“when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including […] facial expressions.”* [@ekman2007directed, 49].
 To cope with critics, several amendments have been made to the BET, increasing the number of basic emotions from six to seven [@ekman1988universality] as well as adding the concept of “display rules” to explain cultural differences in the management of facial expressions [@ekman1987universals].
@@
 
-## The link between emotion and facial expression through the social constructivist approach
+## B. The Social Constructivist Approach
 Detractors of the Basic Emotion Theory consider emotion not as genetically hardwired but as a learnt association between a given situation and an appropriate response [@averill1980constructivist; @barrett2017emotions].
-For the tenants of the constructivist approach, emotions are “concepts” based on past experiences and which are “a collection of embodied, whole brain representations that predict what is about to happen in the sensory environment, what the best action is to deal with impending events, and their consequences for allostasis” [@barrett2017theory, p12].
+For the tenants of the constructivist approach, emotions are “concepts” based on past experiences and which are *“a collection of embodied, whole brain representations that predict what is about to happen in the sensory environment, what the best action is to deal with impending events, and their consequences for allostasis”* [@barrett2017theory, 12].
 Following this assumption, faces are best conceived as tools displaying signals in social interactions [@crivelli2018facial].
-These signals can convey individuals’ motivations and readiness [@frijda1997facial] or social messages [@fridlund1995human].
+These signals can convey individuals’ motivations and readiness [@frijda1997facial] or social messages [@fridlund1994human].
 Therefore, facial expressions are thought as behaviors which meaning is inferred by the observer.
@@
 
-## The recognition of emotional facial expressions
+## C. Emotional Facial Expression Recognition
 Regarding emotional facial expression (EFE) recognition, the behaviorist approach assertions and the constructivist approach claims lead to two divergent predictions.
@@
 Instead, emotions are mentally constructed by the perceiver and mental categories of emotions are needed to accurately categorize facial movements.
-Therefore, since there is no emotional prototypical face, one should expect human observers’ superior ability to accurately recognize EFE as compared to automatic EFE recognition tools.
+Therefore, since there is no emotional prototypical face, one should expect human observers’ superior ability to accurately recognize EFE as compared to automatic EFE recognition tools (Figure \ref{fig:models_img}).
+
+```{r models_img, fig.height=5, fig.cap="\\label{fig:models_img}Comparison of behaviorist and social constructivist approaches according to the expected differences in emotion recognition."}
+# knitr::include_graphics("C:/Users/dupred/Desktop/Projects/dynemo_feel_express/img/dynemo_img.png", auto_pdf = FALSE)
+raster::stack(here::here("./img/models.png")) %>%
+  raster::plotRGB()
+```
 
@@
 
-Method
+II. Method
 =============
@@
 
-## Emotion Elicitation
+## A. Emotion Elicitation
 
@@
 
-For the emotion elicitation experiment, `r nrow(self_report_data)` French participants (`r self_report_gender[self_report_gender$genre_c == "F","n"]` females, `r self_report_gender[self_report_gender$genre_c == "H","n"]` males, *M*age = `r self_report_age[,"Age_m"]`, *SD*age = `r self_report_age[,"Age_sd"]`) were recruited to perform one out of 11 emotion elicitation tasks designed to trigger a positive, a specific negative or a neutral emotional state.
+For the emotion elicitation experiment, `r nrow(self_report_data)` French participants (`r self_report_gender[self_report_gender$genre_c == "F","n"]` females, `r self_report_gender[self_report_gender$genre_c == "H","n"]` males, *M*~age~ = `r self_report_age[,"Age_m"]`, *SD*~age~ = `r self_report_age[,"Age_sd"]`) were recruited to perform one out of 11 emotion elicitation tasks designed to trigger a positive, a specific negative or a neutral emotional state.
 Encoders’ face were recorded using an hidden camera resulting `r nrow(self_report_data)` front facing 768x576 videos varying from `r min(metadata_video$ffprobe_duration)`s to `r max(metadata_video$ffprobe_duration)`s (Figure \ref{fig:dynemo_img}).
@@
 This picture is taken from a pilot with projects collaborators and all gave a consent for the publication of their photos and videos."}
-# knitr::include_graphics("C:/Users/dupred/Desktop/Projects/dynemo_feel_express/img/dynemo_img.png", auto_pdf = FALSE)
-raster::stack(here::here("./img/dynemo_img.jpg")) %>%
+raster::stack(here::here("./img/dynemo_img.png")) %>%
   raster::plotRGB()
@@
 
-After the emotion elicitation task the encoders rated their subjective feeling on Likert scales from 0 ("not at all") to \nolinebreak 5 \nolinebreak ("strongly") related to six “basic” emotion labels (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as six “non-basic” emotion labels (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
+After the emotion elicitation task the encoders rated their subjective feeling on Likert scales from 0 ("not at all") to \nolinebreak 5 \nolinebreak ("strongly") related to six “basic” emotion labels (*i.e.*, *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as six “non-basic” emotion labels (*i.e.*, *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
 
 Finally, a debriefing session was performed to ensure that encoders were not durably affected by the emotion elicitation task.
-The debriefing was also used to check that encoders did not guess the real purpose of the experiment (e.g.
-being filmed while they were performing an emotional elicitation task) to guarantee facial expressions’ genuineness.
+The debriefing was also used to check that encoders did not guess the real purpose of the experiment (*e.g.*, being filmed while they were performing an emotional elicitation task) to guarantee facial expressions’ genuineness.
 All encoders gave their agreement on their data and video to be processed for research purpose only.
 
-## Human Facial Expression Recognition
+## B. Human Facial Expression Recognition
 
@@
 First, the participants had to identify the emotional sequences by pressing the space bar of their keyboard to indicate the beginning and the end of the emotional sequences while watching the video.
-Second, the participants watched each emotional sequence previously identified and labeled the sequence using one of the 12 \nolinebreak emotions proposed including six “basic” emotion labels (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) and six "non-basic" emotion labels (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
+Second, the participants watched each emotional sequence previously identified and labeled the sequence using one of the 12 \nolinebreak emotions proposed including six “basic” emotion labels (*i.e.*, *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) and six "non-basic" emotion labels (*i.e.*, *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
 They also had the possibility to indicate that the sequence was expressing none of the proposed emotion labels.
@@
 
-## Automatic Facial Expression Recognition
+## C. Automatic Facial Expression Recognition
 
@@
 Affdex’s algorithm uses Histogram of Oriented Gradient (HOG) features and Support Vector Machine (SVM) classifiers in order to recognize facial expressions.
-For each video frame, Affdex identifies the probability $p_{video_{i}.label_{j}.t_{k}}$ from 0 to 100 (rescaled to 0 to 1 for the analysis) of the face as expressing one of the six “basic” emotion labels (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as additional psychological states such as *valence*, *engagement* or *contempt*, and facial features such as *cheek raise*, *eye widen* or *jaw drop*.
+For each video frame, Affdex identifies the probability $p_{video_{i}.label_{j}.t_{k}}$ from 0 to 100 (rescaled to 0 to 1 for the analysis) of the face as expressing one of the six “basic” emotion labels (*i.e.*, *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as additional psychological states such as *valence*, *engagement* or *contempt*, and facial features such as *cheek raise*, *eye widen* or *jaw drop*.
 
@@
 
-Results
+III. Results
 ============
@@
 
-## Correlation between self-report and human facial expression recognition
+```{r corr-matrix}
+list_video <- human_recognition_score$C_Video
+list_emotion <- c("disgust","fear","happiness","surprise","sadness","anger")
+
+grid_data <- expand.grid(list_video = list_video, list_emotion = list_emotion)
+
+self_report_result <- self_report_score %>%
+  dplyr::filter(sr_emotion != "undetermined") %>%
+  dplyr::mutate(sr_score = 1)
+
+human_recognition_result <- human_recognition_score %>%
+  dplyr::filter(hr_emotion != "undetermined") %>%
+  dplyr::mutate(hr_score = 1)
+
+automatic_recognition_result <- automatic_recognition_score %>%
+  dplyr::filter(ar_emotion != "undetermined") %>%
+  dplyr::mutate(ar_score = 1)
+
+recognition_result <- grid_data %>%
+  dplyr::left_join(self_report_result, by = c(
+    "list_video" = "C_Video",
+    "list_emotion" = "sr_emotion"
+    )
+  ) %>%
+  dplyr::left_join(human_recognition_result, by = c(
+    "list_video" = "C_Video",
+    "list_emotion" = "hr_emotion"
+    )
+  ) %>%
+  dplyr::left_join(automatic_recognition_result, by = c(
+    "list_video" = "C_Video",
+    "list_emotion" = "ar_emotion"
+    )
+  ) %>%
+  replace(is.na(.), 0)
+
+corr_sr_hr <- cor.test(
+  recognition_result$sr_score, 
+  recognition_result$hr_score, 
+  alternative = "two.sided"
+  ) %>% 
+  papaja::apa_print()
+
+corr_sr_ar <- cor.test(
+  recognition_result$sr_score, 
+  recognition_result$ar_score, 
+  alternative = "two.sided"
+  ) %>% 
+  papaja::apa_print()
+
+corr_hr_ar <- cor.test(
+  recognition_result$hr_score, 
+  recognition_result$ar_score, 
+  alternative = "two.sided"
+  ) %>% 
+  papaja::apa_print()
+```
+
+## A. Human Observers' Accuracy
 
@@
 
-Encoders’ subjective feeling is compared with human observers recognition in a confusion matrix (Figure \ref{fig:confusionMatrix_sr_hr}).
-Each emotion label used to describe encoders’ self-reported subjective feeling (i.e., the label rated with the highest value) is compared with the emotion labels which were rated with the highest score by human observers.
+The overall correlation of recognition and non-recognition between self-reported emotions and human observers recognition is significant but low (`r corr_sr_hr$full_result`).
+In order to identify differences according the emotional labels, encoders’ subjective feeling is compared with human observers recognition in a confusion matrix (Figure \ref{fig:confusionMatrix_sr_hr}).
 
-```{r confusionMatrix_sr_hr, fig.height=5, fig.cap="\\label{fig:confusionMatrix_sr_hr}Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the human observers."}
+```{r confusionMatrix_sr_hr, fig.height=6, fig.cap="\\label{fig:confusionMatrix_sr_hr}Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the human observers."}
 confusionMatrix_sr_hr_freq <- confusionMatrix_sr_hr$table %>% 
@@
 
-Results of the confusion matrix show a moderate agreement between the emotion felt by the encoder during the elicitation and the emotion recognized by the human observers (Accuracy \nolinebreak = \nolinebreak `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *disgust* (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_sr"]))` of the videos self-reported).
+Each emotion label used to describe encoders’ self-reported subjective feeling (*i.e.*, the label rated with the highest value) is compared with the emotion labels which were rated with the highest score by human observers.
+Results of the confusion matrix show a moderate agreement between the emotion felt by the encoder during the elicitation and the emotion recognized by the human observers (Accuracy \nolinebreak = \nolinebreak `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95% CI [`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *disgust* (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_sr"]))` of the videos self-reported).
+These results are far from those classicaly obtained in the literature for emotional facial expression recognition which ranges between 60% and 80% accuracy.
+However these results are mostly obtained with static (*i.e.,* pictures) and posed (*i.e.,* displayed by actors) facial expressions using only 6 emotional labels in a forced-choice paradigm.
+
+Interestingly human observers seem to recognize *surprise* expressed in videos where *anger*, *fear* *happiness* and *sadness* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "anger", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "fear", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "happiness", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "sadness", "prop_sr"]))` of the videos self-reported), and in a lower instance *happiness* was recognized in videos where *fear* and *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "fear", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
+
 Sensitivity, specificity, precision and F1 score for each emotion reveals that *happiness* has the highest coherence ratio whereas *sadness* has the lowest coherence ratio between true positives and false positives (Table \ref{table:confusionTable_sr_hr}).
-Interestingly human observers seem to recognize *surprise* expressed in videos where *anger*, *fear* *happiness* and *sadness* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "anger", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "fear", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "happiness", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "sadness", "prop_sr"]))` of the videos self-reported), and in a lower instance *happiness* was recognized in videos where *fear* and *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "fear", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
 
@@
   kableExtra::kable_styling(
+    latex_options = "HOLD_position",
     font_size = 8,
-    full_width = TRUE,
-    latex_options = c("hold_position")) %>%
+    full_width = TRUE 
+    ) %>%
    add_footnote("Note. \\textit{na.} values are produced when not enough data are available to compute accuracy indicators.", notation="none", escape = FALSE)
+```
 
-# confusionMatrix_sr_hr_table %>%
-#   papaja::apa_table(
-#     caption = table_nums("confusionTable_sr_hr"),
-#     row.names = FALSE,
-#     format = "word",
-#     escape = FALSE
-#     )
-```
+Accuracy metrics by emotional labels indicate a discrepancy in the ratio of true/false positives.
+Whereas *happiness* and *disgust* obtain the highest scores, *anger*, *surprise* and *sadness* have the lowest recognition ratio.
+The underlying effect of expression intensity may explain why *happiness* and *disgust* are easily recognized.
+*Anger* and *sadness* as non-socialy desirable emotion may be have been felt but not expressed.
 
@@
 
-## Correlation between self-report and automatic facial expression recognition
+## B. Automatic Classifier's Accuracy
 
@@
 
-Similarly to the previous analysis, a confusion matrix was used to compare encoders’ subjective feeling with the emotion label recognized by the automatic classifier (Figure \ref{fig:confusionMatrix_sr_ar}).
+Similarly to the previous analysis, the overall correlation of recognition and non-recognition revealed a significant but very low link between self-reported emotions and automatic \nolinebreak classifier's recognition (`r corr_sr_ar$full_result`).
+A confusion matrix was used to compare encoders’ subjective feeling with the emotion label recognized by the automatic classifier (Figure \ref{fig:confusionMatrix_sr_ar}).
 
-```{r confusionMatrix_sr_ar, fig.height=5, fig.cap="\\label{fig:confusionMatrix_sr_ar}Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the automatic classifier."}
+```{r confusionMatrix_sr_ar, fig.height=6, fig.cap="\\label{fig:confusionMatrix_sr_ar}Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the automatic classifier."}
 confusionMatrix_sr_ar_freq <- confusionMatrix_sr_ar$table %>% 
@@
 
-Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are somewhat similar to the ones with human observers (Table \ref{table:confusionTable_sr_ar}).
-Overall, a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy \nolinebreak = \nolinebreak `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_ar$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_ar$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_ar$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` of the video self-reported) is watched.
+Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are somewhat similar to the ones with human observers (Table \nolinebreak \ref{table:confusionTable_sr_ar}).
+Overall, a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy \nolinebreak = \nolinebreak `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`, 95% CI [`r round(confusionMatrix_sr_ar$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_ar$overall["AccuracyUpper"],2)`]; Kappa \nolinebreak = \nolinebreak `r round(confusionMatrix_sr_ar$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` of the video self-reported) is watched.
+
 Surprisingly the automatic classifier incorrectly recognized as *disgust* an important proportion of videos in which *anger*, *happiness* and *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "anger", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
@@
   kableExtra::kable_styling(
+    latex_options = c("HOLD_position"),
     font_size = 8,
-    full_width = TRUE,
-    latex_options = c("hold_position")) %>%
+    full_width = TRUE
+    ) %>%
    add_footnote("Note. \\textit{na.} values are produced when not enough data are available to compute accuracy indicators.", notation="none", escape = FALSE)
-
-# confusionMatrix_sr_ar_table %>%
-#   papaja::apa_table(
-#     caption = table_nums("confusionTable_sr_ar"),
-#     row.names = FALSE,
-#     format = "word",
-#     escape = FALSE
-#     )
 ```
@@
 
-## Comparison between human and automatic recognition
+## C. Comparison Between Human and Automatic Recognition
 
-As previously mentioned, human observers appear to be more accurate than the automatic classifier to recognize individuals’ subjective feeling (human observer Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`; automatic classifier Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`).
+As previously mentioned, human observers appear to be more accurate than the automatic classifier to recognize individuals’ subjective feeling (human observers Accuracy \nolinebreak = \nolinebreak`r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`; automatic classifier Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`; `r corr_hr_ar$full_result`).
 However, both make mistakes.
+
 A third confusion matrix is used to compare similarities (diagonal) and differences between human observers and automatic classifier in classifing the six emotion labels (Figure \nolinebreak \ref{fig:confusionMatrix_hr_ar}).
 
-```{r confusionMatrix_hr_ar, fig.height=5, fig.cap="\\label{fig:confusionMatrix_hr_ar}Proportion of emotion labels classified by human observers which are recognized by the automatif classifier."}
+```{r confusionMatrix_hr_ar, fig.height=6, fig.cap="\\label{fig:confusionMatrix_hr_ar}Proportion of emotion labels classified by human observers which are recognized by the automatif classifier."}
 comparison_hr_ar <- dplyr::inner_join(human_recognition_score,automatic_recognition_score, by = "C_Video") %>% 
@@
 Except for *happiness* and *disgust* (respectively `r scales::percent(as.numeric(confusionMatrix_hr_ar_freq[confusionMatrix_hr_ar_freq$ar_emotion == "happiness" & confusionMatrix_hr_ar_freq$hr_emotion == "happiness", "prop_hr"]))` and `r scales::percent(as.numeric(confusionMatrix_hr_ar_freq[confusionMatrix_hr_ar_freq$ar_emotion == "disgust" & confusionMatrix_hr_ar_freq$hr_emotion == "disgust", "prop_hr"]))` of common labelling), there is no clear common pattern.
-Moreover, the automatic classifier has a tendency to label as *disgust* videos labeled as *sadness* by human observers and as *happiness* videos labaled as *fear* by human observers
+Moreover, the automatic classifier has a tendency to label as *disgust* videos labeled as *sadness* by human observers and as *happiness* videos labeled as *fear* by human observers
 
-Discussion and Conclusion
+IV. Discussion and Conclusion
 ============
@@
 
-Such results support the hypothesis advanced by some authors of a low emotion–expression coherence [e.g., @kappas2003facial].
+Such results support the hypothesis advanced by some authors of a low emotion–expression coherence [*e.g.*, @kappas2003facial].
 In many instances, facial displays are not associated with a concordant emotional state, even any emotional state at all [@bonanno2004brief; @fernandez2013emotion].
 More and more evidences are showing that facial expressions are in reality not expressing emotions [@mckeown2013analogical].
-As well as other nonverbal behaviors, facial movements are not only assumed to be determined by emotion but also by various other causes, such as psychological states (e.g., \nolinebreak motivations or pain), to say nothing of social context and sociocultural norms [such as “display rules”, @ekman1987universals].
+As well as other nonverbal behaviors, facial movements are not only assumed to be determined by emotion but also by various other causes, such as psychological states (*e.g.*, \nolinebreak motivations or pain), to say nothing of social context and sociocultural norms [such as “display rules”, @ekman1987universals].
 This multiple determination excludes any possibility of drawing an inference from facial activity on the underlying psychological state (emotional or other).
+
 Beyond the present observations showing a weak coherence between subjective feelings and spontaneous facial expressions, this study provides some light on the controversy between the behaviorist and the constructivist approaches.
@@
 As a consequence, one can predict from the first line of thinking that individuals’ emotional subjective feeling should be correlated to the recognition of facial expressions from both human observers and automatic classifiers whereas if emotions are social constructs, as stated by the second line of thinking, human observer’s should be better at perceiving emotions in face than automatic classifiers.
+
 Present results plead in favor of the latter stance.
@@
 Moreover the procedure used for human observation can also be open to dispute.
-Instead of asking the human annotators to provide a unique label, a more subtle approach was chosen to mimic results provided by the automatic classifier .
-Whereas this paradigm is longer and more complicated, it can lead to more robust results in reducing the forced-choice biais [@russell1993forced].
+Instead of asking the human annotators to provide a unique label, a more subtle approach was chosen to mimic results provided by the automatic classifier.
+Whereas this paradigm is longer and more complicated, it can lead to more robust results in reducing the forced-choice bais [@russell1993forced].
 However this procedure can also reduce the human observers accuracy.
 In this regard, the results of the human observation could have been more ambiguous because it is not the natural way that people are inferring meaning from facial expressions.
-An alternative explanation relies in reducing the recognition biais involved in the classic recognition paradigm.
+An alternative explanation relies in reducing the recognition bais involved in the classic recognition paradigm.
 Classic forced-choice paradigms obtain artificial high results, thus by using a more evolved approach observers’ accuracy may have been lowered.
@@
 Advancements in identifying “non-basic” emotion labels [@mcduff2016discovering] as well as non-prototypical facial expression have been made in the developpement of automatic facial expression recognition tools.
-However this result suggests that automatic facial expression recognition tools should merely evaluate facial morphology features such as action units [e.g., OpenFace presented in @baltruvsaitis2016openface] rather than inferring supposedly emotional or affective states.
+However this result suggests that automatic facial expression recognition tools should merely evaluate facial morphology features such as action units [*e.g.*, OpenFace presented in @baltruvsaitis2016openface] rather than inferring supposedly emotional or affective states.
 
-Acknowledgment {#acknowledgment}
+V. Acknowledgment {#acknowledgment}
 ==============
@@
 
-References {#references .numbered}
+VI. References {#references .numbered}
 ==========

### 2019-04-15 ACII_2019_paper/ACII_2019_paper.tex
new version

@@
 \usepackage{wrapfig}
+\usepackage[none]{hyphenat}
 \floatplacement{figure}{H}
-\floatplacement{table}{H}
 
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\begin{IEEEkeywords}
+Facial expression, self-report, human observer, automatic recognition.
+\end{IEEEkeywords}
+
+\hypertarget{i.-introduction}{%
+\section{I. Introduction}\label{i.-introduction}}
 
 With the development of commercial automatic facial expression
-recognition tools (see Dupré et al. 2018 for a non-exhaustive list of
-available tools), industries and governments are gradually implementing
-this technology in order to track humans' emotions in various scenarios
-(e.g., marketing, healthcare, automotive to name a few). This
-technoloogy rests on the premise that facial expressions provide a
-direct access to individuals' subjective feeling. Even if this premise
-is central to the modern mainstream approach of human emotion, recent
-research in affective science are challenging it. Once the two competing
-approaches briefly described as well as the predictions they
-respectively entail, an experiment testing these hypotheses will be
-presented and its results analyzed in order to provide empirical
-evidence to contribute to answer the question.
-
-\hypertarget{the-link-between-emotion-and-facial-expression-through-the-behaviorist-approach}{%
-\subsection{The link between emotion and facial expression through the
-behaviorist
-approach}\label{the-link-between-emotion-and-facial-expression-through-the-behaviorist-approach}}
+recognition tools {[}1{]}, industries and governments are gradually
+implementing this technology in order to track humans' emotions in
+various scenarios (\emph{e.g.}, marketing, healthcare, automotive to
+name a few). This technoloogy rests on the premise that facial
+expressions provide a direct access to individuals' subjective feeling.
+Even if this premise is central to the modern mainstream approach of
+human emotion, recent research in affective science are challenging it.
+Once the two competing approaches briefly described as well as the
+predictions they respectively entail, an experiment testing these
+hypotheses will be presented and its results analyzed in order to
+provide empirical evidence to contribute to answer the question.
+
+\hypertarget{a.-the-behaviorist-approach}{%
+\subsection{A. The Behaviorist
+Approach}\label{a.-the-behaviorist-approach}}
 
 Based on the behaviorist approach initiated by Darwin in \emph{The
-Expression of the Emotions in Man and Animals} (Darwin 1872), facial
+Expression of the Emotions in Man and Animals} {[}2{]}, facial
 expressions are conceived as a genuine displays of individuals' inner
@@
 universally displayed and are genetically hardwired not only in humans
-(Ekman 1992) but also in different animal species (Waal 2019). According
-to this view, ``when emotions are aroused by perception of a social
-event, a set of central commands produce patterned emotion-specific
-changes in multiple systems, including {[}\ldots{}{]} facial
-expressions.'' (Ekman 2007, p49). To cope with critics, several
-amendments have been made to the BET, increasing the number of basic
-emotions from six to seven (Ekman and Heider 1988) as well as adding the
-concept of ``display rules'' to explain cultural differences in the
-management of facial expressions (Ekman et al. 1987).
+{[}3{]} but also in different animal species {[}4{]}. According to this
+view, \emph{``when emotions are aroused by perception of a social event,
+a set of central commands produce patterned emotion-specific changes in
+multiple systems, including {[}\ldots{}{]} facial expressions.''} {[}5,
+p. 49{]}. To cope with critics, several amendments have been made to the
+BET, increasing the number of basic emotions from six to seven {[}6{]}
+as well as adding the concept of ``display rules'' to explain cultural
+differences in the management of facial expressions {[}7{]}.
 
@@
 individuals can express emotions without feeling them, in instances in
-which display rules cannot be called upon (Kraut and Johnston 1979;
-Durán, Reisenzein, and Fernández-Dols 2017). These evidences have led to
-an alternative conception, notably the social constructivist approach.
-
-\hypertarget{the-link-between-emotion-and-facial-expression-through-the-social-constructivist-approach}{%
-\subsection{The link between emotion and facial expression through the
-social constructivist
-approach}\label{the-link-between-emotion-and-facial-expression-through-the-social-constructivist-approach}}
+which display rules cannot be called upon {[}8{]}, {[}9{]}. These
+evidences have led to an alternative conception, notably the social
+constructivist approach.
+
+\hypertarget{b.-the-social-constructivist-approach}{%
+\subsection{B. The Social Constructivist
+Approach}\label{b.-the-social-constructivist-approach}}
 
@@
 genetically hardwired but as a learnt association between a given
-situation and an appropriate response (Averill 1980; Barrett 2017a). For
-the tenants of the constructivist approach, emotions are ``concepts''
-based on past experiences and which are ``a collection of embodied,
+situation and an appropriate response {[}10{]}, {[}11{]}. For the
+tenants of the constructivist approach, emotions are ``concepts'' based
+on past experiences and which are \emph{``a collection of embodied,
 whole brain representations that predict what is about to happen in the
 sensory environment, what the best action is to deal with impending
-events, and their consequences for allostasis'' (Barrett 2017b, p12).
+events, and their consequences for allostasis''} {[}12, p. 12{]}.
 Following this assumption, faces are best conceived as tools displaying
-signals in social interactions (Crivelli and Fridlund 2018). These
-signals can convey individuals' motivations and readiness (Frijda and
-Tcherkassof 1997) or social messages (Fridlund 1994). Therefore, facial
-expressions are thought as behaviors which meaning is inferred by the
-observer. Findings support this observer dependence (Lindquist and
-Gendron 2013; Niedenthal et al. 2017). They show that to make meaning of
-another person's facial behavior, the perceiver relies in particular on
-her/his knowledge about emotion categories.
-
-\hypertarget{the-recognition-of-emotional-facial-expressions}{%
-\subsection{The recognition of emotional facial
-expressions}\label{the-recognition-of-emotional-facial-expressions}}
+signals in social interactions {[}13{]}. These signals can convey
+individuals' motivations and readiness {[}14{]} or social messages
+{[}15{]}. Therefore, facial expressions are thought as behaviors which
+meaning is inferred by the observer. Findings support this observer
+dependence {[}16{]}, {[}17{]}. They show that to make meaning of another
+person's facial behavior, the perceiver relies in particular on her/his
+knowledge about emotion categories.
+
+\hypertarget{c.-emotional-facial-expression-recognition}{%
+\subsection{C. Emotional Facial Expression
+Recognition}\label{c.-emotional-facial-expression-recognition}}
 
@@
 episode, facial muscle movements are not linked in a one-to-one manner
-to a specific discrete emotional experience (Doyle and Lindquist 2017,
-p418). Instead, emotions are mentally constructed by the perceiver and
-mental categories of emotions are needed to accurately categorize facial
-movements. Therefore, since there is no emotional prototypical face, one
-should expect human observers' superior ability to accurately recognize
-EFE as compared to automatic EFE recognition tools.
+to a specific discrete emotional experience {[}18, p. p418{]}. Instead,
+emotions are mentally constructed by the perceiver and mental categories
+of emotions are needed to accurately categorize facial movements.
+Therefore, since there is no emotional prototypical face, one should
+expect human observers' superior ability to accurately recognize EFE as
+compared to automatic EFE recognition tools (Figure
+\ref{fig:models_img}).
+
+\begin{figure}
+\centering
+\includegraphics{ACII_2019_paper_files/figure-latex/models_img-1.pdf}
+\caption{\label{fig:models_img}Comparison of behaviorist and social
+constructivist approaches according to the expected differences in
+emotion recognition.}
+\end{figure}
 
@@
 
-\hypertarget{method}{%
-\section{Method}\label{method}}
+\hypertarget{ii.-method}{%
+\section{II. Method}\label{ii.-method}}
 
@@
 
-\hypertarget{emotion-elicitation}{%
-\subsection{Emotion Elicitation}\label{emotion-elicitation}}
+\hypertarget{a.-emotion-elicitation}{%
+\subsection{A. Emotion Elicitation}\label{a.-emotion-elicitation}}
 
 For the emotion elicitation experiment, 358 French participants (182
-females, 176 males, \emph{M}age = 47.9, \emph{SD}age = 9.2) were
-recruited to perform one out of 11 emotion elicitation tasks designed to
-trigger a positive, a specific negative or a neutral emotional state.
-Encoders' face were recorded using an hidden camera resulting 358 front
-facing 768x576 videos varying from 1s to 1479s (Figure
-\ref{fig:dynemo_img}). These recordings form the DynEmo database (see
-Tcherkassof et al. 2013 for a full description of tasks and procedure).
+females, 176 males, \emph{M}\textsubscript{age} = 47.9,
+\emph{SD}\textsubscript{age} = 9.2) were recruited to perform one out of
+11 emotion elicitation tasks designed to trigger a positive, a specific
+negative or a neutral emotional state. Encoders' face were recorded
+using an hidden camera resulting 358 front facing 768x576 videos varying
+from 1s to 1479s (Figure \ref{fig:dynemo_img}). These recordings form
+the DynEmo database {[}19{]}.
 
@@
 \nolinebreak (``strongly'') related to six ``basic'' emotion labels
-(i.e., \emph{anger}, \emph{disgust}, \emph{fear}, \emph{happiness},
-\emph{surprise} and \emph{sadness}) as well as six ``non-basic'' emotion
-labels (i.e., \emph{pride}, \emph{curiosity}, \emph{boredom},
-\emph{shame}, \emph{humiliation}, and \emph{disappointment}).
+(\emph{i.e.}, \emph{anger}, \emph{disgust}, \emph{fear},
+\emph{happiness}, \emph{surprise} and \emph{sadness}) as well as six
+``non-basic'' emotion labels (\emph{i.e.}, \emph{pride},
+\emph{curiosity}, \emph{boredom}, \emph{shame}, \emph{humiliation}, and
+\emph{disappointment}).
 
@@
 also used to check that encoders did not guess the real purpose of the
-experiment (e.g.~being filmed while they were performing an emotional
-elicitation task) to guarantee facial expressions' genuineness. All
-encoders gave their agreement on their data and video to be processed
-for research purpose only.
-
-\hypertarget{human-facial-expression-recognition}{%
-\subsection{Human Facial Expression
-Recognition}\label{human-facial-expression-recognition}}
+experiment (\emph{e.g.}, being filmed while they were performing an
+emotional elicitation task) to guarantee facial expressions'
+genuineness. All encoders gave their agreement on their data and video
+to be processed for research purpose only.
+
+\hypertarget{b.-human-facial-expression-recognition}{%
+\subsection{B. Human Facial Expression
+Recognition}\label{b.-human-facial-expression-recognition}}
 
@@
 \emph{Oudjat}, a software for designing video annotation experiments
-(Dupré et al. 2015). For each video, the annotation procedure followed
-two steps. First, the participants had to identify the emotional
-sequences by pressing the space bar of their keyboard to indicate the
-beginning and the end of the emotional sequences while watching the
-video. Second, the participants watched each emotional sequence
-previously identified and labeled the sequence using one of the 12
-\nolinebreak emotions proposed including six ``basic'' emotion labels
-(i.e., \emph{anger}, \emph{disgust}, \emph{fear}, \emph{happiness},
-\emph{surprise} and \emph{sadness}) and six ``non-basic'' emotion labels
-(i.e., \emph{pride}, \emph{curiosity}, \emph{boredom}, \emph{shame},
+{[}20{]}. For each video, the annotation procedure followed two steps.
+First, the participants had to identify the emotional sequences by
+pressing the space bar of their keyboard to indicate the beginning and
+the end of the emotional sequences while watching the video. Second, the
+participants watched each emotional sequence previously identified and
+labeled the sequence using one of the 12 \nolinebreak emotions proposed
+including six ``basic'' emotion labels (\emph{i.e.}, \emph{anger},
+\emph{disgust}, \emph{fear}, \emph{happiness}, \emph{surprise} and
+\emph{sadness}) and six ``non-basic'' emotion labels (\emph{i.e.},
+\emph{pride}, \emph{curiosity}, \emph{boredom}, \emph{shame},
 \emph{humiliation}, and \emph{disappointment}). They also had the
@@
 
-\hypertarget{automatic-facial-expression-recognition}{%
-\subsection{Automatic Facial Expression
-Recognition}\label{automatic-facial-expression-recognition}}
+\hypertarget{c.-automatic-facial-expression-recognition}{%
+\subsection{C. Automatic Facial Expression
+Recognition}\label{c.-automatic-facial-expression-recognition}}
 
@@
 distributed by Affectiva is a spin-off company resulting from the
-research activities of MIT media lab created in 2009 (McDuff et al.
-2016). Affdex's algorithm uses Histogram of Oriented Gradient (HOG)
-features and Support Vector Machine (SVM) classifiers in order to
-recognize facial expressions. For each video frame, Affdex identifies
-the probability \(p_{video_{i}.label_{j}.t_{k}}\) from 0 to 100
-(rescaled to 0 to 1 for the analysis) of the face as expressing one of
-the six ``basic'' emotion labels (i.e., \emph{anger}, \emph{disgust},
-\emph{fear}, \emph{happiness}, \emph{surprise} and \emph{sadness}) as
-well as additional psychological states such as \emph{valence},
+research activities of MIT media lab created in 2009 {[}21{]}. Affdex's
+algorithm uses Histogram of Oriented Gradient (HOG) features and Support
+Vector Machine (SVM) classifiers in order to recognize facial
+expressions. For each video frame, Affdex identifies the probability
+\(p_{video_{i}.label_{j}.t_{k}}\) from 0 to 100 (rescaled to 0 to 1 for
+the analysis) of the face as expressing one of the six ``basic'' emotion
+labels (\emph{i.e.}, \emph{anger}, \emph{disgust}, \emph{fear},
+\emph{happiness}, \emph{surprise} and \emph{sadness}) as well as
+additional psychological states such as \emph{valence},
 \emph{engagement} or \emph{contempt}, and facial features such as
@@
 probability for each label by frame was converted into odd ratio by
-label (Dente et al. 2017). The highest sum of each odd ratio time-series
-defines the label recognized by the automatic classifier (EQ\ref{eq:2}
-for human recognition and EQ\ref{eq:3} for automatic recognition).
+label {[}22{]}. The highest sum of each odd ratio time-series defines
+the label recognized by the automatic classifier (EQ\ref{eq:2} for human
+recognition and EQ\ref{eq:3} for automatic recognition).
 \begin{equation}
@@
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
+\hypertarget{iii.-results}{%
+\section{III. Results}\label{iii.-results}}
 
@@
 
-\hypertarget{self-reported-emotions-recognized-by-human-observers}{%
-\subsection{Self-reported emotions recognized by human
-observers}\label{self-reported-emotions-recognized-by-human-observers}}
-
-Encoders' subjective feeling is compared with human observers
+\hypertarget{a.-human-observers-accuracy}{%
+\subsection{A. Human Observers'
+Accuracy}\label{a.-human-observers-accuracy}}
+
+The overall correlation of recognition and non-recognition between
+self-reported emotions and human observers recognition is significant
+but low (\(r = .24\), 95\% CI \([.19\), \(.29]\), \(t(1384) = 9.21\),
+\(p < .001\)). In order to identify differences according the emotional
+labels, encoders' subjective feeling is compared with human observers
 recognition in a confusion matrix (Figure
-\ref{fig:confusionMatrix_sr_hr}). Each emotion label used to describe
-encoders' self-reported subjective feeling (i.e., the label rated with
-the highest value) is compared with the emotion labels which were rated
-with the highest score by human observers.
+\ref{fig:confusionMatrix_sr_hr}).
 
@@
 
-Results of the confusion matrix show a moderate agreement between the
-emotion felt by the encoder during the elicitation and the emotion
-recognized by the human observers (Accuracy \nolinebreak =
-\nolinebreak 0.43, 95\%CI{[}0.35,0.52{]}; Kappa = 0.19) except for
-\emph{disgust} (100\% of the videos self-reported). Sensitivity,
-specificity, precision and F1 score for each emotion reveals that
-\emph{happiness} has the highest coherence ratio whereas \emph{sadness}
-has the lowest coherence ratio between true positives and false
-positives (Table \ref{table:confusionTable_sr_hr}). Interestingly human
-observers seem to recognize \emph{surprise} expressed in videos where
-\emph{anger}, \emph{fear} \emph{happiness} and \emph{sadness} was the
-highest self-reported emotion (respectively 28.6\%, 33.3\%, 34.7\% and
-33.3\% of the videos self-reported), and in a lower instance
-\emph{happiness} was recognized in videos where \emph{fear} and
-\emph{surprise} was the highest self-reported emotion (respectively
-50.0\% and 25.6\% of the videos self-reported).
+Each emotion label used to describe encoders' self-reported subjective
+feeling (\emph{i.e.}, the label rated with the highest value) is
+compared with the emotion labels which were rated with the highest score
+by human observers. Results of the confusion matrix show a moderate
+agreement between the emotion felt by the encoder during the elicitation
+and the emotion recognized by the human observers (Accuracy
+\nolinebreak = \nolinebreak 0.43, 95\% CI {[}0.35,0.52{]}; Kappa = 0.19)
+except for \emph{disgust} (100\% of the videos self-reported). These
+results are far from those classicaly obtained in the literature for
+emotional facial expression recognition which ranges between 60\% and
+80\% accuracy. However these results are mostly obtained with static
+(\emph{i.e.,} pictures) and posed (\emph{i.e.,} displayed by actors)
+facial expressions using only 6 emotional labels in a forced-choice
+paradigm.
+
+Interestingly human observers seem to recognize \emph{surprise}
+expressed in videos where \emph{anger}, \emph{fear} \emph{happiness} and
+\emph{sadness} was the highest self-reported emotion (respectively
+28.6\%, 33.3\%, 34.7\% and 33.3\% of the videos self-reported), and in a
+lower instance \emph{happiness} was recognized in videos where
+\emph{fear} and \emph{surprise} was the highest self-reported emotion
+(respectively 50.0\% and 25.6\% of the videos self-reported).
+
+Sensitivity, specificity, precision and F1 score for each emotion
+reveals that \emph{happiness} has the highest coherence ratio whereas
+\emph{sadness} has the lowest coherence ratio between true positives and
+false positives (Table \ref{table:confusionTable_sr_hr}).
 
@@
 
+Accuracy metrics by emotional labels indicate a discrepancy in the ratio
+of true/false positives. Whereas \emph{happiness} and \emph{disgust}
+obtain the highest scores, \emph{anger}, \emph{surprise} and
+\emph{sadness} have the lowest recognition ratio. The underlying effect
+of expression intensity may explain why \emph{happiness} and
+\emph{disgust} are easily recognized. \emph{Anger} and \emph{sadness} as
+non-socialy desirable emotion may be have been felt but not expressed.
+
 However, self-reports show a significant proportion of undetermined
@@
 
-\hypertarget{self-reported-emotions-recognized-by-the-automatic-classifier}{%
-\subsection{Self-reported emotions recognized by the automatic
-classifier}\label{self-reported-emotions-recognized-by-the-automatic-classifier}}
-
-Similarly to the previous analysis, a confusion matrix was used to
+\hypertarget{b.-automatic-classifiers-accuracy}{%
+\subsection{B. Automatic Classifier's
+Accuracy}\label{b.-automatic-classifiers-accuracy}}
+
+Similarly to the previous analysis, the overall correlation of
+recognition and non-recognition revealed a significant but very low link
+between self-reported emotions and automatic \nolinebreak classifier's
+recognition (\(r = .12\), 95\% CI \([.07\), \(.17]\),
+\(t(1384) = 4.50\), \(p < .001\)). A confusion matrix was used to
 compare encoders' subjective feeling with the emotion label recognized
@@
 recognized by the automatic classifier are somewhat similar to the ones
-with human observers (Table \ref{table:confusionTable_sr_ar}). Overall,
-a low agreement between emotion self-reported and emotion recognized by
-the automatic classifier (Accuracy \nolinebreak = \nolinebreak 0.3,
-95\%CI{[}0.22,0.38{]}; Kappa = 0.07) except for \emph{happiness} (44.4\%
-of the video self-reported) is watched. Surprisingly the automatic
-classifier incorrectly recognized as \emph{disgust} an important
-proportion of videos in which \emph{anger}, \emph{happiness} and
-\emph{surprise} was the highest self-reported emotion (respectively
-42.9\%, 25.0\% and 32.6\% of the videos self-reported). In parallel, the
-automatic classifier recognized as \emph{happiness} videos in which
-\emph{fear} and \emph{surprise} was the highest self-reported emotion
-(respectively 66.7\% and 32.6\% of the videos self-reported).
-
-\begin{table}[!h]
+with human observers (Table
+\nolinebreak \ref{table:confusionTable_sr_ar}). Overall, a low agreement
+between emotion self-reported and emotion recognized by the automatic
+classifier (Accuracy \nolinebreak = \nolinebreak 0.3, 95\% CI
+{[}0.22,0.38{]}; Kappa \nolinebreak = \nolinebreak 0.07) except for
+\emph{happiness} (44.4\% of the video self-reported) is watched.
+
+Surprisingly the automatic classifier incorrectly recognized as
+\emph{disgust} an important proportion of videos in which \emph{anger},
+\emph{happiness} and \emph{surprise} was the highest self-reported
+emotion (respectively 42.9\%, 25.0\% and 32.6\% of the videos
+self-reported). In parallel, the automatic classifier recognized as
+\emph{happiness} videos in which \emph{fear} and \emph{surprise} was the
+highest self-reported emotion (respectively 66.7\% and 32.6\% of the
+videos self-reported).
+
+\begin{table}[H]
 
@@
 
-\hypertarget{comparison-between-human-and-automatic-recognition}{%
-\subsection{Comparison between human and automatic
-recognition}\label{comparison-between-human-and-automatic-recognition}}
+\hypertarget{c.-comparison-between-human-and-automatic-recognition}{%
+\subsection{C. Comparison Between Human and Automatic
+Recognition}\label{c.-comparison-between-human-and-automatic-recognition}}
 
@@
 the automatic classifier to recognize individuals' subjective feeling
-(human observer Accuracy = 0.43; automatic classifier Accuracy = 0.3).
-However, both make mistakes. A third confusion matrix is used to compare
-similarities (diagonal) and differences between human observers and
-automatic classifier in classifing the six emotion labels (Figure
+(human observers Accuracy \nolinebreak = \nolinebreak0.43; automatic
+classifier Accuracy = 0.3; \(r = .22\), 95\% CI \([.17\), \(.27]\),
+\(t(1384) = 8.31\), \(p < .001\)). However, both make mistakes.
+
+A third confusion matrix is used to compare similarities (diagonal) and
+differences between human observers and automatic classifier in
+classifing the six emotion labels (Figure
 \nolinebreak \ref{fig:confusionMatrix_hr_ar}).
@@
 
-\hypertarget{discussion-and-conclusion}{%
-\section{Discussion and Conclusion}\label{discussion-and-conclusion}}
+\hypertarget{iv.-discussion-and-conclusion}{%
+\section{IV. Discussion and
+Conclusion}\label{iv.-discussion-and-conclusion}}
 
@@
 Such results support the hypothesis advanced by some authors of a low
-emotion--expression coherence (e.g., Kappas 2003). In many instances,
-facial displays are not associated with a concordant emotional state,
-even any emotional state at all (Bonanno and Keltner 2004;
-Fernández-Dols and Crivelli 2013). More and more evidences are showing
-that facial expressions are in reality not expressing emotions (McKeown
-2013). As well as other nonverbal behaviors, facial movements are not
+emotion--expression coherence {[}23{]}. In many instances, facial
+displays are not associated with a concordant emotional state, even any
+emotional state at all {[}24{]}, {[}25{]}. More and more evidences are
+showing that facial expressions are in reality not expressing emotions
+{[}26{]}. As well as other nonverbal behaviors, facial movements are not
 only assumed to be determined by emotion but also by various other
-causes, such as psychological states (e.g., \nolinebreak motivations or
-pain), to say nothing of social context and sociocultural norms (such as
-``display rules'', Ekman et al. 1987). This multiple determination
-excludes any possibility of drawing an inference from facial activity on
-the underlying psychological state (emotional or other). Beyond the
-present observations showing a weak coherence between subjective
-feelings and spontaneous facial expressions, this study provides some
-light on the controversy between the behaviorist and the constructivist
-approaches. The first of those two lines of thinking (Ekman 1992)
-assumes that expressions of emotion are brief and coherent patterns of
-facial muscle movements that co-vary with discrete subjective
-experiences. Emotions and their related prototypical facial behavior are
-universal because they are considered as innate mechanisms allowing
-individuals to respond adaptively to evolutionary significant events
-(threats, opportunities\ldots{}) encountered in the environment. Instead
-of viewing emotions as natural kinds (Barrett 2017b), the constructivist
-approach supposes that emotions are social constructions. The specific
-emotion categories identified on the face of another person are the
-result of categorization. In other words, the emotions that are
+causes, such as psychological states (\emph{e.g.},
+\nolinebreak motivations or pain), to say nothing of social context and
+sociocultural norms {[}7{]}. This multiple determination excludes any
+possibility of drawing an inference from facial activity on the
+underlying psychological state (emotional or other).
+
+Beyond the present observations showing a weak coherence between
+subjective feelings and spontaneous facial expressions, this study
+provides some light on the controversy between the behaviorist and the
+constructivist approaches. The first of those two lines of thinking
+{[}3{]} assumes that expressions of emotion are brief and coherent
+patterns of facial muscle movements that co-vary with discrete
+subjective experiences. Emotions and their related prototypical facial
+behavior are universal because they are considered as innate mechanisms
+allowing individuals to respond adaptively to evolutionary significant
+events (threats, opportunities\ldots{}) encountered in the environment.
+Instead of viewing emotions as natural kinds {[}12{]}, the
+constructivist approach supposes that emotions are social constructions.
+The specific emotion categories identified on the face of another person
+are the result of categorization. In other words, the emotions that are
 recognized by the observer are constructed in her/his mind, primarily
@@
 of thinking, human observer's should be better at perceiving emotions in
-face than automatic classifiers. Present results plead in favor of the
-latter stance. They show that human observers are more accurate than the
-automatic recognition tool to identify individuals' subjective feeling
-on the basis of their face. Moreover, mistakes made by human observers
-look less arbitrary to the ones made by the classifier. For instance,
-even if mix-up between disgust and anger is sometimes reported in
-recognition studies, confusions such as the ones produced by the
-classifier have never been noted for human observers. A possible
-explanation is that human observers are assessing dynamic expressive
-sequences. They have access to the full video which provides a kinetic
-context whereas the automatic classifier only assesses the videos frame
-by frame. Further research is needed to shed light on this issue.
+face than automatic classifiers.
+
+Present results plead in favor of the latter stance. They show that
+human observers are more accurate than the automatic recognition tool to
+identify individuals' subjective feeling on the basis of their face.
+Moreover, mistakes made by human observers look less arbitrary to the
+ones made by the classifier. For instance, even if mix-up between
+disgust and anger is sometimes reported in recognition studies,
+confusions such as the ones produced by the classifier have never been
+noted for human observers. A possible explanation is that human
+observers are assessing dynamic expressive sequences. They have access
+to the full video which provides a kinetic context whereas the automatic
+classifier only assesses the videos frame by frame. Further research is
+needed to shed light on this issue.
 
@@
 subtle approach was chosen to mimic results provided by the automatic
-classifier . Whereas this paradigm is longer and more complicated, it
-can lead to more robust results in reducing the forced-choice bais
-(Russell 1993). However this procedure can also reduce the human
-observers accuracy. In this regard, the results of the human observation
-could have been more ambiguous because it is not the natural way that
-people are inferring meaning from facial expressions. An alternative
-explanation relies in reducing the recognition bais involved in the
-classic recognition paradigm. Classic forced-choice paradigms obtain
-artificial high results, thus by using a more evolved approach
-observers' accuracy may have been lowered.
+classifier. Whereas this paradigm is longer and more complicated, it can
+lead to more robust results in reducing the forced-choice bais {[}27{]}.
+However this procedure can also reduce the human observers accuracy. In
+this regard, the results of the human observation could have been more
+ambiguous because it is not the natural way that people are inferring
+meaning from facial expressions. An alternative explanation relies in
+reducing the recognition bais involved in the classic recognition
+paradigm. Classic forced-choice paradigms obtain artificial high
+results, thus by using a more evolved approach observers' accuracy may
+have been lowered.
 
@@
 non-prototypical facial expressions. Advancements in identifying
-``non-basic'' emotion labels (McDuff 2016) as well as non-prototypical
-facial expression have been made in the developpement of automatic
-facial expression recognition tools. However this result suggests that
+``non-basic'' emotion labels {[}28{]} as well as non-prototypical facial
+expression have been made in the developpement of automatic facial
+expression recognition tools. However this result suggests that
 automatic facial expression recognition tools should merely evaluate
-facial morphology features such as action units (e.g., OpenFace
-presented in Baltrušaitis, Robinson, and Morency 2016) rather than
+facial morphology features such as action units {[}29{]} rather than
 inferring supposedly emotional or affective states.
@@
 \hypertarget{acknowledgment}{%
-\section{Acknowledgment}\label{acknowledgment}}
+\section{V. Acknowledgment}\label{acknowledgment}}
 
@@
 \hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
+\section*{VI. References}\label{references}}
+\addcontentsline{toc}{section}{VI. References}
 
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-dupre2018accuracy}{}%
+{[}1{]} D. Dupré, N. Andelic, G. Morrison, and G. McKeown, ``Accuracy of
+three commercial automatic emotion recognition systems across different
+individuals and their facial expressions,'' in \emph{Proceedings of the
+international conference on pervasive computing and communications},
+2018, pp. 627--632.
+
+\leavevmode\hypertarget{ref-darwin1872expression}{}%
+{[}2{]} C. Darwin, \emph{The expression of the emotions in man and
+animals}. London, UK: John Murray, 1872.
+
+\leavevmode\hypertarget{ref-ekman1992argument}{}%
+{[}3{]} P. Ekman, ``An argument for basic emotions,'' \emph{Cognition \&
+emotion}, vol. 6, nos. 3-4, pp. 169--200, 1992.
+
+\leavevmode\hypertarget{ref-de2019mama}{}%
+{[}4{]} F. de Waal, \emph{Mama's last hug: Animal emotions and what they
+teach us about ourselves}. London, UK: Granta Books, 2019.
+
+\leavevmode\hypertarget{ref-ekman2007directed}{}%
+{[}5{]} P. Ekman, ``The directed facial action task,'' in \emph{Handbook
+of emotion elicitation and assessment}, J. A. Coan and J. J. B. Allen,
+Eds. New York, NY: Oxford University Press, 2007, pp. 47--53.
+
+\leavevmode\hypertarget{ref-ekman1988universality}{}%
+{[}6{]} P. Ekman and K. G. Heider, ``The universality of a contempt
+expression: A replication,'' \emph{Motivation and Emotion}, vol. 12, no.
+3, pp. 303--308, 1988.
+
+\leavevmode\hypertarget{ref-ekman1987universals}{}%
+{[}7{]} P. Ekman \emph{et al.}, ``Universals and cultural differences in
+the judgments of facial expressions of emotion,'' \emph{Journal of
+Personality and Social Psychology}, vol. 53, no. 4, pp. 712--717, 1987.
+
+\leavevmode\hypertarget{ref-kraut1979social}{}%
+{[}8{]} R. E. Kraut and R. E. Johnston, ``Social and emotional messages
+of smiling: An ethological approach,'' \emph{Journal of Personality and
+Social Psychology}, vol. 37, no. 9, pp. 1539--1553, 1979.
+
+\leavevmode\hypertarget{ref-duran2017coherence}{}%
+{[}9{]} J. I. Durán, R. Reisenzein, and J.-M. Fernández-Dols,
+``Coherence between emotions and facial expressions,'' pp. 107--129,
+2017.
+
 \leavevmode\hypertarget{ref-averill1980constructivist}{}%
-Averill, James R. 1980. ``A Constructivist View of Emotion.'' In
-\emph{Theories of Emotion}, edited by Robert Plutchik and Henry
-Kellerman, 305--39. New York, NY: Academic Press.
+{[}10{]} J. R. Averill, ``A constructivist view of emotion,'' in
+\emph{Theories of emotion}, R. Plutchik and H. Kellerman, Eds. New York,
+NY: Academic Press, 1980, pp. 305--339.
+
+\leavevmode\hypertarget{ref-barrett2017emotions}{}%
+{[}11{]} L. F. Barrett, \emph{How emotions are made: The secret life of
+the brain}. Boston, MA: Houghton Mifflin Harcourt, 2017.
+
+\leavevmode\hypertarget{ref-barrett2017theory}{}%
+{[}12{]} L. F. Barrett, ``The theory of constructed emotion: An active
+inference account of interoception and categorization,'' \emph{Social
+Cognitive and Affective Neuroscience}, vol. 12, no. 1, pp. 1--23, 2017.
+
+\leavevmode\hypertarget{ref-crivelli2018facial}{}%
+{[}13{]} C. Crivelli and A. J. Fridlund, ``Facial displays are tools for
+social influence,'' \emph{Trends in Cognitive Sciences}, vol. 22, no. 5,
+pp. 388--399, 2018.
+
+\leavevmode\hypertarget{ref-frijda1997facial}{}%
+{[}14{]} N. H. Frijda and A. Tcherkassof, ``Facial expressions as modes
+of action readiness,'' in \emph{The psychology of facial expression}, J.
+A. Russell and J.-M. Fernández-Dols, Eds. Cambridge, UK: Cambridge
+University Press, 1997, pp. 78--102.
+
+\leavevmode\hypertarget{ref-fridlund1994human}{}%
+{[}15{]} A. J. Fridlund, \emph{Human facial expression: An evolutionary
+view}. San Diego, CA: Academic Press, 1994.
+
+\leavevmode\hypertarget{ref-lindquist2013s}{}%
+{[}16{]} K. A. Lindquist and M. Gendron, ``What's in a word? Language
+constructs emotion perception,'' \emph{Emotion Review}, vol. 5, no. 1,
+pp. 66--71, 2013.
+
+\leavevmode\hypertarget{ref-niedenthal2017embodied}{}%
+{[}17{]} P. M. Niedenthal, A. Wood, M. Rychlowska, and S. Korb,
+``Embodied simulation in decoding facial expression,'' in \emph{The
+science of facial expression}, J.-M. Fernández-Dols and J. A. Russell,
+Eds. New York, NY: Oxford University Press, 2017, pp. 397--414.
+
+\leavevmode\hypertarget{ref-doyle2017language}{}%
+{[}18{]} C. M. Doyle and K. A. Lindquist, ``Language and emotion:
+Hypotheses on the constructed nature of emotion perception,'' in
+\emph{The science of facial expression}, J.-M. Fernández-Dols and J. A.
+Russell, Eds. New York, NY: Oxford University Press, 2017, pp. 415--432.
+
+\leavevmode\hypertarget{ref-tcherkassof2013dynemo}{}%
+{[}19{]} A. Tcherkassof, D. Dupré, B. Meillon, N. Mandran, M. Dubois,
+and J.-M. Adam, ``DynEmo: A video database of natural facial expressions
+of emotions,'' \emph{The International Journal of Multimedia \& Its
+Applications}, vol. 5, no. 5, pp. 61--80, 2013.
+
+\leavevmode\hypertarget{ref-dupre2015oudjat}{}%
+{[}20{]} D. Dupré \emph{et al.}, ``Oudjat: A configurable and usable
+annotation tool for the study of facial expressions of emotion,''
+\emph{International Journal of Human-Computer Studies}, vol. 83, pp.
+51--61, 2015.
+
+\leavevmode\hypertarget{ref-mcduff2016affdex}{}%
+{[}21{]} D. McDuff, A. Mahmoud, M. Mavadati, M. Amr, J. Turcot, and R.
+el Kaliouby, ``AFFDEX sdk: A cross-platform real-time multi-face
+expression recognition toolkit,'' in \emph{Proceedings of the chi
+conference on human factors in computing systems}, 2016, pp. 3723--3726.
+
+\leavevmode\hypertarget{ref-dente2017measures}{}%
+{[}22{]} P. Dente, D. Küster, L. Skora, and E. Krumhuber, ``Measures and
+metrics for automatic emotion classification via facet,'' in
+\emph{Proceedings of the conference on the study of artificial
+intelligence and simulation of behaviour}, 2017, pp. 160--163.
+
+\leavevmode\hypertarget{ref-kappas2003facial}{}%
+{[}23{]} A. Kappas, ``What facial activity can and cannot tell us about
+emotions,'' in \emph{The human face}, M. Katsikitis, Ed. Boston, MA:
+Springer, 2003, pp. 215--234.
+
+\leavevmode\hypertarget{ref-bonanno2004brief}{}%
+{[}24{]} G. Bonanno and D. Keltner, ``The coherence of emotion systems:
+Comparing `on-line' measures of appraisal and facial expressions, and
+self-report,'' \emph{Cognition and Emotion}, vol. 18, no. 3, pp.
+431--444, 2004.
+
+\leavevmode\hypertarget{ref-fernandez2013emotion}{}%
+{[}25{]} J.-M. Fernández-Dols and C. Crivelli, ``Emotion and expression:
+Naturalistic studies,'' \emph{Emotion Review}, vol. 5, no. 1, pp.
+24--29, 2013.
+
+\leavevmode\hypertarget{ref-mckeown2013analogical}{}%
+{[}26{]} G. J. McKeown, ``The analogical peacock hypothesis: The sexual
+selection of mind-reading and relational cognition in human
+communication,'' \emph{Review of General Psychology}, vol. 17, no. 3,
+pp. 267--287, 2013.
+
+\leavevmode\hypertarget{ref-russell1993forced}{}%
+{[}27{]} J. A. Russell, ``Forced-choice response format in the study of
+facial expression,'' \emph{Motivation and Emotion}, vol. 17, no. 1, pp.
+41--51, 1993.
+
+\leavevmode\hypertarget{ref-mcduff2016discovering}{}%
+{[}28{]} D. McDuff, ``Discovering facial expressions for states of
+amused, persuaded, informed, sentimental and inspired,'' in
+\emph{Proceedings of the international conference on multimodal
+interaction}, 2016, pp. 71--75.
 
 \leavevmode\hypertarget{ref-baltruvsaitis2016openface}{}%
-Baltrušaitis, Tadas, Peter Robinson, and Louis-Philippe Morency. 2016.
-``Openface: An Open Source Facial Behavior Analysis Toolkit.'' In
-\emph{Proceedings of the Winter Conference on Applications of Computer
-Vision}, 1--10. \url{https://doi.org/10.1109/WACV.2016.7477553}.
-
-\leavevmode\hypertarget{ref-barrett2017emotions}{}%
-Barrett, Lisa Feldman. 2017a. \emph{How Emotions Are Made: The Secret
-Life of the Brain}. Boston, MA: Houghton Mifflin Harcourt.
-
-\leavevmode\hypertarget{ref-barrett2017theory}{}%
----------. 2017b. ``The Theory of Constructed Emotion: An Active
-Inference Account of Interoception and Categorization.'' \emph{Social
-Cognitive and Affective Neuroscience} 12 (1): 1--23.
-\url{https://doi.org/10.1093/scan/nsw154}.
-
-\leavevmode\hypertarget{ref-bonanno2004brief}{}%
-Bonanno, George, and Dacher Keltner. 2004. ``The Coherence of Emotion
-Systems: Comparing `on-Line' Measures of Appraisal and Facial
-Expressions, and Self-Report.'' \emph{Cognition and Emotion} 18 (3):
-431--44. \url{https://doi.org/10.1080/02699930341000149}.
-
-\leavevmode\hypertarget{ref-crivelli2018facial}{}%
-Crivelli, Carlos, and Alan J Fridlund. 2018. ``Facial Displays Are Tools
-for Social Influence.'' \emph{Trends in Cognitive Sciences} 22 (5):
-388--99. \url{https://doi.org/10.1016/j.tics.2018.02.006}.
-
-\leavevmode\hypertarget{ref-darwin1872expression}{}%
-Darwin, Charles. 1872. \emph{The Expression of the Emotions in Man and
-Animals}. London, UK: John Murray.
-
-\leavevmode\hypertarget{ref-dente2017measures}{}%
-Dente, Pasquale, Dennis Küster, Lina Skora, and E Krumhuber. 2017.
-``Measures and Metrics for Automatic Emotion Classification via Facet.''
-In \emph{Proceedings of the Conference on the Study of Artificial
-Intelligence and Simulation of Behaviour}, 160--63.
-
-\leavevmode\hypertarget{ref-doyle2017language}{}%
-Doyle, Cameron M, and Kristen A Lindquist. 2017. ``Language and Emotion:
-Hypotheses on the Constructed Nature of Emotion Perception.'' In
-\emph{The Science of Facial Expression}, edited by José-Miguel
-Fernández-Dols and James A Russell, 415--32. New York, NY: Oxford
-University Press.
-
-\leavevmode\hypertarget{ref-dupre2015oudjat}{}%
-Dupré, Damien, Daniel Akpan, Elena Elias, Jean-Michel Adam, Brigitte
-Meillon, Nicolas Bonnefond, Michel Dubois, and Anna Tcherkassof. 2015.
-``Oudjat: A Configurable and Usable Annotation Tool for the Study of
-Facial Expressions of Emotion.'' \emph{International Journal of
-Human-Computer Studies} 83: 51--61.
-\url{https://doi.org/10.1016/j.ijhcs.2015.05.010}.
-
-\leavevmode\hypertarget{ref-dupre2018accuracy}{}%
-Dupré, Damien, Nicole Andelic, Gawain Morrison, and Gary McKeown. 2018.
-``Accuracy of Three Commercial Automatic Emotion Recognition Systems
-Across Different Individuals and Their Facial Expressions.'' In
-\emph{Proceedings of the International Conference on Pervasive Computing
-and Communications}, 627--32.
-\url{https://doi.org/10.1109/PERCOMW.2018.8480127}.
-
-\leavevmode\hypertarget{ref-duran2017coherence}{}%
-Durán, Juan I, Rainer Reisenzein, and José-Miguel Fernández-Dols. 2017.
-``Coherence Between Emotions and Facial Expressions.'' Edited by
-José-Miguel Fernández-Dols and James A Russell, 107--29.
-
-\leavevmode\hypertarget{ref-ekman1992argument}{}%
-Ekman, Paul. 1992. ``An Argument for Basic Emotions.'' \emph{Cognition
-\& Emotion} 6 (3-4): 169--200.
-\url{https://doi.org/10.1080/02699939208411068}.
-
-\leavevmode\hypertarget{ref-ekman2007directed}{}%
----------. 2007. ``The Directed Facial Action Task.'' In \emph{Handbook
-of Emotion Elicitation and Assessment}, edited by James A. Coan and John
-J. B. Allen, 47--53. New York, NY: Oxford University Press.
-
-\leavevmode\hypertarget{ref-ekman1987universals}{}%
-Ekman, Paul, Wallace V Friesen, Maureen O'sullivan, Anthony Chan, Irene
-Diacoyanni-Tarlatzis, Karl Heider, Rainer Krause, et al. 1987.
-``Universals and Cultural Differences in the Judgments of Facial
-Expressions of Emotion.'' \emph{Journal of Personality and Social
-Psychology} 53 (4): 712--17.
-\url{https://doi.org/10.1037/0022-3514.53.4.712}.
-
-\leavevmode\hypertarget{ref-ekman1988universality}{}%
-Ekman, Paul, and Karl G Heider. 1988. ``The Universality of a Contempt
-Expression: A Replication.'' \emph{Motivation and Emotion} 12 (3):
-303--8. \url{https://doi.org/10.1007/BF00993116}.
-
-\leavevmode\hypertarget{ref-fernandez2013emotion}{}%
-Fernández-Dols, José-Miguel, and Carlos Crivelli. 2013. ``Emotion and
-Expression: Naturalistic Studies.'' \emph{Emotion Review} 5 (1): 24--29.
-\url{https://doi.org/10.1177/1754073912457229}.
-
-\leavevmode\hypertarget{ref-fridlund1994human}{}%
-Fridlund, Alan J. 1994. \emph{Human Facial Expression: An Evolutionary
-View}. San Diego, CA: Academic Press.
-
-\leavevmode\hypertarget{ref-frijda1997facial}{}%
-Frijda, Nico H, and Anna Tcherkassof. 1997. ``Facial Expressions as
-Modes of Action Readiness.'' In \emph{The Psychology of Facial
-Expression}, edited by James A Russell and José-Miguel Fernández-Dols,
-78--102. Cambridge, UK: Cambridge University Press.
-
-\leavevmode\hypertarget{ref-kappas2003facial}{}%
-Kappas, Arvid. 2003. ``What Facial Activity Can and Cannot Tell Us About
-Emotions.'' In \emph{The Human Face}, edited by Mary Katsikitis,
-215--34. Boston, MA: Springer.
-
-\leavevmode\hypertarget{ref-kraut1979social}{}%
-Kraut, Robert E, and Robert E Johnston. 1979. ``Social and Emotional
-Messages of Smiling: An Ethological Approach.'' \emph{Journal of
-Personality and Social Psychology} 37 (9): 1539--53.
-
-\leavevmode\hypertarget{ref-lindquist2013s}{}%
-Lindquist, Kristen A, and Maria Gendron. 2013. ``What's in a Word?
-Language Constructs Emotion Perception.'' \emph{Emotion Review} 5 (1):
-66--71. \url{https://doi.org/10.1177/1754073912451351}.
-
-\leavevmode\hypertarget{ref-mcduff2016discovering}{}%
-McDuff, Daniel. 2016. ``Discovering Facial Expressions for States of
-Amused, Persuaded, Informed, Sentimental and Inspired.'' In
-\emph{Proceedings of the International Conference on Multimodal
-Interaction}, 71--75. \url{https://doi.org/10.1145/2993148.2993192}.
-
-\leavevmode\hypertarget{ref-mcduff2016affdex}{}%
-McDuff, Daniel, Abdelrahman Mahmoud, Mohammad Mavadati, May Amr, Jay
-Turcot, and Rana el Kaliouby. 2016. ``AFFDEX Sdk: A Cross-Platform
-Real-Time Multi-Face Expression Recognition Toolkit.'' In
-\emph{Proceedings of the Chi Conference on Human Factors in Computing
-Systems}, 3723--6. \url{https://doi.org/10.1145/2851581.2890247}.
-
-\leavevmode\hypertarget{ref-mckeown2013analogical}{}%
-McKeown, Gary J. 2013. ``The Analogical Peacock Hypothesis: The Sexual
-Selection of Mind-Reading and Relational Cognition in Human
-Communication.'' \emph{Review of General Psychology} 17 (3): 267--87.
-\url{https://doi.org/10.1037/a0032631}.
-
-\leavevmode\hypertarget{ref-niedenthal2017embodied}{}%
-Niedenthal, Paula M, Adrienne Wood, Magdalena Rychlowska, and Sebastian
-Korb. 2017. ``Embodied Simulation in Decoding Facial Expression.'' In
-\emph{The Science of Facial Expression}, edited by José-Miguel
-Fernández-Dols and James A Russell, 397--414. New York, NY: Oxford
-University Press.
-
-\leavevmode\hypertarget{ref-russell1993forced}{}%
-Russell, James A. 1993. ``Forced-Choice Response Format in the Study of
-Facial Expression.'' \emph{Motivation and Emotion} 17 (1): 41--51.
-\url{https://doi.org/10.1007/BF00995206}.
-
-\leavevmode\hypertarget{ref-tcherkassof2013dynemo}{}%
-Tcherkassof, Anna, Damien Dupré, Brigitte Meillon, Nadine Mandran,
-Michel Dubois, and Jean-Michel Adam. 2013. ``DynEmo: A Video Database of
-Natural Facial Expressions of Emotions.'' \emph{The International
-Journal of Multimedia \& Its Applications} 5 (5): 61--80.
-\url{https://doi.org/10.5121/ijma.2013.5505}.
-
-\leavevmode\hypertarget{ref-de2019mama}{}%
-Waal, Frans de. 2019. \emph{Mama's Last Hug: Animal Emotions and What
-They Teach Us About Ourselves}. London, UK: Granta Books.
+{[}29{]} T. Baltrušaitis, P. Robinson, and L.-P. Morency, ``Openface: An
+open source facial behavior analysis toolkit,'' in \emph{Proceedings of
+the winter conference on applications of computer vision}, 2016, pp.
+1--10.
 

### 2019-04-18 ACII_2019_paper/ACII_2019_word.Rmd
word

@@
 ---
-title: "Coherence between subjective feeling and facial expression: a comparison of human and automatic recognition."
-output: word_document
-bibliography: mybibfile.bib
+title: "Coherence between subjective feeling and facial expression: A comparison of human and automatic recognition" 
+affiliation:
+  institution-columnar: true
+
+  institution:
+    - name: Dublin City University
+      location: Dublin, Ireland
+      email: damien.dupre@dcu.ie
+      mark: 1
+      author:
+        - name: Damien Dupré
+    - name: Grenoble Alpes University
+      location: Grenoble, France
+      mark: 2
+      author:
+        - name: Anna Tcherkassof
+          email: anna.tcherkassof@univ-grenoble-alpes.fr
 abstract: |
@@
 
+header-includes:
+  - \usepackage{booktabs}
+  - \usepackage{float}
+  - \usepackage{tabu}
+  - \usepackage{wrapfig}
+  - \usepackage[none]{hyphenat}
+  - \floatplacement{figure}{H}
+
+bibliography: mybibfile.bib
+output: word_document
+csl: http://www.zotero.org/styles/ieee
+fig_caption: yes
+keep_tex: yes
+number_sections: yes
+includes:
+  in_header: preamble-latex.tex
 ---
 
+\begin{IEEEkeywords}
+Facial expression, self-report, human observer, automatic recognition.
+\end{IEEEkeywords}
+
 ```{r setup, include = FALSE}
+# options ----------------------------------------------------------------------
 knitr::opts_chunk$set(
@@
   error = FALSE, 
-  echo = FALSE,
-  fig.pos='H')
-#
+  echo = FALSE)
+
 set.seed(123)# Seed for random number generation
 options(scipen = 999)# Disable scientific number format
-#
-library(kableExtra)
+
+# packages ----------------------------------------------------------------------
+library(kableExtra) # Tables
 library(tidyverse) # Data wrangling & grammar of graphics
 library(here) # Locate files relative to project root
-library(caret)
-library(plotROC)
-library(pROC)
-library(pander)
-library(papaja)
-library(captioner)
-table_nums <- captioner::captioner(prefix = "Table")
-table_nums(
-  name = "confusionTable_sr_hr", 
-  caption = "Agreement accuracy metrics for each emotion.")
-table_nums(
-  name = "confusionTable_sr_ar", 
-  caption = "Agreement accuracy metrics for each emotion.")
-#
-fig_nums <- captioner::captioner()
-fig_nums(
-  name = "dynemo_img", 
-  caption = "Example of a front facing recording synced with the full view of the participant and the elicitation task.
-This picture is taken from a pilot with projects collaborators and all gave a consent for the publication of their photos and videos.")
-fig_nums(
-  name = "confusionMatrix_sr_hr", 
-  caption = "Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the human observers.")
-fig_nums(
-  name = "confusionMatrix_sr_ar", 
-  caption = "Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the automatic classifier.")
-fig_nums(
-  name = "confusionMatrix_hr_ar", 
-  caption = "Confusion matrix of between the human obersers and automatic classifier recognition.")
+library(caret) # Confusion matrices
+library(papaja) # Correlation results
 ```
@@
 
-Introduction
+I. Introduction
 =============
 
-With the development of commercial automatic facial expression recognition tools [see @dupre2018accuracy for a non-exhaustive list of available tools]), industries and governments are gradually implementing this technology in order to track humans’ emotions in various scenarios (e.g., marketing, healthcare, automotive to name a few).
+With the development of commercial automatic facial expression recognition tools [see @dupre2018accuracy for a non-exhaustive list of available tools], industries and governments are gradually implementing this technology in order to track humans’ emotions in various scenarios (*e.g.*, marketing, healthcare, automotive to name a few).
 This technoloogy rests on the premise that facial expressions provide a direct access to individuals’ subjective feeling.
@@
 
-## The link between emotion and facial expression through the behaviorist approach
+## A. The Behaviorist Approach
 Based on the behaviorist approach initiated by Darwin in *The Expression of the Emotions in Man and Animals* [@darwin1872expression], facial expressions are conceived as a genuine displays of individuals’ inner emotional state.
 This hypothesis is used as a basis for the Basic Emotion Theory (BET) which states that a set of six emotions are universally displayed and are genetically hardwired not only in humans [@ekman1992argument] but also in different animal species [@de2019mama].
-According to this view, “when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including […] facial expressions.” [@ekman2007directed, p49]).
-To cope with critics, several amendments have been made to the BET, increasing the number of basic emotions from six to seven [@ekman1988universality] ) as well as adding the concept of “display rules” to explain cultural differences in the management of facial expressions [@ekman1987universals].
+According to this view, *“when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including […] facial expressions.”* [@ekman2007directed, 49].
+To cope with critics, several amendments have been made to the BET, increasing the number of basic emotions from six to seven [@ekman1988universality] as well as adding the concept of “display rules” to explain cultural differences in the management of facial expressions [@ekman1987universals].
 
@@
 
-## The link between emotion and facial expression through the social constructivist approach
-Detractors of the Basic Emotion Theory consider emotion not as genetically hardwired but as a learnt association between a given situation and an appropriate response [@averill1980constructivist; @barrett2017emotions]).
-For the tenants of the constructivist approach, emotions are “concepts” based on past experiences and which are “a collection of embodied, whole brain representations that predict what is about to happen in the sensory environment, what the best action is to deal with impending events, and their consequences for allostasis” [@barrett2017theory, p12].
+## B. The Social Constructivist Approach
+Detractors of the Basic Emotion Theory consider emotion not as genetically hardwired but as a learnt association between a given situation and an appropriate response [@averill1980constructivist; @barrett2017emotions].
+For the tenants of the constructivist approach, emotions are “concepts” based on past experiences and which are *“a collection of embodied, whole brain representations that predict what is about to happen in the sensory environment, what the best action is to deal with impending events, and their consequences for allostasis”* [@barrett2017theory, 12].
 Following this assumption, faces are best conceived as tools displaying signals in social interactions [@crivelli2018facial].
-These signals can convey individuals’ motivations and readiness [@frijda1997facial] or social messages [@fridlund1995human].
+These signals can convey individuals’ motivations and readiness [@frijda1997facial] or social messages [@fridlund1994human].
 Therefore, facial expressions are thought as behaviors which meaning is inferred by the observer.
@@
 
-For the emotion elicitation experiment, `r nrow(self_report_data)` French participants (`r self_report_gender[self_report_gender$genre_c == "F","n"]` females, `r self_report_gender[self_report_gender$genre_c == "H","n"]` males, *M*age = `r self_report_age[,"Age_m"]`, *SD*age = `r self_report_age[,"Age_sd"]`) were recruited to perform one out of 11 emotion elicitation tasks designed to trigger a positive, a specific negative or a neutral emotional state.
-Encoders’ face were recorded using an hidden camera resulting `r nrow(self_report_data)` front facing 768x576 videos varying from `r min(metadata_video$ffprobe_duration)`s to `r max(metadata_video$ffprobe_duration)`s (`r fig_nums("dynemo_img", display = "cite")`).
+For the emotion elicitation experiment, `r nrow(self_report_data)` French participants (`r self_report_gender[self_report_gender$genre_c == "F","n"]` females, `r self_report_gender[self_report_gender$genre_c == "H","n"]` males, *M*~age~ = `r self_report_age[,"Age_m"]`, *SD*~age~ = `r self_report_age[,"Age_sd"]`) were recruited to perform one out of 11 emotion elicitation tasks designed to trigger a positive, a specific negative or a neutral emotional state.
+Encoders’ face were recorded using an hidden camera resulting `r nrow(self_report_data)` front facing 768x576 videos varying from `r min(metadata_video$ffprobe_duration)`s to `r max(metadata_video$ffprobe_duration)`s (Figure \ref{fig:dynemo_img}).
 These recordings form the DynEmo database [see @tcherkassof2013dynemo for a full description of tasks and procedure].
 
-```{r dynemo_img, fig.cap=fig_nums("dynemo_img")}
-knitr::include_graphics(here::here("./img/dynemo_img.jpg"))
+```{r dynemo_img, fig.height=3, fig.cap="\\label{fig:dynemo_img}Example of a front facing recording synced with the full view of the participant and the elicitation task.
+This picture is taken from a pilot with projects collaborators and all gave a consent for the publication of their photos and videos."}
+raster::stack(here::here("./img/dynemo_img.png")) %>%
+  raster::plotRGB()
 ```
 
-After the emotion elicitation task the encoders rated their subjective feeling on Likert scales from 0 ("not at all") to 5 ("strongly") related to six “basic” emotion labels (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as six “non-basic” emotion labels (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
+After the emotion elicitation task the encoders rated their subjective feeling on Likert scales from 0 ("not at all") to \nolinebreak 5 \nolinebreak ("strongly") related to six “basic” emotion labels (*i.e.*, *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as six “non-basic” emotion labels (*i.e.*, *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
 
 Finally, a debriefing session was performed to ensure that encoders were not durably affected by the emotion elicitation task.
-The debriefing was also used to check that encoders did not guess the real purpose of the experiment (e.g.
-being filmed while they were performing an emotional elicitation task) to guarantee facial expressions’ genuineness.
+The debriefing was also used to check that encoders did not guess the real purpose of the experiment (*e.g.*, being filmed while they were performing an emotional elicitation task) to guarantee facial expressions’ genuineness.
 All encoders gave their agreement on their data and video to be processed for research purpose only.
 
-## Human Facial Expression Recognition
+## B. Human Facial Expression Recognition
 
@@
 
-For the human facial expression recognition method, `r nrow(unique_ppt_count)` student participants were recruited to annotate `r nrow(unique_video_count)` out of the `r nrow(self_report_data)` video, therefore only the `r nrow(unique_video_count)` annotated videos will be analysed in this paper.
+For the human facial expression recognition method, `r nrow(unique_ppt_count)` \nolinebreak student participants were recruited to annotate `r nrow(unique_video_count)` out of the `r nrow(self_report_data)` videos, therefore only the `r nrow(unique_video_count)` annotated videos will be analysed in this paper.
 Because videos have different durations, participants had to annotate a series of video corresponding to 30min long in total.
@@
 First, the participants had to identify the emotional sequences by pressing the space bar of their keyboard to indicate the beginning and the end of the emotional sequences while watching the video.
-Second, the participants watched each emotional sequence previously identified and labeled the sequence using one of the 12 emotions proposed including six “basic” emotion labels (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) and six "non-basic" emotion labels (i.e., *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
+Second, the participants watched each emotional sequence previously identified and labeled the sequence using one of the 12 \nolinebreak emotions proposed including six “basic” emotion labels (*i.e.*, *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) and six "non-basic" emotion labels (*i.e.*, *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
 They also had the possibility to indicate that the sequence was expressing none of the proposed emotion labels.
@@
 This annotation procedure results in a uni-dimensional time-series for each video per human observer identifying for each second of the video which emotion was recognized.
-Then, time-series corresponding to the same video were aggregated to calculate the proportion of human observers $x_{video_{i}.label_{j}.t_{k}}$ for each second of the video per emotional label \ref{eq:1}.
+Then, time-series corresponding to the same video were aggregated to calculate the proportion of human observers $x_{video_{i}.label_{j}.t_{k}}$ for each second of the video per emotional label (EQ\ref{eq:1}).
 
@@
 Affdex’s algorithm uses Histogram of Oriented Gradient (HOG) features and Support Vector Machine (SVM) classifiers in order to recognize facial expressions.
-For each video frame, Affdex identifies the probability $p_{video_{i}.label_{j}.t_{k}}$ from 0 to 100 (rescaled to 0 to 1 for the analysis) of the face as expressing one of the six “basic” emotion labels (i.e., *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as additional psychological states such as *valence*, *engagement* or *contempt*, and facial features such as *cheek raise*, *eye widen* or *jaw drop*.
+For each video frame, Affdex identifies the probability $p_{video_{i}.label_{j}.t_{k}}$ from 0 to 100 (rescaled to 0 to 1 for the analysis) of the face as expressing one of the six “basic” emotion labels (*i.e.*, *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as additional psychological states such as *valence*, *engagement* or *contempt*, and facial features such as *cheek raise*, *eye widen* or *jaw drop*.
 
 For both human and automatic recognition, to determine which of the six “basic” emotions can be used to identify each video, the recognition probability for each label by frame was converted into odd ratio by label [@dente2017measures].
-The highest sum of each odd ratio time-series defines the label recognized by the automatic classifier (\ref{eq:1} for human recognition and \ref{eq:1} for automatic recognition).
+The highest sum of each odd ratio time-series defines the label recognized by the automatic classifier (EQ\ref{eq:2} for human recognition and EQ\ref{eq:3} for automatic recognition).
 \begin{equation}
@@
 
-Encoders’ subjective feeling is compared with human observers recognition in a confusion matrix (`r fig_nums("confusionMatrix_sr_hr", display = "cite")`).
-Each emotion label used to describe encoders’ self-reported subjective feeling (i.e., the label rated with the highest value) is compared with the emotion labels which were rated with the highest score by human observers.
+The overall correlation of recognition and non-recognition between self-reported emotions and human observers recognition is significant but low (`r corr_sr_hr$full_result`).
+In order to identify differences according the emotional labels, encoders’ subjective feeling is compared with human observers recognition in a confusion matrix (Figure \ref{fig:confusionMatrix_sr_hr}).
 
-```{r confusionMatrix_sr_hr, fig.height=4, fig.cap=fig_nums("confusionMatrix_sr_hr")}
+```{r confusionMatrix_sr_hr, fig.height=6, fig.cap="\\label{fig:confusionMatrix_sr_hr}Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the human observers."}
 confusionMatrix_sr_hr_freq <- confusionMatrix_sr_hr$table %>% 
@@
 
-Results of the confusion matrix show a moderate agreement between the emotion felt by the encoder during the elicitation and the emotion recognized by the human observers (Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *disgust* (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_sr"]))` of the videos self-reported).
-Sensitivity, specificity, precision and F1 score for each emotion reveals that *happiness* has the highest coherence ratio whereas *sadness* has the lowest coherence ratio between true positives and false positives `r table_nums("confusionTable_sr_hr", display = "cite")`.
+Each emotion label used to describe encoders’ self-reported subjective feeling (*i.e.*, the label rated with the highest value) is compared with the emotion labels which were rated with the highest score by human observers.
+Results of the confusion matrix show a moderate agreement between the emotion felt by the encoder during the elicitation and the emotion recognized by the human observers (Accuracy \nolinebreak = \nolinebreak `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95% CI [`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *disgust* (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_sr"]))` of the videos self-reported).
+These results are far from those classicaly obtained in the literature for emotional facial expression recognition which ranges between 60% and 80% accuracy.
+However these results are mostly obtained with static (*i.e.,* pictures) and posed (*i.e.,* displayed by actors) facial expressions using only 6 emotional labels in a forced-choice paradigm.
+
 Interestingly human observers seem to recognize *surprise* expressed in videos where *anger*, *fear* *happiness* and *sadness* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "anger", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "fear", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "happiness", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "sadness", "prop_sr"]))` of the videos self-reported), and in a lower instance *happiness* was recognized in videos where *fear* and *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "fear", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
+
+Sensitivity, specificity, precision and F1 score for each emotion reveals that *happiness* has the highest coherence ratio whereas *sadness* has the lowest coherence ratio between true positives and false positives (Table \ref{table:confusionTable_sr_hr}).
 
@@
 
-# confusionMatrix_sr_hr_table %>%
-#   knitr::kable(
-#     "latex",
-#     caption = "Agreement accuracy metrics for each emotion. ",
-#     booktabs = TRUE,
-#     digits = 2,
-#     linesep = "",
-#     escape = FALSE,
-#     align = c('l',rep('c', 4))
-#   ) %>%
-#   kableExtra::kable_styling(
-#     font_size = 8,
-#     full_width = TRUE,
-#     latex_options = c("hold_position")) %>%
-#    add_footnote("Note. \\textit{na.} values are produced when not enough data are available to compute accuracy indicators.", notation="none", escape = FALSE)
-
 confusionMatrix_sr_hr_table %>%
-  papaja::apa_table(
-    caption = table_nums("confusionTable_sr_hr"),
-    row.names = FALSE,
-    format = "word",
-    escape = FALSE
-    )
+  knitr::kable(
+    "latex",
+    caption = "\\label{table:confusionTable_sr_hr}Human recognition accuracy metrics for each emotion.",
+    booktabs = TRUE,
+    digits = 2,
+    linesep = "",
+    escape = FALSE,
+    align = c('l',rep('c', 4))
+  ) %>%
+  kableExtra::kable_styling(
+    latex_options = "HOLD_position",
+    font_size = 8,
+    full_width = TRUE 
+    ) %>%
+   add_footnote("Note. \\textit{na.} values are produced when not enough data are available to compute accuracy indicators.", notation="none", escape = FALSE)
 ```
 
-However, self-reports show a very high proportion  of undetermined emotional states which reveals the potential limit of using 6-points likert scales to measure emotional self-reports.
+Accuracy metrics by emotional labels indicate a discrepancy in the ratio of true/false positives.
+Whereas *happiness* and *disgust* obtain the highest scores, *anger*, *surprise* and *sadness* have the lowest recognition ratio.
+The underlying effect of expression intensity may explain why *happiness* and *disgust* are easily recognized.
+*Anger* and *sadness* as non-socialy desirable emotion may be have been felt but not expressed.
+
+However, self-reports show a significant proportion of undetermined emotional states (`r scales::percent(nrow(list_automatic_ties)/358)` of the 358 videos) which reveals the potential limit of using 6-points likert scales to measure emotional self-reports.
 Indeed, encoders can easily score to the maximum for more than one emotion.
 
-## Correlation between self-report and automatic facial expression recognition
+## B. Automatic Classifier's Accuracy
 
@@
 
-Similarly to the previous analysis, a confusion matrix was used to compare encoders’ subjective feeling with the emotion label recognized by the automatic classifier (`r fig_nums("confusionMatrix_sr_ar", display = "cite")`).
+Similarly to the previous analysis, the overall correlation of recognition and non-recognition revealed a significant but very low link between self-reported emotions and automatic \nolinebreak classifier's recognition (`r corr_sr_ar$full_result`).
+A confusion matrix was used to compare encoders’ subjective feeling with the emotion label recognized by the automatic classifier (Figure \ref{fig:confusionMatrix_sr_ar}).
 
-```{r confusionMatrix_sr_ar, fig.height=4, fig.cap=fig_nums("confusionMatrix_sr_ar")}
+```{r confusionMatrix_sr_ar, fig.height=6, fig.cap="\\label{fig:confusionMatrix_sr_ar}Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the automatic classifier."}
 confusionMatrix_sr_ar_freq <- confusionMatrix_sr_ar$table %>% 
@@
 
-Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are somewhat similar to the ones with human observers (`r table_nums("confusionTable_sr_ar", display = "cite")`).
-Overall, a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`, 95%CI[`r round(confusionMatrix_sr_ar$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_ar$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_ar$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` of the video self-reported) is watched.
+Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are somewhat similar to the ones with human observers (Table \nolinebreak \ref{table:confusionTable_sr_ar}).
+Overall, a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy \nolinebreak = \nolinebreak `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`, 95% CI [`r round(confusionMatrix_sr_ar$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_ar$overall["AccuracyUpper"],2)`]; Kappa \nolinebreak = \nolinebreak `r round(confusionMatrix_sr_ar$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` of the video self-reported) is watched.
+
 Surprisingly the automatic classifier incorrectly recognized as *disgust* an important proportion of videos in which *anger*, *happiness* and *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "anger", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
@@
 
-# confusionMatrix_sr_ar_table %>%
-#   knitr::kable(
-#     "latex",
-#     caption = "Agreement accuracy metrics for each emotion.",
-#     booktabs = T,
-#     digits = 2,
-#     linesep = "",
-#     escape = FALSE,
-#     align = c('l',rep('c', 4))
-#   ) %>%
-#   kableExtra::kable_styling(
-#     font_size = 8,
-#     full_width = TRUE,
-#     latex_options = c("hold_position")) %>%
-#    add_footnote("Note. \\textit{na.} values are produced when not enough data are available to compute accuracy indicators.", notation="none", escape = FALSE)
-
 confusionMatrix_sr_ar_table %>%
-  papaja::apa_table(
-    caption = table_nums("confusionTable_sr_ar"),
-    row.names = FALSE,
-    format = "word",
-    escape = FALSE
-    )
+  knitr::kable(
+    "latex",
+    caption = "\\label{table:confusionTable_sr_ar}Autonatic recognition accuracy metrics for each emotion.",
+    booktabs = T,
+    digits = 2,
+    linesep = "",
+    escape = FALSE,
+    align = c('l',rep('c', 4))
+  ) %>%
+  kableExtra::kable_styling(
+    latex_options = c("HOLD_position"),
+    font_size = 8,
+    full_width = TRUE
+    ) %>%
+   add_footnote("Note. \\textit{na.} values are produced when not enough data are available to compute accuracy indicators.", notation="none", escape = FALSE)
 ```
 
-A comparable explanation  can be provided as the level of undetermined emotions are very high for the self reports.
+A comparable explanation involing the amount of undertermined video based on self-reports can be provided as the level of undetermined emotions are very high for the self reports.
 
-## Comparison between human and automatic recognition
+## C. Comparison Between Human and Automatic Recognition
 
-As previously mentioned, human observers appear to be more accurate than the automatic classifier to recognize individuals’ subjective feeling (human observer Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`; automatic classifier Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`.
-).
+As previously mentioned, human observers appear to be more accurate than the automatic classifier to recognize individuals’ subjective feeling (human observers Accuracy \nolinebreak = \nolinebreak`r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`; automatic classifier Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`; `r corr_hr_ar$full_result`).
 However, both make mistakes.
-A third confusion matrix is used to compare similarities between human observers and automatic classifier (`r fig_nums("confusionMatrix_hr_ar", display = "cite")`).
 
-```{r confusionMatrix_hr_ar, fig.height=4, fig.cap=fig_nums("confusionMatrix_hr_ar")}
+A third confusion matrix is used to compare similarities (diagonal) and differences between human observers and automatic classifier in classifing the six emotion labels (Figure \nolinebreak \ref{fig:confusionMatrix_hr_ar}).
+
+```{r confusionMatrix_hr_ar, fig.height=6, fig.cap="\\label{fig:confusionMatrix_hr_ar}Proportion of emotion labels classified by human observers which are recognized by the automatif classifier."}
 comparison_hr_ar <- dplyr::inner_join(human_recognition_score,automatic_recognition_score, by = "C_Video") %>% 
@@
 Except for *happiness* and *disgust* (respectively `r scales::percent(as.numeric(confusionMatrix_hr_ar_freq[confusionMatrix_hr_ar_freq$ar_emotion == "happiness" & confusionMatrix_hr_ar_freq$hr_emotion == "happiness", "prop_hr"]))` and `r scales::percent(as.numeric(confusionMatrix_hr_ar_freq[confusionMatrix_hr_ar_freq$ar_emotion == "disgust" & confusionMatrix_hr_ar_freq$hr_emotion == "disgust", "prop_hr"]))` of common labelling), there is no clear common pattern.
-Moreover, the automatic classifier has a tendency to label as *disgust* videos labeled as *sadness* by human observers and as *happiness* videos labaled as *fear* by human observers
+Moreover, the automatic classifier has a tendency to label as *disgust* videos labeled as *sadness* by human observers and as *happiness* videos labeled as *fear* by human observers
 
-Conclusion
+IV. Discussion and Conclusion
 ============
@@
 
-Such results support the hypothesis advanced by some authors of a low emotion–expression coherence [e.g., @kappas2003facial].
+Such results support the hypothesis advanced by some authors of a low emotion–expression coherence [*e.g.*, @kappas2003facial].
 In many instances, facial displays are not associated with a concordant emotional state, even any emotional state at all [@bonanno2004brief; @fernandez2013emotion].
 More and more evidences are showing that facial expressions are in reality not expressing emotions [@mckeown2013analogical].
-As well as other nonverbal behaviors, facial movements are not only assumed to be determined by emotion but also by various other causes, such as psychological states (e.g., motivations or pain), to say nothing of social context and sociocultural norms [such as “display rules”, @ekman1987universals].
+As well as other nonverbal behaviors, facial movements are not only assumed to be determined by emotion but also by various other causes, such as psychological states (*e.g.*, \nolinebreak motivations or pain), to say nothing of social context and sociocultural norms [such as “display rules”, @ekman1987universals].
 This multiple determination excludes any possibility of drawing an inference from facial activity on the underlying psychological state (emotional or other).
+
 Beyond the present observations showing a weak coherence between subjective feelings and spontaneous facial expressions, this study provides some light on the controversy between the behaviorist and the constructivist approaches.

### 2019-04-18 script/data_wrangling_automatic_recognition.R

@@ -7,7 +7,7 @@ library(tidyverse)
 # 
 # readr::write_rds(automatic_recognition_data, here::here("./data/automatic_recognition_data/automatic_recognition_data.rds"))
 
-automatic_recognition_data <- readr::read_rds(here::here("./data/automatic_recognition_data/automatic_recognition_data.rds"))
+automatic_recognition_data <- readr::read_rds("C:/Users/dupred/OneDrive/data/dynemo_feel_data/automatic_recognition_data/automatic_recognition_data.rds")
 
 automatic_recognition_score <- automatic_recognition_data %>% 
   dplyr::select(C_Video = filename, TimeStamp, happiness = joy, fear, disgust, sadness, anger, surprise) %>% 

### 2019-04-18 script/data_wrangling_human_recognition.R

@@ -2,9 +2,9 @@
 library(tidyverse)
 library(splitstackshape)
 # upload data ------------------------------------------------------------------
-metadata_video <- readr::read_rds(here::here("data/human_recognition_data/metadata_video.rds"))
+metadata_video <- readr::read_rds("C:/Users/dupred/OneDrive/data/dynemo_feel_data/human_recognition_data/metadata_video.rds")
 
-human_recognition_data <- here::here("./data/human_recognition_data") %>% 
+human_recognition_data <- "C:/Users/dupred/OneDrive/data/dynemo_feel_data/human_recognition_data" %>% 
   fs::dir_ls(regexp = "\\.csv$") %>% 
   purrr::map_dfr(readr::read_csv2, .id = "source", col_types = cols(
     C_INDUC = col_character(),

### 2019-04-18 script/data_wrangling_self_report.R

@@ -3,7 +3,7 @@ library(here)
 library(tidyverse)
 library(janitor)
 # upload data ------------------------------------------------------------------
-self_report_data <- readr::read_rds(here::here("data/human_self_report_data/Base_Finale_20090114.rds")) %>% 
+self_report_data <- readr::read_rds("C:/Users/dupred/OneDrive/data/dynemo_feel_data/human_self_report_data/Base_Finale_20090114.rds") %>% 
   janitor::clean_names() %>% 
   dplyr::mutate(C_Video = stringr::str_replace(fichier_video_v,".mpg","")) %>% 
   dplyr::select(

### 2019-04-21 ACII_2019_paper/ACII_2019_paper.Rmd
final version

@@
 ---
-title: "Coherence between subjective feeling and facial expression: A comparison of human and automatic recognition" 
+title: "The Emotional Coherence Controversy: Evidence from Human and Automatic Expression Recognition" 
 affiliation:
@@
     - name: Dublin City University
+      department: Business School
       location: Dublin, Ireland
@@
         - name: Damien Dupré
-    - name: Grenoble Alpes University
+    - name: Université Grenoble Alpes
+      department: Laboratoire Inter-universitaire de Psychologie
       location: Grenoble, France
@@
 abstract: |
-  While it has been taken for granted in the development of several automatic facial expression recognition tools, the question of the link between subjective feelings and facial expressions is still a subject of debate.
-On one hand the behaviorist approach conceives emotions as genetically hardwired and therefore being genuinely displayed through facial expressions.
-On the other hand the constructivist approach conceives emotions as socially constructed, the emotional meaning of a facial expression being inferred by the observer.
-In order to evaluate the link between the subjective feeling of emotions and their recognition based on facial expression, 232 videos of encoders recruited to carry out an emotion elicitation task were annotated by 1383 human observers as well as by an automatic facial expression classifier.
-Results show a low accuracy of human observers and of the automatic classifier to infer the subjective feeling from the facial expressions displayed by encoders .
-They also show a weak link between self-reported emotional states and facial emotional displays.
-Based on these results, the hypothesis of genetically hardwired emotion genuinely display is difficult to support whereas the idea of emotion and facial expression socially constructed appears to be more likely.
-Then automatic emotion recognition tools based on facial expressions should be questioned.
+  While it has been taken for granted in the development of several automatic facial expression recognition tools, the question of the coherence between subjective feelings and facial expressions is still a subject of debate.
+On one hand, the behaviorist approach conceives emotions as genetically hardwired and therefore being genuinely displayed through facial expressions.
+On the other hand, the constructivist approach conceives emotions as socially constructed; the emotional meaning of a facial expression being inferred by the observer.
+In order to evaluate the coherence between the subjective feeling of emotions and their recognition based on facial expression, 232 videos of encoders recruited to carry out an emotion elicitation task were annotated by 1383 human observers as well as by an automatic facial expression classifier.
+Results show low accuracy of human observers and of the automatic classifier to infer the subjective feeling from the facial expressions displayed by encoders.
+They also show a weak coherence between self-reported emotional states and facial emotional displays.
+Based on these results, the hypothesis of genetically hardwired emotion genuinely displayed is difficult to support, whereas the idea of emotion and facial expression socially constructed appears to be more likely.
+Accordingly, automatic emotion recognition tools based on facial expressions should be questioned.
 
@@
   - \floatplacement{figure}{H}
+  - \usepackage{flushend}
+  - \usepackage{biblatex}
 
@@
 
-With the development of commercial automatic facial expression recognition tools [see @dupre2018accuracy for a non-exhaustive list of available tools], industries and governments are gradually implementing this technology in order to track humans’ emotions in various scenarios (*e.g.*, marketing, healthcare, automotive to name a few).
-This technoloogy rests on the premise that facial expressions provide a direct access to individuals’ subjective feeling.
-Even if this premise is central to the modern mainstream approach of human emotion, recent research in affective science are challenging it.
-Once the two competing approaches briefly described as well as the predictions they respectively entail, an experiment testing these hypotheses will be presented and its results analyzed in order to provide empirical evidence to contribute to answer the question.
+With the development of commercial automatic facial expression recognition tools [see @dupre2018accuracy for a non-exhaustive list of available tools], industries and governments are gradually implementing this technology in order to track humans’ emotions in various scenarios (*e.g.*, in marketing, healthcare, and the automotive industry to name a few).
+This technology rests on the premise that facial expressions provide a direct access to individuals’ subjective feelings also called "emotional coherence".
+Even if this premise is central to the modern mainstream approach of human emotion, recent research in affective science is challenging it.
+Once the two competing approaches are briefly described along with the predictions they respectively entail, an experiment testing these hypotheses will be presented and its results analyzed in order to provide empirical evidence to contribute to answer the question.
 
 ## A. The Behaviorist Approach
-Based on the behaviorist approach initiated by Darwin in *The Expression of the Emotions in Man and Animals* [@darwin1872expression], facial expressions are conceived as a genuine displays of individuals’ inner emotional state.
+Based on the behaviorist approach initiated by Darwin in *The Expression of the Emotions in Man and Animals* [@darwin1872expression], facial expressions are conceived as a genuine displays of an individual's inner emotional state.
 This hypothesis is used as a basis for the Basic Emotion Theory (BET) which states that a set of six emotions are universally displayed and are genetically hardwired not only in humans [@ekman1992argument] but also in different animal species [@de2019mama].
 According to this view, *“when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including […] facial expressions.”* [@ekman2007directed, 49].
-To cope with critics, several amendments have been made to the BET, increasing the number of basic emotions from six to seven [@ekman1988universality] as well as adding the concept of “display rules” to explain cultural differences in the management of facial expressions [@ekman1987universals].
+To respond to criticisms, several amendments have been made to the BET, increasing the number of basic emotions from six to seven [@ekman1988universality] as well as adding the concept of “display rules” to explain cultural differences in the management of facial expressions [@ekman1987universals].
 
-Even if this theory obtained a popular support, it fails to explain how individuals can feel emotions without expressing them and how individuals can express emotions without feeling them, in instances in which display rules cannot be called upon [@kraut1979social; @duran2017coherence].
+Even if this theory obtained popular support, it fails to explain how individuals can feel emotions without expressing them, and how individuals can express emotions without feeling them, in instances in which display rules cannot be called upon [@kraut1979social; @duran2017coherence].
 These evidences have led to an alternative conception, notably the social constructivist approach.
@@
 Detractors of the Basic Emotion Theory consider emotion not as genetically hardwired but as a learnt association between a given situation and an appropriate response [@averill1980constructivist; @barrett2017emotions].
-For the tenants of the constructivist approach, emotions are “concepts” based on past experiences and which are *“a collection of embodied, whole brain representations that predict what is about to happen in the sensory environment, what the best action is to deal with impending events, and their consequences for allostasis”* [@barrett2017theory, 12].
+For the proponents of the constructivist approach, emotions are “concepts” based on past experiences and which are *“a collection of embodied, whole brain representations that predict what is about to happen in the sensory environment, what the best action is to deal with impending events, and their consequences for allostasis”* [@barrett2017theory, 12].
 Following this assumption, faces are best conceived as tools displaying signals in social interactions [@crivelli2018facial].
@@
 ## C. Emotional Facial Expression Recognition
-Regarding emotional facial expression (EFE) recognition, the behaviorist approach assertions and the constructivist approach claims lead to two divergent predictions.
-The foremost postulates that, when triggered, each basic emotion is expressed by a prototypical face (non basic emotions being blends of the basic ones).
+Regarding emotional facial expression (EFE) recognition, behaviorist and constructivist approaches lead to two divergent predictions.
+The former postulates that, when triggered, each basic emotion is expressed by a prototypical face (non basic emotions being blends of the basic ones).
 Basic emotions are easily recognized by all human observers and emotional states are accessible by facial measurement.
 As the recognition of facial expressions is based on the identification of specific patterns of facial movements, it implies that EFE recognition of both human observers and automatic classifiers should be as accurate.
-The constructivist approach, in contrast, affirms that facial expressions do not provide a direct access to individuals’ subjective feeling.
+The constructivist approach, in contrast, affirms that facial expressions do not provide a direct access to individuals’ subjective feelings.
 Even if the face possibly moves during an emotional episode, facial muscle movements are not linked in a one-to-one manner to a specific discrete emotional experience [@doyle2017language, p418].
@@
 ```{r models_img, fig.height=5, fig.cap="\\label{fig:models_img}Comparison of behaviorist and social constructivist approaches according to the expected differences in emotion recognition."}
-# knitr::include_graphics("C:/Users/dupred/Desktop/Projects/dynemo_feel_express/img/dynemo_img.png", auto_pdf = FALSE)
 raster::stack(here::here("./img/models.png")) %>%
@@
 
-The aim of the current paper is to investigate the link between the subjective feeling of emotions and its recognition from facial expressions in order to give credit either to the behaviorist plea or to the constructivist’s one.
+The aim of the current paper is to investigate the coherence between the subjective feeling of emotions and its recognition from facial expressions in order to give credit either to the behaviorist plea or to the constructivist’s one.
 Contrary to most studies using posed and static EFE, the present study focuses on natural EFE.
@@
 
-To evaluate the link between subjective feeling of emotions and their recognition from facial expressions, encoders were first recruited to perform an emotion elicitation task while their facial expression was video recorded.
+To evaluate the coherence between subjective feeling of emotions and their recognition from facial expressions, encoders were first recruited to perform an emotion elicitation task while their facial expression was video recorded.
 Then, the videos of the encoders’ faces were shown to human observers and were also analysed by an automatic classifier in order to identify which emotion was displayed.
@@
 ```{r dynemo_img, fig.height=3, fig.cap="\\label{fig:dynemo_img}Example of a front facing recording synced with the full view of the participant and the elicitation task.
-This picture is taken from a pilot with projects collaborators and all gave a consent for the publication of their photos and videos."}
+This picture is taken from a pilot with projects collaborators and all gave consent for the publication of their photos and videos."}
 raster::stack(here::here("./img/dynemo_img.png")) %>%
@@
 The debriefing was also used to check that encoders did not guess the real purpose of the experiment (*e.g.*, being filmed while they were performing an emotional elicitation task) to guarantee facial expressions’ genuineness.
-All encoders gave their agreement on their data and video to be processed for research purpose only.
+All encoders gave their agreement on their data and video to be processed for research purposes only.
 
@@
 The overall correlation of recognition and non-recognition between self-reported emotions and human observers recognition is significant but low (`r corr_sr_hr$full_result`).
-In order to identify differences according the emotional labels, encoders’ subjective feeling is compared with human observers recognition in a confusion matrix (Figure \ref{fig:confusionMatrix_sr_hr}).
+In order to identify differences according to the emotional labels, encoders’ subjective feelings are compared with human observers' recognition in a confusion matrix (Figure \ref{fig:confusionMatrix_sr_hr}).
 
-```{r confusionMatrix_sr_hr, fig.height=6, fig.cap="\\label{fig:confusionMatrix_sr_hr}Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the human observers."}
+```{r confusionMatrix_sr_hr, fig.height=6, fig.cap="\\label{fig:confusionMatrix_sr_hr}Confusion matrix between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the human observers."}
 confusionMatrix_sr_hr_freq <- confusionMatrix_sr_hr$table %>% 
@@
 Results of the confusion matrix show a moderate agreement between the emotion felt by the encoder during the elicitation and the emotion recognized by the human observers (Accuracy \nolinebreak = \nolinebreak `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95% CI [`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *disgust* (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_sr"]))` of the videos self-reported).
-These results are far from those classicaly obtained in the literature for emotional facial expression recognition which ranges between 60% and 80% accuracy.
+These results are far from those classically obtained in the literature for emotional facial expression recognition which ranges between 60% and 80% accuracy.
 However these results are mostly obtained with static (*i.e.,* pictures) and posed (*i.e.,* displayed by actors) facial expressions using only 6 emotional labels in a forced-choice paradigm.
@@
 
-Sensitivity, specificity, precision and F1 score for each emotion reveals that *happiness* has the highest coherence ratio whereas *sadness* has the lowest coherence ratio between true positives and false positives (Table \ref{table:confusionTable_sr_hr}).
+Sensitivity, specificity, precision and F1 scores for each emotion reveals that *happiness* has the highest coherence ratio whereas *sadness* has the lowest coherence ratio between true positives and false positives (Table \ref{table:confusionTable_sr_hr}).
 
@@
 The underlying effect of expression intensity may explain why *happiness* and *disgust* are easily recognized.
-*Anger* and *sadness* as non-socialy desirable emotion may be have been felt but not expressed.
+*Anger* and *sadness* as non-socially desirable emotions may be have been felt but not expressed.
 
@@
 
-Similarly to the previous analysis, the overall correlation of recognition and non-recognition revealed a significant but very low link between self-reported emotions and automatic \nolinebreak classifier's recognition (`r corr_sr_ar$full_result`).
+Similarly to the previous analysis, the overall correlation of recognition and non-recognition revealed a significant but very low coherence between self-reported emotions and automatic \nolinebreak classifier's recognition (`r corr_sr_ar$full_result`).
 A confusion matrix was used to compare encoders’ subjective feeling with the emotion label recognized by the automatic classifier (Figure \ref{fig:confusionMatrix_sr_ar}).
@@
 Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are somewhat similar to the ones with human observers (Table \nolinebreak \ref{table:confusionTable_sr_ar}).
-Overall, a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy \nolinebreak = \nolinebreak `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`, 95% CI [`r round(confusionMatrix_sr_ar$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_ar$overall["AccuracyUpper"],2)`]; Kappa \nolinebreak = \nolinebreak `r round(confusionMatrix_sr_ar$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` of the video self-reported) is watched.
+Overall, a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy \nolinebreak = \nolinebreak `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`, 95% CI [`r round(confusionMatrix_sr_ar$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_ar$overall["AccuracyUpper"],2)`]; Kappa \nolinebreak = \nolinebreak `r round(confusionMatrix_sr_ar$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` of the video self-reported) is evident.
 
-Surprisingly the automatic classifier incorrectly recognized as *disgust* an important proportion of videos in which *anger*, *happiness* and *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "anger", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
+Surprisingly the automatic classifier incorrectly recognized as *disgust* an significant proportion of videos in which *anger*, *happiness* and *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "anger", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
 In parallel, the automatic classifier recognized as *happiness* videos in which *fear* and *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "fear", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
@@
 
-A comparable explanation involing the amount of undertermined video based on self-reports can be provided as the level of undetermined emotions are very high for the self reports.
+A comparable explanation involving the amount of undetermined video based on self-reports can be provided, as the level of undetermined emotions are very high for the self reports.
 
@@
 
-As previously mentioned, human observers appear to be more accurate than the automatic classifier to recognize individuals’ subjective feeling (human observers Accuracy \nolinebreak = \nolinebreak`r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`; automatic classifier Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`; `r corr_hr_ar$full_result`).
+As previously mentioned, human observers appear to be more accurate than the automatic classifier to recognize an individual's subjective feeling (human observers Accuracy \nolinebreak = \nolinebreak`r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`; automatic classifier Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`; `r corr_hr_ar$full_result`).
 However, both make mistakes.
@@
 Except for *happiness* and *disgust* (respectively `r scales::percent(as.numeric(confusionMatrix_hr_ar_freq[confusionMatrix_hr_ar_freq$ar_emotion == "happiness" & confusionMatrix_hr_ar_freq$hr_emotion == "happiness", "prop_hr"]))` and `r scales::percent(as.numeric(confusionMatrix_hr_ar_freq[confusionMatrix_hr_ar_freq$ar_emotion == "disgust" & confusionMatrix_hr_ar_freq$hr_emotion == "disgust", "prop_hr"]))` of common labelling), there is no clear common pattern.
-Moreover, the automatic classifier has a tendency to label as *disgust* videos labeled as *sadness* by human observers and as *happiness* videos labeled as *fear* by human observers
+Moreover, the automatic classifier has a tendency to label as *disgust* videos labeled as *sadness* by human observers, and as *happiness* videos labeled as *fear* by human observers.
 
@@
 
-Despite being one on the most investigated question in affective science, the coherence between emotion felt and facially displayed and facial expression recognized is a hot topic.
-To date no clear evidence has been found to definitely answer it.
-Yet, with the growing interest of industries and government to monitor individual’s psychological states, this issue is under high pressure.
-The present research aims to provide some empirical data to the question.
+Despite being one of the most investigated questions in affective science, the coherence between emotion felt and facially displayed on the one hand, and facial expression recognized on the other is a hot topic.
+To date no clear evidence has been found to definitively solve the questions raised.
+Yet, with the growing interest of industries and government in monitoring individual’s psychological states, this issue is under intense scrutiny.
+The present research aims to provide some empirical data to answer some of the questions posed.
 The faces encoders displayed when confronted with an emotional eliciting task were submitted both to human and to automatic recognition.
-The criterion for recognition accuracy was the subjective feeling self-reported by the encoder once the elicitation task carried out.
+The criterion for recognition accuracy was the subjective feeling self-reported by the encoder once the elicitation task was carried out.
 Results first reveal a low coherence between emotion felt and facial expression displayed.
-Secondly, results show a low accuracy for both humans and automatic classifier in identifying the inner emotional states of these encoders based on their facial expressions.
+Secondly, results show low accuracy rates for both humans and the automatic classifier in identifying the inner emotional states of these encoders based on their facial expressions.
 Thirdly, human observers prove to be better at recognizing the emotion facially expressed than the automatic recognition tool is.
@@
 In many instances, facial displays are not associated with a concordant emotional state, even any emotional state at all [@bonanno2004brief; @fernandez2013emotion].
-More and more evidences are showing that facial expressions are in reality not expressing emotions [@mckeown2013analogical].
+More and more evidence is showing that facial expressions are in reality not expressing emotions [@mckeown2013analogical].
 As well as other nonverbal behaviors, facial movements are not only assumed to be determined by emotion but also by various other causes, such as psychological states (*e.g.*, \nolinebreak motivations or pain), to say nothing of social context and sociocultural norms [such as “display rules”, @ekman1987universals].
@@
 
-Beyond the present observations showing a weak coherence between subjective feelings and spontaneous facial expressions, this study provides some light on the controversy between the behaviorist and the constructivist approaches.
+Beyond the present observations showing a weak coherence between subjective feelings and spontaneous facial expressions, this study sheds some light on the controversy between the behaviorist and the constructivist approaches.
 The first of those two lines of thinking [@ekman1992argument] assumes that expressions of emotion are brief and coherent patterns of facial muscle movements that co-vary with discrete subjective experiences.
@@
 It is the observer that infers the emotional meaning of the facial expression.
-As a consequence, one can predict from the first line of thinking that individuals’ emotional subjective feeling should be correlated to the recognition of facial expressions from both human observers and automatic classifiers whereas if emotions are social constructs, as stated by the second line of thinking, human observer’s should be better at perceiving emotions in face than automatic classifiers.
+As a consequence, one can predict from the first line of thinking that individuals’ emotional subjective feeling should be correlated to the recognition of facial expressions from both human observers and automatic classifiers whereas if emotions are social constructs, as stated by the second line of thinking, human observer’s should be better at perceiving emotions expressed on the face than automatic classifiers.
 
 Present results plead in favor of the latter stance.
-They show that human observers are more accurate than the automatic recognition tool to identify individuals’ subjective feeling on the basis of their face.
+They show that human observers are more accurate than the automatic recognition tool to identify an individual's subjective feeling on the basis of their face.
 Moreover, mistakes made by human observers look less arbitrary to the ones made by the classifier.
-For instance, even if mix-up between disgust and anger is sometimes reported in recognition studies, confusions such as the ones produced by the classifier have never been noted for human observers.
+For instance, even if a mix-up between disgust and anger is sometimes reported in recognition studies, confusions such as the ones produced by the classifier have never been noted for human observers.
 A possible explanation is that human observers are assessing dynamic expressive sequences.
@@
 Instead of asking the human annotators to provide a unique label, a more subtle approach was chosen to mimic results provided by the automatic classifier.
-Whereas this paradigm is longer and more complicated, it can lead to more robust results in reducing the forced-choice bais [@russell1993forced].
-However this procedure can also reduce the human observers accuracy.
+Whereas this paradigm is longer and more complicated, it can lead to more robust results in reducing the forced-choice bias [@russell1993forced].
+However this procedure can also reduce the human observers' accuracy.
 In this regard, the results of the human observation could have been more ambiguous because it is not the natural way that people are inferring meaning from facial expressions.
-An alternative explanation relies in reducing the recognition bais involved in the classic recognition paradigm.
-Classic forced-choice paradigms obtain artificial high results, thus by using a more evolved approach observers’ accuracy may have been lowered.
+An alternative explanation relies in reducing the recognition bias involved in the classic recognition paradigm.
+Classic forced-choice paradigms obtain artificially high results, thus by using a more evolved approach observers’ accuracy may have been lowered.
 
-Considering the above, the results provide additional evidence that individuals’ subjective feeling can not be inferred from facial expressions and in our case invalidate the hypothesis of hardwired emotions unambiguously displayed on the face.
+Considering the above, the results provide additional evidence that an individual's subjective feeling can not be inferred from facial expressions and in our case invalidate the hypothesis of hardwired emotions unambiguously displayed on the face.
 Even if emotions were hardwired, in everyday life one does not observe prototypical facial expressions and therefore research should be focused on analysing non-prototypical facial expressions.
-Advancements in identifying “non-basic” emotion labels [@mcduff2016discovering] as well as non-prototypical facial expression have been made in the developpement of automatic facial expression recognition tools.
-However this result suggests that automatic facial expression recognition tools should merely evaluate facial morphology features such as action units [*e.g.*, OpenFace presented in @baltruvsaitis2016openface] rather than inferring supposedly emotional or affective states.
+Advancements in identifying “non-basic” emotion labels [@mcduff2016discovering] as well as non-prototypical facial expression have been made in the development of automatic facial expression recognition tools.
+However, these results suggest that automatic facial expression recognition tools should merely evaluate facial morphology features such as action units (already evaluated in OpenFace [@baltruvsaitis2016openface], Affectiva's Affdex [@mcduff2016affdex] or Vicar Vision's FaceReader [@den2005facereader] to name a few) rather than inferring supposedly emotional or affective states.
 
@@
 ==========
+\footnotesize

### 2019-04-21 ACII_2019_paper/ACII_2019_paper.tex
final version

@@
 \hypersetup{
-            pdftitle={Coherence between subjective feeling and facial expression: A comparison of human and automatic recognition},
+            pdftitle={The Emotional Coherence Controversy: Evidence from Human and Automatic Expression Recognition},
             pdfborder={0 0 0},
@@
 \floatplacement{figure}{H}
+\usepackage{flushend}
+\usepackage{biblatex}
 
@@
 % Do not put math or special symbols in the title.
-\title{Coherence between subjective feeling and facial expression: A comparison
-of human and automatic recognition}
+\title{The Emotional Coherence Controversy: Evidence from Human and Automatic
+Expression Recognition}
 
@@
 }
-\IEEEauthorblockA{Dublin City University\\
+\IEEEauthorblockA{Business School\\
+Dublin City University\\
 Dublin, Ireland
@@
 }
-\IEEEauthorblockA{Grenoble Alpes University\\
+\IEEEauthorblockA{Laboratoire Inter-universitaire de Psychologie\\
+Université Grenoble Alpes\\
 Grenoble, France
@@
 While it has been taken for granted in the development of several
-automatic facial expression recognition tools, the question of the link
-between subjective feelings and facial expressions is still a subject of
-debate. On one hand the behaviorist approach conceives emotions as
-genetically hardwired and therefore being genuinely displayed through
-facial expressions. On the other hand the constructivist approach
-conceives emotions as socially constructed, the emotional meaning of a
-facial expression being inferred by the observer. In order to evaluate
-the link between the subjective feeling of emotions and their
-recognition based on facial expression, 232 videos of encoders recruited
-to carry out an emotion elicitation task were annotated by 1383 human
-observers as well as by an automatic facial expression classifier.
-Results show a low accuracy of human observers and of the automatic
-classifier to infer the subjective feeling from the facial expressions
-displayed by encoders . They also show a weak link between self-reported
-emotional states and facial emotional displays. Based on these results,
-the hypothesis of genetically hardwired emotion genuinely display is
-difficult to support whereas the idea of emotion and facial expression
-socially constructed appears to be more likely. Then automatic emotion
-recognition tools based on facial expressions should be questioned.
+automatic facial expression recognition tools, the question of the
+coherence between subjective feelings and facial expressions is still a
+subject of debate. On one hand, the behaviorist approach conceives
+emotions as genetically hardwired and therefore being genuinely
+displayed through facial expressions. On the other hand, the
+constructivist approach conceives emotions as socially constructed; the
+emotional meaning of a facial expression being inferred by the observer.
+In order to evaluate the coherence between the subjective feeling of
+emotions and their recognition based on facial expression, 232 videos of
+encoders recruited to carry out an emotion elicitation task were
+annotated by 1383 human observers as well as by an automatic facial
+expression classifier. Results show low accuracy of human observers and
+of the automatic classifier to infer the subjective feeling from the
+facial expressions displayed by encoders. They also show a weak
+coherence between self-reported emotional states and facial emotional
+displays. Based on these results, the hypothesis of genetically
+hardwired emotion genuinely displayed is difficult to support, whereas
+the idea of emotion and facial expression socially constructed appears
+to be more likely. Accordingly, automatic emotion recognition tools
+based on facial expressions should be questioned.
 \end{abstract}
@@
 implementing this technology in order to track humans' emotions in
-various scenarios (\emph{e.g.}, marketing, healthcare, automotive to
-name a few). This technoloogy rests on the premise that facial
-expressions provide a direct access to individuals' subjective feeling.
-Even if this premise is central to the modern mainstream approach of
-human emotion, recent research in affective science are challenging it.
-Once the two competing approaches briefly described as well as the
-predictions they respectively entail, an experiment testing these
-hypotheses will be presented and its results analyzed in order to
-provide empirical evidence to contribute to answer the question.
+various scenarios (\emph{e.g.}, in marketing, healthcare, and the
+automotive industry to name a few). This technology rests on the premise
+that facial expressions provide a direct access to individuals'
+subjective feelings also called ``emotional coherence''. Even if this
+premise is central to the modern mainstream approach of human emotion,
+recent research in affective science is challenging it. Once the two
+competing approaches are briefly described along with the predictions
+they respectively entail, an experiment testing these hypotheses will be
+presented and its results analyzed in order to provide empirical
+evidence to contribute to answer the question.
 
@@
 Expression of the Emotions in Man and Animals} {[}2{]}, facial
-expressions are conceived as a genuine displays of individuals' inner
+expressions are conceived as a genuine displays of an individual's inner
 emotional state. This hypothesis is used as a basis for the Basic
@@
 multiple systems, including {[}\ldots{}{]} facial expressions.''} {[}5,
-p. 49{]}. To cope with critics, several amendments have been made to the
-BET, increasing the number of basic emotions from six to seven {[}6{]}
-as well as adding the concept of ``display rules'' to explain cultural
-differences in the management of facial expressions {[}7{]}.
-
-Even if this theory obtained a popular support, it fails to explain how
-individuals can feel emotions without expressing them and how
+p. 49{]}. To respond to criticisms, several amendments have been made to
+the BET, increasing the number of basic emotions from six to seven
+{[}6{]} as well as adding the concept of ``display rules'' to explain
+cultural differences in the management of facial expressions {[}7{]}.
+
+Even if this theory obtained popular support, it fails to explain how
+individuals can feel emotions without expressing them, and how
 individuals can express emotions without feeling them, in instances in
@@
 situation and an appropriate response {[}10{]}, {[}11{]}. For the
-tenants of the constructivist approach, emotions are ``concepts'' based
-on past experiences and which are \emph{``a collection of embodied,
-whole brain representations that predict what is about to happen in the
-sensory environment, what the best action is to deal with impending
-events, and their consequences for allostasis''} {[}12, p. 12{]}.
-Following this assumption, faces are best conceived as tools displaying
-signals in social interactions {[}13{]}. These signals can convey
-individuals' motivations and readiness {[}14{]} or social messages
-{[}15{]}. Therefore, facial expressions are thought as behaviors which
-meaning is inferred by the observer. Findings support this observer
-dependence {[}16{]}, {[}17{]}. They show that to make meaning of another
-person's facial behavior, the perceiver relies in particular on her/his
-knowledge about emotion categories.
+proponents of the constructivist approach, emotions are ``concepts''
+based on past experiences and which are \emph{``a collection of
+embodied, whole brain representations that predict what is about to
+happen in the sensory environment, what the best action is to deal with
+impending events, and their consequences for allostasis''} {[}12, p.
+12{]}. Following this assumption, faces are best conceived as tools
+displaying signals in social interactions {[}13{]}. These signals can
+convey individuals' motivations and readiness {[}14{]} or social
+messages {[}15{]}. Therefore, facial expressions are thought as
+behaviors which meaning is inferred by the observer. Findings support
+this observer dependence {[}16{]}, {[}17{]}. They show that to make
+meaning of another person's facial behavior, the perceiver relies in
+particular on her/his knowledge about emotion categories.
 
@@
 
-Regarding emotional facial expression (EFE) recognition, the behaviorist
-approach assertions and the constructivist approach claims lead to two
-divergent predictions. The foremost postulates that, when triggered,
-each basic emotion is expressed by a prototypical face (non basic
-emotions being blends of the basic ones). Basic emotions are easily
-recognized by all human observers and emotional states are accessible by
-facial measurement. As the recognition of facial expressions is based on
-the identification of specific patterns of facial movements, it implies
-that EFE recognition of both human observers and automatic classifiers
-should be as accurate. The constructivist approach, in contrast, affirms
-that facial expressions do not provide a direct access to individuals'
-subjective feeling. Even if the face possibly moves during an emotional
-episode, facial muscle movements are not linked in a one-to-one manner
-to a specific discrete emotional experience {[}18, p. p418{]}. Instead,
+Regarding emotional facial expression (EFE) recognition, behaviorist and
+constructivist approaches lead to two divergent predictions. The former
+postulates that, when triggered, each basic emotion is expressed by a
+prototypical face (non basic emotions being blends of the basic ones).
+Basic emotions are easily recognized by all human observers and
+emotional states are accessible by facial measurement. As the
+recognition of facial expressions is based on the identification of
+specific patterns of facial movements, it implies that EFE recognition
+of both human observers and automatic classifiers should be as accurate.
+The constructivist approach, in contrast, affirms that facial
+expressions do not provide a direct access to individuals' subjective
+feelings. Even if the face possibly moves during an emotional episode,
+facial muscle movements are not linked in a one-to-one manner to a
+specific discrete emotional experience {[}18, p. p418{]}. Instead,
 emotions are mentally constructed by the perceiver and mental categories
@@
 
-The aim of the current paper is to investigate the link between the
+The aim of the current paper is to investigate the coherence between the
 subjective feeling of emotions and its recognition from facial
@@
 
-To evaluate the link between subjective feeling of emotions and their
-recognition from facial expressions, encoders were first recruited to
-perform an emotion elicitation task while their facial expression was
+To evaluate the coherence between subjective feeling of emotions and
+their recognition from facial expressions, encoders were first recruited
+to perform an emotion elicitation task while their facial expression was
 video recorded. Then, the videos of the encoders' faces were shown to
@@
 This picture is taken from a pilot with projects collaborators and all
-gave a consent for the publication of their photos and videos.}
+gave consent for the publication of their photos and videos.}
 \end{figure}
@@
 genuineness. All encoders gave their agreement on their data and video
-to be processed for research purpose only.
+to be processed for research purposes only.
 
@@
 but low (\(r = .24\), 95\% CI \([.19\), \(.29]\), \(t(1384) = 9.21\),
-\(p < .001\)). In order to identify differences according the emotional
-labels, encoders' subjective feeling is compared with human observers
-recognition in a confusion matrix (Figure
+\(p < .001\)). In order to identify differences according to the
+emotional labels, encoders' subjective feelings are compared with human
+observers' recognition in a confusion matrix (Figure
 \ref{fig:confusionMatrix_sr_hr}).
@@
 \includegraphics{ACII_2019_paper_files/figure-latex/confusionMatrix_sr_hr-1.pdf}
-\caption{\label{fig:confusionMatrix_sr_hr}Confusion matrix of between
-the emotion self-reported as being characteristic of the elicitation
-with the emotion recognized by the human observers.}
+\caption{\label{fig:confusionMatrix_sr_hr}Confusion matrix between the
+emotion self-reported as being characteristic of the elicitation with
+the emotion recognized by the human observers.}
 \end{figure}
@@
 except for \emph{disgust} (100\% of the videos self-reported). These
-results are far from those classicaly obtained in the literature for
+results are far from those classically obtained in the literature for
 emotional facial expression recognition which ranges between 60\% and
@@
 
-Sensitivity, specificity, precision and F1 score for each emotion
+Sensitivity, specificity, precision and F1 scores for each emotion
 reveals that \emph{happiness} has the highest coherence ratio whereas
@@
 \emph{disgust} are easily recognized. \emph{Anger} and \emph{sadness} as
-non-socialy desirable emotion may be have been felt but not expressed.
+non-socially desirable emotions may be have been felt but not expressed.
 
@@
 Similarly to the previous analysis, the overall correlation of
-recognition and non-recognition revealed a significant but very low link
-between self-reported emotions and automatic \nolinebreak classifier's
-recognition (\(r = .12\), 95\% CI \([.07\), \(.17]\),
-\(t(1384) = 4.50\), \(p < .001\)). A confusion matrix was used to
-compare encoders' subjective feeling with the emotion label recognized
-by the automatic classifier (Figure \ref{fig:confusionMatrix_sr_ar}).
+recognition and non-recognition revealed a significant but very low
+coherence between self-reported emotions and automatic
+\nolinebreak classifier's recognition (\(r = .12\), 95\% CI \([.07\),
+\(.17]\), \(t(1384) = 4.50\), \(p < .001\)). A confusion matrix was used
+to compare encoders' subjective feeling with the emotion label
+recognized by the automatic classifier (Figure
+\ref{fig:confusionMatrix_sr_ar}).
 
@@
 {[}0.22,0.38{]}; Kappa \nolinebreak = \nolinebreak 0.07) except for
-\emph{happiness} (44.4\% of the video self-reported) is watched.
+\emph{happiness} (44.4\% of the video self-reported) is evident.
 
 Surprisingly the automatic classifier incorrectly recognized as
-\emph{disgust} an important proportion of videos in which \emph{anger},
-\emph{happiness} and \emph{surprise} was the highest self-reported
-emotion (respectively 42.9\%, 25.0\% and 32.6\% of the videos
-self-reported). In parallel, the automatic classifier recognized as
-\emph{happiness} videos in which \emph{fear} and \emph{surprise} was the
-highest self-reported emotion (respectively 66.7\% and 32.6\% of the
+\emph{disgust} an significant proportion of videos in which
+\emph{anger}, \emph{happiness} and \emph{surprise} was the highest
+self-reported emotion (respectively 42.9\%, 25.0\% and 32.6\% of the
+videos self-reported). In parallel, the automatic classifier recognized
+as \emph{happiness} videos in which \emph{fear} and \emph{surprise} was
+the highest self-reported emotion (respectively 66.7\% and 32.6\% of the
 videos self-reported).
@@
 
-A comparable explanation involing the amount of undertermined video
-based on self-reports can be provided as the level of undetermined
+A comparable explanation involving the amount of undetermined video
+based on self-reports can be provided, as the level of undetermined
 emotions are very high for the self reports.
@@
 As previously mentioned, human observers appear to be more accurate than
-the automatic classifier to recognize individuals' subjective feeling
+the automatic classifier to recognize an individual's subjective feeling
 (human observers Accuracy \nolinebreak = \nolinebreak0.43; automatic
@@
 automatic classifier has a tendency to label as \emph{disgust} videos
-labeled as \emph{sadness} by human observers and as \emph{happiness}
-videos labeled as \emph{fear} by human observers
+labeled as \emph{sadness} by human observers, and as \emph{happiness}
+videos labeled as \emph{fear} by human observers.
 
@@
 
-Despite being one on the most investigated question in affective
-science, the coherence between emotion felt and facially displayed and
-facial expression recognized is a hot topic. To date no clear evidence
-has been found to definitely answer it. Yet, with the growing interest
-of industries and government to monitor individual's psychological
-states, this issue is under high pressure. The present research aims to
-provide some empirical data to the question. The faces encoders
+Despite being one of the most investigated questions in affective
+science, the coherence between emotion felt and facially displayed on
+the one hand, and facial expression recognized on the other is a hot
+topic. To date no clear evidence has been found to definitively solve
+the questions raised. Yet, with the growing interest of industries and
+government in monitoring individual's psychological states, this issue
+is under intense scrutiny. The present research aims to provide some
+empirical data to answer some of the questions posed. The faces encoders
 displayed when confronted with an emotional eliciting task were
@@
 recognition accuracy was the subjective feeling self-reported by the
-encoder once the elicitation task carried out. Results first reveal a
-low coherence between emotion felt and facial expression displayed.
-Secondly, results show a low accuracy for both humans and automatic
-classifier in identifying the inner emotional states of these encoders
-based on their facial expressions. Thirdly, human observers prove to be
-better at recognizing the emotion facially expressed than the automatic
-recognition tool is.
+encoder once the elicitation task was carried out. Results first reveal
+a low coherence between emotion felt and facial expression displayed.
+Secondly, results show low accuracy rates for both humans and the
+automatic classifier in identifying the inner emotional states of these
+encoders based on their facial expressions. Thirdly, human observers
+prove to be better at recognizing the emotion facially expressed than
+the automatic recognition tool is.
 
@@
 displays are not associated with a concordant emotional state, even any
-emotional state at all {[}24{]}, {[}25{]}. More and more evidences are
+emotional state at all {[}24{]}, {[}25{]}. More and more evidence is
 showing that facial expressions are in reality not expressing emotions
@@
 Beyond the present observations showing a weak coherence between
-subjective feelings and spontaneous facial expressions, this study
-provides some light on the controversy between the behaviorist and the
+subjective feelings and spontaneous facial expressions, this study sheds
+some light on the controversy between the behaviorist and the
 constructivist approaches. The first of those two lines of thinking
@@
 whereas if emotions are social constructs, as stated by the second line
-of thinking, human observer's should be better at perceiving emotions in
-face than automatic classifiers.
+of thinking, human observer's should be better at perceiving emotions
+expressed on the face than automatic classifiers.
 
@@
 human observers are more accurate than the automatic recognition tool to
-identify individuals' subjective feeling on the basis of their face.
+identify an individual's subjective feeling on the basis of their face.
 Moreover, mistakes made by human observers look less arbitrary to the
-ones made by the classifier. For instance, even if mix-up between
+ones made by the classifier. For instance, even if a mix-up between
 disgust and anger is sometimes reported in recognition studies,
@@
 classifier. Whereas this paradigm is longer and more complicated, it can
-lead to more robust results in reducing the forced-choice bais {[}27{]}.
-However this procedure can also reduce the human observers accuracy. In
+lead to more robust results in reducing the forced-choice bias {[}27{]}.
+However this procedure can also reduce the human observers' accuracy. In
 this regard, the results of the human observation could have been more
@@
 meaning from facial expressions. An alternative explanation relies in
-reducing the recognition bais involved in the classic recognition
-paradigm. Classic forced-choice paradigms obtain artificial high
+reducing the recognition bias involved in the classic recognition
+paradigm. Classic forced-choice paradigms obtain artificially high
 results, thus by using a more evolved approach observers' accuracy may
@@
 
-Considering the above, the results provide additional evidence that
-individuals' subjective feeling can not be inferred from facial
+Considering the above, the results provide additional evidence that an
+individual's subjective feeling can not be inferred from facial
 expressions and in our case invalidate the hypothesis of hardwired
@@
 ``non-basic'' emotion labels {[}28{]} as well as non-prototypical facial
-expression have been made in the developpement of automatic facial
-expression recognition tools. However this result suggests that
+expression have been made in the development of automatic facial
+expression recognition tools. However, these results suggest that
 automatic facial expression recognition tools should merely evaluate
-facial morphology features such as action units {[}29{]} rather than
-inferring supposedly emotional or affective states.
+facial morphology features such as action units (already evaluated in
+OpenFace {[}29{]}, Affectiva's Affdex {[}21{]} or Vicar Vision's
+FaceReader {[}30{]} to name a few) rather than inferring supposedly
+emotional or affective states.
 
@@
 \hypertarget{references}{%
-\section*{VI. References}\label{references}}
-\addcontentsline{toc}{section}{VI. References}
+\section{VI. References}\label{references}}
+
+\footnotesize
 
@@
 
+\leavevmode\hypertarget{ref-den2005facereader}{}%
+{[}30{]} M. Den Uyl and H. Van Kuilenburg, ``The facereader: Online
+facial expression recognition,'' in \emph{Proceedings of measuring
+behavior}, 2005, vol. 30, pp. 589--590.
+
 \end{document}

### 2019-04-22 ACII_2019_paper/ACII_2019_paper.Rmd
final

@@
   in_header: preamble-latex.tex
+
+template: C:/Users/dupred/Desktop/Projects/dynemo_feel_express/ACII_2019_paper/conference_041818.tex
 ---

### 2019-04-23 ACII_2019_paper/ACII_2019_paper.Rmd
typo

@@
 Based on the behaviorist approach initiated by Darwin in *The Expression of the Emotions in Man and Animals* [@darwin1872expression], facial expressions are conceived as a genuine displays of an individual's inner emotional state.
-This hypothesis is used as a basis for the Basic Emotion Theory (BET) which states that a set of six emotions are universally displayed and are genetically hardwired not only in humans [@ekman1992argument] but also in different animal species [@de2019mama].
-According to this view, *“when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including […] facial expressions.”* [@ekman2007directed, 49].
-To respond to criticisms, several amendments have been made to the BET, increasing the number of basic emotions from six to seven [@ekman1988universality] as well as adding the concept of “display rules” to explain cultural differences in the management of facial expressions [@ekman1987universals].
+This hypothesis is used as a basis for the Basic Emotion Theory which states that a set of six emotions are universally displayed and are genetically hardwired not only in humans [@ekman1992argument] but also in different animal species [@de2019mama].
+According to the view of the Basic Emotion Theory, *“when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including […] facial expressions.”* [@ekman2007directed, 49].
+To respond to criticisms, several amendments have been made to the Basic Emotion Theory, increasing the number of basic emotions from six to seven [@ekman1988universality] as well as adding the concept of “display rules” to explain cultural differences in the management of facial expressions [@ekman1987universals].
 

### 2019-04-23 ACII_2019_paper/ACII_2019_paper.tex
typo

@@
 emotional state. This hypothesis is used as a basis for the Basic
-Emotion Theory (BET) which states that a set of six emotions are
-universally displayed and are genetically hardwired not only in humans
-{[}3{]} but also in different animal species {[}4{]}. According to this
-view, \emph{``when emotions are aroused by perception of a social event,
-a set of central commands produce patterned emotion-specific changes in
-multiple systems, including {[}\ldots{}{]} facial expressions.''} {[}5,
-p. 49{]}. To respond to criticisms, several amendments have been made to
-the BET, increasing the number of basic emotions from six to seven
-{[}6{]} as well as adding the concept of ``display rules'' to explain
-cultural differences in the management of facial expressions {[}7{]}.
+Emotion Theory which states that a set of six emotions are universally
+displayed and are genetically hardwired not only in humans {[}3{]} but
+also in different animal species {[}4{]}. According to the view of the
+Basic Emotion Theory, \emph{``when emotions are aroused by perception of
+a social event, a set of central commands produce patterned
+emotion-specific changes in multiple systems, including {[}\ldots{}{]}
+facial expressions.''} {[}5, p. 49{]}. To respond to criticisms, several
+amendments have been made to the Basic Emotion Theory, increasing the
+number of basic emotions from six to seven {[}6{]} as well as adding the
+concept of ``display rules'' to explain cultural differences in the
+management of facial expressions {[}7{]}.
 

### 2019-04-23 ACII_2019_paper/ACII_2019_paper.Rmd
submission

@@
   institution:
-    - name: Dublin City University
-      department: Business School
-      location: Dublin, Ireland
-      email: damien.dupre@dcu.ie
+    - name: #Dublin City University
+      department: #Business School
+      location: Blank for review
+      email: #damien.dupre@dcu.ie
       mark: 1
       author:
-        - name: Damien Dupré
-    - name: Université Grenoble Alpes
-      department: Laboratoire Inter-universitaire de Psychologie
-      location: Grenoble, France
+        - name: #Damien Dupré
+    - name: #Université Grenoble Alpes
+      department: #Laboratoire Inter-universitaire de Psychologie
+      location: Blank for review
+      email: #anna.tcherkassof@univ-grenoble-alpes.fr
       mark: 2
       author:
-        - name: Anna Tcherkassof
-          email: anna.tcherkassof@univ-grenoble-alpes.fr
+        - name: #Anna Tcherkassof
+
 abstract: |
@@
 fig_caption: yes
-keep_tex: yes
 number_sections: yes
@@
   in_header: preamble-latex.tex
-
-template: C:/Users/dupred/Desktop/Projects/dynemo_feel_express/ACII_2019_paper/conference_041818.tex
 ---
-
+<!-- template: C:/Users/dupred/Desktop/Projects/dynemo_feel_express/ACII_2019_paper/conference_041818.tex -->
 \begin{IEEEkeywords}
@@
 According to the view of the Basic Emotion Theory, *“when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including […] facial expressions.”* [@ekman2007directed, 49].
-To respond to criticisms, several amendments have been made to the Basic Emotion Theory, increasing the number of basic emotions from six to seven [@ekman1988universality] as well as adding the concept of “display rules” to explain cultural differences in the management of facial expressions [@ekman1987universals].
+To respond to criticisms, several amendments have been made to the Basic Emotion Theory, increasing the number of basic emotions from six to seven [@ekman1988universality] as well as adding the concept of “display rules” to explain cultural differences in the management of facial \linebreak expressions [@ekman1987universals].
 
@@
 
-The authors would like to thank Brigitte Meillon and Jean Michel Adam who developed the software used to collect and preprocess human observer results.
+[blank for review]
+<!-- The authors would like to thank Brigitte Meillon and Jean Michel Adam who developed the software used to collect and preprocess human observer results. -->
 

### 2019-04-23 ACII_2019_paper/ACII_2019_paper.tex
submission

@@
   %% -- beg for/affiliation.institution.author
-Damien Dupré %% -- end for/affiliation.institution.author
+ %% -- end for/affiliation.institution.author
 }
-\IEEEauthorblockA{Business School\\
-Dublin City University\\
-Dublin, Ireland
-\\damien.dupre@dcu.ie
+\IEEEauthorblockA{\\
+Blank for review
+  %% -- beg for/affiliation.institution.author
+ %% -- end for/affiliation.institution.author
 }
@@
   %% -- beg for/affiliation.institution.author
-Anna Tcherkassof %% -- end for/affiliation.institution.author
+ %% -- end for/affiliation.institution.author
 }
-\IEEEauthorblockA{Laboratoire Inter-universitaire de Psychologie\\
-Université Grenoble Alpes\\
-Grenoble, France
+\IEEEauthorblockA{\\
+Blank for review
   %% -- beg for/affiliation.institution.author
-\\anna.tcherkassof@univ-grenoble-alpes.fr
  %% -- end for/affiliation.institution.author
@@
 concept of ``display rules'' to explain cultural differences in the
-management of facial expressions {[}7{]}.
+management of facial \linebreak expressions {[}7{]}.
 
@@
 
-The authors would like to thank Brigitte Meillon and Jean Michel Adam
-who developed the software used to collect and preprocess human observer
-results.
+{[}blank for review{]}
 

### 2019-07-22 ACII_2019_paper/ACII_2019_doc.Rmd
paper doc

--- 2019-04-23 ACII_2019_paper/ACII_2019_paper.Rmd
@@
 bibliography: mybibfile.bib
-output: rticles::ieee_article
+output: word_document
+#output: rticles::ieee_article
 csl: http://www.zotero.org/styles/ieee
@@
 ```{r models_img, fig.height=5, fig.cap="\\label{fig:models_img}Comparison of behaviorist and social constructivist approaches according to the expected differences in emotion recognition."}
-raster::stack(here::here("./img/models.png")) %>%
-  raster::plotRGB()
+# raster::stack(here::here("./img/models.png")) %>%
+#   raster::plotRGB()
+knitr::include_graphics(here::here("./img/models.png"))
 ```
@@
 This picture is taken from a pilot with projects collaborators and all gave consent for the publication of their photos and videos."}
-raster::stack(here::here("./img/dynemo_img.png")) %>%
-  raster::plotRGB()
+# raster::stack(here::here("./img/dynemo_img.png")) %>%
+#   raster::plotRGB()
+knitr::include_graphics(here::here("./img/dynemo_img.png"))
 ```

### 2019-11-01 manuscript_apa.Rmd
references

---
title             : "The Emotion–Facial expression link: Evidence from Human and Automatic Expression Recognition"
shorttitle        : "The Emotion–Facial expression link"

author: 
  - name          : "Anna Tcherkassof"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Bâtiment Michel Dubois, Université Grenoble Alpes, 1251 Avenue Centrale, 38400 Saint-Martin-d'Hères"
    email         : "anna.tcherkassof@univ-grenoble-alpes.fr"
  - name          : "Damien Dupré"
    affiliation   : "2"

affiliation:
  - id            : "1"
    institution   : "University Grenoble Alpes, Grenoble, France"
  - id            : "2"
    institution   : "Dublin City University, Dublin, Ireland"

abstract: |
  While it has been taken for granted in the development of several automatic facial expression recognition tools, the question of the coherence between subjective feelings and facial expressions is still a subject of debate. On one hand, the Basic Emotion View conceives emotions as genetically hardwired and therefore being genuinely displayed through facial expressions. On the other hand, the constructivist approach conceives emotions as socially constructed; the emotional meaning of a facial expression being inferred by the observer. In order to evaluate the coherence between the subjective feeling of emotions and their recognition based on facial expression, 232 videos of encoders recruited to carry out an emotion elicitation task were annotated by 1383 human observers as well as by an automatic facial expression classifier. Results show low accuracy of human observers and of the automatic classifier to infer the subjective feeling from the facial expressions displayed by encoders. They also show a weak consistency between self-reported emotional states and facial emotional displays. Based on these results, the hypothesis of genetically hardwired emotion genuinely displayed is difficult to support, whereas the idea of emotion and facial expression socially constructed appears to be more likely. Accordingly, automatic emotion recognition tools based on facial expressions should be questioned.

keywords          : "Facial expression, self-report, human observer, automatic recognition."
wordcount         : "7920"

bibliography      : ["manuscript_ref.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : 
  bookdown::pdf_book:
    base_format: papaja::apa6_pdf

header-includes:
  - \usepackage{booktabs}
  - \usepackage{float}
  - \usepackage{tabu}
  - \usepackage{wrapfig}
  - \usepackage[none]{hyphenat}
  - \floatplacement{figure}{H}
  - \usepackage{flushend}
  - \usepackage{biblatex}
---

```{r setup, include = FALSE}
library(papaja)
library(kableExtra) # Tables
library(tidyverse) # Data wrangling & grammar of graphics
library(here) # Locate files relative to project root
library(caret) # Confusion matrices
```

```{r analysis-preferences}
set.seed(123) # Seed for random number generation
options(scipen = 999)# Disable scientific number format
knitr::opts_chunk$set(
  cache.extra = knitr::rand_seed, 
  message = FALSE, 
  warning = FALSE, 
  error = FALSE, 
  echo = FALSE
  )
```

```{r data}
# self_report_data
source(here::here("script/data_wrangling_self_report.R"))
self_report_score <- self_report_score %>% 
  dplyr::select(C_Video, sr_emotion = emotion, sr_score = value)
# human_recognition_data
source(here::here("script/data_wrangling_human_recognition.R"))
human_recognition_score <- human_recognition_score %>% 
  dplyr::select(C_Video, hr_emotion = emotion, hr_score = confidence_score)
# automatic_recognition_data
source(here::here("script/data_wrangling_automatic_recognition.R"))
automatic_recognition_score <- automatic_recognition_score %>% 
  dplyr::select(C_Video, ar_emotion = emotion, ar_score = confidence_score) %>% 
  dplyr::filter(C_Video %in% ppt_per_video$C_Video)
```

# Introduction

With the development of commercial automatic facial expression recognition tools [@dupre2018accuracy], industries and governments are gradually implementing this technology in order to track humans’ emotions in various scenarios (*e.g.*, in marketing, healthcare, and the automotive industry to name a few). This technology rests on the premise that facial expressions provide a direct access to individuals’ subjective feelings, and that one can read the emotions displayed on a person’s face as in an open book. This premise is derived from the Basic Emotion View which Paul Ekman is one of the famous proponents of this perspective [@ekman1987universals; @ekman1988universality; @ekman1992argument; @ekman2007directed]. Even if this premise is central to the modern mainstream approach of human emotion, recent research in affective science is challenging it. Once the Basic Emotion View is briefly described, its foremost criticisms are synthetically exposed. Inconsistencies and unpredicted findings research has uncovered come both from field observations and laboratory experiments. They relate to two different issues. The first issue is the one of the sender’s production of spontaneous facial displays. The second issue has to do with the viewer’s interpretation of these facial displays. Building on this evidence, and as a further attempt to shed light on the emotion–face link, the present study is finally exposed. Its main goal is to tackle these two issues at the same time. It aims at identifying the extent to which ordinary people’s experienced emotions are displayed through identifiable emotional facial expressions (EFE) recognized as such both by human individuals and by automatic facial expression recognition tools.

## The Basic Emotion View

The belief that facial expression is linked to emotional states can be traced back to Darwin in *The Expression of the Emotions in Man and Animals* [@darwin1872expression]. Countless studies have fortified the idea that emotional states are inherently coupled to a set of prototypic facial expressions (*e.g.*, [@ekman2011meant]). The Basic Emotion View holds that facial expressions are genuine displays of an individual’s inner emotional state. More specifically, a set of six emotions (*fear*, *anger*, *surprise*, *disgust*, *sadness* and *joy*) are universally displayed and are genetically hardwired not only in humans [@ekman1992argument], but also in different animal species [@de2019mama]. According to the view of the Basic Emotion View, *“when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including […] facial expressions.”* [@ekman2007directed, 49]. To respond to criticisms, several amendments have been made to the Basic Emotion View, increasing the number of basic emotions from six to seven [@ekman1988universality] as well as adding the concept of “display rules” to explain cultural differences in the management of facial expressions [@ekman1987universals].

Relatively few studies have actually tested Basic Emotion View’s fundamental claim regarding the facial production (so-called ‘expression’) of emotions. @rosenberg1994coherence reported the first evidence of coherence between self-report of emotion and displayed facial expressions. Participants were showed emotionally evocative films and reported their own emotions using a moment-by-moment reporting procedure. Analysis of participants’ facial expressions and reports of emotions showed that there was a high degree of temporal linkage and emotional agreement between facial expressions and self-reports. Notwithstanding this study, few have used purely descriptive methods (such as electromyography or objective coding systems – FACS or others) to identify and to measure the actual changes in the face when a given emotion is felt (see @wagner1997methods, for methodological points). Much of the research has focused instead on the recognition of EFE, that is on the issue of the viewer’s interpretation of facial displays. As people have been thought to display their internal state, EFE supposedly serve as a window into the emotions of others. Viewers “recognize” the facial displays and infer the corresponding emotional state. The Readout Hypothesis [@buck1985prime] formalizes this folk theory. According to it, facial expressions have evolved to provide “an external readout of those motivational-emotional processes that have had social implications during the course of evolution” (pp. 396-397). Thus, as readouts of emotional states, spontaneous expressive displays are directly accessible to other organisms. Research on the inference of emotions from facial expressions has established that viewers show considerable agreement that the so-called basic emotions—happiness, anger, fear, disgust, sadness, and surprise—are associated with specific facial displays (*e.g.*, reviews from @elfenbein2002universality; @russell1994there). These findings are considered by many as constituting strong evidence for the Basic Emotion View.

Yet, in spite of the popular support it has received, the empirical data called upon by the Basic Emotion View remains unpersuasive. Notably, it fails to explain, in instances in which display rules cannot be called upon, how individuals can feel emotions without expressing them or how individuals can express emotions without feeling them, and why observers are not totally accurate in recognizing facial expressions of basic emotions, among others. The following  is a summary of key findings and conclusions both from field observations and laboratory experiments on spontaneous expression of emotions, all of them strongly subverting the Basic Emotion viewpoint (for reviews, @fernandez2013emotion; @reisenzein2013coherence; @russell2003facial). Mains issues regarding recognition studies are also swept across afterwards.

## Spontaneous Facial Expression in naturalistic studies

Naturalistic studies look at the ecological frequency of co-occurrence of certain emotions and facial displays [fernandez2013emotion]. Though they cannot be considered as a straightforward test of the triggering role of emotion on the facial behavior, they have the advantage of considering situations that cannot be created in a laboratory. For instance, the ethological study of @kraut1979social contradicts the Basic Emotion View premise that a smile is the major component of a facial display associated with and caused by feelings of joy or happiness. Naturalistic observation at a bowling alley showed that bowlers do not necessarily smile after scoring a spare or a strike (a situation likely to elicit a positive emotion). Rather, they often smile when interacting with other people. More generally, examination of bowlers’ facial display showed that they rarely smiled while facing the pins but often smiled when facing their friends. These findings were confirmed by @ruiz2003spontaneous who analyzed the facial displays of bowlers (after a strike) and soccer fans (after their team scored) reporting happiness in a field study. They found that the probability of smiling was very low when participants were not interacting with someone else. These findings have been supported in other realistic field settings such as the one of @fernandez1995smiles suggesting that happiness is not a sufficient cause of smiling. They watched extremely happy gold medalist athletes displaying facial expressions of sadness (sometimes associated with tears) during their Olympics awards ceremonies. More specifically, winners showed Duchenne smiles and other types of smiles when they were interacting with other people during the awards ceremony, but smiles were scarce or nonexistent when waiting behind the podium and/or when turning toward the flagpoles and focusing their attention on the flags and the national anthem. @crivelli2015smiles also found that the strongest predictor for the occurrence of Duchenne smiles in judo winners is when they are engaged in social interaction and not when they just won their match.

Happiness (or joy) is not the only emotion weakly associated with the predicted facial expression. A naturalistic study conducted by @scherer1997lost in an airport’s baggage handling office showed that passengers claiming for their lost luggage displayed very few facial expressions of negative emotions while self-reporting subjective feeling states of anger or sadness among others. The covariation between passengers’ self-ratings and the claims agents’ attributions of the passengers’ emotions was very low. Another refutation of the Basic Emotion View’s predictions regarding the link between emotions and facial expressions is provided by naturalistic observations of infants’ productions of facial expressions. It is often acknowledged that adults regulate their expressive behavior. In accordance with various display rules (personal, social, cultural ones), they exert a control over the (supposed) automatic readouts of their emotions. Developmental studies provide relevant settings in which display rules are inoperative. Camras and her colleagues observed that facial expressions of negative emotions were displayed in circumstances that were unlikely to have elicited those emotions. For instance, infants displayed "fear" expressions in settings not related to fear [@camras1991development]. @bennett2002facial videotaped a sample of 4-month-old infants during tickle, sour taste, jack-in-the-box, arm restraint, and masked stranger situations. Infants displayed a variety of facial expressions in each eliciting situation. Yet, more infants exhibited positive (joy and surprise) than negative facial expressions (anger, disgust, fear, and sadness) across all situations –except sour taste. No evidence for emotion-specific facial expressions corresponding to anger, fear, and sadness was obtained. Camras and her colleagues also observed that 11-month-old European-American, Chinese, and Japanese infants did not display distinct negative emotion-specific patterns of facial muscles in response to two elicitors meant to induce fear and anger [@camras2007infants]. Thus, these findings in naturalistic settings provide little support for the Basic Emotion View.

## Spontaneous Facial Expression in laboratory studies

Laboratory findings also support field studies. As stressed by @reisenzein2013coherence, experimental studies permit both better control of various factors (*e.g.*, emotion elicitors) and tests about likely moderators of the emotion–facial expression link (*e.g.*, the social context). As an example, a strong disconfirmation of the Basic Emotion View’s premise is put forward by Reisenzein’s studies on surprise (*e.g.*, @reisenzein2000exploring). In eight controlled laboratory situations, surprise was induced by @reisenzein2006evidence by establishing and then invalidating a set of beliefs concerning the experimental events (*e.g*., the unexpected appearance of a picture of one's own face as the last picture in a series of portraits that had to be rated). Visible or EMG-detected facial displays of surprise occurred only in few participants. Yet, most participants reported subjective feelings of surprise and most believed that they had shown a strong surprise facial expression. @schutzwohl2012facial also observed similarly low frequencies of surprise facial expressions when their participants, after leaving the laboratory, unexpectedly found themselves not in the corridor but in a new room with green walls and a red office chair. Less than a quarter of the participants displayed an expression of surprise: only 5% showed widened eyes, raised eyebrows, and opened mouth (*i.e.* the complete expression of surprise according to the Basic Emotion View) and 17%  showed widened eyes and raised eyebrows. Again, participants overestimated their surprise expressivity.

Studies on happiness and related positive emotions (*e.g.*, sensory pleasantness) also show a low coherence between emotion and facial display [@duran2017coherence; @reisenzein2013coherence]. In fact, only experiments on amusement provide a fairly strong association between emotion and smiling. When confronted to humorous events (*e.g.*, jokes, being tickled), the number of facially reactive participants (smiling and laughing) is indeed quite fair, whereas when confronted to other positive emotions (*e.g.*, happiness), few participants show the expected facial expressions [for instance, the Duchenne smile or any kind of “happy” expression yet reporting being happy; see @mehu2007smiles; @lee2002effect, social condition]. More largely, the meta-analysis conducted by @duran2017coherence on the degree of statistical covariation between emotions and facial expressions provides conclusive evidence. First of all, the commonly called basic emotions (happiness/amusement, sadness, fear, anger, disgust and surprise), on average, are weakly correlated with the specific configuration of facial muscles that the Basic Emotion View ascribes to them (the correlation even drops when happiness/amusement studies are excluded). Very few participants who relived an experience of sadness [*e.g.*, @tsai2002emotional] were reactive, only a small number of them showing components of a sad facial expression (*e.g.*, oblique eyebrows, lip corners pulled down). Regarding fear, spider phobic participants recruited by @vernon2002disgust, for instance, were exposed to a live tarantula. About one third of them only displayed some components of the prototypical fear expression (eye widening, brow raising and knitting, etc.). Similar observation is made for anger. The proportion of facially reactive participants (*i.e.* displaying at least one component among frowning, lid/lip tightening...) reporting anger in @johnson2010smile and @tsai2002emotional studies does not exceed 35%. The same proportion of reactive participants confronted to disgusting elicitors is observed [*e.g.*, @ekman1980facial; @fernandez1997spontaneous]. In any case neither an insufficient intensity of the emotion to cause a facial expression nor the intervention of display rules (nor measurement issues) can explain the low emotion–facial expression consistency.

Hence, laboratory studies show that facial expressions of emotion are often not displayed in situations in which the Basic Emotion View would predict them to be expressed. Moreover, when they are indeed displayed, they are more often displayed partially than completely. All in all, research on the spontaneous expression of emotions does not yield strong support for the Basic Emotion View. The available evidence steadily indicates weak links between emotions and their predicted facial expressions both in natural and in semi-naturalistic settings.

## Emotional Facial Expression Recognition studies

The Basic Emotion View postulates that, when triggered, each basic emotion is expressed by a prototypical face (non-basic emotions being blends of the basic ones). In return, the recognition of emotional facial expressions (EFE) is claimed to be based on the identification of specific patterns of facial movements associated with each emotion, as if expression and recognition were the two sides of a same coin. Seemingly compelling evidence support this claim, sustaining the belief that one can read a face in the same way one reads the lines of a book. Besides, recognition systems rely on such a principle and are considered as an objective coding tool because they are based on the identification of specific muscular changes in the face. Many people highly agree that so-called basic emotions are associated with specific facial configurations (Ekman, 2017; Elfenbein & Ambady, 2002; Russell, 1994) and also agree that many believe that it is a strong evidence for the Basic Emotion View. Moreover, it implies that EFE recognition of both human observers and automatic classifiers should be as accurate.

Yet, some researchers have highlighted the limitations of Basic Emotion View empirical research. Among others, evidence has been questioned on methodological grounds [*e.g.*, @russell1994there]. The response format usually used in recognition studies (*i.e*. forced choice: selection of one word from a pre-specified list of emotion labels), notably, leads to a biased consensus as evidenced by @russell1993forced. Depending on the list of emotion labels at participants’ disposal, EFE of sadness can easily be categorized as sad expressions as well as fear expressions... @russell1987relativity also showed the same facial expression can be seen as expressing different types of emotions, depending on what other faces are seen. @digirolamo2017emotion conducted seven experiments that establish that high agreement between participants can be an artifact of the standard method. Thus, results gathered with forced choice cannot demonstrate the univocal link between emotion and facial expressions claimed by the Basic Emotion View. Using alternative recognition methods [emotion satiation procedure, face-matching task, sorting task; *e.g.*, @lindquist2006language; @gendron2014perceptions], it has been shown on the contrary that facial muscle movements are not linked in a one-to-one manner to a specific discrete emotional experience (2017,p. 418). Instead, emotions are probably mentally constructed by the perceiver and mental categories of emotions are needed to accurately categorize facial movements, among other things (*e.g.*, contextual information).

To the methodological limitations contaminating the hundreds of studies apparently supporting the Basic Emotion View, a stimulus bias must be added. Facial stimuli used in experiments also constitute a methodological bias because they are unrepresentative of ordinary facial expressions. Basic Emotion View empirical evidence is based for the most part on methods using a static and unnatural material, namely, still photographs of posed facial expressions of emotion (*e.g.*, intentionally encoded by the sender). This kind of methodology raises questions about its ecological validity and the generalizability of the results to real interpersonal emotional communication [*e.g.*, @tcherkassof2007facial]. Indeed, a number of pieces of evidence indicate that research cannot content itself with data collected with static and posed material. These data come from researches studying the case of dynamic and/or spontaneous facial expressions of emotion. They show that the dynamic aspects of facial movement are likely to be of importance [*e.g.*, @kamachi2013dynamic]. @cohn2003timing have shown that spontaneous smiles are of smaller amplitude and have a more consistent relation between amplitude and duration than deliberate smiles. @hess1990differentiating have also pointed out the importance of the dynamics of facial movements, and particularly the irregularity, or phasic changes, of the expressions’ unfolding. Thus, the motion of facial expression provides observers with other information than the one provided by static expressions. It may be that differences in the social information displayed by static and dynamic expressions leads to facial recognition differential effects. Regarding the issue of spontaneous vs posed expressions (the latter are overused in experiments). As @meillon2010dynemo conclude, EFE have been typically studied as static displays. As a consequence, even though the central role of the dynamics of facial expressions is endorsed, little is still known about the temporal course of facial expressions. Furthermore, studied EFE exhibit emotions simulated or posed by actors. Yet, the lack of spontaneity and naturalness of this material constitutes a serious objection raised against such studies [@kanade2000comprehensive].

All in all, as many doubts can be raised about the standard method, experiments conducted with such a method cannot be considered as providing solid empirical support to the Basic Emotion View. Based on the numerous methodological criticisms, but also theoretical, addressed to this view, alternative conceptions have emerged. Among them, the constructivist approach is gaining in importance. The constructivist approach represents a totally different way of understanding the emotion–facial expression link. It affirms that facial expressions do not provide a direct access to individuals’ subjective feelings. Therefore, instead of considering that emotions can be “read” on facial displays, it claims that the emotion is “in the eye” of the perceiver [@barrett2019emotional].

## Constructivist Approach

Starting from the empirical evidence suggesting that spontaneous facial expressions in ordinary life are equivocal, @dols2017natural argues in favor of a pragmatic conception of natural facial displays. He makes a plea for the idea that natural facial displays, rather than "saying" – because they do not have a specific meaning he claims – “make” things. Facial expressions are actions in a communicative interaction. They do not express emotions but they *"prompt, on the receiver’s side, important inferences about the context, the sender, and the course of the interaction between sender and receiver"* [@dols2017natural, 466, underlined by the author]. As such, the fact that facial displays are able to signal emotions is a byproduct of one of its main function, the one of implementing actions performing practical ends. Therefore, expressive displays hold a motor intention [@pacherie2003modes]. They implement their aim which is their motor intention. They are not primarily communicative signals and even less the outlet of an internal state. Facial displays are (parts of) pragmatic actions aiming at orienting the person’s relation to its environment, for instance maintaining, breaking,  restoring, etc., the relation [@frijda2012recognition]. Facial displays are not “recognized” in semantic terms but are perceived as intentional actions. In the face of the continual flow of uninterrupted facial movements, the perceiver sees behaviors directed towards a goal. S.he translates the continuous flow of movements into coordinated sequences of actions holding a beginning and an end. Facial displays are not simple strings of action units which morphological configuration would be the prototype of a given emotion, and consequently identified as such [@ekman1987universals]. They are best conceived as a Gestalt, the same way as a string of musical notes establishes a melody [@tcherkassof2014emotions]. This is why even unauthentic facial displays can yet be recognized as emotional expressions. Guillaume Duchenne de Boulogne explains that the artist who has shaped the famous Laocoon antic sculpture, exhibited in the Vatican’s museum, has made a modelling mistake since no face can display its emotional expression [@duchenne1876mecanisme]. Indeed, no muscular contraction can produce it. He even rectifies the "mistake" by presenting a statue which face is shaped according to the physiology of facial expressive movements. His demonstration gives food for thought. Even though no objective coding system can correctly code the discordant facial features of the Laocoon’s face, yet anyone can easily recognize the suffering and despair he admirably expresses. This example goes along the lines of the constructivist approach.

The constructivist approach claims that facial displays are behaviors which meaning is inferred by the perceiver. Findings support this observer dependence [@lindquist2013s; niedenthal2017embodied]. They show that to make meaning of another person’s facial behavior, the perceiver relies in particular on her/his knowledge about emotion categories. For instance, Gendron and her colleagues used a face-sorting task allowing them to manipulate the influence of emotion concepts on how facial expressions were perceived. They conducted their experiment among U.S. participants and Himba participants from remote regions of Namibia [@gendron2014perceptions] and Hazda participants of the Eastern Rift Valley of Tanzania [@gendron2018emotion] both groups with limited exposure to Western culture. Gendron and her colleagues demonstrated that facial expressions were not universally “recognized” in discrete emotional terms. Indeed, when Himba and Hazda participants did not have emotion concepts at their disposal to structure perception, they perceived the facial expressions as behaviors (*e.g.*, looking, smelling) that didn’t have a necessary relationship to emotions. They didn’t infer inner states (such as emotional feelings) but they rather proceeded with action identification that pointed out the functions of behaviors [see also @crivelli2017recognizing for similar observations among a small-scale society of Papua New Guinea]. The constructivist approach considers that specific emotion categories, as conceptualized by Western cultures’ knowledge, are casted on the perceived face to make meaning of the sender’s facial displays. Following this approach, faces convey a range of information essential for social communication. They are best conceived as tools displaying signals in social interactions [@crivelli2018facial]. These signals can convey individuals’ motivations and readiness [@frijda1997facial] or social messages [@fridlund1994human]. As for emotional meaning, more specifically, it is shaped by the perceiver according to the specific context in which the facial displays are observed.

Having reached this point, one can assert that numerous questions regarding the link between emotions and facial expressions remain unanswered. The two main competing approaches to facial expressiveness, the Basic Emotion View and the constructivist approach, entail completely opposite predictions regarding the decoding of facial expressions, as evidenced above. The present study aims at examining these predictions in order to provide empirical evidence to allow the discussion to evolve. To date, no systematic study has looked at the same time at facial expressions spontaneously displayed in reaction to emotional triggers and how they are decoded, both by human observers and by automatic emotion recognition tools based on the detection of facial muscular configurations. This study fills that gap. It intends to investigate the consistency between the subjective feeling of emotions and its recognition from facial expressions. Spontaneous and dynamic facial reactions to emotional elicitations are under consideration to ensure the generalizability of the results to emotional behaviors in ordinary life. More specifically, this study aims to examine the recognition of EFE produced by ordinary people during situations judged and/or self-reported to involve different emotions. It (a) examines consistency between ordinary people’s self-reported emotional experience and observers’ judgments of these ordinary people’s EFE, (b) examines consistency between ordinary people’s self-reported emotional experience and an automatic classifier’s analysis of these ordinary people’s EFE. In other words, it is interested in how people actually move their faces to express self-reported emotions, in how human observers accurately infer the expresser’s emotional state, and in how automatic recognition accurately code the expresser’s emotional state. It is expected a human observers’ superior ability to exactly recognize EFE as compared to automatic EFE recognition tools.

# Methods

To evaluate the consistency between subjective feeling of emotions and their recognition from facial expressions, encoders were first recruited to perform an emotion elicitation task while their facial expression was video recorded. In order to reduce the likelihood of facial control, the encoders were alone in the room and were filmed by hidden cameras, so they had no reason to comply with social display rules. Then, the videos of the encoders’ faces were shown to human observers and were also analyzed by an automatic classifier in order to identify which emotion was displayed.

## Emotion Elicitation

```{r self_report_method}
self_report_gender <- self_report_data %>% 
  dplyr::group_by(genre_c) %>% 
  dplyr::summarise(n = n())

self_report_age <- self_report_data %>%
  dplyr::summarise(Age_m = mean(age), Age_sd = sd(age)) %>% 
  dplyr::mutate_all(round, 1)
```

For the emotion elicitation experiment, `r nrow(self_report_data)` encoding participants (`r self_report_gender[self_report_gender$genre_c == "F","n"]` females, `r self_report_gender[self_report_gender$genre_c == "H","n"]` males, *M*~age~ = `r self_report_age[,"Age_m"]`, *SD*~age~ = `r self_report_age[,"Age_sd"]`) were recruited to perform one out of 11 emotion elicitation tasks designed to trigger a positive, a specific negative or a neutral emotional state. Encoders’ faces were recorded using a hidden camera resulting `r nrow(self_report_data)` front facing 768x576 videos varying from `r min(metadata_video$ffprobe_duration)`s to `r max(metadata_video$ffprobe_duration)`s. These recordings form the DynEmo database [@tcherkassof2013dynemo].

After the emotion elicitation task, the encoders rated their subjective feeling on Likert scales from 0 ("not at all") to 5 ("strongly") related to six "basic" emotion labels (*i.e.*, *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as six "non-basic" emotion labels (*i.e.*, *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).

Finally, a debriefing session was performed to ensure that encoders were not durably affected by the emotion elicitation task. The debriefing was also used to check that encoders did not guess the real purpose of the experiment (*e.g.*, being filmed while they were performing an emotional elicitation task) to guarantee facial expressions’ genuineness. All encoders gave their agreement on their data and video to be processed for research purposes only.

## Human Facial Expression Recognition

```{r human_recognition_method}
unique_ppt_count <- human_recognition_data %>% 
  dplyr::group_by(source,SEXE_Juge, C_Juge) %>% 
  dplyr::summarise(n = n())

unique_video_count <- human_recognition_data %>% 
  dplyr::select(source,SEXE_Juge, C_Juge, C_Video) %>% 
  unique() %>% 
  dplyr::group_by(C_Video) %>% 
  dplyr::summarise(n = n())
```

For the human facial expression recognition method, `r nrow(unique_ppt_count)` student participants were recruited to annotate `r nrow(unique_video_count)` out of the `r nrow(self_report_data)` videos, therefore only the `r nrow(unique_video_count)` annotated videos were analyzed in this study. Because videos have different durations, participants had to annotate a series of video corresponding to 30min long in total. Each video was annotated `r mean(unique_video_count$n) %>% round(0)` times on average (*SD* = `r sd(unique_video_count$n) %>% round(0)`).

The annotation of facial expressions was performed on-site using *Oudjat*, a software for designing video annotation experiments [@dupre2015oudjat]. For each video, the annotation procedure followed two steps. First, the participants had to identify the emotional sequences by pressing the space bar of their keyboard to indicate the beginning and the end of the emotional sequences while watching the video. Second, the participants watched each emotional sequence previously identified and labeled the sequence using one of the 12 emotions proposed including six "basic" emotion labels (*i.e.*, *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) and six "non-basic" emotion labels (*i.e.*, *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*). They also had the possibility to indicate that the sequence was expressing none of the proposed emotion labels.

This annotation procedure results in a uni-dimensional time-series for each video per human observer identifying for each second of the video which emotion was recognized. Then, time-series corresponding to the same video were aggregated to calculate the proportion of human observers $x_{video_{i}.label_{j}.t_{k}}$ for each second of the video per emotional label (EQ\ref{eq:1}).

\begin{equation}
\label{eq:1}
x_{video_{i}.label_{j}.t_{k}} = \frac{n_{video_{i}.label_{j}.t_{k}}}{n_{video_{k}}}
\end{equation}

where *i* is one of the `r nrow(unique_video_count)` videos, *j* is one of the six “basic” emotion labels, *k* for each second of the video.

## Automatic Facial Expression Recognition

The `r nrow(unique_video_count)` annotated video were processed with Affdex (SDK v3.4.1). Affdex is an automatic facial expression recognition classifier developed and distributed by Affectiva is a spin-off company resulting from the research activities of MIT media lab created in 2009 [@mcduff2016affdex]. Affdex’s algorithm uses Histogram of Oriented Gradient (HOG) features and Support Vector Machine (SVM) classifiers in order to recognize facial expressions. For each video frame, Affdex identifies the probability $p_{video_{i}.label_{j}.t_{k}}$ from 0 to 100 (rescaled to 0 to 1 for the analysis) of the face as expressing one of the six "basic" emotion labels (*i.e.*, *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as additional psychological states such as *valence*, *engagement* or *contempt*, and facial features such as *cheek raise*, *eye widen* or *jaw drop*.

For both human and automatic recognition, to determine which of the six “basic” emotions can be used to identify each video, the recognition probability for each label by frame was converted into odd ratio by label [@dente2017measures]. The highest sum of each odd ratio time-series defines the label recognized by the automatic classifier (EQ\ref{eq:2} for human recognition and EQ\ref{eq:3} for automatic recognition).

\begin{equation}
\label{eq:2}
video_{i}.label = \max\left(\frac{\sum_{k=1}^{n}x_{video_{i}.label_{j}.t_{k}}}{\sum_{k=1}^{n}x_{video_{i}.t_{k}}}\right)
\end{equation}

\begin{equation}
\label{eq:3}
video_{i}.label = \max\left(\frac{\sum_{k=1}^{f}p_{video_{i}.label_{j}.t_{k}}}{\sum_{k=1}^{f}p_{video_{i}.t_{k}}}\right)
\end{equation}

where *i* is one of the `r nrow(unique_video_count)` videos, *j* is one of the six “basic” emotion labels, *k* for each second of the *n* second video or for each frame of the *f* frame video.

# Results

Since encoders’ self-reports, human annotations and the automatic recognition include data on “non-basic” emotion labels and features, the analysis is performed using only the six "basic" emotion labels in order to compare them. The maximum score for self-reports, human annotations and automatic recognition is used to label the video. In case of more than one label obtaining the maximum value, the video is labeled as undetermined.

```{r corr-matrix}
list_video <- human_recognition_score$C_Video
list_emotion <- c("disgust","fear","happiness","surprise","sadness","anger")

grid_data <- expand.grid(list_video = list_video, list_emotion = list_emotion)

self_report_result <- self_report_score %>%
  dplyr::filter(sr_emotion != "undetermined") %>%
  dplyr::mutate(sr_score = 1)

human_recognition_result <- human_recognition_score %>%
  dplyr::filter(hr_emotion != "undetermined") %>%
  dplyr::mutate(hr_score = 1)

automatic_recognition_result <- automatic_recognition_score %>%
  dplyr::filter(ar_emotion != "undetermined") %>%
  dplyr::mutate(ar_score = 1)

recognition_result <- grid_data %>%
  dplyr::left_join(self_report_result, by = c(
    "list_video" = "C_Video",
    "list_emotion" = "sr_emotion"
    )
  ) %>%
  dplyr::left_join(human_recognition_result, by = c(
    "list_video" = "C_Video",
    "list_emotion" = "hr_emotion"
    )
  ) %>%
  dplyr::left_join(automatic_recognition_result, by = c(
    "list_video" = "C_Video",
    "list_emotion" = "ar_emotion"
    )
  ) %>%
  replace(is.na(.), 0)

corr_sr_hr <- cor.test(
  recognition_result$sr_score, 
  recognition_result$hr_score, 
  alternative = "two.sided"
  ) %>% 
  papaja::apa_print()

corr_sr_ar <- cor.test(
  recognition_result$sr_score, 
  recognition_result$ar_score, 
  alternative = "two.sided"
  ) %>% 
  papaja::apa_print()

corr_hr_ar <- cor.test(
  recognition_result$hr_score, 
  recognition_result$ar_score, 
  alternative = "two.sided"
  ) %>% 
  papaja::apa_print()
```

## Human Observers’ Accuracy

```{r sr_hr}
comparison_sr_hr <- dplyr::inner_join(self_report_score,human_recognition_score, by = "C_Video") %>% 
  dplyr::ungroup() %>% 
  dplyr::filter(hr_emotion != "undetermined") %>% 
  dplyr::filter(sr_emotion != "undetermined") %>% 
  dplyr::mutate_if(is.character,as.factor)

confusionMatrix_sr_hr <- caret::confusionMatrix(
  data = comparison_sr_hr$hr_emotion,
  reference = comparison_sr_hr$sr_emotion
  )
```

The overall correlation of recognition and non-recognition between self-reported emotions and human observers recognition is significant but low (`r corr_sr_hr$full_result`). In order to identify differences according to the emotional labels, encoders’ subjective feelings are compared with human observers’ recognition in a confusion matrix (Figure \ref{fig:confusionMatrix_sr_hr}).

```{r confusionMatrix_sr_hr, fig.height=6, fig.cap="\\label{fig:confusionMatrix_sr_hr}Confusion matrix between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the human observers."}
confusionMatrix_sr_hr_freq <- confusionMatrix_sr_hr$table %>% 
  as.data.frame() %>% 
  dplyr::rename(hr_emotion = Prediction, sr_emotion = Reference) %>% 
  dplyr::mutate(n_video = sum(Freq)) %>% 
  dplyr::mutate(prop_video = Freq/sum(Freq)) %>% 
  dplyr::group_by(sr_emotion) %>% 
  dplyr::mutate(n_sr = sum(Freq)) %>%
  dplyr::mutate(prop_sr = Freq/sum(Freq)) %>% 
  dplyr::ungroup()

confusionMatrix_sr_hr_freq %>% 
  ggplot(mapping = aes(x = sr_emotion, y = hr_emotion)) +
  geom_tile(aes(fill = prop_sr), colour = "white") +
  geom_text(
    aes(label = paste0(scales::percent(prop_sr), "\n(", Freq,"/",n_sr, ")")),
    color = "black",
    size = 4,
    family="serif",
    parse = FALSE,
    lineheight = 0.7)+
  scale_fill_gradient(
    name = "Proportion of Recognized\nSelf-Reported Emotion",
    low = "white", 
    high = "red", 
    labels = scales::percent,
    limits = c(0,1)
  ) +
  scale_x_discrete(name = "Self-Reports") +
  scale_y_discrete(name = "Human Observers") +
  theme_minimal() +
  theme(text = element_text(size=16,family="serif"),
        axis.text.x = element_text(size=16, angle = 45, hjust = 0.75,vjust=0.9),
        axis.text.y = element_text(size=16, hjust = 0.5),
        axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
        legend.text = element_text(size=8),
        legend.position = "bottom")
```

Each emotion label used to describe encoders’ self-reported subjective feeling (*i.e.*, the label rated with the highest value) is compared with the emotion labels which were rated with the highest score by human observers. Results of the confusion matrix show a low agreement between the emotion felt by the encoder during the elicitation and the emotion recognized by the human observers (Accuracy \nolinebreak = \nolinebreak `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95% CI [`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *disgust* (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_sr"]))` of the videos self-reported). These results are far from those classically obtained in the literature for emotional facial expression recognition which ranges between 60% and 80% accuracy. However these results are mostly obtained with static (*i.e.*, pictures) and posed (*i.e.*, displayed by actors) facial expressions using only 6 emotional labels in a forced-choice paradigm.

Interestingly human observers seem to recognize surprise expressed in videos where anger, fear happiness and sadness was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "anger", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "fear", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "happiness", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "sadness", "prop_sr"]))` of the videos self-reported), and in a lower instance *happiness* was recognized in videos where *fear* and *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "fear", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).

Sensitivity, specificity, precision and F1 scores for each emotion reveals that *happiness* has the highest coherence ratio whereas *sadness* has the lowest coherence ratio between true positives and false positives (Table \ref{table:confusionTable_sr_hr}).

```{r confusionTable_sr_hr, results="asis"}
confusionMatrix_sr_hr_table <- confusionMatrix_sr_hr %>%
  magrittr::use_series(byClass) %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Emotion") %>%
  dplyr::mutate(Emotion = gsub(pattern = "Class: ",replacement = "",x = Emotion)) %>% 
  dplyr::select(Emotion, Sensitivity, Specificity, Precision, F1) %>% 
  dplyr::mutate_if(is.numeric,round,2) %>% 
  replace(is.na(.), "\\textit{na.}")

confusionMatrix_sr_hr_table %>%
  knitr::kable(
    "latex",
    caption = "\\label{table:confusionTable_sr_hr}Human recognition accuracy metrics for each emotion.",
    booktabs = TRUE,
    digits = 2,
    linesep = "",
    escape = FALSE,
    align = c('l',rep('c', 4))
  ) %>%
  kableExtra::kable_styling(
    latex_options = "HOLD_position",
    font_size = 8,
    full_width = TRUE 
    ) %>%
   add_footnote("Note. \\textit{na.} values are produced when not enough data are available to compute accuracy indicators.", notation="none", escape = FALSE)
```

Accuracy metrics by emotional labels indicate a discrepancy in the ratio of true/false positives. Whereas *happiness* and *disgust* obtain the highest scores, *anger*, *surprise* and *sadness* have the lowest recognition ratio. The underlying effect of expression intensity may explain why *happiness* and *disgust* are easily recognized. *Anger* and *sadness* as non-socially desirable emotions may be have been felt but not expressed.

However, self-reports show a significant proportion of undetermined emotional states (`r scales::percent(nrow(list_automatic_ties)/358)` of the 358 videos) which reveals the potential limit of using 6-points Likert scales to measure emotional self-reports. Indeed, encoders can easily score to the maximum for more than one emotion.

## Automatic Classifier’s Accuracy

```{r sr_ar}
comparison_sr_ar <- dplyr::inner_join(self_report_score,automatic_recognition_score, by = "C_Video") %>% 
  dplyr::ungroup() %>% 
  dplyr::filter(ar_emotion != "undetermined") %>% 
  dplyr::filter(sr_emotion != "undetermined") %>% 
  dplyr::mutate_if(is.character,as.factor)

confusionMatrix_sr_ar <- caret::confusionMatrix(
  data = comparison_sr_ar$ar_emotion,
  reference = comparison_sr_ar$sr_emotion
  )
```

Similarly to the previous analysis, the overall correlation of recognition and non-recognition revealed a significant but very low coherence between self-reported emotions and automatic classifier’s recognition (`r corr_sr_ar$full_result`). A confusion matrix was used to compare encoders’ subjective feeling with the emotion label recognized by the automatic classifier (Figure \ref{fig:confusionMatrix_sr_ar}).

```{r confusionMatrix_sr_ar, fig.height=6, fig.cap="\\label{fig:confusionMatrix_sr_ar}Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the automatic classifier."}
confusionMatrix_sr_ar_freq <- confusionMatrix_sr_ar$table %>% 
  as.data.frame() %>% 
  dplyr::rename(ar_emotion = Prediction, sr_emotion = Reference) %>% 
  dplyr::mutate(n_video = sum(Freq)) %>% 
  dplyr::mutate(prop_video = Freq/sum(Freq)) %>% 
  dplyr::group_by(sr_emotion) %>% 
  dplyr::mutate(n_sr = sum(Freq)) %>%
  dplyr::mutate(prop_sr = Freq/sum(Freq)) %>% 
  dplyr::ungroup()

confusionMatrix_sr_ar_freq %>% 
  ggplot(mapping = aes(x = sr_emotion, y = ar_emotion)) +
  geom_tile(aes(fill = prop_sr), colour = "white") +
  geom_text(
    aes(label = paste0(scales::percent(prop_sr), "\n(", Freq,"/",n_sr, ")")),
    color = "black",
    size = 4,
    family="serif",
    parse = FALSE,
    lineheight = 0.7)+
  scale_fill_gradient(
    name = "Proportion of Recognized\nSelf-Reported Emotion",
    low = "white", 
    high = "red", 
    labels = scales::percent,
    limits = c(0,1)
  ) +
  scale_x_discrete(name = "Self-Reports") +
  scale_y_discrete(name = "Automatic Classifier") +
  theme_minimal() +
  theme(text = element_text(size=16,family="serif"),
        axis.text.x = element_text(size=16, angle = 45, hjust = 0.75,vjust=0.9),
        axis.text.y = element_text(size=16, hjust = 0.5),
        axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
        legend.text = element_text(size=8),
        legend.position = "bottom")
```

Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are somewhat similar to the ones with human observers (Table \ref{table:confusionTable_sr_ar}). Overall, a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`, 95% CI [`r round(confusionMatrix_sr_ar$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_ar$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_ar$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` of the video self-reported) is evident.

```{r confusionTable_sr_ar, results="asis"}
confusionMatrix_sr_ar_table <- confusionMatrix_sr_ar %>%
  magrittr::use_series(byClass) %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Emotion") %>%
  dplyr::mutate(Emotion = gsub(pattern = "Class: ",replacement = "",x = Emotion)) %>% 
  dplyr::select(Emotion, Sensitivity, Specificity, Precision, F1) %>% 
  dplyr::mutate_if(is.numeric,round,2) %>% 
  replace(is.na(.), "\\textit{na.}")

confusionMatrix_sr_ar_table %>%
  knitr::kable(
    "latex",
    caption = "\\label{table:confusionTable_sr_ar}Autonatic recognition accuracy metrics for each emotion.",
    booktabs = T,
    digits = 2,
    linesep = "",
    escape = FALSE,
    align = c('l',rep('c', 4))
  ) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position"),
    font_size = 8,
    full_width = TRUE
    ) %>%
   add_footnote("Note. \\textit{na.} values are produced when not enough data are available to compute accuracy indicators.", notation="none", escape = FALSE)
```

Surprisingly the automatic classifier incorrectly recognized as *disgust* an significant proportion of videos in which *anger*, *happiness* and *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "anger", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported). In parallel, the automatic classifier recognized as *happiness* videos in which *fear* and *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "fear", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).

A comparable explanation involving the amount of undetermined video based on self-reports can be provided, as the level of undetermined emotions are very high for the self- reports.

## Comparison Between Human and Automatic Recognition

As previously mentioned, human observers appear to be more accurate than the automatic classifier to recognize an individual's subjective feeling (human observers Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`; automatic classifier Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`; `r corr_hr_ar$full_result`). However, both make mistakes.

A third confusion matrix is used to compare similarities (diagonal) and differences between human observers and automatic classifier in classifying the six emotion labels (Figure \ref{fig:confusionMatrix_hr_ar}).

```{r confusionMatrix_hr_ar, fig.height=6, fig.cap="\\label{fig:confusionMatrix_hr_ar}Proportion of emotion labels classified by human observers which are recognized by the automatif classifier."}
comparison_hr_ar <- dplyr::inner_join(human_recognition_score,automatic_recognition_score, by = "C_Video") %>% 
  dplyr::ungroup() %>% 
  dplyr::filter(ar_emotion != "undetermined") %>% 
  dplyr::filter(hr_emotion != "undetermined") %>% 
  dplyr::mutate_if(is.character,as.factor)

confusionMatrix_hr_ar <- caret::confusionMatrix(
  data = comparison_hr_ar$ar_emotion,
  reference = comparison_hr_ar$hr_emotion
)

confusionMatrix_hr_ar_freq <- confusionMatrix_hr_ar$table %>% 
  as.data.frame() %>% 
  dplyr::rename(ar_emotion = Prediction, hr_emotion = Reference) %>% 
  dplyr::mutate(n_video = sum(Freq)) %>% 
  dplyr::mutate(prop_video = Freq/sum(Freq)) %>% 
  dplyr::group_by(hr_emotion) %>% 
  dplyr::mutate(n_hr = sum(Freq)) %>%
  dplyr::mutate(prop_hr = Freq/sum(Freq)) %>% 
  dplyr::ungroup()

confusionMatrix_hr_ar_freq %>% 
  ggplot(mapping = aes(x = hr_emotion, y = ar_emotion)) +
  geom_tile(aes(fill = prop_hr), colour = "white") +
  geom_text(
    aes(label = paste0(scales::percent(prop_hr), "\n(", Freq,"/",n_hr, ")")),
    color = "black",
    size = 4,
    family="serif",
    parse = FALSE,
    lineheight = 0.7)+
  scale_fill_gradient(
    name = "Proportion Recognized Emotion",
    low = "white", 
    high = "red", 
    labels = scales::percent,
    limits = c(0,1)
  ) +
  scale_x_discrete(name = "Human Observers") +
  scale_y_discrete(name = "Automatic Classifier") +
  theme_minimal() +
  theme(text = element_text(size=16,family="serif"),
        axis.text.x = element_text(size=16, angle = 45, hjust = 0.75,vjust=0.9),
        axis.text.y = element_text(size=16, hjust = 0.5),
        axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
        legend.text = element_text(size=8),
        legend.position = "bottom")
```

The overall agreement between human observers and the automatic classifier is in fact very low (Kappa = `r round(confusionMatrix_hr_ar$overall["Kappa"],2)`). Except for *happiness* and *disgust* (respectively `r scales::percent(as.numeric(confusionMatrix_hr_ar_freq[confusionMatrix_hr_ar_freq$ar_emotion == "happiness" & confusionMatrix_hr_ar_freq$hr_emotion == "happiness", "prop_hr"]))` and `r scales::percent(as.numeric(confusionMatrix_hr_ar_freq[confusionMatrix_hr_ar_freq$ar_emotion == "disgust" & confusionMatrix_hr_ar_freq$hr_emotion == "disgust", "prop_hr"]))` of common labelling), there is no clear common pattern. Moreover, the automatic classifier has a tendency to label as *disgust* videos labeled as *sadness* by human observers, and as *happiness* videos labeled as *fear* by human observers.

## Discussion and Conclusion

Despite being one of the most investigated questions in affective science, the consistency between emotion felt and facially displayed on the one hand, and facial expression recognized on the other is a hot topic. To date no clear evidence has been found to definitively solve the questions raised. Yet, with the growing interest of industries and government in monitoring individual’s psychological states, this issue is under intense scrutiny. The present research aims to provide some empirical data to answer some of the questions posed. The faces encoders spontaneously displayed when confronted with an emotional eliciting task were submitted both to human and to automatic recognition. The criterion for recognition accuracy was the subjective feeling self-reported by the encoder once the elicitation task was carried out. Results first reveal a low consistency between emotion felt and facial expression displayed. They show that facial expressions of emotion are often not displayed when the Basic Emotion View would predict them to be expressed. Secondly, results show low accuracy rates for both humans and the automatic classifier in identifying the inner emotional states of these encoders based on their facial expressions. Thirdly, human observers prove to be better at recognizing the emotion facially expressed than the automatic recognition tool is. Such results support the hypothesis advanced by some authors of low emotion–expression coherence (Kappas 2003). In many instances, facial displays are not associated with a concordant emotional state, even any emotional state at all (Bonanno & Keltner 2004; Fernández-Dols & Crivelli, 2013]. More and more evidence is showing that facial expressions are in reality not expressing emotions (McKeown 2013]. Increasing studies show that, for most emotions, the EFE elicited by emotional triggers are scarce and partial, even when micro-expressions are taken into account (Dúran et al., 2017). These studies are conducted either in laboratory settings or in the field. This is not to say that facial expression is not an informative modality for understanding the emotional state of a person. There is indeed an affinity between emotion and facial display (Frijda & Tcherkassof, 1997). However, claiming that there are unique “signatures” allowing using configurations of facial muscles to infer the presence of a specific emotion is misleading. As well as other nonverbal behaviors, facial movements are not only assumed to be determined by emotion but also by various other causes, such as psychological states (e.g., motivations or pain), to say nothing of social context and sociocultural norms (Ekman et al. 1987). Hence, facial movements have causes and functions other than the expression of emotion. This multiple determination excludes any possibility of drawing a linear inference from facial activity on the underlying psychological state (emotional or other). Beyond the present observations showing a weak consistency between subjective feelings and spontaneous facial expressions, this study sheds some light on the controversy between the Basic Emotion View and the constructivist approach as to the facial recognition issue. The former (Ekman 1992) assumes that expressions of emotion are brief and coherent patterns of facial muscle movements that co-vary with discrete subjective experiences. In return, this information displayed by the face corresponds to the one extracted by perceivers. Instead of viewing emotions as natural kinds (Barrett, 2006), the constructivist approach supposes that emotions are social constructions and that facial behaviors intrinsically situated. The emotions that are recognized by the observer are constructed in her/his mind. Therefore, facial movements do not express specific emotions because they do not carry intrinsic emotional signification. It is the observer that infers the emotional meaning of the facial expression, this interpretation depending on availability of different kinds of information (contextual information, linguistic categories…). As a consequence, one can predict from the first line of thinking that individuals’ emotional subjective feeling should be correlated to the recognition of facial expressions from both human observers and automatic classifiers whereas if emotions are social constructs, as stated by the second line of thinking, human observer’s should be better at perceiving emotions expressed on the face than automatic classifiers.

The present findings speak against any strong version of the Basic Emotion View. The correlations between the self-measured emotions and the observed facial behaviors are low and the latter are weakly recognized. Present results plead instead in favor of the constructivist stance. They show that human observers are more accurate than the automatic recognition tool to identify an individual’s subjective feeling on the basis of their face. Moreover, mistakes made by human observers look less arbitrary to the ones made by the classifier. For instance, even if a mix-up between disgust and anger is sometimes reported in recognition studies, odd confusions such as the present ones produced by the classifier have never been noted for human observers. The latter obviously make sense of the facial behavior they are witnessing. However, the human decoders in the present study were presented with faces without any contextual information which could have helped them to shape more precisely their interpretation. Hassin, Aviezer and Bentin (2013) reviewed evidence to demonstrate that faces are inherently ambiguous and that observers rely on situational cues when they process facial displays (see also Aviezer, Hassin, Ryan, Grady, Susskind, Anderson, Moscovitch & Bentin, 2008; Barrett, Mesquita, Gendron, 2011). It happens that contextual information even shifts emotion perception. In the present case, the facial stimuli displayed to the decoders were as equivocal as are real-life facial expressions (Aviezer, Ensenberg, & Hassin, 2017). Yet, decoders had no cue at disposal. So, without any possibility of integrating faces and context, their decoding accuracy is pretty much degraded as it is usual in such cases (e.g., Wagner, MacDonald, & Manstead, 1986). However, they demonstrate their superiority on the classifier tool probably thanks to their capacity to rely on their previous personal experience to invent a context in which a face could display such an expression. Once a credible context retrieved, they can affix an emotional label to the facial behavior.

Several limitations should be stated, highlighting the need for further research. One of them is effectiveness of the emotion elicitation tasks. One can consider that an intensity threshold needs to be exceeded for a visible emotional expression to occur. In the case of the present study, it may be that insufficient emotion intensity accounts for the low number of visibly reactive participants. However, such line of reasoning fits badly with the Basic Emotion View according to which emotions and their related prototypical facial behavior are universal because they are considered as innate mechanisms allowing individuals to respond adaptively to evolutionary significant events (threats, opportunities…) encountered in the environment, whatever their potency. Moreover, various studies have used fairly strong emotion inductions without obtaining any visible facial display of any kind of emotion (Dúran et al, 2017). Besides, the present results show a quite good correlation between reported emotion and facial behavior for happiness. Thus, there are no reasons to believe that an explanation in terms of a too low emotion threshold applies for the other emotions under consideration. The latter were triggered and measured with conceptually identical elicitation and assessment procedures (Reisenzein et al., 2013). The use of self-reports to evaluate encoders’ subjective feelings is another limitation that can also be put forward because of the numerous cognitive biases they entail. Well known problems with the reliability of self-reports are, among others, the reconstructive nature of memory, the influence of attentional biases on reports, demand characteristics, distorting effects of implicit causal theories and personal motives, as stressed by Nielsen & Kaszniak (2007). It is clearly obvious that self-reports are not simple outlets of inner mental processes but personal constructions and they are affected by many factors (Kappas, 2003). The emotional feeling echoes motivational tendencies, bodily changes, and cognitive appraisals of events (Sander & Scherer, 2009). All these are encapsulated into semantic categories referred to by labels. As things stand currently, there is unfortunately no objective way for accessing and assessing inner subjective emotional feelings except for asking people to report their subjective experience in words. The procedure used for human recognition can also be open to dispute. First, one can criticize the decoders' expertise level since they were not FACS coders but untrained students. Of course one can expect a difference between skilled annotators and novice ones regarding the assessment of emotional facial displays. This said, in everyday life, few people are specialist coders yet the quantity of successful social interactions proves lay persons recognize pretty well others’ facial behaviors. Therefore, and especially in the interest of generalizability, asking inexpert people seems relevant. Secondly, instead of using a classic forced-choice procedure, a more subtle approach was chosen to mimic results provided by the automatic classifier. As explained above, annotators first delimitated a temporal sequence during which they noticed an emotional display on the face, and then attributed an emotion label to this behavioral sequence in a second step. Whereas this paradigm is longer and more complicated, it can lead to more robust results in reducing the forced-choice bias (Russell, 1993). However this procedure can also reduce the human observers’ accuracy. In this regard, the results of the human observation could have been more ambiguous because it is not the natural way that people are inferring meaning from facial expressions. An alternative explanation relies in reducing the recognition bias involved in the classic recognition paradigm. Classic forced-choice paradigms obtain artificially high results, thus by using a more evolved approach observers’ accuracy may have been lowered. Another flaw is the lack of comparison with various facial expression recognition methods. Human recognition has only been compared to the Affdex classifier. Future studies are needed to confront human assessments with different automatic recognition methods, both frame based methods and sequence based automatic ones. This latter issue is particularly decisive. Indeed, the issue of the recognition of dynamic expressive sequences is essential because ordinary facial behavior is made up of dynamically shifting morphological features (Krumhuber, Kappas & Manstead, 2013). This temporal information is indisputably a key feature of facial activity. Not only observer-based judgements of facial displays must be compared to automated facial analysis, but also different kinds of human recognition measurements should be undertaken. The challenge researchers are especially confronted with is to find ways to appropriately collect data regarding the perception of spontaneous and dynamic facial behavior. Finally, our understanding of facial displays as they occur in everyday interactions requires a strong emphasis on ecological concerns. The present study is a laboratory experiment.  It has the advantage of controlling different parameters of the emotions investigated, such as intensity, quality, and temporal (onset, duration) features. Moreover, as encoders are alone, facing an emotional trigger, it also controls for the social context, removing its possible influence on their facial behaviors. However, it is known that encoders’ imagination can influence their expressiveness, for instance when they believe that their friend is doing the same (vs a different) emotional task in another room (Jakobs, Manstead, & Fischer, 1999). Trying to exclude social influence by leaving encoders alone may be illusionary then. From an ecological perspective, it is even a mistake to exonerate behavioral observations from social contexts.  Facial activity measurement in dyadic interactions has shown that the facial behavior of the perceiver reflects sometimes more what the expresser is experiencing than what the perceiver is feeling. It is the case, for instance, of emotional mimicry in dynamic social interactions (Hess & Fischer, 2016).  It is also the case of healthy partners interacting with schizophrenic patients whose facial activity is pretty much identical (Krause, Steimer-Krause, Merten & Burkhard, 1998). Hence, in order to better comprehend emotional communication in human relationships, experimental research should be corroborated with more ecological protocols.

Nowadays, automatic recognition systems are based on the coding of the facial muscular activations from which they infer the expressed emotion. Such automatic classifier tools take for granted that, when experienced, firstly an emotion is displayed on the face, secondly, in the form of a configuration of facial muscles that is all his own, and thirdly, which is recognized by the perceiver (be it a human being or an automatic system). These are the Basic Emotion View assertions, jeopardized by field observations and laboratory experiments on spontaneous expression of emotions, such as the present study. All raise serious objections to the supposed close relation between emotion and face. They bring up several questions regarding the role the context plays in the emission and interpretation of the so-called facial expression of emotion. The finalization of operational and effective “reading emotional faces” devices rests on their answer. As a result, despite being one of the most investigated questions in affective science, the growing interest of industries and governments in tracking individual’s psychological states is hardly given satisfaction. Considering the above, the present results provide additional evidence that an individual’s subjective feeling cannot be inferred from facial expressions and in our case invalidate the hypothesis of hardwired emotions unambiguously displayed on the face. Even if emotions were hardwired, in everyday life one does not observe prototypical facial expressions because of their rarity and therefore research should be focused on analyzing non-prototypical facial expressions. Advancements in identifying “non-basic” emotion labels (McDuff, 2016) as well as non-prototypical facial expression have indeed been made in the development of automatic facial expression recognition tools. However, these present results suggest that automatic facial expression recognition tools should merely evaluate facial morphology features such as action units (already evaluated in OpenFace, Baltrušaitis, Robinson, & Morency, 2016, Affectiva’s Affdex, McDuff, Mahmoud, Mavadati, Amr, Turcot, & el Kaliouby, 2016, or Vicar Vision’s FaceReader, Den Uyl & Van Kuilenburg, 2005, to name a few) rather than inferring supposedly emotional or affective states. Trying to interpret facial displays as a means of determining underlying emotional state, in all likelihood, remains vain.

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup

### 2019-11-01 manuscript_apa.tex
references

\documentclass[man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={The Emotion--Facial expression link: Evidence from Human and Automatic Expression Recognition},
            pdfauthor={Anna Tcherkassof~\& Damien Dupré},
            pdfkeywords={Facial expression, self-report, human observer, automatic recognition.},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

  \title{The Emotion--Facial expression link: Evidence from Human and Automatic Expression Recognition}
    \author{Anna Tcherkassof\textsuperscript{1}~\& Damien Dupré\textsuperscript{2}}
    \date{}

\shorttitle{The Emotion–Facial expression link}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} University Grenoble Alpes, Grenoble, France\\\textsuperscript{2} Dublin City University, Dublin, Ireland}
\keywords{Facial expression, self-report, human observer, automatic recognition.\newline\indent Word count: 7920}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

\linenumbers
\usepackage{booktabs}
\usepackage{float}
\usepackage{tabu}
\usepackage{wrapfig}
\usepackage[none]{hyphenat}
\floatplacement{figure}{H}
\usepackage{flushend}
\usepackage{biblatex}

\authornote{

Correspondence concerning this article should be addressed to Anna Tcherkassof, Bâtiment Michel Dubois, Université Grenoble Alpes, 1251 Avenue Centrale, 38400 Saint-Martin-d'Hères. E-mail: \href{mailto:anna.tcherkassof@univ-grenoble-alpes.fr}{\nolinkurl{anna.tcherkassof@univ-grenoble-alpes.fr}}}

\abstract{
While it has been taken for granted in the development of several automatic facial expression recognition tools, the question of the coherence between subjective feelings and facial expressions is still a subject of debate. On one hand, the Basic Emotion View conceives emotions as genetically hardwired and therefore being genuinely displayed through facial expressions. On the other hand, the constructivist approach conceives emotions as socially constructed; the emotional meaning of a facial expression being inferred by the observer. In order to evaluate the coherence between the subjective feeling of emotions and their recognition based on facial expression, 232 videos of encoders recruited to carry out an emotion elicitation task were annotated by 1383 human observers as well as by an automatic facial expression classifier. Results show low accuracy of human observers and of the automatic classifier to infer the subjective feeling from the facial expressions displayed by encoders. They also show a weak consistency between self-reported emotional states and facial emotional displays. Based on these results, the hypothesis of genetically hardwired emotion genuinely displayed is difficult to support, whereas the idea of emotion and facial expression socially constructed appears to be more likely. Accordingly, automatic emotion recognition tools based on facial expressions should be questioned.

}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

With the development of commercial automatic facial expression recognition tools (Dupré, Andelic, Morrison, \& McKeown, 2018), industries and governments are gradually implementing this technology in order to track humans' emotions in various scenarios (\emph{e.g.}, in marketing, healthcare, and the automotive industry to name a few). This technology rests on the premise that facial expressions provide a direct access to individuals' subjective feelings, and that one can read the emotions displayed on a person's face as in an open book. This premise is derived from the Basic Emotion View which Paul Ekman is one of the famous proponents of this perspective (Ekman, 1992, 2007; Ekman et al., 1987; Ekman \& Heider, 1988). Even if this premise is central to the modern mainstream approach of human emotion, recent research in affective science is challenging it. Once the Basic Emotion View is briefly described, its foremost criticisms are synthetically exposed. Inconsistencies and unpredicted findings research has uncovered come both from field observations and laboratory experiments. They relate to two different issues. The first issue is the one of the sender's production of spontaneous facial displays. The second issue has to do with the viewer's interpretation of these facial displays. Building on this evidence, and as a further attempt to shed light on the emotion--face link, the present study is finally exposed. Its main goal is to tackle these two issues at the same time. It aims at identifying the extent to which ordinary people's experienced emotions are displayed through identifiable emotional facial expressions (EFE) recognized as such both by human individuals and by automatic facial expression recognition tools.

\hypertarget{the-basic-emotion-view}{%
\subsection{The Basic Emotion View}\label{the-basic-emotion-view}}

The belief that facial expression is linked to emotional states can be traced back to Darwin in \emph{The Expression of the Emotions in Man and Animals} (Darwin, 1872). Countless studies have fortified the idea that emotional states are inherently coupled to a set of prototypic facial expressions (\emph{e.g.}, (Ekman \& Cordaro, 2011)). The Basic Emotion View holds that facial expressions are genuine displays of an individual's inner emotional state. More specifically, a set of six emotions (\emph{fear}, \emph{anger}, \emph{surprise}, \emph{disgust}, \emph{sadness} and \emph{joy}) are universally displayed and are genetically hardwired not only in humans (Ekman, 1992), but also in different animal species (Waal, 2019). According to the view of the Basic Emotion View, \emph{\enquote{when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including {[}\ldots{}{]} facial expressions.}} (Ekman, 2007, p. 49). To respond to criticisms, several amendments have been made to the Basic Emotion View, increasing the number of basic emotions from six to seven (Ekman \& Heider, 1988) as well as adding the concept of \enquote{display rules} to explain cultural differences in the management of facial expressions (Ekman et al., 1987).

Relatively few studies have actually tested Basic Emotion View's fundamental claim regarding the facial production (so-called \enquote{expression}) of emotions. Rosenberg \& Ekman (1994) reported the first evidence of coherence between self-report of emotion and displayed facial expressions. Participants were showed emotionally evocative films and reported their own emotions using a moment-by-moment reporting procedure. Analysis of participants' facial expressions and reports of emotions showed that there was a high degree of temporal linkage and emotional agreement between facial expressions and self-reports. Notwithstanding this study, few have used purely descriptive methods (such as electromyography or objective coding systems -- FACS or others) to identify and to measure the actual changes in the face when a given emotion is felt (see Wagner (1997), for methodological points). Much of the research has focused instead on the recognition of EFE, that is on the issue of the viewer's interpretation of facial displays. As people have been thought to display their internal state, EFE supposedly serve as a window into the emotions of others. Viewers \enquote{recognize} the facial displays and infer the corresponding emotional state. The Readout Hypothesis (Buck, 1985) formalizes this folk theory. According to it, facial expressions have evolved to provide \enquote{an external readout of those motivational-emotional processes that have had social implications during the course of evolution} (pp.~396-397). Thus, as readouts of emotional states, spontaneous expressive displays are directly accessible to other organisms. Research on the inference of emotions from facial expressions has established that viewers show considerable agreement that the so-called basic emotions---happiness, anger, fear, disgust, sadness, and surprise---are associated with specific facial displays (\emph{e.g.}, reviews from Elfenbein \& Ambady (2002); Russell (1994)). These findings are considered by many as constituting strong evidence for the Basic Emotion View.

Yet, in spite of the popular support it has received, the empirical data called upon by the Basic Emotion View remains unpersuasive. Notably, it fails to explain, in instances in which display rules cannot be called upon, how individuals can feel emotions without expressing them or how individuals can express emotions without feeling them, and why observers are not totally accurate in recognizing facial expressions of basic emotions, among others. The following is a summary of key findings and conclusions both from field observations and laboratory experiments on spontaneous expression of emotions, all of them strongly subverting the Basic Emotion viewpoint (for reviews, Fernández-Dols \& Crivelli (2013); Reisenzein, Studtmann, \& Horstmann (2013); Russell, Bachorowski, \& Fernández-Dols (2003)). Mains issues regarding recognition studies are also swept across afterwards.

\hypertarget{spontaneous-facial-expression-in-naturalistic-studies}{%
\subsection{Spontaneous Facial Expression in naturalistic studies}\label{spontaneous-facial-expression-in-naturalistic-studies}}

Naturalistic studies look at the ecological frequency of co-occurrence of certain emotions and facial displays {[}fernandez2013emotion{]}. Though they cannot be considered as a straightforward test of the triggering role of emotion on the facial behavior, they have the advantage of considering situations that cannot be created in a laboratory. For instance, the ethological study of Kraut \& Johnston (1979) contradicts the Basic Emotion View premise that a smile is the major component of a facial display associated with and caused by feelings of joy or happiness. Naturalistic observation at a bowling alley showed that bowlers do not necessarily smile after scoring a spare or a strike (a situation likely to elicit a positive emotion). Rather, they often smile when interacting with other people. More generally, examination of bowlers' facial display showed that they rarely smiled while facing the pins but often smiled when facing their friends. These findings were confirmed by Ruiz-Belda, Fernández-Dols, Carrera, \& Barchard (2003) who analyzed the facial displays of bowlers (after a strike) and soccer fans (after their team scored) reporting happiness in a field study. They found that the probability of smiling was very low when participants were not interacting with someone else. These findings have been supported in other realistic field settings such as the one of Fernández-Dols \& Ruiz-Belda (1995) suggesting that happiness is not a sufficient cause of smiling. They watched extremely happy gold medalist athletes displaying facial expressions of sadness (sometimes associated with tears) during their Olympics awards ceremonies. More specifically, winners showed Duchenne smiles and other types of smiles when they were interacting with other people during the awards ceremony, but smiles were scarce or nonexistent when waiting behind the podium and/or when turning toward the flagpoles and focusing their attention on the flags and the national anthem. Crivelli, Carrera, \& Fernández-Dols (2015) also found that the strongest predictor for the occurrence of Duchenne smiles in judo winners is when they are engaged in social interaction and not when they just won their match.

Happiness (or joy) is not the only emotion weakly associated with the predicted facial expression. A naturalistic study conducted by Scherer \& Ceschi (1997) in an airport's baggage handling office showed that passengers claiming for their lost luggage displayed very few facial expressions of negative emotions while self-reporting subjective feeling states of anger or sadness among others. The covariation between passengers' self-ratings and the claims agents' attributions of the passengers' emotions was very low. Another refutation of the Basic Emotion View's predictions regarding the link between emotions and facial expressions is provided by naturalistic observations of infants' productions of facial expressions. It is often acknowledged that adults regulate their expressive behavior. In accordance with various display rules (personal, social, cultural ones), they exert a control over the (supposed) automatic readouts of their emotions. Developmental studies provide relevant settings in which display rules are inoperative. Camras and her colleagues observed that facial expressions of negative emotions were displayed in circumstances that were unlikely to have elicited those emotions. For instance, infants displayed \enquote{fear} expressions in settings not related to fear (Camras, Malatesta, \& Izard, 1991). Bennett, Bendersky, \& Lewis (2002) videotaped a sample of 4-month-old infants during tickle, sour taste, jack-in-the-box, arm restraint, and masked stranger situations. Infants displayed a variety of facial expressions in each eliciting situation. Yet, more infants exhibited positive (joy and surprise) than negative facial expressions (anger, disgust, fear, and sadness) across all situations --except sour taste. No evidence for emotion-specific facial expressions corresponding to anger, fear, and sadness was obtained. Camras and her colleagues also observed that 11-month-old European-American, Chinese, and Japanese infants did not display distinct negative emotion-specific patterns of facial muscles in response to two elicitors meant to induce fear and anger (Camras et al., 2007). Thus, these findings in naturalistic settings provide little support for the Basic Emotion View.

\hypertarget{spontaneous-facial-expression-in-laboratory-studies}{%
\subsection{Spontaneous Facial Expression in laboratory studies}\label{spontaneous-facial-expression-in-laboratory-studies}}

Laboratory findings also support field studies. As stressed by Reisenzein et al. (2013), experimental studies permit both better control of various factors (\emph{e.g.}, emotion elicitors) and tests about likely moderators of the emotion--facial expression link (\emph{e.g.}, the social context). As an example, a strong disconfirmation of the Basic Emotion View's premise is put forward by Reisenzein's studies on surprise (\emph{e.g.}, Reisenzein (2000)). In eight controlled laboratory situations, surprise was induced by Reisenzein, Bördgen, Holtbernd, \& Matz (2006) by establishing and then invalidating a set of beliefs concerning the experimental events (\emph{e.g}., the unexpected appearance of a picture of one's own face as the last picture in a series of portraits that had to be rated). Visible or EMG-detected facial displays of surprise occurred only in few participants. Yet, most participants reported subjective feelings of surprise and most believed that they had shown a strong surprise facial expression. Schützwohl \& Reisenzein (2012) also observed similarly low frequencies of surprise facial expressions when their participants, after leaving the laboratory, unexpectedly found themselves not in the corridor but in a new room with green walls and a red office chair. Less than a quarter of the participants displayed an expression of surprise: only 5\% showed widened eyes, raised eyebrows, and opened mouth (\emph{i.e.} the complete expression of surprise according to the Basic Emotion View) and 17\% showed widened eyes and raised eyebrows. Again, participants overestimated their surprise expressivity.

Studies on happiness and related positive emotions (\emph{e.g.}, sensory pleasantness) also show a low coherence between emotion and facial display (Durán, Reisenzein, \& Fernández-Dols, 2017; Reisenzein et al., 2013). In fact, only experiments on amusement provide a fairly strong association between emotion and smiling. When confronted to humorous events (\emph{e.g.}, jokes, being tickled), the number of facially reactive participants (smiling and laughing) is indeed quite fair, whereas when confronted to other positive emotions (\emph{e.g.}, happiness), few participants show the expected facial expressions (for instance, the Duchenne smile or any kind of ``happy'' expression yet reporting being happy; see Mehu, Grammer, \& Dunbar, 2007; Lee \& Wagner, 2002, social condition). More largely, the meta-analysis conducted by Durán et al. (2017) on the degree of statistical covariation between emotions and facial expressions provides conclusive evidence. First of all, the commonly called basic emotions (happiness/amusement, sadness, fear, anger, disgust and surprise), on average, are weakly correlated with the specific configuration of facial muscles that the Basic Emotion View ascribes to them (the correlation even drops when happiness/amusement studies are excluded). Very few participants who relived an experience of sadness (\emph{e.g.}, Tsai, Chentsova-Dutton, Freire-Bebeau, \& Przymus, 2002) were reactive, only a small number of them showing components of a sad facial expression (\emph{e.g.}, oblique eyebrows, lip corners pulled down). Regarding fear, spider phobic participants recruited by Vernon \& Berenbaum (2002), for instance, were exposed to a live tarantula. About one third of them only displayed some components of the prototypical fear expression (eye widening, brow raising and knitting, etc.). Similar observation is made for anger. The proportion of facially reactive participants (\emph{i.e.} displaying at least one component among frowning, lid/lip tightening\ldots{}) reporting anger in Johnson, Waugh, \& Fredrickson (2010) and Tsai et al. (2002) studies does not exceed 35\%. The same proportion of reactive participants confronted to disgusting elicitors is observed (\emph{e.g.}, Ekman, Freisen, \& Ancoli, 1980; Fernandez-Dols, Sanchez, Carrera, \& Ruiz-Belda, 1997). In any case neither an insufficient intensity of the emotion to cause a facial expression nor the intervention of display rules (nor measurement issues) can explain the low emotion--facial expression consistency.

Hence, laboratory studies show that facial expressions of emotion are often not displayed in situations in which the Basic Emotion View would predict them to be expressed. Moreover, when they are indeed displayed, they are more often displayed partially than completely. All in all, research on the spontaneous expression of emotions does not yield strong support for the Basic Emotion View. The available evidence steadily indicates weak links between emotions and their predicted facial expressions both in natural and in semi-naturalistic settings.

\hypertarget{emotional-facial-expression-recognition-studies}{%
\subsection{Emotional Facial Expression Recognition studies}\label{emotional-facial-expression-recognition-studies}}

The Basic Emotion View postulates that, when triggered, each basic emotion is expressed by a prototypical face (non-basic emotions being blends of the basic ones). In return, the recognition of emotional facial expressions (EFE) is claimed to be based on the identification of specific patterns of facial movements associated with each emotion, as if expression and recognition were the two sides of a same coin. Seemingly compelling evidence support this claim, sustaining the belief that one can read a face in the same way one reads the lines of a book. Besides, recognition systems rely on such a principle and are considered as an objective coding tool because they are based on the identification of specific muscular changes in the face. Many people highly agree that so-called basic emotions are associated with specific facial configurations (Ekman, 2017; Elfenbein \& Ambady, 2002; Russell, 1994) and also agree that many believe that it is a strong evidence for the Basic Emotion View. Moreover, it implies that EFE recognition of both human observers and automatic classifiers should be as accurate.

Yet, some researchers have highlighted the limitations of Basic Emotion View empirical research. Among others, evidence has been questioned on methodological grounds (\emph{e.g.}, Russell, 1994). The response format usually used in recognition studies (\emph{i.e}. forced choice: selection of one word from a pre-specified list of emotion labels), notably, leads to a biased consensus as evidenced by Russell (1993). Depending on the list of emotion labels at participants' disposal, EFE of sadness can easily be categorized as sad expressions as well as fear expressions\ldots{} Russell \& Fehr (1987) also showed the same facial expression can be seen as expressing different types of emotions, depending on what other faces are seen. DiGirolamo \& Russell (2017) conducted seven experiments that establish that high agreement between participants can be an artifact of the standard method. Thus, results gathered with forced choice cannot demonstrate the univocal link between emotion and facial expressions claimed by the Basic Emotion View. Using alternative recognition methods (emotion satiation procedure, face-matching task, sorting task; \emph{e.g.}, Lindquist, Barrett, Bliss-Moreau, \& Russell, 2006; Gendron, Roberson, Vyver, \& Barrett, 2014), it has been shown on the contrary that facial muscle movements are not linked in a one-to-one manner to a specific discrete emotional experience (2017,p.~418). Instead, emotions are probably mentally constructed by the perceiver and mental categories of emotions are needed to accurately categorize facial movements, among other things (\emph{e.g.}, contextual information).

To the methodological limitations contaminating the hundreds of studies apparently supporting the Basic Emotion View, a stimulus bias must be added. Facial stimuli used in experiments also constitute a methodological bias because they are unrepresentative of ordinary facial expressions. Basic Emotion View empirical evidence is based for the most part on methods using a static and unnatural material, namely, still photographs of posed facial expressions of emotion (\emph{e.g.}, intentionally encoded by the sender). This kind of methodology raises questions about its ecological validity and the generalizability of the results to real interpersonal emotional communication (\emph{e.g.}, Tcherkassof, Bollon, Dubois, Pansu, \& Adam, 2007). Indeed, a number of pieces of evidence indicate that research cannot content itself with data collected with static and posed material. These data come from researches studying the case of dynamic and/or spontaneous facial expressions of emotion. They show that the dynamic aspects of facial movement are likely to be of importance (\emph{e.g.}, Kamachi et al., 2013). Cohn \& Schmidt (2003) have shown that spontaneous smiles are of smaller amplitude and have a more consistent relation between amplitude and duration than deliberate smiles. Hess \& Kleck (1990) have also pointed out the importance of the dynamics of facial movements, and particularly the irregularity, or phasic changes, of the expressions' unfolding. Thus, the motion of facial expression provides observers with other information than the one provided by static expressions. It may be that differences in the social information displayed by static and dynamic expressions leads to facial recognition differential effects. Regarding the issue of spontaneous vs posed expressions (the latter are overused in experiments). As Meillon et al. (2010) conclude, EFE have been typically studied as static displays. As a consequence, even though the central role of the dynamics of facial expressions is endorsed, little is still known about the temporal course of facial expressions. Furthermore, studied EFE exhibit emotions simulated or posed by actors. Yet, the lack of spontaneity and naturalness of this material constitutes a serious objection raised against such studies (Kanade, Cohn, \& Tian, 2000).

All in all, as many doubts can be raised about the standard method, experiments conducted with such a method cannot be considered as providing solid empirical support to the Basic Emotion View. Based on the numerous methodological criticisms, but also theoretical, addressed to this view, alternative conceptions have emerged. Among them, the constructivist approach is gaining in importance. The constructivist approach represents a totally different way of understanding the emotion--facial expression link. It affirms that facial expressions do not provide a direct access to individuals' subjective feelings. Therefore, instead of considering that emotions can be \enquote{read} on facial displays, it claims that the emotion is \enquote{in the eye} of the perceiver (Barrett, Adolphs, Marsella, Martinez, \& Pollak, 2019).

\hypertarget{constructivist-approach}{%
\subsection{Constructivist Approach}\label{constructivist-approach}}

Starting from the empirical evidence suggesting that spontaneous facial expressions in ordinary life are equivocal, Fernández-Dols (2017) argues in favor of a pragmatic conception of natural facial displays. He makes a plea for the idea that natural facial displays, rather than \enquote{saying} -- because they do not have a specific meaning he claims -- \enquote{make} things. Facial expressions are actions in a communicative interaction. They do not express emotions but they \enquote{prompt, on the receiver's side, important inferences about the context, the sender, and the course of the interaction between sender and receiver} (p.466, underlined by the author). As such, the fact that facial displays are able to signal emotions is a byproduct of one of its main function, the one of implementing actions performing practical ends. Therefore, expressive displays hold a motor intention (Pacherie, 2003). They implement their aim which is their motor intention. They are not primarily communicative signals and even less the outlet of an internal state. Facial displays are (parts of) pragmatic actions aiming at orienting the person's relation to its environment, for instance maintaining, breaking, restoring, etc., the relation (Frijda, 2012). Facial displays are not \enquote{recognized} in semantic terms but are perceived as intentional actions. In the face of the continual flow of uninterrupted facial movements, the perceiver sees behaviors directed towards a goal. S.he translates the continuous flow of movements into coordinated sequences of actions holding a beginning and an end. Facial displays are not simple strings of action units which morphological configuration would be the prototype of a given emotion, and consequently identified as such (Ekman, 1982). They are best conceived as a Gestalt, the same way as a string of musical notes establishes a melody (Tcherkassof \& Frijda, 2014). This is why even unauthentic facial displays can yet be recognized as emotional expressions. Guillaume Duchenne de Boulogne explains that the artist who has shaped the famous Laocoon antic sculpture, exhibited in the Vatican's museum, has made a modelling mistake since no face can display its emotional expression. Indeed, no muscular contraction can produce it. He even rectifies the « mistake » by presenting a statue which face is shaped according to the physiology of facial expressive movements. His demonstration gives food for thought. Even though no objective coding system can correctly code the discordant facial features of the Laocoon's face, yet anyone can easily recognize the suffering and despair he admirably expresses. This example goes along the lines of the constructivist approach.
The constructivist approach claims that facial displays are behaviors which meaning is inferred by the perceiver. Findings support this observer dependence (Lindquist \& Gendron, 2013; Niedenthal, Wood, Rychlowska \& Korn, 2017). They show that to make meaning of another person's facial behavior, the perceiver relies in particular on her/his knowledge about emotion categories. For instance, Gendron and her colleagues used a face-sorting task allowing them to manipulate the influence of emotion concepts on how facial expressions were perceived. They conducted their experiment among U.S. participants and Himba participants from remote regions of Namibia (Gendron, Roberson, van der Vyver \& Barrett, 2014) and Hazda participants of the Eastern Rift Valley of Tanzania (Gendron, Hoemann, Crittenden, Msafiri, Ruark, \& Barrett, 2018) both groups with limited exposure to Western culture. Gendron and her colleagues demonstrated that facial expressions were not universally \enquote{recognized} in discrete emotional terms. Indeed, when Himba and Hazda participants did not have emotion concepts at their disposal to structure perception, they perceived the facial expressions as behaviors (e.g., looking, smelling) that didn't have a necessary relationship to emotions. They didn't infer inner states (such as emotional feelings) but they rather proceeded with action identification that pointed out the functions of behaviors (see also Crivelli, Russell, Jarillo, \& Fernández-Dols, 2017 for similar observations among a small-scale society of Papua New Guinea). The constructivist approach considers that specific emotion categories, as conceptualized by Western cultures' knowledge, are casted on the perceived face to make meaning of the sender's facial displays. Following this approach, faces convey a range of information essential for social communication. They are best conceived as tools displaying signals in social interactions (Crivelli \& Fridlund, 2018). These signals can convey individuals' motivations and readiness (Frijda \& Tcherkassof, 1997) or social messages (Fridlund, 1994). As for emotional meaning, more specifically, it is shaped by the perceiver according to the specific context in which the facial displays are observed.

Having reached this point, one can assert that numerous questions regarding the link between emotions and facial expressions remain unanswered. The two main competing approaches to facial expressiveness, the Basic Emotion View and the constructivist approach, entail completely opposite predictions regarding the decoding of facial expressions, as evidenced above. The present study aims at examining these predictions in order to provide empirical evidence to allow the discussion to evolve. To date, no systematic study has looked at the same time at facial expressions spontaneously displayed in reaction to emotional triggers and how they are decoded, both by human observers and by automatic emotion recognition tools based on the detection of facial muscular configurations. This study fills that gap. It intends to investigate the consistency between the subjective feeling of emotions and its recognition from facial expressions. Spontaneous and dynamic facial reactions to emotional elicitations are under consideration to ensure the generalizability of the results to emotional behaviors in ordinary life. More specifically, this study aims to examine the recognition of EFE produced by ordinary people during situations judged and/or self-reported to involve different emotions. It (a) examines consistency between ordinary people's self-reported emotional experience and observers' judgments of these ordinary people's EFE, (b) examines consistency between ordinary people's self-reported emotional experience and an automatic classifier's analysis of these ordinary people's EFE. In other words, it is interested in how people actually move their faces to express self-reported emotions, in how human observers accurately infer the expresser's emotional state, and in how automatic recognition accurately code the expresser's emotional state. It is expected a human observers' superior ability to exactly recognize EFE as compared to automatic EFE recognition tools.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

To evaluate the consistency between subjective feeling of emotions and their recognition from facial expressions, encoders were first recruited to perform an emotion elicitation task while their facial expression was video recorded. In order to reduce the likelihood of facial control, the encoders were alone in the room and were filmed by hidden cameras, so they had no reason to comply with social display rules. Then, the videos of the encoders' faces were shown to human observers and were also analyzed by an automatic classifier in order to identify which emotion was displayed.

\hypertarget{emotion-elicitation}{%
\subsection{Emotion Elicitation}\label{emotion-elicitation}}

For the emotion elicitation experiment, 358 encoding participants (182 females, 176 males, Mage = 47.9, SDage = 9.2) were recruited to perform one out of 11 emotion elicitation tasks designed to trigger a positive, a specific negative or a neutral emotional state. Encoders' faces were recorded using a hidden camera resulting 358 front facing 768x576 videos varying from 1s to 1479s. These recordings form the DynEmo database (Tcherkassof, Dupré, Meillon, Mandran, Dubois, \& Adam, 2013).

After the emotion elicitation task, the encoders rated their subjective feeling on Likert scales from 0 (\enquote{not at all}) to 5 (\enquote{strongly}) related to six \enquote{basic} emotion labels (i.e., anger, disgust, fear, happiness, surprise and sadness) as well as six \enquote{non-basic} emotion labels (i.e., pride, curiosity, boredom, shame, humiliation, and disappointment).

Finally, a debriefing session was performed to ensure that encoders were not durably affected by the emotion elicitation task. The debriefing was also used to check that encoders did not guess the real purpose of the experiment (e.g., being filmed while they were performing an emotional elicitation task) to guarantee facial expressions' genuineness. All encoders gave their agreement on their data and video to be processed for research purposes only.

\hypertarget{human-facial-expression-recognition}{%
\subsection{Human Facial Expression Recognition}\label{human-facial-expression-recognition}}

For the human facial expression recognition method, 1383 student participants were recruited to annotate 232 out of the 358 videos, therefore only the 232 annotated videos were analyzed in this study. Because videos have different durations, participants had to annotate a series of video corresponding to 30min long in total. Each video was annotated 29 times on average (SD = 12).
The annotation of facial expressions was performed on-site using Oudjat, a software for designing video annotation experiments (Dupré, Akpan, Elias, Adam, Meillon, Bonnefond, Dubois, \& Tcherkassof (2015). For each video, the annotation procedure followed two steps. First, the participants had to identify the emotional sequences by pressing the space bar of their keyboard to indicate the beginning and the end of the emotional sequences while watching the video. Second, the participants watched each emotional sequence previously identified and labeled the sequence using one of the 12 emotions proposed including six \enquote{basic} emotion labels (i.e., anger, disgust, fear, happiness, surprise and sadness) and six \enquote{non-basic} emotion labels (i.e., pride, curiosity, boredom, shame, humiliation, and disappointment). They also had the possibility to indicate that the sequence was expressing none of the proposed emotion labels.
This annotation procedure results in a uni-dimensional time-series for each video per human observer identifying for each second of the video which emotion was recognized. Then, time-series corresponding to the same video were aggregated to calculate the proportion of human observers \(x_{video_{i}.label_{j}.t_{k}}\) for each second of the video per emotional label (EQ\ref{eq:1}).

\begin{equation}
\label{eq:1}
x_{video_{i}.label_{j}.t_{k}} = \frac{n_{video_{i}.label_{j}.t_{k}}}{n_{video_{k}}}
\end{equation}

where \emph{i} is one of the 232 videos, \emph{j} is one of the six \enquote{basic} emotion labels, \emph{k} for each second of the video.

\hypertarget{automatic-facial-expression-recognition}{%
\subsection{Automatic Facial Expression Recognition}\label{automatic-facial-expression-recognition}}

The 232 annotated video were processed with Affdex (SDK v3.4.1). Affdex is an automatic facial expression recognition classifier developed and distributed by Affectiva is a spin-off company resulting from the research activities of MIT media lab created in 2009 (McDuff, Mahmoud, Mavadati, Amr, Turcot, \& el Kaliouby, 2016). Affdex's algorithm uses Histogram of Oriented Gradient (HOG) features and Support Vector Machine (SVM) classifiers in order to recognize facial expressions. For each video frame, Affdex identifies the probability p\_(video\_i.label\_j.t\_k ) from 0 to 100 (rescaled to 0 to 1 for the analysis) of the face as expressing one of the six \enquote{basic} emotion labels (i.e., anger, disgust, fear, happiness, surprise and sadness) as well as additional psychological states such as valence, engagement or contempt, and facial features such as cheek raise, eye widen or jaw drop.

For both human and automatic recognition, to determine which of the six \enquote{basic} emotions can be used to identify each video, the recognition probability for each label by frame was converted into odd ratio by label {[}22{]}. The highest sum of each odd ratio time-series defines the label recognized by the automatic classifier (EQ\ref{eq:2} for human recognition and EQ\ref{eq:3} for automatic recognition).
\begin{equation}
\label{eq:2}
video_{i}.label = \max\left(\frac{\sum_{k=1}^{n}x_{video_{i}.label_{j}.t_{k}}}{\sum_{k=1}^{n}x_{video_{i}.t_{k}}}\right)
\end{equation}

\begin{equation}
\label{eq:3}
video_{i}.label = \max\left(\frac{\sum_{k=1}^{f}p_{video_{i}.label_{j}.t_{k}}}{\sum_{k=1}^{f}p_{video_{i}.t_{k}}}\right)
\end{equation}

where \emph{i} is one of the 232 videos, \emph{j} is one of the six \enquote{basic} emotion labels, \emph{k} for each second of the \emph{n} second video or for each frame of the \emph{f} frame video.

\hypertarget{results}{%
\section{Results}\label{results}}

Since encoders' self-reports, human annotations and the automatic recognition include data on \enquote{non-basic} emotion labels and features, the analysis is performed using only the six \enquote{basic} emotion labels in order to compare them. The maximum score for self-reports, human annotations and automatic recognition is used to label the video. In case of more than one label obtaining the maximum value, the video is labeled as undetermined.

\hypertarget{human-observers-accuracy}{%
\subsection{Human Observers' Accuracy}\label{human-observers-accuracy}}

The overall correlation of recognition and non-recognition between self-reported emotions and human observers recognition is significant but low (\(r = .24\), 95\% CI \([.19\), \(.29]\), \(t(1384) = 9.21\), \(p < .001\)). In order to identify differences according to the emotional labels, encoders' subjective feelings are compared with human observers' recognition in a confusion matrix (Figure \ref{fig:confusionMatrix_sr_hr}).

\begin{figure}
\centering
\includegraphics{manuscript_apa_files/figure-latex/confusionMatrix_sr_hr-1.pdf}
\caption{(\#fig:confusionMatrix\_sr\_hr)\label{fig:confusionMatrix_sr_hr}Confusion matrix between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the human observers.}
\end{figure}

Each emotion label used to describe encoders' self-reported subjective feeling (\emph{i.e.}, the label rated with the highest value) is compared with the emotion labels which were rated with the highest score by human observers. Results of the confusion matrix show a low agreement between the emotion felt by the encoder during the elicitation and the emotion recognized by the human observers (Accuracy \nolinebreak = \nolinebreak 0.43, 95\% CI {[}0.35,0.52{]}; Kappa = 0.19) except for \emph{disgust} (100\% of the videos self-reported). These results are far from those classically obtained in the literature for emotional facial expression recognition which ranges between 60\% and 80\% accuracy. However these results are mostly obtained with static (\emph{i.e.}, pictures) and posed (\emph{i.e.}, displayed by actors) facial expressions using only 6 emotional labels in a forced-choice paradigm.

Interestingly human observers seem to recognize surprise expressed in videos where anger, fear happiness and sadness was the highest self-reported emotion (respectively 28.6\%, 33.3\%, 34.7\% and 33.3\% of the videos self-reported), and in a lower instance \emph{happiness} was recognized in videos where \emph{fear} and \emph{surprise} was the highest self-reported emotion (respectively 50.0\% and 25.6\% of the videos self-reported).

Sensitivity, specificity, precision and F1 scores for each emotion reveals that \emph{happiness} has the highest coherence ratio whereas \emph{sadness} has the lowest coherence ratio between true positives and false positives (Table \ref{table:confusionTable_sr_hr}).

\textbackslash{}begin\{table\}{[}H{]}

\textbackslash{}caption\{(\#tab:confusionTable\_sr\_hr)\label{table:confusionTable_sr_hr}Human recognition accuracy metrics for each emotion.\}
\centering
\fontsize{8}{10}\selectfont

\begin{tabu} to \linewidth {>{\raggedright}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
\toprule
Emotion & Sensitivity & Specificity & Precision & F1\\
\midrule
anger & 0.43 & 0.93 & 0.23 & 0.3\\
disgust & 1.00 & 0.94 & 0.33 & 0.5\\
fear & 0.00 & 0.97 & 0.00 & \textit{na.}\\
happiness & 0.49 & 0.78 & 0.69 & 0.57\\
sadness & 0.08 & 0.92 & 0.08 & 0.08\\
surprise & 0.44 & 0.67 & 0.37 & 0.4\\
\bottomrule
\multicolumn{5}{l}{\textsuperscript{} Note. \textit{na.} values are produced when not enough}\\
\multicolumn{5}{l}{data are available to compute accuracy indicators.}\\
\end{tabu}

\textbackslash{}end\{table\}

Accuracy metrics by emotional labels indicate a discrepancy in the ratio of true/false positives. Whereas \emph{happiness} and \emph{disgust} obtain the highest scores, \emph{anger}, \emph{surprise} and \emph{sadness} have the lowest recognition ratio. The underlying effect of expression intensity may explain why \emph{happiness} and \emph{disgust} are easily recognized. \emph{Anger} and \emph{sadness} as non-socially desirable emotions may be have been felt but not expressed.

However, self-reports show a significant proportion of undetermined emotional states (35.2\% of the 358 videos) which reveals the potential limit of using 6-points Likert scales to measure emotional self-reports. Indeed, encoders can easily score to the maximum for more than one emotion.

\hypertarget{automatic-classifiers-accuracy}{%
\subsection{Automatic Classifier's Accuracy}\label{automatic-classifiers-accuracy}}

Similarly to the previous analysis, the overall correlation of recognition and non-recognition revealed a significant but very low coherence between self-reported emotions and automatic classifier's recognition (\(r = .12\), 95\% CI \([.07\), \(.17]\), \(t(1384) = 4.50\), \(p < .001\)). A confusion matrix was used to compare encoders' subjective feeling with the emotion label recognized by the automatic classifier (Figure \ref{fig:confusionMatrix_sr_ar}).

\begin{figure}
\centering
\includegraphics{manuscript_apa_files/figure-latex/confusionMatrix_sr_ar-1.pdf}
\caption{(\#fig:confusionMatrix\_sr\_ar)\label{fig:confusionMatrix_sr_ar}Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the automatic classifier.}
\end{figure}

Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are somewhat similar to the ones with human observers (Table \ref{table:confusionTable_sr_ar}). Overall, a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy = 0.30, 95\% CI {[}0.22,0.38{]}; Kappa = 0.07) except for \emph{happiness} (44.4\% of the video self-reported) is evident.

\textbackslash{}begin\{table\}{[}H{]}

\textbackslash{}caption\{(\#tab:confusionTable\_sr\_ar)\label{table:confusionTable_sr_ar}Autonatic recognition accuracy metrics for each emotion.\}
\centering
\fontsize{8}{10}\selectfont

\begin{tabu} to \linewidth {>{\raggedright}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
\toprule
Emotion & Sensitivity & Specificity & Precision & F1\\
\midrule
anger & 0.00 & 0.90 & 0.00 & \textit{na.}\\
disgust & 0.25 & 0.72 & 0.02 & 0.04\\
fear & 0.00 & 0.94 & 0.00 & \textit{na.}\\
happiness & 0.44 & 0.73 & 0.62 & 0.52\\
sadness & 0.15 & 0.95 & 0.25 & 0.19\\
surprise & 0.19 & 0.86 & 0.36 & 0.25\\
\bottomrule
\multicolumn{5}{l}{\textsuperscript{} Note. \textit{na.} values are produced when not enough}\\
\multicolumn{5}{l}{data are available to compute accuracy indicators.}\\
\end{tabu}

\textbackslash{}end\{table\}

Surprisingly the automatic classifier incorrectly recognized as \emph{disgust} an significant proportion of videos in which \emph{anger}, \emph{happiness} and \emph{surprise} was the highest self-reported emotion (respectively 42.9\%, 25.0\% and 32.6\% of the videos self-reported). In parallel, the automatic classifier recognized as \emph{happiness} videos in which \emph{fear} and \emph{surprise} was the highest self-reported emotion (respectively 66.7\% and 32.6\% of the videos self-reported).

A comparable explanation involving the amount of undetermined video based on self-reports can be provided, as the level of undetermined emotions are very high for the self- reports.

\hypertarget{comparison-between-human-and-automatic-recognition}{%
\subsection{Comparison Between Human and Automatic Recognition}\label{comparison-between-human-and-automatic-recognition}}

As previously mentioned, human observers appear to be more accurate than the automatic classifier to recognize an individual's subjective feeling (human observers Accuracy = 0.43; automatic classifier Accuracy = 0.30; \(r = .22\), 95\% CI \([.17\), \(.27]\), \(t(1384) = 8.31\), \(p < .001\)). However, both make mistakes.

A third confusion matrix is used to compare similarities (diagonal) and differences between human observers and automatic classifier in classifying the six emotion labels (Figure \ref{fig:confusionMatrix_hr_ar}).

\begin{figure}
\centering
\includegraphics{manuscript_apa_files/figure-latex/confusionMatrix_hr_ar-1.pdf}
\caption{(\#fig:confusionMatrix\_hr\_ar)\label{fig:confusionMatrix_hr_ar}Proportion of emotion labels classified by human observers which are recognized by the automatif classifier.}
\end{figure}

The overall agreement between human observers and the automatic classifier is in fact very low (Kappa = 0.18). Except for \emph{happiness} and \emph{disgust} (respectively 64.9\% and 50.0\% of common labelling), there is no clear common pattern. Moreover, the automatic classifier has a tendency to label as \emph{disgust} videos labeled as \emph{sadness} by human observers, and as \emph{happiness} videos labeled as \emph{fear} by human observers.

\hypertarget{discussion-and-conclusion}{%
\subsection{Discussion and Conclusion}\label{discussion-and-conclusion}}

Despite being one of the most investigated questions in affective science, the consistency between emotion felt and facially displayed on the one hand, and facial expression recognized on the other is a hot topic. To date no clear evidence has been found to definitively solve the questions raised. Yet, with the growing interest of industries and government in monitoring individual's psychological states, this issue is under intense scrutiny. The present research aims to provide some empirical data to answer some of the questions posed. The faces encoders spontaneously displayed when confronted with an emotional eliciting task were submitted both to human and to automatic recognition. The criterion for recognition accuracy was the subjective feeling self-reported by the encoder once the elicitation task was carried out. Results first reveal a low consistency between emotion felt and facial expression displayed. They show that facial expressions of emotion are often not displayed when the Basic Emotion View would predict them to be expressed. Secondly, results show low accuracy rates for both humans and the automatic classifier in identifying the inner emotional states of these encoders based on their facial expressions. Thirdly, human observers prove to be better at recognizing the emotion facially expressed than the automatic recognition tool is. Such results support the hypothesis advanced by some authors of low emotion--expression coherence (Kappas 2003). In many instances, facial displays are not associated with a concordant emotional state, even any emotional state at all (Bonanno \& Keltner 2004; Fernández-Dols \& Crivelli, 2013{]}. More and more evidence is showing that facial expressions are in reality not expressing emotions (McKeown 2013{]}. Increasing studies show that, for most emotions, the EFE elicited by emotional triggers are scarce and partial, even when micro-expressions are taken into account (Dúran et al., 2017). These studies are conducted either in laboratory settings or in the field. This is not to say that facial expression is not an informative modality for understanding the emotional state of a person. There is indeed an affinity between emotion and facial display (Frijda \& Tcherkassof, 1997). However, claiming that there are unique \enquote{signatures} allowing using configurations of facial muscles to infer the presence of a specific emotion is misleading. As well as other nonverbal behaviors, facial movements are not only assumed to be determined by emotion but also by various other causes, such as psychological states (e.g., motivations or pain), to say nothing of social context and sociocultural norms (Ekman et al.~1987). Hence, facial movements have causes and functions other than the expression of emotion. This multiple determination excludes any possibility of drawing a linear inference from facial activity on the underlying psychological state (emotional or other). Beyond the present observations showing a weak consistency between subjective feelings and spontaneous facial expressions, this study sheds some light on the controversy between the Basic Emotion View and the constructivist approach as to the facial recognition issue. The former (Ekman 1992) assumes that expressions of emotion are brief and coherent patterns of facial muscle movements that co-vary with discrete subjective experiences. In return, this information displayed by the face corresponds to the one extracted by perceivers. Instead of viewing emotions as natural kinds (Barrett, 2006), the constructivist approach supposes that emotions are social constructions and that facial behaviors intrinsically situated. The emotions that are recognized by the observer are constructed in her/his mind. Therefore, facial movements do not express specific emotions because they do not carry intrinsic emotional signification. It is the observer that infers the emotional meaning of the facial expression, this interpretation depending on availability of different kinds of information (contextual information, linguistic categories\ldots{}). As a consequence, one can predict from the first line of thinking that individuals' emotional subjective feeling should be correlated to the recognition of facial expressions from both human observers and automatic classifiers whereas if emotions are social constructs, as stated by the second line of thinking, human observer's should be better at perceiving emotions expressed on the face than automatic classifiers.

The present findings speak against any strong version of the Basic Emotion View. The correlations between the self-measured emotions and the observed facial behaviors are low and the latter are weakly recognized. Present results plead instead in favor of the constructivist stance. They show that human observers are more accurate than the automatic recognition tool to identify an individual's subjective feeling on the basis of their face. Moreover, mistakes made by human observers look less arbitrary to the ones made by the classifier. For instance, even if a mix-up between disgust and anger is sometimes reported in recognition studies, odd confusions such as the present ones produced by the classifier have never been noted for human observers. The latter obviously make sense of the facial behavior they are witnessing. However, the human decoders in the present study were presented with faces without any contextual information which could have helped them to shape more precisely their interpretation. Hassin, Aviezer and Bentin (2013) reviewed evidence to demonstrate that faces are inherently ambiguous and that observers rely on situational cues when they process facial displays (see also Aviezer, Hassin, Ryan, Grady, Susskind, Anderson, Moscovitch \& Bentin, 2008; Barrett, Mesquita, Gendron, 2011). It happens that contextual information even shifts emotion perception. In the present case, the facial stimuli displayed to the decoders were as equivocal as are real-life facial expressions (Aviezer, Ensenberg, \& Hassin, 2017). Yet, decoders had no cue at disposal. So, without any possibility of integrating faces and context, their decoding accuracy is pretty much degraded as it is usual in such cases (e.g., Wagner, MacDonald, \& Manstead, 1986). However, they demonstrate their superiority on the classifier tool probably thanks to their capacity to rely on their previous personal experience to invent a context in which a face could display such an expression. Once a credible context retrieved, they can affix an emotional label to the facial behavior.

Several limitations should be stated, highlighting the need for further research. One of them is effectiveness of the emotion elicitation tasks. One can consider that an intensity threshold needs to be exceeded for a visible emotional expression to occur. In the case of the present study, it may be that insufficient emotion intensity accounts for the low number of visibly reactive participants. However, such line of reasoning fits badly with the Basic Emotion View according to which emotions and their related prototypical facial behavior are universal because they are considered as innate mechanisms allowing individuals to respond adaptively to evolutionary significant events (threats, opportunities\ldots{}) encountered in the environment, whatever their potency. Moreover, various studies have used fairly strong emotion inductions without obtaining any visible facial display of any kind of emotion (Dúran et al, 2017). Besides, the present results show a quite good correlation between reported emotion and facial behavior for happiness. Thus, there are no reasons to believe that an explanation in terms of a too low emotion threshold applies for the other emotions under consideration. The latter were triggered and measured with conceptually identical elicitation and assessment procedures (Reisenzein et al., 2013). The use of self-reports to evaluate encoders' subjective feelings is another limitation that can also be put forward because of the numerous cognitive biases they entail. Well known problems with the reliability of self-reports are, among others, the reconstructive nature of memory, the influence of attentional biases on reports, demand characteristics, distorting effects of implicit causal theories and personal motives, as stressed by Nielsen \& Kaszniak (2007). It is clearly obvious that self-reports are not simple outlets of inner mental processes but personal constructions and they are affected by many factors (Kappas, 2003). The emotional feeling echoes motivational tendencies, bodily changes, and cognitive appraisals of events (Sander \& Scherer, 2009). All these are encapsulated into semantic categories referred to by labels. As things stand currently, there is unfortunately no objective way for accessing and assessing inner subjective emotional feelings except for asking people to report their subjective experience in words. The procedure used for human recognition can also be open to dispute. First, one can criticize the decoders' expertise level since they were not FACS coders but untrained students. Of course one can expect a difference between skilled annotators and novice ones regarding the assessment of emotional facial displays. This said, in everyday life, few people are specialist coders yet the quantity of successful social interactions proves lay persons recognize pretty well others' facial behaviors. Therefore, and especially in the interest of generalizability, asking inexpert people seems relevant. Secondly, instead of using a classic forced-choice procedure, a more subtle approach was chosen to mimic results provided by the automatic classifier. As explained above, annotators first delimitated a temporal sequence during which they noticed an emotional display on the face, and then attributed an emotion label to this behavioral sequence in a second step. Whereas this paradigm is longer and more complicated, it can lead to more robust results in reducing the forced-choice bias (Russell, 1993). However this procedure can also reduce the human observers' accuracy. In this regard, the results of the human observation could have been more ambiguous because it is not the natural way that people are inferring meaning from facial expressions. An alternative explanation relies in reducing the recognition bias involved in the classic recognition paradigm. Classic forced-choice paradigms obtain artificially high results, thus by using a more evolved approach observers' accuracy may have been lowered. Another flaw is the lack of comparison with various facial expression recognition methods. Human recognition has only been compared to the Affdex classifier. Future studies are needed to confront human assessments with different automatic recognition methods, both frame based methods and sequence based automatic ones. This latter issue is particularly decisive. Indeed, the issue of the recognition of dynamic expressive sequences is essential because ordinary facial behavior is made up of dynamically shifting morphological features (Krumhuber, Kappas \& Manstead, 2013). This temporal information is indisputably a key feature of facial activity. Not only observer-based judgements of facial displays must be compared to automated facial analysis, but also different kinds of human recognition measurements should be undertaken. The challenge researchers are especially confronted with is to find ways to appropriately collect data regarding the perception of spontaneous and dynamic facial behavior. Finally, our understanding of facial displays as they occur in everyday interactions requires a strong emphasis on ecological concerns. The present study is a laboratory experiment. It has the advantage of controlling different parameters of the emotions investigated, such as intensity, quality, and temporal (onset, duration) features. Moreover, as encoders are alone, facing an emotional trigger, it also controls for the social context, removing its possible influence on their facial behaviors. However, it is known that encoders' imagination can influence their expressiveness, for instance when they believe that their friend is doing the same (vs a different) emotional task in another room (Jakobs, Manstead, \& Fischer, 1999). Trying to exclude social influence by leaving encoders alone may be illusionary then. From an ecological perspective, it is even a mistake to exonerate behavioral observations from social contexts. Facial activity measurement in dyadic interactions has shown that the facial behavior of the perceiver reflects sometimes more what the expresser is experiencing than what the perceiver is feeling. It is the case, for instance, of emotional mimicry in dynamic social interactions (Hess \& Fischer, 2016). It is also the case of healthy partners interacting with schizophrenic patients whose facial activity is pretty much identical (Krause, Steimer-Krause, Merten \& Burkhard, 1998). Hence, in order to better comprehend emotional communication in human relationships, experimental research should be corroborated with more ecological protocols.

Nowadays, automatic recognition systems are based on the coding of the facial muscular activations from which they infer the expressed emotion. Such automatic classifier tools take for granted that, when experienced, firstly an emotion is displayed on the face, secondly, in the form of a configuration of facial muscles that is all his own, and thirdly, which is recognized by the perceiver (be it a human being or an automatic system). These are the Basic Emotion View assertions, jeopardized by field observations and laboratory experiments on spontaneous expression of emotions, such as the present study. All raise serious objections to the supposed close relation between emotion and face. They bring up several questions regarding the role the context plays in the emission and interpretation of the so-called facial expression of emotion. The finalization of operational and effective \enquote{reading emotional faces} devices rests on their answer. As a result, despite being one of the most investigated questions in affective science, the growing interest of industries and governments in tracking individual's psychological states is hardly given satisfaction. Considering the above, the present results provide additional evidence that an individual's subjective feeling cannot be inferred from facial expressions and in our case invalidate the hypothesis of hardwired emotions unambiguously displayed on the face. Even if emotions were hardwired, in everyday life one does not observe prototypical facial expressions because of their rarity and therefore research should be focused on analyzing non-prototypical facial expressions. Advancements in identifying \enquote{non-basic} emotion labels (McDuff, 2016) as well as non-prototypical facial expression have indeed been made in the development of automatic facial expression recognition tools. However, these present results suggest that automatic facial expression recognition tools should merely evaluate facial morphology features such as action units (already evaluated in OpenFace, Baltrušaitis, Robinson, \& Morency, 2016, Affectiva's Affdex, McDuff, Mahmoud, Mavadati, Amr, Turcot, \& el Kaliouby, 2016, or Vicar Vision's FaceReader, Den Uyl \& Van Kuilenburg, 2005, to name a few) rather than inferring supposedly emotional or affective states. Trying to interpret facial displays as a means of determining underlying emotional state, in all likelihood, remains vain.

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-barrett2019emotional}{}%
Barrett, L. F., Adolphs, R., Marsella, S., Martinez, A. M., \& Pollak, S. D. (2019). Emotional expressions reconsidered: Challenges to inferring emotion from human facial movements. \emph{Psychological Science in the Public Interest}, \emph{20}(1), 1--68.

\leavevmode\hypertarget{ref-bennett2002facial}{}%
Bennett, D. S., Bendersky, M., \& Lewis, M. (2002). Facial expressivity at 4 months: A context by expression analysis. \emph{Infancy}, \emph{3}(1), 97--113.

\leavevmode\hypertarget{ref-buck1985prime}{}%
Buck, R. (1985). Prime theory: An integrated view of motivation and emotion. \emph{Psychological Review}, \emph{92}(3), 389.

\leavevmode\hypertarget{ref-camras1991development}{}%
Camras, L. A., Malatesta, C., \& Izard, C. E. (1991). The development of facial expressions in infancy.

\leavevmode\hypertarget{ref-camras2007infants}{}%
Camras, L. A., Oster, H., Bakeman, R., Meng, Z., Ujiie, T., \& Campos, J. J. (2007). Do infants show distinct negative facial expressions for fear and anger? Emotional expression in 11-month-old european american, chinese, and japanese infants. \emph{Infancy}, \emph{11}(2), 131--155.

\leavevmode\hypertarget{ref-cohn2003timing}{}%
Cohn, J. F., \& Schmidt, K. (2003). The timing of facial motion in posed and spontaneous smiles. In \emph{Active media technology} (pp. 57--69). World Scientific.

\leavevmode\hypertarget{ref-crivelli2015smiles}{}%
Crivelli, C., Carrera, P., \& Fernández-Dols, J.-M. (2015). Are smiles a sign of happiness? Spontaneous expressions of judo winners. \emph{Evolution and Human Behavior}, \emph{36}(1), 52--58.

\leavevmode\hypertarget{ref-darwin1872expression}{}%
Darwin, C. (1872). \emph{The expression of the emotions in man and animals}. London, UK: John Murray.

\leavevmode\hypertarget{ref-digirolamo2017emotion}{}%
DiGirolamo, M. A., \& Russell, J. A. (2017). The emotion seen in a face can be a methodological artifact: The process of elimination hypothesis. \emph{Emotion}, \emph{17}(3), 538.

\leavevmode\hypertarget{ref-dupre2018accuracy}{}%
Dupré, D., Andelic, N., Morrison, G., \& McKeown, G. (2018). Accuracy of three commercial automatic emotion recognition systems across different individuals and their facial expressions. In \emph{Proceedings of the international conference on pervasive computing and communications} (pp. 627--632). \url{https://doi.org/10.1109/PERCOMW.2018.8480127}

\leavevmode\hypertarget{ref-duran2017coherence}{}%
Durán, J. I., Reisenzein, R., \& Fernández-Dols, J.-M. (2017). Coherence between emotions and facial expressions, 107--129.

\leavevmode\hypertarget{ref-ekman1992argument}{}%
Ekman, P. (1992). An argument for basic emotions. \emph{Cognition \& Emotion}, \emph{6}(3-4), 169--200. \url{https://doi.org/10.1080/02699939208411068}

\leavevmode\hypertarget{ref-ekman2007directed}{}%
Ekman, P. (2007). The directed facial action task. In J. A. Coan \& J. J. B. Allen (Eds.), \emph{Handbook of emotion elicitation and assessment} (pp. 47--53). New York, NY: Oxford University Press.

\leavevmode\hypertarget{ref-ekman2011meant}{}%
Ekman, P., \& Cordaro, D. (2011). What is meant by calling emotions basic. \emph{Emotion Review}, \emph{3}(4), 364--370.

\leavevmode\hypertarget{ref-ekman1980facial}{}%
Ekman, P., Freisen, W. V., \& Ancoli, S. (1980). Facial signs of emotional experience. \emph{Journal of Personality and Social Psychology}, \emph{39}(6), 1125.

\leavevmode\hypertarget{ref-ekman1987universals}{}%
Ekman, P., Friesen, W. V., O'sullivan, M., Chan, A., Diacoyanni-Tarlatzis, I., Heider, K., \ldots{} Tomita, M. (1987). Universals and cultural differences in the judgments of facial expressions of emotion. \emph{Journal of Personality and Social Psychology}, \emph{53}(4), 712--717. \url{https://doi.org/10.1037/0022-3514.53.4.712}

\leavevmode\hypertarget{ref-ekman1988universality}{}%
Ekman, P., \& Heider, K. G. (1988). The universality of a contempt expression: A replication. \emph{Motivation and Emotion}, \emph{12}(3), 303--308. \url{https://doi.org/10.1007/BF00993116}

\leavevmode\hypertarget{ref-elfenbein2002universality}{}%
Elfenbein, H. A., \& Ambady, N. (2002). On the universality and cultural specificity of emotion recognition: A meta-analysis. \emph{Psychological Bulletin}, \emph{128}(2), 203.

\leavevmode\hypertarget{ref-fernandez1997spontaneous}{}%
Fernandez-Dols, J.-M., Sanchez, F., Carrera, P., \& Ruiz-Belda, M.-A. (1997). Are spontaneous expressions and emotions linked? An experimental test of coherence. \emph{Journal of Nonverbal Behavior}, \emph{21}(3), 163--177.

\leavevmode\hypertarget{ref-fernandez2013emotion}{}%
Fernández-Dols, J.-M., \& Crivelli, C. (2013). Emotion and expression: Naturalistic studies. \emph{Emotion Review}, \emph{5}(1), 24--29. \url{https://doi.org/10.1177/1754073912457229}

\leavevmode\hypertarget{ref-fernandez1995smiles}{}%
Fernández-Dols, J.-M., \& Ruiz-Belda, M.-A. (1995). Are smiles a sign of happiness? Gold medal winners at the olympic games. \emph{Journal of Personality and Social Psychology}, \emph{69}(6), 1113.

\leavevmode\hypertarget{ref-gendron2014perceptions}{}%
Gendron, M., Roberson, D., Vyver, J. M. van der, \& Barrett, L. F. (2014). Perceptions of emotion from facial expressions are not culturally universal: Evidence from a remote culture. \emph{Emotion}, \emph{14}(2), 251.

\leavevmode\hypertarget{ref-hess1990differentiating}{}%
Hess, U., \& Kleck, R. E. (1990). Differentiating emotion elicited and deliberate emotional facial expressions. \emph{European Journal of Social Psychology}, \emph{20}(5), 369--385.

\leavevmode\hypertarget{ref-johnson2010smile}{}%
Johnson, K. J., Waugh, C. E., \& Fredrickson, B. L. (2010). Smile to see the forest: Facially expressed positive emotions broaden cognition. \emph{Cognition and Emotion}, \emph{24}(2), 299--321.

\leavevmode\hypertarget{ref-kamachi2013dynamic}{}%
Kamachi, M., Bruce, V., Mukaida, S., Gyoba, J., Yoshikawa, S., \& Akamatsu, S. (2013). Dynamic properties influence the perception of facial expressions. \emph{Perception}, \emph{42}(11), 1266--1278.

\leavevmode\hypertarget{ref-kanade2000comprehensive}{}%
Kanade, T., Cohn, J. F., \& Tian, Y. (2000). Comprehensive database for facial expression analysis. In \emph{Proceedings fourth ieee international conference on automatic face and gesture recognition (cat. No. PR00580)} (pp. 46--53). IEEE.

\leavevmode\hypertarget{ref-kraut1979social}{}%
Kraut, R. E., \& Johnston, R. E. (1979). Social and emotional messages of smiling: An ethological approach. \emph{Journal of Personality and Social Psychology}, \emph{37}(9), 1539--1553.

\leavevmode\hypertarget{ref-lee2002effect}{}%
Lee, V., \& Wagner, H. (2002). The effect of social presence on the facial and verbal expression of emotion and the interrelationships among emotion components. \emph{Journal of Nonverbal Behavior}, \emph{26}(1), 3--25.

\leavevmode\hypertarget{ref-lindquist2006language}{}%
Lindquist, K. A., Barrett, L. F., Bliss-Moreau, E., \& Russell, J. A. (2006). Language and the perception of emotion. \emph{Emotion}, \emph{6}(1), 125.

\leavevmode\hypertarget{ref-mehu2007smiles}{}%
Mehu, M., Grammer, K., \& Dunbar, R. I. (2007). Smiles when sharing. \emph{Evolution and Human Behavior}, \emph{28}(6), 415--422.

\leavevmode\hypertarget{ref-meillon2010dynemo}{}%
Meillon, B., Tcherkassof, A., Mandran, N., Adam, J.-M., Dubois, M., Dupré, D., \ldots{} Caplier, A. (2010). DynEmo: A corpus of dynamic and spontaneous emotional facial expressions. In \emph{Proceedings of international workshop series on multimodal corpora, tools and resources. Multimodal corpora: Advances in capturing, coding and analyzing multimodality} (pp. 31--36).

\leavevmode\hypertarget{ref-reisenzein2000exploring}{}%
Reisenzein, R. (2000). Exploring the strength of association between the components of emotion syndromes: The case of surprise. \emph{Cognition \& Emotion}, \emph{14}(1), 1--38.

\leavevmode\hypertarget{ref-reisenzein2006evidence}{}%
Reisenzein, R., Bördgen, S., Holtbernd, T., \& Matz, D. (2006). Evidence for strong dissociation between emotion and facial displays: The case of surprise. \emph{Journal of Personality and Social Psychology}, \emph{91}(2), 295.

\leavevmode\hypertarget{ref-reisenzein2013coherence}{}%
Reisenzein, R., Studtmann, M., \& Horstmann, G. (2013). Coherence between emotion and facial expression: Evidence from laboratory experiments. \emph{Emotion Review}, \emph{5}(1), 16--23.

\leavevmode\hypertarget{ref-rosenberg1994coherence}{}%
Rosenberg, E. L., \& Ekman, P. (1994). Coherence between expressive and experiential systems in emotion. \emph{Cognition \& Emotion}, \emph{8}(3), 201--229.

\leavevmode\hypertarget{ref-ruiz2003spontaneous}{}%
Ruiz-Belda, M.-A., Fernández-Dols, J.-M., Carrera, P., \& Barchard, K. (2003). Spontaneous facial expressions of happy bowlers and soccer fans. \emph{Cognition and Emotion}, \emph{17}(2), 315--326.

\leavevmode\hypertarget{ref-russell1993forced}{}%
Russell, J. A. (1993). Forced-choice response format in the study of facial expression. \emph{Motivation and Emotion}, \emph{17}(1), 41--51. \url{https://doi.org/10.1007/BF00995206}

\leavevmode\hypertarget{ref-russell1994there}{}%
Russell, J. A. (1994). Is there universal recognition of emotion from facial expression? A review of the cross-cultural studies. \emph{Psychological Bulletin}, \emph{115}(1), 102.

\leavevmode\hypertarget{ref-russell2003facial}{}%
Russell, J. A., Bachorowski, J.-A., \& Fernández-Dols, J.-M. (2003). Facial and vocal expressions of emotion. \emph{Annual Review of Psychology}, \emph{54}(1), 329--349.

\leavevmode\hypertarget{ref-russell1987relativity}{}%
Russell, J. A., \& Fehr, B. (1987). Relativity in the perception of emotion in facial expressions. \emph{Journal of Experimental Psychology: General}, \emph{116}(3), 223.

\leavevmode\hypertarget{ref-scherer1997lost}{}%
Scherer, K. R., \& Ceschi, G. (1997). Lost luggage: A field study of emotion--antecedent appraisal. \emph{Motivation and Emotion}, \emph{21}(3), 211--235.

\leavevmode\hypertarget{ref-schutzwohl2012facial}{}%
Schützwohl, A., \& Reisenzein, R. (2012). Facial expressions in response to a highly surprising event exceeding the field of vision: A test of darwin's theory of surprise. \emph{Evolution and Human Behavior}, \emph{33}(6), 657--664.

\leavevmode\hypertarget{ref-tcherkassof2007facial}{}%
Tcherkassof, A., Bollon, T., Dubois, M., Pansu, P., \& Adam, J.-M. (2007). Facial expressions of emotions: A methodological contribution to the study of spontaneous and dynamic emotional faces. \emph{European Journal of Social Psychology}, \emph{37}(6), 1325--1345.

\leavevmode\hypertarget{ref-tsai2002emotional}{}%
Tsai, J. L., Chentsova-Dutton, Y., Freire-Bebeau, L., \& Przymus, D. E. (2002). Emotional expression and physiology in european americans and hmong americans. \emph{Emotion}, \emph{2}(4), 380.

\leavevmode\hypertarget{ref-vernon2002disgust}{}%
Vernon, L. L., \& Berenbaum, H. (2002). Disgust and fear in response to spiders. \emph{Cognition \& Emotion}, \emph{16}(6), 809--830.

\leavevmode\hypertarget{ref-de2019mama}{}%
Waal, F. de. (2019). \emph{Mama's last hug: Animal emotions and what they teach us about ourselves}. London, UK: Granta Books.

\leavevmode\hypertarget{ref-wagner1997methods}{}%
Wagner, H. L. (1997). Methods for the study of facial behavior. In J. A. Russell \& J.-M. Fernández-Dols (Eds.), \emph{The psychology of facial expression} (pp. 31--54). Cambridge, UK: Cambridge University Press.

\leavevmode\hypertarget{ref-barrett2019emotional}{}%
Barrett, L. F., Adolphs, R., Marsella, S., Martinez, A. M., \& Pollak, S. D. (2019). Emotional expressions reconsidered: Challenges to inferring emotion from human facial movements. \emph{Psychological Science in the Public Interest}, \emph{20}(1), 1--68.

\leavevmode\hypertarget{ref-bennett2002facial}{}%
Bennett, D. S., Bendersky, M., \& Lewis, M. (2002). Facial expressivity at 4 months: A context by expression analysis. \emph{Infancy}, \emph{3}(1), 97--113.

\leavevmode\hypertarget{ref-buck1985prime}{}%
Buck, R. (1985). Prime theory: An integrated view of motivation and emotion. \emph{Psychological Review}, \emph{92}(3), 389.

\leavevmode\hypertarget{ref-camras1991development}{}%
Camras, L. A., Malatesta, C., \& Izard, C. E. (1991). The development of facial expressions in infancy.

\leavevmode\hypertarget{ref-camras2007infants}{}%
Camras, L. A., Oster, H., Bakeman, R., Meng, Z., Ujiie, T., \& Campos, J. J. (2007). Do infants show distinct negative facial expressions for fear and anger? Emotional expression in 11-month-old european american, chinese, and japanese infants. \emph{Infancy}, \emph{11}(2), 131--155.

\leavevmode\hypertarget{ref-cohn2003timing}{}%
Cohn, J. F., \& Schmidt, K. (2003). The timing of facial motion in posed and spontaneous smiles. In \emph{Active media technology} (pp. 57--69). World Scientific.

\leavevmode\hypertarget{ref-crivelli2015smiles}{}%
Crivelli, C., Carrera, P., \& Fernández-Dols, J.-M. (2015). Are smiles a sign of happiness? Spontaneous expressions of judo winners. \emph{Evolution and Human Behavior}, \emph{36}(1), 52--58.

\leavevmode\hypertarget{ref-darwin1872expression}{}%
Darwin, C. (1872). \emph{The expression of the emotions in man and animals}. London, UK: John Murray.

\leavevmode\hypertarget{ref-digirolamo2017emotion}{}%
DiGirolamo, M. A., \& Russell, J. A. (2017). The emotion seen in a face can be a methodological artifact: The process of elimination hypothesis. \emph{Emotion}, \emph{17}(3), 538.

\leavevmode\hypertarget{ref-dupre2018accuracy}{}%
Dupré, D., Andelic, N., Morrison, G., \& McKeown, G. (2018). Accuracy of three commercial automatic emotion recognition systems across different individuals and their facial expressions. In \emph{Proceedings of the international conference on pervasive computing and communications} (pp. 627--632). \url{https://doi.org/10.1109/PERCOMW.2018.8480127}

\leavevmode\hypertarget{ref-duran2017coherence}{}%
Durán, J. I., Reisenzein, R., \& Fernández-Dols, J.-M. (2017). Coherence between emotions and facial expressions, 107--129.

\leavevmode\hypertarget{ref-ekman1992argument}{}%
Ekman, P. (1992). An argument for basic emotions. \emph{Cognition \& Emotion}, \emph{6}(3-4), 169--200. \url{https://doi.org/10.1080/02699939208411068}

\leavevmode\hypertarget{ref-ekman2007directed}{}%
Ekman, P. (2007). The directed facial action task. In J. A. Coan \& J. J. B. Allen (Eds.), \emph{Handbook of emotion elicitation and assessment} (pp. 47--53). New York, NY: Oxford University Press.

\leavevmode\hypertarget{ref-ekman2011meant}{}%
Ekman, P., \& Cordaro, D. (2011). What is meant by calling emotions basic. \emph{Emotion Review}, \emph{3}(4), 364--370.

\leavevmode\hypertarget{ref-ekman1980facial}{}%
Ekman, P., Freisen, W. V., \& Ancoli, S. (1980). Facial signs of emotional experience. \emph{Journal of Personality and Social Psychology}, \emph{39}(6), 1125.

\leavevmode\hypertarget{ref-ekman1987universals}{}%
Ekman, P., Friesen, W. V., O'sullivan, M., Chan, A., Diacoyanni-Tarlatzis, I., Heider, K., \ldots{} Tomita, M. (1987). Universals and cultural differences in the judgments of facial expressions of emotion. \emph{Journal of Personality and Social Psychology}, \emph{53}(4), 712--717. \url{https://doi.org/10.1037/0022-3514.53.4.712}

\leavevmode\hypertarget{ref-ekman1988universality}{}%
Ekman, P., \& Heider, K. G. (1988). The universality of a contempt expression: A replication. \emph{Motivation and Emotion}, \emph{12}(3), 303--308. \url{https://doi.org/10.1007/BF00993116}

\leavevmode\hypertarget{ref-elfenbein2002universality}{}%
Elfenbein, H. A., \& Ambady, N. (2002). On the universality and cultural specificity of emotion recognition: A meta-analysis. \emph{Psychological Bulletin}, \emph{128}(2), 203.

\leavevmode\hypertarget{ref-fernandez1997spontaneous}{}%
Fernandez-Dols, J.-M., Sanchez, F., Carrera, P., \& Ruiz-Belda, M.-A. (1997). Are spontaneous expressions and emotions linked? An experimental test of coherence. \emph{Journal of Nonverbal Behavior}, \emph{21}(3), 163--177.

\leavevmode\hypertarget{ref-fernandez2013emotion}{}%
Fernández-Dols, J.-M., \& Crivelli, C. (2013). Emotion and expression: Naturalistic studies. \emph{Emotion Review}, \emph{5}(1), 24--29. \url{https://doi.org/10.1177/1754073912457229}

\leavevmode\hypertarget{ref-fernandez1995smiles}{}%
Fernández-Dols, J.-M., \& Ruiz-Belda, M.-A. (1995). Are smiles a sign of happiness? Gold medal winners at the olympic games. \emph{Journal of Personality and Social Psychology}, \emph{69}(6), 1113.

\leavevmode\hypertarget{ref-gendron2014perceptions}{}%
Gendron, M., Roberson, D., Vyver, J. M. van der, \& Barrett, L. F. (2014). Perceptions of emotion from facial expressions are not culturally universal: Evidence from a remote culture. \emph{Emotion}, \emph{14}(2), 251.

\leavevmode\hypertarget{ref-hess1990differentiating}{}%
Hess, U., \& Kleck, R. E. (1990). Differentiating emotion elicited and deliberate emotional facial expressions. \emph{European Journal of Social Psychology}, \emph{20}(5), 369--385.

\leavevmode\hypertarget{ref-johnson2010smile}{}%
Johnson, K. J., Waugh, C. E., \& Fredrickson, B. L. (2010). Smile to see the forest: Facially expressed positive emotions broaden cognition. \emph{Cognition and Emotion}, \emph{24}(2), 299--321.

\leavevmode\hypertarget{ref-kamachi2013dynamic}{}%
Kamachi, M., Bruce, V., Mukaida, S., Gyoba, J., Yoshikawa, S., \& Akamatsu, S. (2013). Dynamic properties influence the perception of facial expressions. \emph{Perception}, \emph{42}(11), 1266--1278.

\leavevmode\hypertarget{ref-kanade2000comprehensive}{}%
Kanade, T., Cohn, J. F., \& Tian, Y. (2000). Comprehensive database for facial expression analysis. In \emph{Proceedings fourth ieee international conference on automatic face and gesture recognition (cat. No. PR00580)} (pp. 46--53). IEEE.

\leavevmode\hypertarget{ref-kraut1979social}{}%
Kraut, R. E., \& Johnston, R. E. (1979). Social and emotional messages of smiling: An ethological approach. \emph{Journal of Personality and Social Psychology}, \emph{37}(9), 1539--1553.

\leavevmode\hypertarget{ref-lee2002effect}{}%
Lee, V., \& Wagner, H. (2002). The effect of social presence on the facial and verbal expression of emotion and the interrelationships among emotion components. \emph{Journal of Nonverbal Behavior}, \emph{26}(1), 3--25.

\leavevmode\hypertarget{ref-lindquist2006language}{}%
Lindquist, K. A., Barrett, L. F., Bliss-Moreau, E., \& Russell, J. A. (2006). Language and the perception of emotion. \emph{Emotion}, \emph{6}(1), 125.

\leavevmode\hypertarget{ref-mehu2007smiles}{}%
Mehu, M., Grammer, K., \& Dunbar, R. I. (2007). Smiles when sharing. \emph{Evolution and Human Behavior}, \emph{28}(6), 415--422.

\leavevmode\hypertarget{ref-meillon2010dynemo}{}%
Meillon, B., Tcherkassof, A., Mandran, N., Adam, J.-M., Dubois, M., Dupré, D., \ldots{} Caplier, A. (2010). DynEmo: A corpus of dynamic and spontaneous emotional facial expressions. In \emph{Proceedings of international workshop series on multimodal corpora, tools and resources. Multimodal corpora: Advances in capturing, coding and analyzing multimodality} (pp. 31--36).

\leavevmode\hypertarget{ref-reisenzein2000exploring}{}%
Reisenzein, R. (2000). Exploring the strength of association between the components of emotion syndromes: The case of surprise. \emph{Cognition \& Emotion}, \emph{14}(1), 1--38.

\leavevmode\hypertarget{ref-reisenzein2006evidence}{}%
Reisenzein, R., Bördgen, S., Holtbernd, T., \& Matz, D. (2006). Evidence for strong dissociation between emotion and facial displays: The case of surprise. \emph{Journal of Personality and Social Psychology}, \emph{91}(2), 295.

\leavevmode\hypertarget{ref-reisenzein2013coherence}{}%
Reisenzein, R., Studtmann, M., \& Horstmann, G. (2013). Coherence between emotion and facial expression: Evidence from laboratory experiments. \emph{Emotion Review}, \emph{5}(1), 16--23.

\leavevmode\hypertarget{ref-rosenberg1994coherence}{}%
Rosenberg, E. L., \& Ekman, P. (1994). Coherence between expressive and experiential systems in emotion. \emph{Cognition \& Emotion}, \emph{8}(3), 201--229.

\leavevmode\hypertarget{ref-ruiz2003spontaneous}{}%
Ruiz-Belda, M.-A., Fernández-Dols, J.-M., Carrera, P., \& Barchard, K. (2003). Spontaneous facial expressions of happy bowlers and soccer fans. \emph{Cognition and Emotion}, \emph{17}(2), 315--326.

\leavevmode\hypertarget{ref-russell1993forced}{}%
Russell, J. A. (1993). Forced-choice response format in the study of facial expression. \emph{Motivation and Emotion}, \emph{17}(1), 41--51. \url{https://doi.org/10.1007/BF00995206}

\leavevmode\hypertarget{ref-russell1994there}{}%
Russell, J. A. (1994). Is there universal recognition of emotion from facial expression? A review of the cross-cultural studies. \emph{Psychological Bulletin}, \emph{115}(1), 102.

\leavevmode\hypertarget{ref-russell2003facial}{}%
Russell, J. A., Bachorowski, J.-A., \& Fernández-Dols, J.-M. (2003). Facial and vocal expressions of emotion. \emph{Annual Review of Psychology}, \emph{54}(1), 329--349.

\leavevmode\hypertarget{ref-russell1987relativity}{}%
Russell, J. A., \& Fehr, B. (1987). Relativity in the perception of emotion in facial expressions. \emph{Journal of Experimental Psychology: General}, \emph{116}(3), 223.

\leavevmode\hypertarget{ref-scherer1997lost}{}%
Scherer, K. R., \& Ceschi, G. (1997). Lost luggage: A field study of emotion--antecedent appraisal. \emph{Motivation and Emotion}, \emph{21}(3), 211--235.

\leavevmode\hypertarget{ref-schutzwohl2012facial}{}%
Schützwohl, A., \& Reisenzein, R. (2012). Facial expressions in response to a highly surprising event exceeding the field of vision: A test of darwin's theory of surprise. \emph{Evolution and Human Behavior}, \emph{33}(6), 657--664.

\leavevmode\hypertarget{ref-tcherkassof2007facial}{}%
Tcherkassof, A., Bollon, T., Dubois, M., Pansu, P., \& Adam, J.-M. (2007). Facial expressions of emotions: A methodological contribution to the study of spontaneous and dynamic emotional faces. \emph{European Journal of Social Psychology}, \emph{37}(6), 1325--1345.

\leavevmode\hypertarget{ref-tsai2002emotional}{}%
Tsai, J. L., Chentsova-Dutton, Y., Freire-Bebeau, L., \& Przymus, D. E. (2002). Emotional expression and physiology in european americans and hmong americans. \emph{Emotion}, \emph{2}(4), 380.

\leavevmode\hypertarget{ref-vernon2002disgust}{}%
Vernon, L. L., \& Berenbaum, H. (2002). Disgust and fear in response to spiders. \emph{Cognition \& Emotion}, \emph{16}(6), 809--830.

\leavevmode\hypertarget{ref-de2019mama}{}%
Waal, F. de. (2019). \emph{Mama's last hug: Animal emotions and what they teach us about ourselves}. London, UK: Granta Books.

\leavevmode\hypertarget{ref-wagner1997methods}{}%
Wagner, H. L. (1997). Methods for the study of facial behavior. In J. A. Russell \& J.-M. Fernández-Dols (Eds.), \emph{The psychology of facial expression} (pp. 31--54). Cambridge, UK: Cambridge University Press.

\endgroup

\end{document}

### 2019-11-04 manuscript_apa.Rmd
text before modif

@@
 keywords          : "Facial expression, self-report, human observer, automatic recognition."
-wordcount         : "7920"
+wordcount         : "7998"
 
@@
 ---
+(ref:confusionMatrix-sr-hr) Confusion matrix between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the human observers.
+(ref:confusionMatrix-sr-ar) Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the automatic classifier.
+(ref:confusionMatrix-hr-ar) Proportion of emotion labels classified by human observers which are recognized by the automatif classifier.
+(ref:confusionTable-sr-hr) Human recognition accuracy metrics for each emotion.
+(ref:confusionTable-sr-ar) Autonatic recognition accuracy metrics for each emotion.
 
@@
 
-With the development of commercial automatic facial expression recognition tools [@dupre2018accuracy], industries and governments are gradually implementing this technology in order to track humans’ emotions in various scenarios (*e.g.*, in marketing, healthcare, and the automotive industry to name a few).
+With the development of commercial automatic facial expression recognition tools [@dupre2018accuracy], industries and governments are gradually implementing this technology in order to monitor humans’ emotions in various scenarios (*e.g.*, in marketing, healthcare, and the automotive industry to name a few).
 This technology rests on the premise that facial expressions provide a direct access to individuals’ subjective feelings, and that one can read the emotions displayed on a person’s face as in an open book.
@@
 
-```{r self_report_method}
+```{r self-report-method}
 self_report_gender <- self_report_data %>% 
@@
 
-```{r human_recognition_method}
+```{r human-recognition-method}
 unique_ppt_count <- human_recognition_data %>% 
@@
 
-```{r sr_hr}
+```{r sr-hr}
 comparison_sr_hr <- dplyr::inner_join(self_report_score,human_recognition_score, by = "C_Video") %>% 
@@
 The overall correlation of recognition and non-recognition between self-reported emotions and human observers recognition is significant but low (`r corr_sr_hr$full_result`).
-In order to identify differences according to the emotional labels, encoders’ subjective feelings are compared with human observers’ recognition in a confusion matrix (Figure \ref{fig:confusionMatrix_sr_hr}).
+In order to identify differences according to the emotional labels, encoders’ subjective feelings are compared with human observers’ recognition in a confusion matrix (Figure \@ref(fig:confusionMatrix-sr-hr).
 
-```{r confusionMatrix_sr_hr, fig.height=6, fig.cap="\\label{fig:confusionMatrix_sr_hr}Confusion matrix between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the human observers."}
+```{r confusionMatrix-sr-hr, fig.height=6, out.width ="\\textwidth", fig.pos = "!h", fig.cap="(ref:confusionMatrix-sr-hr)"}
 confusionMatrix_sr_hr_freq <- confusionMatrix_sr_hr$table %>% 
@@
 Each emotion label used to describe encoders’ self-reported subjective feeling (*i.e.*, the label rated with the highest value) is compared with the emotion labels which were rated with the highest score by human observers.
-Results of the confusion matrix show a low agreement between the emotion felt by the encoder during the elicitation and the emotion recognized by the human observers (Accuracy \nolinebreak = \nolinebreak `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95% CI [`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *disgust* (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_sr"]))` of the videos self-reported).
+Results of the confusion matrix show a low agreement between the emotion felt by the encoder during the elicitation and the emotion recognized by the human observers (Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95% CI [`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *disgust* (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_sr"]))` of the videos self-reported).
 These results are far from those classically obtained in the literature for emotional facial expression recognition which ranges between 60% and 80% accuracy.
@@
 
-Sensitivity, specificity, precision and F1 scores for each emotion reveals that *happiness* has the highest coherence ratio whereas *sadness* has the lowest coherence ratio between true positives and false positives (Table \ref{table:confusionTable_sr_hr}).
+Sensitivity, specificity, precision and F1 scores for each emotion reveals that *happiness* has the highest coherence ratio whereas *sadness* has the lowest coherence ratio between true positives and false positives (Table \@ref(tab:confusionTable-sr-hr)).
 
-```{r confusionTable_sr_hr, results="asis"}
+```{r confusionTable-sr-hr, results="asis"}
 confusionMatrix_sr_hr_table <- confusionMatrix_sr_hr %>%
@@
     "latex",
-    caption = "\\label{table:confusionTable_sr_hr}Human recognition accuracy metrics for each emotion.",
+    caption = "(ref:confusionTable-sr-hr)",
     booktabs = TRUE,
@@
     escape = FALSE,
-    align = c('l',rep('c', 4))
+    align = c('l', rep('c', 4)),
+    longtable = T
   ) %>%
@@
   ) %>%
-   add_footnote("Note. \\textit{na.} values are produced when not enough data are available to compute accuracy indicators.", notation="none", escape = FALSE)
+  kableExtra::footnote(
+    general_title = "",
+    general = "Note. *na.* values are produced when not enough data are available to compute accuracy indicators.", 
+    escape = FALSE,
+    footnote_as_chunk = TRUE,
+    threeparttable = TRUE
+  )
 ```
@@
 
-```{r sr_ar}
+```{r sr-ar}
 comparison_sr_ar <- dplyr::inner_join(self_report_score,automatic_recognition_score, by = "C_Video") %>% 
@@
 Similarly to the previous analysis, the overall correlation of recognition and non-recognition revealed a significant but very low coherence between self-reported emotions and automatic classifier’s recognition (`r corr_sr_ar$full_result`).
-A confusion matrix was used to compare encoders’ subjective feeling with the emotion label recognized by the automatic classifier (Figure \ref{fig:confusionMatrix_sr_ar}).
+A confusion matrix was used to compare encoders’ subjective feeling with the emotion label recognized by the automatic classifier (Figure \@ref(fig:confusionMatrix-sr-ar)).
 
-```{r confusionMatrix_sr_ar, fig.height=6, fig.cap="\\label{fig:confusionMatrix_sr_ar}Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the automatic classifier."}
+```{r confusionMatrix-sr-ar, fig.height=6, out.width ="\\textwidth", fig.pos = "!h", fig.cap="(ref:confusionMatrix-sr-ar)"}
 confusionMatrix_sr_ar_freq <- confusionMatrix_sr_ar$table %>% 
@@
 
-Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are somewhat similar to the ones with human observers (Table \ref{table:confusionTable_sr_ar}).
+Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are somewhat similar to the ones with human observers (Table \@ref(tab:confusionTable-sr-ar)).
 Overall, a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`, 95% CI [`r round(confusionMatrix_sr_ar$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_ar$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_ar$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` of the video self-reported) is evident.
 
-```{r confusionTable_sr_ar, results="asis"}
+```{r confusionTable-sr-ar, results="asis"}
 confusionMatrix_sr_ar_table <- confusionMatrix_sr_ar %>%
@@
     "latex",
-    caption = "\\label{table:confusionTable_sr_ar}Autonatic recognition accuracy metrics for each emotion.",
+    caption = "(ref:confusionTable-sr-ar)",
     booktabs = T,
@@
     escape = FALSE,
-    align = c('l',rep('c', 4))
+    align = c('l',rep('c', 4)),
+    longtable = T
   ) %>%
@@
     ) %>%
-   add_footnote("Note. \\textit{na.} values are produced when not enough data are available to compute accuracy indicators.", notation="none", escape = FALSE)
+  kableExtra::footnote(
+    general_title = "",
+    general = "Note. *na.* values are produced when not enough data are available to compute accuracy indicators.", 
+    escape = FALSE,
+    footnote_as_chunk = TRUE,
+    threeparttable = TRUE
+    )
 ```
@@
 
-A third confusion matrix is used to compare similarities (diagonal) and differences between human observers and automatic classifier in classifying the six emotion labels (Figure \ref{fig:confusionMatrix_hr_ar}).
+A third confusion matrix is used to compare similarities (diagonal) and differences between human observers and automatic classifier in classifying the six emotion labels (Figure \@ref(fig:confusionMatrix-hr-ar)).
 
-```{r confusionMatrix_hr_ar, fig.height=6, fig.cap="\\label{fig:confusionMatrix_hr_ar}Proportion of emotion labels classified by human observers which are recognized by the automatif classifier."}
+```{r confusionMatrix-hr-ar, fig.height=6, out.width ="\\textwidth", fig.pos = "!h", fig.cap="(ref:confusionMatrix-hr-ar)"}
 comparison_hr_ar <- dplyr::inner_join(human_recognition_score,automatic_recognition_score, by = "C_Video") %>% 
@@
   dplyr::filter(hr_emotion != "undetermined") %>% 
-  dplyr::mutate_if(is.character,as.factor)
+  dplyr::mutate_if(is.character, as.factor)
 
@@
 Thirdly, human observers prove to be better at recognizing the emotion facially expressed than the automatic recognition tool is.
-Such results support the hypothesis advanced by some authors of low emotion–expression coherence (Kappas 2003).
-In many instances, facial displays are not associated with a concordant emotional state, even any emotional state at all (Bonanno & Keltner 2004; Fernández-Dols & Crivelli, 2013].
-More and more evidence is showing that facial expressions are in reality not expressing emotions (McKeown 2013].
-Increasing studies show that, for most emotions, the EFE elicited by emotional triggers are scarce and partial, even when micro-expressions are taken into account (Dúran et al., 2017).
+Such results support the hypothesis advanced by some authors of low emotion–expression coherence [@kappas2003facial].
+In many instances, facial displays are not associated with a concordant emotional state, even any emotional state at all [@bonanno2004brief; @fernandez2013emotion].
+More and more evidence is showing that facial expressions are in reality not expressing emotions [@mckeown2013analogical].
+Increasing studies show that, for most emotions, the EFE elicited by emotional triggers are scarce and partial, even when micro-expressions are taken into account [@duran2017coherence].
 These studies are conducted either in laboratory settings or in the field.
 This is not to say that facial expression is not an informative modality for understanding the emotional state of a person.
-There is indeed an affinity between emotion and facial display (Frijda & Tcherkassof, 1997).
+There is indeed an affinity between emotion and facial display [@frijda1997facial].
 However, claiming that there are unique “signatures” allowing using configurations of facial muscles to infer the presence of a specific emotion is misleading.
-As well as other nonverbal behaviors, facial movements are not only assumed to be determined by emotion but also by various other causes, such as psychological states (e.g., motivations or pain), to say nothing of social context and sociocultural norms (Ekman et al.
-1987).
+As well as other nonverbal behaviors, facial movements are not only assumed to be determined by emotion but also by various other causes, such as psychological states (*e.g.*, motivations or pain), to say nothing of social context and sociocultural norms [@ekman1987universals].
 Hence, facial movements have causes and functions other than the expression of emotion.
@@
 Beyond the present observations showing a weak consistency between subjective feelings and spontaneous facial expressions, this study sheds some light on the controversy between the Basic Emotion View and the constructivist approach as to the facial recognition issue.
-The former (Ekman 1992) assumes that expressions of emotion are brief and coherent patterns of facial muscle movements that co-vary with discrete subjective experiences.
+The former [@ekman1992argument] assumes that expressions of emotion are brief and coherent patterns of facial muscle movements that co-vary with discrete subjective experiences.
 In return, this information displayed by the face corresponds to the one extracted by perceivers.
-Instead of viewing emotions as natural kinds (Barrett, 2006), the constructivist approach supposes that emotions are social constructions and that facial behaviors intrinsically situated.
+Instead of viewing emotions as natural kinds [@barrett2006emotions], the constructivist approach supposes that emotions are social constructions and that facial behaviors intrinsically situated.
 The emotions that are recognized by the observer are constructed in her/his mind.
 Therefore, facial movements do not express specific emotions because they do not carry intrinsic emotional signification.
-It is the observer that infers the emotional meaning of the facial expression, this interpretation depending on availability of different kinds of information (contextual information, linguistic categories…).
+It is the observer that infers the emotional meaning of the facial expression, this interpretation depending on availability of different kinds of information (contextual information, linguistic categories...).
 As a consequence, one can predict from the first line of thinking that individuals’ emotional subjective feeling should be correlated to the recognition of facial expressions from both human observers and automatic classifiers whereas if emotions are social constructs, as stated by the second line of thinking, human observer’s should be better at perceiving emotions expressed on the face than automatic classifiers.
@@
 In the case of the present study, it may be that insufficient emotion intensity accounts for the low number of visibly reactive participants.
-However, such line of reasoning fits badly with the Basic Emotion View according to which emotions and their related prototypical facial behavior are universal because they are considered as innate mechanisms allowing individuals to respond adaptively to evolutionary significant events (threats, opportunities…) encountered in the environment, whatever their potency.
-Moreover, various studies have used fairly strong emotion inductions without obtaining any visible facial display of any kind of emotion (Dúran et al, 2017).
+However, such line of reasoning fits badly with the Basic Emotion View according to which emotions and their related prototypical facial behavior are universal because they are considered as innate mechanisms allowing individuals to respond adaptively to evolutionary significant events (threats, opportunities...) encountered in the environment, whatever their potency.
+Moreover, various studies have used fairly strong emotion inductions without obtaining any visible facial display of any kind of emotion [@duran2017coherence].
 Besides, the present results show a quite good correlation between reported emotion and facial behavior for happiness.
 Thus, there are no reasons to believe that an explanation in terms of a too low emotion threshold applies for the other emotions under consideration.
-The latter were triggered and measured with conceptually identical elicitation and assessment procedures (Reisenzein et al., 2013).
+The latter were triggered and measured with conceptually identical elicitation and assessment procedures [@reisenzein2013coherence].
 The use of self-reports to evaluate encoders’ subjective feelings is another limitation that can also be put forward because of the numerous cognitive biases they entail.
-Well known problems with the reliability of self-reports are, among others, the reconstructive nature of memory, the influence of attentional biases on reports, demand characteristics, distorting effects of implicit causal theories and personal motives, as stressed by Nielsen & Kaszniak (2007).
-It is clearly obvious that self-reports are not simple outlets of inner mental processes but personal constructions and they are affected by many factors (Kappas, 2003).
-The emotional feeling echoes motivational tendencies, bodily changes, and cognitive appraisals of events (Sander & Scherer, 2009).
+Well known problems with the reliability of self-reports are, among others, the reconstructive nature of memory, the influence of attentional biases on reports, demand characteristics, distorting effects of implicit causal theories and personal motives, as stressed by @nielsen2007conceptual.
+It is clearly obvious that self-reports are not simple outlets of inner mental processes but personal constructions and they are affected by many factors [@kappas2003facial].
+The emotional feeling echoes motivational tendencies, bodily changes, and cognitive appraisals of events [@sander2014oxford].
 All these are encapsulated into semantic categories referred to by labels.
@@
 As explained above, annotators first delimitated a temporal sequence during which they noticed an emotional display on the face, and then attributed an emotion label to this behavioral sequence in a second step.
-Whereas this paradigm is longer and more complicated, it can lead to more robust results in reducing the forced-choice bias (Russell, 1993).
+Whereas this paradigm is longer and more complicated, it can lead to more robust results in reducing the forced-choice bias [@russell1993forced].
 However this procedure can also reduce the human observers’ accuracy.
@@
 This latter issue is particularly decisive.
-Indeed, the issue of the recognition of dynamic expressive sequences is essential because ordinary facial behavior is made up of dynamically shifting morphological features (Krumhuber, Kappas & Manstead, 2013).
+Indeed, the issue of the recognition of dynamic expressive sequences is essential because ordinary facial behavior is made up of dynamically shifting morphological features [@krumhuber2013effects].
 This temporal information is indisputably a key feature of facial activity.
@@
 Moreover, as encoders are alone, facing an emotional trigger, it also controls for the social context, removing its possible influence on their facial behaviors.
-However, it is known that encoders’ imagination can influence their expressiveness, for instance when they believe that their friend is doing the same (vs a different) emotional task in another room (Jakobs, Manstead, & Fischer, 1999).
+However, it is known that encoders’ imagination can influence their expressiveness, for instance when they believe that their friend is doing the same (*vs.* a different) emotional task in another room [@jakobs1999social].
 Trying to exclude social influence by leaving encoders alone may be illusionary then.
@@
 Facial activity measurement in dyadic interactions has shown that the facial behavior of the perceiver reflects sometimes more what the expresser is experiencing than what the perceiver is feeling.
-It is the case, for instance, of emotional mimicry in dynamic social interactions (Hess & Fischer, 2016).
-It is also the case of healthy partners interacting with schizophrenic patients whose facial activity is pretty much identical (Krause, Steimer-Krause, Merten & Burkhard, 1998).
+It is the case, for instance, of emotional mimicry in dynamic social interactions [@hess2016emotional].
+It is also the case of healthy partners interacting with schizophrenic patients whose facial activity is pretty much identical [@krause1998dyadic].
 Hence, in order to better comprehend emotional communication in human relationships, experimental research should be corroborated with more ecological protocols.
@@
 Even if emotions were hardwired, in everyday life one does not observe prototypical facial expressions because of their rarity and therefore research should be focused on analyzing non-prototypical facial expressions.
-Advancements in identifying “non-basic” emotion labels (McDuff, 2016) as well as non-prototypical facial expression have indeed been made in the development of automatic facial expression recognition tools.
-However, these present results suggest that automatic facial expression recognition tools should merely evaluate facial morphology features such as action units (already evaluated in OpenFace, Baltrušaitis, Robinson, & Morency, 2016, Affectiva’s Affdex, McDuff, Mahmoud, Mavadati, Amr, Turcot, & el Kaliouby, 2016, or Vicar Vision’s FaceReader, Den Uyl & Van Kuilenburg, 2005, to name a few) rather than inferring supposedly emotional or affective states.
+Advancements in identifying “non-basic” emotion labels as well as non-prototypical facial expression have indeed been made in the development of automatic facial expression recognition tools [@mcduff2016discovering].
+However, these present results suggest that automatic facial expression recognition tools should merely evaluate facial morphology features such as action units [already evaluated in OpenFace, @baltruvsaitis2016openface, Affectiva’s Affdex, @mcduff2016affdex, or Vicar Vision’s FaceReader, @den2005facereader, to name a few] rather than inferring supposedly emotional or affective states.
 Trying to interpret facial displays as a means of determining underlying emotional state, in all likelihood, remains vain.

### 2019-11-04 manuscript_apa.tex
text before modif

@@
 \textsuperscript{1} University Grenoble Alpes, Grenoble, France\\\textsuperscript{2} Dublin City University, Dublin, Ireland}
-\keywords{Facial expression, self-report, human observer, automatic recognition.\newline\indent Word count: 7920}
+\keywords{Facial expression, self-report, human observer, automatic recognition.\newline\indent Word count: 7998}
 \usepackage{csquotes}
@@
 }
+
 \hypertarget{introduction}{%
@@
 
-With the development of commercial automatic facial expression recognition tools (Dupré, Andelic, Morrison, \& McKeown, 2018), industries and governments are gradually implementing this technology in order to track humans' emotions in various scenarios (\emph{e.g.}, in marketing, healthcare, and the automotive industry to name a few).
+With the development of commercial automatic facial expression recognition tools (Dupré, Andelic, Morrison, \& McKeown, 2018), industries and governments are gradually implementing this technology in order to monitor humans' emotions in various scenarios (\emph{e.g.}, in marketing, healthcare, and the automotive industry to name a few).
 This technology rests on the premise that facial expressions provide a direct access to individuals' subjective feelings, and that one can read the emotions displayed on a person's face as in an open book.
@@
 
-Starting from the empirical evidence suggesting that spontaneous facial expressions in ordinary life are equivocal, Fernández-Dols (2017) argues in favor of a pragmatic conception of natural facial displays.
+Starting from the empirical evidence suggesting that spontaneous facial expressions in ordinary life are equivocal, Dols (2017) argues in favor of a pragmatic conception of natural facial displays.
 He makes a plea for the idea that natural facial displays, rather than \enquote{saying} -- because they do not have a specific meaning he claims -- \enquote{make} things.
 Facial expressions are actions in a communicative interaction.
-They do not express emotions but they \enquote{prompt, on the receiver's side, important inferences about the context, the sender, and the course of the interaction between sender and receiver} (p.466, underlined by the author).
+They do not express emotions but they \emph{\enquote{prompt, on the receiver's side, important inferences about the context, the sender, and the course of the interaction between sender and receiver}} (Dols, 2017, p.
+466, underlined by the author).
 As such, the fact that facial displays are able to signal emotions is a byproduct of one of its main function, the one of implementing actions performing practical ends.
@@
 S.he translates the continuous flow of movements into coordinated sequences of actions holding a beginning and an end.
-Facial displays are not simple strings of action units which morphological configuration would be the prototype of a given emotion, and consequently identified as such (Ekman, 1982).
+Facial displays are not simple strings of action units which morphological configuration would be the prototype of a given emotion, and consequently identified as such (Ekman et al., 1987).
 They are best conceived as a Gestalt, the same way as a string of musical notes establishes a melody (Tcherkassof \& Frijda, 2014).
 This is why even unauthentic facial displays can yet be recognized as emotional expressions.
-Guillaume Duchenne de Boulogne explains that the artist who has shaped the famous Laocoon antic sculpture, exhibited in the Vatican's museum, has made a modelling mistake since no face can display its emotional expression.
+Guillaume Duchenne de Boulogne explains that the artist who has shaped the famous Laocoon antic sculpture, exhibited in the Vatican's museum, has made a modelling mistake since no face can display its emotional expression (Duchenne, 1876).
 Indeed, no muscular contraction can produce it.
-He even rectifies the « mistake » by presenting a statue which face is shaped according to the physiology of facial expressive movements.
+He even rectifies the \enquote{mistake} by presenting a statue which face is shaped according to the physiology of facial expressive movements.
 His demonstration gives food for thought.
@@
 This example goes along the lines of the constructivist approach.
+
 The constructivist approach claims that facial displays are behaviors which meaning is inferred by the perceiver.
-Findings support this observer dependence (Lindquist \& Gendron, 2013; Niedenthal, Wood, Rychlowska \& Korn, 2017).
+Findings support this observer dependence {[}Lindquist \& Gendron (2013); niedenthal2017embodied{]}.
 They show that to make meaning of another person's facial behavior, the perceiver relies in particular on her/his knowledge about emotion categories.
@@
 They conducted their experiment among U.S.
-participants and Himba participants from remote regions of Namibia (Gendron, Roberson, van der Vyver \& Barrett, 2014) and Hazda participants of the Eastern Rift Valley of Tanzania (Gendron, Hoemann, Crittenden, Msafiri, Ruark, \& Barrett, 2018) both groups with limited exposure to Western culture.
+participants and Himba participants from remote regions of Namibia (Gendron et al., 2014) and Hazda participants of the Eastern Rift Valley of Tanzania (Gendron et al., 2018) both groups with limited exposure to Western culture.
 Gendron and her colleagues demonstrated that facial expressions were not universally \enquote{recognized} in discrete emotional terms.
-Indeed, when Himba and Hazda participants did not have emotion concepts at their disposal to structure perception, they perceived the facial expressions as behaviors (e.g., looking, smelling) that didn't have a necessary relationship to emotions.
+Indeed, when Himba and Hazda participants did not have emotion concepts at their disposal to structure perception, they perceived the facial expressions as behaviors (\emph{e.g.}, looking, smelling) that didn't have a necessary relationship to emotions.
 They didn't infer inner states (such as emotional feelings) but they rather proceeded with action identification that pointed out the functions of behaviors (see also Crivelli, Russell, Jarillo, \& Fernández-Dols, 2017 for similar observations among a small-scale society of Papua New Guinea).
@@
 
-For the emotion elicitation experiment, 358 encoding participants (182 females, 176 males, Mage = 47.9, SDage = 9.2) were recruited to perform one out of 11 emotion elicitation tasks designed to trigger a positive, a specific negative or a neutral emotional state.
+For the emotion elicitation experiment, 358 encoding participants (182 females, 176 males, \emph{M}\textsubscript{age} = 47.9, \emph{SD}\textsubscript{age} = 9.2) were recruited to perform one out of 11 emotion elicitation tasks designed to trigger a positive, a specific negative or a neutral emotional state.
 Encoders' faces were recorded using a hidden camera resulting 358 front facing 768x576 videos varying from 1s to 1479s.
-These recordings form the DynEmo database (Tcherkassof, Dupré, Meillon, Mandran, Dubois, \& Adam, 2013).
-
-After the emotion elicitation task, the encoders rated their subjective feeling on Likert scales from 0 (\enquote{not at all}) to 5 (\enquote{strongly}) related to six \enquote{basic} emotion labels (i.e., anger, disgust, fear, happiness, surprise and sadness) as well as six \enquote{non-basic} emotion labels (i.e., pride, curiosity, boredom, shame, humiliation, and disappointment).
+These recordings form the DynEmo database (Tcherkassof et al., 2013).
+
+After the emotion elicitation task, the encoders rated their subjective feeling on Likert scales from 0 (\enquote{not at all}) to 5 (\enquote{strongly}) related to six \enquote{basic} emotion labels (\emph{i.e.}, \emph{anger}, \emph{disgust}, \emph{fear}, \emph{happiness}, \emph{surprise} and \emph{sadness}) as well as six \enquote{non-basic} emotion labels (\emph{i.e.}, \emph{pride}, \emph{curiosity}, \emph{boredom}, \emph{shame}, \emph{humiliation}, and \emph{disappointment}).
 
 Finally, a debriefing session was performed to ensure that encoders were not durably affected by the emotion elicitation task.
-The debriefing was also used to check that encoders did not guess the real purpose of the experiment (e.g., being filmed while they were performing an emotional elicitation task) to guarantee facial expressions' genuineness.
+The debriefing was also used to check that encoders did not guess the real purpose of the experiment (\emph{e.g.}, being filmed while they were performing an emotional elicitation task) to guarantee facial expressions' genuineness.
 All encoders gave their agreement on their data and video to be processed for research purposes only.
@@
 Because videos have different durations, participants had to annotate a series of video corresponding to 30min long in total.
-Each video was annotated 29 times on average (SD = 12).
-The annotation of facial expressions was performed on-site using Oudjat, a software for designing video annotation experiments (Dupré, Akpan, Elias, Adam, Meillon, Bonnefond, Dubois, \& Tcherkassof (2015).
+Each video was annotated 29 times on average (\emph{SD} = 12).
+
+The annotation of facial expressions was performed on-site using \emph{Oudjat}, a software for designing video annotation experiments (Dupré et al., 2015).
 For each video, the annotation procedure followed two steps.
 First, the participants had to identify the emotional sequences by pressing the space bar of their keyboard to indicate the beginning and the end of the emotional sequences while watching the video.
-Second, the participants watched each emotional sequence previously identified and labeled the sequence using one of the 12 emotions proposed including six \enquote{basic} emotion labels (i.e., anger, disgust, fear, happiness, surprise and sadness) and six \enquote{non-basic} emotion labels (i.e., pride, curiosity, boredom, shame, humiliation, and disappointment).
+Second, the participants watched each emotional sequence previously identified and labeled the sequence using one of the 12 emotions proposed including six \enquote{basic} emotion labels (\emph{i.e.}, \emph{anger}, \emph{disgust}, \emph{fear}, \emph{happiness}, \emph{surprise} and \emph{sadness}) and six \enquote{non-basic} emotion labels (\emph{i.e.}, \emph{pride}, \emph{curiosity}, \emph{boredom}, \emph{shame}, \emph{humiliation}, and \emph{disappointment}).
 They also had the possibility to indicate that the sequence was expressing none of the proposed emotion labels.
+
 This annotation procedure results in a uni-dimensional time-series for each video per human observer identifying for each second of the video which emotion was recognized.
@@
 The 232 annotated video were processed with Affdex (SDK v3.4.1).
-Affdex is an automatic facial expression recognition classifier developed and distributed by Affectiva is a spin-off company resulting from the research activities of MIT media lab created in 2009 (McDuff, Mahmoud, Mavadati, Amr, Turcot, \& el Kaliouby, 2016).
+Affdex is an automatic facial expression recognition classifier developed and distributed by Affectiva is a spin-off company resulting from the research activities of MIT media lab created in 2009 (McDuff et al., 2016).
 Affdex's algorithm uses Histogram of Oriented Gradient (HOG) features and Support Vector Machine (SVM) classifiers in order to recognize facial expressions.
-For each video frame, Affdex identifies the probability p\_(video\_i.label\_j.t\_k ) from 0 to 100 (rescaled to 0 to 1 for the analysis) of the face as expressing one of the six \enquote{basic} emotion labels (i.e., anger, disgust, fear, happiness, surprise and sadness) as well as additional psychological states such as valence, engagement or contempt, and facial features such as cheek raise, eye widen or jaw drop.
-
-For both human and automatic recognition, to determine which of the six \enquote{basic} emotions can be used to identify each video, the recognition probability for each label by frame was converted into odd ratio by label {[}22{]}.
+For each video frame, Affdex identifies the probability \(p_{video_{i}.label_{j}.t_{k}}\) from 0 to 100 (rescaled to 0 to 1 for the analysis) of the face as expressing one of the six \enquote{basic} emotion labels (\emph{i.e.}, \emph{anger}, \emph{disgust}, \emph{fear}, \emph{happiness}, \emph{surprise} and \emph{sadness}) as well as additional psychological states such as \emph{valence}, \emph{engagement} or \emph{contempt}, and facial features such as \emph{cheek raise}, \emph{eye widen} or \emph{jaw drop}.
+
+For both human and automatic recognition, to determine which of the six \enquote{basic} emotions can be used to identify each video, the recognition probability for each label by frame was converted into odd ratio by label (Dente, Küster, Skora, \& Krumhuber, 2017).
 The highest sum of each odd ratio time-series defines the label recognized by the automatic classifier (EQ\ref{eq:2} for human recognition and EQ\ref{eq:3} for automatic recognition).
+
 \begin{equation}
@@
 The overall correlation of recognition and non-recognition between self-reported emotions and human observers recognition is significant but low (\(r = .24\), 95\% CI \([.19\), \(.29]\), \(t(1384) = 9.21\), \(p < .001\)).
-In order to identify differences according to the emotional labels, encoders' subjective feelings are compared with human observers' recognition in a confusion matrix (Figure \ref{fig:confusionMatrix_sr_hr}).
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_apa_files/figure-latex/confusionMatrix_sr_hr-1.pdf}
-\caption{(\#fig:confusionMatrix\_sr\_hr)\label{fig:confusionMatrix_sr_hr}Confusion matrix between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the human observers.}
+In order to identify differences according to the emotional labels, encoders' subjective feelings are compared with human observers' recognition in a confusion matrix (Figure \ref{fig:confusionMatrix-sr-hr}.
+
+\begin{figure}[!h]
+\includegraphics[width=\textwidth]{manuscript_apa_files/figure-latex/confusionMatrix-sr-hr-1} \caption{Confusion matrix between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the human observers.}\label{fig:confusionMatrix-sr-hr}
 \end{figure}
@@
 Each emotion label used to describe encoders' self-reported subjective feeling (\emph{i.e.}, the label rated with the highest value) is compared with the emotion labels which were rated with the highest score by human observers.
-Results of the confusion matrix show a low agreement between the emotion felt by the encoder during the elicitation and the emotion recognized by the human observers (Accuracy \nolinebreak = \nolinebreak 0.43, 95\% CI {[}0.35,0.52{]}; Kappa = 0.19) except for \emph{disgust} (100\% of the videos self-reported).
+Results of the confusion matrix show a low agreement between the emotion felt by the encoder during the elicitation and the emotion recognized by the human observers (Accuracy = 0.43, 95\% CI {[}0.35,0.52{]}; Kappa = 0.19) except for \emph{disgust} (100\% of the videos self-reported).
 These results are far from those classically obtained in the literature for emotional facial expression recognition which ranges between 60\% and 80\% accuracy.
@@
 
-Sensitivity, specificity, precision and F1 scores for each emotion reveals that \emph{happiness} has the highest coherence ratio whereas \emph{sadness} has the lowest coherence ratio between true positives and false positives (Table \ref{table:confusionTable_sr_hr}).
-
-\textbackslash{}begin\{table\}{[}H{]}
-
-\textbackslash{}caption\{(\#tab:confusionTable\_sr\_hr)\label{table:confusionTable_sr_hr}Human recognition accuracy metrics for each emotion.\}
-\centering
-\fontsize{8}{10}\selectfont
-
-\begin{tabu} to \linewidth {>{\raggedright}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
+Sensitivity, specificity, precision and F1 scores for each emotion reveals that \emph{happiness} has the highest coherence ratio whereas \emph{sadness} has the lowest coherence ratio between true positives and false positives (Table \ref{tab:confusionTable-sr-hr}).
+
+\begingroup\fontsize{8}{10}\selectfont
+
+\begin{ThreePartTable}
+\begin{TableNotes}[para]
+\item Note. textit{na.} values are produced when not enough data are available to compute accuracy indicators.
+\end{TableNotes}
+\begin{longtabu} to \linewidth {>{\raggedright}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
+\caption{\label{tab:confusionTable-sr-hr}Human recognition accuracy metrics for each emotion.}\\
 \toprule
@@
 \bottomrule
-\multicolumn{5}{l}{\textsuperscript{} Note. \textit{na.} values are produced when not enough}\\
-\multicolumn{5}{l}{data are available to compute accuracy indicators.}\\
-\end{tabu}
-
-\textbackslash{}end\{table\}
+\insertTableNotes
+\end{longtabu}
+\end{ThreePartTable}
+\endgroup{}
 
@@
 Similarly to the previous analysis, the overall correlation of recognition and non-recognition revealed a significant but very low coherence between self-reported emotions and automatic classifier's recognition (\(r = .12\), 95\% CI \([.07\), \(.17]\), \(t(1384) = 4.50\), \(p < .001\)).
-A confusion matrix was used to compare encoders' subjective feeling with the emotion label recognized by the automatic classifier (Figure \ref{fig:confusionMatrix_sr_ar}).
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_apa_files/figure-latex/confusionMatrix_sr_ar-1.pdf}
-\caption{(\#fig:confusionMatrix\_sr\_ar)\label{fig:confusionMatrix_sr_ar}Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the automatic classifier.}
+A confusion matrix was used to compare encoders' subjective feeling with the emotion label recognized by the automatic classifier (Figure \ref{fig:confusionMatrix-sr-ar}).
+
+\begin{figure}[!h]
+\includegraphics[width=\textwidth]{manuscript_apa_files/figure-latex/confusionMatrix-sr-ar-1} \caption{Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the automatic classifier.}\label{fig:confusionMatrix-sr-ar}
 \end{figure}
 
-Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are somewhat similar to the ones with human observers (Table \ref{table:confusionTable_sr_ar}).
+Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are somewhat similar to the ones with human observers (Table \ref{tab:confusionTable-sr-ar}).
 Overall, a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy = 0.30, 95\% CI {[}0.22,0.38{]}; Kappa = 0.07) except for \emph{happiness} (44.4\% of the video self-reported) is evident.
 
-\textbackslash{}begin\{table\}{[}H{]}
-
-\textbackslash{}caption\{(\#tab:confusionTable\_sr\_ar)\label{table:confusionTable_sr_ar}Autonatic recognition accuracy metrics for each emotion.\}
-\centering
-\fontsize{8}{10}\selectfont
-
-\begin{tabu} to \linewidth {>{\raggedright}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
+\begingroup\fontsize{8}{10}\selectfont
+
+\begin{ThreePartTable}
+\begin{TableNotes}[para]
+\item Note. textit{na.} values are produced when not enough data are available to compute accuracy indicators.
+\end{TableNotes}
+\begin{longtabu} to \linewidth {>{\raggedright}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
+\caption{\label{tab:confusionTable-sr-ar}Autonatic recognition accuracy metrics for each emotion.}\\
 \toprule
@@
 
-A third confusion matrix is used to compare similarities (diagonal) and differences between human observers and automatic classifier in classifying the six emotion labels (Figure \ref{fig:confusionMatrix_hr_ar}).
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_apa_files/figure-latex/confusionMatrix_hr_ar-1.pdf}
-\caption{(\#fig:confusionMatrix\_hr\_ar)\label{fig:confusionMatrix_hr_ar}Proportion of emotion labels classified by human observers which are recognized by the automatif classifier.}
+A third confusion matrix is used to compare similarities (diagonal) and differences between human observers and automatic classifier in classifying the six emotion labels (Figure \ref{fig:confusionMatrix-hr-ar}).
+
+\begin{figure}[!h]
+\includegraphics[width=\textwidth]{manuscript_apa_files/figure-latex/confusionMatrix-hr-ar-1} \caption{Proportion of emotion labels classified by human observers which are recognized by the automatif classifier.}\label{fig:confusionMatrix-hr-ar}
 \end{figure}
@@
 Thirdly, human observers prove to be better at recognizing the emotion facially expressed than the automatic recognition tool is.
-Such results support the hypothesis advanced by some authors of low emotion--expression coherence (Kappas 2003).
-In many instances, facial displays are not associated with a concordant emotional state, even any emotional state at all (Bonanno \& Keltner 2004; Fernández-Dols \& Crivelli, 2013{]}.
-More and more evidence is showing that facial expressions are in reality not expressing emotions (McKeown 2013{]}.
-Increasing studies show that, for most emotions, the EFE elicited by emotional triggers are scarce and partial, even when micro-expressions are taken into account (Dúran et al., 2017).
+Such results support the hypothesis advanced by some authors of low emotion--expression coherence (Kappas, 2003).
+In many instances, facial displays are not associated with a concordant emotional state, even any emotional state at all (Bonanno \& Keltner, 2004; Fernández-Dols \& Crivelli, 2013).
+More and more evidence is showing that facial expressions are in reality not expressing emotions (McKeown, 2013).
+Increasing studies show that, for most emotions, the EFE elicited by emotional triggers are scarce and partial, even when micro-expressions are taken into account (Durán et al., 2017).
 These studies are conducted either in laboratory settings or in the field.
@@
 However, claiming that there are unique \enquote{signatures} allowing using configurations of facial muscles to infer the presence of a specific emotion is misleading.
-As well as other nonverbal behaviors, facial movements are not only assumed to be determined by emotion but also by various other causes, such as psychological states (e.g., motivations or pain), to say nothing of social context and sociocultural norms (Ekman et al.~1987).
+As well as other nonverbal behaviors, facial movements are not only assumed to be determined by emotion but also by various other causes, such as psychological states (\emph{e.g.}, motivations or pain), to say nothing of social context and sociocultural norms (Ekman et al., 1987).
 Hence, facial movements have causes and functions other than the expression of emotion.
@@
 Beyond the present observations showing a weak consistency between subjective feelings and spontaneous facial expressions, this study sheds some light on the controversy between the Basic Emotion View and the constructivist approach as to the facial recognition issue.
-The former (Ekman 1992) assumes that expressions of emotion are brief and coherent patterns of facial muscle movements that co-vary with discrete subjective experiences.
+The former (Ekman, 1992) assumes that expressions of emotion are brief and coherent patterns of facial muscle movements that co-vary with discrete subjective experiences.
 In return, this information displayed by the face corresponds to the one extracted by perceivers.
@@
 However, such line of reasoning fits badly with the Basic Emotion View according to which emotions and their related prototypical facial behavior are universal because they are considered as innate mechanisms allowing individuals to respond adaptively to evolutionary significant events (threats, opportunities\ldots{}) encountered in the environment, whatever their potency.
-Moreover, various studies have used fairly strong emotion inductions without obtaining any visible facial display of any kind of emotion (Dúran et al, 2017).
+Moreover, various studies have used fairly strong emotion inductions without obtaining any visible facial display of any kind of emotion (Durán et al., 2017).
 Besides, the present results show a quite good correlation between reported emotion and facial behavior for happiness.
@@
 It is clearly obvious that self-reports are not simple outlets of inner mental processes but personal constructions and they are affected by many factors (Kappas, 2003).
-The emotional feeling echoes motivational tendencies, bodily changes, and cognitive appraisals of events (Sander \& Scherer, 2009).
+The emotional feeling echoes motivational tendencies, bodily changes, and cognitive appraisals of events (Sander \& Scherer, 2014).
 All these are encapsulated into semantic categories referred to by labels.
@@
 This latter issue is particularly decisive.
-Indeed, the issue of the recognition of dynamic expressive sequences is essential because ordinary facial behavior is made up of dynamically shifting morphological features (Krumhuber, Kappas \& Manstead, 2013).
+Indeed, the issue of the recognition of dynamic expressive sequences is essential because ordinary facial behavior is made up of dynamically shifting morphological features (Krumhuber, Kappas, \& Manstead, 2013).
 This temporal information is indisputably a key feature of facial activity.
@@
 Moreover, as encoders are alone, facing an emotional trigger, it also controls for the social context, removing its possible influence on their facial behaviors.
-However, it is known that encoders' imagination can influence their expressiveness, for instance when they believe that their friend is doing the same (vs a different) emotional task in another room (Jakobs, Manstead, \& Fischer, 1999).
+However, it is known that encoders' imagination can influence their expressiveness, for instance when they believe that their friend is doing the same (\emph{vs.} a different) emotional task in another room (Jakobs, Manstead, \& Fischer, 1999).
 Trying to exclude social influence by leaving encoders alone may be illusionary then.
@@
 It is the case, for instance, of emotional mimicry in dynamic social interactions (Hess \& Fischer, 2016).
-It is also the case of healthy partners interacting with schizophrenic patients whose facial activity is pretty much identical (Krause, Steimer-Krause, Merten \& Burkhard, 1998).
+It is also the case of healthy partners interacting with schizophrenic patients whose facial activity is pretty much identical (Krause, Steimer-Krause, Merten, \& Ullrich, 1998).
 Hence, in order to better comprehend emotional communication in human relationships, experimental research should be corroborated with more ecological protocols.
@@
 Even if emotions were hardwired, in everyday life one does not observe prototypical facial expressions because of their rarity and therefore research should be focused on analyzing non-prototypical facial expressions.
-Advancements in identifying \enquote{non-basic} emotion labels (McDuff, 2016) as well as non-prototypical facial expression have indeed been made in the development of automatic facial expression recognition tools.
-However, these present results suggest that automatic facial expression recognition tools should merely evaluate facial morphology features such as action units (already evaluated in OpenFace, Baltrušaitis, Robinson, \& Morency, 2016, Affectiva's Affdex, McDuff, Mahmoud, Mavadati, Amr, Turcot, \& el Kaliouby, 2016, or Vicar Vision's FaceReader, Den Uyl \& Van Kuilenburg, 2005, to name a few) rather than inferring supposedly emotional or affective states.
+Advancements in identifying \enquote{non-basic} emotion labels as well as non-prototypical facial expression have indeed been made in the development of automatic facial expression recognition tools (McDuff, 2016).
+However, these present results suggest that automatic facial expression recognition tools should merely evaluate facial morphology features such as action units (already evaluated in OpenFace, Baltrušaitis, Robinson, \& Morency, 2016, Affectiva's Affdex, McDuff et al.
+(2016), or Vicar Vision's FaceReader, Den Uyl and Van Kuilenburg (2005), to name a few) rather than inferring supposedly emotional or affective states.
 Trying to interpret facial displays as a means of determining underlying emotional state, in all likelihood, remains vain.
@@
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-baltruvsaitis2016openface}{}%
+Baltrušaitis, T., Robinson, P., \& Morency, L.-P.
+(2016).
+Openface: An open source facial behavior analysis toolkit.
+In \emph{2016 ieee winter conference on applications of computer vision (wacv)} (pp.
+1--10).
+IEEE.
+
+\leavevmode\hypertarget{ref-barrett2006emotions}{}%
+Barrett, L. F. (2006). Are emotions natural kinds? \emph{Perspectives on Psychological Science}, \emph{1}(1), 28--58.
+
 \leavevmode\hypertarget{ref-barrett2019emotional}{}%
@@
 
+\leavevmode\hypertarget{ref-bonanno2004brief}{}%
+Bonanno, G., \& Keltner, D.
+(2004).
+The coherence of emotion systems: Comparing ``on-line'' measures of appraisal and facial expressions, and self-report.
+\emph{Cognition and Emotion}, \emph{18}(3), 431--444.
+\url{https://doi.org/10.1080/02699930341000149}
+
 \leavevmode\hypertarget{ref-buck1985prime}{}%
@@
 
+\leavevmode\hypertarget{ref-crivelli2018facial}{}%
+Crivelli, C., \& Fridlund, A. J. (2018). Facial displays are tools for social influence. \emph{Trends in Cognitive Sciences}, \emph{22}(5), 388--399. \url{https://doi.org/10.1016/j.tics.2018.02.006}
+
+\leavevmode\hypertarget{ref-crivelli2017recognizing}{}%
+Crivelli, C., Russell, J.
+A., Jarillo, S., \& Fernández-Dols, J.-M.
+(2017).
+Recognizing spontaneous facial expressions of emotion in a small-scale society of papua new guinea.
+\emph{Emotion}, \emph{17}(2), 337.
+
 \leavevmode\hypertarget{ref-darwin1872expression}{}%
@@
 
+\leavevmode\hypertarget{ref-dente2017measures}{}%
+Dente, P., Küster, D., Skora, L., \& Krumhuber, E.
+(2017).
+Measures and metrics for automatic emotion classification via facet.
+In \emph{Proceedings of the conference on the study of artificial intelligence and simulation of behaviour} (pp.
+160--163).
+
+\leavevmode\hypertarget{ref-den2005facereader}{}%
+Den Uyl, M., \& Van Kuilenburg, H. (2005). The facereader: Online facial expression recognition. In \emph{Proceedings of measuring behavior} (Vol. 30, pp. 589--590). Citeseer.
+
 \leavevmode\hypertarget{ref-digirolamo2017emotion}{}%
 DiGirolamo, M. A., \& Russell, J. A. (2017). The emotion seen in a face can be a methodological artifact: The process of elimination hypothesis. \emph{Emotion}, \emph{17}(3), 538.
+
+\leavevmode\hypertarget{ref-dols2017natural}{}%
+Dols, J.
+M.
+F.
+(2017).
+Natural facial expression: A view from psychological constructionism and pragmatics.
+In J.
+M.
+F.
+Dols \& J.
+A.
+Russell (Eds.), \emph{The science of facial expression} (pp.
+457--478).
+New York, NY: Oxford University Press.
+
+\leavevmode\hypertarget{ref-duchenne1876mecanisme}{}%
+Duchenne, G.-B. (1876). \emph{Mécanisme de la physionomie humaine: Où, analyse électro-physiologique de l'expression des passions}. J.-B. Baillière.
+
+\leavevmode\hypertarget{ref-dupre2015oudjat}{}%
+Dupré, D., Akpan, D., Elias, E., Adam, J.-M., Meillon, B., Bonnefond, N., \ldots{} Tcherkassof, A.
+(2015).
+Oudjat: A configurable and usable annotation tool for the study of facial expressions of emotion.
+\emph{International Journal of Human-Computer Studies}, \emph{83}, 51--61.
+\url{https://doi.org/10.1016/j.ijhcs.2015.05.010}
 
@@
 Fernández-Dols, J.-M., \& Ruiz-Belda, M.-A. (1995). Are smiles a sign of happiness? Gold medal winners at the olympic games. \emph{Journal of Personality and Social Psychology}, \emph{69}(6), 1113.
+
+\leavevmode\hypertarget{ref-fridlund1994human}{}%
+Fridlund, A. J. (1994). \emph{Human facial expression: An evolutionary view}. San Diego, CA: Academic Press.
+
+\leavevmode\hypertarget{ref-frijda2012recognition}{}%
+Frijda, N.
+(2012).
+Recognition of emotion in others.
+In F.
+Paglieri (Ed.), \emph{Consciousness in interaction.
+The role of the natural and social context in shaping consciousness} (pp.
+237--256).
+Amsterdam, NL: J.Benjamins.
+
+\leavevmode\hypertarget{ref-frijda1997facial}{}%
+Frijda, N.
+H., \& Tcherkassof, A.
+(1997).
+Facial expressions as modes of action readiness.
+In J.
+A.
+Russell \& J.-M.
+Fernández-Dols (Eds.), \emph{The psychology of facial expression} (pp.
+78--102).
+New York, NY: Cambridge University Press.
+
+\leavevmode\hypertarget{ref-gendron2018emotion}{}%
+Gendron, M., Hoemann, K., Crittenden, A. N., Msafiri, S., Ruark, G. A., \& Barrett, L. F. (2018). Emotion perception in hadza hunter gatherers.
 
@@
 
+\leavevmode\hypertarget{ref-hess2016emotional}{}%
+Hess, U., \& Fischer, A. H. (2016). \emph{Emotional mimicry in social context}. Cambridge University Press.
+
 \leavevmode\hypertarget{ref-hess1990differentiating}{}%
 Hess, U., \& Kleck, R. E. (1990). Differentiating emotion elicited and deliberate emotional facial expressions. \emph{European Journal of Social Psychology}, \emph{20}(5), 369--385.
+
+\leavevmode\hypertarget{ref-jakobs1999social}{}%
+Jakobs, E., Manstead, A.
+S., \& Fischer, A.
+H.
+(1999).
+Social motives and emotional feelings as determinants of facial displays: The case of smiling.
+\emph{Personality and Social Psychology Bulletin}, \emph{25}(4), 424--435.
 
@@
 
+\leavevmode\hypertarget{ref-kappas2003facial}{}%
+Kappas, A. (2003). What facial activity can and cannot tell us about emotions. In M. Katsikitis (Ed.), \emph{The human face} (pp. 215--234). Boston, MA: Springer.
+
+\leavevmode\hypertarget{ref-krause1998dyadic}{}%
+Krause, R., Steimer-Krause, E., Merten, J., \& Ullrich, B. (1998). Dyadic interaction regulation, emotion, and psychopathology. \emph{Emotions and Psychopathology: Theory and Research}, 70--80.
+
 \leavevmode\hypertarget{ref-kraut1979social}{}%
 Kraut, R. E., \& Johnston, R. E. (1979). Social and emotional messages of smiling: An ethological approach. \emph{Journal of Personality and Social Psychology}, \emph{37}(9), 1539--1553.
+
+\leavevmode\hypertarget{ref-krumhuber2013effects}{}%
+Krumhuber, E. G., Kappas, A., \& Manstead, A. S. (2013). Effects of dynamic aspects of facial expressions: A review. \emph{Emotion Review}, \emph{5}(1), 41--46.
 
@@
 Lindquist, K. A., Barrett, L. F., Bliss-Moreau, E., \& Russell, J. A. (2006). Language and the perception of emotion. \emph{Emotion}, \emph{6}(1), 125.
+
+\leavevmode\hypertarget{ref-lindquist2013s}{}%
+Lindquist, K. A., \& Gendron, M. (2013). What's in a word? Language constructs emotion perception. \emph{Emotion Review}, \emph{5}(1), 66--71. \url{https://doi.org/10.1177/1754073912451351}
+
+\leavevmode\hypertarget{ref-mcduff2016discovering}{}%
+McDuff, D.
+(2016).
+Discovering facial expressions for states of amused, persuaded, informed, sentimental and inspired.
+In \emph{Proceedings of the international conference on multimodal interaction} (pp.
+71--75).
+\url{https://doi.org/10.1145/2993148.2993192}
+
+\leavevmode\hypertarget{ref-mcduff2016affdex}{}%
+McDuff, D., Mahmoud, A., Mavadati, M., Amr, M., Turcot, J., \& Kaliouby, R.
+el.
+(2016).
+AFFDEX sdk: A cross-platform real-time multi-face expression recognition toolkit.
+In \emph{Proceedings of the chi conference on human factors in computing systems} (pp.
+3723--3726).
+\url{https://doi.org/10.1145/2851581.2890247}
+
+\leavevmode\hypertarget{ref-mckeown2013analogical}{}%
+McKeown, G.
+J.
+(2013).
+The analogical peacock hypothesis: The sexual selection of mind-reading and relational cognition in human communication.
+\emph{Review of General Psychology}, \emph{17}(3), 267--287.
+\url{https://doi.org/10.1037/a0032631}
 
@@
 
+\leavevmode\hypertarget{ref-nielsen2007conceptual}{}%
+Nielsen, L., \& Kaszniak, A. W. (2007). Conceptual, theoretical, and methodological issues in inferring subjective emotion experience. \emph{Handbook of Emotion Elicitation and Assessment}, 361--375.
+
+\leavevmode\hypertarget{ref-pacherie2003modes}{}%
+Pacherie, E.
+(2003).
+Modes de structuration des contenus perceptifs visuels.
+\emph{J.
+Bouveresse \& J.-J.
+Rosat (Eds), Philosophies de La Perception.
+Phénoménologie, Grammaire et Sciences Cognitives, Paris: Odile Jacob}, 263--289.
+
 \leavevmode\hypertarget{ref-reisenzein2000exploring}{}%
@@
 
+\leavevmode\hypertarget{ref-sander2014oxford}{}%
+Sander, D., \& Scherer, K. (2014). \emph{Oxford companion to emotion and the affective sciences}. OUP Oxford.
+
 \leavevmode\hypertarget{ref-scherer1997lost}{}%
@@
 \emph{European Journal of Social Psychology}, \emph{37}(6), 1325--1345.
+
+\leavevmode\hypertarget{ref-tcherkassof2013dynemo}{}%
+Tcherkassof, A., Dupré, D., Meillon, B., Mandran, N., Dubois, M., \& Adam, J.-M.
+(2013).
+DynEmo: A video database of natural facial expressions of emotions.
+\emph{The International Journal of Multimedia \& Its Applications}, \emph{5}(5), 61--80.
+\url{https://doi.org/10.5121/ijma.2013.5505}
+
+\leavevmode\hypertarget{ref-tcherkassof2014emotions}{}%
+Tcherkassof, A., \& Frijda, N. H. (2014). Les émotions: Une conception relationnelle. \emph{LAnnee Psychologique}, \emph{114}(3), 501--535.
 
@@
 
+\leavevmode\hypertarget{ref-baltruvsaitis2016openface}{}%
+Baltrušaitis, T., Robinson, P., \& Morency, L.-P.
+(2016).
+Openface: An open source facial behavior analysis toolkit.
+In \emph{2016 ieee winter conference on applications of computer vision (wacv)} (pp.
+1--10).
+IEEE.
+
+\leavevmode\hypertarget{ref-barrett2006emotions}{}%
+Barrett, L. F. (2006). Are emotions natural kinds? \emph{Perspectives on Psychological Science}, \emph{1}(1), 28--58.
+
 \leavevmode\hypertarget{ref-barrett2019emotional}{}%

### 2019-11-04 manuscript_apa.Rmd
manuscript doc

@@
 
-keywords          : "Facial expression, self-report, human observer, automatic recognition."
-wordcount         : "7998"
+keywords          : "emotion, facial expression, self-report, human observer, automatic recognition."
+wordcount         : "7924"
 
@@
 classoption       : "man"
-output            : 
-  bookdown::pdf_book:
-    base_format: papaja::apa6_pdf
+output: papaja::apa6_word
+# output            : 
+#   bookdown::pdf_book:
+#     base_format: papaja::apa6_pdf
 
@@
 This technology rests on the premise that facial expressions provide a direct access to individuals’ subjective feelings, and that one can read the emotions displayed on a person’s face as in an open book.
-This premise is derived from the Basic Emotion View which Paul Ekman is one of the famous proponents of this perspective [@ekman1987universals; @ekman1988universality; @ekman1992argument; @ekman2007directed].
+This premise is derived from the Basic Emotion View which suggests a one-to-one mapping between subective feeling and facial expression [@ekman1987universals; @ekman1988universality; @ekman1992argument; @ekman2007directed].
 Even if this premise is central to the modern mainstream approach of human emotion, recent research in affective science is challenging it.
@@
 The belief that facial expression is linked to emotional states can be traced back to Darwin in *The Expression of the Emotions in Man and Animals* [@darwin1872expression].
-Countless studies have fortified the idea that emotional states are inherently coupled to a set of prototypic facial expressions (*e.g.*, [@ekman2011meant]).
+Countless studies have fortified the idea that emotional states are inherently coupled to a set of prototypic facial expressions [*e.g.*, @ekman2011meant].
 The Basic Emotion View holds that facial expressions are genuine displays of an individual’s inner emotional state.
@@
 
-Relatively few studies have actually tested Basic Emotion View’s fundamental claim regarding the facial production (so-called ‘expression’) of emotions.
+Relatively few studies have actually tested Basic Emotion View’s fundamental claim regarding the facial production (so-called "expression") of emotions.
 @rosenberg1994coherence reported the first evidence of coherence between self-report of emotion and displayed facial expressions.
@@
 Analysis of participants’ facial expressions and reports of emotions showed that there was a high degree of temporal linkage and emotional agreement between facial expressions and self-reports.
-Notwithstanding this study, few have used purely descriptive methods (such as electromyography or objective coding systems – FACS or others) to identify and to measure the actual changes in the face when a given emotion is felt (see @wagner1997methods, for methodological points).
+Notwithstanding this study, few have used purely descriptive methods such as electromyography or objective face coding systems to identify and to measure the actual changes in the face when a given emotion is felt [see @wagner1997methods, for methodological points].
 Much of the research has focused instead on the recognition of EFE, that is on the issue of the viewer’s interpretation of facial displays.
@@
 The Readout Hypothesis [@buck1985prime] formalizes this folk theory.
-According to it, facial expressions have evolved to provide “an external readout of those motivational-emotional processes that have had social implications during the course of evolution” (pp.
-396-397).
+According to it, facial expressions have evolved to provide *“an external readout of those motivational-emotional processes that have had social implications during the course of evolution”* [@buck1985prime, 396-397].
 Thus, as readouts of emotional states, spontaneous expressive displays are directly accessible to other organisms.
-Research on the inference of emotions from facial expressions has established that viewers show considerable agreement that the so-called basic emotions—happiness, anger, fear, disgust, sadness, and surprise—are associated with specific facial displays (*e.g.*, reviews from @elfenbein2002universality; @russell1994there).
-These findings are considered by many as constituting strong evidence for the Basic Emotion View.
+Research on the inference of emotions from facial expressions has established that viewers show considerable agreement that the so-called basic emotions—happiness, anger, fear, disgust, sadness, and surprise—are associated with specific facial displays [*e.g.*, reviews from @elfenbein2002universality; @russell1994there].
 
@@
 Notably, it fails to explain, in instances in which display rules cannot be called upon, how individuals can feel emotions without expressing them or how individuals can express emotions without feeling them, and why observers are not totally accurate in recognizing facial expressions of basic emotions, among others.
-The following  is a summary of key findings and conclusions both from field observations and laboratory experiments on spontaneous expression of emotions, all of them strongly subverting the Basic Emotion viewpoint (for reviews, @fernandez2013emotion; @reisenzein2013coherence; @russell2003facial).
+The following  is a summary of key findings and conclusions both from field observations and laboratory experiments on spontaneous expression of emotions, all of them strongly subverting the Basic Emotion viewpoint [for reviews, @fernandez2013emotion; @reisenzein2013coherence; @russell2003facial].
 Mains issues regarding recognition studies are also swept across afterwards.
@@
 
-Naturalistic studies look at the ecological frequency of co-occurrence of certain emotions and facial displays [fernandez2013emotion].
+Naturalistic studies look at the ecological frequency of co-occurrence of certain emotions and facial displays [@fernandez2013emotion].
 Though they cannot be considered as a straightforward test of the triggering role of emotion on the facial behavior, they have the advantage of considering situations that cannot be created in a laboratory.
@@
 More generally, examination of bowlers’ facial display showed that they rarely smiled while facing the pins but often smiled when facing their friends.
-These findings were confirmed by @ruiz2003spontaneous who analyzed the facial displays of bowlers (after a strike) and soccer fans (after their team scored) reporting happiness in a field study.
-They found that the probability of smiling was very low when participants were not interacting with someone else.
+These findings were confirmed by @ruiz2003spontaneous who analyzed the facial displays of bowlers after a strike and soccer fans after their team scored.
+Their results show a low probability of smiling when participants were not interacting with someone else.
 These findings have been supported in other realistic field settings such as the one of @fernandez1995smiles suggesting that happiness is not a sufficient cause of smiling.
-They watched extremely happy gold medalist athletes displaying facial expressions of sadness (sometimes associated with tears) during their Olympics awards ceremonies.
+@fernandez1995smiles watched extremely happy gold medalist athletes displaying facial expressions of sadness (sometimes associated with tears) during their Olympics awards ceremonies.
 More specifically, winners showed Duchenne smiles and other types of smiles when they were interacting with other people during the awards ceremony, but smiles were scarce or nonexistent when waiting behind the podium and/or when turning toward the flagpoles and focusing their attention on the flags and the national anthem.
@@
 
-Happiness (or joy) is not the only emotion weakly associated with the predicted facial expression.
+Happiness/joy is not the only emotion weakly associated with the predicted facial expression.
 A naturalistic study conducted by @scherer1997lost in an airport’s baggage handling office showed that passengers claiming for their lost luggage displayed very few facial expressions of negative emotions while self-reporting subjective feeling states of anger or sadness among others.
@@
 It is often acknowledged that adults regulate their expressive behavior.
-In accordance with various display rules (personal, social, cultural ones), they exert a control over the (supposed) automatic readouts of their emotions.
+In accordance with various display rules (personal, social, cultural ones), they exert a control over the supposed automatic readouts of their emotions.
 Developmental studies provide relevant settings in which display rules are inoperative.
@@
 Infants displayed a variety of facial expressions in each eliciting situation.
-Yet, more infants exhibited positive (joy and surprise) than negative facial expressions (anger, disgust, fear, and sadness) across all situations –except sour taste.
+Yet, more infants exhibited positive than negative facial expressions across all situations –except sour taste.
 No evidence for emotion-specific facial expressions corresponding to anger, fear, and sadness was obtained.
 Camras and her colleagues also observed that 11-month-old European-American, Chinese, and Japanese infants did not display distinct negative emotion-specific patterns of facial muscles in response to two elicitors meant to induce fear and anger [@camras2007infants].
-Thus, these findings in naturalistic settings provide little support for the Basic Emotion View.
+Thus, these findings in naturalistic settings provide little support for the one-to-one mapping of subjective feeling and facial expression.
 
@@
 As stressed by @reisenzein2013coherence, experimental studies permit both better control of various factors (*e.g.*, emotion elicitors) and tests about likely moderators of the emotion–facial expression link (*e.g.*, the social context).
-As an example, a strong disconfirmation of the Basic Emotion View’s premise is put forward by Reisenzein’s studies on surprise (*e.g.*, @reisenzein2000exploring).
-In eight controlled laboratory situations, surprise was induced by @reisenzein2006evidence by establishing and then invalidating a set of beliefs concerning the experimental events (*e.g*., the unexpected appearance of a picture of one's own face as the last picture in a series of portraits that had to be rated).
-Visible or EMG-detected facial displays of surprise occurred only in few participants.
+As an example, a strong disconfirmation of the Basic Emotion View’s premise is put forward by Reisenzein’s studies on surprise [*e.g.*, @reisenzein2000exploring].
+In eight controlled laboratory situations, surprise was induced by @reisenzein2006evidence by establishing and then invalidating a set of beliefs concerning the experimental events such as the unexpected appearance of a picture of one's own face as the last picture in a series of portraits that had to be rated.
+Visible or electromyography-detected facial displays of surprise occurred only in few participants.
 Yet, most participants reported subjective feelings of surprise and most believed that they had shown a strong surprise facial expression.
 @schutzwohl2012facial also observed similarly low frequencies of surprise facial expressions when their participants, after leaving the laboratory, unexpectedly found themselves not in the corridor but in a new room with green walls and a red office chair.
-Less than a quarter of the participants displayed an expression of surprise: only 5% showed widened eyes, raised eyebrows, and opened mouth (*i.e.* the complete expression of surprise according to the Basic Emotion View) and 17%  showed widened eyes and raised eyebrows.
+Less than a quarter of the participants displayed an expression of surprise: only 5% showed widened eyes, raised eyebrows, and opened mouth which correspond to the complete expression of surprise according to the Basic Emotion View and 17%  showed widened eyes and raised eyebrows.
 Again, participants overestimated their surprise expressivity.
 
-Studies on happiness and related positive emotions (*e.g.*, sensory pleasantness) also show a low coherence between emotion and facial display [@duran2017coherence; @reisenzein2013coherence].
+Studies on happiness and related positive emotions such as sensory pleasantness also show a low coherence between emotion and facial display [@duran2017coherence; @reisenzein2013coherence].
 In fact, only experiments on amusement provide a fairly strong association between emotion and smiling.
-When confronted to humorous events (*e.g.*, jokes, being tickled), the number of facially reactive participants (smiling and laughing) is indeed quite fair, whereas when confronted to other positive emotions (*e.g.*, happiness), few participants show the expected facial expressions [for instance, the Duchenne smile or any kind of “happy” expression yet reporting being happy; see @mehu2007smiles; @lee2002effect, social condition].
+When confronted to humorous events (*e.g.*, jokes, being tickled), the number of participants smiling and laughing is indeed quite fair, whereas when confronted to other positive emotions (*e.g.*, happiness), few participants show the expected facial expressions [for instance, the Duchenne smile or any kind of expression related to happiness; see @mehu2007smiles; @lee2002effect].
 More largely, the meta-analysis conducted by @duran2017coherence on the degree of statistical covariation between emotions and facial expressions provides conclusive evidence.
-First of all, the commonly called basic emotions (happiness/amusement, sadness, fear, anger, disgust and surprise), on average, are weakly correlated with the specific configuration of facial muscles that the Basic Emotion View ascribes to them (the correlation even drops when happiness/amusement studies are excluded).
-Very few participants who relived an experience of sadness [*e.g.*, @tsai2002emotional] were reactive, only a small number of them showing components of a sad facial expression (*e.g.*, oblique eyebrows, lip corners pulled down).
-Regarding fear, spider phobic participants recruited by @vernon2002disgust, for instance, were exposed to a live tarantula.
-About one third of them only displayed some components of the prototypical fear expression (eye widening, brow raising and knitting, etc.).
+First of all, the basic emotions are weakly correlated with the specific configuration of facial muscles that the Basic Emotion View ascribes to them (the correlation drops when happiness/amusement studies are excluded).
+Very few participants who relived an experience of sadness were reactive, only a small number of them showing components of a sad facial expression with oblique eyebrows or lip corners pulled down [@tsai2002emotional].
+Regarding fear, about one third of spider phobic participants exposed to a live tarantula displayed some components of the prototypical fear expression such as eye widening, brow raising and knitting [@vernon2002disgust].
 Similar observation is made for anger.
-The proportion of facially reactive participants (*i.e.* displaying at least one component among frowning, lid/lip tightening...) reporting anger in @johnson2010smile and @tsai2002emotional studies does not exceed 35%.
+The proportion of facially reactive participants displaying at least one component among frowning or lid/lip tightening and reporting anger does not exceed 35% [@johnson2010smile; @tsai2002emotional].
 The same proportion of reactive participants confronted to disgusting elicitors is observed [*e.g.*, @ekman1980facial; @fernandez1997spontaneous].
@@
 Hence, laboratory studies show that facial expressions of emotion are often not displayed in situations in which the Basic Emotion View would predict them to be expressed.
-Moreover, when they are indeed displayed, they are more often displayed partially than completely.
+Moreover, when corresponding facial expressions are indeed displayed, they are only partially displayed.
 All in all, research on the spontaneous expression of emotions does not yield strong support for the Basic Emotion View.
@@
 The Basic Emotion View postulates that, when triggered, each basic emotion is expressed by a prototypical face (non-basic emotions being blends of the basic ones).
-In return, the recognition of emotional facial expressions (EFE) is claimed to be based on the identification of specific patterns of facial movements associated with each emotion, as if expression and recognition were the two sides of a same coin.
-Seemingly compelling evidence support this claim, sustaining the belief that one can read a face in the same way one reads the lines of a book.
+In return, the recognition of EFE is claimed to be based on the identification of specific patterns of facial movements associated with each emotion, as if expression and recognition were the two sides of a same coin.
+Seemingly compelling evidence support this claim, sustaining the possibility of a clear readout of subjective feelings from facial expressions.
 Besides, recognition systems rely on such a principle and are considered as an objective coding tool because they are based on the identification of specific muscular changes in the face.
-Many people highly agree that so-called basic emotions are associated with specific facial configurations (Ekman, 2017; Elfenbein & Ambady, 2002; Russell, 1994) and also agree that many believe that it is a strong evidence for the Basic Emotion View.
+Many people highly agree that so-called basic emotions are associated with specific facial configurations [@ekman2017facial; @elfenbein2002universality] and also agree that many believe that it is a strong evidence for the Basic Emotion View.
 Moreover, it implies that EFE recognition of both human observers and automatic classifiers should be as accurate.
@@
 Among others, evidence has been questioned on methodological grounds [*e.g.*, @russell1994there].
-The response format usually used in recognition studies (*i.e*.
-forced choice: selection of one word from a pre-specified list of emotion labels), notably, leads to a biased consensus as evidenced by @russell1993forced.
-Depending on the list of emotion labels at participants’ disposal, EFE of sadness can easily be categorized as sad expressions as well as fear expressions...
+The response format usually used in recognition studies (*i.e*, forced choice: selection of one word from a pre-specified list of emotion labels), notably, leads to a biased consensus[@russell1993forced].
+Depending on the list of emotion labels at participants’ disposal, EFE of sadness can easily be categorized as sad expressions as well as fear expressions for example.
 @russell1987relativity also showed the same facial expression can be seen as expressing different types of emotions, depending on what other faces are seen.
@@
 418).
-Instead, emotions are probably mentally constructed by the perceiver and mental categories of emotions are needed to accurately categorize facial movements, among other things (*e.g.*, contextual information).
+Instead, emotions are probably mentally constructed by the perceiver and mental categories of emotions are needed to accurately categorize facial movements among contextual information.
 
@@
 It may be that differences in the social information displayed by static and dynamic expressions leads to facial recognition differential effects.
-Regarding the issue of spontaneous vs posed expressions (the latter are overused in experiments).
+Regarding the issue of spontaneous *vs.* posed expressions (the latter are overused in experiments).
 As @meillon2010dynemo conclude, EFE have been typically studied as static displays.
@@
 
-All in all, as many doubts can be raised about the standard method, experiments conducted with such a method cannot be considered as providing solid empirical support to the Basic Emotion View.
+Finally, as many doubts can be raised about the standard method, experiments conducted with such a method cannot be considered as providing solid empirical support to the Basic Emotion View.
 Based on the numerous methodological criticisms, but also theoretical, addressed to this view, alternative conceptions have emerged.
@@
 Starting from the empirical evidence suggesting that spontaneous facial expressions in ordinary life are equivocal, @dols2017natural argues in favor of a pragmatic conception of natural facial displays.
-He makes a plea for the idea that natural facial displays, rather than "saying" – because they do not have a specific meaning he claims – “make” things.
+He makes a plea for the idea that natural facial displays, rather than "saying"--because facial expressions do not have a specific meaning--“make” things.
 Facial expressions are actions in a communicative interaction.
-They do not express emotions but they *"prompt, on the receiver’s side, important inferences about the context, the sender, and the course of the interaction between sender and receiver"* [@dols2017natural, 466, underlined by the author].
+They do not express emotions but they *"prompt, on the receiver’s side, important inferences about the context, the sender, and the course of the interaction between sender and receiver"* [@dols2017natural, 466].
 As such, the fact that facial displays are able to signal emotions is a byproduct of one of its main function, the one of implementing actions performing practical ends.
@@
 They are not primarily communicative signals and even less the outlet of an internal state.
-Facial displays are (parts of) pragmatic actions aiming at orienting the person’s relation to its environment, for instance maintaining, breaking,  restoring, etc., the relation [@frijda2012recognition].
-Facial displays are not “recognized” in semantic terms but are perceived as intentional actions.
-In the face of the continual flow of uninterrupted facial movements, the perceiver sees behaviors directed towards a goal.
-S.he translates the continuous flow of movements into coordinated sequences of actions holding a beginning and an end.
+Facial displays are parts of pragmatic actions aiming at orienting the person’s relation to its environment.
+For instance facial displays are maintaining, breaking or restoring the relationship between sender and perceiver [@frijda2012recognition].
+Facial displays are not recognized in semantic terms but are perceived as intentional actions.
+In the face of the continual flow of uninterrupted facial movements, perceivers see behaviors directed towards a goal.
+They translate the continuous flow of movements into coordinated sequences of actions holding a beginning and an end.
 Facial displays are not simple strings of action units which morphological configuration would be the prototype of a given emotion, and consequently identified as such [@ekman1987universals].
@@
 This is why even unauthentic facial displays can yet be recognized as emotional expressions.
-Guillaume Duchenne de Boulogne explains that the artist who has shaped the famous Laocoon antic sculpture, exhibited in the Vatican’s museum, has made a modelling mistake since no face can display its emotional expression [@duchenne1876mecanisme].
+Duchenne de Boulogne explains that the artist who has shaped the famous Laocoon antic sculpture, exhibited in the Vatican’s museum, has made a modelling mistake since no face can display its emotional expression [@duchenne1876mecanisme].
 Indeed, no muscular contraction can produce it.
@@
 His demonstration gives food for thought.
-Even though no objective coding system can correctly code the discordant facial features of the Laocoon’s face, yet anyone can easily recognize the suffering and despair he admirably expresses.
+Even though no objective coding system can correctly code the discordant facial features of the Laocoon’s face, yet anyone can easily recognize the suffering and despair he expresses.
 This example goes along the lines of the constructivist approach.
 
-The constructivist approach claims that facial displays are behaviors which meaning is inferred by the perceiver.
+The constructivist approach claims that facial displays are behaviors which meaning is inferred by perceivers.
 Findings support this observer dependence [@lindquist2013s; niedenthal2017embodied].
@@
 participants and Himba participants from remote regions of Namibia [@gendron2014perceptions] and Hazda participants of the Eastern Rift Valley of Tanzania [@gendron2018emotion] both groups with limited exposure to Western culture.
-Gendron and her colleagues demonstrated that facial expressions were not universally “recognized” in discrete emotional terms.
-Indeed, when Himba and Hazda participants did not have emotion concepts at their disposal to structure perception, they perceived the facial expressions as behaviors (*e.g.*, looking, smelling) that didn’t have a necessary relationship to emotions.
-They didn’t infer inner states (such as emotional feelings) but they rather proceeded with action identification that pointed out the functions of behaviors [see also @crivelli2017recognizing for similar observations among a small-scale society of Papua New Guinea].
+Gendron and her colleagues demonstrated that facial expressions were not universally recognized in discrete emotional terms.
+Indeed, when Himba and Hazda participants did not have emotion concepts at their disposal to structure perception, they perceived the facial expressions as behaviors, such as looking or smelling, that didn’t have a necessary relationship to emotions.
+They did not infer inner states but they rather proceeded with action identification that pointed out the functions of behaviors [see also @crivelli2017recognizing for similar observations among a small-scale society of Papua New Guinea].
 The constructivist approach considers that specific emotion categories, as conceptualized by Western cultures’ knowledge, are casted on the perceived face to make meaning of the sender’s facial displays.
@@
 First, the participants had to identify the emotional sequences by pressing the space bar of their keyboard to indicate the beginning and the end of the emotional sequences while watching the video.
-Second, the participants watched each emotional sequence previously identified and labeled the sequence using one of the 12 emotions proposed including six "basic" emotion labels (*i.e.*, *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) and six "non-basic" emotion labels (*i.e.*, *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
+Second, the participants watched each emotional sequence previously identified and labeled the sequence using one of the 12 emotions proposed including six basic emotion labels (*i.e.*, *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) and six "non-basic" emotion labels (*i.e.*, *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
 They also had the possibility to indicate that the sequence was expressing none of the proposed emotion labels.
@@
 Affdex’s algorithm uses Histogram of Oriented Gradient (HOG) features and Support Vector Machine (SVM) classifiers in order to recognize facial expressions.
-For each video frame, Affdex identifies the probability $p_{video_{i}.label_{j}.t_{k}}$ from 0 to 100 (rescaled to 0 to 1 for the analysis) of the face as expressing one of the six "basic" emotion labels (*i.e.*, *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as additional psychological states such as *valence*, *engagement* or *contempt*, and facial features such as *cheek raise*, *eye widen* or *jaw drop*.
+For each video frame, Affdex identifies the probability $p_{video_{i}.label_{j}.t_{k}}$ from 0 to 100 (rescaled to 0 to 1 for the analysis) of the face as expressing one of the six basic emotion labels (*i.e.*, *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as additional psychological states such as *valence*, *engagement* or *contempt*, and facial features such as *cheek raise*, *eye widen* or *jaw drop*.
 
-For both human and automatic recognition, to determine which of the six “basic” emotions can be used to identify each video, the recognition probability for each label by frame was converted into odd ratio by label [@dente2017measures].
+For both human and automatic recognition, to determine which of the six basic emotions can be used to identify each video, the recognition probability for each label by frame was converted into odd ratio by label [@dente2017measures].
 The highest sum of each odd ratio time-series defines the label recognized by the automatic classifier (EQ\ref{eq:2} for human recognition and EQ\ref{eq:3} for automatic recognition).
@@
 
-Since encoders’ self-reports, human annotations and the automatic recognition include data on “non-basic” emotion labels and features, the analysis is performed using only the six "basic" emotion labels in order to compare them.
+Since encoders’ self-reports, human annotations and the automatic recognition include data on “non-basic” emotion labels and features, the analysis is performed using only the six basic emotion labels in order to compare them.
 The maximum score for self-reports, human annotations and automatic recognition is used to label the video.
@@
 Therefore, facial movements do not express specific emotions because they do not carry intrinsic emotional signification.
-It is the observer that infers the emotional meaning of the facial expression, this interpretation depending on availability of different kinds of information (contextual information, linguistic categories...).
+It is the observer that infers the emotional meaning of the facial expression, this interpretation depending on availability of different information such as context or linguistic categories.
 As a consequence, one can predict from the first line of thinking that individuals’ emotional subjective feeling should be correlated to the recognition of facial expressions from both human observers and automatic classifiers whereas if emotions are social constructs, as stated by the second line of thinking, human observer’s should be better at perceiving emotions expressed on the face than automatic classifiers.
@@
 However, the human decoders in the present study were presented with faces without any contextual information which could have helped them to shape more precisely their interpretation.
-Hassin, Aviezer and Bentin (2013) reviewed evidence to demonstrate that faces are inherently ambiguous and that observers rely on situational cues when they process facial displays (see also Aviezer, Hassin, Ryan, Grady, Susskind, Anderson, Moscovitch & Bentin, 2008; Barrett, Mesquita, Gendron, 2011).
+@hassin2013inherently reviewed evidence to demonstrate that faces are inherently ambiguous and that observers rely on situational cues when they process facial displays [see also @aviezer2008angry; @barrett2011context].
 It happens that contextual information even shifts emotion perception.
-In the present case, the facial stimuli displayed to the decoders were as equivocal as are real-life facial expressions (Aviezer, Ensenberg, & Hassin, 2017).
+In the present case, the facial stimuli displayed to the decoders were as equivocal as are real-life facial expressions [@aviezer2017inherently].
 Yet, decoders had no cue at disposal.
-So, without any possibility of integrating faces and context, their decoding accuracy is pretty much degraded as it is usual in such cases (e.g., Wagner, MacDonald, & Manstead, 1986).
+So, without any possibility of integrating faces and context, their decoding accuracy is pretty much degraded as it is usual in such cases [*e.g.*, @wagner1986communication].
 However, they demonstrate their superiority on the classifier tool probably thanks to their capacity to rely on their previous personal experience to invent a context in which a face could display such an expression.

### 2019-11-04 manuscript_apa.tex
manuscript doc

@@
 This technology rests on the premise that facial expressions provide a direct access to individuals' subjective feelings, and that one can read the emotions displayed on a person's face as in an open book.
-This premise is derived from the Basic Emotion View which Paul Ekman is one of the famous proponents of this perspective (Ekman, 1992, 2007; Ekman et al., 1987; Ekman \& Heider, 1988).
+This premise is derived from the Basic Emotion View which suggests a one-to-one mapping between subective feeling and facial expression (Ekman, 1992, 2007; Ekman et al., 1987; Ekman \& Heider, 1988).
 Even if this premise is central to the modern mainstream approach of human emotion, recent research in affective science is challenging it.
@@
 The belief that facial expression is linked to emotional states can be traced back to Darwin in \emph{The Expression of the Emotions in Man and Animals} (Darwin, 1872).
-Countless studies have fortified the idea that emotional states are inherently coupled to a set of prototypic facial expressions (\emph{e.g.}, (Ekman \& Cordaro, 2011)).
+Countless studies have fortified the idea that emotional states are inherently coupled to a set of prototypic facial expressions (\emph{e.g.}, Ekman \& Cordaro, 2011).
 The Basic Emotion View holds that facial expressions are genuine displays of an individual's inner emotional state.
@@
 Analysis of participants' facial expressions and reports of emotions showed that there was a high degree of temporal linkage and emotional agreement between facial expressions and self-reports.
-Notwithstanding this study, few have used purely descriptive methods (such as electromyography or objective coding systems -- FACS or others) to identify and to measure the actual changes in the face when a given emotion is felt (see Wagner (1997), for methodological points).
+Notwithstanding this study, few have used purely descriptive methods such as electromyography or objective face coding systems to identify and to measure the actual changes in the face when a given emotion is felt (see Wagner, 1997, for methodological points).
 Much of the research has focused instead on the recognition of EFE, that is on the issue of the viewer's interpretation of facial displays.
@@
 The Readout Hypothesis (Buck, 1985) formalizes this folk theory.
-According to it, facial expressions have evolved to provide \enquote{an external readout of those motivational-emotional processes that have had social implications during the course of evolution} (pp.~396-397).
+According to it, facial expressions have evolved to provide \emph{\enquote{an external readout of those motivational-emotional processes that have had social implications during the course of evolution}} (Buck, 1985, pp.
+396--397).
 Thus, as readouts of emotional states, spontaneous expressive displays are directly accessible to other organisms.
-Research on the inference of emotions from facial expressions has established that viewers show considerable agreement that the so-called basic emotions---happiness, anger, fear, disgust, sadness, and surprise---are associated with specific facial displays (\emph{e.g.}, reviews from Elfenbein \& Ambady (2002); Russell (1994)).
-These findings are considered by many as constituting strong evidence for the Basic Emotion View.
+Research on the inference of emotions from facial expressions has established that viewers show considerable agreement that the so-called basic emotions---happiness, anger, fear, disgust, sadness, and surprise---are associated with specific facial displays (\emph{e.g.}, reviews from Elfenbein \& Ambady, 2002; Russell, 1994).
 
@@
 Notably, it fails to explain, in instances in which display rules cannot be called upon, how individuals can feel emotions without expressing them or how individuals can express emotions without feeling them, and why observers are not totally accurate in recognizing facial expressions of basic emotions, among others.
-The following is a summary of key findings and conclusions both from field observations and laboratory experiments on spontaneous expression of emotions, all of them strongly subverting the Basic Emotion viewpoint (for reviews, Fernández-Dols \& Crivelli (2013); Reisenzein, Studtmann, \& Horstmann (2013); Russell, Bachorowski, \& Fernández-Dols (2003)).
+The following is a summary of key findings and conclusions both from field observations and laboratory experiments on spontaneous expression of emotions, all of them strongly subverting the Basic Emotion viewpoint (for reviews, Fernández-Dols \& Crivelli, 2013; Reisenzein, Studtmann, \& Horstmann, 2013; Russell, Bachorowski, \& Fernández-Dols, 2003).
 Mains issues regarding recognition studies are also swept across afterwards.
@@
 
-Naturalistic studies look at the ecological frequency of co-occurrence of certain emotions and facial displays {[}fernandez2013emotion{]}.
+Naturalistic studies look at the ecological frequency of co-occurrence of certain emotions and facial displays (Fernández-Dols \& Crivelli, 2013).
 Though they cannot be considered as a straightforward test of the triggering role of emotion on the facial behavior, they have the advantage of considering situations that cannot be created in a laboratory.
@@
 More generally, examination of bowlers' facial display showed that they rarely smiled while facing the pins but often smiled when facing their friends.
-These findings were confirmed by Ruiz-Belda, Fernández-Dols, Carrera, \& Barchard (2003) who analyzed the facial displays of bowlers (after a strike) and soccer fans (after their team scored) reporting happiness in a field study.
-They found that the probability of smiling was very low when participants were not interacting with someone else.
+These findings were confirmed by Ruiz-Belda, Fernández-Dols, Carrera, \& Barchard (2003) who analyzed the facial displays of bowlers after a strike and soccer fans after their team scored.
+Their results show a low probability of smiling when participants were not interacting with someone else.
 These findings have been supported in other realistic field settings such as the one of Fernández-Dols \& Ruiz-Belda (1995) suggesting that happiness is not a sufficient cause of smiling.
-They watched extremely happy gold medalist athletes displaying facial expressions of sadness (sometimes associated with tears) during their Olympics awards ceremonies.
+Fernández-Dols \& Ruiz-Belda (1995) watched extremely happy gold medalist athletes displaying facial expressions of sadness (sometimes associated with tears) during their Olympics awards ceremonies.
 More specifically, winners showed Duchenne smiles and other types of smiles when they were interacting with other people during the awards ceremony, but smiles were scarce or nonexistent when waiting behind the podium and/or when turning toward the flagpoles and focusing their attention on the flags and the national anthem.
@@
 
-Happiness (or joy) is not the only emotion weakly associated with the predicted facial expression.
+Happiness/joy is not the only emotion weakly associated with the predicted facial expression.
 A naturalistic study conducted by Scherer \& Ceschi (1997) in an airport's baggage handling office showed that passengers claiming for their lost luggage displayed very few facial expressions of negative emotions while self-reporting subjective feeling states of anger or sadness among others.
@@
 Infants displayed a variety of facial expressions in each eliciting situation.
-Yet, more infants exhibited positive (joy and surprise) than negative facial expressions (anger, disgust, fear, and sadness) across all situations --except sour taste.
+Yet, more infants exhibited positive than negative facial expressions across all situations --except sour taste.
 No evidence for emotion-specific facial expressions corresponding to anger, fear, and sadness was obtained.
 Camras and her colleagues also observed that 11-month-old European-American, Chinese, and Japanese infants did not display distinct negative emotion-specific patterns of facial muscles in response to two elicitors meant to induce fear and anger (Camras et al., 2007).
-Thus, these findings in naturalistic settings provide little support for the Basic Emotion View.
+Thus, these findings in naturalistic settings provide little support for the one-to-one mapping of subjective feeling and facial expression.
 
@@
 (2013), experimental studies permit both better control of various factors (\emph{e.g.}, emotion elicitors) and tests about likely moderators of the emotion--facial expression link (\emph{e.g.}, the social context).
-As an example, a strong disconfirmation of the Basic Emotion View's premise is put forward by Reisenzein's studies on surprise (\emph{e.g.}, Reisenzein (2000)).
-In eight controlled laboratory situations, surprise was induced by Reisenzein, Bördgen, Holtbernd, \& Matz (2006) by establishing and then invalidating a set of beliefs concerning the experimental events (\emph{e.g}., the unexpected appearance of a picture of one's own face as the last picture in a series of portraits that had to be rated).
-Visible or EMG-detected facial displays of surprise occurred only in few participants.
+As an example, a strong disconfirmation of the Basic Emotion View's premise is put forward by Reisenzein's studies on surprise (\emph{e.g.}, Reisenzein, 2000).
+In eight controlled laboratory situations, surprise was induced by Reisenzein, Bördgen, Holtbernd, \& Matz (2006) by establishing and then invalidating a set of beliefs concerning the experimental events such as the unexpected appearance of a picture of one's own face as the last picture in a series of portraits that had to be rated.
+Visible or electromyography-detected facial displays of surprise occurred only in few participants.
 Yet, most participants reported subjective feelings of surprise and most believed that they had shown a strong surprise facial expression.
 Schützwohl \& Reisenzein (2012) also observed similarly low frequencies of surprise facial expressions when their participants, after leaving the laboratory, unexpectedly found themselves not in the corridor but in a new room with green walls and a red office chair.
-Less than a quarter of the participants displayed an expression of surprise: only 5\% showed widened eyes, raised eyebrows, and opened mouth (\emph{i.e.} the complete expression of surprise according to the Basic Emotion View) and 17\% showed widened eyes and raised eyebrows.
+Less than a quarter of the participants displayed an expression of surprise: only 5\% showed widened eyes, raised eyebrows, and opened mouth which correspond to the complete expression of surprise according to the Basic Emotion View and 17\% showed widened eyes and raised eyebrows.
 Again, participants overestimated their surprise expressivity.
 
-Studies on happiness and related positive emotions (\emph{e.g.}, sensory pleasantness) also show a low coherence between emotion and facial display (Durán, Reisenzein, \& Fernández-Dols, 2017; Reisenzein et al., 2013).
+Studies on happiness and related positive emotions such as sensory pleasantness also show a low coherence between emotion and facial display (Durán, Reisenzein, \& Fernández-Dols, 2017; Reisenzein et al., 2013).
 In fact, only experiments on amusement provide a fairly strong association between emotion and smiling.
-When confronted to humorous events (\emph{e.g.}, jokes, being tickled), the number of facially reactive participants (smiling and laughing) is indeed quite fair, whereas when confronted to other positive emotions (\emph{e.g.}, happiness), few participants show the expected facial expressions (for instance, the Duchenne smile or any kind of ``happy'' expression yet reporting being happy; see Mehu, Grammer, \& Dunbar, 2007; Lee \& Wagner, 2002, social condition).
+When confronted to humorous events (\emph{e.g.}, jokes, being tickled), the number of participants smiling and laughing is indeed quite fair, whereas when confronted to other positive emotions (\emph{e.g.}, happiness), few participants show the expected facial expressions (for instance, the Duchenne smile or any kind of expression related to happiness; see Mehu, Grammer, \& Dunbar, 2007; Lee \& Wagner, 2002).
 More largely, the meta-analysis conducted by Durán et al.
 (2017) on the degree of statistical covariation between emotions and facial expressions provides conclusive evidence.
-First of all, the commonly called basic emotions (happiness/amusement, sadness, fear, anger, disgust and surprise), on average, are weakly correlated with the specific configuration of facial muscles that the Basic Emotion View ascribes to them (the correlation even drops when happiness/amusement studies are excluded).
-Very few participants who relived an experience of sadness (\emph{e.g.}, Tsai, Chentsova-Dutton, Freire-Bebeau, \& Przymus, 2002) were reactive, only a small number of them showing components of a sad facial expression (\emph{e.g.}, oblique eyebrows, lip corners pulled down).
-Regarding fear, spider phobic participants recruited by Vernon \& Berenbaum (2002), for instance, were exposed to a live tarantula.
-About one third of them only displayed some components of the prototypical fear expression (eye widening, brow raising and knitting, etc.).
+First of all, the basic emotions are weakly correlated with the specific configuration of facial muscles that the Basic Emotion View ascribes to them (the correlation drops when happiness/amusement studies are excluded).
+Very few participants who relived an experience of sadness were reactive, only a small number of them showing components of a sad facial expression with oblique eyebrows or lip corners pulled down (Tsai, Chentsova-Dutton, Freire-Bebeau, \& Przymus, 2002).
+Regarding fear, about one third of spider phobic participants exposed to a live tarantula displayed some components of the prototypical fear expression such as eye widening, brow raising and knitting (Vernon \& Berenbaum, 2002).
 Similar observation is made for anger.
-The proportion of facially reactive participants (\emph{i.e.} displaying at least one component among frowning, lid/lip tightening\ldots{}) reporting anger in Johnson, Waugh, \& Fredrickson (2010) and Tsai et al.
-(2002) studies does not exceed 35\%.
+The proportion of facially reactive participants displaying at least one component among frowning or lid/lip tightening and reporting anger does not exceed 35\% (Johnson, Waugh, \& Fredrickson, 2010; Tsai et al., 2002).
 The same proportion of reactive participants confronted to disgusting elicitors is observed (\emph{e.g.}, Ekman, Freisen, \& Ancoli, 1980; Fernandez-Dols, Sanchez, Carrera, \& Ruiz-Belda, 1997).
@@
 The Basic Emotion View postulates that, when triggered, each basic emotion is expressed by a prototypical face (non-basic emotions being blends of the basic ones).
-In return, the recognition of emotional facial expressions (EFE) is claimed to be based on the identification of specific patterns of facial movements associated with each emotion, as if expression and recognition were the two sides of a same coin.
-Seemingly compelling evidence support this claim, sustaining the belief that one can read a face in the same way one reads the lines of a book.
+In return, the recognition of EFE is claimed to be based on the identification of specific patterns of facial movements associated with each emotion, as if expression and recognition were the two sides of a same coin.
+Seemingly compelling evidence support this claim, sustaining the possibility of a clear readout of subjective feelings from facial expressions.
 Besides, recognition systems rely on such a principle and are considered as an objective coding tool because they are based on the identification of specific muscular changes in the face.
-Many people highly agree that so-called basic emotions are associated with specific facial configurations (Ekman, 2017; Elfenbein \& Ambady, 2002; Russell, 1994) and also agree that many believe that it is a strong evidence for the Basic Emotion View.
+Many people highly agree that so-called basic emotions are associated with specific facial configurations (Ekman, 2017; Elfenbein \& Ambady, 2002) and also agree that many believe that it is a strong evidence for the Basic Emotion View.
 Moreover, it implies that EFE recognition of both human observers and automatic classifiers should be as accurate.
@@
 Among others, evidence has been questioned on methodological grounds (\emph{e.g.}, Russell, 1994).
-The response format usually used in recognition studies (\emph{i.e}.
-forced choice: selection of one word from a pre-specified list of emotion labels), notably, leads to a biased consensus as evidenced by Russell (1993).
-Depending on the list of emotion labels at participants' disposal, EFE of sadness can easily be categorized as sad expressions as well as fear expressions\ldots{} Russell \& Fehr (1987) also showed the same facial expression can be seen as expressing different types of emotions, depending on what other faces are seen.
+The response format usually used in recognition studies (\emph{i.e}, forced choice: selection of one word from a pre-specified list of emotion labels), notably, leads to a biased consensus(Russell, 1993).
+Depending on the list of emotion labels at participants' disposal, EFE of sadness can easily be categorized as sad expressions as well as fear expressions for example.
+Russell \& Fehr (1987) also showed the same facial expression can be seen as expressing different types of emotions, depending on what other faces are seen.
 DiGirolamo \& Russell (2017) conducted seven experiments that establish that high agreement between participants can be an artifact of the standard method.
@@
 Using alternative recognition methods (emotion satiation procedure, face-matching task, sorting task; \emph{e.g.}, Lindquist, Barrett, Bliss-Moreau, \& Russell, 2006; Gendron, Roberson, Vyver, \& Barrett, 2014), it has been shown on the contrary that facial muscle movements are not linked in a one-to-one manner to a specific discrete emotional experience (2017,p.~418).
-Instead, emotions are probably mentally constructed by the perceiver and mental categories of emotions are needed to accurately categorize facial movements, among other things (\emph{e.g.}, contextual information).
+Instead, emotions are probably mentally constructed by the perceiver and mental categories of emotions are needed to accurately categorize facial movements among contextual information.
 
@@
 It may be that differences in the social information displayed by static and dynamic expressions leads to facial recognition differential effects.
-Regarding the issue of spontaneous vs posed expressions (the latter are overused in experiments).
+Regarding the issue of spontaneous \emph{vs.} posed expressions (the latter are overused in experiments).
 As Meillon et al.
@@
 Starting from the empirical evidence suggesting that spontaneous facial expressions in ordinary life are equivocal, Dols (2017) argues in favor of a pragmatic conception of natural facial displays.
-He makes a plea for the idea that natural facial displays, rather than \enquote{saying} -- because they do not have a specific meaning he claims -- \enquote{make} things.
+He makes a plea for the idea that natural facial displays, rather than \enquote{saying}--because facial expressions do not have a specific meaning--\enquote{make} things.
 Facial expressions are actions in a communicative interaction.
 They do not express emotions but they \emph{\enquote{prompt, on the receiver's side, important inferences about the context, the sender, and the course of the interaction between sender and receiver}} (Dols, 2017, p.
-466, underlined by the author).
+466).
 As such, the fact that facial displays are able to signal emotions is a byproduct of one of its main function, the one of implementing actions performing practical ends.
@@
 They are not primarily communicative signals and even less the outlet of an internal state.
-Facial displays are (parts of) pragmatic actions aiming at orienting the person's relation to its environment, for instance maintaining, breaking, restoring, etc., the relation (Frijda, 2012).
-Facial displays are not \enquote{recognized} in semantic terms but are perceived as intentional actions.
-In the face of the continual flow of uninterrupted facial movements, the perceiver sees behaviors directed towards a goal.
-S.he translates the continuous flow of movements into coordinated sequences of actions holding a beginning and an end.
+Facial displays are parts of pragmatic actions aiming at orienting the person's relation to its environment.
+For instance facial displays are maintaining, breaking or restoring the relationship between sender and perceiver (Frijda, 2012).
+Facial displays are not recognized in semantic terms but are perceived as intentional actions.
+In the face of the continual flow of uninterrupted facial movements, perceivers see behaviors directed towards a goal.
+They translate the continuous flow of movements into coordinated sequences of actions holding a beginning and an end.
 Facial displays are not simple strings of action units which morphological configuration would be the prototype of a given emotion, and consequently identified as such (Ekman et al., 1987).
@@
 This is why even unauthentic facial displays can yet be recognized as emotional expressions.
-Guillaume Duchenne de Boulogne explains that the artist who has shaped the famous Laocoon antic sculpture, exhibited in the Vatican's museum, has made a modelling mistake since no face can display its emotional expression (Duchenne, 1876).
+Duchenne de Boulogne explains that the artist who has shaped the famous Laocoon antic sculpture, exhibited in the Vatican's museum, has made a modelling mistake since no face can display its emotional expression (Duchenne, 1876).
 Indeed, no muscular contraction can produce it.
@@
 His demonstration gives food for thought.
-Even though no objective coding system can correctly code the discordant facial features of the Laocoon's face, yet anyone can easily recognize the suffering and despair he admirably expresses.
+Even though no objective coding system can correctly code the discordant facial features of the Laocoon's face, yet anyone can easily recognize the suffering and despair he expresses.
 This example goes along the lines of the constructivist approach.
 
-The constructivist approach claims that facial displays are behaviors which meaning is inferred by the perceiver.
+The constructivist approach claims that facial displays are behaviors which meaning is inferred by perceivers.
 Findings support this observer dependence {[}Lindquist \& Gendron (2013); niedenthal2017embodied{]}.
@@
 participants and Himba participants from remote regions of Namibia (Gendron et al., 2014) and Hazda participants of the Eastern Rift Valley of Tanzania (Gendron et al., 2018) both groups with limited exposure to Western culture.
-Gendron and her colleagues demonstrated that facial expressions were not universally \enquote{recognized} in discrete emotional terms.
-Indeed, when Himba and Hazda participants did not have emotion concepts at their disposal to structure perception, they perceived the facial expressions as behaviors (\emph{e.g.}, looking, smelling) that didn't have a necessary relationship to emotions.
-They didn't infer inner states (such as emotional feelings) but they rather proceeded with action identification that pointed out the functions of behaviors (see also Crivelli, Russell, Jarillo, \& Fernández-Dols, 2017 for similar observations among a small-scale society of Papua New Guinea).
+Gendron and her colleagues demonstrated that facial expressions were not universally recognized in discrete emotional terms.
+Indeed, when Himba and Hazda participants did not have emotion concepts at their disposal to structure perception, they perceived the facial expressions as behaviors, such as looking or smelling, that didn't have a necessary relationship to emotions.
+They did not infer inner states but they rather proceeded with action identification that pointed out the functions of behaviors (see also Crivelli, Russell, Jarillo, \& Fernández-Dols, 2017 for similar observations among a small-scale society of Papua New Guinea).
 The constructivist approach considers that specific emotion categories, as conceptualized by Western cultures' knowledge, are casted on the perceived face to make meaning of the sender's facial displays.
@@
 First, the participants had to identify the emotional sequences by pressing the space bar of their keyboard to indicate the beginning and the end of the emotional sequences while watching the video.
-Second, the participants watched each emotional sequence previously identified and labeled the sequence using one of the 12 emotions proposed including six \enquote{basic} emotion labels (\emph{i.e.}, \emph{anger}, \emph{disgust}, \emph{fear}, \emph{happiness}, \emph{surprise} and \emph{sadness}) and six \enquote{non-basic} emotion labels (\emph{i.e.}, \emph{pride}, \emph{curiosity}, \emph{boredom}, \emph{shame}, \emph{humiliation}, and \emph{disappointment}).
+Second, the participants watched each emotional sequence previously identified and labeled the sequence using one of the 12 emotions proposed including six basic emotion labels (\emph{i.e.}, \emph{anger}, \emph{disgust}, \emph{fear}, \emph{happiness}, \emph{surprise} and \emph{sadness}) and six \enquote{non-basic} emotion labels (\emph{i.e.}, \emph{pride}, \emph{curiosity}, \emph{boredom}, \emph{shame}, \emph{humiliation}, and \emph{disappointment}).
 They also had the possibility to indicate that the sequence was expressing none of the proposed emotion labels.
@@
 Affdex's algorithm uses Histogram of Oriented Gradient (HOG) features and Support Vector Machine (SVM) classifiers in order to recognize facial expressions.
-For each video frame, Affdex identifies the probability \(p_{video_{i}.label_{j}.t_{k}}\) from 0 to 100 (rescaled to 0 to 1 for the analysis) of the face as expressing one of the six \enquote{basic} emotion labels (\emph{i.e.}, \emph{anger}, \emph{disgust}, \emph{fear}, \emph{happiness}, \emph{surprise} and \emph{sadness}) as well as additional psychological states such as \emph{valence}, \emph{engagement} or \emph{contempt}, and facial features such as \emph{cheek raise}, \emph{eye widen} or \emph{jaw drop}.
-
-For both human and automatic recognition, to determine which of the six \enquote{basic} emotions can be used to identify each video, the recognition probability for each label by frame was converted into odd ratio by label (Dente, Küster, Skora, \& Krumhuber, 2017).
+For each video frame, Affdex identifies the probability \(p_{video_{i}.label_{j}.t_{k}}\) from 0 to 100 (rescaled to 0 to 1 for the analysis) of the face as expressing one of the six basic emotion labels (\emph{i.e.}, \emph{anger}, \emph{disgust}, \emph{fear}, \emph{happiness}, \emph{surprise} and \emph{sadness}) as well as additional psychological states such as \emph{valence}, \emph{engagement} or \emph{contempt}, and facial features such as \emph{cheek raise}, \emph{eye widen} or \emph{jaw drop}.
+
+For both human and automatic recognition, to determine which of the six basic emotions can be used to identify each video, the recognition probability for each label by frame was converted into odd ratio by label (Dente, Küster, Skora, \& Krumhuber, 2017).
 The highest sum of each odd ratio time-series defines the label recognized by the automatic classifier (EQ\ref{eq:2} for human recognition and EQ\ref{eq:3} for automatic recognition).
@@
 
-Since encoders' self-reports, human annotations and the automatic recognition include data on \enquote{non-basic} emotion labels and features, the analysis is performed using only the six \enquote{basic} emotion labels in order to compare them.
+Since encoders' self-reports, human annotations and the automatic recognition include data on \enquote{non-basic} emotion labels and features, the analysis is performed using only the six basic emotion labels in order to compare them.
 The maximum score for self-reports, human annotations and automatic recognition is used to label the video.
@@
 \begin{TableNotes}[para]
-\item Note. textit{na.} values are produced when not enough data are available to compute accuracy indicators.
+\item Note. *na.* values are produced when not enough data are available to compute accuracy indicators.
 \end{TableNotes}
@@
 Therefore, facial movements do not express specific emotions because they do not carry intrinsic emotional signification.
-It is the observer that infers the emotional meaning of the facial expression, this interpretation depending on availability of different kinds of information (contextual information, linguistic categories\ldots{}).
+It is the observer that infers the emotional meaning of the facial expression, this interpretation depending on availability of different information such as context or linguistic categories.
 As a consequence, one can predict from the first line of thinking that individuals' emotional subjective feeling should be correlated to the recognition of facial expressions from both human observers and automatic classifiers whereas if emotions are social constructs, as stated by the second line of thinking, human observer's should be better at perceiving emotions expressed on the face than automatic classifiers.
@@
 However, the human decoders in the present study were presented with faces without any contextual information which could have helped them to shape more precisely their interpretation.
-Hassin, Aviezer and Bentin (2013) reviewed evidence to demonstrate that faces are inherently ambiguous and that observers rely on situational cues when they process facial displays (see also Aviezer, Hassin, Ryan, Grady, Susskind, Anderson, Moscovitch \& Bentin, 2008; Barrett, Mesquita, Gendron, 2011).
+Hassin, Aviezer, \& Bentin (2013) reviewed evidence to demonstrate that faces are inherently ambiguous and that observers rely on situational cues when they process facial displays (see also Aviezer et al., 2008; Barrett, Mesquita, \& Gendron, 2011).
 It happens that contextual information even shifts emotion perception.
@@
 Yet, decoders had no cue at disposal.
-So, without any possibility of integrating faces and context, their decoding accuracy is pretty much degraded as it is usual in such cases (e.g., Wagner, MacDonald, \& Manstead, 1986).
+So, without any possibility of integrating faces and context, their decoding accuracy is pretty much degraded as it is usual in such cases (\emph{e.g.}, Wagner, MacDonald, \& Manstead, 1986).
 However, they demonstrate their superiority on the classifier tool probably thanks to their capacity to rely on their previous personal experience to invent a context in which a face could display such an expression.
@@
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-aviezer2017inherently}{}%
+Aviezer, H., Ensenberg, N., \& Hassin, R. R. (2017). The inherently contextualized nature of facial emotion perception. \emph{Current Opinion in Psychology}, \emph{17}, 47--54.
+
+\leavevmode\hypertarget{ref-aviezer2008angry}{}%
+Aviezer, H., Hassin, R.
+R., Ryan, J., Grady, C., Susskind, J., Anderson, A., \ldots{} Bentin, S.
+(2008).
+Angry, disgusted, or afraid?
+Studies on the malleability of emotion perception.
+\emph{Psychological Science}, \emph{19}(7), 724--732.
+
 \leavevmode\hypertarget{ref-baltruvsaitis2016openface}{}%
@@
 \emph{Psychological Science in the Public Interest}, \emph{20}(1), 1--68.
+
+\leavevmode\hypertarget{ref-barrett2011context}{}%
+Barrett, L. F., Mesquita, B., \& Gendron, M. (2011). Context in emotion perception. \emph{Current Directions in Psychological Science}, \emph{20}(5), 286--290.
 
@@
 
+\leavevmode\hypertarget{ref-ekman2017facial}{}%
+Ekman, P. (2017). Facial expressions. In J.-M. Fernández-Dols \& J. A. Russell (Eds.), \emph{The science of facial expression} (pp. 39--56). New York, NY: Oxford University Press.
+
 \leavevmode\hypertarget{ref-ekman2011meant}{}%
@@
 
+\leavevmode\hypertarget{ref-hassin2013inherently}{}%
+Hassin, R. R., Aviezer, H., \& Bentin, S. (2013). Inherently ambiguous: Facial expressions of emotions, in context. \emph{Emotion Review}, \emph{5}(1), 60--65.
+
 \leavevmode\hypertarget{ref-hess2016emotional}{}%
@@
 
+\leavevmode\hypertarget{ref-wagner1986communication}{}%
+Wagner, H., MacDonald, C., \& Manstead, A. (1986). Communication of individual emotions by spontaneous facial expressions. \emph{Journal of Personality and Social Psychology}, \emph{50}(4), 737.
+
+\leavevmode\hypertarget{ref-aviezer2017inherently}{}%
+Aviezer, H., Ensenberg, N., \& Hassin, R. R. (2017). The inherently contextualized nature of facial emotion perception. \emph{Current Opinion in Psychology}, \emph{17}, 47--54.
+
+\leavevmode\hypertarget{ref-aviezer2008angry}{}%
+Aviezer, H., Hassin, R.
+R., Ryan, J., Grady, C., Susskind, J., Anderson, A., \ldots{} Bentin, S.
+(2008).
+Angry, disgusted, or afraid?
+Studies on the malleability of emotion perception.
+\emph{Psychological Science}, \emph{19}(7), 724--732.
+
 \leavevmode\hypertarget{ref-baltruvsaitis2016openface}{}%
@@
 
+\leavevmode\hypertarget{ref-wagner1986communication}{}%
+Wagner, H., MacDonald, C., \& Manstead, A. (1986). Communication of individual emotions by spontaneous facial expressions. \emph{Journal of Personality and Social Psychology}, \emph{50}(4), 737.
+
 \endgroup

### 2019-11-06 manuscript_apa.Rmd
manuscript apa revised

@@
 ---
-title             : "The Emotion–Facial expression link: Evidence from Human and Automatic Expression Recognition"
-shorttitle        : "The Emotion–Facial expression link"
+title             : "The emotion–facial expression link: Evidence from human and automatic expression recognition"
+shorttitle        : "The emotion–facial expression link"
 
@@
   While it has been taken for granted in the development of several automatic facial expression recognition tools, the question of the coherence between subjective feelings and facial expressions is still a subject of debate.
-On one hand, the Basic Emotion View conceives emotions as genetically hardwired and therefore being genuinely displayed through facial expressions.
+On one hand, the "Basic Emotion View" conceives emotions as genetically hardwired and therefore being genuinely displayed through facial expressions.
 On the other hand, the constructivist approach conceives emotions as socially constructed; the emotional meaning of a facial expression being inferred by the observer.
@@
 They also show a weak consistency between self-reported emotional states and facial emotional displays.
-Based on these results, the hypothesis of genetically hardwired emotion genuinely displayed is difficult to support, whereas the idea of emotion and facial expression socially constructed appears to be more likely.
+Based on these results, the hypothesis of genetically hardwired emotion genuinely displayed is difficult to support, whereas the idea of emotion and facial expression as being socially constructed appears to be more likely.
 Accordingly, automatic emotion recognition tools based on facial expressions should be questioned.
@@
 keywords          : "emotion, facial expression, self-report, human observer, automatic recognition."
-wordcount         : "7924"
+wordcount         : "7941"
 
@@
 floatsintext      : no
-figurelist        : no
-tablelist         : no
+figurelist        : yes
+tablelist         : yes
 footnotelist      : no
@@
 classoption       : "man"
-output: papaja::apa6_word
-# output            : 
-#   bookdown::pdf_book:
-#     base_format: papaja::apa6_pdf
+
+# output: papaja::apa6_word
+
+output            :
+  bookdown::pdf_book:
+    toc: false
+    number_sections: false
+    base_format: papaja::apa6_pdf
 
@@
 ```{r data}
-# self_report_data
+## self_report_data
 source(here::here("script/data_wrangling_self_report.R"))
@@
   dplyr::select(C_Video, sr_emotion = emotion, sr_score = value)
-# human_recognition_data
+
+## human_recognition_data
 source(here::here("script/data_wrangling_human_recognition.R"))
@@
   dplyr::select(C_Video, hr_emotion = emotion, hr_score = confidence_score)
-# automatic_recognition_data
+
+## automatic_recognition_data
 source(here::here("script/data_wrangling_automatic_recognition.R"))
@@
 With the development of commercial automatic facial expression recognition tools [@dupre2018accuracy], industries and governments are gradually implementing this technology in order to monitor humans’ emotions in various scenarios (*e.g.*, in marketing, healthcare, and the automotive industry to name a few).
-This technology rests on the premise that facial expressions provide a direct access to individuals’ subjective feelings, and that one can read the emotions displayed on a person’s face as in an open book.
-This premise is derived from the Basic Emotion View which suggests a one-to-one mapping between subective feeling and facial expression [@ekman1987universals; @ekman1988universality; @ekman1992argument; @ekman2007directed].
+This technology rests on the premise that facial expressions provide direct access to individuals’ subjective feelings, and that one can read the emotions displayed on a person’s face as one would an open book.
+This premise is derived from the "Basic Emotion View", a theory which suggests a one-to-one mapping between subective feeling and facial expression [@ekman1987universals; @ekman1988universality; @ekman1992argument; @ekman2007directed].
 Even if this premise is central to the modern mainstream approach of human emotion, recent research in affective science is challenging it.
-Once the Basic Emotion View is briefly described, its foremost criticisms are synthetically exposed.
-Inconsistencies and unpredicted findings research has uncovered come both from field observations and laboratory experiments.
+Once the Basic Emotion View is briefly described, its foremost criticisms will be exposed.
+Inconsistencies and unpredicted findings come both from field observations and laboratory experiments.
 They relate to two different issues.
-The first issue is the one of the sender’s production of spontaneous facial displays.
-The second issue has to do with the viewer’s interpretation of these facial displays.
-Building on this evidence, and as a further attempt to shed light on the emotion–face link, the present study is finally exposed.
-Its main goal is to tackle these two issues at the same time.
-It aims at identifying the extent to which ordinary people’s experienced emotions are displayed through identifiable emotional facial expressions (EFE) recognized as such both by human individuals and by automatic facial expression recognition tools.
+The first issue is the sender’s production of spontaneous facial displays.
+The second issue relates to the receiver’s interpretation of these facial displays.
+Building on this evidence, and as an attempt to shed further light on the emotion–face link, the present study's main goal is to tackle these two issues at the same time.
+It aims at identifying the extent to which ordinary people’s experienced emotions are displayed through identifiable Emotional Facial Expressions (EFE) recognized as such both by human individuals, and by automatic facial expression recognition tools.
 
@@
 More specifically, a set of six emotions (*fear*, *anger*, *surprise*, *disgust*, *sadness* and *joy*) are universally displayed and are genetically hardwired not only in humans [@ekman1992argument], but also in different animal species [@de2019mama].
-According to the view of the Basic Emotion View, *“when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including […] facial expressions.”* [@ekman2007directed, 49].
+According to the Basic Emotion View, *“when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including […] facial expressions.”* [@ekman2007directed, 49].
 To respond to criticisms, several amendments have been made to the Basic Emotion View, increasing the number of basic emotions from six to seven [@ekman1988universality] as well as adding the concept of “display rules” to explain cultural differences in the management of facial expressions [@ekman1987universals].
@@
 @rosenberg1994coherence reported the first evidence of coherence between self-report of emotion and displayed facial expressions.
-Participants were showed emotionally evocative films and reported their own emotions using a moment-by-moment reporting procedure.
+Participants were shown emotionally evocative films and reported their own emotions using a moment-by-moment reporting procedure.
 Analysis of participants’ facial expressions and reports of emotions showed that there was a high degree of temporal linkage and emotional agreement between facial expressions and self-reports.
@@
 The following  is a summary of key findings and conclusions both from field observations and laboratory experiments on spontaneous expression of emotions, all of them strongly subverting the Basic Emotion viewpoint [for reviews, @fernandez2013emotion; @reisenzein2013coherence; @russell2003facial].
-Mains issues regarding recognition studies are also swept across afterwards.
+Mains issues regarding recognition studies are also briefly discussed afterwards.
 
@@
 Naturalistic studies look at the ecological frequency of co-occurrence of certain emotions and facial displays [@fernandez2013emotion].
-Though they cannot be considered as a straightforward test of the triggering role of emotion on the facial behavior, they have the advantage of considering situations that cannot be created in a laboratory.
+Though they cannot be considered as a straightforward test of the triggering role of emotion on facial behavior, they have the advantage of considering situations that cannot be created in a laboratory.
 For instance, the ethological study of @kraut1979social contradicts the Basic Emotion View premise that a smile is the major component of a facial display associated with and caused by feelings of joy or happiness.
@@
 The Basic Emotion View postulates that, when triggered, each basic emotion is expressed by a prototypical face (non-basic emotions being blends of the basic ones).
-In return, the recognition of EFE is claimed to be based on the identification of specific patterns of facial movements associated with each emotion, as if expression and recognition were the two sides of a same coin.
-Seemingly compelling evidence support this claim, sustaining the possibility of a clear readout of subjective feelings from facial expressions.
-Besides, recognition systems rely on such a principle and are considered as an objective coding tool because they are based on the identification of specific muscular changes in the face.
-Many people highly agree that so-called basic emotions are associated with specific facial configurations [@ekman2017facial; @elfenbein2002universality] and also agree that many believe that it is a strong evidence for the Basic Emotion View.
-Moreover, it implies that EFE recognition of both human observers and automatic classifiers should be as accurate.
+In return, the recognition of EFE is claimed to be based on the identification of specific patterns of facial movements associated with each emotion, as if expression and recognition were the two sides of the same coin.
+Seemingly compelling evidence supports this claim, sustaining the possibility of a clear readout of subjective feelings from facial expressions.
+Furthermore, recognition systems rely on such a principle and are considered as an objective coding tool because they are based on the identification of specific muscular changes in the face.
+Many people strongly agree that so-called basic emotions are associated with specific facial configurations [@ekman2017facial; @elfenbein2002universality] and also argue that this is strong evidence for the Basic Emotion View.
+Moreover, it implies that EFE recognition of human observers should be as accurate as automatic classifiers'.
 
@@
 The response format usually used in recognition studies (*i.e*, forced choice: selection of one word from a pre-specified list of emotion labels), notably, leads to a biased consensus[@russell1993forced].
-Depending on the list of emotion labels at participants’ disposal, EFE of sadness can easily be categorized as sad expressions as well as fear expressions for example.
-@russell1987relativity also showed the same facial expression can be seen as expressing different types of emotions, depending on what other faces are seen.
-@digirolamo2017emotion conducted seven experiments that establish that high agreement between participants can be an artifact of the standard method.
-Thus, results gathered with forced choice cannot demonstrate the univocal link between emotion and facial expressions claimed by the Basic Emotion View.
-Using alternative recognition methods [emotion satiation procedure, face-matching task, sorting task; *e.g.*, @lindquist2006language; @gendron2014perceptions], it has been shown on the contrary that facial muscle movements are not linked in a one-to-one manner to a specific discrete emotional experience (2017,p.
-418).
+Depending on the list of emotion labels at participants’ disposal, EFE of sadness can easily be categorized as sad expressions as well as fear expressions, as one example.
+@russell1987relativity also shown that the same facial expression can be seen as expressing different types of emotions, depending on what other faces are seen.
+@digirolamo2017emotion conducted seven experiments that establish that high agreement between participants can be an artifact of the standard method used by EFE recognition studies.
+Thus, results gathered with forced choice cannot demonstrate the unequivocal link between emotion and facial expressions claimed by the Basic Emotion View.
+Using alternative recognition methods [emotion satiation procedure, face-matching task, sorting task; *e.g.*, @lindquist2006language; @gendron2014perceptions], it has been shown on the contrary that facial muscle movements are not linked in a one-to-one manner to a specific discrete emotional experience.
 Instead, emotions are probably mentally constructed by the perceiver and mental categories of emotions are needed to accurately categorize facial movements among contextual information.
@@
 It may be that differences in the social information displayed by static and dynamic expressions leads to facial recognition differential effects.
-Regarding the issue of spontaneous *vs.* posed expressions (the latter are overused in experiments).
-As @meillon2010dynemo conclude, EFE have been typically studied as static displays.
+Regarding the issue of spontaneous *vs.* posed expressions (the latter are overused in experiments), as @meillon2010dynemo conclude, EFE have been typically studied as static displays.
 As a consequence, even though the central role of the dynamics of facial expressions is endorsed, little is still known about the temporal course of facial expressions.
@@
 They do not express emotions but they *"prompt, on the receiver’s side, important inferences about the context, the sender, and the course of the interaction between sender and receiver"* [@dols2017natural, 466].
-As such, the fact that facial displays are able to signal emotions is a byproduct of one of its main function, the one of implementing actions performing practical ends.
+As such, the fact that facial displays are able to signal emotions is a byproduct of one of their main functions; implementing actions performing practical ends.
 Therefore, expressive displays hold a motor intention [@pacherie2003modes].
 They implement their aim which is their motor intention.
-They are not primarily communicative signals and even less the outlet of an internal state.
-Facial displays are parts of pragmatic actions aiming at orienting the person’s relation to its environment.
+They are not primarily communicative signals, and even less so can they be considered the outlet of an internal state.
+Facial displays are parts of pragmatic actions aiming at orienting the person’s relation to their environment.
 For instance facial displays are maintaining, breaking or restoring the relationship between sender and perceiver [@frijda2012recognition].
@@
 They translate the continuous flow of movements into coordinated sequences of actions holding a beginning and an end.
-Facial displays are not simple strings of action units which morphological configuration would be the prototype of a given emotion, and consequently identified as such [@ekman1987universals].
+Facial displays are not simple strings of action units, the morphological configuration of which would be the prototype of a given emotion, and consequently identified as such [@ekman1987universals].
 They are best conceived as a Gestalt, the same way as a string of musical notes establishes a melody [@tcherkassof2014emotions].
-This is why even unauthentic facial displays can yet be recognized as emotional expressions.
+This is why even inauthentic facial displays can still be recognized as emotional expressions.
 Duchenne de Boulogne explains that the artist who has shaped the famous Laocoon antic sculpture, exhibited in the Vatican’s museum, has made a modelling mistake since no face can display its emotional expression [@duchenne1876mecanisme].
@@
 His demonstration gives food for thought.
-Even though no objective coding system can correctly code the discordant facial features of the Laocoon’s face, yet anyone can easily recognize the suffering and despair he expresses.
-This example goes along the lines of the constructivist approach.
+Even though no objective coding system can correctly code the discordant facial features of the Laocoon’s face, anyone can easily recognize the suffering and despair he expresses.
+This example is aligned with the constructivist approach.
 
-The constructivist approach claims that facial displays are behaviors which meaning is inferred by perceivers.
+The constructivist approach claims that facial displays are behaviors whose meaning is inferred by perceivers.
 Findings support this observer dependence [@lindquist2013s; niedenthal2017embodied].
@@
 Indeed, when Himba and Hazda participants did not have emotion concepts at their disposal to structure perception, they perceived the facial expressions as behaviors, such as looking or smelling, that didn’t have a necessary relationship to emotions.
-They did not infer inner states but they rather proceeded with action identification that pointed out the functions of behaviors [see also @crivelli2017recognizing for similar observations among a small-scale society of Papua New Guinea].
-The constructivist approach considers that specific emotion categories, as conceptualized by Western cultures’ knowledge, are casted on the perceived face to make meaning of the sender’s facial displays.
+They did not infer inner states, rather they proceeded with action identification that pointed out the functions of behaviors [see also @crivelli2017recognizing for similar observations among a small-scale society of Papua New Guinea].
+The constructivist approach considers that specific emotion categories, as conceptualized by Western cultures’ knowledge, are cast on the perceived face to make meaning of the sender’s facial displays.
 Following this approach, faces convey a range of information essential for social communication.
@@
 These signals can convey individuals’ motivations and readiness [@frijda1997facial] or social messages [@fridlund1994human].
-As for emotional meaning, more specifically, it is shaped by the perceiver according to the specific context in which the facial displays are observed.
+As for emotional meaning, more specifically, this is shaped by the perceiver according to the specific context in which the facial displays are observed.
 
@@
 The two main competing approaches to facial expressiveness, the Basic Emotion View and the constructivist approach, entail completely opposite predictions regarding the decoding of facial expressions, as evidenced above.
-The present study aims at examining these predictions in order to provide empirical evidence to allow the discussion to evolve.
-To date, no systematic study has looked at the same time at facial expressions spontaneously displayed in reaction to emotional triggers and how they are decoded, both by human observers and by automatic emotion recognition tools based on the detection of facial muscular configurations.
+The present study aims to examine these predictions in order to provide empirical evidence to allow the discussion to evolve.
+To date, no systematic study has looked at facial expressions spontaneously displayed in reaction to emotional triggers and how they are decoded, both by human observers and by automatic emotion recognition tools based on the detection of facial muscular configurations.
 This study fills that gap.
@@
 More specifically, this study aims to examine the recognition of EFE produced by ordinary people during situations judged and/or self-reported to involve different emotions.
-It (a) examines consistency between ordinary people’s self-reported emotional experience and observers’ judgments of these ordinary people’s EFE, (b) examines consistency between ordinary people’s self-reported emotional experience and an automatic classifier’s analysis of these ordinary people’s EFE.
-In other words, it is interested in how people actually move their faces to express self-reported emotions, in how human observers accurately infer the expresser’s emotional state, and in how automatic recognition accurately code the expresser’s emotional state.
-It is expected a human observers’ superior ability to exactly recognize EFE as compared to automatic EFE recognition tools.
+It (a) examines consistency between ordinary people’s self-reported emotional experience and observers’ judgments of these ordinary people’s EFE, and (b) examines consistency between ordinary people’s self-reported emotional experience and an automatic classifier’s analysis of these ordinary people’s EFE.
+In other words, it is interested in how people actually move their faces to express self-reported emotions, in how human observers accurately infer the expresser’s emotional state, and in how automatic recognition accurately codes the expresser’s emotional state.
+Because of their superior ability to exactly recognize EFE, it is expected that human observers’ will have a higher accuracy to identify senders' subjective feeling than automatic EFE recognition tools.
 
@@
 The debriefing was also used to check that encoders did not guess the real purpose of the experiment (*e.g.*, being filmed while they were performing an emotional elicitation task) to guarantee facial expressions’ genuineness.
-All encoders gave their agreement on their data and video to be processed for research purposes only.
+All encoders gave their agreement on their data and recordings being processed for research purposes only.
 
@@
 
-This annotation procedure results in a uni-dimensional time-series for each video per human observer identifying for each second of the video which emotion was recognized.
+This annotation procedure resulted in a uni-dimensional time-series for each video per human observer, identifying for each second of the video which emotion was recognized.
 Then, time-series corresponding to the same video were aggregated to calculate the proportion of human observers $x_{video_{i}.label_{j}.t_{k}}$ for each second of the video per emotional label (EQ\ref{eq:1}).
@@
 The `r nrow(unique_video_count)` annotated video were processed with Affdex (SDK v3.4.1).
-Affdex is an automatic facial expression recognition classifier developed and distributed by Affectiva is a spin-off company resulting from the research activities of MIT media lab created in 2009 [@mcduff2016affdex].
+Affdex is an automatic facial expression recognition classifier developed and distributed by Affectiva, which is a spin-off company resulting from the research activities of MIT media lab created in 2009 [@mcduff2016affdex].
 Affdex’s algorithm uses Histogram of Oriented Gradient (HOG) features and Support Vector Machine (SVM) classifiers in order to recognize facial expressions.
@@
 
+Since encoders’ self-reports, human annotations and the automatic recognition include data on “non-basic” emotion labels and features, the analysis is performed using only the six basic emotion labels in order to compare them.
+The maximum rating for self-reports, human annotations and automatic recognition is used to label the video.
+In case of more than one label obtaining the maximum rating, the video is labeled as undetermined.
+
+The confusion matrices obtained by crossing encoders’ self-reports, human annotations and the automatic recognition are analysed by calulating two-sided Pearson correlation, accuracy which corresponds to the overall agreement rate averaged over cross-validation iterations and Cohen’s (unweighted) Kappa.
+
 # Results
-
-Since encoders’ self-reports, human annotations and the automatic recognition include data on “non-basic” emotion labels and features, the analysis is performed using only the six basic emotion labels in order to compare them.
-The maximum score for self-reports, human annotations and automatic recognition is used to label the video.
-In case of more than one label obtaining the maximum value, the video is labeled as undetermined.
 
@@
   recognition_result$hr_score,
+  method = "pearson",
   alternative = "two.sided"
@@
   recognition_result$ar_score,
+  method = "pearson",
   alternative = "two.sided"
@@
   recognition_result$ar_score, 
+  method = "pearson",
   alternative = "two.sided"
@@
   papaja::apa_print()
+
+# sr_n <- recognition_result %>% 
+#   dplyr::filter(sr_score == 1) %>% 
+#   dplyr::group_by(list_emotion) %>% 
+#   dplyr::tally()
 ```
@@
   dplyr::filter(sr_emotion != "undetermined") %>% 
-  dplyr::mutate_if(is.character,as.factor)
+  dplyr::mutate_if(is.character, as.factor)
 
@@
 The overall correlation of recognition and non-recognition between self-reported emotions and human observers recognition is significant but low (`r corr_sr_hr$full_result`).
-In order to identify differences according to the emotional labels, encoders’ subjective feelings are compared with human observers’ recognition in a confusion matrix (Figure \@ref(fig:confusionMatrix-sr-hr).
+In order to identify differences according to the emotional labels, encoders’ subjective feelings are compared with human observers’ recognition in a confusion matrix (Figure \@ref(fig:confusionMatrix-sr-hr)).
 
@@
     general_title = "",
-    general = "Note. *na.* values are produced when not enough data are available to compute accuracy indicators.", 
+    general = "Note. na. values are produced when not enough data are available to compute accuracy indicators.", 
     escape = FALSE,
@@
 
-Surprisingly the automatic classifier incorrectly recognized as *disgust* an significant proportion of videos in which *anger*, *happiness* and *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "anger", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
+Surprisingly the automatic classifier incorrectly recognized as *disgust* a significant proportion of videos in which *anger*, *happiness* and *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "anger", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
 In parallel, the automatic classifier recognized as *happiness* videos in which *fear* and *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "fear", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "disgust" & confusionMatrix_sr_ar_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
@@
 There is indeed an affinity between emotion and facial display [@frijda1997facial].
-However, claiming that there are unique “signatures” allowing using configurations of facial muscles to infer the presence of a specific emotion is misleading.
+However, claiming that unique facial muscles configurations are used both to express and to infer the presence of a specific emotion is misleading.
 As well as other nonverbal behaviors, facial movements are not only assumed to be determined by emotion but also by various other causes, such as psychological states (*e.g.*, motivations or pain), to say nothing of social context and sociocultural norms [@ekman1987universals].
@@
 It is the observer that infers the emotional meaning of the facial expression, this interpretation depending on availability of different information such as context or linguistic categories.
-As a consequence, one can predict from the first line of thinking that individuals’ emotional subjective feeling should be correlated to the recognition of facial expressions from both human observers and automatic classifiers whereas if emotions are social constructs, as stated by the second line of thinking, human observer’s should be better at perceiving emotions expressed on the face than automatic classifiers.
+As a consequence, one can predict from the first line of thinking that individuals’ emotional subjective feeling should be correlated to the recognition of facial expressions from both human observers and automatic classifiers whereas if emotions are social constructs, as stated by the second line of thinking, human observers should be better at perceiving emotions expressed on the face than automatic classifiers.
 
@@
 They show that human observers are more accurate than the automatic recognition tool to identify an individual’s subjective feeling on the basis of their face.
-Moreover, mistakes made by human observers look less arbitrary to the ones made by the classifier.
+Moreover, mistakes made by human observers look less arbitrary than the ones made by the classifier.
 For instance, even if a mix-up between disgust and anger is sometimes reported in recognition studies, odd confusions such as the present ones produced by the classifier have never been noted for human observers.
@@
 It happens that contextual information even shifts emotion perception.
-In the present case, the facial stimuli displayed to the decoders were as equivocal as are real-life facial expressions [@aviezer2017inherently].
-Yet, decoders had no cue at disposal.
-So, without any possibility of integrating faces and context, their decoding accuracy is pretty much degraded as it is usual in such cases [*e.g.*, @wagner1986communication].
-However, they demonstrate their superiority on the classifier tool probably thanks to their capacity to rely on their previous personal experience to invent a context in which a face could display such an expression.
-Once a credible context retrieved, they can affix an emotional label to the facial behavior.
+In the present case, the facial stimuli displayed to the decoders were as equivocal as real-life facial expressions [@aviezer2017inherently].
+Yet, decoders had no cue at their disposal.
+So, without any possibility of integrating faces and context, their decoding accuracy is reduced [*e.g.*, @wagner1986communication].
+However, their superiority to the classifier tool is probably owing to their capacity to rely on their previous personal experience to invent a context in which a face could display such an expression.
+Once a credible context is retrieved, they can affix an emotional label to the facial behavior.
 
@@
 In the case of the present study, it may be that insufficient emotion intensity accounts for the low number of visibly reactive participants.
-However, such line of reasoning fits badly with the Basic Emotion View according to which emotions and their related prototypical facial behavior are universal because they are considered as innate mechanisms allowing individuals to respond adaptively to evolutionary significant events (threats, opportunities...) encountered in the environment, whatever their potency.
+However, such line of reasoning fits badly with the Basic Emotion View according to which emotions and their related prototypical facial behavior are universal because they are considered as innate mechanisms allowing individuals to respond adaptively to evolutionary significant events (threats, opportunities *etc.*) encountered in the environment, whatever their potency.
 Moreover, various studies have used fairly strong emotion inductions without obtaining any visible facial display of any kind of emotion [@duran2017coherence].
-Besides, the present results show a quite good correlation between reported emotion and facial behavior for happiness.
-Thus, there are no reasons to believe that an explanation in terms of a too low emotion threshold applies for the other emotions under consideration.
+Moreover, the present results show a quite good correlation between reported emotion and facial behavior for happiness.
+Thus, there are no reasons to believe that an explanation in terms of too low an emotion threshold applies for the other emotions under consideration.
 The latter were triggered and measured with conceptually identical elicitation and assessment procedures [@reisenzein2013coherence].
@@
 Of course one can expect a difference between skilled annotators and novice ones regarding the assessment of emotional facial displays.
-This said, in everyday life, few people are specialist coders yet the quantity of successful social interactions proves lay persons recognize pretty well others’ facial behaviors.
+This said, in everyday life, few people are specialist coders yet the quantity of successful social interactions proves lay persons recognize quite well others’ facial behaviors.
 Therefore, and especially in the interest of generalizability, asking inexpert people seems relevant.
@@
 An alternative explanation relies in reducing the recognition bias involved in the classic recognition paradigm.
-Classic forced-choice paradigms obtain artificially high results, thus by using a more evolved approach observers’ accuracy may have been lowered.
+Classic forced-choice paradigms obtain artificially high results, thus by using a more evolved approach, observers’ accuracy may have been lowered.
 Another flaw is the lack of comparison with various facial expression recognition methods.
@@
 This temporal information is indisputably a key feature of facial activity.
-Not only observer-based judgements of facial displays must be compared to automated facial analysis, but also different kinds of human recognition measurements should be undertaken.
+It is not only observer-based judgements of facial displays which must be compared to automated facial analysis, but also different kinds of human recognition measurements should be undertaken.
 The challenge researchers are especially confronted with is to find ways to appropriately collect data regarding the perception of spontaneous and dynamic facial behavior.
@@
 However, it is known that encoders’ imagination can influence their expressiveness, for instance when they believe that their friend is doing the same (*vs.* a different) emotional task in another room [@jakobs1999social].
-Trying to exclude social influence by leaving encoders alone may be illusionary then.
+Trying to exclude social influence by leaving encoders alone may be illusionary.
 From an ecological perspective, it is even a mistake to exonerate behavioral observations from social contexts.
@@
 It is the case, for instance, of emotional mimicry in dynamic social interactions [@hess2016emotional].
-It is also the case of healthy partners interacting with schizophrenic patients whose facial activity is pretty much identical [@krause1998dyadic].
+It is also the case of healthy patients interacting with schizophrenic patients whose facial activity is almost identical [@krause1998dyadic].
 Hence, in order to better comprehend emotional communication in human relationships, experimental research should be corroborated with more ecological protocols.
@@
 Nowadays, automatic recognition systems are based on the coding of the facial muscular activations from which they infer the expressed emotion.
-Such automatic classifier tools take for granted that, when experienced, firstly an emotion is displayed on the face, secondly, in the form of a configuration of facial muscles that is all his own, and thirdly, which is recognized by the perceiver (be it a human being or an automatic system).
+Such automatic classifier tools take for granted that, when experienced, an emotion is: firstly, displayed on the face; secondly, in the form of a configuration of facial muscles that is unique to a person; and thirdly, recognized by the perceiver (human being or automatic classifier).
 These are the Basic Emotion View assertions, jeopardized by field observations and laboratory experiments on spontaneous expression of emotions, such as the present study.
@@
 They bring up several questions regarding the role the context plays in the emission and interpretation of the so-called facial expression of emotion.
-The finalization of operational and effective “reading emotional faces” devices rests on their answer.
-As a result, despite being one of the most investigated questions in affective science, the growing interest of industries and governments in tracking individual’s psychological states is hardly given satisfaction.
-Considering the above, the present results provide additional evidence that an individual’s subjective feeling cannot be inferred from facial expressions and in our case invalidate the hypothesis of hardwired emotions unambiguously displayed on the face.
-Even if emotions were hardwired, in everyday life one does not observe prototypical facial expressions because of their rarity and therefore research should be focused on analyzing non-prototypical facial expressions.
-Advancements in identifying “non-basic” emotion labels as well as non-prototypical facial expression have indeed been made in the development of automatic facial expression recognition tools [@mcduff2016discovering].
-However, these present results suggest that automatic facial expression recognition tools should merely evaluate facial morphology features such as action units [already evaluated in OpenFace, @baltruvsaitis2016openface, Affectiva’s Affdex, @mcduff2016affdex, or Vicar Vision’s FaceReader, @den2005facereader, to name a few] rather than inferring supposedly emotional or affective states.
-Trying to interpret facial displays as a means of determining underlying emotional state, in all likelihood, remains vain.
+The finalization of operational and effective “reading emotional faces” devices rests on the answers, if any, to the questions raised.
+As a result, despite being one of the most investigated questions in affective science, the growing interest of industries and governments in tracking individual’s psychological states is supported by controversial assumptions.
+Considering the above, the present results provide additional evidence that an individual’s subjective feelings cannot be inferred from facial expressions, and they invalidate the hypothesis of hardwired emotions unambiguously displayed on the face.
+Even if emotions were hardwired, in everyday life one does not observe prototypical facial expressions because of their rarity, and therefore research should be focused on analyzing non-prototypical facial expressions.
+Advancements in identifying “non-basic” emotion labels as well as non-prototypical facial expression have indeed occured with automatic facial expression recognition tools [@mcduff2016discovering].
+However, these present results suggest that automatic facial expression recognition tools should merely evaluate facial morphology features such as action units [already evaluated in OpenFace, @baltruvsaitis2016openface; Affectiva’s Affdex, @mcduff2016affdex; Vicar Vision’s FaceReader, @den2005facereader, to name a few] rather than inferring supposedly emotional or affective states.
+Trying to interpret facial displays as a means of determining underlying emotional state, in all likelihood, remains in vain.
 

### 2019-11-06 script/corr_test.R

@@ -1,90 +1,0 @@
-list_video <- self_report_score$C_Video
-list_emotion <- c("disgust","fear","happiness","surprise","sadness","anger")
-
-grid_data <- expand.grid(list_video = list_video, list_emotion = list_emotion)
-
-self_report_result <- self_report_score %>% 
-  dplyr::filter(sr_emotion != "undetermined") %>% 
-  dplyr::mutate(sr_score = 1)
-
-human_recognition_result <- human_recognition_score %>% 
-  dplyr::filter(hr_emotion != "undetermined") %>% 
-  dplyr::mutate(hr_score = 1)
-
-automatic_recognition_result <- automatic_recognition_score %>% 
-  dplyr::filter(ar_emotion != "undetermined") %>% 
-  dplyr::mutate(ar_score = 1)
-
-recognition_result <- grid_data %>% 
-  dplyr::left_join(self_report_result, by = c(
-    "list_video" = "C_Video",
-    "list_emotion" = "sr_emotion"
-    )
-  ) %>% 
-  dplyr::left_join(human_recognition_result, by = c(
-    "list_video" = "C_Video",
-    "list_emotion" = "hr_emotion"
-    )
-  ) %>% 
-  dplyr::left_join(automatic_recognition_result, by = c(
-    "list_video" = "C_Video",
-    "list_emotion" = "ar_emotion"
-    )
-  ) %>% 
-  replace(is.na(.), 0)
-
-recognition_result %>% 
-  dplyr::select(-list_video, -list_emotion) %>% 
-  cor(method = c("pearson")) %>%
-  ggcorrplot::ggcorrplot(
-    type = "lower",
-    outline.col = "white", 
-    lab = TRUE, 
-    legend.title = "Correlation\n(Pearson)", 
-    lab_size = 4)
-
-recognition_result %>% 
-  ggplot(aes(d = sr_score, m = hr_score)) + 
-  plotROC::geom_roc(labels = FALSE,pointsize = 0) 
-
-recognition_result %>% 
-  ggplot(aes(d = sr_score, m = ar_score)) + 
-  plotROC::geom_roc(labels = FALSE,pointsize = 0) 
-
-
-human_roc <- pROC::roc(recognition_result$sr_score, recognition_result$hr_score)
-human_auc <- pROC::auc(human_roc)
-human_auc
-
-automatic_roc <- pROC::roc(recognition_result$sr_score, recognition_result$ar_score)
-automatic_auc <- pROC::auc(automatic_roc)
[…]

### 2019-11-06 script/corr_test_all.R

@@ -1,51 +1,0 @@
-self_report_score_all <- self_report_tidy %>% 
-  dplyr::select(C_Video, emotion, sr_value = value)
-
-human_recognition_score_all <- average_timeline %>%
-  tidyr::gather(emotion,value, c("disgust","sadness","anger", "fear", "happiness", "surprise")) %>% 
-  dplyr::group_by(C_Video, emotion) %>%
-  dplyr::summarise(sum_emotion = sum(value)) %>% 
-  dplyr::group_by(C_Video) %>% 
-  dplyr::mutate(sum_video = sum(sum_emotion)) %>% 
-  dplyr::mutate(confidence_score = sum_emotion/sum_video) %>% 
-  dplyr::select(C_Video, emotion, hr_value = confidence_score)
-
-automatic_recognition_score_all <- automatic_recognition_data %>% 
-  dplyr::select(C_Video = filename, TimeStamp, happiness = joy, fear, disgust, sadness, anger, surprise) %>% 
-  dplyr::mutate(C_Video = stringr::str_replace(C_Video,".csv","")) %>% 
-  na.omit() %>% 
-  tidyr::gather(emotion, value, c("happiness", "fear", "disgust", "sadness", "anger", "surprise")) %>% 
-  dplyr::mutate(probabilities = value/100) %>% 
-  dplyr::mutate(odds = probabilities/(1-probabilities)) %>% 
-  dplyr::select(C_Video, TimeStamp, emotion, odds) %>% 
-  dplyr::group_by(C_Video, emotion) %>%
-  dplyr::summarise(sum_emotion = sum(odds)) %>% 
-  dplyr::group_by(C_Video) %>% 
-  dplyr::mutate(sum_video = sum(sum_emotion)) %>% 
-  dplyr::mutate(confidence_score = sum_emotion/sum_video) %>% 
-  dplyr::select(C_Video, emotion, ar_value = confidence_score)
-
-recognition_result_all <- self_report_score_all %>% 
-  dplyr::left_join(human_recognition_score_all, by = c(
-    "C_Video",
-    "emotion"
-  )
-  ) %>% 
-  dplyr::left_join(automatic_recognition_score_all, by = c(
-    "C_Video",
-    "emotion"
-  )
-  ) %>% 
-  replace(is.na(.), 0)
-
-recognition_result_all %>% 
-  dplyr::select(-C_Video, -emotion) %>% 
-  cor(method = c("pearson")) %>%
[…]

### 2019-11-06 script/corr_test_probabilities.R

@@ -1,112 +1,0 @@
-list_video <- self_report_score$C_Video
-list_emotion <- c("disgust","fear","happiness","surprise","sadness","anger")
-
-grid_data <- expand.grid(list_video = list_video, list_emotion = list_emotion)
-
-self_report_result <- self_report_score %>% 
-  dplyr::filter(sr_emotion != "undetermined") %>% 
-  dplyr::mutate(sr_score = 1)
-
-human_recognition_score_all <- average_timeline %>%
-  tidyr::gather(emotion,value, c("disgust","sadness","anger", "fear", "happiness", "surprise")) %>% 
-  dplyr::group_by(C_Video, emotion) %>%
-  dplyr::summarise(sum_emotion = sum(value)) %>% 
-  dplyr::group_by(C_Video) %>% 
-  dplyr::mutate(sum_video = sum(sum_emotion)) %>% 
-  dplyr::mutate(confidence_score = sum_emotion/sum_video) %>% 
-  dplyr::select(C_Video, emotion, hr_value = confidence_score)
-
-automatic_recognition_score_all <- automatic_recognition_data %>% 
-  dplyr::select(C_Video = filename, TimeStamp, happiness = joy, fear, disgust, sadness, anger, surprise) %>% 
-  dplyr::mutate(C_Video = stringr::str_replace(C_Video,".csv","")) %>% 
-  na.omit() %>% 
-  tidyr::gather(emotion, value, c("happiness", "fear", "disgust", "sadness", "anger", "surprise")) %>% 
-  dplyr::mutate(probabilities = value/100) %>% 
-  dplyr::mutate(odds = probabilities/(1-probabilities)) %>% 
-  dplyr::select(C_Video, TimeStamp, emotion, odds) %>% 
-  dplyr::group_by(C_Video, emotion) %>%
-  dplyr::summarise(sum_emotion = sum(odds)) %>% 
-  dplyr::group_by(C_Video) %>% 
-  dplyr::mutate(sum_video = sum(sum_emotion)) %>% 
-  dplyr::mutate(confidence_score = sum_emotion/sum_video) %>% 
-  dplyr::select(C_Video, emotion, ar_value = confidence_score)
-
-recognition_result_probabilities <- grid_data %>% 
-  dplyr::left_join(self_report_result, by = c(
-    "list_video" = "C_Video",
-    "list_emotion" = "sr_emotion"
-  )
-  ) %>% 
-  dplyr::left_join(human_recognition_score_all, by = c(
[…]

### 2019-11-06 script/data_wrangling_automatic_recognition.R

@@ -1,19 +1,24 @@
+################################################################################
+#                  Data Wrangling: Automatic Recognition                       #
+################################################################################
+
+# setup ------------------------------------------------------------------------
 library(tidyverse)
 
-# automatic_recognition_data <- here::here("./data/automatic_recognition_data/automatic_recognition_data") %>% 
-#   fs::dir_ls(regexp = "\\.csv$") %>% 
-#   purrr::map_dfr(readr::read_csv, .id = "source") %>% 
-#   dplyr::mutate(filename = base::basename(source))
-# 
-# readr::write_rds(automatic_recognition_data, here::here("./data/automatic_recognition_data/automatic_recognition_data.rds"))
+# upload data ------------------------------------------------------------------
+automatic_recognition_data <- "../../data/dynemo_feel_data/automatic_recognition_data/automatic_recognition_data.rds" %>% 
+  readr::read_rds()
 
-automatic_recognition_data <- readr::read_rds("C:/Users/dupred/OneDrive/data/dynemo_feel_data/automatic_recognition_data/automatic_recognition_data.rds")
-
+# data wrangling ---------------------------------------------------------------
 automatic_recognition_score <- automatic_recognition_data %>% 
   dplyr::select(C_Video = filename, TimeStamp, happiness = joy, fear, disgust, sadness, anger, surprise) %>% 
-  dplyr::mutate(C_Video = stringr::str_replace(C_Video,".csv","")) %>% 
+  dplyr::mutate(C_Video = stringr::str_replace(C_Video, ".csv", "")) %>% 
   na.omit() %>% 
-  tidyr::gather(emotion, value, c("happiness", "fear", "disgust", "sadness", "anger", "surprise")) %>% 
+  tidyr::gather(
+    emotion, 
+    value, 
+    c("happiness", "fear", "disgust", "sadness", "anger", "surprise")
+  ) %>% 
   dplyr::mutate(probabilities = value/100) %>% 
   dplyr::mutate(odds = probabilities/(1-probabilities)) %>% 
   dplyr::select(C_Video, TimeStamp, emotion, odds) %>% 
@@ -31,10 +36,12 @@ list_ties <- automatic_recognition_score %>%
   dplyr::mutate(ties = "yes") %>% 
   dplyr::select(-n)
 
-automatic_recognition_score <- dplyr::left_join(automatic_recognition_score, list_ties, by = c("C_Video")) %>% 
[…]

### 2019-11-06 script/data_wrangling_human_recognition.R

@@ -1,10 +1,16 @@
+################################################################################
+#                     Data Wrangling: Human Recognition                        #
+################################################################################
+
 # setup ------------------------------------------------------------------------
 library(tidyverse)
 library(splitstackshape)
-# upload data ------------------------------------------------------------------
-metadata_video <- readr::read_rds("C:/Users/dupred/OneDrive/data/dynemo_feel_data/human_recognition_data/metadata_video.rds")
 
-human_recognition_data <- "C:/Users/dupred/OneDrive/data/dynemo_feel_data/human_recognition_data" %>% 
+# upload data ------------------------------------------------------------------
+metadata_video <- "../../data/dynemo_feel_data/human_recognition_data/metadata_video.rds" %>% 
+  readr::read_rds()
+
+human_recognition_data <- "../../data/dynemo_feel_data/human_recognition_data" %>% 
   fs::dir_ls(regexp = "\\.csv$") %>% 
   purrr::map_dfr(readr::read_csv2, .id = "source", col_types = cols(
     C_INDUC = col_character(),
@@ -50,21 +56,25 @@ data_timeline <- human_recognition_data %>%
   unnest() %>% 
   dplyr::mutate_if(is.factor,as.character)
 
-final_timeline <- dplyr::left_join(full_timeline,data_timeline, by = c("source", "C_INDUC", "SEX_Sujet", "SEXE_Juge", "C_Juge", "C_Video", "timeline")) %>% 
-  replace(is.na(.),"no_rec") %>% 
-  dplyr::mutate(Curiosite = ifelse(E_Detectee == "Curiosite",1,0)) %>% 
-  dplyr::mutate(Deception = ifelse(E_Detectee == "Deception",1,0)) %>% 
-  dplyr::mutate(Degout = ifelse(E_Detectee == "Degout",1,0)) %>% 
-  dplyr::mutate(Emu = ifelse(E_Detectee == "Emu",1,0)) %>% 
-  dplyr::mutate(Enervement = ifelse(E_Detectee == "Enervement",1,0)) %>% 
[…]

### 2019-11-06 script/data_wrangling_self_report.R

@@ -1,9 +1,15 @@
+################################################################################
+#                  Data Wrangling: Automatic Recognition                       #
+################################################################################
+
 # setup ------------------------------------------------------------------------
 library(here)
 library(tidyverse)
 library(janitor)
+
 # upload data ------------------------------------------------------------------
-self_report_data <- readr::read_rds("C:/Users/dupred/OneDrive/data/dynemo_feel_data/human_self_report_data/Base_Finale_20090114.rds") %>% 
+self_report_data <- "../../data/dynemo_feel_data/human_self_report_data/Base_Finale_20090114.rds" %>% 
+  readr::read_rds() %>% 
   janitor::clean_names() %>% 
   dplyr::mutate(C_Video = stringr::str_replace(fichier_video_v,".mpg","")) %>% 
   dplyr::select(
@@ -78,7 +84,6 @@ self_report_tidy <- self_report_data %>%
   )
 
 # emotion score ----------------------------------------------------------------
-
 self_report_score <- self_report_tidy %>%
   dplyr::group_by(C_Video) %>%
   dplyr::filter(value == max(value)) %>% 
@@ -100,76 +105,3 @@ self_report_score <- dplyr::left_join(self_report_score, list_automatic_ties, by
     TRUE ~ emotion
   )) %>% 
   dplyr::select(-ties)
-
-
-# self_report_tidy <- self_report_tidy %>%
-#   dplyr::group_by(induction_c, emotion) %>%
-#   dplyr::summarise(mean_value = mean(value))
-
-# self_report_tidy %>% 
-#   ggplot(mapping = aes(x =  induction_c, y = emotion)) +
-#   geom_tile(aes(fill = mean_value), colour = "white") +
-#   geom_text(aes(label = round(mean_value,1)),
-#             color = "black",
-#             size = 3,
-#             #family="serif",
-#             #fontface = "bold",
-#             parse = FALSE,
-#             lineheight = 0.7
-#   ) +
-#   scale_x_discrete(name = "Elicitation Task") +
-#   scale_y_discrete(name = "Emotion Felt") +
[…]

### 2019-11-06 script/get_video_duration.R

@@ -1,33 +1,0 @@
-library(tidyverse)
-
-################################################################################
-get_video_duration <- function(file_path) {
-  
-  ffprob_path <- "C:\\Users\\dupred\\Desktop\\ffmpeg-20190312-d227ed5-win64-static\\ffmpeg-20190312-d227ed5-win64-static\\bin\\ffprobe.exe"
-  system_code <- "-show_entries format=duration"
-  
-  ffprobe_duration <- paste(ffprob_path, system_code,file_path)
-  
-  system(command = ffprobe_duration, intern = TRUE)[2]
-  
-}
-################################################################################
-
-list_video <- "C:\\Users\\dupred\\Desktop\\DynEmo_Database_mp4" %>% 
-  list.files(pattern = "\\.mp4$",full.names = TRUE)
-
-ffprobe_duration <- list_video %>% 
-  purrr::map(get_video_duration) %>% 
-  plyr::ldply(data.frame)%>% 
-  dplyr::rename_at(1,~"ffprobe_duration")
-
-metadata_video <- data.frame(list_video, ffprobe_duration) %>% 
-  dplyr::mutate_if(is.factor,as.character) %>% 
-  dplyr::mutate(C_Video = base::basename(list_video) %>% stringr::str_replace(".mp4", "")) %>% 
-  dplyr::mutate(ffprobe_duration = stringr::str_replace(ffprobe_duration, "duration=", "") %>% as.numeric()) %>% #as.numeric()*1000 for millisecond 
-  dplyr::mutate(ffprobe_duration = round(ffprobe_duration,0)) %>% 
-  dplyr::select(-list_video)
-
-# write_rds(metadata_video, "metadata_video.rds")
-
-

### 2019-11-06 manuscript_apa.Rmd
manuscript doc

@@
 
-# output: papaja::apa6_word
+output: papaja::apa6_word
 
-output            :
-  bookdown::pdf_book:
-    toc: false
-    number_sections: false
-    base_format: papaja::apa6_pdf
+# output            :
+#   bookdown::pdf_book:
+#     toc: false
+#     number_sections: false
+#     base_format: papaja::apa6_pdf
 
@@
 This annotation procedure resulted in a uni-dimensional time-series for each video per human observer, identifying for each second of the video which emotion was recognized.
-Then, time-series corresponding to the same video were aggregated to calculate the proportion of human observers $x_{video_{i}.label_{j}.t_{k}}$ for each second of the video per emotional label (EQ\ref{eq:1}).
+Then, time-series corresponding to the same video were aggregated to calculate the proportion of human observers $x_{video_{i}.label_{j}.t_{k}}$ for each second of the video per emotional label shown in Equation \@ref(eq:one).
 
 \begin{equation}
-\label{eq:1}
-x_{video_{i}.label_{j}.t_{k}} = \frac{n_{video_{i}.label_{j}.t_{k}}}{n_{video_{k}}}
+x_{video_{i}.label_{j}.t_{k}} = \frac{n_{video_{i}.label_{j}.t_{k}}}{n_{video_{k}}}(\#eq:one)
 \end{equation}
@@
 For both human and automatic recognition, to determine which of the six basic emotions can be used to identify each video, the recognition probability for each label by frame was converted into odd ratio by label [@dente2017measures].
-The highest sum of each odd ratio time-series defines the label recognized by the automatic classifier (EQ\ref{eq:2} for human recognition and EQ\ref{eq:3} for automatic recognition).
+The highest sum of each odd ratio time-series defines the label recognized (see Equation \@ref(eq:two) for human recognition and Equation \@ref(eq:three) for automatic recognition).
 
 \begin{equation}
-\label{eq:2}
-video_{i}.label = \max\left(\frac{\sum_{k=1}^{n}x_{video_{i}.label_{j}.t_{k}}}{\sum_{k=1}^{n}x_{video_{i}.t_{k}}}\right)
+video_{i}.label = \max\left(\frac{\sum_{k=1}^{n}x_{video_{i}.label_{j}.t_{k}}}{\sum_{k=1}^{n}x_{video_{i}.t_{k}}}\right)(\#eq:two)
 \end{equation}
@@
 \begin{equation}
-\label{eq:3}
-video_{i}.label = \max\left(\frac{\sum_{k=1}^{f}p_{video_{i}.label_{j}.t_{k}}}{\sum_{k=1}^{f}p_{video_{i}.t_{k}}}\right)
+video_{i}.label = \max\left(\frac{\sum_{k=1}^{f}p_{video_{i}.label_{j}.t_{k}}}{\sum_{k=1}^{f}p_{video_{i}.t_{k}}}\right)(\#eq:three)
 \end{equation}
@@
 recognition_result <- grid_data %>%
-  dplyr::left_join(self_report_result, by = c(
+  dplyr::left_join(
+    self_report_result,
+    by = c(
       "list_video" = "C_Video",
@@
     ) %>%
-  dplyr::left_join(human_recognition_result, by = c(
+  dplyr::left_join(
+    human_recognition_result,
+    by = c(
       "list_video" = "C_Video",
@@
     ) %>%
-  dplyr::left_join(automatic_recognition_result, by = c(
+  dplyr::left_join(
+    automatic_recognition_result,
+    by = c(
       "list_video" = "C_Video",
@@
   dplyr::mutate_if(is.numeric,round,2) %>% 
-  replace(is.na(.), "\\textit{na.}")
+  # replace(is.na(.), "\\textit{na.}")
+  replace(is.na(.), "na.")
 
 confusionMatrix_sr_hr_table %>%
-  knitr::kable(
-    "latex",
+  # knitr::kable(
+  #   "latex",
+  #   caption = "(ref:confusionTable-sr-hr)",
+  #   booktabs = TRUE,
+  #   digits = 2,
+  #   linesep = "",
+  #   escape = FALSE,
+  #   align = c('l', rep('c', 4)),
+  #   longtable = T
+  # ) %>%
+  # kableExtra::kable_styling(
+  #   latex_options = "HOLD_position",
+  #   font_size = 8,
+  #   full_width = TRUE 
+  # ) %>%
+  # kableExtra::footnote(
+  #   general_title = "",
+  #   general = "Note. na. values are produced when not enough data are available to compute accuracy indicators.", 
+  #   escape = FALSE,
+  #   footnote_as_chunk = TRUE,
+  #   threeparttable = TRUE
+  # )
+  papaja::apa_table(
     caption = "(ref:confusionTable-sr-hr)",
-    booktabs = TRUE,
-    digits = 2,
-    linesep = "",
-    escape = FALSE,
-    align = c('l', rep('c', 4)),
-    longtable = T
-  ) %>%
-  kableExtra::kable_styling(
-    latex_options = "HOLD_position",
-    font_size = 8,
-    full_width = TRUE 
-  ) %>%
-  kableExtra::footnote(
-    general_title = "",
-    general = "Note. na. values are produced when not enough data are available to compute accuracy indicators.", 
-    escape = FALSE,
-    footnote_as_chunk = TRUE,
-    threeparttable = TRUE
+    note = "na. values are produced when not enough data are available to compute accuracy indicators.",
+    escape = TRUE
   )
@@
   tibble::rownames_to_column("Emotion") %>%
-  dplyr::mutate(Emotion = gsub(pattern = "Class: ",replacement = "",x = Emotion)) %>% 
+  dplyr::mutate(Emotion = gsub(pattern = "Class: ", replacement = "", x = Emotion)) %>% 
   dplyr::select(Emotion, Sensitivity, Specificity, Precision, F1) %>% 
-  dplyr::mutate_if(is.numeric,round,2) %>% 
-  replace(is.na(.), "\\textit{na.}")
+  dplyr::mutate_if(is.numeric, round, 2) %>% 
+  #replace(is.na(.), "\\textit{na.}")
+  replace(is.na(.), "na.")
 
 confusionMatrix_sr_ar_table %>%
-  knitr::kable(
-    "latex",
+  # knitr::kable(
+  #   "latex",
+  #   caption = "(ref:confusionTable-sr-ar)",
+  #   booktabs = T,
+  #   digits = 2,
+  #   linesep = "",
+  #   escape = FALSE,
+  #   align = c('l',rep('c', 4)),
+  #   longtable = T
+  # ) %>%
+  # kableExtra::kable_styling(
+  #   latex_options = c("HOLD_position"),
+  #   font_size = 8,
+  #   full_width = TRUE
+  #   ) %>%
+  # kableExtra::footnote(
+  #   general_title = "",
+  #   general = "Note. na. values are produced when not enough data are available to compute accuracy indicators.", 
+  #   escape = FALSE,
+  #   footnote_as_chunk = TRUE,
+  #   threeparttable = TRUE
+  #   ) 
+  papaja::apa_table(
     caption = "(ref:confusionTable-sr-ar)",
-    booktabs = T,
-    digits = 2,
-    linesep = "",
-    escape = FALSE,
-    align = c('l',rep('c', 4)),
-    longtable = T
-  ) %>%
-  kableExtra::kable_styling(
-    latex_options = c("HOLD_position"),
-    font_size = 8,
-    full_width = TRUE
-    ) %>%
-  kableExtra::footnote(
-    general_title = "",
-    general = "Note. na. values are produced when not enough data are available to compute accuracy indicators.", 
-    escape = FALSE,
-    footnote_as_chunk = TRUE,
-    threeparttable = TRUE
+    note = "na. values are produced when not enough data are available to compute accuracy indicators.",
+    escape = TRUE
   )
@@
   geom_text(
-    aes(label = paste0(scales::percent(prop_hr), "\n(", Freq,"/",n_hr, ")")),
+    aes(label = paste0(scales::percent(prop_hr), "\n(", Freq, "/", n_hr, ")")),
     color = "black",

### 2019-11-06 manuscript_apa.tex
manuscript doc

@@
 \hypersetup{unicode=true,
-            pdftitle={The Emotion--Facial expression link: Evidence from Human and Automatic Expression Recognition},
+            pdftitle={The emotion--facial expression link: Evidence from human and automatic expression recognition},
             pdfauthor={Anna Tcherkassof~\& Damien Dupré},
-            pdfkeywords={Facial expression, self-report, human observer, automatic recognition.},
+            pdfkeywords={emotion, facial expression, self-report, human observer, automatic recognition.},
             pdfborder={0 0 0},
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
+\setcounter{secnumdepth}{0}
 % Redefines (sub)paragraphs to behave more like sections
@@
 
-  \title{The Emotion--Facial expression link: Evidence from Human and Automatic Expression Recognition}
+  \title{The emotion--facial expression link: Evidence from human and automatic expression recognition}
     \author{Anna Tcherkassof\textsuperscript{1}~\& Damien Dupré\textsuperscript{2}}
@@
 
-\shorttitle{The Emotion–Facial expression link}
+\shorttitle{The emotion–facial expression link}
 \affiliation{
@@
 \textsuperscript{1} University Grenoble Alpes, Grenoble, France\\\textsuperscript{2} Dublin City University, Dublin, Ireland}
-\keywords{Facial expression, self-report, human observer, automatic recognition.\newline\indent Word count: 7998}
+\keywords{emotion, facial expression, self-report, human observer, automatic recognition.\newline\indent Word count: 7941}
 \usepackage{csquotes}
@@
 \linenumbers
+\usepackage[titles]{tocloft}
+\cftpagenumbersoff{figure}
+\renewcommand{\cftfigpresnum}{\itshape\figurename\enspace}
+\renewcommand{\cftfigaftersnum}{.\space}
+\setlength{\cftfigindent}{0pt}
+\setlength{\cftafterloftitleskip}{0pt}
+\settowidth{\cftfignumwidth}{Figure 10.\qquad}
+\cftpagenumbersoff{table}
+\renewcommand{\cfttabpresnum}{\itshape\tablename\enspace}
+\renewcommand{\cfttabaftersnum}{.\space}
+\setlength{\cfttabindent}{0pt}
+\setlength{\cftafterloftitleskip}{0pt}
+\settowidth{\cfttabnumwidth}{Table 10.\qquad}
 \usepackage{booktabs}
@@
 While it has been taken for granted in the development of several automatic facial expression recognition tools, the question of the coherence between subjective feelings and facial expressions is still a subject of debate.
-On one hand, the Basic Emotion View conceives emotions as genetically hardwired and therefore being genuinely displayed through facial expressions.
+On one hand, the ``Basic Emotion View'' conceives emotions as genetically hardwired and therefore being genuinely displayed through facial expressions.
 On the other hand, the constructivist approach conceives emotions as socially constructed; the emotional meaning of a facial expression being inferred by the observer.
@@
 
-{
-\setcounter{tocdepth}{2}
-\tableofcontents
-}
-
 \hypertarget{introduction}{%
@@
 With the development of commercial automatic facial expression recognition tools (Dupré, Andelic, Morrison, \& McKeown, 2018), industries and governments are gradually implementing this technology in order to monitor humans' emotions in various scenarios (\emph{e.g.}, in marketing, healthcare, and the automotive industry to name a few).
-This technology rests on the premise that facial expressions provide a direct access to individuals' subjective feelings, and that one can read the emotions displayed on a person's face as in an open book.
-This premise is derived from the Basic Emotion View which suggests a one-to-one mapping between subective feeling and facial expression (Ekman, 1992, 2007; Ekman et al., 1987; Ekman \& Heider, 1988).
+This technology rests on the premise that facial expressions provide direct access to individuals' subjective feelings, and that one can read the emotions displayed on a person's face as one would an open book.
+This premise is derived from the \enquote{Basic Emotion View}, a theory which suggests a one-to-one mapping between subective feeling and facial expression (Ekman, 1992, 2007; Ekman et al., 1987; Ekman \& Heider, 1988).
 Even if this premise is central to the modern mainstream approach of human emotion, recent research in affective science is challenging it.
-Once the Basic Emotion View is briefly described, its foremost criticisms are synthetically exposed.
-Inconsistencies and unpredicted findings research has uncovered come both from field observations and laboratory experiments.
+Once the Basic Emotion View is briefly described, its foremost criticisms will be exposed.
+Inconsistencies and unpredicted findings come both from field observations and laboratory experiments.
 They relate to two different issues.
-The first issue is the one of the sender's production of spontaneous facial displays.
-The second issue has to do with the viewer's interpretation of these facial displays.
-Building on this evidence, and as a further attempt to shed light on the emotion--face link, the present study is finally exposed.
-Its main goal is to tackle these two issues at the same time.
-It aims at identifying the extent to which ordinary people's experienced emotions are displayed through identifiable emotional facial expressions (EFE) recognized as such both by human individuals and by automatic facial expression recognition tools.
+The first issue is the sender's production of spontaneous facial displays.
+The second issue relates to the receiver's interpretation of these facial displays.
+Building on this evidence, and as an attempt to shed further light on the emotion--face link, the present study's main goal is to tackle these two issues at the same time.
+It aims at identifying the extent to which ordinary people's experienced emotions are displayed through identifiable Emotional Facial Expressions (EFE) recognized as such both by human individuals, and by automatic facial expression recognition tools.
 
@@
 More specifically, a set of six emotions (\emph{fear}, \emph{anger}, \emph{surprise}, \emph{disgust}, \emph{sadness} and \emph{joy}) are universally displayed and are genetically hardwired not only in humans (Ekman, 1992), but also in different animal species (Waal, 2019).
-According to the view of the Basic Emotion View, \emph{\enquote{when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including {[}\ldots{}{]} facial expressions.}} (Ekman, 2007, p.
+According to the Basic Emotion View, \emph{\enquote{when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including {[}\ldots{}{]} facial expressions.}} (Ekman, 2007, p.
 49).
@@
 Rosenberg \& Ekman (1994) reported the first evidence of coherence between self-report of emotion and displayed facial expressions.
-Participants were showed emotionally evocative films and reported their own emotions using a moment-by-moment reporting procedure.
+Participants were shown emotionally evocative films and reported their own emotions using a moment-by-moment reporting procedure.
 Analysis of participants' facial expressions and reports of emotions showed that there was a high degree of temporal linkage and emotional agreement between facial expressions and self-reports.
@@
 The following is a summary of key findings and conclusions both from field observations and laboratory experiments on spontaneous expression of emotions, all of them strongly subverting the Basic Emotion viewpoint (for reviews, Fernández-Dols \& Crivelli, 2013; Reisenzein, Studtmann, \& Horstmann, 2013; Russell, Bachorowski, \& Fernández-Dols, 2003).
-Mains issues regarding recognition studies are also swept across afterwards.
+Mains issues regarding recognition studies are also briefly discussed afterwards.
 
@@
 Naturalistic studies look at the ecological frequency of co-occurrence of certain emotions and facial displays (Fernández-Dols \& Crivelli, 2013).
-Though they cannot be considered as a straightforward test of the triggering role of emotion on the facial behavior, they have the advantage of considering situations that cannot be created in a laboratory.
+Though they cannot be considered as a straightforward test of the triggering role of emotion on facial behavior, they have the advantage of considering situations that cannot be created in a laboratory.
 For instance, the ethological study of Kraut \& Johnston (1979) contradicts the Basic Emotion View premise that a smile is the major component of a facial display associated with and caused by feelings of joy or happiness.
@@
 The Basic Emotion View postulates that, when triggered, each basic emotion is expressed by a prototypical face (non-basic emotions being blends of the basic ones).
-In return, the recognition of EFE is claimed to be based on the identification of specific patterns of facial movements associated with each emotion, as if expression and recognition were the two sides of a same coin.
-Seemingly compelling evidence support this claim, sustaining the possibility of a clear readout of subjective feelings from facial expressions.
-Besides, recognition systems rely on such a principle and are considered as an objective coding tool because they are based on the identification of specific muscular changes in the face.
-Many people highly agree that so-called basic emotions are associated with specific facial configurations (Ekman, 2017; Elfenbein \& Ambady, 2002) and also agree that many believe that it is a strong evidence for the Basic Emotion View.
-Moreover, it implies that EFE recognition of both human observers and automatic classifiers should be as accurate.
+In return, the recognition of EFE is claimed to be based on the identification of specific patterns of facial movements associated with each emotion, as if expression and recognition were the two sides of the same coin.
+Seemingly compelling evidence supports this claim, sustaining the possibility of a clear readout of subjective feelings from facial expressions.
+Furthermore, recognition systems rely on such a principle and are considered as an objective coding tool because they are based on the identification of specific muscular changes in the face.
+Many people strongly agree that so-called basic emotions are associated with specific facial configurations (Ekman, 2017; Elfenbein \& Ambady, 2002) and also argue that this is strong evidence for the Basic Emotion View.
+Moreover, it implies that EFE recognition of human observers should be as accurate as automatic classifiers'.
 
@@
 The response format usually used in recognition studies (\emph{i.e}, forced choice: selection of one word from a pre-specified list of emotion labels), notably, leads to a biased consensus(Russell, 1993).
-Depending on the list of emotion labels at participants' disposal, EFE of sadness can easily be categorized as sad expressions as well as fear expressions for example.
-Russell \& Fehr (1987) also showed the same facial expression can be seen as expressing different types of emotions, depending on what other faces are seen.
-DiGirolamo \& Russell (2017) conducted seven experiments that establish that high agreement between participants can be an artifact of the standard method.
-Thus, results gathered with forced choice cannot demonstrate the univocal link between emotion and facial expressions claimed by the Basic Emotion View.
-Using alternative recognition methods (emotion satiation procedure, face-matching task, sorting task; \emph{e.g.}, Lindquist, Barrett, Bliss-Moreau, \& Russell, 2006; Gendron, Roberson, Vyver, \& Barrett, 2014), it has been shown on the contrary that facial muscle movements are not linked in a one-to-one manner to a specific discrete emotional experience (2017,p.~418).
+Depending on the list of emotion labels at participants' disposal, EFE of sadness can easily be categorized as sad expressions as well as fear expressions, as one example.
+Russell \& Fehr (1987) also shown that the same facial expression can be seen as expressing different types of emotions, depending on what other faces are seen.
+DiGirolamo \& Russell (2017) conducted seven experiments that establish that high agreement between participants can be an artifact of the standard method used by EFE recognition studies.
+Thus, results gathered with forced choice cannot demonstrate the unequivocal link between emotion and facial expressions claimed by the Basic Emotion View.
+Using alternative recognition methods (emotion satiation procedure, face-matching task, sorting task; \emph{e.g.}, Lindquist, Barrett, Bliss-Moreau, \& Russell, 2006; Gendron, Roberson, Vyver, \& Barrett, 2014), it has been shown on the contrary that facial muscle movements are not linked in a one-to-one manner to a specific discrete emotional experience.
 Instead, emotions are probably mentally constructed by the perceiver and mental categories of emotions are needed to accurately categorize facial movements among contextual information.
@@
 They show that the dynamic aspects of facial movement are likely to be of importance (\emph{e.g.}, Kamachi et al., 2013).
-Cohn \& Schmidt (2003) have shown that spontaneous smiles are of smaller amplitude and have a more consistent relation between amplitude and duration than deliberate smiles.
+Cohn \& Schmidt (2004) have shown that spontaneous smiles are of smaller amplitude and have a more consistent relation between amplitude and duration than deliberate smiles.
 Hess \& Kleck (1990) have also pointed out the importance of the dynamics of facial movements, and particularly the irregularity, or phasic changes, of the expressions' unfolding.
@@
 It may be that differences in the social information displayed by static and dynamic expressions leads to facial recognition differential effects.
-Regarding the issue of spontaneous \emph{vs.} posed expressions (the latter are overused in experiments).
-As Meillon et al.
+Regarding the issue of spontaneous \emph{vs.} posed expressions (the latter are overused in experiments), as Meillon et al.
 (2010) conclude, EFE have been typically studied as static displays.
@@
 466).
-As such, the fact that facial displays are able to signal emotions is a byproduct of one of its main function, the one of implementing actions performing practical ends.
+As such, the fact that facial displays are able to signal emotions is a byproduct of one of their main functions; implementing actions performing practical ends.
 Therefore, expressive displays hold a motor intention (Pacherie, 2003).
 They implement their aim which is their motor intention.
-They are not primarily communicative signals and even less the outlet of an internal state.
-Facial displays are parts of pragmatic actions aiming at orienting the person's relation to its environment.
+They are not primarily communicative signals, and even less so can they be considered the outlet of an internal state.
+Facial displays are parts of pragmatic actions aiming at orienting the person's relation to their environment.
 For instance facial displays are maintaining, breaking or restoring the relationship between sender and perceiver (Frijda, 2012).
@@
 They translate the continuous flow of movements into coordinated sequences of actions holding a beginning and an end.
-Facial displays are not simple strings of action units which morphological configuration would be the prototype of a given emotion, and consequently identified as such (Ekman et al., 1987).
+Facial displays are not simple strings of action units, the morphological configuration of which would be the prototype of a given emotion, and consequently identified as such (Ekman et al., 1987).
 They are best conceived as a Gestalt, the same way as a string of musical notes establishes a melody (Tcherkassof \& Frijda, 2014).
-This is why even unauthentic facial displays can yet be recognized as emotional expressions.
+This is why even inauthentic facial displays can still be recognized as emotional expressions.
 Duchenne de Boulogne explains that the artist who has shaped the famous Laocoon antic sculpture, exhibited in the Vatican's museum, has made a modelling mistake since no face can display its emotional expression (Duchenne, 1876).
@@
 His demonstration gives food for thought.
-Even though no objective coding system can correctly code the discordant facial features of the Laocoon's face, yet anyone can easily recognize the suffering and despair he expresses.
-This example goes along the lines of the constructivist approach.
-
-The constructivist approach claims that facial displays are behaviors which meaning is inferred by perceivers.
+Even though no objective coding system can correctly code the discordant facial features of the Laocoon's face, anyone can easily recognize the suffering and despair he expresses.
+This example is aligned with the constructivist approach.
+
+The constructivist approach claims that facial displays are behaviors whose meaning is inferred by perceivers.
 Findings support this observer dependence {[}Lindquist \& Gendron (2013); niedenthal2017embodied{]}.
@@
 Indeed, when Himba and Hazda participants did not have emotion concepts at their disposal to structure perception, they perceived the facial expressions as behaviors, such as looking or smelling, that didn't have a necessary relationship to emotions.
-They did not infer inner states but they rather proceeded with action identification that pointed out the functions of behaviors (see also Crivelli, Russell, Jarillo, \& Fernández-Dols, 2017 for similar observations among a small-scale society of Papua New Guinea).
-The constructivist approach considers that specific emotion categories, as conceptualized by Western cultures' knowledge, are casted on the perceived face to make meaning of the sender's facial displays.
+They did not infer inner states, rather they proceeded with action identification that pointed out the functions of behaviors (see also Crivelli, Russell, Jarillo, \& Fernández-Dols, 2017 for similar observations among a small-scale society of Papua New Guinea).
+The constructivist approach considers that specific emotion categories, as conceptualized by Western cultures' knowledge, are cast on the perceived face to make meaning of the sender's facial displays.
 Following this approach, faces convey a range of information essential for social communication.
@@
 These signals can convey individuals' motivations and readiness (Frijda \& Tcherkassof, 1997) or social messages (Fridlund, 1994).
-As for emotional meaning, more specifically, it is shaped by the perceiver according to the specific context in which the facial displays are observed.
+As for emotional meaning, more specifically, this is shaped by the perceiver according to the specific context in which the facial displays are observed.
 
@@
 More specifically, this study aims to examine the recognition of EFE produced by ordinary people during situations judged and/or self-reported to involve different emotions.
-It (a) examines consistency between ordinary people's self-reported emotional experience and observers' judgments of these ordinary people's EFE, (b) examines consistency between ordinary people's self-reported emotional experience and an automatic classifier's analysis of these ordinary people's EFE.
-In other words, it is interested in how people actually move their faces to express self-reported emotions, in how human observers accurately infer the expresser's emotional state, and in how automatic recognition accurately code the expresser's emotional state.
-It is expected a human observers' superior ability to exactly recognize EFE as compared to automatic EFE recognition tools.
+It (a) examines consistency between ordinary people's self-reported emotional experience and observers' judgments of these ordinary people's EFE, and (b) examines consistency between ordinary people's self-reported emotional experience and an automatic classifier's analysis of these ordinary people's EFE.
+In other words, it is interested in how people actually move their faces to express self-reported emotions, in how human observers accurately infer the expresser's emotional state, and in how automatic recognition accurately codes the expresser's emotional state.
+Because of their superior ability to exactly recognize EFE, it is expected that human observers' will have a higher accuracy to identify senders' subjective feeling than automatic EFE recognition tools.
 
@@
 The debriefing was also used to check that encoders did not guess the real purpose of the experiment (\emph{e.g.}, being filmed while they were performing an emotional elicitation task) to guarantee facial expressions' genuineness.
-All encoders gave their agreement on their data and video to be processed for research purposes only.
+All encoders gave their agreement on their data and recordings being processed for research purposes only.
 
@@
 
-This annotation procedure results in a uni-dimensional time-series for each video per human observer identifying for each second of the video which emotion was recognized.
+This annotation procedure resulted in a uni-dimensional time-series for each video per human observer, identifying for each second of the video which emotion was recognized.
 Then, time-series corresponding to the same video were aggregated to calculate the proportion of human observers \(x_{video_{i}.label_{j}.t_{k}}\) for each second of the video per emotional label (EQ\ref{eq:1}).
@@
 The 232 annotated video were processed with Affdex (SDK v3.4.1).
-Affdex is an automatic facial expression recognition classifier developed and distributed by Affectiva is a spin-off company resulting from the research activities of MIT media lab created in 2009 (McDuff et al., 2016).
+Affdex is an automatic facial expression recognition classifier developed and distributed by Affectiva, which is a spin-off company resulting from the research activities of MIT media lab created in 2009 (McDuff et al., 2016).
 Affdex's algorithm uses Histogram of Oriented Gradient (HOG) features and Support Vector Machine (SVM) classifiers in order to recognize facial expressions.
@@
 
+Since encoders' self-reports, human annotations and the automatic recognition include data on \enquote{non-basic} emotion labels and features, the analysis is performed using only the six basic emotion labels in order to compare them.
+The maximum rating for self-reports, human annotations and automatic recognition is used to label the video.
+In case of more than one label obtaining the maximum rating, the video is labeled as undetermined.
+
+The confusion matrices obtained by crossing encoders' self-reports, human annotations and the automatic recognition are analysed by calulating two-sided Pearson correlation, accuracy which corresponds to the overall agreement rate averaged over cross-validation iterations and Cohen's (unweighted) Kappa.
+
 \hypertarget{results}{%
@@
 
-Since encoders' self-reports, human annotations and the automatic recognition include data on \enquote{non-basic} emotion labels and features, the analysis is performed using only the six basic emotion labels in order to compare them.
-The maximum score for self-reports, human annotations and automatic recognition is used to label the video.
-In case of more than one label obtaining the maximum value, the video is labeled as undetermined.
-
 \hypertarget{human-observers-accuracy}{%
@@
 The overall correlation of recognition and non-recognition between self-reported emotions and human observers recognition is significant but low (\(r = .24\), 95\% CI \([.19\), \(.29]\), \(t(1384) = 9.21\), \(p < .001\)).
-In order to identify differences according to the emotional labels, encoders' subjective feelings are compared with human observers' recognition in a confusion matrix (Figure \ref{fig:confusionMatrix-sr-hr}.
+In order to identify differences according to the emotional labels, encoders' subjective feelings are compared with human observers' recognition in a confusion matrix (Figure \ref{fig:confusionMatrix-sr-hr}).
 
@@
 \begin{TableNotes}[para]
-\item Note. *na.* values are produced when not enough data are available to compute accuracy indicators.
+\item Note. na. values are produced when not enough data are available to compute accuracy indicators.
 \end{TableNotes}
@@
 
-Surprisingly the automatic classifier incorrectly recognized as \emph{disgust} an significant proportion of videos in which \emph{anger}, \emph{happiness} and \emph{surprise} was the highest self-reported emotion (respectively 42.9\%, 25.0\% and 32.6\% of the videos self-reported).
+Surprisingly the automatic classifier incorrectly recognized as \emph{disgust} a significant proportion of videos in which \emph{anger}, \emph{happiness} and \emph{surprise} was the highest self-reported emotion (respectively 42.9\%, 25.0\% and 32.6\% of the videos self-reported).
 In parallel, the automatic classifier recognized as \emph{happiness} videos in which \emph{fear} and \emph{surprise} was the highest self-reported emotion (respectively 66.7\% and 32.6\% of the videos self-reported).
@@
 There is indeed an affinity between emotion and facial display (Frijda \& Tcherkassof, 1997).
-However, claiming that there are unique \enquote{signatures} allowing using configurations of facial muscles to infer the presence of a specific emotion is misleading.
+However, claiming that unique facial muscles configurations are used both to express and to infer the presence of a specific emotion is misleading.
 As well as other nonverbal behaviors, facial movements are not only assumed to be determined by emotion but also by various other causes, such as psychological states (\emph{e.g.}, motivations or pain), to say nothing of social context and sociocultural norms (Ekman et al., 1987).
@@
 It is the observer that infers the emotional meaning of the facial expression, this interpretation depending on availability of different information such as context or linguistic categories.
-As a consequence, one can predict from the first line of thinking that individuals' emotional subjective feeling should be correlated to the recognition of facial expressions from both human observers and automatic classifiers whereas if emotions are social constructs, as stated by the second line of thinking, human observer's should be better at perceiving emotions expressed on the face than automatic classifiers.
+As a consequence, one can predict from the first line of thinking that individuals' emotional subjective feeling should be correlated to the recognition of facial expressions from both human observers and automatic classifiers whereas if emotions are social constructs, as stated by the second line of thinking, human observers should be better at perceiving emotions expressed on the face than automatic classifiers.
 
@@
 They show that human observers are more accurate than the automatic recognition tool to identify an individual's subjective feeling on the basis of their face.
-Moreover, mistakes made by human observers look less arbitrary to the ones made by the classifier.
+Moreover, mistakes made by human observers look less arbitrary than the ones made by the classifier.
 For instance, even if a mix-up between disgust and anger is sometimes reported in recognition studies, odd confusions such as the present ones produced by the classifier have never been noted for human observers.
@@
 It happens that contextual information even shifts emotion perception.
-In the present case, the facial stimuli displayed to the decoders were as equivocal as are real-life facial expressions (Aviezer, Ensenberg, \& Hassin, 2017).
-Yet, decoders had no cue at disposal.
-So, without any possibility of integrating faces and context, their decoding accuracy is pretty much degraded as it is usual in such cases (\emph{e.g.}, Wagner, MacDonald, \& Manstead, 1986).
-However, they demonstrate their superiority on the classifier tool probably thanks to their capacity to rely on their previous personal experience to invent a context in which a face could display such an expression.
-Once a credible context retrieved, they can affix an emotional label to the facial behavior.
+In the present case, the facial stimuli displayed to the decoders were as equivocal as real-life facial expressions (Aviezer, Ensenberg, \& Hassin, 2017).
+Yet, decoders had no cue at their disposal.
+So, without any possibility of integrating faces and context, their decoding accuracy is reduced (\emph{e.g.}, Wagner, MacDonald, \& Manstead, 1986).
+However, their superiority to the classifier tool is probably owing to their capacity to rely on their previous personal experience to invent a context in which a face could display such an expression.
+Once a credible context is retrieved, they can affix an emotional label to the facial behavior.
 
@@
 In the case of the present study, it may be that insufficient emotion intensity accounts for the low number of visibly reactive participants.
-However, such line of reasoning fits badly with the Basic Emotion View according to which emotions and their related prototypical facial behavior are universal because they are considered as innate mechanisms allowing individuals to respond adaptively to evolutionary significant events (threats, opportunities\ldots{}) encountered in the environment, whatever their potency.
+However, such line of reasoning fits badly with the Basic Emotion View according to which emotions and their related prototypical facial behavior are universal because they are considered as innate mechanisms allowing individuals to respond adaptively to evolutionary significant events (threats, opportunities \emph{etc.}) encountered in the environment, whatever their potency.
 Moreover, various studies have used fairly strong emotion inductions without obtaining any visible facial display of any kind of emotion (Durán et al., 2017).
-Besides, the present results show a quite good correlation between reported emotion and facial behavior for happiness.
-Thus, there are no reasons to believe that an explanation in terms of a too low emotion threshold applies for the other emotions under consideration.
+Moreover, the present results show a quite good correlation between reported emotion and facial behavior for happiness.
+Thus, there are no reasons to believe that an explanation in terms of too low an emotion threshold applies for the other emotions under consideration.
 The latter were triggered and measured with conceptually identical elicitation and assessment procedures (Reisenzein et al., 2013).
@@
 Of course one can expect a difference between skilled annotators and novice ones regarding the assessment of emotional facial displays.
-This said, in everyday life, few people are specialist coders yet the quantity of successful social interactions proves lay persons recognize pretty well others' facial behaviors.
+This said, in everyday life, few people are specialist coders yet the quantity of successful social interactions proves lay persons recognize quite well others' facial behaviors.
 Therefore, and especially in the interest of generalizability, asking inexpert people seems relevant.
@@
 An alternative explanation relies in reducing the recognition bias involved in the classic recognition paradigm.
-Classic forced-choice paradigms obtain artificially high results, thus by using a more evolved approach observers' accuracy may have been lowered.
+Classic forced-choice paradigms obtain artificially high results, thus by using a more evolved approach, observers' accuracy may have been lowered.
 Another flaw is the lack of comparison with various facial expression recognition methods.
@@
 However, it is known that encoders' imagination can influence their expressiveness, for instance when they believe that their friend is doing the same (\emph{vs.} a different) emotional task in another room (Jakobs, Manstead, \& Fischer, 1999).
-Trying to exclude social influence by leaving encoders alone may be illusionary then.
+Trying to exclude social influence by leaving encoders alone may be illusionary.
 From an ecological perspective, it is even a mistake to exonerate behavioral observations from social contexts.
@@
 It is the case, for instance, of emotional mimicry in dynamic social interactions (Hess \& Fischer, 2016).
-It is also the case of healthy partners interacting with schizophrenic patients whose facial activity is pretty much identical (Krause, Steimer-Krause, Merten, \& Ullrich, 1998).
+It is also the case of healthy patients interacting with schizophrenic patients whose facial activity is almost identical (Krause, Steimer-Krause, Merten, \& Ullrich, 1998).
 Hence, in order to better comprehend emotional communication in human relationships, experimental research should be corroborated with more ecological protocols.
@@
 They bring up several questions regarding the role the context plays in the emission and interpretation of the so-called facial expression of emotion.
-The finalization of operational and effective \enquote{reading emotional faces} devices rests on their answer.
-As a result, despite being one of the most investigated questions in affective science, the growing interest of industries and governments in tracking individual's psychological states is hardly given satisfaction.
-Considering the above, the present results provide additional evidence that an individual's subjective feeling cannot be inferred from facial expressions and in our case invalidate the hypothesis of hardwired emotions unambiguously displayed on the face.
-Even if emotions were hardwired, in everyday life one does not observe prototypical facial expressions because of their rarity and therefore research should be focused on analyzing non-prototypical facial expressions.
-Advancements in identifying \enquote{non-basic} emotion labels as well as non-prototypical facial expression have indeed been made in the development of automatic facial expression recognition tools (McDuff, 2016).
-However, these present results suggest that automatic facial expression recognition tools should merely evaluate facial morphology features such as action units (already evaluated in OpenFace, Baltrušaitis, Robinson, \& Morency, 2016, Affectiva's Affdex, McDuff et al.
-(2016), or Vicar Vision's FaceReader, Den Uyl and Van Kuilenburg (2005), to name a few) rather than inferring supposedly emotional or affective states.
-Trying to interpret facial displays as a means of determining underlying emotional state, in all likelihood, remains vain.
+The finalization of operational and effective \enquote{reading emotional faces} devices rests on the answers, if any, to the questions raised.
+As a result, despite being one of the most investigated questions in affective science, the growing interest of industries and governments in tracking individual's psychological states is supported by controversial assumptions.
+Considering the above, the present results provide additional evidence that an individual's subjective feelings cannot be inferred from facial expressions, and they invalidate the hypothesis of hardwired emotions unambiguously displayed on the face.
+Even if emotions were hardwired, in everyday life one does not observe prototypical facial expressions because of their rarity, and therefore research should be focused on analyzing non-prototypical facial expressions.
+Advancements in identifying \enquote{non-basic} emotion labels as well as non-prototypical facial expression have indeed occured with automatic facial expression recognition tools (McDuff, 2016).
+However, these present results suggest that automatic facial expression recognition tools should merely evaluate facial morphology features such as action units (already evaluated in OpenFace, Baltrušaitis, Robinson, \& Morency, 2016; Affectiva's Affdex, McDuff et al., 2016; Vicar Vision's FaceReader, Den Uyl \& Van Kuilenburg, 2005, to name a few) rather than inferring supposedly emotional or affective states.
+Trying to interpret facial displays as a means of determining underlying emotional state, in all likelihood, remains in vain.
 
@@
 \leavevmode\hypertarget{ref-aviezer2017inherently}{}%
-Aviezer, H., Ensenberg, N., \& Hassin, R. R. (2017). The inherently contextualized nature of facial emotion perception. \emph{Current Opinion in Psychology}, \emph{17}, 47--54.
+Aviezer, H., Ensenberg, N., \& Hassin, R.
+R.
+(2017).
+The inherently contextualized nature of facial emotion perception.
+\emph{Current Opinion in Psychology}, \emph{17}, 47--54.
+\url{https://doi.org/10.1016/j.copsyc.2017.06.006}
 
@@
 \emph{Psychological Science}, \emph{19}(7), 724--732.
+\url{https://doi.org/10.1111/j.1467-9280.2008.02148.x}
 
@@
 Openface: An open source facial behavior analysis toolkit.
-In \emph{2016 ieee winter conference on applications of computer vision (wacv)} (pp.
+In \emph{Winter conference on applications of computer vision} (pp.
 1--10).
-IEEE.
+\url{https://doi.org/10.1109/WACV.2016.7477553}
 
 \leavevmode\hypertarget{ref-barrett2006emotions}{}%
-Barrett, L. F. (2006). Are emotions natural kinds? \emph{Perspectives on Psychological Science}, \emph{1}(1), 28--58.
+Barrett, L. F. (2006). Are emotions natural kinds? \emph{Perspectives on Psychological Science}, \emph{1}(1), 28--58. \url{https://doi.org/10.1111/j.1745-6916.2006.00003.x}
 
@@
 \emph{Psychological Science in the Public Interest}, \emph{20}(1), 1--68.
+\url{https://doi.org/10.1177/1529100619832930}
 
 \leavevmode\hypertarget{ref-barrett2011context}{}%
-Barrett, L. F., Mesquita, B., \& Gendron, M. (2011). Context in emotion perception. \emph{Current Directions in Psychological Science}, \emph{20}(5), 286--290.
+Barrett, L.
+F., Mesquita, B., \& Gendron, M.
+(2011).
+Context in emotion perception.
+\emph{Current Directions in Psychological Science}, \emph{20}(5), 286--290.
+\url{https://doi.org/10.1177/0963721411422522}
 
 \leavevmode\hypertarget{ref-bennett2002facial}{}%
-Bennett, D. S., Bendersky, M., \& Lewis, M. (2002). Facial expressivity at 4 months: A context by expression analysis. \emph{Infancy}, \emph{3}(1), 97--113.
+Bennett, D.
+S., Bendersky, M., \& Lewis, M.
+(2002).
+Facial expressivity at 4 months: A context by expression analysis.
+\emph{Infancy}, \emph{3}(1), 97--113.
+\url{https://doi.org/10.1207/S15327078IN0301_5}
 
@@
 \leavevmode\hypertarget{ref-buck1985prime}{}%
-Buck, R. (1985). Prime theory: An integrated view of motivation and emotion. \emph{Psychological Review}, \emph{92}(3), 389.
+Buck, R. (1985). Prime theory: An integrated view of motivation and emotion. \emph{Psychological Review}, \emph{92}(3), 389--413. \url{https://doi.org/10.1037/0033-295X.92.3.389}
 
 \leavevmode\hypertarget{ref-camras1991development}{}%
-Camras, L. A., Malatesta, C., \& Izard, C. E. (1991). The development of facial expressions in infancy.
+Camras, L.
+A., Malatesta, C., \& Izard, C.
+E.
+(1991).
+The development of facial expressions in infancy.
+In R.
+S.
+Feldman \& B.
+Rimé (Eds.), \emph{Studies in emotion \& social interaction.
+Fundamentals of nonverbal behavior} (pp.
+31--54).
+New York, NY: Cambridge University Press.
 
@@
 \emph{Infancy}, \emph{11}(2), 131--155.
+\url{https://doi.org/10.1111/j.1532-7078.2007.tb00219.x}
 
 \leavevmode\hypertarget{ref-cohn2003timing}{}%
-Cohn, J. F., \& Schmidt, K. (2003). The timing of facial motion in posed and spontaneous smiles. In \emph{Active media technology} (pp. 57--69). World Scientific.
+Cohn, J.
+F., \& Schmidt, K.
+(2004).
+The timing of facial motion in posed and spontaneous smiles.
+\emph{International Journal of Wavelets, Multiresolution and Information Processing}, \emph{2}(2), 121--132.
+\url{https://doi.org/10.1142/S021969130400041X}
 
 \leavevmode\hypertarget{ref-crivelli2015smiles}{}%
-Crivelli, C., Carrera, P., \& Fernández-Dols, J.-M. (2015). Are smiles a sign of happiness? Spontaneous expressions of judo winners. \emph{Evolution and Human Behavior}, \emph{36}(1), 52--58.
+Crivelli, C., Carrera, P., \& Fernández-Dols, J.-M.
+(2015).
+Are smiles a sign of happiness?
+Spontaneous expressions of judo winners.
+\emph{Evolution and Human Behavior}, \emph{36}(1), 52--58.
+\url{https://doi.org/10.1016/j.evolhumbehav.2014.08.009}
 
@@
 Recognizing spontaneous facial expressions of emotion in a small-scale society of papua new guinea.
-\emph{Emotion}, \emph{17}(2), 337.
+\emph{Emotion}, \emph{17}(2), 337--347.
+\url{https://doi.org/10.1037/emo0000236}
 
@@
 \leavevmode\hypertarget{ref-den2005facereader}{}%
-Den Uyl, M., \& Van Kuilenburg, H. (2005). The facereader: Online facial expression recognition. In \emph{Proceedings of measuring behavior} (Vol. 30, pp. 589--590). Citeseer.
+Den Uyl, M., \& Van Kuilenburg, H. (2005). The facereader: Online facial expression recognition. In \emph{Proceedings of measuring behavior} (pp. 589--590).
 
 \leavevmode\hypertarget{ref-digirolamo2017emotion}{}%
-DiGirolamo, M. A., \& Russell, J. A. (2017). The emotion seen in a face can be a methodological artifact: The process of elimination hypothesis. \emph{Emotion}, \emph{17}(3), 538.
+DiGirolamo, M.
+A., \& Russell, J.
+A.
+(2017).
+The emotion seen in a face can be a methodological artifact: The process of elimination hypothesis.
+\emph{Emotion}, \emph{17}(3), 538--546.
+\url{https://doi.org/10.1037/emo0000247}
 
@@
 \leavevmode\hypertarget{ref-duchenne1876mecanisme}{}%
-Duchenne, G.-B. (1876). \emph{Mécanisme de la physionomie humaine: Où, analyse électro-physiologique de l'expression des passions}. J.-B. Baillière.
+Duchenne, G.-B. (1876). \emph{Mécanisme de la physionomie humaine: Où, analyse électro-physiologique de l'expression des passions}. Paris, France: J.-B. Baillière.
 
@@
 \leavevmode\hypertarget{ref-duran2017coherence}{}%
-Durán, J. I., Reisenzein, R., \& Fernández-Dols, J.-M. (2017). Coherence between emotions and facial expressions, 107--129.
+Durán, J.
+I., Reisenzein, R., \& Fernández-Dols, J.-M.
+(2017).
+Coherence between emotions and facial expressions.
+In J.-M.
+Fernández-Dols \& J.
+A.
+Russell (Eds.), \emph{The science of facial expression} (pp.
+107--129).
+New York, NY: Oxford University Press.
 
 \leavevmode\hypertarget{ref-ekman1992argument}{}%
-Ekman, P. (1992). An argument for basic emotions. \emph{Cognition \& Emotion}, \emph{6}(3-4), 169--200. \url{https://doi.org/10.1080/02699939208411068}
+Ekman, P. (1992). An argument for basic emotions. \emph{Cognition and Emotion}, \emph{6}(3-4), 169--200. \url{https://doi.org/10.1080/02699939208411068}
 
@@
 \leavevmode\hypertarget{ref-ekman2011meant}{}%
-Ekman, P., \& Cordaro, D. (2011). What is meant by calling emotions basic. \emph{Emotion Review}, \emph{3}(4), 364--370.
+Ekman, P., \& Cordaro, D. (2011). What is meant by calling emotions basic. \emph{Emotion Review}, \emph{3}(4), 364--370. \url{https://doi.org/10.1177/1754073911410740}
 
 \leavevmode\hypertarget{ref-ekman1980facial}{}%
-Ekman, P., Freisen, W. V., \& Ancoli, S. (1980). Facial signs of emotional experience. \emph{Journal of Personality and Social Psychology}, \emph{39}(6), 1125.
+Ekman, P., Freisen, W.
+V., \& Ancoli, S.
+(1980).
+Facial signs of emotional experience.
+\emph{Journal of Personality and Social Psychology}, \emph{39}(6), 1125--1134.
+\url{https://doi.org/10.1037/h0077722}
 
@@
 \leavevmode\hypertarget{ref-elfenbein2002universality}{}%
-Elfenbein, H. A., \& Ambady, N. (2002). On the universality and cultural specificity of emotion recognition: A meta-analysis. \emph{Psychological Bulletin}, \emph{128}(2), 203.
+Elfenbein, H.
+A., \& Ambady, N.
+(2002).
+On the universality and cultural specificity of emotion recognition: A meta-analysis.
+\emph{Psychological Bulletin}, \emph{128}(2), 203--235.
+\url{https://doi.org/10.1037/0033-2909.128.2.203}
 
@@
 \emph{Journal of Nonverbal Behavior}, \emph{21}(3), 163--177.
+\url{https://doi.org/10.1023/A:1024917530100}
 
@@
 \leavevmode\hypertarget{ref-fernandez1995smiles}{}%
-Fernández-Dols, J.-M., \& Ruiz-Belda, M.-A. (1995). Are smiles a sign of happiness? Gold medal winners at the olympic games. \emph{Journal of Personality and Social Psychology}, \emph{69}(6), 1113.
+Fernández-Dols, J.-M., \& Ruiz-Belda, M.-A.
+(1995).
+Are smiles a sign of happiness?
+Gold medal winners at the olympic games.
+\emph{Journal of Personality and Social Psychology}, \emph{69}(6), 1113--1119.
+\url{https://doi.org/10.1037/0022-3514.69.6.1113}
 
@@
 Frijda, N.
+H.
 (2012).
@@
 \leavevmode\hypertarget{ref-gendron2018emotion}{}%
-Gendron, M., Hoemann, K., Crittenden, A. N., Msafiri, S., Ruark, G. A., \& Barrett, L. F. (2018). Emotion perception in hadza hunter gatherers.
+Gendron, M., Hoemann, K., Crittenden, A. N., Msafiri, S., Ruark, G. A., \& Barrett, L. F. (2018). Emotion perception in hadza hunter gatherers. \url{https://doi.org/10.31234/osf.io/pf2q3}
 
@@
 Perceptions of emotion from facial expressions are not culturally universal: Evidence from a remote culture.
-\emph{Emotion}, \emph{14}(2), 251.
+\emph{Emotion}, \emph{14}(2), 251--262.
+\url{https://doi.org/10.1037/a0036052}
 
 \leavevmode\hypertarget{ref-hassin2013inherently}{}%
-Hassin, R. R., Aviezer, H., \& Bentin, S. (2013). Inherently ambiguous: Facial expressions of emotions, in context. \emph{Emotion Review}, \emph{5}(1), 60--65.
+Hassin, R.
+R., Aviezer, H., \& Bentin, S.
+(2013).
+Inherently ambiguous: Facial expressions of emotions, in context.
+\emph{Emotion Review}, \emph{5}(1), 60--65.
+\url{https://doi.org/10.1177/1754073912451331}
 
 \leavevmode\hypertarget{ref-hess2016emotional}{}%
-Hess, U., \& Fischer, A. H. (2016). \emph{Emotional mimicry in social context}. Cambridge University Press.
+Hess, U., \& Fischer, A. H. (2016). \emph{Emotional mimicry in social context}. New York, NY: Cambridge University Press.
 
 \leavevmode\hypertarget{ref-hess1990differentiating}{}%
-Hess, U., \& Kleck, R. E. (1990). Differentiating emotion elicited and deliberate emotional facial expressions. \emph{European Journal of Social Psychology}, \emph{20}(5), 369--385.
+Hess, U., \& Kleck, R.
+E.
+(1990).
+Differentiating emotion elicited and deliberate emotional facial expressions.
+\emph{European Journal of Social Psychology}, \emph{20}(5), 369--385.
+\url{https://doi.org/10.1002/ejsp.2420200502}
 
@@
 \emph{Personality and Social Psychology Bulletin}, \emph{25}(4), 424--435.
+\url{https://doi.org/10.1177/0146167299025004003}
 
 \leavevmode\hypertarget{ref-johnson2010smile}{}%
-Johnson, K. J., Waugh, C. E., \& Fredrickson, B. L. (2010). Smile to see the forest: Facially expressed positive emotions broaden cognition. \emph{Cognition and Emotion}, \emph{24}(2), 299--321.
+Johnson, K.
+J., Waugh, C.
+E., \& Fredrickson, B.
+L.
+(2010).
+Smile to see the forest: Facially expressed positive emotions broaden cognition.
+\emph{Cognition and Emotion}, \emph{24}(2), 299--321.
+\url{https://doi.org/10.1080/02699930903384667}
 
 \leavevmode\hypertarget{ref-kamachi2013dynamic}{}%
-Kamachi, M., Bruce, V., Mukaida, S., Gyoba, J., Yoshikawa, S., \& Akamatsu, S. (2013). Dynamic properties influence the perception of facial expressions. \emph{Perception}, \emph{42}(11), 1266--1278.
+Kamachi, M., Bruce, V., Mukaida, S., Gyoba, J., Yoshikawa, S., \& Akamatsu, S.
+(2013).
+Dynamic properties influence the perception of facial expressions.
+\emph{Perception}, \emph{42}(11), 1266--1278.
+\url{https://doi.org/10.1068/p3131n}
 
@@
 Comprehensive database for facial expression analysis.
-In \emph{Proceedings fourth ieee international conference on automatic face and gesture recognition (cat.
-No.
-PR00580)} (pp.
+In \emph{Proceedings fourth international conference on automatic face and gesture recognition} (pp.
 46--53).
-IEEE.
 
@@
 \leavevmode\hypertarget{ref-krause1998dyadic}{}%
-Krause, R., Steimer-Krause, E., Merten, J., \& Ullrich, B. (1998). Dyadic interaction regulation, emotion, and psychopathology. \emph{Emotions and Psychopathology: Theory and Research}, 70--80.
+Krause, R., Steimer-Krause, E., Merten, J., \& Ullrich, B.
+(1998).
+Dyadic interaction regulation, emotion, and psychopathology.
+In W.
+F.
+J.
+Flack \& J.
+D.
+Laird (Eds.), \emph{Emotions and psychopathology: Theory and research} (pp.
+70--80).
+New York, NY: Oxford University Press.
 
 \leavevmode\hypertarget{ref-kraut1979social}{}%
-Kraut, R. E., \& Johnston, R. E. (1979). Social and emotional messages of smiling: An ethological approach. \emph{Journal of Personality and Social Psychology}, \emph{37}(9), 1539--1553.
+Kraut, R.
+E., \& Johnston, R.
+E.
+(1979).
+Social and emotional messages of smiling: An ethological approach.
+\emph{Journal of Personality and Social Psychology}, \emph{37}(9), 1539--1553.
+\url{https://doi.org/10.1037/0022-3514.37.9.1539}
 
 \leavevmode\hypertarget{ref-krumhuber2013effects}{}%
-Krumhuber, E. G., Kappas, A., \& Manstead, A. S. (2013). Effects of dynamic aspects of facial expressions: A review. \emph{Emotion Review}, \emph{5}(1), 41--46.
+Krumhuber, E.
+G., Kappas, A., \& Manstead, A.
+S.
+R.
+(2013).
+Effects of dynamic aspects of facial expressions: A review.
+\emph{Emotion Review}, \emph{5}(1), 41--46.
+\url{https://doi.org/10.1177/1754073912451349}
 
@@
 \emph{Journal of Nonverbal Behavior}, \emph{26}(1), 3--25.
+\url{https://doi.org/10.1023/A:1014479919684}
 
 \leavevmode\hypertarget{ref-lindquist2006language}{}%
-Lindquist, K. A., Barrett, L. F., Bliss-Moreau, E., \& Russell, J. A. (2006). Language and the perception of emotion. \emph{Emotion}, \emph{6}(1), 125.
+Lindquist, K.
+A., Barrett, L.
+F., Bliss-Moreau, E., \& Russell, J.
+A.
+(2006).
+Language and the perception of emotion.
+\emph{Emotion}, \emph{6}(1), 125--138.
+\url{https://doi.org/10.1037/1528-3542.6.1.125}
 
@@
 AFFDEX sdk: A cross-platform real-time multi-face expression recognition toolkit.
-In \emph{Proceedings of the chi conference on human factors in computing systems} (pp.
+In \emph{Proceedings of the conference on human factors in computing systems} (pp.
 3723--3726).
@@
 \leavevmode\hypertarget{ref-mehu2007smiles}{}%
-Mehu, M., Grammer, K., \& Dunbar, R. I. (2007). Smiles when sharing. \emph{Evolution and Human Behavior}, \emph{28}(6), 415--422.
+Mehu, M., Grammer, K., \& Dunbar, R. I. (2007). Smiles when sharing. \emph{Evolution and Human Behavior}, \emph{28}(6), 415--422. \url{https://doi.org/10.1016/j.evolhumbehav.2007.05.010}
 
@@
 \leavevmode\hypertarget{ref-nielsen2007conceptual}{}%
-Nielsen, L., \& Kaszniak, A. W. (2007). Conceptual, theoretical, and methodological issues in inferring subjective emotion experience. \emph{Handbook of Emotion Elicitation and Assessment}, 361--375.
+Nielsen, L., \& Kaszniak, A.
+W.
+(2007).
+Conceptual, theoretical, and methodological issues in inferring subjective emotion experience.
+In J.
+A.
+Coan \& J.
+J.
+B.
+Allen (Eds.), \emph{Handbook of emotion elicitation and assessment} (pp.
+361--375).
+New York, NY: Oxford University Press.
 
@@
 Modes de structuration des contenus perceptifs visuels.
-\emph{J.
+In J.
 Bouveresse \& J.-J.
-Rosat (Eds), Philosophies de La Perception.
-Phénoménologie, Grammaire et Sciences Cognitives, Paris: Odile Jacob}, 263--289.
+Rosat (Eds.), \emph{Philosophies de la perception.
+Phénoménologie, grammaire et sciences cognitives} (pp.
+263--289).
+Paris, France: Odile Jacob.
 
 \leavevmode\hypertarget{ref-reisenzein2000exploring}{}%
-Reisenzein, R. (2000). Exploring the strength of association between the components of emotion syndromes: The case of surprise. \emph{Cognition \& Emotion}, \emph{14}(1), 1--38.
+Reisenzein, R.
+(2000).
+Exploring the strength of association between the components of emotion syndromes: The case of surprise.
+\emph{Cognition and Emotion}, \emph{14}(1), 1--38.
+\url{https://doi.org/10.1080/026999300378978}
 
@@
 Evidence for strong dissociation between emotion and facial displays: The case of surprise.
-\emph{Journal of Personality and Social Psychology}, \emph{91}(2), 295.
+\emph{Journal of Personality and Social Psychology}, \emph{91}(2), 295--315.
+\url{https://doi.org/10.1037/0022-3514.91.2.295}
 
 \leavevmode\hypertarget{ref-reisenzein2013coherence}{}%
-Reisenzein, R., Studtmann, M., \& Horstmann, G. (2013). Coherence between emotion and facial expression: Evidence from laboratory experiments. \emph{Emotion Review}, \emph{5}(1), 16--23.
+Reisenzein, R., Studtmann, M., \& Horstmann, G.
+(2013).
+Coherence between emotion and facial expression: Evidence from laboratory experiments.
+\emph{Emotion Review}, \emph{5}(1), 16--23.
+\url{https://doi.org/10.1177/1754073912457228}
 
 \leavevmode\hypertarget{ref-rosenberg1994coherence}{}%
-Rosenberg, E. L., \& Ekman, P. (1994). Coherence between expressive and experiential systems in emotion. \emph{Cognition \& Emotion}, \emph{8}(3), 201--229.
+Rosenberg, E.
+L., \& Ekman, P.
+(1994).
+Coherence between expressive and experiential systems in emotion.
+\emph{Cognition and Emotion}, \emph{8}(3), 201--229.
+\url{https://doi.org/10.1080/02699939408408938}
 
 \leavevmode\hypertarget{ref-ruiz2003spontaneous}{}%
-Ruiz-Belda, M.-A., Fernández-Dols, J.-M., Carrera, P., \& Barchard, K. (2003). Spontaneous facial expressions of happy bowlers and soccer fans. \emph{Cognition and Emotion}, \emph{17}(2), 315--326.
+Ruiz-Belda, M.-A., Fernández-Dols, J.-M., Carrera, P., \& Barchard, K.
+(2003).
+Spontaneous facial expressions of happy bowlers and soccer fans.
+\emph{Cognition and Emotion}, \emph{17}(2), 315--326.
+\url{https://doi.org/10.1080/02699930302288}
 
@@
 \leavevmode\hypertarget{ref-russell1994there}{}%
-Russell, J. A. (1994). Is there universal recognition of emotion from facial expression? A review of the cross-cultural studies. \emph{Psychological Bulletin}, \emph{115}(1), 102.
+Russell, J.
+A.
+(1994).
+Is there universal recognition of emotion from facial expression?
+A review of the cross-cultural studies.
+\emph{Psychological Bulletin}, \emph{115}(1), 102--141.
+\url{https://doi.org/10.1037/0033-2909.115.1.102}
 
 \leavevmode\hypertarget{ref-russell2003facial}{}%
-Russell, J. A., Bachorowski, J.-A., \& Fernández-Dols, J.-M. (2003). Facial and vocal expressions of emotion. \emph{Annual Review of Psychology}, \emph{54}(1), 329--349.
+Russell, J.
+A., Bachorowski, J.-A., \& Fernández-Dols, J.-M.
+(2003).
+Facial and vocal expressions of emotion.
+\emph{Annual Review of Psychology}, \emph{54}(1), 329--349.
+\url{https://doi.org/10.1146/annurev.psych.54.101601.145102}
 
 \leavevmode\hypertarget{ref-russell1987relativity}{}%
-Russell, J. A., \& Fehr, B. (1987). Relativity in the perception of emotion in facial expressions. \emph{Journal of Experimental Psychology: General}, \emph{116}(3), 223.
+Russell, J.
+A., \& Fehr, B.
+(1987).
+Relativity in the perception of emotion in facial expressions.
+\emph{Journal of Experimental Psychology: General}, \emph{116}(3), 223--237.
+\url{https://doi.org/10.1037/0096-3445.116.3.223}
 
 \leavevmode\hypertarget{ref-sander2014oxford}{}%
-Sander, D., \& Scherer, K. (2014). \emph{Oxford companion to emotion and the affective sciences}. OUP Oxford.
+Sander, D., \& Scherer, K. (2014). \emph{Oxford companion to emotion and the affective sciences}. New York, NY: Oxford University Press.
 
 \leavevmode\hypertarget{ref-scherer1997lost}{}%
-Scherer, K. R., \& Ceschi, G. (1997). Lost luggage: A field study of emotion--antecedent appraisal. \emph{Motivation and Emotion}, \emph{21}(3), 211--235.
+Scherer, K. R., \& Ceschi, G. (1997). Lost luggage: A field study of emotion--antecedent appraisal. \emph{Motivation and Emotion}, \emph{21}(3), 211--235. \url{https://doi.org/10.1023/A:1024498629430}
 
@@
 \emph{Evolution and Human Behavior}, \emph{33}(6), 657--664.
+\url{https://doi.org/10.1016/j.evolhumbehav.2012.04.003}
 
@@
 \emph{European Journal of Social Psychology}, \emph{37}(6), 1325--1345.
+\url{https://doi.org/10.1002/ejsp.427}
 
@@
 \leavevmode\hypertarget{ref-tcherkassof2014emotions}{}%
-Tcherkassof, A., \& Frijda, N. H. (2014). Les émotions: Une conception relationnelle. \emph{LAnnee Psychologique}, \emph{114}(3), 501--535.
+Tcherkassof, A., \& Frijda, N. H. (2014). Les émotions: Une conception relationnelle. \emph{L'Année Psychologique}, \emph{114}(3), 501--535.
 
 \leavevmode\hypertarget{ref-tsai2002emotional}{}%
-Tsai, J. L., Chentsova-Dutton, Y., Freire-Bebeau, L., \& Przymus, D. E. (2002). Emotional expression and physiology in european americans and hmong americans. \emph{Emotion}, \emph{2}(4), 380.
+Tsai, J.
+L., Chentsova-Dutton, Y., Freire-Bebeau, L., \& Przymus, D.
+E.
+(2002).
+Emotional expression and physiology in european americans and hmong americans.
+\emph{Emotion}, \emph{2}(4), 380--397.
+\url{https://doi.org/10.1037/1528-3542.2.4.380}
 
 \leavevmode\hypertarget{ref-vernon2002disgust}{}%
-Vernon, L. L., \& Berenbaum, H. (2002). Disgust and fear in response to spiders. \emph{Cognition \& Emotion}, \emph{16}(6), 809--830.
+Vernon, L. L., \& Berenbaum, H. (2002). Disgust and fear in response to spiders. \emph{Cognition and Emotion}, \emph{16}(6), 809--830. \url{https://doi.org/10.1080/02699930143000464}
 
@@
 \leavevmode\hypertarget{ref-wagner1986communication}{}%
-Wagner, H., MacDonald, C., \& Manstead, A. (1986). Communication of individual emotions by spontaneous facial expressions. \emph{Journal of Personality and Social Psychology}, \emph{50}(4), 737.
+Wagner, H.
+L., MacDonald, C.
+J., \& Manstead, A.
+S.
+(1986).
+Communication of individual emotions by spontaneous facial expressions.
+\emph{Journal of Personality and Social Psychology}, \emph{50}(4), 737--743.
+\url{https://doi.org/10.1037/0022-3514.50.4.737}
 
 \leavevmode\hypertarget{ref-aviezer2017inherently}{}%
-Aviezer, H., Ensenberg, N., \& Hassin, R. R. (2017). The inherently contextualized nature of facial emotion perception. \emph{Current Opinion in Psychology}, \emph{17}, 47--54.
+Aviezer, H., Ensenberg, N., \& Hassin, R.
+R.
+(2017).
+The inherently contextualized nature of facial emotion perception.
+\emph{Current Opinion in Psychology}, \emph{17}, 47--54.
+\url{https://doi.org/10.1016/j.copsyc.2017.06.006}
 
@@
 \leavevmode\hypertarget{ref-wagner1986communication}{}%
-Wagner, H., MacDonald, C., \& Manstead, A. (1986). Communication of individual emotions by spontaneous facial expressions. \emph{Journal of Personality and Social Psychology}, \emph{50}(4), 737.
+Wagner, H.
+L., MacDonald, C.
+J., \& Manstead, A.
+S.
+(1986).
+Communication of individual emotions by spontaneous facial expressions.
+\emph{Journal of Personality and Social Psychology}, \emph{50}(4), 737--743.
+\url{https://doi.org/10.1037/0022-3514.50.4.737}
 
@@
 
+\clearpage
+
+\renewcommand{\listfigurename}{Figure captions}
+
+\listoffigures
+
+\clearpage
+
+\renewcommand{\listtablename}{Table captions}
+
+\listoftables
+
 \end{document}

### 2020-08-27 manuscript/manuscript_apa.Rmd
before revision

--- 2019-11-06 manuscript_apa.Rmd
@@
 title             : "The emotion–facial expression link: Evidence from human and automatic expression recognition"
-shorttitle        : "The emotion–facial expression link"
+shorttitle        : "The emotion facial expression link"
 
@@
 keywords          : "emotion, facial expression, self-report, human observer, automatic recognition."
-wordcount         : "7941"
+wordcount         : "7943"
 
-bibliography      : ["manuscript_ref.bib"]
+bibliography      : ["bib/manuscript_ref.bib"]
 
 floatsintext      : no
-figurelist        : yes
-tablelist         : yes
+figurelist        : no
+tablelist         : no
 footnotelist      : no
-linenumbers       : yes
+linenumbers       : no
 mask              : no
@@
 
-documentclass     : "apa6"
-classoption       : "man"
+documentclass     : "apa7"
+classoption       : "doc"
 
-output: papaja::apa6_word
+#output: papaja::apa6_word
 
-# output            :
-#   bookdown::pdf_book:
-#     toc: false
-#     number_sections: false
-#     base_format: papaja::apa6_pdf
+output            :
+  bookdown::pdf_book:
+    toc: false
+    number_sections: false
+    base_format: papaja::apa6_pdf
 
@@
 ```{r setup, include = FALSE}
-library(papaja)
-library(kableExtra) # Tables
 library(tidyverse) # Data wrangling & grammar of graphics
 library(here) # Locate files relative to project root
+library(papaja) # Manuscript design and results
+library(kableExtra) # Tables
 library(caret) # Confusion matrices
@@
   cache.extra = knitr::rand_seed,
+  cache = TRUE,
   message = FALSE, 
@@
 This technology rests on the premise that facial expressions provide direct access to individuals’ subjective feelings, and that one can read the emotions displayed on a person’s face as one would an open book.
-This premise is derived from the "Basic Emotion View", a theory which suggests a one-to-one mapping between subective feeling and facial expression [@ekman1987universals; @ekman1988universality; @ekman1992argument; @ekman2007directed].
+This premise is derived from the "Basic Emotion View", a theory which suggests a one-to-one mapping between subjective feeling and facial expression [@ekman1987universals; @ekman1988universality; @ekman1992argument; @ekman2007directed].
 Even if this premise is central to the modern mainstream approach of human emotion, recent research in affective science is challenging it.
@@
 self_report_gender <- self_report_data %>% 
-  dplyr::group_by(genre_c) %>% 
-  dplyr::summarise(n = n())
+  dplyr::count(genre_c) %>% 
+  dplyr::pull(n, genre_c)
 
 self_report_age <- self_report_data %>%
-  dplyr::summarise(Age_m = mean(age), Age_sd = sd(age)) %>% 
-  dplyr::mutate_all(round, 1)
+  dplyr::summarise(
+    age_m = mean(age) %>% round(1), 
+    age_sd = sd(age) %>% round(1)
+  )
 ```
 
-For the emotion elicitation experiment, `r nrow(self_report_data)` encoding participants (`r self_report_gender[self_report_gender$genre_c == "F","n"]` females, `r self_report_gender[self_report_gender$genre_c == "H","n"]` males, *M*~age~ = `r self_report_age[,"Age_m"]`, *SD*~age~ = `r self_report_age[,"Age_sd"]`) were recruited to perform one out of 11 emotion elicitation tasks designed to trigger a positive, a specific negative or a neutral emotional state.
+For the emotion elicitation experiment, `r nrow(self_report_data)` encoding participants (`r self_report_gender["F"]` females, `r self_report_gender["H"]` males, *M*~age~ = `r self_report_age["age_m"]`, *SD*~age~ = `r self_report_age["age_sd"]`) were recruited to perform one out of 11 emotion elicitation tasks designed to trigger a positive, a specific negative or a neutral emotional state.
 Encoders’ faces were recorded using a hidden camera resulting `r nrow(self_report_data)` front facing 768x576 videos varying from `r min(metadata_video$ffprobe_duration)`s to `r max(metadata_video$ffprobe_duration)`s.
@@
 unique_ppt_count <- human_recognition_data %>% 
-  dplyr::group_by(source,SEXE_Juge, C_Juge) %>% 
-  dplyr::summarise(n = n())
+  dplyr::count(source, SEXE_Juge, C_Juge)
 
 unique_video_count <- human_recognition_data %>% 
-  dplyr::select(source,SEXE_Juge, C_Juge, C_Video) %>% 
-  unique() %>% 
-  dplyr::group_by(C_Video) %>% 
-  dplyr::summarise(n = n())
+  dplyr::distinct(source, SEXE_Juge, C_Juge, C_Video) %>% 
+  dplyr::count(C_Video) %>% 
+  dplyr::summarise(
+    vid_n = n(),
+    vid_m = mean(n) %>% round(0), 
+    vid_sd = sd(n) %>% round(0)
+  )
 ```
 
-For the human facial expression recognition method, `r nrow(unique_ppt_count)` student participants were recruited to annotate `r nrow(unique_video_count)` out of the `r nrow(self_report_data)` videos, therefore only the `r nrow(unique_video_count)` annotated videos were analyzed in this study.
-Because videos have different durations, participants had to annotate a series of video corresponding to 30min long in total.
-Each video was annotated `r mean(unique_video_count$n) %>% round(0)` times on average (*SD* = `r sd(unique_video_count$n) %>% round(0)`).
+For the human facial expression recognition method, `r nrow(unique_ppt_count)` student participants were recruited to annotate `r unique_video_count["vid_n"]` out of the `r nrow(self_report_data)` videos, therefore only the `r unique_video_count["vid_n"]` annotated videos were analyzed in this study.
+Because videos have different duration, participants had to annotate a series of video corresponding to 30min long in total.
+Each video was annotated `r unique_video_count["vid_m"]` times on average (*SD* = `r unique_video_count["vid_sd"]`).
 
@@
 
-where *i* is one of the `r nrow(unique_video_count)` videos, *j* is one of the six “basic” emotion labels, *k* for each second of the video.
+where *i* is one of the `r unique_video_count["vid_n"]` videos, *j* is one of the six “basic” emotion labels, *k* for each second of the video.
 
@@
 
-The `r nrow(unique_video_count)` annotated video were processed with Affdex (SDK v3.4.1).
+The `r unique_video_count["vid_n"]` annotated video were processed with Affdex (SDK v3.4.1).
 Affdex is an automatic facial expression recognition classifier developed and distributed by Affectiva, which is a spin-off company resulting from the research activities of MIT media lab created in 2009 [@mcduff2016affdex].
@@
 
-where *i* is one of the `r nrow(unique_video_count)` videos, *j* is one of the six “basic” emotion labels, *k* for each second of the *n* second video or for each frame of the *f* frame video.
+where *i* is one of the `r unique_video_count["vid_n"]` videos, *j* is one of the six “basic” emotion labels, *k* for each second of the *n* second video or for each frame of the *f* frame video.
 
@@
 
-corr_sr_hr <- cor.test(
+corr_sr_hr <- 
+  cor.test(
     recognition_result$sr_score, 
@@
 
-corr_sr_ar <- cor.test(
+corr_sr_ar <- 
+  cor.test(
     recognition_result$sr_score, 
@@
 
-corr_hr_ar <- cor.test(
+corr_hr_ar <- 
+  cor.test(
     recognition_result$hr_score, 
@@
   papaja::apa_print()
-
-# sr_n <- recognition_result %>% 
-#   dplyr::filter(sr_score == 1) %>% 
-#   dplyr::group_by(list_emotion) %>% 
-#   dplyr::tally()
 ```
@@
 ```{r sr-hr}
-comparison_sr_hr <- dplyr::inner_join(self_report_score,human_recognition_score, by = "C_Video") %>% 
+comparison_sr_hr <- self_report_score %>% 
+  dplyr::inner_join(human_recognition_score, by = "C_Video") %>% 
   dplyr::ungroup() %>% 
@@
 
-confusionMatrix_sr_hr <- caret::confusionMatrix(
+confusionMatrix_sr_hr <- 
+  caret::confusionMatrix(
     data = comparison_sr_hr$hr_emotion,
@@
   geom_text(
-    aes(label = paste0(scales::percent(prop_sr), "\n(", Freq,"/",n_sr, ")")),
+    aes(label = paste0(
+      scales::percent(prop_sr, accuracy = 1), "\n(", Freq, "/", n_sr, ")"
+    )),
     color = "black",
     size = 4,
-    family="serif",
+    family = "serif",
     parse = FALSE,
-    lineheight = 0.7)+
+    lineheight = 0.7
+  ) +
   scale_fill_gradient(
@@
     labels = scales::percent,
-    limits = c(0,1)
+    limits = c(0, 1)
   ) +
@@
   theme_minimal() +
-  theme(text = element_text(size=16,family="serif"),
-        axis.text.x = element_text(size=16, angle = 45, hjust = 0.75,vjust=0.9),
-        axis.text.y = element_text(size=16, hjust = 0.5),
+  theme(
+    text = element_text(size = 16, family = "serif"),
+    axis.text.x = element_text(size = 16, angle = 45, hjust = 0.75, vjust = 0.9),
+    axis.text.y = element_text(size = 16, hjust = 0.5),
     axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
-        legend.text = element_text(size=8),
-        legend.position = "bottom")
+    legend.text = element_text(size = 8),
+    legend.position = "bottom"
+  )
 ```
@@
   tibble::rownames_to_column("Emotion") %>%
-  dplyr::mutate(Emotion = gsub(pattern = "Class: ",replacement = "",x = Emotion)) %>% 
+  dplyr::mutate(Emotion = gsub(pattern = "Class: ", replacement = "", x = Emotion)) %>% 
   dplyr::select(Emotion, Sensitivity, Specificity, Precision, F1) %>% 
-  dplyr::mutate_if(is.numeric,round,2) %>% 
-  # replace(is.na(.), "\\textit{na.}")
+  dplyr::mutate_if(is.numeric, round,2) %>% 
   replace(is.na(.), "na.")
@@
 confusionMatrix_sr_hr_table %>%
-  # knitr::kable(
-  #   "latex",
-  #   caption = "(ref:confusionTable-sr-hr)",
-  #   booktabs = TRUE,
-  #   digits = 2,
-  #   linesep = "",
-  #   escape = FALSE,
-  #   align = c('l', rep('c', 4)),
-  #   longtable = T
-  # ) %>%
-  # kableExtra::kable_styling(
-  #   latex_options = "HOLD_position",
-  #   font_size = 8,
-  #   full_width = TRUE 
-  # ) %>%
-  # kableExtra::footnote(
-  #   general_title = "",
-  #   general = "Note. na. values are produced when not enough data are available to compute accuracy indicators.", 
-  #   escape = FALSE,
-  #   footnote_as_chunk = TRUE,
-  #   threeparttable = TRUE
-  # )
   papaja::apa_table(
@@
 ```{r sr-ar}
-comparison_sr_ar <- dplyr::inner_join(self_report_score,automatic_recognition_score, by = "C_Video") %>% 
+comparison_sr_ar <- self_report_score %>% 
+  dplyr::inner_join(automatic_recognition_score, by = "C_Video") %>% 
   dplyr::ungroup() %>% 
@@
   dplyr::filter(sr_emotion != "undetermined") %>% 
-  dplyr::mutate_if(is.character,as.factor)
+  dplyr::mutate_if(is.character, as.factor)
 
-confusionMatrix_sr_ar <- caret::confusionMatrix(
+confusionMatrix_sr_ar <- 
+  caret::confusionMatrix(
     data = comparison_sr_ar$ar_emotion,
@@
   theme_minimal() +
-  theme(text = element_text(size=16,family="serif"),
-        axis.text.x = element_text(size=16, angle = 45, hjust = 0.75,vjust=0.9),
-        axis.text.y = element_text(size=16, hjust = 0.5),
-        axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
-        legend.text = element_text(size=8),
-        legend.position = "bottom")
+  theme(
+    text = element_text(size = 16, family = "serif"),
+    axis.text.x = element_text(size = 16, angle = 45, hjust = 0.75, vjust = 0.9 ),
+    axis.text.y = element_text(size = 16, hjust = 0.5),
+    axis.title.y = element_text(margin = margin( t = 0, r = 20, b = 0, l = 0)),
+    legend.text = element_text(size = 8),
+    legend.position = "bottom"
+  )
 ```
@@
   dplyr::mutate_if(is.numeric, round, 2) %>% 
-  #replace(is.na(.), "\\textit{na.}")
   replace(is.na(.), "na.")
@@
 confusionMatrix_sr_ar_table %>%
-  # knitr::kable(
-  #   "latex",
-  #   caption = "(ref:confusionTable-sr-ar)",
-  #   booktabs = T,
-  #   digits = 2,
-  #   linesep = "",
-  #   escape = FALSE,
-  #   align = c('l',rep('c', 4)),
-  #   longtable = T
-  # ) %>%
-  # kableExtra::kable_styling(
-  #   latex_options = c("HOLD_position"),
-  #   font_size = 8,
-  #   full_width = TRUE
-  #   ) %>%
-  # kableExtra::footnote(
-  #   general_title = "",
-  #   general = "Note. na. values are produced when not enough data are available to compute accuracy indicators.", 
-  #   escape = FALSE,
-  #   footnote_as_chunk = TRUE,
-  #   threeparttable = TRUE
-  #   ) 
   papaja::apa_table(
@@
 ```{r confusionMatrix-hr-ar, fig.height=6, out.width ="\\textwidth", fig.pos = "!h", fig.cap="(ref:confusionMatrix-hr-ar)"}
-comparison_hr_ar <- dplyr::inner_join(human_recognition_score,automatic_recognition_score, by = "C_Video") %>% 
+comparison_hr_ar <- human_recognition_score %>% 
+  dplyr::inner_join(automatic_recognition_score, by = "C_Video") %>% 
   dplyr::ungroup() %>% 
@@
 
-confusionMatrix_hr_ar <- caret::confusionMatrix(
+confusionMatrix_hr_ar <- 
+  caret::confusionMatrix(
     data = comparison_hr_ar$ar_emotion,
@@
   geom_text(
-    aes(label = paste0(scales::percent(prop_hr), "\n(", Freq, "/", n_hr, ")")),
+    aes(label = paste0(
+      scales::percent(prop_hr, accuracy = 1), "\n(", Freq, "/", n_hr, ")"
+    )),
     color = "black",
     size = 4,
-    family="serif",
+    family = "serif",
     parse = FALSE,
-    lineheight = 0.7)+
+    lineheight = 0.7
+  ) +
   scale_fill_gradient(
@@
   theme_minimal() +
-  theme(text = element_text(size=16,family="serif"),
-        axis.text.x = element_text(size=16, angle = 45, hjust = 0.75,vjust=0.9),
-        axis.text.y = element_text(size=16, hjust = 0.5),
-        axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
-        legend.text = element_text(size=8),
-        legend.position = "bottom")
+  theme(
+    text = element_text(size = 16, family = "serif"),
+    axis.text.x = element_text(size = 16, angle = 45, hjust = 0.75, vjust = 0.9),
+    axis.text.y = element_text(size = 16, hjust = 0.5),
+    axis.title.y = element_text(margin = margin(t = 0,r = 20, b = 0, l = 0)),
+    legend.text = element_text(size = 8),
+    legend.position = "bottom"
+  )
 ```
@@
 ```{r create_r-references}
-r_refs(file = "r-references.bib")
+#r_refs(file = "bib/r-references.bib")
 ```

### 2020-08-27 revision/prpf_1/revision_letter.Rmd
before revision

---
title       : "The emotion–facial expression link: Evidence from human and automatic expression recognition"
journal     : "Psychological Research"
manuscript  : "PRPF-D-20-00087"
bibliography: "../../manuscript/bib/manuscript_ref.bib"
output      : "papaja::revision_letter_pdf"
#output      : word_document
---

# Editor Comments

\RC{Dear Dr Tcherkassof,

We have received the reports from our advisers on your manuscript, "The emotion–facial expression link: Evidence from human and automatic expression recognition", which you submitted to Psychological Research.

Based on the advice received, I feel that your manuscript could be reconsidered for publication should you be prepared to incorporate major revisions. YOU ARE KINDLY REQUESTED TO ALSO CHECK THE WEBSITE FOR POSSIBLE REVIEWER ATTACHMENTS!

In order to submit your revised manuscript, please access the journal's website.

We look forward to receiving your revised manuscript before 25 Aug 2020.

With kind regards,
Ruth Krebs
Associate Editor
Psychological Research
}

Dear Dr. Ruth Krebs,

We would like thank you for taking the time to consider our manuscript for publication at _Psychological Research_, and the opportunity to resubmit a revised copy of this manuscript. We would also like to take this opportunity to express our thanks to the reviewers for the positive feedback and helpful comments for correction.

We believe such changes have resulted in an improved revised manuscript, which you will find uploaded alongside this document. The manuscript has been revised to address the reviewer comments, which are appended alongside our responses to this letter.

In the following we address the respect points made by both the editor and the reviewers.

# Reviewer \#1

## Main Points

\RC{The work of this article gives new insights about the connection of emotional self-reports and observer categorizations and automatic recognition of facial movements. The data are very interesting and this work should be published. But off course some things could be made better in the paper. The most important things are mainly in the methods and results section.}

## Method section

\RC{Page 15, line 57: in 2-3 or more lines should be described which kind of emotion inductions have been used for the data base.}

\RC{Page 16 line 39: why did the annotators only rate 232 of the 358 videos? And how have been the videos selected? Randomized selection or other criteria? Which? Why is not every video rated the same time? This especially is a point which looks not understandable.}

## Results section

\RC{Page 19 line 17 to 20: Which kind of correlation was computed here? A pearson correlation? It should be a point-biserial correlation as one of the two variables is dichotomous.}

\RC{Page 21 line 32: "expression intensity", where are expression intensity scores? Could you please bring the data into the paper?}

\RC{Page 21 line 40ff: Why should undetermined emotional states reveal that a 6-point Likert scale has a limit? What kind of limit? What would be better to do in future? Please explain in the text.}

## Minor points

\RC{Page 4 Line 44/45: one would not assume total agreement among all people based on basic emotion theory, so "totally" should be changed by another term like "mainly" or something similar.}

\RC{Page 5 Line 16/17: Isn´t the point of Kraut and Jonston that according Basic Emotion Theory the smile is only a sign of expression of joy? So I would write: "that a smile is only the major".}

\RC{Page 5 line 58 ff: maybe smiling is not the right sign to be looked directly after winning a contest. Triumpf is usually exposed immediately after the contest and other emotions like joy or pride come a little later (see the work of Matsumoto and colleagues on that). So this argument from Crivelli et al. (2015) does not hold in that context.}

\RC{Page 18: line 56: "two-sided pearson correlation" probably means two-tailed t-tests of pearson-correlations, or not?}

\RC{Page 20 Line25: "60% and 80% accuracy"; please give a citation.}

\RC{Page 23 Line 56: what is correlated here (r=.22)? A correlation of two accuracies does not make sense in my mind. If yes, what are the units? Emotion categories, events or…???}

\RC{Page 31 line 9ff: I do not understand the point under secondly, automatic classifiers assume for every person unique facial expression for one emotion?? Did I understand that right? Could you explain that more in detail. So maybe in the methods section a more detailled description of the algorithm and how it works, would help…; as this is a psychological journal a principal description of the characteristics of the algorithm would help (that does not mean to explai}

# Reviewer \#2

\RC{1. The paper did not distinguish clearly between opinion and empirical evidence. Please revise.}

\RC{2. The paper's arguments must be built on an appropriate base of theory, concepts, or other ideas. It is recommended that the authors refer to Eigen Face Approach. }

\RC{3. The authors did not make clear emotion taxonomy. Ekman's basic set of emotions and Russell's circumflex model of affect are recommended to elaborate.}

\RC{4.  What is the significance of this study? must be explained with a sensible reason.}

\RC{5. The paper must contribute to a critical understanding of the issues.}

\RC{6. You're highly advised to show yourselves as authors while explaining theme of research in the paper. Keep the available references and at the same time let's the readers identify your own ideas and voice concerning what you cite in Introduction. }

\RC{7. The selection of the participants, to me, is vague.}

\RC{8. I wonder why the authors did not implement Principal Component Analysis.}

\RC{9. Discussion needs a deeper look. The plausible reasons behind obtaining these results are not satisfactory.  Moreover, the author/s did not discuss the validity and accuracy of the data. They did not state what their study adds to the body of literature. The last point is that in the discussion, the authors must compare/contrast their results with similar studies, mentioned in literature.}

\RC{10. It is recommended that authors consider APA, 7th edition in in-text citations. }

# References

\setlength{\leftskip}{0.4in}
\setlength{\parindent}{-0.4in}

### 2020-08-27 script/data_wrangling_automatic_recognition.R

@@ -4,9 +4,16 @@
 
 # setup ------------------------------------------------------------------------
 library(tidyverse)
+library(here)
+
+# data path --------------------------------------------------------------------
+data_path <- here::here() %>% 
+  dirname() %>% 
+  dirname() %>% 
+  paste0("/data/dynemo_feel_data/automatic_recognition_data")
 
 # upload data ------------------------------------------------------------------
-automatic_recognition_data <- "../../data/dynemo_feel_data/automatic_recognition_data/automatic_recognition_data.rds" %>% 
+automatic_recognition_data <- file.path(data_path,  "automatic_recognition_data.rds") %>% 
   readr::read_rds()
 
 # data wrangling ---------------------------------------------------------------

### 2020-08-27 script/data_wrangling_human_recognition.R

@@ -5,12 +5,19 @@
 # setup ------------------------------------------------------------------------
 library(tidyverse)
 library(splitstackshape)
+library(here)
+
+# data path --------------------------------------------------------------------
+data_path <- here::here() %>% 
+  dirname() %>% 
+  dirname() %>% 
+  paste0("/data/dynemo_feel_data/human_recognition_data")
 
 # upload data ------------------------------------------------------------------
-metadata_video <- "../../data/dynemo_feel_data/human_recognition_data/metadata_video.rds" %>% 
+metadata_video <- file.path(data_path,  "metadata_video.rds") %>% 
   readr::read_rds()
 
-human_recognition_data <- "../../data/dynemo_feel_data/human_recognition_data" %>% 
+human_recognition_data <- data_path %>% 
   fs::dir_ls(regexp = "\\.csv$") %>% 
   purrr::map_dfr(readr::read_csv2, .id = "source", col_types = cols(
     C_INDUC = col_character(),

### 2020-08-27 script/data_wrangling_self_report.R

@@ -7,8 +7,14 @@ library(here)
 library(tidyverse)
 library(janitor)
 
+# data path --------------------------------------------------------------------
+data_path <- here::here() %>% 
+  dirname() %>% 
+  dirname() %>% 
+  paste0("/data/dynemo_feel_data/human_self_report_data")
+
 # upload data ------------------------------------------------------------------
-self_report_data <- "../../data/dynemo_feel_data/human_self_report_data/Base_Finale_20090114.rds" %>% 
+self_report_data <- file.path(data_path,  "Base_Finale_20090114.rds") %>% 
   readr::read_rds() %>% 
   janitor::clean_names() %>% 
   dplyr::mutate(C_Video = stringr::str_replace(fichier_video_v,".mpg","")) %>% 

### 2020-08-28 manuscript/manuscript_apa.Rmd
renv

@@
 (ref:confusionMatrix-sr-ar) Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the automatic classifier.
-(ref:confusionMatrix-hr-ar) Proportion of emotion labels classified by human observers which are recognized by the automatif classifier.
+(ref:confusionMatrix-hr-ar) Proportion of emotion labels classified by human observers which are recognized by the automatic classifier.
 (ref:confusionTable-sr-hr) Human recognition accuracy metrics for each emotion.
-(ref:confusionTable-sr-ar) Autonatic recognition accuracy metrics for each emotion.
+(ref:confusionTable-sr-ar) Automatic recognition accuracy metrics for each emotion.
 
@@
   cache.extra = knitr::rand_seed,
-  cache = TRUE,
+  cache = FALSE,
   message = FALSE, 

### 2020-08-28 manuscript/manuscript_apa.tex
renv

--- 2019-11-06 manuscript_apa.tex
@@
-\documentclass[man]{apa6}
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+  english,
+  doc]{apa7}
 \usepackage{lmodern}
@@
 \usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
 \ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
@@
   \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
 \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
   \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
-\usepackage{hyperref}
-\hypersetup{unicode=true,
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
   pdftitle={The emotion--facial expression link: Evidence from human and automatic expression recognition},
-            pdfauthor={Anna Tcherkassof~\& Damien Dupré},
+  pdfauthor={Anna Tcherkassof1 \& Damien Dupré2},
+  pdflang={en-EN},
   pdfkeywords={emotion, facial expression, self-report, human observer, automatic recognition.},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
 \usepackage{longtable,booktabs}
-\usepackage{graphicx,grffile}
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
 \makeatletter
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+% Make \paragraph and \subparagraph free-standing
 \ifx\paragraph\undefined\else
@@
 \fi
-
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-  \title{The emotion--facial expression link: Evidence from human and automatic expression recognition}
-    \author{Anna Tcherkassof\textsuperscript{1}~\& Damien Dupré\textsuperscript{2}}
-    \date{}
-
-\shorttitle{The emotion–facial expression link}
-\affiliation{
-\vspace{0.5cm}
-\textsuperscript{1} University Grenoble Alpes, Grenoble, France\\\textsuperscript{2} Dublin City University, Dublin, Ireland}
-\keywords{emotion, facial expression, self-report, human observer, automatic recognition.\newline\indent Word count: 7941}
-\usepackage{csquotes}
+% Manuscript styling
 \usepackage{upgreek}
@@
 
+% Table formatting
 \usepackage{longtable}
 \usepackage{lscape}
-\usepackage{multirow}
-\usepackage{tabularx}
-\usepackage[flushleft]{threeparttable}
-\usepackage{threeparttablex}
-
+% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
+\usepackage{multirow}		% Table styling
+\usepackage{tabularx}		% Control Column width
+\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
+\usepackage{threeparttablex}            % Lets threeparttable work with longtable
+
+% Create new environments so endfloat can handle them
+% \newenvironment{ltable}
+%   {\begin{landscape}\begin{center}\begin{threeparttable}}
+%   {\end{threeparttable}\end{center}\end{landscape}}
 \newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}
 
+% Enables adjusting longtable caption width to table width
+% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
 \makeatletter
@@
 
-\DeclareDelayedFloatFlavor{ThreePartTable}{table}
-\DeclareDelayedFloatFlavor{lltable}{table}
-\DeclareDelayedFloatFlavor*{longtable}{table}
+% \setlength{\parindent}{0.5in}
+% \setlength{\parskip}{0pt plus 0pt minus 0pt}
+
+% \usepackage{etoolbox}
 \makeatletter
-\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
+\patchcmd{\HyOrg@maketitle}
+  {\section{\normalfont\normalsize\abstractname}}
+  {\section*{\normalfont\normalsize\abstractname}}
+  {}{\typeout{Failed to patch abstract.}}
+\patchcmd{\HyOrg@maketitle}
+  {\section{\protect\normalfont{\@title}}}
+  {\section*{\protect\normalfont{\@title}}}
+  {}{\typeout{Failed to patch title.}}
 \makeatother
-\usepackage{lineno}
-
-\linenumbers
-\usepackage[titles]{tocloft}
-\cftpagenumbersoff{figure}
-\renewcommand{\cftfigpresnum}{\itshape\figurename\enspace}
-\renewcommand{\cftfigaftersnum}{.\space}
-\setlength{\cftfigindent}{0pt}
-\setlength{\cftafterloftitleskip}{0pt}
-\settowidth{\cftfignumwidth}{Figure 10.\qquad}
-\cftpagenumbersoff{table}
-\renewcommand{\cfttabpresnum}{\itshape\tablename\enspace}
-\renewcommand{\cfttabaftersnum}{.\space}
-\setlength{\cfttabindent}{0pt}
-\setlength{\cftafterloftitleskip}{0pt}
-\settowidth{\cfttabnumwidth}{Table 10.\qquad}
+\shorttitle{The emotion facial expression link}
+\keywords{emotion, facial expression, self-report, human observer, automatic recognition.\newline\indent Word count: 7943}
+\usepackage{csquotes}
 \usepackage{booktabs}
@@
 \usepackage{biblatex}
-
-\authornote{
-
-Correspondence concerning this article should be addressed to Anna Tcherkassof, Bâtiment Michel Dubois, Université Grenoble Alpes, 1251 Avenue Centrale, 38400 Saint-Martin-d'Hères.
-E-mail: \href{mailto:anna.tcherkassof@univ-grenoble-alpes.fr}{\nolinkurl{anna.tcherkassof@univ-grenoble-alpes.fr}}}
+\ifxetex
+  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
+  \usepackage{polyglossia}
+  \setmainlanguage[]{english}
+\else
+  \usepackage[shorthands=off,main=english]{babel}
+\fi
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newenvironment{cslreferences}%
+  {\setlength{\parindent}{0pt}%
+  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
+  {\par}
+
+\title{The emotion--facial expression link: Evidence from human and automatic expression recognition}
+\author{Anna Tcherkassof\textsuperscript{1} \& Damien Dupré\textsuperscript{2}}
+\date{}
+
+\affiliation{\vspace{0.5cm}\textsuperscript{1} University Grenoble Alpes, Grenoble, France\\\textsuperscript{2} Dublin City University, Dublin, Ireland}
 
@@
 Accordingly, automatic emotion recognition tools based on facial expressions should be questioned.
-
 }
@@
 This technology rests on the premise that facial expressions provide direct access to individuals' subjective feelings, and that one can read the emotions displayed on a person's face as one would an open book.
-This premise is derived from the \enquote{Basic Emotion View}, a theory which suggests a one-to-one mapping between subective feeling and facial expression (Ekman, 1992, 2007; Ekman et al., 1987; Ekman \& Heider, 1988).
+This premise is derived from the ``Basic Emotion View'', a theory which suggests a one-to-one mapping between subjective feeling and facial expression (Ekman, 1992, 2007; Ekman et al., 1987; Ekman \& Heider, 1988).
 Even if this premise is central to the modern mainstream approach of human emotion, recent research in affective science is challenging it.
@@
 More specifically, a set of six emotions (\emph{fear}, \emph{anger}, \emph{surprise}, \emph{disgust}, \emph{sadness} and \emph{joy}) are universally displayed and are genetically hardwired not only in humans (Ekman, 1992), but also in different animal species (Waal, 2019).
-According to the Basic Emotion View, \emph{\enquote{when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including {[}\ldots{}{]} facial expressions.}} (Ekman, 2007, p.
+According to the Basic Emotion View, \emph{``when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including {[}\ldots{]} facial expressions.''} (Ekman, 2007, p.
 49).
-To respond to criticisms, several amendments have been made to the Basic Emotion View, increasing the number of basic emotions from six to seven (Ekman \& Heider, 1988) as well as adding the concept of \enquote{display rules} to explain cultural differences in the management of facial expressions (Ekman et al., 1987).
-
-Relatively few studies have actually tested Basic Emotion View's fundamental claim regarding the facial production (so-called \enquote{expression}) of emotions.
+To respond to criticisms, several amendments have been made to the Basic Emotion View, increasing the number of basic emotions from six to seven (Ekman \& Heider, 1988) as well as adding the concept of ``display rules'' to explain cultural differences in the management of facial expressions (Ekman et al., 1987).
+
+Relatively few studies have actually tested Basic Emotion View's fundamental claim regarding the facial production (so-called ``expression'') of emotions.
 Rosenberg \& Ekman (1994) reported the first evidence of coherence between self-report of emotion and displayed facial expressions.
@@
 As people have been thought to display their internal state, EFE supposedly serve as a window into the emotions of others.
-Viewers \enquote{recognize} the facial displays and infer the corresponding emotional state.
+Viewers ``recognize'' the facial displays and infer the corresponding emotional state.
 The Readout Hypothesis (Buck, 1985) formalizes this folk theory.
-According to it, facial expressions have evolved to provide \emph{\enquote{an external readout of those motivational-emotional processes that have had social implications during the course of evolution}} (Buck, 1985, pp.
+According to it, facial expressions have evolved to provide \emph{``an external readout of those motivational-emotional processes that have had social implications during the course of evolution''} (Buck, 1985, pp.
 396--397).
@@
 Camras and her colleagues observed that facial expressions of negative emotions were displayed in circumstances that were unlikely to have elicited those emotions.
-For instance, infants displayed \enquote{fear} expressions in settings not related to fear (Camras, Malatesta, \& Izard, 1991).
+For instance, infants displayed ``fear'' expressions in settings not related to fear (Camras, Malatesta, \& Izard, 1991).
 Bennett, Bendersky, \& Lewis (2002) videotaped a sample of 4-month-old infants during tickle, sour taste, jack-in-the-box, arm restraint, and masked stranger situations.
@@
 It affirms that facial expressions do not provide a direct access to individuals' subjective feelings.
-Therefore, instead of considering that emotions can be \enquote{read} on facial displays, it claims that the emotion is \enquote{in the eye} of the perceiver (Barrett, Adolphs, Marsella, Martinez, \& Pollak, 2019).
+Therefore, instead of considering that emotions can be ``read'' on facial displays, it claims that the emotion is ``in the eye'' of the perceiver (Barrett, Adolphs, Marsella, Martinez, \& Pollak, 2019).
 
@@
 Starting from the empirical evidence suggesting that spontaneous facial expressions in ordinary life are equivocal, Dols (2017) argues in favor of a pragmatic conception of natural facial displays.
-He makes a plea for the idea that natural facial displays, rather than \enquote{saying}--because facial expressions do not have a specific meaning--\enquote{make} things.
+He makes a plea for the idea that natural facial displays, rather than ``saying''--because facial expressions do not have a specific meaning--``make'' things.
 Facial expressions are actions in a communicative interaction.
-They do not express emotions but they \emph{\enquote{prompt, on the receiver's side, important inferences about the context, the sender, and the course of the interaction between sender and receiver}} (Dols, 2017, p.
+They do not express emotions but they \emph{``prompt, on the receiver's side, important inferences about the context, the sender, and the course of the interaction between sender and receiver''} (Dols, 2017, p.
 466).
@@
 Indeed, no muscular contraction can produce it.
-He even rectifies the \enquote{mistake} by presenting a statue which face is shaped according to the physiology of facial expressive movements.
+He even rectifies the ``mistake'' by presenting a statue which face is shaped according to the physiology of facial expressive movements.
 His demonstration gives food for thought.
@@
 
-After the emotion elicitation task, the encoders rated their subjective feeling on Likert scales from 0 (\enquote{not at all}) to 5 (\enquote{strongly}) related to six \enquote{basic} emotion labels (\emph{i.e.}, \emph{anger}, \emph{disgust}, \emph{fear}, \emph{happiness}, \emph{surprise} and \emph{sadness}) as well as six \enquote{non-basic} emotion labels (\emph{i.e.}, \emph{pride}, \emph{curiosity}, \emph{boredom}, \emph{shame}, \emph{humiliation}, and \emph{disappointment}).
+After the emotion elicitation task, the encoders rated their subjective feeling on Likert scales from 0 (``not at all'') to 5 (``strongly'') related to six ``basic'' emotion labels (\emph{i.e.}, \emph{anger}, \emph{disgust}, \emph{fear}, \emph{happiness}, \emph{surprise} and \emph{sadness}) as well as six ``non-basic'' emotion labels (\emph{i.e.}, \emph{pride}, \emph{curiosity}, \emph{boredom}, \emph{shame}, \emph{humiliation}, and \emph{disappointment}).
 
@@
 For the human facial expression recognition method, 1383 student participants were recruited to annotate 232 out of the 358 videos, therefore only the 232 annotated videos were analyzed in this study.
-Because videos have different durations, participants had to annotate a series of video corresponding to 30min long in total.
+Because videos have different duration, participants had to annotate a series of video corresponding to 30min long in total.
 Each video was annotated 29 times on average (\emph{SD} = 12).
@@
 First, the participants had to identify the emotional sequences by pressing the space bar of their keyboard to indicate the beginning and the end of the emotional sequences while watching the video.
-Second, the participants watched each emotional sequence previously identified and labeled the sequence using one of the 12 emotions proposed including six basic emotion labels (\emph{i.e.}, \emph{anger}, \emph{disgust}, \emph{fear}, \emph{happiness}, \emph{surprise} and \emph{sadness}) and six \enquote{non-basic} emotion labels (\emph{i.e.}, \emph{pride}, \emph{curiosity}, \emph{boredom}, \emph{shame}, \emph{humiliation}, and \emph{disappointment}).
+Second, the participants watched each emotional sequence previously identified and labeled the sequence using one of the 12 emotions proposed including six basic emotion labels (\emph{i.e.}, \emph{anger}, \emph{disgust}, \emph{fear}, \emph{happiness}, \emph{surprise} and \emph{sadness}) and six ``non-basic'' emotion labels (\emph{i.e.}, \emph{pride}, \emph{curiosity}, \emph{boredom}, \emph{shame}, \emph{humiliation}, and \emph{disappointment}).
 They also had the possibility to indicate that the sequence was expressing none of the proposed emotion labels.
@@
 This annotation procedure resulted in a uni-dimensional time-series for each video per human observer, identifying for each second of the video which emotion was recognized.
-Then, time-series corresponding to the same video were aggregated to calculate the proportion of human observers \(x_{video_{i}.label_{j}.t_{k}}\) for each second of the video per emotional label (EQ\ref{eq:1}).
+Then, time-series corresponding to the same video were aggregated to calculate the proportion of human observers \(x_{video_{i}.label_{j}.t_{k}}\) for each second of the video per emotional label shown in Equation \eqref{eq:one}.
 
 \begin{equation}
-\label{eq:1}
-x_{video_{i}.label_{j}.t_{k}} = \frac{n_{video_{i}.label_{j}.t_{k}}}{n_{video_{k}}}
+x_{video_{i}.label_{j}.t_{k}} = \frac{n_{video_{i}.label_{j}.t_{k}}}{n_{video_{k}}}\label{eq:one}
 \end{equation}
 
-where \emph{i} is one of the 232 videos, \emph{j} is one of the six \enquote{basic} emotion labels, \emph{k} for each second of the video.
+where \emph{i} is one of the 232 videos, \emph{j} is one of the six ``basic'' emotion labels, \emph{k} for each second of the video.
 
@@
 For both human and automatic recognition, to determine which of the six basic emotions can be used to identify each video, the recognition probability for each label by frame was converted into odd ratio by label (Dente, Küster, Skora, \& Krumhuber, 2017).
-The highest sum of each odd ratio time-series defines the label recognized by the automatic classifier (EQ\ref{eq:2} for human recognition and EQ\ref{eq:3} for automatic recognition).
+The highest sum of each odd ratio time-series defines the label recognized (see Equation \eqref{eq:two} for human recognition and Equation \eqref{eq:three} for automatic recognition).
 
 \begin{equation}
-\label{eq:2}
-video_{i}.label = \max\left(\frac{\sum_{k=1}^{n}x_{video_{i}.label_{j}.t_{k}}}{\sum_{k=1}^{n}x_{video_{i}.t_{k}}}\right)
+video_{i}.label = \max\left(\frac{\sum_{k=1}^{n}x_{video_{i}.label_{j}.t_{k}}}{\sum_{k=1}^{n}x_{video_{i}.t_{k}}}\right)\label{eq:two}
 \end{equation}
@@
 \begin{equation}
-\label{eq:3}
-video_{i}.label = \max\left(\frac{\sum_{k=1}^{f}p_{video_{i}.label_{j}.t_{k}}}{\sum_{k=1}^{f}p_{video_{i}.t_{k}}}\right)
+video_{i}.label = \max\left(\frac{\sum_{k=1}^{f}p_{video_{i}.label_{j}.t_{k}}}{\sum_{k=1}^{f}p_{video_{i}.t_{k}}}\right)\label{eq:three}
 \end{equation}
 
-where \emph{i} is one of the 232 videos, \emph{j} is one of the six \enquote{basic} emotion labels, \emph{k} for each second of the \emph{n} second video or for each frame of the \emph{f} frame video.
-
-Since encoders' self-reports, human annotations and the automatic recognition include data on \enquote{non-basic} emotion labels and features, the analysis is performed using only the six basic emotion labels in order to compare them.
+where \emph{i} is one of the 232 videos, \emph{j} is one of the six ``basic'' emotion labels, \emph{k} for each second of the \emph{n} second video or for each frame of the \emph{f} frame video.
+
+Since encoders' self-reports, human annotations and the automatic recognition include data on ``non-basic'' emotion labels and features, the analysis is performed using only the six basic emotion labels in order to compare them.
 The maximum rating for self-reports, human annotations and automatic recognition is used to label the video.
@@
 
-Interestingly human observers seem to recognize surprise expressed in videos where anger, fear happiness and sadness was the highest self-reported emotion (respectively 28.6\%, 33.3\%, 34.7\% and 33.3\% of the videos self-reported), and in a lower instance \emph{happiness} was recognized in videos where \emph{fear} and \emph{surprise} was the highest self-reported emotion (respectively 50.0\% and 25.6\% of the videos self-reported).
+Interestingly human observers seem to recognize surprise expressed in videos where anger, fear happiness and sadness was the highest self-reported emotion (respectively 29\%, 33\%, 35\% and 33\% of the videos self-reported), and in a lower instance \emph{happiness} was recognized in videos where \emph{fear} and \emph{surprise} was the highest self-reported emotion (respectively 50\% and 26\% of the videos self-reported).
 
@@
 
-\begingroup\fontsize{8}{10}\selectfont
-
-\begin{ThreePartTable}
-\begin{TableNotes}[para]
-\item Note. na. values are produced when not enough data are available to compute accuracy indicators.
-\end{TableNotes}
-\begin{longtabu} to \linewidth {>{\raggedright}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
-\caption{\label{tab:confusionTable-sr-hr}Human recognition accuracy metrics for each emotion.}\\
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:confusionTable-sr-hr}Human recognition accuracy metrics for each emotion.}
+
+\begin{tabular}{lllll}
 \toprule
-Emotion & Sensitivity & Specificity & Precision & F1\\
+Emotion & \multicolumn{1}{c}{Sensitivity} & \multicolumn{1}{c}{Specificity} & \multicolumn{1}{c}{Precision} & \multicolumn{1}{c}{F1}\\
 \midrule
@@
 disgust & 1.00 & 0.94 & 0.33 & 0.5\\
-fear & 0.00 & 0.97 & 0.00 & \textit{na.}\\
+fear & 0.00 & 0.97 & 0.00 & na.\\
 happiness & 0.49 & 0.78 & 0.69 & 0.57\\
@@
 \bottomrule
-\insertTableNotes
-\end{longtabu}
-\end{ThreePartTable}
-\endgroup{}
+\addlinespace
+\end{tabular}
+
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} na. values are produced when not enough data are available to compute accuracy indicators.}
+\end{tablenotes}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
 
@@
 
-However, self-reports show a significant proportion of undetermined emotional states (35.2\% of the 358 videos) which reveals the potential limit of using 6-points Likert scales to measure emotional self-reports.
+However, self-reports show a significant proportion of undetermined emotional states (35\% of the 358 videos) which reveals the potential limit of using 6-points Likert scales to measure emotional self-reports.
 Indeed, encoders can easily score to the maximum for more than one emotion.
@@
 Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are somewhat similar to the ones with human observers (Table \ref{tab:confusionTable-sr-ar}).
-Overall, a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy = 0.30, 95\% CI {[}0.22,0.38{]}; Kappa = 0.07) except for \emph{happiness} (44.4\% of the video self-reported) is evident.
-
-\begingroup\fontsize{8}{10}\selectfont
-
-\begin{ThreePartTable}
-\begin{TableNotes}[para]
-\item Note. na. values are produced when not enough data are available to compute accuracy indicators.
-\end{TableNotes}
-\begin{longtabu} to \linewidth {>{\raggedright}X>{\centering}X>{\centering}X>{\centering}X>{\centering}X}
-\caption{\label{tab:confusionTable-sr-ar}Autonatic recognition accuracy metrics for each emotion.}\\
+Overall, a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy = 0.30, 95\% CI {[}0.22,0.38{]}; Kappa = 0.07) except for \emph{happiness} (44\% of the video self-reported) is evident.
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:confusionTable-sr-ar}Automatic recognition accuracy metrics for each emotion.}
+
+\begin{tabular}{lllll}
 \toprule
-Emotion & Sensitivity & Specificity & Precision & F1\\
+Emotion & \multicolumn{1}{c}{Sensitivity} & \multicolumn{1}{c}{Specificity} & \multicolumn{1}{c}{Precision} & \multicolumn{1}{c}{F1}\\
 \midrule
-anger & 0.00 & 0.90 & 0.00 & \textit{na.}\\
+anger & 0.00 & 0.90 & 0.00 & na.\\
 disgust & 0.25 & 0.72 & 0.02 & 0.04\\
-fear & 0.00 & 0.94 & 0.00 & \textit{na.}\\
+fear & 0.00 & 0.94 & 0.00 & na.\\
 happiness & 0.44 & 0.73 & 0.62 & 0.52\\
@@
 \bottomrule
-\insertTableNotes
-\end{longtabu}
-\end{ThreePartTable}
-\endgroup{}
-
-Surprisingly the automatic classifier incorrectly recognized as \emph{disgust} a significant proportion of videos in which \emph{anger}, \emph{happiness} and \emph{surprise} was the highest self-reported emotion (respectively 42.9\%, 25.0\% and 32.6\% of the videos self-reported).
-In parallel, the automatic classifier recognized as \emph{happiness} videos in which \emph{fear} and \emph{surprise} was the highest self-reported emotion (respectively 66.7\% and 32.6\% of the videos self-reported).
+\addlinespace
+\end{tabular}
+
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} na. values are produced when not enough data are available to compute accuracy indicators.}
+\end{tablenotes}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+Surprisingly the automatic classifier incorrectly recognized as \emph{disgust} a significant proportion of videos in which \emph{anger}, \emph{happiness} and \emph{surprise} was the highest self-reported emotion (respectively 43\%, 25\% and 33\% of the videos self-reported).
+In parallel, the automatic classifier recognized as \emph{happiness} videos in which \emph{fear} and \emph{surprise} was the highest self-reported emotion (respectively 67\% and 33\% of the videos self-reported).
 
@@
 \begin{figure}[!h]
-\includegraphics[width=\textwidth]{manuscript_apa_files/figure-latex/confusionMatrix-hr-ar-1} \caption{Proportion of emotion labels classified by human observers which are recognized by the automatif classifier.}\label{fig:confusionMatrix-hr-ar}
+\includegraphics[width=\textwidth]{manuscript_apa_files/figure-latex/confusionMatrix-hr-ar-1} \caption{Proportion of emotion labels classified by human observers which are recognized by the automatic classifier.}\label{fig:confusionMatrix-hr-ar}
 \end{figure}
@@
 The overall agreement between human observers and the automatic classifier is in fact very low (Kappa = 0.18).
-Except for \emph{happiness} and \emph{disgust} (respectively 64.9\% and 50.0\% of common labelling), there is no clear common pattern.
+Except for \emph{happiness} and \emph{disgust} (respectively 65\% and 50\% of common labelling), there is no clear common pattern.
 Moreover, the automatic classifier has a tendency to label as \emph{disgust} videos labeled as \emph{sadness} by human observers, and as \emph{happiness} videos labeled as \emph{fear} by human observers.
@@
 They bring up several questions regarding the role the context plays in the emission and interpretation of the so-called facial expression of emotion.
-The finalization of operational and effective \enquote{reading emotional faces} devices rests on the answers, if any, to the questions raised.
+The finalization of operational and effective ``reading emotional faces'' devices rests on the answers, if any, to the questions raised.
 As a result, despite being one of the most investigated questions in affective science, the growing interest of industries and governments in tracking individual's psychological states is supported by controversial assumptions.
@@
 Even if emotions were hardwired, in everyday life one does not observe prototypical facial expressions because of their rarity, and therefore research should be focused on analyzing non-prototypical facial expressions.
-Advancements in identifying \enquote{non-basic} emotion labels as well as non-prototypical facial expression have indeed occured with automatic facial expression recognition tools (McDuff, 2016).
+Advancements in identifying ``non-basic'' emotion labels as well as non-prototypical facial expression have indeed occured with automatic facial expression recognition tools (McDuff, 2016).
 However, these present results suggest that automatic facial expression recognition tools should merely evaluate facial morphology features such as action units (already evaluated in OpenFace, Baltrušaitis, Robinson, \& Morency, 2016; Affectiva's Affdex, McDuff et al., 2016; Vicar Vision's FaceReader, Den Uyl \& Van Kuilenburg, 2005, to name a few) rather than inferring supposedly emotional or affective states.
@@
 \hypertarget{refs}{}
+\begin{cslreferences}
 \leavevmode\hypertarget{ref-aviezer2017inherently}{}%
@@
 \url{https://doi.org/10.1037/0022-3514.50.4.737}
+\end{cslreferences}
 
@@
 
-\clearpage
-
-\renewcommand{\listfigurename}{Figure captions}
-
-\listoffigures
-
-\clearpage
-
-\renewcommand{\listtablename}{Table captions}
-
-\listoftables
-
 \end{document}

### 2020-08-28 manuscript/manuscript_apa.Rmd
review

@@
 
-The confusion matrices obtained by crossing encoders’ self-reports, human annotations and the automatic recognition are analysed by calulating two-sided Pearson correlation, accuracy which corresponds to the overall agreement rate averaged over cross-validation iterations and Cohen’s (unweighted) Kappa.
+The confusion matrices obtained by crossing encoders’ self-reports, human annotations and the automatic recognition are analysed by calculating two-sided Pearson correlation, accuracy which corresponds to the overall agreement rate averaged over cross-validation iterations and Cohen’s (unweighted) Kappa.
 

### 2020-08-28 revision/prpf_1/revision_letter.Rmd
review

@@
 ---
+
+```{r setup, include = FALSE}
+# libraries --------------------------------------------------------------------
+library(tidyverse)
+library(magrittr)
+library(here)
+library(ltm)
+
+# options ----------------------------------------------------------------------
+set.seed(123)
+options(scipen = 999)
+knitr::opts_chunk$set(
+  cache.extra = knitr::rand_seed,
+  cache = FALSE,
+  message = FALSE, 
+  warning = FALSE, 
+  error = FALSE, 
+  echo = FALSE
+  )
+
+# data -------------------------------------------------------------------------
+
+## self_report_data
+source(here::here("script/data_wrangling_self_report.R"))
+self_report_score <- self_report_score %>%
+  dplyr::select(C_Video, sr_emotion = emotion, sr_score = value)
+
+## human_recognition_data
+source(here::here("script/data_wrangling_human_recognition.R"))
+human_recognition_score <- human_recognition_score %>%
+  dplyr::select(C_Video, hr_emotion = emotion, hr_score = confidence_score)
+
+## automatic_recognition_data
+source(here::here("script/data_wrangling_automatic_recognition.R"))
+automatic_recognition_score <- automatic_recognition_score %>%
+  dplyr::select(C_Video, ar_emotion = emotion, ar_score = confidence_score) %>%
+  dplyr::filter(C_Video %in% ppt_per_video$C_Video)
+
+list_video <- human_recognition_score$C_Video
+list_emotion <- c("disgust","fear","happiness","surprise","sadness","anger")
+
+grid_data <- expand.grid(list_video = list_video, list_emotion = list_emotion)
+
+self_report_result <- self_report_score %>%
+  dplyr::filter(sr_emotion != "undetermined") %>%
+  dplyr::mutate(sr_score = 1)
+
+human_recognition_result <- human_recognition_score %>%
+  dplyr::filter(hr_emotion != "undetermined") %>%
+  dplyr::mutate(hr_score = 1)
+
+automatic_recognition_result <- automatic_recognition_score %>%
+  dplyr::filter(ar_emotion != "undetermined") %>%
+  dplyr::mutate(ar_score = 1)
+
+recognition_result <- grid_data %>%
+  dplyr::left_join(
+    self_report_result,
+    by = c(
+      "list_video" = "C_Video",
+      "list_emotion" = "sr_emotion"
+      )
+    ) %>%
+  dplyr::left_join(
+    human_recognition_result,
+    by = c(
+      "list_video" = "C_Video",
+      "list_emotion" = "hr_emotion"
+      )
+    ) %>%
+  dplyr::left_join(
+    automatic_recognition_result,
+    by = c(
+      "list_video" = "C_Video",
+      "list_emotion" = "ar_emotion"
+      )
+    ) %>%
+  replace(is.na(.), 0)
+```
 
@@
 
+We greatly appreciate the Reviewer’s positive comments. We have carefully revised the manuscript according to the Reviewer’s insightful comments and provided point-by-point responses as follows.
+
 ## Method section
@@
 \RC{Page 19 line 17 to 20: Which kind of correlation was computed here? A pearson correlation? It should be a point-biserial correlation as one of the two variables is dichotomous.}
+
+```{r echo=TRUE}
+recognition_result %$%
+  ltm::biserial.cor(sr_score, hr_score, level = 2)
+
+recognition_result %$%
+  cor.test(sr_score, hr_score, method = "pearson", alternative = "two.sided") %>% 
+  use_series("estimate")
+```
 

### 2020-08-28 script/data_wrangling_automatic_recognition.R

@@ -10,7 +10,7 @@ library(here)
 data_path <- here::here() %>% 
   dirname() %>% 
   dirname() %>% 
-  paste0("/data/dynemo_feel_data/automatic_recognition_data")
+  file.path("data/dynemo_feel_data/automatic_recognition_data")
 
 # upload data ------------------------------------------------------------------
 automatic_recognition_data <- file.path(data_path,  "automatic_recognition_data.rds") %>% 

### 2020-08-28 script/data_wrangling_human_recognition.R

@@ -11,7 +11,7 @@ library(here)
 data_path <- here::here() %>% 
   dirname() %>% 
   dirname() %>% 
-  paste0("/data/dynemo_feel_data/human_recognition_data")
+  file.path("data/dynemo_feel_data/human_recognition_data")
 
 # upload data ------------------------------------------------------------------
 metadata_video <- file.path(data_path,  "metadata_video.rds") %>% 
@@ -58,9 +58,9 @@ full_timeline <- unique_ppt %>%
   dplyr::mutate_if(is.factor,as.character)
 
 data_timeline <- human_recognition_data %>% 
-  mutate(timeline = map2(TC_Debut, TC_Fin, `:`)) %>% 
-  select(-TC_Debut, -TC_Fin) %>% 
-  unnest() %>% 
+  dplyr::mutate(timeline = map2(TC_Debut, TC_Fin, `:`)) %>% 
+  dplyr::select(-TC_Debut, -TC_Fin) %>% 
+  tidyr::unnest() %>% 
   dplyr::mutate_if(is.factor,as.character)
 
 final_timeline <- full_timeline %>% 

### 2020-08-28 script/data_wrangling_self_report.R

@@ -11,7 +11,7 @@ library(janitor)
 data_path <- here::here() %>% 
   dirname() %>% 
   dirname() %>% 
-  paste0("/data/dynemo_feel_data/human_self_report_data")
+  file.path("data/dynemo_feel_data/human_self_report_data")
 
 # upload data ------------------------------------------------------------------
 self_report_data <- file.path(data_path,  "Base_Finale_20090114.rds") %>% 

### 2020-09-02 manuscript/manuscript_apa.Rmd
sent to anna

--- 2020-08-28 manuscript/manuscript_apa.Rmd
@@
 bibliography      : ["bib/manuscript_ref.bib"]
+csl               : ["csl/apa7.csl"]
 
@@
 documentclass     : "apa7"
-classoption       : "doc"
+classoption       : "man"
 
-#output: papaja::apa6_word
+output: papaja::apa6_word
 
-output            :
-  bookdown::pdf_book:
-    toc: false
-    number_sections: false
-    base_format: papaja::apa6_pdf
+# output            :
+#   bookdown::pdf_book:
+#     toc: false
+#     number_sections: false
+#     base_format: papaja::apa6_pdf
 
@@
 
-With the development of commercial automatic facial expression recognition tools [@dupre2018accuracy], industries and governments are gradually implementing this technology in order to monitor humans’ emotions in various scenarios (*e.g.*, in marketing, healthcare, and the automotive industry to name a few).
+With the development of commercial automatic facial expression recognition tools [@dupre2018accuracy; @dupre2020performance], industries and governments are gradually implementing this technology in order to monitor humans’ emotions in various scenarios (*e.g.*, in marketing, healthcare, and the automotive industry to name a few).
 This technology rests on the premise that facial expressions provide direct access to individuals’ subjective feelings, and that one can read the emotions displayed on a person’s face as one would an open book.
@@
 Yet, in spite of the popular support it has received, the empirical data called upon by the Basic Emotion View remains unpersuasive.
-Notably, it fails to explain, in instances in which display rules cannot be called upon, how individuals can feel emotions without expressing them or how individuals can express emotions without feeling them, and why observers are not totally accurate in recognizing facial expressions of basic emotions, among others.
+Notably, it fails to explain, in instances in which display rules cannot be called upon, how individuals can feel emotions without expressing them or how individuals can express emotions without feeling them, and why observers can make mistakes in recognizing facial expressions of basic emotions, among others.
 The following  is a summary of key findings and conclusions both from field observations and laboratory experiments on spontaneous expression of emotions, all of them strongly subverting the Basic Emotion viewpoint [for reviews, @fernandez2013emotion; @reisenzein2013coherence; @russell2003facial].
@@
 Though they cannot be considered as a straightforward test of the triggering role of emotion on facial behavior, they have the advantage of considering situations that cannot be created in a laboratory.
-For instance, the ethological study of @kraut1979social contradicts the Basic Emotion View premise that a smile is the major component of a facial display associated with and caused by feelings of joy or happiness.
+For instance, the ethological study of @kraut1979social contradicts the Basic Emotion View premise that a smile is only the major component of a facial display associated with and caused by feelings of joy or happiness.
 Naturalistic observation at a bowling alley showed that bowlers do not necessarily smile after scoring a spare or a strike (a situation likely to elicit a positive emotion).
@@
 Among them, the constructivist approach is gaining in importance.
-The constructivist approach represents a totally different way of understanding the emotion–facial expression link.
+The constructivist approach represents a different way of understanding the emotion–facial expression link.
 It affirms that facial expressions do not provide a direct access to individuals’ subjective feelings.
@@
 For the emotion elicitation experiment, `r nrow(self_report_data)` encoding participants (`r self_report_gender["F"]` females, `r self_report_gender["H"]` males, *M*~age~ = `r self_report_age["age_m"]`, *SD*~age~ = `r self_report_age["age_sd"]`) were recruited to perform one out of 11 emotion elicitation tasks designed to trigger a positive, a specific negative or a neutral emotional state.
+These participants were recruited by a private company for a study supposedly devoted to an "ergonomic visual task" (cover story).
+After a description of the general aims, participants agreed and signed the experiment consent form.
+At the end of the elicitation task, they received an equivalent of €50 in voucher for their participation.
+A second consent form was signed by the participants to allow their video to be processed for research purposes.
+
+Two different types of task were used to elicit emotions.
+Some tasks were passive and consisted in the participant watching videos selected to trigger emotions (e.g.
+television commercials).
+Other tasks were active; they required the participant to interact with a computer (e.g.
+to answer questions or to test a flawed software).
 Encoders’ faces were recorded using a hidden camera resulting `r nrow(self_report_data)` front facing 768x576 videos varying from `r min(metadata_video$ffprobe_duration)`s to `r max(metadata_video$ffprobe_duration)`s.
@@
 
-For the human facial expression recognition method, `r nrow(unique_ppt_count)` student participants were recruited to annotate `r unique_video_count["vid_n"]` out of the `r nrow(self_report_data)` videos, therefore only the `r unique_video_count["vid_n"]` annotated videos were analyzed in this study.
-Because videos have different duration, participants had to annotate a series of video corresponding to 30min long in total.
+To analyse how the recorded facial expressions are perceived, `r nrow(unique_ppt_count)` participants were recruited among social science under- and post-graduates through advertising to watch and to annotate these recordings.
+They received a course credit for their participation.
+After a description of the general aims, participants agreed and signed the experiment consent form.
+An iterative procedure was setup to insure that each recording has been annotated by at least 20 participants: 10% of the video database were randomly selected and annotated until all the videos reached the amount of annotation required, then another section of the database is randomly selected among the remaining recordings to be annotated.
+To avoid potential decrease in the quality of the annotations due to cognitive fatigue, each participant only annotated recordings during a maximum of 30min.
+As a results only `r unique_video_count["vid_n"]` out of the `r nrow(self_report_data)` videos have been annotated.
 Each video was annotated `r unique_video_count["vid_m"]` times on average (*SD* = `r unique_video_count["vid_sd"]`).
@@
 Affdex’s algorithm uses Histogram of Oriented Gradient (HOG) features and Support Vector Machine (SVM) classifiers in order to recognize facial expressions.
-For each video frame, Affdex identifies the probability $p_{video_{i}.label_{j}.t_{k}}$ from 0 to 100 (rescaled to 0 to 1 for the analysis) of the face as expressing one of the six basic emotion labels (*i.e.*, *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as additional psychological states such as *valence*, *engagement* or *contempt*, and facial features such as *cheek raise*, *eye widen* or *jaw drop*.
+For each video frame, Affdex identifies the probability $p_{video_{i}.label_{j}.t_{k}}$ of expressing each of the six basic emotion labels (*i.e.*, *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as additional psychological states such as *valence*, *engagement* or *contempt*, and facial features such as *cheek raise*, *eye widen* or *jaw drop*.
+The result of Affdex is a mutivariate timeseries output which provides for each label a probability from 0 to 100 (rescaled from 0 to 1 for the analysis) of expressing the corresponding label.
+For this study only the six basic emotion labels are analysed to match with self-reports and human observers.
 
@@
 
-The confusion matrices obtained by crossing encoders’ self-reports, human annotations and the automatic recognition are analysed by calulating two-sided Pearson correlation, accuracy which corresponds to the overall agreement rate averaged over cross-validation iterations and Cohen’s (unweighted) Kappa.
+The confusion matrices obtained by crossing encoders’ self-reports, human annotations and the automatic recognition are analysed by calculating two-tailed t-tests of Pearson's product-moment correlation between dichotomous variables, accuracy which corresponds to the overall agreement rate averaged over cross-validation iterations and Cohen’s (unweighted) Kappa.
 
@@
 Results of the confusion matrix show a low agreement between the emotion felt by the encoder during the elicitation and the emotion recognized by the human observers (Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95% CI [`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *disgust* (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_sr"]))` of the videos self-reported).
-These results are far from those classically obtained in the literature for emotional facial expression recognition which ranges between 60% and 80% accuracy.
-However these results are mostly obtained with static (*i.e.*, pictures) and posed (*i.e.*, displayed by actors) facial expressions using only 6 emotional labels in a forced-choice paradigm.
+These results are far from those classically obtained in the literature for emotional facial expression recognition [@krumhuber2020human, report an average accuracy of 65%].
+However these results are mostly obtained with posed facial expressions (*i.e.*, displayed by actors) in a forced-choice recognition paradigm.
 
@@
 Whereas *happiness* and *disgust* obtain the highest scores, *anger*, *surprise* and *sadness* have the lowest recognition ratio.
-The underlying effect of expression intensity may explain why *happiness* and *disgust* are easily recognized.
-*Anger* and *sadness* as non-socially desirable emotions may be have been felt but not expressed.
+This difference between emotions is usually observed in the literature [e.g.
+@krumhuber2020human].
+Multiple reasons can explain why *happiness* and *disgust* are more easily recognized than *anger*, *surprise* and *sadness*.
+It is possible that these expressions are usually more intense than others.
+In addition, *anger* and *sadness* as non-socially desirable emotions may be have been felt but not expressed.
 
 However, self-reports show a significant proportion of undetermined emotional states (`r scales::percent(nrow(list_automatic_ties)/358)` of the 358 videos) which reveals the potential limit of using 6-points Likert scales to measure emotional self-reports.
-Indeed, encoders can easily score to the maximum for more than one emotion.
+Indeed, Likert scales may not be able to discriminate between two dominant emotion felt with different intensities due to their reduced number of possibilities.
+To better discriminate between emotions, an alternative would be to use a continuous slider from 0 to 100 [see for example @lottridge2009emotional].
 
@@
 
-As previously mentioned, human observers appear to be more accurate than the automatic classifier to recognize an individual's subjective feeling (human observers Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`; automatic classifier Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`; `r corr_hr_ar$full_result`).
-However, both make mistakes.
+As previously mentioned, human observers appear to be more accurate than the automatic classifier to recognize an individual's subjective feeling (human observers Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`; automatic classifier Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`).
+However, both make similar mistakes as the two-tailed t-tests of Pearson's product-moment correlation between human observers and automatic classifier recognition is significant (`r corr_hr_ar$full_result`).
 

### 2020-09-02 manuscript/manuscript_apa.tex
sent to anna

@@
 
-With the development of commercial automatic facial expression recognition tools (Dupré, Andelic, Morrison, \& McKeown, 2018), industries and governments are gradually implementing this technology in order to monitor humans' emotions in various scenarios (\emph{e.g.}, in marketing, healthcare, and the automotive industry to name a few).
+With the development of commercial automatic facial expression recognition tools (Dupré et al., 2020, 2018), industries and governments are gradually implementing this technology in order to monitor humans' emotions in various scenarios (\emph{e.g.}, in marketing, healthcare, and the automotive industry to name a few).
 This technology rests on the premise that facial expressions provide direct access to individuals' subjective feelings, and that one can read the emotions displayed on a person's face as one would an open book.
-This premise is derived from the ``Basic Emotion View'', a theory which suggests a one-to-one mapping between subjective feeling and facial expression (Ekman, 1992, 2007; Ekman et al., 1987; Ekman \& Heider, 1988).
+This premise is derived from the ``Basic Emotion View'', a theory which suggests a one-to-one mapping between subjective feeling and facial expression (Ekman et al., 1987; Ekman, 1992, 2007; Ekman \& Heider, 1988).
 Even if this premise is central to the modern mainstream approach of human emotion, recent research in affective science is challenging it.
@@
 Yet, in spite of the popular support it has received, the empirical data called upon by the Basic Emotion View remains unpersuasive.
-Notably, it fails to explain, in instances in which display rules cannot be called upon, how individuals can feel emotions without expressing them or how individuals can express emotions without feeling them, and why observers are not totally accurate in recognizing facial expressions of basic emotions, among others.
-The following is a summary of key findings and conclusions both from field observations and laboratory experiments on spontaneous expression of emotions, all of them strongly subverting the Basic Emotion viewpoint (for reviews, Fernández-Dols \& Crivelli, 2013; Reisenzein, Studtmann, \& Horstmann, 2013; Russell, Bachorowski, \& Fernández-Dols, 2003).
+Notably, it fails to explain, in instances in which display rules cannot be called upon, how individuals can feel emotions without expressing them or how individuals can express emotions without feeling them, and why observers can make mistakes in recognizing facial expressions of basic emotions, among others.
+The following is a summary of key findings and conclusions both from field observations and laboratory experiments on spontaneous expression of emotions, all of them strongly subverting the Basic Emotion viewpoint (for reviews, Fernández-Dols \& Crivelli, 2013; Reisenzein et al., 2013; Russell et al., 2003).
 Mains issues regarding recognition studies are also briefly discussed afterwards.
@@
 Though they cannot be considered as a straightforward test of the triggering role of emotion on facial behavior, they have the advantage of considering situations that cannot be created in a laboratory.
-For instance, the ethological study of Kraut \& Johnston (1979) contradicts the Basic Emotion View premise that a smile is the major component of a facial display associated with and caused by feelings of joy or happiness.
+For instance, the ethological study of Kraut \& Johnston (1979) contradicts the Basic Emotion View premise that a smile is only the major component of a facial display associated with and caused by feelings of joy or happiness.
 Naturalistic observation at a bowling alley showed that bowlers do not necessarily smile after scoring a spare or a strike (a situation likely to elicit a positive emotion).
@@
 More generally, examination of bowlers' facial display showed that they rarely smiled while facing the pins but often smiled when facing their friends.
-These findings were confirmed by Ruiz-Belda, Fernández-Dols, Carrera, \& Barchard (2003) who analyzed the facial displays of bowlers after a strike and soccer fans after their team scored.
+These findings were confirmed by Ruiz-Belda et al.
+(2003) who analyzed the facial displays of bowlers after a strike and soccer fans after their team scored.
 Their results show a low probability of smiling when participants were not interacting with someone else.
@@
 More specifically, winners showed Duchenne smiles and other types of smiles when they were interacting with other people during the awards ceremony, but smiles were scarce or nonexistent when waiting behind the podium and/or when turning toward the flagpoles and focusing their attention on the flags and the national anthem.
-Crivelli, Carrera, \& Fernández-Dols (2015) also found that the strongest predictor for the occurrence of Duchenne smiles in judo winners is when they are engaged in social interaction and not when they just won their match.
+Crivelli et al.
+(2015) also found that the strongest predictor for the occurrence of Duchenne smiles in judo winners is when they are engaged in social interaction and not when they just won their match.
 
@@
 Camras and her colleagues observed that facial expressions of negative emotions were displayed in circumstances that were unlikely to have elicited those emotions.
-For instance, infants displayed ``fear'' expressions in settings not related to fear (Camras, Malatesta, \& Izard, 1991).
-Bennett, Bendersky, \& Lewis (2002) videotaped a sample of 4-month-old infants during tickle, sour taste, jack-in-the-box, arm restraint, and masked stranger situations.
+For instance, infants displayed ``fear'' expressions in settings not related to fear (Camras et al., 1991).
+Bennett et al.
+(2002) videotaped a sample of 4-month-old infants during tickle, sour taste, jack-in-the-box, arm restraint, and masked stranger situations.
 Infants displayed a variety of facial expressions in each eliciting situation.
@@
 As an example, a strong disconfirmation of the Basic Emotion View's premise is put forward by Reisenzein's studies on surprise (\emph{e.g.}, Reisenzein, 2000).
-In eight controlled laboratory situations, surprise was induced by Reisenzein, Bördgen, Holtbernd, \& Matz (2006) by establishing and then invalidating a set of beliefs concerning the experimental events such as the unexpected appearance of a picture of one's own face as the last picture in a series of portraits that had to be rated.
+In eight controlled laboratory situations, surprise was induced by Reisenzein et al.
+(2006) by establishing and then invalidating a set of beliefs concerning the experimental events such as the unexpected appearance of a picture of one's own face as the last picture in a series of portraits that had to be rated.
 Visible or electromyography-detected facial displays of surprise occurred only in few participants.
@@
 
-Studies on happiness and related positive emotions such as sensory pleasantness also show a low coherence between emotion and facial display (Durán, Reisenzein, \& Fernández-Dols, 2017; Reisenzein et al., 2013).
+Studies on happiness and related positive emotions such as sensory pleasantness also show a low coherence between emotion and facial display (Durán et al., 2017; Reisenzein et al., 2013).
 In fact, only experiments on amusement provide a fairly strong association between emotion and smiling.
-When confronted to humorous events (\emph{e.g.}, jokes, being tickled), the number of participants smiling and laughing is indeed quite fair, whereas when confronted to other positive emotions (\emph{e.g.}, happiness), few participants show the expected facial expressions (for instance, the Duchenne smile or any kind of expression related to happiness; see Mehu, Grammer, \& Dunbar, 2007; Lee \& Wagner, 2002).
+When confronted to humorous events (\emph{e.g.}, jokes, being tickled), the number of participants smiling and laughing is indeed quite fair, whereas when confronted to other positive emotions (\emph{e.g.}, happiness), few participants show the expected facial expressions (for instance, the Duchenne smile or any kind of expression related to happiness; see Mehu et al., 2007; Lee \& Wagner, 2002).
 More largely, the meta-analysis conducted by Durán et al.
@@
 First of all, the basic emotions are weakly correlated with the specific configuration of facial muscles that the Basic Emotion View ascribes to them (the correlation drops when happiness/amusement studies are excluded).
-Very few participants who relived an experience of sadness were reactive, only a small number of them showing components of a sad facial expression with oblique eyebrows or lip corners pulled down (Tsai, Chentsova-Dutton, Freire-Bebeau, \& Przymus, 2002).
+Very few participants who relived an experience of sadness were reactive, only a small number of them showing components of a sad facial expression with oblique eyebrows or lip corners pulled down (Tsai et al., 2002).
 Regarding fear, about one third of spider phobic participants exposed to a live tarantula displayed some components of the prototypical fear expression such as eye widening, brow raising and knitting (Vernon \& Berenbaum, 2002).
 Similar observation is made for anger.
-The proportion of facially reactive participants displaying at least one component among frowning or lid/lip tightening and reporting anger does not exceed 35\% (Johnson, Waugh, \& Fredrickson, 2010; Tsai et al., 2002).
-The same proportion of reactive participants confronted to disgusting elicitors is observed (\emph{e.g.}, Ekman, Freisen, \& Ancoli, 1980; Fernandez-Dols, Sanchez, Carrera, \& Ruiz-Belda, 1997).
+The proportion of facially reactive participants displaying at least one component among frowning or lid/lip tightening and reporting anger does not exceed 35\% (Johnson et al., 2010; Tsai et al., 2002).
+The same proportion of reactive participants confronted to disgusting elicitors is observed (\emph{e.g.}, Ekman et al., 1980; Fernandez-Dols et al., 1997).
 In any case neither an insufficient intensity of the emotion to cause a facial expression nor the intervention of display rules (nor measurement issues) can explain the low emotion--facial expression consistency.
@@
 Thus, results gathered with forced choice cannot demonstrate the unequivocal link between emotion and facial expressions claimed by the Basic Emotion View.
-Using alternative recognition methods (emotion satiation procedure, face-matching task, sorting task; \emph{e.g.}, Lindquist, Barrett, Bliss-Moreau, \& Russell, 2006; Gendron, Roberson, Vyver, \& Barrett, 2014), it has been shown on the contrary that facial muscle movements are not linked in a one-to-one manner to a specific discrete emotional experience.
+Using alternative recognition methods (emotion satiation procedure, face-matching task, sorting task; \emph{e.g.}, Lindquist et al., 2006; Gendron et al., 2014), it has been shown on the contrary that facial muscle movements are not linked in a one-to-one manner to a specific discrete emotional experience.
 Instead, emotions are probably mentally constructed by the perceiver and mental categories of emotions are needed to accurately categorize facial movements among contextual information.
@@
 Basic Emotion View empirical evidence is based for the most part on methods using a static and unnatural material, namely, still photographs of posed facial expressions of emotion (\emph{e.g.}, intentionally encoded by the sender).
-This kind of methodology raises questions about its ecological validity and the generalizability of the results to real interpersonal emotional communication (\emph{e.g.}, Tcherkassof, Bollon, Dubois, Pansu, \& Adam, 2007).
+This kind of methodology raises questions about its ecological validity and the generalizability of the results to real interpersonal emotional communication (\emph{e.g.}, Tcherkassof et al., 2007).
 Indeed, a number of pieces of evidence indicate that research cannot content itself with data collected with static and posed material.
@@
 Furthermore, studied EFE exhibit emotions simulated or posed by actors.
-Yet, the lack of spontaneity and naturalness of this material constitutes a serious objection raised against such studies (Kanade, Cohn, \& Tian, 2000).
+Yet, the lack of spontaneity and naturalness of this material constitutes a serious objection raised against such studies (Kanade et al., 2000).
 
@@
 Among them, the constructivist approach is gaining in importance.
-The constructivist approach represents a totally different way of understanding the emotion--facial expression link.
+The constructivist approach represents a different way of understanding the emotion--facial expression link.
 It affirms that facial expressions do not provide a direct access to individuals' subjective feelings.
-Therefore, instead of considering that emotions can be ``read'' on facial displays, it claims that the emotion is ``in the eye'' of the perceiver (Barrett, Adolphs, Marsella, Martinez, \& Pollak, 2019).
+Therefore, instead of considering that emotions can be ``read'' on facial displays, it claims that the emotion is ``in the eye'' of the perceiver (Barrett et al., 2019).
 
@@
 Indeed, when Himba and Hazda participants did not have emotion concepts at their disposal to structure perception, they perceived the facial expressions as behaviors, such as looking or smelling, that didn't have a necessary relationship to emotions.
-They did not infer inner states, rather they proceeded with action identification that pointed out the functions of behaviors (see also Crivelli, Russell, Jarillo, \& Fernández-Dols, 2017 for similar observations among a small-scale society of Papua New Guinea).
+They did not infer inner states, rather they proceeded with action identification that pointed out the functions of behaviors (see also Crivelli et al., 2017 for similar observations among a small-scale society of Papua New Guinea).
 The constructivist approach considers that specific emotion categories, as conceptualized by Western cultures' knowledge, are cast on the perceived face to make meaning of the sender's facial displays.
@@
 For the emotion elicitation experiment, 358 encoding participants (182 females, 176 males, \emph{M}\textsubscript{age} = 47.9, \emph{SD}\textsubscript{age} = 9.2) were recruited to perform one out of 11 emotion elicitation tasks designed to trigger a positive, a specific negative or a neutral emotional state.
+These participants were recruited by a private company for a study supposedly devoted to an ``ergonomic visual task'' (cover story).
+After a description of the general aims, participants agreed and signed the experiment consent form.
+At the end of the elicitation task, they received an equivalent of €50 in voucher for their participation.
+A second consent form was signed by the participants to allow their video to be processed for research purposes.
+
+Two different types of task were used to elicit emotions.
+Some tasks were passive and consisted in the participant watching videos selected to trigger emotions (e.g.~television commercials).
+Other tasks were active; they required the participant to interact with a computer (e.g.~to answer questions or to test a flawed software).
 Encoders' faces were recorded using a hidden camera resulting 358 front facing 768x576 videos varying from 1s to 1479s.
@@
 
-For the human facial expression recognition method, 1383 student participants were recruited to annotate 232 out of the 358 videos, therefore only the 232 annotated videos were analyzed in this study.
-Because videos have different duration, participants had to annotate a series of video corresponding to 30min long in total.
+To analyse how the recorded facial expressions are perceived, 1383 participants were recruited among social science under- and post-graduates through advertising to watch and to annotate these recordings.
+They received a course credit for their participation.
+After a description of the general aims, participants agreed and signed the experiment consent form.
+An iterative procedure was setup to insure that each recording has been annotated by at least 20 participants: 10\% of the video database were randomly selected and annotated until all the videos reached the amount of annotation required, then another section of the database is randomly selected among the remaining recordings to be annotated.
+To avoid potential decrease in the quality of the annotations due to cognitive fatigue, each participant only annotated recordings during a maximum of 30min.
+As a results only 232 out of the 358 videos have been annotated.
 Each video was annotated 29 times on average (\emph{SD} = 12).
@@
 Affdex's algorithm uses Histogram of Oriented Gradient (HOG) features and Support Vector Machine (SVM) classifiers in order to recognize facial expressions.
-For each video frame, Affdex identifies the probability \(p_{video_{i}.label_{j}.t_{k}}\) from 0 to 100 (rescaled to 0 to 1 for the analysis) of the face as expressing one of the six basic emotion labels (\emph{i.e.}, \emph{anger}, \emph{disgust}, \emph{fear}, \emph{happiness}, \emph{surprise} and \emph{sadness}) as well as additional psychological states such as \emph{valence}, \emph{engagement} or \emph{contempt}, and facial features such as \emph{cheek raise}, \emph{eye widen} or \emph{jaw drop}.
-
-For both human and automatic recognition, to determine which of the six basic emotions can be used to identify each video, the recognition probability for each label by frame was converted into odd ratio by label (Dente, Küster, Skora, \& Krumhuber, 2017).
+For each video frame, Affdex identifies the probability \(p_{video_{i}.label_{j}.t_{k}}\) of expressing each of the six basic emotion labels (\emph{i.e.}, \emph{anger}, \emph{disgust}, \emph{fear}, \emph{happiness}, \emph{surprise} and \emph{sadness}) as well as additional psychological states such as \emph{valence}, \emph{engagement} or \emph{contempt}, and facial features such as \emph{cheek raise}, \emph{eye widen} or \emph{jaw drop}.
+The result of Affdex is a mutivariate timeseries output which provides for each label a probability from 0 to 100 (rescaled from 0 to 1 for the analysis) of expressing the corresponding label.
+For this study only the six basic emotion labels are analysed to match with self-reports and human observers.
+
+For both human and automatic recognition, to determine which of the six basic emotions can be used to identify each video, the recognition probability for each label by frame was converted into odd ratio by label (Dente et al., 2017).
 The highest sum of each odd ratio time-series defines the label recognized (see Equation \eqref{eq:two} for human recognition and Equation \eqref{eq:three} for automatic recognition).
@@
 
-The confusion matrices obtained by crossing encoders' self-reports, human annotations and the automatic recognition are analysed by calulating two-sided Pearson correlation, accuracy which corresponds to the overall agreement rate averaged over cross-validation iterations and Cohen's (unweighted) Kappa.
+The confusion matrices obtained by crossing encoders' self-reports, human annotations and the automatic recognition are analysed by calculating two-tailed t-tests of Pearson's product-moment correlation between dichotomous variables, accuracy which corresponds to the overall agreement rate averaged over cross-validation iterations and Cohen's (unweighted) Kappa.
 
@@
 Results of the confusion matrix show a low agreement between the emotion felt by the encoder during the elicitation and the emotion recognized by the human observers (Accuracy = 0.43, 95\% CI {[}0.35,0.52{]}; Kappa = 0.19) except for \emph{disgust} (100\% of the videos self-reported).
-These results are far from those classically obtained in the literature for emotional facial expression recognition which ranges between 60\% and 80\% accuracy.
-However these results are mostly obtained with static (\emph{i.e.}, pictures) and posed (\emph{i.e.}, displayed by actors) facial expressions using only 6 emotional labels in a forced-choice paradigm.
+These results are far from those classically obtained in the literature for emotional facial expression recognition (Krumhuber et al., 2020, report an average accuracy of 65\%).
+However these results are mostly obtained with posed facial expressions (\emph{i.e.}, displayed by actors) in a forced-choice recognition paradigm.
 
@@
 Whereas \emph{happiness} and \emph{disgust} obtain the highest scores, \emph{anger}, \emph{surprise} and \emph{sadness} have the lowest recognition ratio.
-The underlying effect of expression intensity may explain why \emph{happiness} and \emph{disgust} are easily recognized.
-\emph{Anger} and \emph{sadness} as non-socially desirable emotions may be have been felt but not expressed.
+This difference between emotions is usually observed in the literature (e.g.~Krumhuber et al., 2020).
+Multiple reasons can explain why \emph{happiness} and \emph{disgust} are more easily recognized than \emph{anger}, \emph{surprise} and \emph{sadness}.
+It is possible that these expressions are usually more intense than others.
+In addition, \emph{anger} and \emph{sadness} as non-socially desirable emotions may be have been felt but not expressed.
 
 However, self-reports show a significant proportion of undetermined emotional states (35\% of the 358 videos) which reveals the potential limit of using 6-points Likert scales to measure emotional self-reports.
-Indeed, encoders can easily score to the maximum for more than one emotion.
+Indeed, Likert scales may not be able to discriminate between two dominant emotion felt with different intensities due to their reduced number of possibilities.
+To better discriminate between emotions, an alternative would be to use a continuous slider from 0 to 100 (see for example Lottridge \& Chignell, 2009).
 
@@
 
-As previously mentioned, human observers appear to be more accurate than the automatic classifier to recognize an individual's subjective feeling (human observers Accuracy = 0.43; automatic classifier Accuracy = 0.30; \(r = .22\), 95\% CI \([.17\), \(.27]\), \(t(1384) = 8.31\), \(p < .001\)).
-However, both make mistakes.
+As previously mentioned, human observers appear to be more accurate than the automatic classifier to recognize an individual's subjective feeling (human observers Accuracy = 0.43; automatic classifier Accuracy = 0.30).
+However, both make similar mistakes as the two-tailed t-tests of Pearson's product-moment correlation between human observers and automatic classifier recognition is significant (\(r = .22\), 95\% CI \([.17\), \(.27]\), \(t(1384) = 8.31\), \(p < .001\)).
 
@@
 However, the human decoders in the present study were presented with faces without any contextual information which could have helped them to shape more precisely their interpretation.
-Hassin, Aviezer, \& Bentin (2013) reviewed evidence to demonstrate that faces are inherently ambiguous and that observers rely on situational cues when they process facial displays (see also Aviezer et al., 2008; Barrett, Mesquita, \& Gendron, 2011).
+Hassin et al.
+(2013) reviewed evidence to demonstrate that faces are inherently ambiguous and that observers rely on situational cues when they process facial displays (see also Aviezer et al., 2008; Barrett et al., 2011).
 It happens that contextual information even shifts emotion perception.
-In the present case, the facial stimuli displayed to the decoders were as equivocal as real-life facial expressions (Aviezer, Ensenberg, \& Hassin, 2017).
+In the present case, the facial stimuli displayed to the decoders were as equivocal as real-life facial expressions (Aviezer et al., 2017).
 Yet, decoders had no cue at their disposal.
-So, without any possibility of integrating faces and context, their decoding accuracy is reduced (\emph{e.g.}, Wagner, MacDonald, \& Manstead, 1986).
+So, without any possibility of integrating faces and context, their decoding accuracy is reduced (\emph{e.g.}, Wagner et al., 1986).
 However, their superiority to the classifier tool is probably owing to their capacity to rely on their previous personal experience to invent a context in which a face could display such an expression.
@@
 This latter issue is particularly decisive.
-Indeed, the issue of the recognition of dynamic expressive sequences is essential because ordinary facial behavior is made up of dynamically shifting morphological features (Krumhuber, Kappas, \& Manstead, 2013).
+Indeed, the issue of the recognition of dynamic expressive sequences is essential because ordinary facial behavior is made up of dynamically shifting morphological features (Krumhuber et al., 2013).
 This temporal information is indisputably a key feature of facial activity.
@@
 Moreover, as encoders are alone, facing an emotional trigger, it also controls for the social context, removing its possible influence on their facial behaviors.
-However, it is known that encoders' imagination can influence their expressiveness, for instance when they believe that their friend is doing the same (\emph{vs.} a different) emotional task in another room (Jakobs, Manstead, \& Fischer, 1999).
+However, it is known that encoders' imagination can influence their expressiveness, for instance when they believe that their friend is doing the same (\emph{vs.} a different) emotional task in another room (Jakobs et al., 1999).
 Trying to exclude social influence by leaving encoders alone may be illusionary.
@@
 It is the case, for instance, of emotional mimicry in dynamic social interactions (Hess \& Fischer, 2016).
-It is also the case of healthy patients interacting with schizophrenic patients whose facial activity is almost identical (Krause, Steimer-Krause, Merten, \& Ullrich, 1998).
+It is also the case of healthy patients interacting with schizophrenic patients whose facial activity is almost identical (Krause et al., 1998).
 Hence, in order to better comprehend emotional communication in human relationships, experimental research should be corroborated with more ecological protocols.
@@
 Advancements in identifying ``non-basic'' emotion labels as well as non-prototypical facial expression have indeed occured with automatic facial expression recognition tools (McDuff, 2016).
-However, these present results suggest that automatic facial expression recognition tools should merely evaluate facial morphology features such as action units (already evaluated in OpenFace, Baltrušaitis, Robinson, \& Morency, 2016; Affectiva's Affdex, McDuff et al., 2016; Vicar Vision's FaceReader, Den Uyl \& Van Kuilenburg, 2005, to name a few) rather than inferring supposedly emotional or affective states.
+However, these present results suggest that automatic facial expression recognition tools should merely evaluate facial morphology features such as action units (already evaluated in OpenFace, Baltrušaitis et al., 2016; Affectiva's Affdex, McDuff et al., 2016; Vicar Vision's FaceReader, Den Uyl \& Van Kuilenburg, 2005, to name a few) rather than inferring supposedly emotional or affective states.
 Trying to interpret facial displays as a means of determining underlying emotional state, in all likelihood, remains in vain.
@@
 Aviezer, H., Hassin, R.
-R., Ryan, J., Grady, C., Susskind, J., Anderson, A., \ldots{} Bentin, S.
+R., Ryan, J., Grady, C., Susskind, J., Anderson, A., Moscovitch, M., \& Bentin, S.
 (2008).
@@
 Openface: An open source facial behavior analysis toolkit.
-In \emph{Winter conference on applications of computer vision} (pp.
-1--10).
+\emph{Winter Conference on Applications of Computer Vision}, 1--10.
 \url{https://doi.org/10.1109/WACV.2016.7477553}
@@
 31--54).
-New York, NY: Cambridge University Press.
+Cambridge University Press.
 
@@
 \leavevmode\hypertarget{ref-darwin1872expression}{}%
-Darwin, C. (1872). \emph{The expression of the emotions in man and animals}. London, UK: John Murray.
+Darwin, C. (1872). \emph{The expression of the emotions in man and animals}. John Murray.
 
@@
 Measures and metrics for automatic emotion classification via facet.
-In \emph{Proceedings of the conference on the study of artificial intelligence and simulation of behaviour} (pp.
-160--163).
+\emph{Proceedings of the Conference on the Study of Artificial Intelligence and Simulation of Behaviour}, 160--163.
 
 \leavevmode\hypertarget{ref-den2005facereader}{}%
-Den Uyl, M., \& Van Kuilenburg, H. (2005). The facereader: Online facial expression recognition. In \emph{Proceedings of measuring behavior} (pp. 589--590).
+Den Uyl, M., \& Van Kuilenburg, H. (2005). The facereader: Online facial expression recognition. \emph{Proceedings of Measuring Behavior}, 589--590.
 
@@
 457--478).
-New York, NY: Oxford University Press.
+Oxford University Press.
 
 \leavevmode\hypertarget{ref-duchenne1876mecanisme}{}%
-Duchenne, G.-B. (1876). \emph{Mécanisme de la physionomie humaine: Où, analyse électro-physiologique de l'expression des passions}. Paris, France: J.-B. Baillière.
+Duchenne, G.-B. (1876). \emph{Mécanisme de la physionomie humaine: Où, analyse électro-physiologique de l'expression des passions}. J.-B. Baillière.
 
 \leavevmode\hypertarget{ref-dupre2015oudjat}{}%
-Dupré, D., Akpan, D., Elias, E., Adam, J.-M., Meillon, B., Bonnefond, N., \ldots{} Tcherkassof, A.
+Dupré, D., Akpan, D., Elias, E., Adam, J.-M., Meillon, B., Bonnefond, N., Dubois, M., \& Tcherkassof, A.
 (2015).
@@
 Accuracy of three commercial automatic emotion recognition systems across different individuals and their facial expressions.
-In \emph{Proceedings of the international conference on pervasive computing and communications} (pp.
-627--632).
+\emph{Proceedings of the International Conference on Pervasive Computing and Communications}, 627--632.
 \url{https://doi.org/10.1109/PERCOMW.2018.8480127}
+
+\leavevmode\hypertarget{ref-dupre2020performance}{}%
+Dupré, D., Krumhuber, E.
+G., Küster, D., \& McKeown, G.
+J.
+(2020).
+A performance comparison of eight commercially available automatic classifiers for facial affect recognition.
+\emph{Plos One}, \emph{15}(4), e0231968.
+\url{https://doi.org/10.1371/journal.pone.0231968}
 
@@
 107--129).
-New York, NY: Oxford University Press.
+Oxford University Press.
 
@@
 \leavevmode\hypertarget{ref-ekman2007directed}{}%
-Ekman, P. (2007). The directed facial action task. In J. A. Coan \& J. J. B. Allen (Eds.), \emph{Handbook of emotion elicitation and assessment} (pp. 47--53). New York, NY: Oxford University Press.
+Ekman, P. (2007). The directed facial action task. In J. A. Coan \& J. J. B. Allen (Eds.), \emph{Handbook of emotion elicitation and assessment} (pp. 47--53). Oxford University Press.
 
 \leavevmode\hypertarget{ref-ekman2017facial}{}%
-Ekman, P. (2017). Facial expressions. In J.-M. Fernández-Dols \& J. A. Russell (Eds.), \emph{The science of facial expression} (pp. 39--56). New York, NY: Oxford University Press.
+Ekman, P. (2017). Facial expressions. In J.-M. Fernández-Dols \& J. A. Russell (Eds.), \emph{The science of facial expression} (pp. 39--56). Oxford University Press.
 
@@
 Ekman, P., Friesen, W.
-V., O'sullivan, M., Chan, A., Diacoyanni-Tarlatzis, I., Heider, K., \ldots{} Tomita, M.
+V., O'sullivan, M., Chan, A., Diacoyanni-Tarlatzis, I., Heider, K., Krause, R., LeCompte, W.
+A., Pitcairn, T., Ricci-Bitti, S., Pio E, \& Tomita, M.
 (1987).
@@
 \leavevmode\hypertarget{ref-fridlund1994human}{}%
-Fridlund, A. J. (1994). \emph{Human facial expression: An evolutionary view}. San Diego, CA: Academic Press.
+Fridlund, A. J. (1994). \emph{Human facial expression: An evolutionary view}. Academic Press.
 
@@
 237--256).
-Amsterdam, NL: J.Benjamins.
+J.Benjamins.
 
@@
 78--102).
-New York, NY: Cambridge University Press.
+Cambridge University Press.
 
 \leavevmode\hypertarget{ref-gendron2018emotion}{}%
-Gendron, M., Hoemann, K., Crittenden, A. N., Msafiri, S., Ruark, G. A., \& Barrett, L. F. (2018). Emotion perception in hadza hunter gatherers. \url{https://doi.org/10.31234/osf.io/pf2q3}
+Gendron, M., Hoemann, K., Crittenden, A. N., Msafiri, S., Ruark, G. A., \& Barrett, L. F. (2018). \emph{Emotion perception in hadza hunter gatherers}. \url{https://doi.org/10.31234/osf.io/pf2q3}
 
@@
 \leavevmode\hypertarget{ref-hess2016emotional}{}%
-Hess, U., \& Fischer, A. H. (2016). \emph{Emotional mimicry in social context}. New York, NY: Cambridge University Press.
+Hess, U., \& Fischer, A. H. (2016). \emph{Emotional mimicry in social context}. Cambridge University Press.
 
@@
 Comprehensive database for facial expression analysis.
-In \emph{Proceedings fourth international conference on automatic face and gesture recognition} (pp.
-46--53).
+\emph{Proceedings Fourth International Conference on Automatic Face and Gesture Recognition}, 46--53.
 
 \leavevmode\hypertarget{ref-kappas2003facial}{}%
-Kappas, A. (2003). What facial activity can and cannot tell us about emotions. In M. Katsikitis (Ed.), \emph{The human face} (pp. 215--234). Boston, MA: Springer.
+Kappas, A. (2003). What facial activity can and cannot tell us about emotions. In M. Katsikitis (Ed.), \emph{The human face} (pp. 215--234). Springer.
 
@@
 70--80).
-New York, NY: Oxford University Press.
+Oxford University Press.
 
@@
 
+\leavevmode\hypertarget{ref-krumhuber2020human}{}%
+Krumhuber, E.
+G., Küster, D., Namba, S., \& Skora, L.
+(2020).
+Human and machine validation of 14 databases of dynamic facial expressions.
+\emph{Behavior Research Methods}, 1--16.
+\url{https://doi.org/https://doi.org/10.3758/s13428-020-01443-y}
+
 \leavevmode\hypertarget{ref-lee2002effect}{}%
@@
 
+\leavevmode\hypertarget{ref-lottridge2009emotional}{}%
+Lottridge, D., \& Chignell, M.
+(2009).
+Emotional bandwidth: Information theory analysis of affective response ratings using a continuous slider.
+\emph{IFIP Conference on Human-Computer Interaction}, 111--114.
+
 \leavevmode\hypertarget{ref-mcduff2016discovering}{}%
@@
 Discovering facial expressions for states of amused, persuaded, informed, sentimental and inspired.
-In \emph{Proceedings of the international conference on multimodal interaction} (pp.
-71--75).
+\emph{Proceedings of the International Conference on Multimodal Interaction}, 71--75.
 \url{https://doi.org/10.1145/2993148.2993192}
@@
 AFFDEX sdk: A cross-platform real-time multi-face expression recognition toolkit.
-In \emph{Proceedings of the conference on human factors in computing systems} (pp.
-3723--3726).
+\emph{Proceedings of the Conference on Human Factors in Computing Systems}, 3723--3726.
 \url{https://doi.org/10.1145/2851581.2890247}
@@
 \leavevmode\hypertarget{ref-meillon2010dynemo}{}%
-Meillon, B., Tcherkassof, A., Mandran, N., Adam, J.-M., Dubois, M., Dupré, D., \ldots{} Caplier, A.
+Meillon, B., Tcherkassof, A., Mandran, N., Adam, J.-M., Dubois, M., Dupré, D., Benoit, A.-M., Guérin-Dugué, A., \& Caplier, A.
 (2010).
 DynEmo: A corpus of dynamic and spontaneous emotional facial expressions.
-In \emph{Proceedings of international workshop series on multimodal corpora, tools and resources.
-Multimodal corpora: Advances in capturing, coding and analyzing multimodality} (pp.
-31--36).
+\emph{Proceedings of International Workshop Series on Multimodal Corpora, Tools and Resources.
+Multimodal Corpora: Advances in Capturing, Coding and Analyzing Multimodality}, 31--36.
 
@@
 361--375).
-New York, NY: Oxford University Press.
+Oxford University Press.
 
@@
 263--289).
-Paris, France: Odile Jacob.
+Odile Jacob.
 
@@
 \leavevmode\hypertarget{ref-sander2014oxford}{}%
-Sander, D., \& Scherer, K. (2014). \emph{Oxford companion to emotion and the affective sciences}. New York, NY: Oxford University Press.
+Sander, D., \& Scherer, K. (2014). \emph{Oxford companion to emotion and the affective sciences}. Oxford University Press.
 
@@
 \leavevmode\hypertarget{ref-de2019mama}{}%
-Waal, F. de. (2019). \emph{Mama's last hug: Animal emotions and what they teach us about ourselves}. London, UK: Granta Books.
+Waal, F. de. (2019). \emph{Mama's last hug: Animal emotions and what they teach us about ourselves}. Granta Books.
 
 \leavevmode\hypertarget{ref-wagner1997methods}{}%
-Wagner, H.
-L.
-(1997).
-Methods for the study of facial behavior.
-In J.
-A.
-Russell \& J.-M.
-Fernández-Dols (Eds.), \emph{The psychology of facial expression} (pp.
-31--54).
-Cambridge, UK: Cambridge University Press.
+Wagner, H. L. (1997). Methods for the study of facial behavior. In J. A. Russell \& J.-M. Fernández-Dols (Eds.), \emph{The psychology of facial expression} (pp. 31--54). Cambridge University Press.
 

### 2020-09-02 revision/prpf_1/revision_letter.Rmd
sent to anna

@@
 bibliography: "../../manuscript/bib/manuscript_ref.bib"
-output      : "papaja::revision_letter_pdf"
-#output      : word_document
+#output      : "papaja::revision_letter_pdf"
+output      : word_document
 ---
@@
 ```{r setup, include = FALSE}
+# warning ----------------------------------------------------------------------
+# by default "papaja::revision_letter_pdf" doesn't display code chunks when echo
+# is set to TRUE. To solve this problem, insert the following code in the 
+# revision_letter.tex document just below \documentclass:
+## $if(highlighting-macros)$
+## $highlighting-macros$
+## $endif$
+
 # libraries --------------------------------------------------------------------
@@
   replace(is.na(.), 0)
+
+unique_video_count <- human_recognition_data %>% 
+  dplyr::distinct(source, SEXE_Juge, C_Juge, C_Video) %>% 
+  dplyr::count(C_Video) %>% 
+  dplyr::summarise(
+    vid_n = n(),
+    vid_m = mean(n) %>% round(0), 
+    vid_sd = sd(n) %>% round(0)
+  )
+
+comparison_sr_hr <- self_report_score %>% 
+  dplyr::inner_join(human_recognition_score, by = "C_Video") %>% 
+  dplyr::ungroup() %>% 
+  dplyr::filter(hr_emotion != "undetermined") %>% 
+  dplyr::filter(sr_emotion != "undetermined") %>% 
+  dplyr::mutate_if(is.character, as.factor)
+
+confusionMatrix_sr_hr <- 
+  caret::confusionMatrix(
+    data = comparison_sr_hr$hr_emotion,
+    reference = comparison_sr_hr$sr_emotion
+  )
+
+comparison_sr_ar <- self_report_score %>% 
+  dplyr::inner_join(automatic_recognition_score, by = "C_Video") %>% 
+  dplyr::ungroup() %>% 
+  dplyr::filter(ar_emotion != "undetermined") %>% 
+  dplyr::filter(sr_emotion != "undetermined") %>% 
+  dplyr::mutate_if(is.character, as.factor)
+
+confusionMatrix_sr_ar <- 
+  caret::confusionMatrix(
+    data = comparison_sr_ar$ar_emotion,
+    reference = comparison_sr_ar$sr_emotion
+  )
+
+corr_hr_ar <- 
+  cor.test(
+    recognition_result$hr_score, 
+    recognition_result$ar_score, 
+    method = "pearson",
+    alternative = "two.sided"
+  ) %>% 
+  papaja::apa_print()
+
+self_report_gender <- self_report_data %>% 
+  dplyr::count(genre_c) %>% 
+  dplyr::pull(n, genre_c)
+
+self_report_age <- self_report_data %>%
+  dplyr::summarise(
+    age_m = mean(age) %>% round(1), 
+    age_sd = sd(age) %>% round(1)
+  )
 ```
@@
 
-\RC{Dear Dr Tcherkassof,
+\\RC{Dear Dr Tcherkassof,
 
@@
 
-\RC{The work of this article gives new insights about the connection of emotional self-reports and observer categorizations and automatic recognition of facial movements.
+\\RC{The work of this article gives new insights about the connection of emotional self-reports and observer categorizations and automatic recognition of facial movements.
 The data are very interesting and this work should be published.
@@
 
-\RC{Page 15, line 57: in 2-3 or more lines should be described which kind of emotion inductions have been used for the data base.}
+\\RC{Page 15, line 57: in 2-3 or more lines should be described which kind of emotion inductions have been used for the data base.}
 
-\RC{Page 16 line 39: why did the annotators only rate 232 of the 358 videos?
+As requested a short presentation of emotion elicitation task has been added as follow:
+
+> Two different types of task were used to elicit emotions.
+Some tasks were passive and consisted in the participant watching videos selected to trigger emotions (e.g.
+television commercials).
+Other tasks were active; they required the participant to interact with a computer (e.g.
+to answer questions or to test a flawed software).
+
+\\RC{Page 16 line 39: why did the annotators only rate 232 of the 358 videos?
 And how have been the videos selected?
@@
 
+The annotation process was quite complicated.
+Because the total duration of the database is XXX min, we could not ask each participant to annotate it entirely.
+In addition we had to insure that each recording is annotated by at least 20 participants to obtain an sufficient agreement.
+Therefore, we have randomly selected a batch of 10% of the recordings to be annotated until all the videos are annotated at least 20 times.
+Then, another batch of 10% of the database was randomly selected.
+However, to avoid their cognitive fatigue, we have limited participant's experiement duration to 30min.
+In order to add this element and to make it clearer, the annotation procedure has been completely rewritten as follow:
+
+> To analyse how the recorded facial expressions are perceived, `r nrow(unique_ppt_count)` participants were recruited to watch and to annotate these recordings.
+An iterative procedure was setup to insure that each recording has been annotated by at least 20 participants: 10% of the video database were randomly selected and annotated until all the videos reached the amount of annotation required, then another section of the database is randomly selected among the remaining recordings to be annotated.
+To avoid potential decrease in the quality of the annotations due to cognitive fatigue, each participant only annotated recordings during a maximum of 30min.
+As a results only `r unique_video_count["vid_n"]` out of the `r nrow(self_report_data)` videos have been annotated.
+Each video was annotated `r unique_video_count["vid_m"]` times on average (*SD* = `r unique_video_count["vid_sd"]`).
+
 ## Results section
 
-\RC{Page 19 line 17 to 20: Which kind of correlation was computed here? A pearson correlation? It should be a point-biserial correlation as one of the two variables is dichotomous.}
+\\RC{Page 19 line 17 to 20: Which kind of correlation was computed here? A pearson correlation? It should be a point-biserial correlation as one of the two variables is dichotomous.}
+
+It appears that point-biserial correlation for two dichotomous variables is mathematically equivalent to calculating the Pearson's product-moment correlation and produces the same results [@bass1991correcting].
+
+The following reproducible example in R shows that results are identical between point-biserial correlation using the function `biserial.cor()` from the package {ltm} and Pearson's product-moment correlation using the function `cor.test()` from the package {stats}:
+
+```{r echo=TRUE}
+x <- c(0, 1, 0, 1, 0, 1, 0, 1, 0 , 1)
+y <- c(0, 0, 0, 0, 0, 1, 1, 1, 1 , 1)
+
+ltm::biserial.cor(x, y, level = 2)
+
+cor.test(x, y, method = "pearson", alternative = "two.sided")$estimate
+```
+
+The following example tests the computation of the point-biserial correlation and the Pearson's product-moment correlation with our data (i.e.
+correlation between participant's self-report and observers' annotation):
 
@@
 
-\RC{Page 21 line 32: "expression intensity", where are expression intensity scores? Could you please bring the data into the paper?}
+As the two functions are providing the exact same result we can insure the integrity of our results to the reviewers using the Pearson's product-moment correlation.
+However, as our previous manuscript only indicate Pearson's correlation, the new manuscript now mentions Pearson's product-moment correlation in order to disambiguate this point.
 
-\RC{Page 21 line 40ff: Why should undetermined emotional states reveal that a 6-point Likert scale has a limit? What kind of limit? What would be better to do in future? Please explain in the text.}
+\\RC{Page 21 line 32: "expression intensity", where are expression intensity scores? Could you please bring the data into the paper?}
+
+In the previous version of the manuscript, "expression intensity" was referring to a result usually obtained in facial expression recognition studies [see the recent @krumhuber2020human for example].
+However this assumption is just an interpretation and is not supported by data.
+In order to correct this point, the sentence has been rewritten as follow:
+
+> This difference between emotions is usually observed in the literature [e.g.
+@krumhuber2020human].
+Multiple reasons can explain why *happiness* and *disgust* are more easily recognized than *anger*, *surprise* and *sadness*.
+It is possible that these expressions are usually more intense than others.
+In addition, *anger* and *sadness* as non-socially desirable emotions may be have been felt but not expressed.
+
+\\RC{Page 21 line 40ff: Why should undetermined emotional states reveal that a 6-point Likert scale has a limit? What kind of limit? What would be better to do in future? Please explain in the text.}
+
+Many thanks, indeed this point requires some explanations.
+"undetermined" is used when more than one label obtain the highest score.
+As there is a tie between two or more label, it is not possible to identify a dominant one.
+In term of affective states, two emotions can be felt simultaneously with the same intensity.
+However, if one emotions is felt with a lower intensity, it is possible that likert scales are not discriminant enough with 6-point and results in an "undetermined" labels even if the two dominant emotions are not felt with the same intensity.
+To better discriminate between labels, an alternative is to use a continuous slider from 0 to 100.
+This point has been added to the new version of the manuscript as follow:
+
+> "Indeed, Likert scales may not be able to discriminate between two dominant emotion felt with different intensities due to their reduced number of possibilities.
+To better discriminate between emotions, an alternative would be to use a continuous slider from 0 to 100 [see for example @lottridge2009emotional]."
 
@@
 
-\RC{Page 4 Line 44/45: one would not assume total agreement among all people based on basic emotion theory, so "totally" should be changed by another term like "mainly" or something similar.}
+\\RC{Page 4 Line 44/45: one would not assume total agreement among all people based on basic emotion theory, so "totally" should be changed by another term like "mainly" or something similar.}
 
-\RC{Page 5 Line 16/17: Isn´t the point of Kraut and Jonston that according Basic Emotion Theory the smile is only a sign of expression of joy? So I would write: "that a smile is only the major".}
+The end of the sentence has been rewritten as follow:
 
-\RC{Page 5 line 58 ff: maybe smiling is not the right sign to be looked directly after winning a contest.
+> "... and why observers can make mistakes in recognizing facial expressions of basic emotions, among others."
+
+\\RC{Page 5 Line 16/17: Isn´t the point of Kraut and Jonston that according Basic Emotion Theory the smile is only a sign of expression of joy? So I would write: "that a smile is only the major".}
+
+The sentence has been changed as suggested.
+
+\\RC{Page 5 line 58 ff: maybe smiling is not the right sign to be looked directly after winning a contest.
 Triumpf is usually exposed immediately after the contest and other emotions like joy or pride come a little later (see the work of Matsumoto and colleagues on that).
@@
 
-\RC{Page 18: line 56: "two-sided pearson correlation" probably means two-tailed t-tests of pearson-correlations, or not?}
+To see with Anna
 
-\RC{Page 20 Line25: "60% and 80% accuracy"; please give a citation.}
+\\RC{Page 18: line 56: "two-sided pearson correlation" probably means two-tailed t-tests of pearson-correlations, or not?}
 
-\RC{Page 23 Line 56: what is correlated here (r=.22)? A correlation of two accuracies does not make sense in my mind. If yes, what are the units? Emotion categories, events or…???}
+Many thanks for pointing our mistake, it is indeed a "two-tailed t-tests of pearson-correlations" and it has been changed in the manuscript.
 
-\RC{Page 31 line 9ff: I do not understand the point under secondly, automatic classifiers assume for every person unique facial expression for one emotion??
+\\RC{Page 20 Line25: "60% and 80% accuracy"; please give a citation.}
+
+This information has been precised with a reference to @krumhuber2020human as follow:
+
+> "These results are far from those classically obtained in the literature for emotional facial expression recognition [@krumhuber2020human, report an average accuracy of 65%]."
+
+\\RC{Page 23 Line 56: what is correlated here (r=.22)? A correlation of two accuracies does not make sense in my mind. If yes, what are the units? Emotion categories, events or…???}
+
+Our formulation was indeed incomplete.
+For each of the video and emotion categories, all three of self-report, human observers and automatic recognition are encoded as 1 for the emotion category recognised and 0 otherwise.
+Therefore the correlation corresponds to the correlation between human observers and automatic classifier recognition of emotion categories.
+This element has been added to the manuscript as follow:
+
+> "As previously mentioned, human observers appear to be more accurate than the automatic classifier to recognize an individual's subjective feeling (human observers Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`; automatic classifier Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`).
+However, both make similar mistakes as the two-tailed t-tests of Pearson's product-moment correlation between human observers and automatic classifier recognition of emotion categories is significant (`r corr_hr_ar$full_result`)."
+
+\\RC{Page 31 line 9ff: I do not understand the point under secondly, automatic classifiers assume for every person unique facial expression for one emotion??
 Did I understand that right?
@@
 
+The output of the automatic classifier is a multivariate timeseries matrix with, for every labels, a probability of expressing the corresponding label for each video frame.
+Therefore, the equation used to identify which label corresponds the most to the overall video is crucial.
+In order to make that point clearer, the following sentences have been added to the method section:
+
+> "For each video frame, Affdex identifies the probability $p_{video_{i}.label_{j}.t_{k}}$ of expressing each of the six basic emotion labels (*i.e.*, *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as additional psychological states such as *valence*, *engagement* or *contempt*, and facial features such as *cheek raise*, *eye widen* or *jaw drop*.
+The result of Affdex is a mutivariate timeseries output which provides for each label a probability from 0 to 100 (rescaled from 0 to 1 for the analysis) of expressing the corresponding label.
+For this study only the six basic emotion labels are analysed to match with self-reports and human observers."
+
 # Reviewer \#2
 
-\RC{1. The paper did not distinguish clearly between opinion and empirical evidence. Please revise.}
+\\RC{1. The paper did not distinguish clearly between opinion and empirical evidence. Please revise.}
 
-\RC{2. The paper's arguments must be built on an appropriate base of theory, concepts, or other ideas. It is recommended that the authors refer to Eigen Face Approach. }
+See with Anna
 
-\RC{3. The authors did not make clear emotion taxonomy. Ekman's basic set of emotions and Russell's circumflex model of affect are recommended to elaborate.}
+\\RC{2. The paper's arguments must be built on an appropriate base of theory, concepts, or other ideas. It is recommended that the authors refer to Eigen Face Approach. }
 
-\RC{4.  What is the significance of this study? must be explained with a sensible reason.}
+The Eigen Face Approach is used in the development of the automatic classifier's algorithm to maximise emotion classification accuracy [@turk1991eigenfaces; @cendrillon2000real].
+Here, the automatic classifier has already been developed and our team is not involved in Affdex's development.
+Consequently, it appears that introducing the theory and concepts behind the Eigen Face Approach is not relevant in the context of this study.
 
-\RC{5. The paper must contribute to a critical understanding of the issues.}
+\\RC{3. The authors did not make clear emotion taxonomy. Ekman's basic set of emotions and Russell's circumflex model of affect are recommended to elaborate.}
 
-\RC{6.
+See with Anna
+
+\\RC{4.  What is the significance of this study? must be explained with a sensible reason.}
+
+See with Anna
+
+\\RC{5. The paper must contribute to a critical understanding of the issues.}
+
+See with Anna
+
+\\RC{6.
 You're highly advised to show yourselves as authors while explaining theme of research in the paper.
@@
 
-\RC{7. The selection of the participants, to me, is vague.}
+See with Anna
 
-\RC{8. I wonder why the authors did not implement Principal Component Analysis.}
+\\RC{7. The selection of the participants, to me, is vague.}
 
-\RC{9.
+In order to precise the procedure used for the selection of participants, the following sentences have been added to the "Emotion Elicitation" section, and to the "Human Facial Expression Recognition" section:
+
+> "For the emotion elicitation experiment, `r nrow(self_report_data)` encoding participants (`r self_report_gender["F"]` females, `r self_report_gender["H"]` males, *M*~age~ = `r self_report_age["age_m"]`, *SD*~age~ = `r self_report_age["age_sd"]`) were recruited to perform one out of 11 emotion elicitation tasks designed to trigger a positive, a specific negative or a neutral emotional state.
+These participants were recruited by a private company for a study supposedly devoted to an "ergonomic visual task" (cover story).
+After a description of the general aims, participants agreed and signed the experiment consent form.
+At the end of the elicitation task, they received an equivalent of €50 in voucher for their participation.
+A second consent form was signed by the participants to allow their video to be processed for research purposes."
+
+> "To analyse how the recorded facial expressions are perceived, `r nrow(unique_ppt_count)` participants were recruited among social science under- and post-graduates through advertising to watch and to annotate these recordings.
+They received a course credit for their participation.
+After a description of the general aims, participants agreed and signed the experiment consent form.
+"
+
+\\RC{8. I wonder why the authors did not implement Principal Component Analysis.}
+
+PCA are used to infer one or multiple latent constructs from a series of variables.
+Here, the purpose of this paper is to investigate the relationship between self-reported emotions and emotion recognised by human observers and by an automatic classifier.
+Consequently, an approach based on correlations appeared to be more relevant than a PCA.
+
+\\RC{9.
 Discussion needs a deeper look.
@@
 
-\RC{10. It is recommended that authors consider APA, 7th edition in in-text citations. }
+See with Anna
+
+\\RC{10. It is recommended that authors consider APA, 7th edition in in-text citations. }
+
+As requested the in-text citations now uses APA, 7th edition reference style.
 

### 2020-11-04 manuscript/manuscript_apa.Rmd
revision 2

@@
     email         : "anna.tcherkassof@univ-grenoble-alpes.fr"
+    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
+      - Conceptualization
+      - Project administration
+      - Validation
+      - Writing - Original Draft Preparation
+      - Writing - Review & Editing
   - name          : "Damien Dupré"
     affiliation   : "2"
+    role:
+      - Investigation
+      - Data curation
+      - Formal Analysis
+      - Software
+      - Visualization
+      - Writing - Original Draft Preparation
+      - Writing - Review & Editing
 
@@
 
+authornote: |
+  We thank anonymous reviewers for their insightful comments on a previous version of the manuscript.
+
 abstract: |
   While it has been taken for granted in the development of several automatic facial expression recognition tools, the question of the coherence between subjective feelings and facial expressions is still a subject of debate.
-On one hand, the "Basic Emotion View" conceives emotions as genetically hardwired and therefore being genuinely displayed through facial expressions.
-On the other hand, the constructivist approach conceives emotions as socially constructed; the emotional meaning of a facial expression being inferred by the observer.
-In order to evaluate the coherence between the subjective feeling of emotions and their recognition based on facial expression, 232 videos of encoders recruited to carry out an emotion elicitation task were annotated by 1383 human observers as well as by an automatic facial expression classifier.
-Results show low accuracy of human observers and of the automatic classifier to infer the subjective feeling from the facial expressions displayed by encoders.
-They also show a weak consistency between self-reported emotional states and facial emotional displays.
+On one hand, the “Basic Emotion View” conceives emotions as genetically hardwired and therefore being genuinely displayed through facial expressions.
+Consequently, emotion recognition is perceiver-independent.
+On the other hand, the constructivist approach conceives emotions as socially constructed, the emotional meaning of a facial expression being inferred by the perceiver.
+Hence, emotion recognition is perceiver-dependent.
+In order 1) to evaluate the coherence between the subjective feeling of emotions and their spontaneous facial displays, and 2) to compare the recognition of such displays by human perceivers and by an automatic facial expression classifier, 232 videos of expressers recruited to carry out an emotion elicitation task were annotated by 1383 human perceivers as well as by Affdex, an automatic classifier.
+Results show a weak consistency between self-reported emotional states by expressers and their facial emotional displays.
+They also show low accuracy both of human perceivers and of the automatic classifier to infer the subjective feeling from the spontaneous facial expressions displayed by expressers.
+However, the results are more in favor of a perceiver-dependent view.
 Based on these results, the hypothesis of genetically hardwired emotion genuinely displayed is difficult to support, whereas the idea of emotion and facial expression as being socially constructed appears to be more likely.
@@
 
-keywords          : "emotion, facial expression, self-report, human observer, automatic recognition."
+keywords          : "emotion, facial expression, self-report, human perceiver, automatic recognition."
+
 wordcount         : "7943"
-
 bibliography      : ["bib/manuscript_ref.bib"]
@@
 ---
-(ref:confusionMatrix-sr-hr) Confusion matrix between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the human observers.
+(ref:confusionMatrix-sr-hr) Confusion matrix between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the human perceivers.
 (ref:confusionMatrix-sr-ar) Confusion matrix of between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the automatic classifier.
-(ref:confusionMatrix-hr-ar) Proportion of emotion labels classified by human observers which are recognized by the automatic classifier.
+(ref:confusionMatrix-hr-ar) Proportion of emotion labels classified by human perceivers which are recognized by the automatic classifier.
 (ref:confusionTable-sr-hr) Human recognition accuracy metrics for each emotion.
@@
 ```{r data}
-## self_report_data
+## self_report_data ------------------------------------------------------------
 source(here::here("script/data_wrangling_self_report.R"))
@@
 
-## human_recognition_data
+## human_recognition_data ------------------------------------------------------
 source(here::here("script/data_wrangling_human_recognition.R"))
@@
 
-## automatic_recognition_data
+## automatic_recognition_data --------------------------------------------------
 source(here::here("script/data_wrangling_automatic_recognition.R"))
@@
 With the development of commercial automatic facial expression recognition tools [@dupre2018accuracy; @dupre2020performance], industries and governments are gradually implementing this technology in order to monitor humans’ emotions in various scenarios (*e.g.*, in marketing, healthcare, and the automotive industry to name a few).
-This technology rests on the premise that facial expressions provide direct access to individuals’ subjective feelings, and that one can read the emotions displayed on a person’s face as one would an open book.
-This premise is derived from the "Basic Emotion View", a theory which suggests a one-to-one mapping between subjective feeling and facial expression [@ekman1987universals; @ekman1988universality; @ekman1992argument; @ekman2007directed].
-Even if this premise is central to the modern mainstream approach of human emotion, recent research in affective science is challenging it.
-Once the Basic Emotion View is briefly described, its foremost criticisms will be exposed.
-Inconsistencies and unpredicted findings come both from field observations and laboratory experiments.
-They relate to two different issues.
-The first issue is the sender’s production of spontaneous facial displays.
-The second issue relates to the receiver’s interpretation of these facial displays.
-Building on this evidence, and as an attempt to shed further light on the emotion–face link, the present study's main goal is to tackle these two issues at the same time.
-It aims at identifying the extent to which ordinary people’s experienced emotions are displayed through identifiable Emotional Facial Expressions (EFE) recognized as such both by human individuals, and by automatic facial expression recognition tools.
+This technology rests on the premise that emotional state is revealed by the measurement of facial features thanks to a one-to-one mapping between subjective feeling and facial expression.
+This premise comes from the Basic Emotion View [@ekman1987universals; @ekman1988universality; @ekman1992argument; @ekman2007directed].
+By contrast, for the constructivist approach [barrett2017emotions; @gendron2017facing], emotions being socially constructed, a making-meaning process takes part in the link between emotion and facial expression.
+Emotion recognition is considered as resulting from the perceiver making inferences about the expresser’s underlying emotions.
+The constructivist approach is therefore a perceiver-*dependent* view whereas the Basic Emotion View is a perceiver-*independent* view [@gendron2017facing].
+In the emotion–facial expression link contention, the issue of facial expressions’ naturalness must also be considered.
+Since automatic recognition tools are meant to be used in real-life conditions, spontaneous displays of emotional feelings are the heart of the matter as well as their accurate recognition.
+The potential significance of the present study is that it aims at providing data, not only on the emotion–facial expression link, but also on the perceiver-independent *vs.* perceiver-dependent question of emotion recognition as regard to spontaneous facial expressions of self-reported emotions.
+It looks at the triangulation emotional feeling—spontaneous facial display—emotion recognition within a single paradigm comparing the recognition accuracy between human and machines.
 
@@
 
-The belief that facial expression is linked to emotional states can be traced back to Darwin in *The Expression of the Emotions in Man and Animals* [@darwin1872expression].
-Countless studies have fortified the idea that emotional states are inherently coupled to a set of prototypic facial expressions [*e.g.*, @ekman2011meant].
+Since Darwin’s book *The Expression of the Emotions in Man and Animals* [@darwin1872expression], countless studies have fortified the idea that emotional states are inherently coupled to a set of prototypic facial expressions [*e.g.*, @ekman2011meant].
 The Basic Emotion View holds that facial expressions are genuine displays of an individual’s inner emotional state.
-More specifically, a set of six emotions (*fear*, *anger*, *surprise*, *disgust*, *sadness* and *joy*) are universally displayed and are genetically hardwired not only in humans [@ekman1992argument], but also in different animal species [@de2019mama].
+More specifically, a set of six so-called basic emotions (*fear*, *anger*, *surprise*, *disgust*, *sadness* and *happiness*) are universally displayed and are genetically hardwired not only in humans [@ekman1992argument], but also in different animal species [@de2019mama].
 According to the Basic Emotion View, *“when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including […] facial expressions.”* [@ekman2007directed, 49].
 To respond to criticisms, several amendments have been made to the Basic Emotion View, increasing the number of basic emotions from six to seven [@ekman1988universality] as well as adding the concept of “display rules” to explain cultural differences in the management of facial expressions [@ekman1987universals].
+The most recent versions of the Basic Emotion Theory even distinguish some twenty expressions of emotions, including love, pride, embarrassment, shame, among others [@keltner2019emotional].
 
-Relatively few studies have actually tested Basic Emotion View’s fundamental claim regarding the facial production (so-called "expression") of emotions.
+Relatively few studies have actually tested Basic Emotion View’s fundamental claim regarding the facial production (or “expression”) of emotions.
 @rosenberg1994coherence reported the first evidence of coherence between self-report of emotion and displayed facial expressions.
@@
 Notwithstanding this study, few have used purely descriptive methods such as electromyography or objective face coding systems to identify and to measure the actual changes in the face when a given emotion is felt [see @wagner1997methods, for methodological points].
-Much of the research has focused instead on the recognition of EFE, that is on the issue of the viewer’s interpretation of facial displays.
-As people have been thought to display their internal state, EFE supposedly serve as a window into the emotions of others.
-Viewers “recognize” the facial displays and infer the corresponding emotional state.
-The Readout Hypothesis [@buck1985prime] formalizes this folk theory.
-According to it, facial expressions have evolved to provide *“an external readout of those motivational-emotional processes that have had social implications during the course of evolution”* [@buck1985prime, 396-397].
+
+Much of the research has focused instead on the recognition of emotional facial expressions (EFE), that is on the issue of the perceiver’s interpretation of facial displays.
+According to the Readout Hypothesis (Buck, 1985) which formalizes the basic Emotion View, facial expressions have evolved to provide *“an external readout of those motivational-emotional processes that have had social implications during the course of evolution”* [@buck1985prime, 396-397].
 Thus, as readouts of emotional states, spontaneous expressive displays are directly accessible to other organisms.
-Research on the inference of emotions from facial expressions has established that viewers show considerable agreement that the so-called basic emotions—happiness, anger, fear, disgust, sadness, and surprise—are associated with specific facial displays [*e.g.*, reviews from @elfenbein2002universality; @russell1994there].
+Research on the inference of emotions from facial expressions has established that perceivers show considerable agreement that the so-called basic emotions are associated with specific facial displays [*e.g.*, reviews from @elfenbein2002universality; @russell1994there].
 
-Yet, in spite of the popular support it has received, the empirical data called upon by the Basic Emotion View remains unpersuasive.
-Notably, it fails to explain, in instances in which display rules cannot be called upon, how individuals can feel emotions without expressing them or how individuals can express emotions without feeling them, and why observers can make mistakes in recognizing facial expressions of basic emotions, among others.
-The following  is a summary of key findings and conclusions both from field observations and laboratory experiments on spontaneous expression of emotions, all of them strongly subverting the Basic Emotion viewpoint [for reviews, @fernandez2013emotion; @reisenzein2013coherence; @russell2003facial].
-Mains issues regarding recognition studies are also briefly discussed afterwards.
+Yet, the empirical data called upon by the Basic Emotion View remains unpersuasive.
+Notably, it fails to explain, in instances in which display rules cannot be called upon, how individuals can feel emotions without expressing them or how individuals can express emotions without feeling them, and why perceivers can make mistakes in recognizing facial expressions of basic emotions, among others.
+All these findings subverting the Basic Emotion viewpoint come from field observations and laboratory experiments on spontaneous expression of emotions [for reviews, @fernandez2013emotion; @reisenzein2013coherence; @russell2003facial].
+The key findings are summarized below as well as the main issues regarding recognition studies.
 
@@
 Though they cannot be considered as a straightforward test of the triggering role of emotion on facial behavior, they have the advantage of considering situations that cannot be created in a laboratory.
-For instance, the ethological study of @kraut1979social contradicts the Basic Emotion View premise that a smile is only the major component of a facial display associated with and caused by feelings of joy or happiness.
+For instance, the ethological study of Kraut and @kraut1979social contradicts the Basic Emotion View premise that a smile is only the major component of a facial display associated with and caused by feelings of happiness.
 Naturalistic observation at a bowling alley showed that bowlers do not necessarily smile after scoring a spare or a strike (a situation likely to elicit a positive emotion).
@@
 More specifically, winners showed Duchenne smiles and other types of smiles when they were interacting with other people during the awards ceremony, but smiles were scarce or nonexistent when waiting behind the podium and/or when turning toward the flagpoles and focusing their attention on the flags and the national anthem.
-@crivelli2015smiles also found that the strongest predictor for the occurrence of Duchenne smiles in judo winners is when they are engaged in social interaction and not when they just won their match.
+Other naturalistic studies have also found that happiness cannot be considered the strongest predictor for the occurrence of Duchenne smiles [@crivelli2015smiles].
 
-Happiness/joy is not the only emotion weakly associated with the predicted facial expression.
+Happiness is not the only emotion weakly associated with the predicted facial expression.
 A naturalistic study conducted by @scherer1997lost in an airport’s baggage handling office showed that passengers claiming for their lost luggage displayed very few facial expressions of negative emotions while self-reporting subjective feeling states of anger or sadness among others.
@@
 Camras and her colleagues observed that facial expressions of negative emotions were displayed in circumstances that were unlikely to have elicited those emotions.
-For instance, infants displayed "fear" expressions in settings not related to fear [@camras1991development].
+For instance, infants displayed “fear” expressions in settings not related to fear [@camras1991development].
 @bennett2002facial videotaped a sample of 4-month-old infants during tickle, sour taste, jack-in-the-box, arm restraint, and masked stranger situations.
@@
 As an example, a strong disconfirmation of the Basic Emotion View’s premise is put forward by Reisenzein’s studies on surprise [*e.g.*, @reisenzein2000exploring].
-In eight controlled laboratory situations, surprise was induced by @reisenzein2006evidence by establishing and then invalidating a set of beliefs concerning the experimental events such as the unexpected appearance of a picture of one's own face as the last picture in a series of portraits that had to be rated.
+In eight controlled laboratory situations, surprise was induced by @reisenzein2006evidence by establishing and then invalidating a set of beliefs concerning the experimental events such as the unexpected appearance of a picture of one’s own face as the last picture in a series of portraits that had to be rated.
 Visible or electromyography-detected facial displays of surprise occurred only in few participants.
@@
 Many people strongly agree that so-called basic emotions are associated with specific facial configurations [@ekman2017facial; @elfenbein2002universality] and also argue that this is strong evidence for the Basic Emotion View.
-Moreover, it implies that EFE recognition of human observers should be as accurate as automatic classifiers'.
+Moreover, it implies that EFE recognition of human perceivers should be as accurate as automatic classifiers.
 
@@
 Facial stimuli used in experiments also constitute a methodological bias because they are unrepresentative of ordinary facial expressions.
-Basic Emotion View empirical evidence is based for the most part on methods using a static and unnatural material, namely, still photographs of posed facial expressions of emotion (*e.g.*, intentionally encoded by the sender).
+Basic Emotion View empirical evidence is based for the most part on methods using a static and unnatural material, namely, still photographs of posed facial expressions of emotion (*e.g.*, intentionally encoded by the expresser).
 This kind of methodology raises questions about its ecological validity and the generalizability of the results to real interpersonal emotional communication [*e.g.*, @tcherkassof2007facial].
@@
 @hess1990differentiating have also pointed out the importance of the dynamics of facial movements, and particularly the irregularity, or phasic changes, of the expressions’ unfolding.
-Thus, the motion of facial expression provides observers with other information than the one provided by static expressions.
+Thus, the motion of facial expression provides perceivers with other information than the one provided by static expressions.
 It may be that differences in the social information displayed by static and dynamic expressions leads to facial recognition differential effects.
@@
 Facial expressions are actions in a communicative interaction.
-They do not express emotions but they *"prompt, on the receiver’s side, important inferences about the context, the sender, and the course of the interaction between sender and receiver"* [@dols2017natural, 466].
+They do not express emotions but they *"prompt, on the receiver’s side, important inferences about the context, the sender [expresser], and the course of the interaction between sender [expresser] and receiver"* [@dols2017natural, 466].
 As such, the fact that facial displays are able to signal emotions is a byproduct of one of their main functions; implementing actions performing practical ends.
-Therefore, expressive displays hold a motor intention [@pacherie2003modes].
+Therefore, a certain number of expressive displays (*e.g.*, surprise, disgust, fear, anger) hold a motor intention [@pacherie2003modes].
 They implement their aim which is their motor intention.
@@
 Facial displays are parts of pragmatic actions aiming at orienting the person’s relation to their environment.
-For instance facial displays are maintaining, breaking or restoring the relationship between sender and perceiver [@frijda2012recognition].
+For instance, facial displays are maintaining, breaking or restoring the relationship between expresser and perceiver [@frijda2012recognition].
 Facial displays are not recognized in semantic terms but are perceived as intentional actions.
@@
 The constructivist approach claims that facial displays are behaviors whose meaning is inferred by perceivers.
-Findings support this observer dependence [@lindquist2013s; niedenthal2017embodied].
+Findings support this perceiver dependence [@lindquist2013s; @niedenthal2017embodied].
 They show that to make meaning of another person’s facial behavior, the perceiver relies in particular on her/his knowledge about emotion categories.
@@
 They did not infer inner states, rather they proceeded with action identification that pointed out the functions of behaviors [see also @crivelli2017recognizing for similar observations among a small-scale society of Papua New Guinea].
-The constructivist approach considers that specific emotion categories, as conceptualized by Western cultures’ knowledge, are cast on the perceived face to make meaning of the sender’s facial displays.
+The constructivist approach considers that specific emotion categories, as conceptualized by Western cultures’ knowledge, are cast on the perceived face to make meaning of the expresser’s facial displays.
 Following this approach, faces convey a range of information essential for social communication.
@@
 The present study aims to examine these predictions in order to provide empirical evidence to allow the discussion to evolve.
-To date, no systematic study has looked at facial expressions spontaneously displayed in reaction to emotional triggers and how they are decoded, both by human observers and by automatic emotion recognition tools based on the detection of facial muscular configurations.
+To date, no systematic study has looked at facial expressions spontaneously displayed in reaction to emotional triggers and how they are decoded, both by human perceivers and by automatic emotion recognition tools based on the detection of facial muscular configurations.
 This study fills that gap.
@@
 More specifically, this study aims to examine the recognition of EFE produced by ordinary people during situations judged and/or self-reported to involve different emotions.
-It (a) examines consistency between ordinary people’s self-reported emotional experience and observers’ judgments of these ordinary people’s EFE, and (b) examines consistency between ordinary people’s self-reported emotional experience and an automatic classifier’s analysis of these ordinary people’s EFE.
-In other words, it is interested in how people actually move their faces to express self-reported emotions, in how human observers accurately infer the expresser’s emotional state, and in how automatic recognition accurately codes the expresser’s emotional state.
-Because of their superior ability to exactly recognize EFE, it is expected that human observers’ will have a higher accuracy to identify senders' subjective feeling than automatic EFE recognition tools.
+It (a) examines consistency between ordinary people’s self-reported emotional experience and perceivers’ judgments of these ordinary people’s EFE, and (b) examines consistency between ordinary people’s self-reported emotional experience and an automatic classifier’s analysis of these ordinary people’s EFE.
+In other words, it is interested in how people actually move their faces to express self-reported emotions, in how human perceivers accurately infer the expresser’s emotional state, and in how automatic recognition accurately codes the expresser’s emotional state.
+Because of their superior ability to exactly recognize EFE, it is expected that human perceivers’ will have a higher accuracy to identify expressers' subjective feeling than automatic EFE recognition tools.
 
@@
 
-To evaluate the consistency between subjective feeling of emotions and their recognition from facial expressions, encoders were first recruited to perform an emotion elicitation task while their facial expression was video recorded.
-In order to reduce the likelihood of facial control, the encoders were alone in the room and were filmed by hidden cameras, so they had no reason to comply with social display rules.
-Then, the videos of the encoders’ faces were shown to human observers and were also analyzed by an automatic classifier in order to identify which emotion was displayed.
+To evaluate the consistency between subjective feeling of emotions and their recognition from facial expressions, expressers were first recruited to perform an emotion elicitation task while their facial expression was video recorded.
+In order to reduce the likelihood of facial control, the expressers were alone in the room and were filmed by hidden cameras, so they had no reason to comply with social display rules.
+Then, the videos of the expressers’ faces were shown to human perceivers and were also analyzed by an automatic classifier in order to identify which emotion was displayed.
 
@@
 to answer questions or to test a flawed software).
-Encoders’ faces were recorded using a hidden camera resulting `r nrow(self_report_data)` front facing 768x576 videos varying from `r min(metadata_video$ffprobe_duration)`s to `r max(metadata_video$ffprobe_duration)`s.
+Expressers’ faces were recorded using a hidden camera resulting `r nrow(self_report_data)` front facing 768x576 videos varying from `r min(metadata_video$ffprobe_duration)`s to `r max(metadata_video$ffprobe_duration)`s.
 These recordings form the DynEmo database [@tcherkassof2013dynemo].
 
-After the emotion elicitation task, the encoders rated their subjective feeling on Likert scales from 0 ("not at all") to 5 ("strongly") related to six "basic" emotion labels (*i.e.*, *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as six "non-basic" emotion labels (*i.e.*, *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
+After the emotion elicitation task, the expressers rated their subjective feeling on Likert scales from 0 ("not at all") to 5 ("strongly") related to six "basic" emotion labels (*i.e.*, *anger*, *disgust*, *fear*, *happiness*, *surprise* and *sadness*) as well as six "non-basic" emotion labels (*i.e.*, *pride*, *curiosity*, *boredom*, *shame*, *humiliation*, and *disappointment*).
 
-Finally, a debriefing session was performed to ensure that encoders were not durably affected by the emotion elicitation task.
-The debriefing was also used to check that encoders did not guess the real purpose of the experiment (*e.g.*, being filmed while they were performing an emotional elicitation task) to guarantee facial expressions’ genuineness.
-All encoders gave their agreement on their data and recordings being processed for research purposes only.
+Finally, a debriefing session was performed to ensure that expressers were not durably affected by the emotion elicitation task.
+The debriefing was also used to check that expressers did not guess the real purpose of the experiment (*e.g.*, being filmed while they were performing an emotional elicitation task) to guarantee facial expressions’ genuineness.
+All expressers gave their agreement on their data and recordings being processed for research purposes only.
 
@@
 
-This annotation procedure resulted in a uni-dimensional time-series for each video per human observer, identifying for each second of the video which emotion was recognized.
-Then, time-series corresponding to the same video were aggregated to calculate the proportion of human observers $x_{video_{i}.label_{j}.t_{k}}$ for each second of the video per emotional label shown in Equation \@ref(eq:one).
+This annotation procedure resulted in a uni-dimensional time-series for each video per human perceiver, identifying for each second of the video which emotion was recognized.
+Then, time-series corresponding to the same video were aggregated to calculate the proportion of human perceivers $x_{video_{i}.label_{j}.t_{k}}$ for each second of the video per emotional label shown in Equation \@ref(eq:one).
 
@@
 The result of Affdex is a mutivariate timeseries output which provides for each label a probability from 0 to 100 (rescaled from 0 to 1 for the analysis) of expressing the corresponding label.
-For this study only the six basic emotion labels are analysed to match with self-reports and human observers.
+For this study only the six basic emotion labels are analysed to match with self-reports and human perceivers.
 
@@
 
-Since encoders’ self-reports, human annotations and the automatic recognition include data on “non-basic” emotion labels and features, the analysis is performed using only the six basic emotion labels in order to compare them.
+Since expressers’ self-reports, human annotations and the automatic recognition include data on “non-basic” emotion labels and features, the analysis is performed using only the six basic emotion labels in order to compare them.
 The maximum rating for self-reports, human annotations and automatic recognition is used to label the video.
@@
 
-The confusion matrices obtained by crossing encoders’ self-reports, human annotations and the automatic recognition are analysed by calculating two-tailed t-tests of Pearson's product-moment correlation between dichotomous variables, accuracy which corresponds to the overall agreement rate averaged over cross-validation iterations and Cohen’s (unweighted) Kappa.
+The confusion matrices obtained by crossing expressers’ self-reports, human annotations and the automatic recognition are analysed by calculating two-tailed t-tests of Pearson's product-moment correlation between dichotomous variables, accuracy which corresponds to the overall agreement rate averaged over cross-validation iterations and Cohen’s (unweighted) Kappa.
 
@@
 
-## Human Observers’ Accuracy
+## Human Perceivers’ Accuracy
 
@@
 
-The overall correlation of recognition and non-recognition between self-reported emotions and human observers recognition is significant but low (`r corr_sr_hr$full_result`).
-In order to identify differences according to the emotional labels, encoders’ subjective feelings are compared with human observers’ recognition in a confusion matrix (Figure \@ref(fig:confusionMatrix-sr-hr)).
+The overall correlation of recognition and non-recognition between self-reported emotions and human perceivers recognition is significant but low (`r corr_sr_hr$full_result`).
+In order to identify differences according to the emotional labels, expressers’ subjective feelings are compared with human perceivers’ recognition in a confusion matrix (Figure \@ref(fig:confusionMatrix-sr-hr)).
 
@@
   scale_x_discrete(name = "Self-Reports") +
-  scale_y_discrete(name = "Human Observers") +
+  scale_y_discrete(name = "Human Perceivers") +
   theme_minimal() +
@@
 
-Each emotion label used to describe encoders’ self-reported subjective feeling (*i.e.*, the label rated with the highest value) is compared with the emotion labels which were rated with the highest score by human observers.
-Results of the confusion matrix show a low agreement between the emotion felt by the encoder during the elicitation and the emotion recognized by the human observers (Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95% CI [`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *disgust* (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_sr"]))` of the videos self-reported).
+Each emotion label used to describe expressers’ self-reported subjective feeling (*i.e.*, the label rated with the highest value) is compared with the emotion labels which were rated with the highest score by human perceivers.
+Results of the confusion matrix show a low agreement between the emotion felt by the expresser during the elicitation and the emotion recognized by the human perceivers (Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`, 95% CI [`r round(confusionMatrix_sr_hr$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_hr$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_hr$overall["Kappa"],2)`) except for *disgust* (`r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "disgust" & confusionMatrix_sr_hr_freq$sr_emotion == "disgust", "prop_sr"]))` of the videos self-reported).
 These results are far from those classically obtained in the literature for emotional facial expression recognition [@krumhuber2020human, report an average accuracy of 65%].
@@
 
-Interestingly human observers seem to recognize surprise expressed in videos where anger, fear happiness and sadness was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "anger", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "fear", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "happiness", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "sadness", "prop_sr"]))` of the videos self-reported), and in a lower instance *happiness* was recognized in videos where *fear* and *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "fear", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
+Interestingly human perceivers seem to recognize surprise expressed in videos where anger, fear happiness and sadness was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "anger", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "fear", "prop_sr"]))`, `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "happiness", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "surprise" & confusionMatrix_sr_hr_freq$sr_emotion == "sadness", "prop_sr"]))` of the videos self-reported), and in a lower instance *happiness* was recognized in videos where *fear* and *surprise* was the highest self-reported emotion (respectively `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "fear", "prop_sr"]))` and `r scales::percent(as.numeric(confusionMatrix_sr_hr_freq[confusionMatrix_sr_hr_freq$hr_emotion == "happiness" & confusionMatrix_sr_hr_freq$sr_emotion == "surprise", "prop_sr"]))` of the videos self-reported).
 
@@
 Similarly to the previous analysis, the overall correlation of recognition and non-recognition revealed a significant but very low coherence between self-reported emotions and automatic classifier’s recognition (`r corr_sr_ar$full_result`).
-A confusion matrix was used to compare encoders’ subjective feeling with the emotion label recognized by the automatic classifier (Figure \@ref(fig:confusionMatrix-sr-ar)).
+A confusion matrix was used to compare expressers’ subjective feeling with the emotion label recognized by the automatic classifier (Figure \@ref(fig:confusionMatrix-sr-ar)).
 
@@
 
-Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are somewhat similar to the ones with human observers (Table \@ref(tab:confusionTable-sr-ar)).
+Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are somewhat similar to the ones with human perceivers (Table \@ref(tab:confusionTable-sr-ar)).
 Overall, a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`, 95% CI [`r round(confusionMatrix_sr_ar$overall["AccuracyLower"],2)`,`r round(confusionMatrix_sr_ar$overall["AccuracyUpper"],2)`]; Kappa = `r round(confusionMatrix_sr_ar$overall["Kappa"],2)`) except for *happiness* (`r scales::percent(as.numeric(confusionMatrix_sr_ar_freq[confusionMatrix_sr_ar_freq$ar_emotion == "happiness" & confusionMatrix_sr_ar_freq$sr_emotion == "happiness", "prop_sr"]))` of the video self-reported) is evident.
@@
 
-As previously mentioned, human observers appear to be more accurate than the automatic classifier to recognize an individual's subjective feeling (human observers Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`; automatic classifier Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`).
-However, both make similar mistakes as the two-tailed t-tests of Pearson's product-moment correlation between human observers and automatic classifier recognition is significant (`r corr_hr_ar$full_result`).
+As previously mentioned, human perceivers appear to be more accurate than the automatic classifier to recognize an individual's subjective feeling (human perceivers Accuracy = `r round(confusionMatrix_sr_hr$overall["Accuracy"],2)`; automatic classifier Accuracy = `r round(confusionMatrix_sr_ar$overall["Accuracy"],2)`).
+However, both make similar mistakes as the two-tailed t-tests of Pearson's product-moment correlation between human perceivers and automatic classifier recognition is significant (`r corr_hr_ar$full_result`).
 
-A third confusion matrix is used to compare similarities (diagonal) and differences between human observers and automatic classifier in classifying the six emotion labels (Figure \@ref(fig:confusionMatrix-hr-ar)).
+A third confusion matrix is used to compare similarities (diagonal) and differences between human perceivers and automatic classifier in classifying the six emotion labels (Figure \@ref(fig:confusionMatrix-hr-ar)).
 
@@
   ) +
-  scale_x_discrete(name = "Human Observers") +
+  scale_x_discrete(name = "Human Perceivers") +
   scale_y_discrete(name = "Automatic Classifier") +
@@
 
-The overall agreement between human observers and the automatic classifier is in fact very low (Kappa = `r round(confusionMatrix_hr_ar$overall["Kappa"],2)`).
+The overall agreement between human perceivers and the automatic classifier is in fact very low (Kappa = `r round(confusionMatrix_hr_ar$overall["Kappa"],2)`).
 Except for *happiness* and *disgust* (respectively `r scales::percent(as.numeric(confusionMatrix_hr_ar_freq[confusionMatrix_hr_ar_freq$ar_emotion == "happiness" & confusionMatrix_hr_ar_freq$hr_emotion == "happiness", "prop_hr"]))` and `r scales::percent(as.numeric(confusionMatrix_hr_ar_freq[confusionMatrix_hr_ar_freq$ar_emotion == "disgust" & confusionMatrix_hr_ar_freq$hr_emotion == "disgust", "prop_hr"]))` of common labelling), there is no clear common pattern.
-Moreover, the automatic classifier has a tendency to label as *disgust* videos labeled as *sadness* by human observers, and as *happiness* videos labeled as *fear* by human observers.
+Moreover, the automatic classifier has a tendency to label as *disgust* videos labeled as *sadness* by human perceivers, and as *happiness* videos labeled as *fear* by human perceivers.
 
@@
 The present research aims to provide some empirical data to answer some of the questions posed.
-The faces encoders spontaneously displayed when confronted with an emotional eliciting task were submitted both to human and to automatic recognition.
-The criterion for recognition accuracy was the subjective feeling self-reported by the encoder once the elicitation task was carried out.
+The faces expressers spontaneously displayed when confronted with an emotional eliciting task were submitted both to human and to automatic recognition.
+The criterion for recognition accuracy was the subjective feeling self-reported by the expresser once the elicitation task was carried out.
 Results first reveal a low consistency between emotion felt and facial expression displayed.
 They show that facial expressions of emotion are often not displayed when the Basic Emotion View would predict them to be expressed.
-Secondly, results show low accuracy rates for both humans and the automatic classifier in identifying the inner emotional states of these encoders based on their facial expressions.
-Thirdly, human observers prove to be better at recognizing the emotion facially expressed than the automatic recognition tool is.
+Secondly, results show low accuracy rates for both humans and the automatic classifier in identifying the inner emotional states of these expressers based on their facial expressions.
+Thirdly, human perceivers prove to be better at recognizing the emotion facially expressed than the automatic recognition tool is.
 Such results support the hypothesis advanced by some authors of low emotion–expression coherence [@kappas2003facial].
@@
 Instead of viewing emotions as natural kinds [@barrett2006emotions], the constructivist approach supposes that emotions are social constructions and that facial behaviors intrinsically situated.
-The emotions that are recognized by the observer are constructed in her/his mind.
+The emotions that are recognized by the perceiver are constructed in her/his mind.
 Therefore, facial movements do not express specific emotions because they do not carry intrinsic emotional signification.
-It is the observer that infers the emotional meaning of the facial expression, this interpretation depending on availability of different information such as context or linguistic categories.
-As a consequence, one can predict from the first line of thinking that individuals’ emotional subjective feeling should be correlated to the recognition of facial expressions from both human observers and automatic classifiers whereas if emotions are social constructs, as stated by the second line of thinking, human observers should be better at perceiving emotions expressed on the face than automatic classifiers.
+It is the perceiver that infers the emotional meaning of the facial expression, this interpretation depending on availability of different information such as context or linguistic categories.
+As a consequence, one can predict from the first line of thinking that individuals’ emotional subjective feeling should be correlated to the recognition of facial expressions from both human perceivers and automatic classifiers whereas if emotions are social constructs, as stated by the second line of thinking, human perceivers should be better at perceiving emotions expressed on the face than automatic classifiers.
 
@@
 The correlations between the self-measured emotions and the observed facial behaviors are low and the latter are weakly recognized.
-Present results plead instead in favor of the constructivist stance.
-They show that human observers are more accurate than the automatic recognition tool to identify an individual’s subjective feeling on the basis of their face.
-Moreover, mistakes made by human observers look less arbitrary than the ones made by the classifier.
-For instance, even if a mix-up between disgust and anger is sometimes reported in recognition studies, odd confusions such as the present ones produced by the classifier have never been noted for human observers.
-The latter obviously make sense of the facial behavior they are witnessing.
-However, the human decoders in the present study were presented with faces without any contextual information which could have helped them to shape more precisely their interpretation.
-@hassin2013inherently reviewed evidence to demonstrate that faces are inherently ambiguous and that observers rely on situational cues when they process facial displays [see also @aviezer2008angry; @barrett2011context].
+Results show that human perceivers are more accurate than the automatic recognition tool to identify an individual’s subjective feeling on the basis of their face.
+Moreover, mistakes made by human perceivers look less arbitrary than the ones made by the classifier.
+For instance, even if a mix-up between disgust and anger is sometimes reported in recognition studies, odd confusions such as the present ones produced by the classifier have never been noted for human perceivers.
+The latter obviously make sense of the facial behavior they are witnessing, as claimed by the constructivist stance.
+That said, these results do not unequivocally argue in favor of the latter.
+Human perceivers in the present study were far from being perfectly accurate.
+Of course, one could argue that they were presented with faces without any contextual information which could have helped them to shape more precisely their interpretation.
+@hassin2013inherently reviewed evidence to demonstrate that faces are inherently ambiguous and that perceivers rely on situational cues when they process facial displays [see also @aviezer2008angry; @barrett2011context].
 It happens that contextual information even shifts emotion perception.
-In the present case, the facial stimuli displayed to the decoders were as equivocal as real-life facial expressions [@aviezer2017inherently].
-Yet, decoders had no cue at their disposal.
-So, without any possibility of integrating faces and context, their decoding accuracy is reduced [*e.g.*, @wagner1986communication].
-However, their superiority to the classifier tool is probably owing to their capacity to rely on their previous personal experience to invent a context in which a face could display such an expression.
+In the present case, the facial stimuli displayed to the decoders were as equivocal as real-life facial expressions [@aviezer2017inherently], yet perceivers having no cue at their disposal.
+So, without any possibility of integrating faces and context, their decoding accuracy is necessarily reduced [*e.g.*, @wagner1986communication].
+Their superiority to the classifier tool is probably owing to their capacity to rely on their previous personal experience to invent a context in which a face could display such an expression.
 Once a credible context is retrieved, they can affix an emotional label to the facial behavior.
@@
 The latter were triggered and measured with conceptually identical elicitation and assessment procedures [@reisenzein2013coherence].
-The use of self-reports to evaluate encoders’ subjective feelings is another limitation that can also be put forward because of the numerous cognitive biases they entail.
+The use of self-reports to evaluate expressers’ subjective feelings is another limitation that can also be put forward because of the numerous cognitive biases they entail.
 Well known problems with the reliability of self-reports are, among others, the reconstructive nature of memory, the influence of attentional biases on reports, demand characteristics, distorting effects of implicit causal theories and personal motives, as stressed by @nielsen2007conceptual.
@@
 Whereas this paradigm is longer and more complicated, it can lead to more robust results in reducing the forced-choice bias [@russell1993forced].
-However this procedure can also reduce the human observers’ accuracy.
+However this procedure can also reduce the human perceivers’ accuracy.
 In this regard, the results of the human observation could have been more ambiguous because it is not the natural way that people are inferring meaning from facial expressions.
 An alternative explanation relies in reducing the recognition bias involved in the classic recognition paradigm.
-Classic forced-choice paradigms obtain artificially high results, thus by using a more evolved approach, observers’ accuracy may have been lowered.
+Classic forced-choice paradigms obtain artificially high results, thus by using a more evolved approach, perceivers’ accuracy may have been lowered.
 Another flaw is the lack of comparison with various facial expression recognition methods.
@@
 This temporal information is indisputably a key feature of facial activity.
-It is not only observer-based judgements of facial displays which must be compared to automated facial analysis, but also different kinds of human recognition measurements should be undertaken.
+It is not only perceiver-based judgements of facial displays which must be compared to automated facial analysis, but also different kinds of human recognition measurements should be undertaken.
 The challenge researchers are especially confronted with is to find ways to appropriately collect data regarding the perception of spontaneous and dynamic facial behavior.
@@
 It has the advantage of controlling different parameters of the emotions investigated, such as intensity, quality, and temporal (onset, duration) features.
-Moreover, as encoders are alone, facing an emotional trigger, it also controls for the social context, removing its possible influence on their facial behaviors.
-However, it is known that encoders’ imagination can influence their expressiveness, for instance when they believe that their friend is doing the same (*vs.* a different) emotional task in another room [@jakobs1999social].
-Trying to exclude social influence by leaving encoders alone may be illusionary.
+Moreover, as expressers are alone, facing an emotional trigger, it also controls for the social context, removing its possible influence on their facial behaviors.
+However, it is known that expressers’ imagination can influence their expressiveness, for instance when they believe that their friend is doing the same (*vs.* a different) emotional task in another room [@jakobs1999social].
+Trying to exclude social influence by leaving expressers alone may be illusionary.
 From an ecological perspective, it is even a mistake to exonerate behavioral observations from social contexts.
@@
 All raise serious objections to the supposed close relation between emotion and face.
-They bring up several questions regarding the role the context plays in the emission and interpretation of the so-called facial expression of emotion.
-The finalization of operational and effective “reading emotional faces” devices rests on the answers, if any, to the questions raised.
+They bring up several questions, in particular regarding the perceiver’s role in the interpretation of the expresser’s emotional face.
+Is EFE recognition a perceiver-independent or perceiver-dependent process?
+Although the present results tend to favor the second option, the finalization of operational and effective “reading emotional faces” devices rests on the answer, if any, to the question raised.
+In any case, as stressed by @krumhuber2020human, machines’ accurate emotion classification is predicted by the prototypicality of a facial display.
+More prototypical is the facial behavior, more accurate is the classification.
+However, ordinary emotional facial behaviors are seldom prototypical.
 As a result, despite being one of the most investigated questions in affective science, the growing interest of industries and governments in tracking individual’s psychological states is supported by controversial assumptions.
-Considering the above, the present results provide additional evidence that an individual’s subjective feelings cannot be inferred from facial expressions, and they invalidate the hypothesis of hardwired emotions unambiguously displayed on the face.
+
+Considering the above, the present results add to the body of literature evidence that an individual’s subjective feelings are hardly inferred from spontaneous facial expressions in the absence of contextual cues, and they invalidate the hypothesis of hardwired emotions unambiguously displayed on the face.
+This study is the first one, to our knowledge, to look simultaneously at facial expressions spontaneously displayed in reaction to emotional triggers and their accurate decoding by human perceivers as compared to automatic emotion recognition tools.
+Results provide no support for the one-to-one mapping of subjective feelings and facial behaviors.
 Even if emotions were hardwired, in everyday life one does not observe prototypical facial expressions because of their rarity, and therefore research should be focused on analyzing non-prototypical facial expressions.
+Indeed, this study provide additional corroboration that facial displays of emotion are not exhibited in circumstances in which the Basic Emotion View would predict them to be expressed [@duran2017coherence].
 Advancements in identifying “non-basic” emotion labels as well as non-prototypical facial expression have indeed occured with automatic facial expression recognition tools [@mcduff2016discovering].
-However, these present results suggest that automatic facial expression recognition tools should merely evaluate facial morphology features such as action units [already evaluated in OpenFace, @baltruvsaitis2016openface; Affectiva’s Affdex, @mcduff2016affdex; Vicar Vision’s FaceReader, @den2005facereader, to name a few] rather than inferring supposedly emotional or affective states.
-Trying to interpret facial displays as a means of determining underlying emotional state, in all likelihood, remains in vain.
+However, these present results contribute to a critical understanding of the issues at stake.
+They suggest that automatic facial expression recognition tools should merely evaluate facial morphology features such as action units [already evaluated in OpenFace, @baltruvsaitis2016openface; Affectiva’s Affdex, @mcduff2016affdex; Vicar Vision’s FaceReader, @den2005facereader, to name a few] rather than inferring supposedly emotional or affective states.
+Trying to interpret facial displays as a means of determining underlying emotional state (especially without any information about the context, nor person specifications of the expresser, *etc.*), in all likelihood, remains dubious methods.
 

### 2020-11-04 manuscript/manuscript_apa.Rmd
acknowledgment

@@
 
-wordcount         : "7943"
-bibliography      : ["bib/manuscript_ref.bib"]
+wordcount         : "8558"
+bibliography      : ["bib/manuscript_ref.bib", "bib/r-references.bib"]
 csl               : ["csl/apa7.csl"]
@@
 
-output: papaja::apa6_word
+#output           : papaja::apa6_word
 
-# output            :
-#   bookdown::pdf_book:
-#     toc: false
-#     number_sections: false
-#     base_format: papaja::apa6_pdf
+output            : 
+  bookdown::pdf_book:
+    toc: false
+    number_sections: false
+    base_format: papaja::apa6_pdf
 
@@
 
+# Acknowledgement
+
+The authors would like to thank the developers of the following R packages used to process, analyze, display and report data: `r papaja::cite_r(file = "bib/r-references.bib")`.
+
+# Data availability
+
+The R code and relevant data for statistical computing are available on the "The emotion–facial expression link (data and code)" repository from the Center for Open Science [@Dupré_Tcherkassof_2020].
+
 \newpage

### 2020-11-04 manuscript/manuscript_apa.tex
acknowledgment

@@
   english,
-  doc]{apa7}
+  man]{apa7}
 \usepackage{lmodern}
@@
   pdflang={en-EN},
-  pdfkeywords={emotion, facial expression, self-report, human observer, automatic recognition.},
+  pdfkeywords={emotion, facial expression, self-report, human perceiver, automatic recognition.},
   hidelinks,
@@
 \shorttitle{The emotion facial expression link}
-\keywords{emotion, facial expression, self-report, human observer, automatic recognition.\newline\indent Word count: 7943}
+\keywords{emotion, facial expression, self-report, human perceiver, automatic recognition.\newline\indent Word count: 8558}
+\DeclareDelayedFloatFlavor{ThreePartTable}{table}
+\DeclareDelayedFloatFlavor{lltable}{table}
+\DeclareDelayedFloatFlavor*{longtable}{table}
+\makeatletter
+\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
+\makeatother
 \usepackage{csquotes}
@@
 
+\authornote{
+
+We thank anonymous reviewers for their insightful comments on a previous version of the manuscript.
+
+The authors made the following contributions.
+Anna Tcherkassof: Conceptualization, Project administration, Validation, Writing - Original Draft Preparation, Writing - Review \& Editing; Damien Dupré: Investigation, Data curation, Formal Analysis, Software, Visualization, Writing - Original Draft Preparation, Writing - Review \& Editing.
+
+Correspondence concerning this article should be addressed to Anna Tcherkassof, Bâtiment Michel Dubois, Université Grenoble Alpes, 1251 Avenue Centrale, 38400 Saint-Martin-d'Hères.
+E-mail: \href{mailto:anna.tcherkassof@univ-grenoble-alpes.fr}{\nolinkurl{anna.tcherkassof@univ-grenoble-alpes.fr}}
+
+}
+
 \affiliation{\vspace{0.5cm}\textsuperscript{1} University Grenoble Alpes, Grenoble, France\\\textsuperscript{2} Dublin City University, Dublin, Ireland}
@@
 On one hand, the ``Basic Emotion View'' conceives emotions as genetically hardwired and therefore being genuinely displayed through facial expressions.
-On the other hand, the constructivist approach conceives emotions as socially constructed; the emotional meaning of a facial expression being inferred by the observer.
-In order to evaluate the coherence between the subjective feeling of emotions and their recognition based on facial expression, 232 videos of encoders recruited to carry out an emotion elicitation task were annotated by 1383 human observers as well as by an automatic facial expression classifier.
-Results show low accuracy of human observers and of the automatic classifier to infer the subjective feeling from the facial expressions displayed by encoders.
-They also show a weak consistency between self-reported emotional states and facial emotional displays.
+Consequently, emotion recognition is perceiver-independent.
+On the other hand, the constructivist approach conceives emotions as socially constructed, the emotional meaning of a facial expression being inferred by the perceiver.
+Hence, emotion recognition is perceiver-dependent.
+In order 1) to evaluate the coherence between the subjective feeling of emotions and their spontaneous facial displays, and 2) to compare the recognition of such displays by human perceivers and by an automatic facial expression classifier, 232 videos of expressers recruited to carry out an emotion elicitation task were annotated by 1383 human perceivers as well as by Affdex, an automatic classifier.
+Results show a weak consistency between self-reported emotional states by expressers and their facial emotional displays.
+They also show low accuracy both of human perceivers and of the automatic classifier to infer the subjective feeling from the spontaneous facial expressions displayed by expressers.
+However, the results are more in favor of a perceiver-dependent view.
 Based on these results, the hypothesis of genetically hardwired emotion genuinely displayed is difficult to support, whereas the idea of emotion and facial expression as being socially constructed appears to be more likely.
@@
 With the development of commercial automatic facial expression recognition tools (Dupré et al., 2020, 2018), industries and governments are gradually implementing this technology in order to monitor humans' emotions in various scenarios (\emph{e.g.}, in marketing, healthcare, and the automotive industry to name a few).
-This technology rests on the premise that facial expressions provide direct access to individuals' subjective feelings, and that one can read the emotions displayed on a person's face as one would an open book.
-This premise is derived from the ``Basic Emotion View'', a theory which suggests a one-to-one mapping between subjective feeling and facial expression (Ekman et al., 1987; Ekman, 1992, 2007; Ekman \& Heider, 1988).
-Even if this premise is central to the modern mainstream approach of human emotion, recent research in affective science is challenging it.
-Once the Basic Emotion View is briefly described, its foremost criticisms will be exposed.
-Inconsistencies and unpredicted findings come both from field observations and laboratory experiments.
-They relate to two different issues.
-The first issue is the sender's production of spontaneous facial displays.
-The second issue relates to the receiver's interpretation of these facial displays.
-Building on this evidence, and as an attempt to shed further light on the emotion--face link, the present study's main goal is to tackle these two issues at the same time.
-It aims at identifying the extent to which ordinary people's experienced emotions are displayed through identifiable Emotional Facial Expressions (EFE) recognized as such both by human individuals, and by automatic facial expression recognition tools.
+This technology rests on the premise that emotional state is revealed by the measurement of facial features thanks to a one-to-one mapping between subjective feeling and facial expression.
+This premise comes from the Basic Emotion View (Ekman et al., 1987; Ekman, 1992, 2007; Ekman \& Heider, 1988).
+By contrast, for the constructivist approach (barrett2017emotions; Gendron \& Barrett, 2017), emotions being socially constructed, a making-meaning process takes part in the link between emotion and facial expression.
+Emotion recognition is considered as resulting from the perceiver making inferences about the expresser's underlying emotions.
+The constructivist approach is therefore a perceiver-\emph{dependent} view whereas the Basic Emotion View is a perceiver-\emph{independent} view (Gendron \& Barrett, 2017).
+In the emotion--facial expression link contention, the issue of facial expressions' naturalness must also be considered.
+Since automatic recognition tools are meant to be used in real-life conditions, spontaneous displays of emotional feelings are the heart of the matter as well as their accurate recognition.
+The potential significance of the present study is that it aims at providing data, not only on the emotion--facial expression link, but also on the perceiver-independent \emph{vs.} perceiver-dependent question of emotion recognition as regard to spontaneous facial expressions of self-reported emotions.
+It looks at the triangulation emotional feeling---spontaneous facial display---emotion recognition within a single paradigm comparing the recognition accuracy between human and machines.
 
@@
 
-The belief that facial expression is linked to emotional states can be traced back to Darwin in \emph{The Expression of the Emotions in Man and Animals} (Darwin, 1872).
-Countless studies have fortified the idea that emotional states are inherently coupled to a set of prototypic facial expressions (\emph{e.g.}, Ekman \& Cordaro, 2011).
+Since Darwin's book \emph{The Expression of the Emotions in Man and Animals} (Darwin, 1872), countless studies have fortified the idea that emotional states are inherently coupled to a set of prototypic facial expressions (\emph{e.g.}, Ekman \& Cordaro, 2011).
 The Basic Emotion View holds that facial expressions are genuine displays of an individual's inner emotional state.
-More specifically, a set of six emotions (\emph{fear}, \emph{anger}, \emph{surprise}, \emph{disgust}, \emph{sadness} and \emph{joy}) are universally displayed and are genetically hardwired not only in humans (Ekman, 1992), but also in different animal species (Waal, 2019).
+More specifically, a set of six so-called basic emotions (\emph{fear}, \emph{anger}, \emph{surprise}, \emph{disgust}, \emph{sadness} and \emph{happiness}) are universally displayed and are genetically hardwired not only in humans (Ekman, 1992), but also in different animal species (Waal, 2019).
 According to the Basic Emotion View, \emph{``when emotions are aroused by perception of a social event, a set of central commands produce patterned emotion-specific changes in multiple systems, including {[}\ldots{]} facial expressions.''} (Ekman, 2007, p.
@@
 To respond to criticisms, several amendments have been made to the Basic Emotion View, increasing the number of basic emotions from six to seven (Ekman \& Heider, 1988) as well as adding the concept of ``display rules'' to explain cultural differences in the management of facial expressions (Ekman et al., 1987).
-
-Relatively few studies have actually tested Basic Emotion View's fundamental claim regarding the facial production (so-called ``expression'') of emotions.
+The most recent versions of the Basic Emotion Theory even distinguish some twenty expressions of emotions, including love, pride, embarrassment, shame, among others (Keltner et al., 2019).
+
+Relatively few studies have actually tested Basic Emotion View's fundamental claim regarding the facial production (or ``expression'') of emotions.
 Rosenberg \& Ekman (1994) reported the first evidence of coherence between self-report of emotion and displayed facial expressions.
@@
 Notwithstanding this study, few have used purely descriptive methods such as electromyography or objective face coding systems to identify and to measure the actual changes in the face when a given emotion is felt (see Wagner, 1997, for methodological points).
-Much of the research has focused instead on the recognition of EFE, that is on the issue of the viewer's interpretation of facial displays.
-As people have been thought to display their internal state, EFE supposedly serve as a window into the emotions of others.
-Viewers ``recognize'' the facial displays and infer the corresponding emotional state.
-The Readout Hypothesis (Buck, 1985) formalizes this folk theory.
-According to it, facial expressions have evolved to provide \emph{``an external readout of those motivational-emotional processes that have had social implications during the course of evolution''} (Buck, 1985, pp.
+
+Much of the research has focused instead on the recognition of emotional facial expressions (EFE), that is on the issue of the perceiver's interpretation of facial displays.
+According to the Readout Hypothesis (Buck, 1985) which formalizes the basic Emotion View, facial expressions have evolved to provide \emph{``an external readout of those motivational-emotional processes that have had social implications during the course of evolution''} (Buck, 1985, pp.
 396--397).
 Thus, as readouts of emotional states, spontaneous expressive displays are directly accessible to other organisms.
-Research on the inference of emotions from facial expressions has established that viewers show considerable agreement that the so-called basic emotions---happiness, anger, fear, disgust, sadness, and surprise---are associated with specific facial displays (\emph{e.g.}, reviews from Elfenbein \& Ambady, 2002; Russell, 1994).
-
-Yet, in spite of the popular support it has received, the empirical data called upon by the Basic Emotion View remains unpersuasive.
-Notably, it fails to explain, in instances in which display rules cannot be called upon, how individuals can feel emotions without expressing them or how individuals can express emotions without feeling them, and why observers can make mistakes in recognizing facial expressions of basic emotions, among others.
-The following is a summary of key findings and conclusions both from field observations and laboratory experiments on spontaneous expression of emotions, all of them strongly subverting the Basic Emotion viewpoint (for reviews, Fernández-Dols \& Crivelli, 2013; Reisenzein et al., 2013; Russell et al., 2003).
-Mains issues regarding recognition studies are also briefly discussed afterwards.
+Research on the inference of emotions from facial expressions has established that perceivers show considerable agreement that the so-called basic emotions are associated with specific facial displays (\emph{e.g.}, reviews from Elfenbein \& Ambady, 2002; Russell, 1994).
+
+Yet, the empirical data called upon by the Basic Emotion View remains unpersuasive.
+Notably, it fails to explain, in instances in which display rules cannot be called upon, how individuals can feel emotions without expressing them or how individuals can express emotions without feeling them, and why perceivers can make mistakes in recognizing facial expressions of basic emotions, among others.
+All these findings subverting the Basic Emotion viewpoint come from field observations and laboratory experiments on spontaneous expression of emotions (for reviews, Fernández-Dols \& Crivelli, 2013; Reisenzein et al., 2013; Russell et al., 2003).
+The key findings are summarized below as well as the main issues regarding recognition studies.
 
@@
 Though they cannot be considered as a straightforward test of the triggering role of emotion on facial behavior, they have the advantage of considering situations that cannot be created in a laboratory.
-For instance, the ethological study of Kraut \& Johnston (1979) contradicts the Basic Emotion View premise that a smile is only the major component of a facial display associated with and caused by feelings of joy or happiness.
+For instance, the ethological study of Kraut and Kraut \& Johnston (1979) contradicts the Basic Emotion View premise that a smile is only the major component of a facial display associated with and caused by feelings of happiness.
 Naturalistic observation at a bowling alley showed that bowlers do not necessarily smile after scoring a spare or a strike (a situation likely to elicit a positive emotion).
@@
 More specifically, winners showed Duchenne smiles and other types of smiles when they were interacting with other people during the awards ceremony, but smiles were scarce or nonexistent when waiting behind the podium and/or when turning toward the flagpoles and focusing their attention on the flags and the national anthem.
-Crivelli et al.
-(2015) also found that the strongest predictor for the occurrence of Duchenne smiles in judo winners is when they are engaged in social interaction and not when they just won their match.
-
-Happiness/joy is not the only emotion weakly associated with the predicted facial expression.
+Other naturalistic studies have also found that happiness cannot be considered the strongest predictor for the occurrence of Duchenne smiles (Crivelli et al., 2015).
+
+Happiness is not the only emotion weakly associated with the predicted facial expression.
 A naturalistic study conducted by Scherer \& Ceschi (1997) in an airport's baggage handling office showed that passengers claiming for their lost luggage displayed very few facial expressions of negative emotions while self-reporting subjective feeling states of anger or sadness among others.
@@
 Many people strongly agree that so-called basic emotions are associated with specific facial configurations (Ekman, 2017; Elfenbein \& Ambady, 2002) and also argue that this is strong evidence for the Basic Emotion View.
-Moreover, it implies that EFE recognition of human observers should be as accurate as automatic classifiers'.
+Moreover, it implies that EFE recognition of human perceivers should be as accurate as automatic classifiers.
 
@@
 Facial stimuli used in experiments also constitute a methodological bias because they are unrepresentative of ordinary facial expressions.
-Basic Emotion View empirical evidence is based for the most part on methods using a static and unnatural material, namely, still photographs of posed facial expressions of emotion (\emph{e.g.}, intentionally encoded by the sender).
+Basic Emotion View empirical evidence is based for the most part on methods using a static and unnatural material, namely, still photographs of posed facial expressions of emotion (\emph{e.g.}, intentionally encoded by the expresser).
 This kind of methodology raises questions about its ecological validity and the generalizability of the results to real interpersonal emotional communication (\emph{e.g.}, Tcherkassof et al., 2007).
@@
 Hess \& Kleck (1990) have also pointed out the importance of the dynamics of facial movements, and particularly the irregularity, or phasic changes, of the expressions' unfolding.
-Thus, the motion of facial expression provides observers with other information than the one provided by static expressions.
+Thus, the motion of facial expression provides perceivers with other information than the one provided by static expressions.
 It may be that differences in the social information displayed by static and dynamic expressions leads to facial recognition differential effects.
@@
 
-Starting from the empirical evidence suggesting that spontaneous facial expressions in ordinary life are equivocal, Dols (2017) argues in favor of a pragmatic conception of natural facial displays.
+Starting from the empirical evidence suggesting that spontaneous facial expressions in ordinary life are equivocal, Fernández-Dols (2017) argues in favor of a pragmatic conception of natural facial displays.
 He makes a plea for the idea that natural facial displays, rather than ``saying''--because facial expressions do not have a specific meaning--``make'' things.
 Facial expressions are actions in a communicative interaction.
-They do not express emotions but they \emph{``prompt, on the receiver's side, important inferences about the context, the sender, and the course of the interaction between sender and receiver''} (Dols, 2017, p.
+They do not express emotions but they \emph{``prompt, on the receiver's side, important inferences about the context, the sender {[}expresser{]}, and the course of the interaction between sender {[}expresser{]} and receiver''} (Fernández-Dols, 2017, p.
 466).
 As such, the fact that facial displays are able to signal emotions is a byproduct of one of their main functions; implementing actions performing practical ends.
-Therefore, expressive displays hold a motor intention (Pacherie, 2003).
+Therefore, a certain number of expressive displays (\emph{e.g.}, surprise, disgust, fear, anger) hold a motor intention (Pacherie, 2003).
 They implement their aim which is their motor intention.
@@
 Facial displays are parts of pragmatic actions aiming at orienting the person's relation to their environment.
-For instance facial displays are maintaining, breaking or restoring the relationship between sender and perceiver (Frijda, 2012).
+For instance, facial displays are maintaining, breaking or restoring the relationship between expresser and perceiver (Frijda, 2012).
 Facial displays are not recognized in semantic terms but are perceived as intentional actions.
@@
 The constructivist approach claims that facial displays are behaviors whose meaning is inferred by perceivers.
-Findings support this observer dependence {[}Lindquist \& Gendron (2013); niedenthal2017embodied{]}.
+Findings support this perceiver dependence (Lindquist \& Gendron, 2013; Niedenthal et al., 2017).
 They show that to make meaning of another person's facial behavior, the perceiver relies in particular on her/his knowledge about emotion categories.
@@
 They did not infer inner states, rather they proceeded with action identification that pointed out the functions of behaviors (see also Crivelli et al., 2017 for similar observations among a small-scale society of Papua New Guinea).
-The constructivist approach considers that specific emotion categories, as conceptualized by Western cultures' knowledge, are cast on the perceived face to make meaning of the sender's facial displays.
+The constructivist approach considers that specific emotion categories, as conceptualized by Western cultures' knowledge, are cast on the perceived face to make meaning of the expresser's facial displays.
 Following this approach, faces convey a range of information essential for social communication.
@@
 More specifically, this study aims to examine the recognition of EFE produced by ordinary people during situations judged and/or self-reported to involve different emotions.
-It (a) examines consistency between ordinary people's self-reported emotional experience and observers' judgments of these ordinary people's EFE, and (b) examines consistency between ordinary people's self-reported emotional experience and an automatic classifier's analysis of these ordinary people's EFE.
-In other words, it is interested in how people actually move their faces to express self-reported emotions, in how human observers accurately infer the expresser's emotional state, and in how automatic recognition accurately codes the expresser's emotional state.
-Because of their superior ability to exactly recognize EFE, it is expected that human observers' will have a higher accuracy to identify senders' subjective feeling than automatic EFE recognition tools.
+It (a) examines consistency between ordinary people's self-reported emotional experience and perceivers' judgments of these ordinary people's EFE, and (b) examines consistency between ordinary people's self-reported emotional experience and an automatic classifier's analysis of these ordinary people's EFE.
+In other words, it is interested in how people actually move their faces to express self-reported emotions, in how human perceivers accurately infer the expresser's emotional state, and in how automatic recognition accurately codes the expresser's emotional state.
+Because of their superior ability to exactly recognize EFE, it is expected that human perceivers' will have a higher accuracy to identify expressers' subjective feeling than automatic EFE recognition tools.
 
@@
 
-To evaluate the consistency between subjective feeling of emotions and their recognition from facial expressions, encoders were first recruited to perform an emotion elicitation task while their facial expression was video recorded.
-In order to reduce the likelihood of facial control, the encoders were alone in the room and were filmed by hidden cameras, so they had no reason to comply with social display rules.
-Then, the videos of the encoders' faces were shown to human observers and were also analyzed by an automatic classifier in order to identify which emotion was displayed.
+To evaluate the consistency between subjective feeling of emotions and their recognition from facial expressions, expressers were first recruited to perform an emotion elicitation task while their facial expression was video recorded.
+In order to reduce the likelihood of facial control, the expressers were alone in the room and were filmed by hidden cameras, so they had no reason to comply with social display rules.
+Then, the videos of the expressers' faces were shown to human perceivers and were also analyzed by an automatic classifier in order to identify which emotion was displayed.
 
@@
 Other tasks were active; they required the participant to interact with a computer (e.g.~to answer questions or to test a flawed software).
-Encoders' faces were recorded using a hidden camera resulting 358 front facing 768x576 videos varying from 1s to 1479s.
+Expressers' faces were recorded using a hidden camera resulting 358 front facing 768x576 videos varying from 1s to 1479s.
 These recordings form the DynEmo database (Tcherkassof et al., 2013).
 
-After the emotion elicitation task, the encoders rated their subjective feeling on Likert scales from 0 (``not at all'') to 5 (``strongly'') related to six ``basic'' emotion labels (\emph{i.e.}, \emph{anger}, \emph{disgust}, \emph{fear}, \emph{happiness}, \emph{surprise} and \emph{sadness}) as well as six ``non-basic'' emotion labels (\emph{i.e.}, \emph{pride}, \emph{curiosity}, \emph{boredom}, \emph{shame}, \emph{humiliation}, and \emph{disappointment}).
-
-Finally, a debriefing session was performed to ensure that encoders were not durably affected by the emotion elicitation task.
-The debriefing was also used to check that encoders did not guess the real purpose of the experiment (\emph{e.g.}, being filmed while they were performing an emotional elicitation task) to guarantee facial expressions' genuineness.
-All encoders gave their agreement on their data and recordings being processed for research purposes only.
+After the emotion elicitation task, the expressers rated their subjective feeling on Likert scales from 0 (``not at all'') to 5 (``strongly'') related to six ``basic'' emotion labels (\emph{i.e.}, \emph{anger}, \emph{disgust}, \emph{fear}, \emph{happiness}, \emph{surprise} and \emph{sadness}) as well as six ``non-basic'' emotion labels (\emph{i.e.}, \emph{pride}, \emph{curiosity}, \emph{boredom}, \emph{shame}, \emph{humiliation}, and \emph{disappointment}).
+
+Finally, a debriefing session was performed to ensure that expressers were not durably affected by the emotion elicitation task.
+The debriefing was also used to check that expressers did not guess the real purpose of the experiment (\emph{e.g.}, being filmed while they were performing an emotional elicitation task) to guarantee facial expressions' genuineness.
+All expressers gave their agreement on their data and recordings being processed for research purposes only.
 
@@
 
-This annotation procedure resulted in a uni-dimensional time-series for each video per human observer, identifying for each second of the video which emotion was recognized.
-Then, time-series corresponding to the same video were aggregated to calculate the proportion of human observers \(x_{video_{i}.label_{j}.t_{k}}\) for each second of the video per emotional label shown in Equation \eqref{eq:one}.
+This annotation procedure resulted in a uni-dimensional time-series for each video per human perceiver, identifying for each second of the video which emotion was recognized.
+Then, time-series corresponding to the same video were aggregated to calculate the proportion of human perceivers \(x_{video_{i}.label_{j}.t_{k}}\) for each second of the video per emotional label shown in Equation \eqref{eq:one}.
 
@@
 
-Since encoders' self-reports, human annotations and the automatic recognition include data on ``non-basic'' emotion labels and features, the analysis is performed using only the six basic emotion labels in order to compare them.
+Since expressers' self-reports, human annotations and the automatic recognition include data on ``non-basic'' emotion labels and features, the analysis is performed using only the six basic emotion labels in order to compare them.
 The maximum rating for self-reports, human annotations and automatic recognition is used to label the video.
@@
 
-The confusion matrices obtained by crossing encoders' self-reports, human annotations and the automatic recognition are analysed by calculating two-tailed t-tests of Pearson's product-moment correlation between dichotomous variables, accuracy which corresponds to the overall agreement rate averaged over cross-validation iterations and Cohen's (unweighted) Kappa.
+The confusion matrices obtained by crossing expressers' self-reports, human annotations and the automatic recognition are analysed by calculating two-tailed t-tests of Pearson's product-moment correlation between dichotomous variables, accuracy which corresponds to the overall agreement rate averaged over cross-validation iterations and Cohen's (unweighted) Kappa.
 
@@
 
-\hypertarget{human-observers-accuracy}{%
-\subsection{Human Observers' Accuracy}\label{human-observers-accuracy}}
-
-The overall correlation of recognition and non-recognition between self-reported emotions and human observers recognition is significant but low (\(r = .24\), 95\% CI \([.19\), \(.29]\), \(t(1384) = 9.21\), \(p < .001\)).
-In order to identify differences according to the emotional labels, encoders' subjective feelings are compared with human observers' recognition in a confusion matrix (Figure \ref{fig:confusionMatrix-sr-hr}).
+\hypertarget{human-perceivers-accuracy}{%
+\subsection{Human Perceivers' Accuracy}\label{human-perceivers-accuracy}}
+
+The overall correlation of recognition and non-recognition between self-reported emotions and human perceivers recognition is significant but low (\(r = .24\), 95\% CI \([.19\), \(.29]\), \(t(1384) = 9.21\), \(p < .001\)).
+In order to identify differences according to the emotional labels, expressers' subjective feelings are compared with human perceivers' recognition in a confusion matrix (Figure \ref{fig:confusionMatrix-sr-hr}).
 
 \begin{figure}[!h]
-\includegraphics[width=\textwidth]{manuscript_apa_files/figure-latex/confusionMatrix-sr-hr-1} \caption{Confusion matrix between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the human observers.}\label{fig:confusionMatrix-sr-hr}
+\includegraphics[width=\textwidth]{manuscript_apa_files/figure-latex/confusionMatrix-sr-hr-1} \caption{Confusion matrix between the emotion self-reported as being characteristic of the elicitation with the emotion recognized by the human perceivers.}\label{fig:confusionMatrix-sr-hr}
 \end{figure}
 
-Each emotion label used to describe encoders' self-reported subjective feeling (\emph{i.e.}, the label rated with the highest value) is compared with the emotion labels which were rated with the highest score by human observers.
-Results of the confusion matrix show a low agreement between the emotion felt by the encoder during the elicitation and the emotion recognized by the human observers (Accuracy = 0.43, 95\% CI {[}0.35,0.52{]}; Kappa = 0.19) except for \emph{disgust} (100\% of the videos self-reported).
+Each emotion label used to describe expressers' self-reported subjective feeling (\emph{i.e.}, the label rated with the highest value) is compared with the emotion labels which were rated with the highest score by human perceivers.
+Results of the confusion matrix show a low agreement between the emotion felt by the expresser during the elicitation and the emotion recognized by the human perceivers (Accuracy = 0.43, 95\% CI {[}0.35,0.52{]}; Kappa = 0.19) except for \emph{disgust} (100\% of the videos self-reported).
 These results are far from those classically obtained in the literature for emotional facial expression recognition (Krumhuber et al., 2020, report an average accuracy of 65\%).
@@
 
-Interestingly human observers seem to recognize surprise expressed in videos where anger, fear happiness and sadness was the highest self-reported emotion (respectively 29\%, 33\%, 35\% and 33\% of the videos self-reported), and in a lower instance \emph{happiness} was recognized in videos where \emph{fear} and \emph{surprise} was the highest self-reported emotion (respectively 50\% and 26\% of the videos self-reported).
+Interestingly human perceivers seem to recognize surprise expressed in videos where anger, fear happiness and sadness was the highest self-reported emotion (respectively 29\%, 33\%, 35\% and 33\% of the videos self-reported), and in a lower instance \emph{happiness} was recognized in videos where \emph{fear} and \emph{surprise} was the highest self-reported emotion (respectively 50\% and 26\% of the videos self-reported).
 
@@
 Similarly to the previous analysis, the overall correlation of recognition and non-recognition revealed a significant but very low coherence between self-reported emotions and automatic classifier's recognition (\(r = .12\), 95\% CI \([.07\), \(.17]\), \(t(1384) = 4.50\), \(p < .001\)).
-A confusion matrix was used to compare encoders' subjective feeling with the emotion label recognized by the automatic classifier (Figure \ref{fig:confusionMatrix-sr-ar}).
+A confusion matrix was used to compare expressers' subjective feeling with the emotion label recognized by the automatic classifier (Figure \ref{fig:confusionMatrix-sr-ar}).
 
@@
 
-Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are somewhat similar to the ones with human observers (Table \ref{tab:confusionTable-sr-ar}).
+Results obtained for the comparison between emotions self-reported and recognized by the automatic classifier are somewhat similar to the ones with human perceivers (Table \ref{tab:confusionTable-sr-ar}).
 Overall, a low agreement between emotion self-reported and emotion recognized by the automatic classifier (Accuracy = 0.30, 95\% CI {[}0.22,0.38{]}; Kappa = 0.07) except for \emph{happiness} (44\% of the video self-reported) is evident.
@@
 
-As previously mentioned, human observers appear to be more accurate than the automatic classifier to recognize an individual's subjective feeling (human observers Accuracy = 0.43; automatic classifier Accuracy = 0.30).
-However, both make similar mistakes as the two-tailed t-tests of Pearson's product-moment correlation between human observers and automatic classifier recognition is significant (\(r = .22\), 95\% CI \([.17\), \(.27]\), \(t(1384) = 8.31\), \(p < .001\)).
-
-A third confusion matrix is used to compare similarities (diagonal) and differences between human observers and automatic classifier in classifying the six emotion labels (Figure \ref{fig:confusionMatrix-hr-ar}).
+As previously mentioned, human perceivers appear to be more accurate than the automatic classifier to recognize an individual's subjective feeling (human perceivers Accuracy = 0.43; automatic classifier Accuracy = 0.30).
+However, both make similar mistakes as the two-tailed t-tests of Pearson's product-moment correlation between human perceivers and automatic classifier recognition is significant (\(r = .22\), 95\% CI \([.17\), \(.27]\), \(t(1384) = 8.31\), \(p < .001\)).
+
+A third confusion matrix is used to compare similarities (diagonal) and differences between human perceivers and automatic classifier in classifying the six emotion labels (Figure \ref{fig:confusionMatrix-hr-ar}).
 
 \begin{figure}[!h]
-\includegraphics[width=\textwidth]{manuscript_apa_files/figure-latex/confusionMatrix-hr-ar-1} \caption{Proportion of emotion labels classified by human observers which are recognized by the automatic classifier.}\label{fig:confusionMatrix-hr-ar}
+\includegraphics[width=\textwidth]{manuscript_apa_files/figure-latex/confusionMatrix-hr-ar-1} \caption{Proportion of emotion labels classified by human perceivers which are recognized by the automatic classifier.}\label{fig:confusionMatrix-hr-ar}
 \end{figure}
 
-The overall agreement between human observers and the automatic classifier is in fact very low (Kappa = 0.18).
+The overall agreement between human perceivers and the automatic classifier is in fact very low (Kappa = 0.18).
 Except for \emph{happiness} and \emph{disgust} (respectively 65\% and 50\% of common labelling), there is no clear common pattern.
-Moreover, the automatic classifier has a tendency to label as \emph{disgust} videos labeled as \emph{sadness} by human observers, and as \emph{happiness} videos labeled as \emph{fear} by human observers.
+Moreover, the automatic classifier has a tendency to label as \emph{disgust} videos labeled as \emph{sadness} by human perceivers, and as \emph{happiness} videos labeled as \emph{fear} by human perceivers.
 
@@
 The present research aims to provide some empirical data to answer some of the questions posed.
-The faces encoders spontaneously displayed when confronted with an emotional eliciting task were submitted both to human and to automatic recognition.
-The criterion for recognition accuracy was the subjective feeling self-reported by the encoder once the elicitation task was carried out.
+The faces expressers spontaneously displayed when confronted with an emotional eliciting task were submitted both to human and to automatic recognition.
+The criterion for recognition accuracy was the subjective feeling self-reported by the expresser once the elicitation task was carried out.
 Results first reveal a low consistency between emotion felt and facial expression displayed.
 They show that facial expressions of emotion are often not displayed when the Basic Emotion View would predict them to be expressed.
-Secondly, results show low accuracy rates for both humans and the automatic classifier in identifying the inner emotional states of these encoders based on their facial expressions.
-Thirdly, human observers prove to be better at recognizing the emotion facially expressed than the automatic recognition tool is.
+Secondly, results show low accuracy rates for both humans and the automatic classifier in identifying the inner emotional states of these expressers based on their facial expressions.
+Thirdly, human perceivers prove to be better at recognizing the emotion facially expressed than the automatic recognition tool is.
 Such results support the hypothesis advanced by some authors of low emotion--expression coherence (Kappas, 2003).
@@
 Instead of viewing emotions as natural kinds (Barrett, 2006), the constructivist approach supposes that emotions are social constructions and that facial behaviors intrinsically situated.
-The emotions that are recognized by the observer are constructed in her/his mind.
+The emotions that are recognized by the perceiver are constructed in her/his mind.
 Therefore, facial movements do not express specific emotions because they do not carry intrinsic emotional signification.
-It is the observer that infers the emotional meaning of the facial expression, this interpretation depending on availability of different information such as context or linguistic categories.
-As a consequence, one can predict from the first line of thinking that individuals' emotional subjective feeling should be correlated to the recognition of facial expressions from both human observers and automatic classifiers whereas if emotions are social constructs, as stated by the second line of thinking, human observers should be better at perceiving emotions expressed on the face than automatic classifiers.
+It is the perceiver that infers the emotional meaning of the facial expression, this interpretation depending on availability of different information such as context or linguistic categories.
+As a consequence, one can predict from the first line of thinking that individuals' emotional subjective feeling should be correlated to the recognition of facial expressions from both human perceivers and automatic classifiers whereas if emotions are social constructs, as stated by the second line of thinking, human perceivers should be better at perceiving emotions expressed on the face than automatic classifiers.
 
@@
 The correlations between the self-measured emotions and the observed facial behaviors are low and the latter are weakly recognized.
-Present results plead instead in favor of the constructivist stance.
-They show that human observers are more accurate than the automatic recognition tool to identify an individual's subjective feeling on the basis of their face.
-Moreover, mistakes made by human observers look less arbitrary than the ones made by the classifier.
-For instance, even if a mix-up between disgust and anger is sometimes reported in recognition studies, odd confusions such as the present ones produced by the classifier have never been noted for human observers.
-The latter obviously make sense of the facial behavior they are witnessing.
-However, the human decoders in the present study were presented with faces without any contextual information which could have helped them to shape more precisely their interpretation.
+Results show that human perceivers are more accurate than the automatic recognition tool to identify an individual's subjective feeling on the basis of their face.
+Moreover, mistakes made by human perceivers look less arbitrary than the ones made by the classifier.
+For instance, even if a mix-up between disgust and anger is sometimes reported in recognition studies, odd confusions such as the present ones produced by the classifier have never been noted for human perceivers.
+The latter obviously make sense of the facial behavior they are witnessing, as claimed by the constructivist stance.
+That said, these results do not unequivocally argue in favor of the latter.
+Human perceivers in the present study were far from being perfectly accurate.
+Of course, one could argue that they were presented with faces without any contextual information which could have helped them to shape more precisely their interpretation.
 Hassin et al.
-(2013) reviewed evidence to demonstrate that faces are inherently ambiguous and that observers rely on situational cues when they process facial displays (see also Aviezer et al., 2008; Barrett et al., 2011).
+(2013) reviewed evidence to demonstrate that faces are inherently ambiguous and that perceivers rely on situational cues when they process facial displays (see also Aviezer et al., 2008; Barrett et al., 2011).
 It happens that contextual information even shifts emotion perception.
-In the present case, the facial stimuli displayed to the decoders were as equivocal as real-life facial expressions (Aviezer et al., 2017).
-Yet, decoders had no cue at their disposal.
-So, without any possibility of integrating faces and context, their decoding accuracy is reduced (\emph{e.g.}, Wagner et al., 1986).
-However, their superiority to the classifier tool is probably owing to their capacity to rely on their previous personal experience to invent a context in which a face could display such an expression.
+In the present case, the facial stimuli displayed to the decoders were as equivocal as real-life facial expressions (Aviezer et al., 2017), yet perceivers having no cue at their disposal.
+So, without any possibility of integrating faces and context, their decoding accuracy is necessarily reduced (\emph{e.g.}, Wagner et al., 1986).
+Their superiority to the classifier tool is probably owing to their capacity to rely on their previous personal experience to invent a context in which a face could display such an expression.
 Once a credible context is retrieved, they can affix an emotional label to the facial behavior.
@@
 The latter were triggered and measured with conceptually identical elicitation and assessment procedures (Reisenzein et al., 2013).
-The use of self-reports to evaluate encoders' subjective feelings is another limitation that can also be put forward because of the numerous cognitive biases they entail.
+The use of self-reports to evaluate expressers' subjective feelings is another limitation that can also be put forward because of the numerous cognitive biases they entail.
 Well known problems with the reliability of self-reports are, among others, the reconstructive nature of memory, the influence of attentional biases on reports, demand characteristics, distorting effects of implicit causal theories and personal motives, as stressed by Nielsen \& Kaszniak (2007).
@@
 Whereas this paradigm is longer and more complicated, it can lead to more robust results in reducing the forced-choice bias (Russell, 1993).
-However this procedure can also reduce the human observers' accuracy.
+However this procedure can also reduce the human perceivers' accuracy.
 In this regard, the results of the human observation could have been more ambiguous because it is not the natural way that people are inferring meaning from facial expressions.
 An alternative explanation relies in reducing the recognition bias involved in the classic recognition paradigm.
-Classic forced-choice paradigms obtain artificially high results, thus by using a more evolved approach, observers' accuracy may have been lowered.
+Classic forced-choice paradigms obtain artificially high results, thus by using a more evolved approach, perceivers' accuracy may have been lowered.
 Another flaw is the lack of comparison with various facial expression recognition methods.
@@
 It has the advantage of controlling different parameters of the emotions investigated, such as intensity, quality, and temporal (onset, duration) features.
-Moreover, as encoders are alone, facing an emotional trigger, it also controls for the social context, removing its possible influence on their facial behaviors.
-However, it is known that encoders' imagination can influence their expressiveness, for instance when they believe that their friend is doing the same (\emph{vs.} a different) emotional task in another room (Jakobs et al., 1999).
-Trying to exclude social influence by leaving encoders alone may be illusionary.
+Moreover, as expressers are alone, facing an emotional trigger, it also controls for the social context, removing its possible influence on their facial behaviors.
+However, it is known that expressers' imagination can influence their expressiveness, for instance when they believe that their friend is doing the same (\emph{vs.} a different) emotional task in another room (Jakobs et al., 1999).
+Trying to exclude social influence by leaving expressers alone may be illusionary.
 From an ecological perspective, it is even a mistake to exonerate behavioral observations from social contexts.
@@
 All raise serious objections to the supposed close relation between emotion and face.
-They bring up several questions regarding the role the context plays in the emission and interpretation of the so-called facial expression of emotion.
-The finalization of operational and effective ``reading emotional faces'' devices rests on the answers, if any, to the questions raised.
+They bring up several questions, in particular regarding the perceiver's role in the interpretation of the expresser's emotional face.
+Is EFE recognition a perceiver-independent or perceiver-dependent process?
+Although the present results tend to favor the second option, the finalization of operational and effective ``reading emotional faces'' devices rests on the answer, if any, to the question raised.
+In any case, as stressed by Krumhuber et al.
+(2020), machines' accurate emotion classification is predicted by the prototypicality of a facial display.
+More prototypical is the facial behavior, more accurate is the classification.
+However, ordinary emotional facial behaviors are seldom prototypical.
 As a result, despite being one of the most investigated questions in affective science, the growing interest of industries and governments in tracking individual's psychological states is supported by controversial assumptions.
-Considering the above, the present results provide additional evidence that an individual's subjective feelings cannot be inferred from facial expressions, and they invalidate the hypothesis of hardwired emotions unambiguously displayed on the face.
+
+Considering the above, the present results add to the body of literature evidence that an individual's subjective feelings are hardly inferred from spontaneous facial expressions in the absence of contextual cues, and they invalidate the hypothesis of hardwired emotions unambiguously displayed on the face.
+This study is the first one, to our knowledge, to look simultaneously at facial expressions spontaneously displayed in reaction to emotional triggers and their accurate decoding by human perceivers as compared to automatic emotion recognition tools.
+Results provide no support for the one-to-one mapping of subjective feelings and facial behaviors.
 Even if emotions were hardwired, in everyday life one does not observe prototypical facial expressions because of their rarity, and therefore research should be focused on analyzing non-prototypical facial expressions.
+Indeed, this study provide additional corroboration that facial displays of emotion are not exhibited in circumstances in which the Basic Emotion View would predict them to be expressed (Durán et al., 2017).
 Advancements in identifying ``non-basic'' emotion labels as well as non-prototypical facial expression have indeed occured with automatic facial expression recognition tools (McDuff, 2016).
-However, these present results suggest that automatic facial expression recognition tools should merely evaluate facial morphology features such as action units (already evaluated in OpenFace, Baltrušaitis et al., 2016; Affectiva's Affdex, McDuff et al., 2016; Vicar Vision's FaceReader, Den Uyl \& Van Kuilenburg, 2005, to name a few) rather than inferring supposedly emotional or affective states.
-Trying to interpret facial displays as a means of determining underlying emotional state, in all likelihood, remains in vain.
+However, these present results contribute to a critical understanding of the issues at stake.
+They suggest that automatic facial expression recognition tools should merely evaluate facial morphology features such as action units (already evaluated in OpenFace, Baltrušaitis et al., 2016; Affectiva's Affdex, McDuff et al., 2016; Vicar Vision's FaceReader, Den Uyl \& Van Kuilenburg, 2005, to name a few) rather than inferring supposedly emotional or affective states.
+Trying to interpret facial displays as a means of determining underlying emotional state (especially without any information about the context, nor person specifications of the expresser, \emph{etc.}), in all likelihood, remains dubious methods.
+
+\hypertarget{acknowledgement}{%
+\section{Acknowledgement}\label{acknowledgement}}
+
+The authors would like to thank the developers of the following R packages used to process, analyze, display and report data: R (Version 4.0.2; R Core Team, 2019) and the R-packages \emph{caret} (Version 6.0.86; Jed Wing et al., 2019), \emph{here} (Version 0.1; Müller, 2017), \emph{janitor} (Version 2.0.1; Firke, 2019), \emph{kableExtra} (Version 1.2.1; Zhu, 2019), \emph{papaja} (Version 0.1.0.9997; Aust \& Barth, 2018), and \emph{tidyverse} (Version 1.3.0; Wickham, 2017).
+
+\hypertarget{data-availability}{%
+\section{Data availability}\label{data-availability}}
+
+The R code and relevant data for statistical computing are available on the ``The emotion--facial expression link (data and code)'' repository from the Center for Open Science (Dupré \& Tcherkassof, 2020).
 
@@
 \begin{cslreferences}
+\leavevmode\hypertarget{ref-R-papaja}{}%
+Aust, F., \& Barth, M. (2018). \emph{papaja: Create APA manuscripts with R Markdown}. \url{https://github.com/crsh/papaja}
+
 \leavevmode\hypertarget{ref-aviezer2017inherently}{}%
@@
 
-\leavevmode\hypertarget{ref-dols2017natural}{}%
-Dols, J.
-M.
-F.
-(2017).
-Natural facial expression: A view from psychological constructionism and pragmatics.
-In J.
-M.
-F.
-Dols \& J.
-A.
-Russell (Eds.), \emph{The science of facial expression} (pp.
-457--478).
-Oxford University Press.
-
 \leavevmode\hypertarget{ref-duchenne1876mecanisme}{}%
@@
 \url{https://doi.org/10.1371/journal.pone.0231968}
+
+\leavevmode\hypertarget{ref-Dupruxe9_Tcherkassof_2020}{}%
+Dupré, D., \& Tcherkassof, A. (2020). \emph{The emotion--facial expression link (data and code)}. OSF. \url{https://doi.org/10.17605/OSF.IO/ERUA5}
 
@@
 
+\leavevmode\hypertarget{ref-dols2017natural}{}%
+Fernández-Dols, J.
+M.
+(2017).
+Natural facial expression: A view from psychological constructionism and pragmatics.
+In J.
+M.
+Fernández-Dols \& J.
+A.
+Russell (Eds.), \emph{The science of facial expression} (pp.
+457--478).
+Oxford University Press.
+
 \leavevmode\hypertarget{ref-fernandez2013emotion}{}%
@@
 \url{https://doi.org/10.1037/0022-3514.69.6.1113}
+
+\leavevmode\hypertarget{ref-R-janitor}{}%
+Firke, S. (2019). \emph{Janitor: Simple tools for examining and cleaning dirty data}. \url{https://CRAN.R-project.org/package=janitor}
 
@@
 
+\leavevmode\hypertarget{ref-gendron2017facing}{}%
+Gendron, M., \& Barrett, L.
+F.
+(2017).
+Facing the past: A history of the face in psychological research on emotion perception.
+In J.-M.
+Fernández-Dols \& J.
+A.
+Russell (Eds.), \emph{The science of facial expression} (pp.
+15--36).
+Oxford University Press.
+
 \leavevmode\hypertarget{ref-gendron2018emotion}{}%
@@
 
+\leavevmode\hypertarget{ref-R-caret}{}%
+Jed Wing, M.
+K.
+C.
+from, Weston, S., Williams, A., Keefer, C., Engelhardt, A., Cooper, T., Mayer, Z., Kenkel, B., R Core Team, Benesty, M., Lescarbeau, R., Ziem, A., Scrucca, L., Tang, Y., Candan, C., \& Hunt., T.
+(2019).
+\emph{Caret: Classification and regression training}.
+\url{https://CRAN.R-project.org/package=caret}
+
 \leavevmode\hypertarget{ref-johnson2010smile}{}%
@@
 Kappas, A. (2003). What facial activity can and cannot tell us about emotions. In M. Katsikitis (Ed.), \emph{The human face} (pp. 215--234). Springer.
+
+\leavevmode\hypertarget{ref-keltner2019emotional}{}%
+Keltner, D., Sauter, D., Tracy, J., \& Cowen, A.
+(2019).
+Emotional expression: Advances in basic emotion theory.
+\emph{Journal of Nonverbal Behavior}, \emph{43}, 133--160.
+\url{https://doi.org/https://doi.org/10.1007/s10919-019-00293-3}
 
@@
 Multimodal Corpora: Advances in Capturing, Coding and Analyzing Multimodality}, 31--36.
+
+\leavevmode\hypertarget{ref-R-here}{}%
+Müller, K. (2017). \emph{Here: A simpler way to find your files}. \url{https://CRAN.R-project.org/package=here}
+
+\leavevmode\hypertarget{ref-niedenthal2017embodied}{}%
+Niedenthal, P.
+M., Wood, A., Rychlowska, M., \& Korb, S.
+(2017).
+Embodied simulation in decoding facial expression.
+In J.-M.
+Fernández-Dols \& J.
+A.
+Russell (Eds.), \emph{The science of facial expression} (pp.
+397--414).
+Oxford University Press.
 
@@
 
+\leavevmode\hypertarget{ref-R-base}{}%
+R Core Team. (2019). \emph{R: A language and environment for statistical computing}. R Foundation for Statistical Computing. \url{https://www.R-project.org/}
+
 \leavevmode\hypertarget{ref-reisenzein2000exploring}{}%
@@
 \url{https://doi.org/10.1037/0022-3514.50.4.737}
+
+\leavevmode\hypertarget{ref-R-tidyverse}{}%
+Wickham, H. (2017). \emph{Tidyverse: Easily install and load the 'tidyverse'}. \url{https://CRAN.R-project.org/package=tidyverse}
+
+\leavevmode\hypertarget{ref-R-kableExtra}{}%
+Zhu, H. (2019). \emph{KableExtra: Construct complex table with 'kable' and pipe syntax}. \url{https://CRAN.R-project.org/package=kableExtra}
+
+\leavevmode\hypertarget{ref-R-papaja}{}%
+Aust, F., \& Barth, M. (2018). \emph{papaja: Create APA manuscripts with R Markdown}. \url{https://github.com/crsh/papaja}
 
@@
 \url{https://doi.org/10.1037/0022-3514.50.4.737}
+
+\leavevmode\hypertarget{ref-R-tidyverse}{}%
+Wickham, H. (2017). \emph{Tidyverse: Easily install and load the 'tidyverse'}. \url{https://CRAN.R-project.org/package=tidyverse}
+
+\leavevmode\hypertarget{ref-R-kableExtra}{}%
+Zhu, H. (2019). \emph{KableExtra: Construct complex table with 'kable' and pipe syntax}. \url{https://CRAN.R-project.org/package=kableExtra}
 \end{cslreferences}

### 2020-11-06 manuscript/manuscript_apa.Rmd
osf link

@@
 
-The R code and relevant data for statistical computing are available on the "The emotion–facial expression link (data and code)" repository from the Center for Open Science [@Dupré_Tcherkassof_2020].
+The R code and relevant data for statistical computing are available on *The emotion–facial expression link (data and code)* repository from the Center for Open Science at https://osf.io/erua5/ [@Dupré_Tcherkassof_2020].
 

### 2020-11-06 manuscript/manuscript_apa.tex
osf link

@@
 
-The R code and relevant data for statistical computing are available on the ``The emotion--facial expression link (data and code)'' repository from the Center for Open Science (Dupré \& Tcherkassof, 2020).
+The R code and relevant data for statistical computing are available on \emph{The emotion--facial expression link (data and code)} repository from the Center for Open Science at \url{https://osf.io/erua5/} (Dupré \& Tcherkassof, 2020).
 

### 2020-11-21 manuscript/manuscript_apa.tex
final

@@
 They conducted their experiment among U.S.
-participants and Himba participants from remote regions of Namibia (Gendron et al., 2014) and Hazda participants of the Eastern Rift Valley of Tanzania (Gendron et al., 2018) both groups with limited exposure to Western culture.
+participants and Himba participants from remote regions of Namibia (Gendron et al., 2014) and Hazda participants of the Eastern Rift Valley of Tanzania (Gendron et al., 2020) both groups with limited exposure to Western culture.
 Gendron and her colleagues demonstrated that facial expressions were not universally recognized in discrete emotional terms.
@@
 \leavevmode\hypertarget{ref-gendron2018emotion}{}%
-Gendron, M., Hoemann, K., Crittenden, A. N., Msafiri, S., Ruark, G. A., \& Barrett, L. F. (2018). \emph{Emotion perception in hadza hunter gatherers}. \url{https://doi.org/10.31234/osf.io/pf2q3}
+Gendron, M., Hoemann, K., Crittenden, A.
+N., Mangola, S.
+M., Ruark, G.
+A., \& Barrett, L.
+F.
+(2020).
+Emotion perception in hadza hunter-gatherers.
+\emph{Scientific Reports}, \emph{10}(1), 1--17.
+\url{https://doi.org/10.1038/s41598-020-60257-2}
 
